### 2017-10-02 OaklandSubmission/main.tex
starting repo

\documentclass[conference,compsoc]{IEEEtran}

\usepackage{epsfig,endnotes}
\usepackage{varwidth}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{balance}
\usepackage{color}
\usepackage{nicefrac}
\usepackage{amsmath}
\usepackage{braket}
\usepackage{bm}
\usepackage{mathtools}
\usepackage{multirow}
\usepackage{bigdelim}
\usepackage{mathtools}
\usepackage{amssymb}
\usepackage{indentfirst}
\usepackage{booktabs}
\usepackage{enumitem}
\usepackage{boxedminipage}
\usepackage{float}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{cleveref}
\usepackage{subcaption}
\usepackage[normalem]{ulem}
\usepackage{xpatch}
\usepackage{stmaryrd}

\makeatletter
\xpatchcmd{\algorithmic}{\itemsep\z@}{\itemsep=0.3ex plus2pt}{}{}
\makeatother

\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
\newcommand{\crefrangeconjunction}{ to~}
\newcommand{\minus}{\scalebox{0.75}[1.0]{$-$}}
\newcommand{\sameer}[1]{{\leavevmode\color{red}{#1}}}
\setlength{\textfloatsep}{0.3cm}
\newcommand{\ourprotocol}{Tuffy}

\newcommand{\bits}{\{0,1\}}
\newcommand{\Ex}{\mathbb{E}}
\newcommand{\To}{\rightarrow}
\newcommand{\R}{\mathbb{R}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\F}{\mathbb{F}}
\newcommand{\GF}{\mathsf{GF}}

\usepackage{MnSymbol}
\newcommand{\maxpr}{\text{\rm max-pr}}
\newcommand{\cclass}[1]{\mathbf{#1}}
\renewcommand{\P}{\cclass{P}}
\newtheorem{problem}{Problem}
\newtheorem{theorem}{Theorem}
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{claim}[theorem]{Claim}
\newtheorem{fact}[theorem]{Fact}
\newtheorem{remk}[theorem]{Remark}
\newtheorem{apdxlemma}{Lemma}

\newcommand{\etal}{{\it et~al.\ }}
\newcommand{\ie} {{\it i.e.,\ }}
\newcommand{\eg} {{\it e.g.,\ }}
\newcommand{\cf}{{\it cf.,\ }}
\newcommand{\suchthat}{{\;\; : \;\;}}
\newcommand{\pr}[1]{\Pr\left[#1\right]}

\newcommand{\class}[1]{\mathbf{#1}}
\newcommand{\SZK}{\class{SZK}}
\newcommand{\BPP}{\class{BPP}}
\newcommand{\NP}{\class{NP}}
\renewcommand{\P}{\class{P}}
\newcommand{\negl}{{\mathrm{neg}}}

\newcommand{\Enc}{E}
\newcommand{\Dec}{D}
\newcommand{\Gen}{G}

\newcommand{\PK}{\mathit{PK}}
\newcommand{\SK}{\mathit{SK}}
\newcommand{\pk}{\mathit{pk}}
\newcommand{\sk}{\mathit{sk}}
\newcommand{\MD}[1]{\mathrm{MD{#1}}}
\newcommand{\SHA}{\mbox{SHA-1}}

\begin{document}
\title{SecureDNN: Efficient Analytics over Encrypted Data}

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
% \author{\IEEEauthorblockN{Michael Shell}
% \IEEEauthorblockA{School of Electrical and\\Computer Engineering\\
% Georgia Institute of Technology\\
% Atlanta, Georgia 30332--0250\\
% Email: http://www.michaelshell.org/contact.html}
% \and
% \IEEEauthorblockN{Homer Simpson}
% \IEEEauthorblockA{Twentieth Century Fox\\
% Springfield, USA\\
% Email: homer@thesimpsons.com}
% \and
% \IEEEauthorblockN{James Kirk\\ and Montgomery Scott}
% \IEEEauthorblockA{Starfleet Academy\\
% San Francisco, California 96678-2391\\
% Telephone: (800) 555--1212\\
% Fax: (888) 555--1212}}

\maketitle
% \begin{abstract}
% The abstract goes here.
% \end{abstract}
\IEEEpeerreviewmaketitle

\graphicspath{{./Images/}}
\DeclareGraphicsExtensions{.pdf,.jpg,.png}

\input{./Sections/CompleteProtocol.tex}
\newpage
\clearpage
\input{./Sections/ReLU.tex}
\input{./Sections/Independent.tex}
\input{./Sections/Introduction.tex}

\newpage
\bibliographystyle{abbrv}
\bibliography{./bib}
\balance

\end{document}

### 2017-10-02 OaklandSubmission/eval.tex
minor

We first implement benchmarks of ~\cite{shafindss} and~\cite{minionn} in \tool.
Next, we show the generality of \tool by implementing state-of-the-art machine learning models
. 
lin
nb
tree\\
minion-cnn
secureml
cryptonet\\
logistic
cnn
bonsai\\
cifar

### 2017-10-02 OaklandSubmission/intro.tex
init

\section{Introduction}
\label{sec:intro}

Start here \cite{lambdaps,wysteria}

### 2017-10-03 OaklandSubmission/eval.tex
minor

@@
+\section{Evaluation}
+\label{sec:eval}
+
 We first implement benchmarks of ~\cite{shafindss} and~\cite{minionn} in \tool.
-Next, we show the generality of \tool by implementing state-of-the-art machine learning models
-. 
+Next, we show the generality of \tool by implementing state-of-the-art machine learning models.
+Finally, we discuss a model that requires pipelinig.
+Our results are in two settings, a LAN setting and a cross-continent WAN setting.
+The round trip time for two machines in a LAN setting is 1ms and in WAN setting is 40ms.
+
+\subsection{Standard models}
+\label{sec:shallow}
+We consider three types of classifiers: linear, Naive Bayes, and decision trees. 
+These classifiers are trained on the followinng data sets from the UCI  machine learning repository:
+ the Wisconsin Breast Cancer data set, 
+Credit Approval data set, Audiology (Standardized) data set, Nursery data set, and
+ECG (electrocardiogram) classification data from~\cite{barni}.
+
+%dataset  | size | gates | and | mul | depth
+
+%dataset | time | comm | timeL | commL | timeW | commW
+
 lin

### 2017-10-03 OaklandSubmission/intro.tex
intro-flow-1

@@
 
-Start here \cite{lambdaps,wysteria}
+Secure multiparty computation \cite{yao,gmw} (SMC) is a powerful cryptographic tool that allows a set of mutually distrusting parties to compute a joint function of their secret inputs.
+Examples of these include running machine learning prediction algorithm where one party holds the secret medical report and the other party holds the model for desease.
+Since the introduction of SMC in 1980's, there has been a large body of work \cite{..} that has transformed SMC from a mere theoretical tool to something that can be used in practice.
+
+Unfortunately, implementing fuctionalities using SMC protocols requires thorogh understanding of cryptography.
+To allow for widespread use of SMC, it is critical that SMC protocols are programmable by non-cryptographic experts.
+To cater to this, there have been several efforts developing domain-specific languages that are programmer friendly and compile to a SMC back-end.
+Most notable of these systems include \cite{...}.
+A common feature of all of these works is that they use a boolean circuit based cryptographic backend such as garbled circuits.
+As one can expect, compiling a function to a boolean circuit is one of the biggest efficiency bottleneck.
+
+Recent works \cite{aby,secureml,minion} have shown that great performance benefit can be obtained by mixing arithmetic and boolean computations in the back-end.
+Certain computations such as integer multiplication are efficient when implemented using arithmetic SMC \cite{gmw} whereas max(x,0) needs to be computed using a boolean back-end \cite{yao}.
+More details later..
+
+\cite{lambdaps,wysteria}

### 2017-10-03 OaklandSubmission/intro.tex
intro-flow-1

@@
 More details later..
+ABY \cite{aby} requires programmer to be aware of arithmetic and boolean trade-offs and  write high-level circuits consisting of both arithmetic and boolean gates and share-conversion gates.
+Other works such as \cite{secureml,minion} build on ideas from \cite{aby} and develop tailor made algorithms for neural network training and prediction \cite{ml} and claim huge improvements over only boolean implementations.
+As is already mentioned in these works, their systems are mere proofs-of-concept and far from being implementable.
+
+In this work, we develop and implement our framework \tool where the programmer writes a high level C-like program (instead of a circuit) to describe the function to be computed.
+The programmer is oblivious of the Our compiler automatically compiles it to a circuit framework consisting to both arithmetic and boolean gates as well share-conversion gates wherever required.
+We use ABY as the cryptographic backend.
+We give a type system, correctness .....
+We evaluate our framework on ...
+and show performance comparable and even better than tailor made protocols.
 

### 2017-10-03 OaklandSubmission/intro.tex
intro-flow-2

@@
 In this work, we develop and implement our framework \tool where the programmer writes a high level C-like program (instead of a circuit) to describe the function to be computed.
-The programmer is oblivious of the Our compiler automatically compiles it to a circuit framework consisting to both arithmetic and boolean gates as well share-conversion gates wherever required.
+The programmer is oblivious of the cryptographic backend being used.
+Our compiler automatically compiles it to a circuit framework consisting to both arithmetic and boolean gates as well share-conversion gates wherever required.
 We use ABY as the cryptographic backend.
@@
 and show performance comparable and even better than tailor made protocols.
+It is simple to program in our framework and our work provides a meaningful baseline for future works designing tailor made protocols for specific functionalities.
 
+\subsection{\tool: Overview and Contributions}
+
+\noindent\textbf{Scaling to large programs.} Describe pipeling here...
+
+\noindent\textbf{Related Work}
 \cite{lambdaps,wysteria}

### 2017-10-03 OaklandSubmission/eval.tex

@@
 
-%dataset  | size | gates | and | mul | depth
+For a  (binary) linear classifier of model size $n$ we output $w^Tx>0$ where $w$ and $x$ are vectors
+of length $n$. In Naive Bayes classification, we are given a vector $P$ of log-probabilities
+$\log p(C=c_i)$, where $p(C=c_i)$ is the likelihood that the output class is $c_i$ for $i\in[n]$. 
+We are also given a matrix $T$ s.t. $T[i][j]=\log p(X_j=x_j|C=c_i)$, i.e., the log-probability that 
+the $j^{th}$ input feature is $x_j$ (where $j\in [d]$) given that the output class is $c_i$. The output is
+\[
+{\it argmax}_i P[i]+\sum_{j=1}^d\log p(X_j=x_j|C=c_i)
+\]
+A decision tree has a boolean predicate $b_i(x) = x_i\leq w_i$ at each internal node and classes as leaves.
+We start from root and if $b_i(x)$ evaluates to true then we go to the right node else
+we go to the right node. This process is repeated until we reach a leaf that provides
+the predicted label. Such a decision tree can be encoded as a polynomial and then prediction
+reduces to polynomial evaluation. E.g., a tree with depth one can be encoded as the polynomial
+$(1-b(x))c_l + b(x)c_r$ where $b(x)$ is the predicate at the root, $c_l$ is the label of the left leaf
+and $c_r$ is the label of the right leaf. The trees in our benchmark set have a depth of 4.
 
-%dataset | time | comm | timeL | commL | timeW | commW
+\begin{table*}
+\begin{tabular}{c|c|c|c|c |c|c|c|c|c|c}
+Dataset & $n$  & Prev. time (s) & Prev. Comm (kb) & LAN time (s) & WAN time (s) & Comm. (kb)  & \#And & \#Mul & \#Gates & size\\
+\hline
+Breast cancer & 30 & 0.3 & 36 & 0.1 & 0.3 & 25 & 95 & 30 & 727 & 20\\
+\hline
+Credit & 47 & 0.3 & 41 & 0.1 & 0.3 & 36 & 95 & 47 & 795 & 20\\
+\hline
+\end{tabular}
 
-lin
-nb
-tree\\
-minion-cnn
-secureml
-cryptonet\\
-logistic
-cnn
-bonsai\\
-cifar
+ \caption{Linear classification benchmarks}
+ \label{tab:lcd} 
+\end{table*}
+
+\begin{table*}
+\begin{tabular}{c|c|c|c|c|c |c|c|c|c|c|c}
+Dataset & $|P|$ & $|x|$ & Prev. time (s) & Prev. Comm (Mb) & LAN time (s) & WAN time (s) & Comm. (Mb)  & \#And & \#Mul & \#Gates  & size\\
+\hline
+Nursery & 5 & 9 & 1.5 & 0.2 & 0.1 & 0.4 & 0.6 & 13k & 0 & 73k & 50\\
+\hline
+Audiology & 24 & 70 & 3.9 & 2.0 & 1.5 & 2.9 & 37 & 750k & 0 & 4219k & 50\\
+\hline
+\end{tabular}
+
+ \caption{Naive Bayes benchmarks}
+ \label{tab:lcd} 
+\end{table*}
+
+\begin{table*}
+\begin{tabular}{c|c|c|c |c|c|c|c|c|c}
+Dataset  & Prev. time (s) & Prev. Comm (kb) & LAN time (s) & WAN time (s) & Comm. (kb)  & \#And & \#Mul & \#Gates & size\\
+\hline
+Nursery &  0.3 & 102 & 0.1 & 0.3 & 32 & 504 & 3 & 3324 & 20\\
+\hline
+ECG &  0.4 & 102 & 0.1 & 0.4 & 49 & 756 & 5 & 5002 & 20\\
+\hline
+\end{tabular}
+
+ \caption{Decision tree benchmarks}
+ \label{tab:lcd} 
+\end{table*}
+
+%dataset  | size | gates | and | mul | depth | length
+
+%dataset | time | comm | timeL | timeW | comm
+\subsection{Neural networks}
+We evaluate on the benchmarks described in \minion~\cite{minionn}.
+The first is the neural network described in SecureML~\cite{secureml}.
+It has three fully connected layers with square as the activation function.
+Next, we implement the neural network of Cryptonets~\cite{cryptonets} in \tool.
+This network also uses square as the activation function and has one convolution (with 5 output channels), one maxpool, and one fully connected layer.
+Finally, we compare against\minion on a convolutional neural network
+with two convolutions (with 16 output channels), two maxpool, and two fully connected layers.
+In contrast to the previous two, this network uses ReLU for activation.
+In our results, we only compare with \minion, as it outperforms the earlier systems of
+Cryptonets and SecureML.
+%net | minionT | minionC | timeL | timeW | comm | gates | and | mul | length
+
+\begin{table*}
+\begin{tabular}{c|c|c|c |c|c|c|c|c|c}
+Name  & Prev. time (s) & Prev. Comm (Mb) & LAN time (s) & WAN time (s) & Comm. (Mb)  & \#And & \#Mul & \#Gates & size\\
+\hline
+SecureML   &  1.1 & 15.8 & 0.7 & 1.7  & 76   &  2k   & 119k & 366k   & 78\\
+\hline
+Cryptonets &  1.3 & 47.6 & 0.6 & 1.6  & 7    & 2k    & 108k & 316k & 88\\
+\hline
+CNN        &  9.4 & 657.5& 5.1 & 11.6 & 501  & 1640k & 667k & 9480k & 154\\
+\hline
+\end{tabular}
+
+ \caption{Neural network benchmarks}
+ \label{tab:lcd} 
+\end{table*}
+
+\subsection{Practical models}
+Tensorflow~\cite{tensorflow} is a standard machine learning toolkit.
+Its tutorial describes two prediction models for handwritten digit recognition
+using the MNIST dataset~\cite{mnist} which has 55,000 images as training data,
+10,000 images as test data (mnist.test), and 5,000 images as validation data.
+Each image is a greyscale image of size $28\times 28$.
+The first model that the tutorial describes is a softmax regression
+which provides an accuracy of 92\%. Since this model is over floating-point,
+as is standard~\cite{secureml,minionn}, we port it to a model over 32-bit integers manually. After our port, the accuracy is still preserved.
+Next, we implement this model in \tool and the results are shown below.
+
+The next model in the Tensorflow tutorial is a convolution neural net with two convolutions
+(with 32 output channels), two maxpool, and two fully connected layers.
+This network is both bigger and more accurate than the networks presented in the previous section.
+In particular, it has an accuracy of 99.2\%. Again, since this model is in floating-point,
+we port it to integers while maintaining the accuracy\footnote{The accuracy of the integer model is 0.04\% higher than the floating-point model.}.
+
+% benchmark | size | gates | and | mul | depth  | length | comm | timeL | timeW
+
+\begin{table*}
+\begin{tabular}{c|c|c|c |c|c|c|c|c|c | c}
+Name       & LAN time (s) & WAN time (s) & Comm. (Mb)  & \#And & \#Mul & \#Gates & Model size & Program size\\
+\hline
+Regression &  0.1         & 0.7         & 5            & 2k    & 8k    &  35k    & 8k   & 38\\
+\hline
+CNN        &  30.5        & 60.3        & 2955         & 6082k & 4163k &  42104k & 3226k& 172\\
+\hline
+\end{tabular}
+
+ \caption{Tensorflow tutorial benchmarks}
+ \label{tab:lcd} 
+\end{table*}
+
+Finally, there are a class of machine learning models that are much more efficient than
+deep neural nets and provide good accuracies on standard learning tasks. One such model
+is \bonsai~\cite{bonsai}, which is a decision tree with more powerful nodes.
+In particular, each node $m_k$ evaluates a predicate $\theta_k^TZx > 0$ to decide whether
+to pass $x$ to the left branch or the right branch. and the predicted value is
+\[
+{\it argmax} \sum_k I_k(x)W_k^TZx\circ f(\sigma V_k^TZx) 
+\]
+Here, $I_k$ is 1 if the $k^{th}$ node is present on the path traversed by $x$
+and is zero otherwise. 
+The operation $\circ$ is a pointwise dot product and
+the function $f$ is given by $f(x) = x$ if $|x| < 1$ and $\mathrm{signum}(x)$ otherwise.
+We show results on three standard data sets: character  (Chars4k~\cite{campos}), text (USPS~\cite{hull}), and object categorization (WARD~\cite{yang}).
+There are other other benchmarks in~\cite{bonsai} that do not add any new insights here.
+Moreover, the dataset WARD requires the largest model.
+These models are already over integers and no port from floating-point is required.
+%dataset | size | gates | and | mul | depth | length | comm | timeL | timeW
+
+\begin{table*}
+\begin{tabular}{c|c|c|c |c|c|c|c|c|c | c}
+Name       & LAN time (s) & WAN time (s) & Comm. (Mb)  & \#And & \#Mul & \#Gates & depth & Program size\\
+\hline
+Chars4k    &  0.1         & 0.7         & 2            & 18k    & 3k    &  85k     & 1   & 89\\
+\hline
+USPS       &  0.2         & 0.9         & 4            & 62k    & 2k    &  285k    & 2   & 156\\
+\hline
+WARD       &  0.3         & 1.1         & 9            & 106k    & 8k    &  506k    & 3   & 283\\
+\hline
+\end{tabular}
+
+ \caption{Bonsai benchmarks}
+ \label{tab:lcd} 
+\end{table*}
+
+\subsection{Pipelining}
+The largest benchmark of \minion is a neural network for CIFAR-10 dataset~\cite{cifar}: label colored images for objects in 10 classes. This network needs more memory than what is available
+on our machines. Therefore, we use pipelining and divide the computation into seven stages.
+The first step does a convolution with 64 output channels and a ReLU activation.
+The next four stages perform a convolution that involves multiplying a $64\times 576$ matrix with a
+$576\times 1024$ matrix. The next stage performs a ReLU, a meanpool, and a convolution.
+The final stage performs four convolutions, a mean pool, five ReLU and a fully connected layer.
+The results show the end-to-end results as well as the statistics for the heaviest stage.
+
+\begin{table*}
+\begin{tabular}{c|c|c|c |c|c|c}
+           & LAN time (s) & WAN time (s) & Comm. (Mb)  & \#And & \#Mul & \#Gates \\
+\hline
+Total      &  265.6       & 647.5        & 40683       & 21m    & 61m    &  337m  \\
+\hline
+Stage 6    &  55.2        & 122.6        & 6744        & 12m    & 10m   &  98m  \\
+\hline
+\end{tabular}
+
+ \caption{Pipelining results}
+ \label{tab:lcd} 
+\end{table*}
+
+%stage | time | comm | gates | mul | add | depth

### 2017-10-05 OaklandSubmission/intro.tex
intro-ideas

@@
 We use ABY as the cryptographic backend.
-We give a type system, correctness .....
+\divya{Say something about what kind of backend we want.
+provides SMC for a mix of arithmetic and boolean circuit with appropriate secure conversion between two types and ABY provides such a framework for semi-honest 2pc.} Our work is compatible with malicious, multiparty as well...
+etc We give a type system, correctness .....
 We evaluate our framework on ...
@@
 
-\subsection{\tool: Overview and Contributions}
+\subsection{\tool: Overview and Contributions} Write high level overview of programming language and tool..
 
-\noindent\textbf{Scaling to large programs.} Describe pipeling here...
+\divya{ideally have small paragraphs about key features}
+1. ease to programming.. refer to section with example in our language; contrast to verilog paper; programmer is completely oblivious of security; looks like a normal C program
+
+2. first programmable system that supports a mix of arithmetic and boolean for efficiency give example of multiplication of integers; linear algebra; logistic regression
+
+3.
+prove correctness, type system guarantees correct termination, progress etc..; In contrast with most systems that require a programmer to label variables as public secret, we automatically infer them.
+leads to more efficiency compared to default secret
+
+4. Scaling to large programs. Describe pipeling here...
+
+5. can serve as a baseline for future works on hand-crafted algorithms for specific applications
+
+6.
+Suitable to use multiple crypto backends; We formalize the guarantees needed from crypto backend; it is easy for expert programmers to to extend our framework for example add another SMC backend.
+say more about this later...
+there say that possible to integrate minion's more efficient matrix based preprocessing into our framework and matrix ops as native types and instructions in our language
+
+7. experiment results...
 
@@
 \cite{lambdaps,wysteria}
+
+\divya{Useful phrases:} progress in practical efficiency; non-specialist programmers; fundamental representation gap; circuits as an abstraction of computation;
+
+\divya{useful thoughts} It is widely believed that garbled circuit based 2pc is most practical; but not quite true..
+
+\divya{points that have to be there...} Security; traces are ``oblivious'' of secret values if observables/outputs are same; formal cryptographic security guarantee uses simulation paradigm \cite{canetti2000}
+
+\divya{not sure if we wanna say anything about these} 1.
+all the programs written in our language are memory trace oblivious because array indices are public 2.
+instruction trace obliviousness handled by using a multiplexor or executing both branches..
+plus all loops are public counters.
+Our type system rejects programs that loop on secret variables
+
+\divya{questions:} Does ABY allow secret array indices? or does it allow arrays? thats not so important actually
+
+\divya{Related Work: writing some citations here to be used appropriately.}
+\begin{enumerate}
+\item Secure Computation of MIPS Machine Code %https://eprint.iacr.org/2015/547
+
+\end{enumerate}
+\newpage

### 2017-10-05 OaklandSubmission/eval.tex
warn fix
start eval

@@
 
-We first implement benchmarks of ~\cite{shafindss} and~\cite{minionn} in \tool.
+We first implement benchmarks of ~\cite{shafindss} and~\cite{minionn} in \tool to compare gainst 
+previous work.
 Next, we show the generality of \tool by implementing state-of-the-art machine learning models.
@@
 The round trip time for two machines in a LAN setting is 1ms and in WAN setting is 40ms.
+Each machine has a Intel(R) Xeon(R) CPU E5-2673 v3 processor running at 2.40GHz with
+28 GBs of RAM. Since most of our benchmarks are related to machine learning,
+we set up some notation standard to machine learning next.
+
+\subsection{Benchmarks}
+In this paper, we focus on machine learning models for {\it classification}.
+A {\it classifier} $C$ uses a trained model to predict a {\it label} $\ell$ for an input data point $x$. For example, given a data point which is a tuple of height and weight of an individual,
+a classifier can predict a label ``male'' or ``female''.
+
+A binary linear classifier is one of the simplest classifers. Here, the input data point $x\in\R^d$
+and the model is a vector $w\in\R^d$. The possible labels $\ell\in\{\mathit{true},\mathit{false}\}$
+and the classifier $C_w\equiv w^Tx>0$, where $T$ denotes the matrix transpose operator.
+This classifier requires $n$ multiplications, $n-1$ additions, and a comparison.
+
+A more interesting classifier is Naive Bayes that predicts labels from the set $[n]$.
+Here, the input data point is a {\it feature}
+vector $x=(x_1,x_2,\ldots,x_d)^T$ where each $x_i\in [F]$.
+The model has two matrices: a vector $P$ with $|P|=n$ and $P(i) \equiv \log p(\ell=i)$, where $p(\ell=i)$ is the likelihood that the output label is $i$ for $i\in[n]$.
+The other matrix $T$ has size $n\times d\times F$ and $T(i)(j)(x_j)\equiv\log p(x_j|\ell=i)$, i.e., the log-probability that 
+the $j^{th}$ input feature is $x_j$ (where $j\in [d]$) given that the output label is $i$.
+The classifier $C_{P,T}(x)$ outputs
+\[
+{\it argmax}_i P(i)+\sum_{j=1}^d T(i)(j)(x_j)
+\]
+This classifier requires $ndF$ comparisons and additions.
+
+A decision tree of size $N$ takes as input $x\in\R^n$ and the model consists of a boolean predicate
+assigned to each internal node of the tree.
+In particular, a node $j\in[N]$ at depth $i$ has a predicate $b_{j}^i\equiv x_i\leq w_i$.
+We start evaluating the tree from the root and if the predicate at the current node $b_j^i$ is false (true) then $x$ is passed to the left (right) node $b_{2j+1}^{i+1}$ ($b_{2j+2}^{i+1}$).
+This process is repeated till we reach a leaf.
+The leaves of the tree have labels and the output label is the leaf visited by this traversal.
+Such a (binary) decision tree can be encoded as a polynomial and then prediction
+reduces to polynomial evaluation. E.g., a binary tree with depth one and size $N=3$ can be encoded as the polynomial
+$(1-b(x))\ell_1 + b(x)\ell_2$ where $b(x)$ is the predicate at the root, $\ell_1$ is the label of the left leaf and $\ell_r$ is the label of the right leaf.
+
+The next class of classifiers that we discuss are deep neural nets or DNNs. 
+A DNN has multiple layers that each compute a matrix multiplication followed by
+an {\it activation} function. The most common activation functions are 
+squaring $f(x)=x^2$ and ReLU $f(x)=\mathit{max}(x,0)$. 
+If $A$ is a matrix then the output of $f(A)$ is also a matrix that contains $f$  applied
+ to each entry of $A$ pointwise.
+Given an input vector $x$, the predicted label of a DNN is
+\[
+ \mathit{argmax} W_n\cdot f_{N-1}(\ldots f_1(W_1\cdot x)\ldots)
+\]
+Here, $W_i$'s are matrices that are part of the model and $f_i$'s are the activation functions.
+The operator $\cdot$ denotes a matrix multiplication.
+Neural nets usually have one or more fully connected layers.
+These layers multiply a matrix with a vector.
+Some neural nets have convolution layers and these are called CNNs.
+For the purpose of this paper, a convolution can be considered as a (heavy) matrix-matrix multiplication. Therefore, fully connected layers are easier computation-wise compared
+to convolution layers. Such DNNs and CNNs are computationaly heavy but provide
+much better accuracies on computer vision tasks than the classifiers discussed above.
+
+Finally, there are a class of machine learning models that are much more efficient than
+deep neural nets and provide reasonably good accuracies on standard learning tasks. One such model
+is \bonsai~\cite{bonsai}, which is a variant of the decision trees described above.
+In particular, each node $m_k$ evaluates a predicate $\theta_k^TZx > 0$ to decide whether
+to pass $x$ to the left branch or the right branch ($\theta_k$ is a vector and $Z$ is a matrix).
+The predicted value is
+\[
+{\it argmax} \sum_k I_k(x)W_k^TZx\circ f(V_k^TZx) 
+\]
+Here, $I_k$ is 1 if the $k^{th}$ node is present on the path traversed by $x$
+and is zero otherwise. 
+The operation $\circ$ is a pointwise dot product, $W_k$'s and $V_k$'s are vectors,  and
+the function $f$ is given by $f(x) = x$ if $|x| < 1$ and $\mathrm{signum}(x)$ otherwise.
+
+In the following, we implement these classifiers in \tool and report the time taken for making predictions securely.
 
@@
 
-For a  (binary) linear classifier of model size $n$ we output $w^Tx>0$ where $w$ and $x$ are vectors
-of length $n$. In Naive Bayes classification, we are given a vector $P$ of log-probabilities
-$\log p(C=c_i)$, where $p(C=c_i)$ is the likelihood that the output class is $c_i$ for $i\in[n]$. 
-We are also given a matrix $T$ s.t. $T[i][j]=\log p(X_j=x_j|C=c_i)$, i.e., the log-probability that 
-the $j^{th}$ input feature is $x_j$ (where $j\in [d]$) given that the output class is $c_i$. The output is
-\[
-{\it argmax}_i P[i]+\sum_{j=1}^d\log p(X_j=x_j|C=c_i)
-\]
-A decision tree has a boolean predicate $b_i(x) = x_i\leq w_i$ at each internal node and classes as leaves.
-We start from root and if $b_i(x)$ evaluates to true then we go to the right node else
-we go to the right node. This process is repeated until we reach a leaf that provides
-the predicted label. Such a decision tree can be encoded as a polynomial and then prediction
-reduces to polynomial evaluation. E.g., a tree with depth one can be encoded as the polynomial
-$(1-b(x))c_l + b(x)c_r$ where $b(x)$ is the predicate at the root, $c_l$ is the label of the left leaf
-and $c_r$ is the label of the right leaf. The trees in our benchmark set have a depth of 4.
-
 \begin{table*}
 \begin{tabular}{c|c|c|c|c |c|c|c|c|c|c}
-Dataset & $n$  & Prev. time (s) & Prev. Comm (kb) & LAN time (s) & WAN time (s) & Comm. (kb)  & \#And & \#Mul & \#Gates & size\\
+Dataset & $d$  & Prev. time (s) & Prev. Comm (kb) & LAN time (s) & WAN time (s) & Comm. (kb)  & \#And & \#Mul & \#Gates & size\\
 \hline
@@
  \caption{Linear classification benchmarks}
- \label{tab:lcd} 
+ \label{tab:lc} 
 \end{table*}
@@
 \begin{tabular}{c|c|c|c|c|c |c|c|c|c|c|c}
-Dataset & $|P|$ & $|x|$ & Prev. time (s) & Prev. Comm (Mb) & LAN time (s) & WAN time (s) & Comm. (Mb)  & \#And & \#Mul & \#Gates  & size\\
+Dataset & $n$ & $F$ & Prev. time (s) & Prev. Comm (Mb) & LAN time (s) & WAN time (s) & Comm. (Mb)  & \#And  & \#Gates  & size\\
 \hline
-Nursery & 5 & 9 & 1.5 & 0.2 & 0.1 & 0.4 & 0.6 & 13k & 0 & 73k & 50\\
+Nursery & 5 & 9 & 1.5 & 0.2 & 0.1 & 0.4 & 0.6 & 13k  & 73k & 50\\
 \hline
-Audiology & 24 & 70 & 3.9 & 2.0 & 1.5 & 2.9 & 37 & 750k & 0 & 4219k & 50\\
+Audiology & 24 & 70 & 3.9 & 2.0 & 1.5 & 2.9 & 37 & 750k & 4219k & 50\\
 \hline
@@
  \caption{Naive Bayes benchmarks}
- \label{tab:lcd} 
+ \label{tab:nb} 
 \end{table*}
@@
 \begin{table*}
-\begin{tabular}{c|c|c|c |c|c|c|c|c|c}
-Dataset  & Prev. time (s) & Prev. Comm (kb) & LAN time (s) & WAN time (s) & Comm. (kb)  & \#And & \#Mul & \#Gates & size\\
+\begin{tabular}{c|c|c|c |c|c|c|c|c|c|c}
+Dataset  & $N$ & Prev. time (s) & Prev. Comm (kb) & LAN time (s) & WAN time (s) & Comm. (kb)  & \#And & \#Mul & \#Gates & size\\
 \hline
-Nursery &  0.3 & 102 & 0.1 & 0.3 & 32 & 504 & 3 & 3324 & 20\\
+Nursery & 4 & 0.3 & 102 & 0.1 & 0.3 & 32 & 504 & 3 & 3324 & 20\\
 \hline
-ECG &  0.4 & 102 & 0.1 & 0.4 & 49 & 756 & 5 & 5002 & 20\\
+ECG &  6 & 0.4 & 102 & 0.1 & 0.4 & 49 & 756 & 5 & 5002 & 20\\
 \hline
@@
  \caption{Decision tree benchmarks}
- \label{tab:lcd} 
+ \label{tab:dt} 
 \end{table*}
@@
  \caption{Neural network benchmarks}
- \label{tab:lcd} 
+ \label{tab:nn} 
 \end{table*}
@@
  \caption{Tensorflow tutorial benchmarks}
- \label{tab:lcd} 
+ \label{tab:tf} 
 \end{table*}
 
-Finally, there are a class of machine learning models that are much more efficient than
-deep neural nets and provide good accuracies on standard learning tasks. One such model
-is \bonsai~\cite{bonsai}, which is a decision tree with more powerful nodes.
-In particular, each node $m_k$ evaluates a predicate $\theta_k^TZx > 0$ to decide whether
-to pass $x$ to the left branch or the right branch. and the predicted value is
-\[
-{\it argmax} \sum_k I_k(x)W_k^TZx\circ f(\sigma V_k^TZx) 
-\]
-Here, $I_k$ is 1 if the $k^{th}$ node is present on the path traversed by $x$
-and is zero otherwise. 
-The operation $\circ$ is a pointwise dot product and
-the function $f$ is given by $f(x) = x$ if $|x| < 1$ and $\mathrm{signum}(x)$ otherwise.
-We show results on three standard data sets: character  (Chars4k~\cite{campos}), text (USPS~\cite{hull}), and object categorization (WARD~\cite{yang}).
-There are other other benchmarks in~\cite{bonsai} that do not add any new insights here.
-Moreover, the dataset WARD requires the largest model.
+We show \bonsai results on three standard data sets: character  (Chars4k~\cite{campos}), text (USPS~\cite{hull}), and object categorization (WARD~\cite{yang}).
+There are other other benchmarks in~\cite{bonsai} that do not add any new insights and we do not report them here.
+We remark that the dataset WARD requires the largest model.
 These models are already over integers and no port from floating-point is required.
@@
  \caption{Bonsai benchmarks}
- \label{tab:lcd} 
+ \label{tab:bonsai} 
 \end{table*}
@@
 The results show the end-to-end results as well as the statistics for the heaviest stage.
+In contrast, \minion takes 544 seconds and 9272 Mb for this model. As before, we are faster than
+\minion in the LAN setting and slower in the WAN setting.
 
@@
  \caption{Pipelining results}
- \label{tab:lcd} 
+ \label{tab:cifar} 
 \end{table*}

### 2017-10-05 OaklandSubmission/eval.tex
eval pass

@@
 Each machine has a Intel(R) Xeon(R) CPU E5-2673 v3 processor running at 2.40GHz with
-28 GBs of RAM. Since most of our benchmarks are related to machine learning,
-we set up some notation standard to machine learning next.
+28 GBs of RAM. The \tool compiler is written in Python and compiles each of our benchmarks
+in under a second to ABY code~\cite{aby}. Since most of our benchmarks are related to machine learning, we set up some notation standard to machine learning next.
 
@@
 Some neural nets have convolution layers and these are called CNNs.
-For the purpose of this paper, a convolution can be considered as a (heavy) matrix-matrix multiplication. Therefore, fully connected layers are easier computation-wise compared
+For the purpose of this paper, a convolution can be considered as a (heavy) matrix-matrix multiplication.
+The size of matrices manipulated by convolution layers grows linearly with {\it window size} (typically 9 or 25) and the number of {\it output channels} (typically 16, 32, or 64).
+Therefore, fully connected layers are easier computation-wise compared
 to convolution layers. Such DNNs and CNNs are computationaly heavy but provide
@@
 We consider three types of classifiers: linear, Naive Bayes, and decision trees. 
-These classifiers are trained on the followinng data sets from the UCI  machine learning repository:
+These classifiers are trained on the followinng data sets from the UCI  machine learning repository~\cite{uci}:
  the Wisconsin Breast Cancer data set, 
@@
 
-\begin{table*}
+The results for linear classification are in Table~\ref{tab:lc}.
+The input and model are both vectors of length $d$.
+The time in columns ``Prev.
+time'' and communication ``Prev.Comm'' is the time and total network communication reported in~\cite{shafindss}.
+They remark that these times are obtained by mimicing a network with 40ms round trip time, which is the same as our WAN setting.
+The total execution time of \tool generated code in the LAN and the WAN setting is reported next followed by the total communication.
+The execution time is dominated by multiplications and boolean-ands.
+The number of boolean-and gates (\#And) remains constant for both the benchmnarks as they reflect the number of and gates required for a single 32-bit integer comparison with zero.
+The number of multiplication gates (\#Mul) is the same as the model size $d$.
+The last column shows the lines of code of \tool program, which is also independent of the model size.
+
+The results for Naive Bayes are in Table~\ref{tab:nb}. As before, $n$ denotes the length of the input feature vector $x=(x_1,\ldots,x_n)$ and $F$ is the number of possible values of each $x_i$.
+Although, there are no multiplications in these benchmarks, these do have a significant number of comparisons that result in many boolean-and gates (13k/750k denote 13000/75000). 
+Here, again columns 4 and 5 show the results reported in~\cite{shafindss} with 40ms network round trip time.
+We observe that \tool generated code has competitive performance on this task.
+This improvement is significant as~\cite{shafindss} uses custom designed protocols and we are executing a generic two party computation (2PC).
+Moreover,~\cite{shafindss} remarks that generic 2PC did not scale to the smallest of their Naive Bayes classifiers and they had to scale down the prediction task and there a Yao-based implementation was 500 times slower.
+We have shown that by using \tool, generic 2PC can scale to real prediction tasks and provide performance competitive with specialized protocols.
+
+This claim is further validated by Table~\ref{tab:dt}.
+In~\cite{wu}, Wu et al. describe specialized protocols for trees and forests (that perfrom better than~\cite{shafindss}) and we report their performance in columns 3 and 4 of Table~\ref{tab:dt}.
+Again, we observe that the performance of \tool generated generic 2PC code is competitive with specialized protocols.
+Therefore, we believe that \tool can be a good baseline to compare against and can aid the development of specialized protocols.
+
+\begin{table*}[ht]
 \begin{tabular}{c|c|c|c|c |c|c|c|c|c|c}
-Dataset & $d$  & Prev. time (s) & Prev. Comm (kb) & LAN time (s) & WAN time (s) & Comm. (kb)  & \#And & \#Mul & \#Gates & size\\
+Dataset & $d$  & Prev. time (s) & Prev. Comm (kb) & LAN time (s) & WAN time (s) & Comm. (kb)  & \#And & \#Mul & \#Gates & LOC\\
 \hline
@@
 
- \caption{Linear classification benchmarks}
+ \caption{Linear classification results}
  \label{tab:lc} 
@@
 
-\begin{table*}
+\begin{table*}[ht]
 \begin{tabular}{c|c|c|c|c|c |c|c|c|c|c|c}
-Dataset & $n$ & $F$ & Prev. time (s) & Prev. Comm (Mb) & LAN time (s) & WAN time (s) & Comm. (Mb)  & \#And  & \#Gates  & size\\
+Dataset & $n$ & $F$ & Prev. time (s) & Prev. Comm (Mb) & LAN time (s) & WAN time (s) & Comm. (Mb)  & \#And & \#Mul & \#Gates  & LOC\\
 \hline
-Nursery & 5 & 9 & 1.5 & 0.2 & 0.1 & 0.4 & 0.6 & 13k  & 73k & 50\\
+Nursery & 5 & 9 & 1.5 & 0.2 & 0.1 & 0.4 & 0.6 & 13k & 0 & 73k & 50\\
 \hline
-Audiology & 24 & 70 & 3.9 & 2.0 & 1.5 & 2.9 & 37 & 750k & 4219k & 50\\
+Audiology & 24 & 70 & 3.9 & 2.0 & 1.5 & 2.9 & 37 & 750k & 0 & 4219k & 50\\
 \hline
@@
 
- \caption{Naive Bayes benchmarks}
+ \caption{Naive Bayes results}
  \label{tab:nb} 
@@
 
-\begin{table*}
+\begin{table*}[ht]
 \begin{tabular}{c|c|c|c |c|c|c|c|c|c|c}
-Dataset  & $N$ & Prev. time (s) & Prev. Comm (kb) & LAN time (s) & WAN time (s) & Comm. (kb)  & \#And & \#Mul & \#Gates & size\\
+Dataset  & $N$ & Prev. time (s) & Prev. Comm (kb) & LAN time (s) & WAN time (s) & Comm. (kb)  & \#And & \#Mul & \#Gates & LOC\\
 \hline
@@
 \subsection{Neural networks}
-We evaluate on the benchmarks described in \minion~\cite{minionn}.
-The first is the neural network described in SecureML~\cite{secureml}.
+We evaluate \tool on the DNNs described in \minion~\cite{minionn} and the results are reported in Table~\ref{tab:nn}\footnote{~\cite{minionn} do not report the network round-trip time.}.
+Our goal here is to demonstrate that \tool
+can provide performance competitive to specialized protocols for DNNs described in~\cite{secureml,cryptonets,minionn}. 
+The first benchmark  is the DNN described in SecureML~\cite{secureml}.
 It has three fully connected layers with square as the activation function.
+This DNN is light-weight as it requires only arithmetic computations (additions and multiplications)
+except for the {\it argmax} at the end.
 Next, we implement the neural network of Cryptonets~\cite{cryptonets} in \tool.
-This network also uses square as the activation function and has one convolution (with 5 output channels), one maxpool, and one fully connected layer.
-Finally, we compare against\minion on a convolutional neural network
-with two convolutions (with 16 output channels), two maxpool, and two fully connected layers.
-In contrast to the previous two, this network uses ReLU for activation.
-In our results, we only compare with \minion, as it outperforms the earlier systems of
-Cryptonets and SecureML.
+This network also uses square as the activation function and has one convolution (with 5 output channels) and one fully connected layer.
+Finally, we compare against\minion on a CNN
+with two convolutions (with 16 output channels) and two fully connected layers.
+In contrast to the previous two DNNs, this DNN uses ReLU for activation and
+has significantly higher number of boolean-and gates.
+Recall, that square activation can be implemented using arithmetic gates but ReLU requires boolean-and gates.
+For a complete description of these benchmarks,  the reader is referred to the original references.
+
+For the benchmarks in Table~\ref{tab:nn}, \minion
+outperforms SecureML and Cryptonets and columns 2 and 3 show the time taken by \minion.
+We observe that our performance is competitive with \minion in the LAN and the WAN settings and the lines of code required are still small.
+However, \minion also reports performance results on a neural network with 7 convolution layers.
+\tool does not scale to this network without pipelining and we discuss this network in Section~\ref{sec:pipeeval}.
 %net | minionT | minionC | timeL | timeW | comm | gates | and | mul | length
@@
 \begin{tabular}{c|c|c|c |c|c|c|c|c|c}
-Name  & Prev. time (s) & Prev. Comm (Mb) & LAN time (s) & WAN time (s) & Comm. (Mb)  & \#And & \#Mul & \#Gates & size\\
+Name  & Prev. time (s) & Prev. Comm (Mb) & LAN time (s) & WAN time (s) & Comm. (Mb)  & \#And & \#Mul & \#Gates & LOC\\
 \hline
@@
 Tensorflow~\cite{tensorflow} is a standard machine learning toolkit.
-Its tutorial describes two prediction models for handwritten digit recognition
-using the MNIST dataset~\cite{mnist} which has 55,000 images as training data,
-10,000 images as test data (mnist.test), and 5,000 images as validation data.
-Each image is a greyscale image of size $28\times 28$.
-The first model that the tutorial describes is a softmax regression
+Its introductory tutorial describes two prediction models for handwritten digit recognition
+using the MNIST dataset~\cite{mnist}.
+Each image in this dataset is a greyscale image of digits 0 to 9 of size $28\times 28$.
+The first model that the tutorial describes is softmax regression
 which provides an accuracy of 92\%. Since this model is over floating-point,
-as is standard~\cite{secureml,minionn}, we port it to a model over 32-bit integers manually. After our port, the accuracy is still preserved.
-Next, we implement this model in \tool and the results are shown below.
+as is standard (\cite{secureml,minionn}), we port it to a model over 32-bit integers manually. After our port, the prediction accuracy is still preserved.
+Next, we implement this ported model in \tool and the results are shown in the first row of Table~\ref{tab:tf}.
+Since, we are not familiar with any other tools that have used this model as a benchmark,
+the columns ``Prev. Time'' and ``Prev. Comm.'' are missing.
+The column ``Model size'' is the number of 32 bit integers in the trained model.
 
 The next model in the Tensorflow tutorial is a convolution neural net with two convolutions
-(with 32 output channels), two maxpool, and two fully connected layers.
+(with 32 output channels) and two fully connected layers with RELU as the activation function.
 This network is both bigger and more accurate than the networks presented in the previous section.
@@
 we port it to integers while maintaining the accuracy\footnote{The accuracy of the integer model is 0.04\% higher than the floating-point model.}.
+We observe that this CNN can take a minute per prediction in the WAN setting and is the largest
+benchmark that we have evaluated without pipelining.
 
@@
 \begin{tabular}{c|c|c|c |c|c|c|c|c|c | c}
-Name       & LAN time (s) & WAN time (s) & Comm. (Mb)  & \#And & \#Mul & \#Gates & Model size & Program size\\
+Name       & LAN time (s) & WAN time (s) & Comm. (Mb)  & \#And & \#Mul & \#Gates & Model size & LOC\\
 \hline
@@
 
-We show \bonsai results on three standard data sets: character  (Chars4k~\cite{campos}), text (USPS~\cite{hull}), and object categorization (WARD~\cite{yang}).
-There are other other benchmarks in~\cite{bonsai} that do not add any new insights and we do not report them here.
-We remark that the dataset WARD requires the largest model.
-These models are already over integers and no port from floating-point is required.
+Before discussing pipelining, we show \bonsai results on three standard data sets: character  (Chars4k~\cite{campos}, accuracy 74.71\%) recognition, text (USPS~\cite{hull}, accuracy 94.4\%) recognition, and object categorization (WARD~\cite{yang}, accuracy 95.7\%).
+The \bonsai models are already over integers and no port from floating-point is required.
+We implement the trained models in \tool for all benchmarks in~\cite{bonsai}
+and show a subset of our results in Table~\ref{tab:bonsai}.
+In particular, we do not report the results that do not add any additional insight.
+Out of all benchmarks in~\cite{bonsai}, the dataset WARD requires the largest model.
+The column depth shows the depth of the tree used by \bonsai. The size of \tool
+program grows with the depth of the tree as the \tool implementation requires the
+loop over nodes of the tree to be unrolled. 
 %dataset | size | gates | and | mul | depth | length | comm | timeL | timeW
@@
 \begin{tabular}{c|c|c|c |c|c|c|c|c|c | c}
-Name       & LAN time (s) & WAN time (s) & Comm. (Mb)  & \#And & \#Mul & \#Gates & depth & Program size\\
+Name       & LAN time (s) & WAN time (s) & Comm. (Mb)  & \#And & \#Mul & \#Gates & depth & LOC\\
 \hline
@@
 \subsection{Pipelining}
-The largest benchmark of \minion is a neural network for CIFAR-10 dataset~\cite{cifar}: label colored images for objects in 10 classes. This network needs more memory than what is available
+\label{sec:pipeeval}
+The largest benchmark of \minion is a DNN for CIFAR-10 dataset~\cite{cifar}.
+The predictor's task is to label colored ($32\times 32$) images for objects in 10 classes. A secure evaluation of this DNN needs more memory than what is available
 on our machines. Therefore, we use pipelining and divide the computation into seven stages.
 The first step does a convolution with 64 output channels and a ReLU activation.
-The next four stages perform a convolution that involves multiplying a $64\times 576$ matrix with a
-$576\times 1024$ matrix. The next stage performs a ReLU, a meanpool, and a convolution.
-The final stage performs four convolutions, a mean pool, five ReLU and a fully connected layer.
-The results show the end-to-end results as well as the statistics for the heaviest stage.
-In contrast, \minion takes 544 seconds and 9272 Mb for this model. As before, we are faster than
-\minion in the LAN setting and slower in the WAN setting.
+The next four stages together perform a convolution that involves multiplying a $64\times 576$ matrix with a
+$576\times 1024$ matrix. The next stage performs a ReLU and a convolution.
+The final stage has four convolutions, five ReLUs, and a fully connected layer.
+The total number of lines of \tool code for this benchamrk is 336 lines.
 
+The results in Table~\ref{tab:cifar} show the end-to-end results as well as the statistics for the heaviest stage (the sixth stage).
+The number of gates are in millions, hence the suffix 'm' in the last three columns.
+For this DNN, \minion takes 544 seconds and communicates 9272 Mb. 
+Like the previous comparisons in Table~\ref{tab:nn}, \tool generated implementations are competitive with \minion here as well. 
 \begin{table*}

### 2017-10-05 OaklandSubmission/intro.tex
initial intro

@@
 Secure multiparty computation \cite{yao,gmw} (SMC) is a powerful cryptographic tool that allows a set of mutually distrusting parties to compute a joint function of their secret inputs.
-Examples of these include running machine learning prediction algorithm where one party holds the secret medical report and the other party holds the model for desease.
-Since the introduction of SMC in 1980's, there has been a large body of work \cite{..} that has transformed SMC from a mere theoretical tool to something that can be used in practice.
+Examples of these include various machine learning prediction algorithm where one party holds the secret input such as a medical report and the other party holds the ML model for disease prediction.
+Since the introduction of SMC in 1980's, there has been a large body of work \cite{..} that has transformed SMC from a mere theoretical tool to something practically efficient/usable.
 
-Unfortunately, implementing fuctionalities using SMC protocols requires thorogh understanding of cryptography.
+Unfortunately, implementing functionalities using SMC protocols requires thorough understanding of cryptography.
 To allow for widespread use of SMC, it is critical that SMC protocols are programmable by non-cryptographic experts.
 To cater to this, there have been several efforts developing domain-specific languages that are programmer friendly and compile to a SMC back-end.
-Most notable of these systems include \cite{...}.
-A common feature of all of these works is that they use a boolean circuit based cryptographic backend such as garbled circuits.
-As one can expect, compiling a function to a boolean circuit is one of the biggest efficiency bottleneck.
+Most notable of these domain specific languages and systems include \cite{...}.
+All of these works build on a cryptographic back-end that is either entirely boolean \cite{yao,gmw} or arithmetic \cite{homo}.
+In fact, most of these works use boolean circuits based scheme for completeness.
+However, most interesting functions require a mix of arithmetic and boolean computations.
+Examples include .........
+As one can expect, expressing these computations as boolean circuits is one of the biggest efficiency bottleneck.
+\divya{Note that comparisons cannot be expressed in arithmetic circuits}
 
-Recent works \cite{aby,secureml,minion} have shown that great performance benefit can be obtained by mixing arithmetic and boolean computations in the back-end.
+To address this issue, recently Demmler et al.
+\cite{aby} gave a framework for SMC where the parties can mix arithmetic and boolean computations and demonstrated great performance benefits \divya{on small examples}.
+However, using their system requires the programmer to be aware of trade-offs of arithmetic and boolean cryptographic schemes.
+The programmer writes circuits consisting of a mix of arithmetic and boolean gates along with appropriate conversion gates.
+In short, as is also mentioned by the authors themselves, \cite{aby} is not suitable to be used by non-specialist programmers.
+
+%Recent works \cite{aby,secureml,minion} have shown that great performance benefit can be obtained by mixing arithmetic and boolean computations in the back-end.
 Certain computations such as integer multiplication are efficient when implemented using arithmetic SMC \cite{gmw} whereas max(x,0) needs to be computed using a boolean back-end \cite{yao}.
@@
 In this work, we develop and implement our framework \tool where the programmer writes a high level C-like program (instead of a circuit) to describe the function to be computed.
-The programmer is oblivious of the cryptographic backend being used.
+Our compiler automatically converts this program to a suitable SMC protocol that uses both arithmetic and boolean secure computation components.
+Our framework can work with any secure implementation of a mix of arithmetic and boolean computations.
+In our work, we focus on semi-honest secure two-party computation and  ABY \cite{aby} provides the suitable cryptographic back-end.
+\divya{We provide a formal type system....
+} We evaluate our system on benchmarks from \cite{} and show that automatically generated protocols match the performance of hand-crafted protocols in most cases (see \ref{sec:eval}).
+\divya{Below, we will  give an overview of \tool and describe our contributions.}
+
+%The programmer is oblivious of the cryptographic back-end being used.
 Our compiler automatically compiles it to a circuit framework consisting to both arithmetic and boolean gates as well share-conversion gates wherever required.
@@
 
-5. can serve as a baseline for future works on hand-crafted algorithms for specific applications
+5. can serve as a baseline for future works on hand-crafted algorithms for specific applications cite minion etc here...
 
@@
 \item Secure Computation of MIPS Machine Code %https://eprint.iacr.org/2015/547
+\item TASTY https://eprint.iacr.org/2010/365.pdf
 
 \end{enumerate}
+
+There have been many languages and compilers that compile to cryptographic backends
+exception of TASTY
 \newpage

### 2017-10-05 OaklandSubmission/eval.tex
minor

@@
 and the model is a vector $w\in\R^d$. The possible labels $\ell\in\{\mathit{true},\mathit{false}\}$
-and the classifier $C_w\equiv w^Tx>0$, where $T$ denotes the matrix transpose operator.
+and the classifier $C_w\equiv w^Tx>0$, where $\cdot^T$ denotes the matrix transpose operator.
 This classifier requires $n$ multiplications, $n-1$ additions, and a comparison.
@@
 This process is repeated till we reach a leaf.
-The leaves of the tree have labels and the output label is the leaf visited by this traversal.
+The leaves of the tree are labels and the output label is the leaf visited by this traversal.
 Such a (binary) decision tree can be encoded as a polynomial and then prediction
@@
 These layers multiply a matrix with a vector.
-Some neural nets have convolution layers and these are called CNNs.
+Some neural nets have convolution layers and such DNNs are also called CNNs.
 For the purpose of this paper, a convolution can be considered as a (heavy) matrix-matrix multiplication.
-The size of matrices manipulated by convolution layers grows linearly with {\it window size} (typically 9 or 25) and the number of {\it output channels} (typically 16, 32, or 64).
+The size of matrices manipulated by a convolution layer grows linearly with {\it window size} (typically 9 or 25), the number of {\it output channels} (typically 16, 32, or 64),
+and the size of the matrix input to this layer. 
 Therefore, fully connected layers are easier computation-wise compared

### 2017-10-06 OaklandSubmission/intro.tex
intro

@@
 
-Secure multiparty computation \cite{yao,gmw} (SMC) is a powerful cryptographic tool that allows a set of mutually distrusting parties to compute a joint function of their secret inputs.
+Most cryptographic tools rely on circuit as abstractions for any general computation but real-life developers write programs.
+It is critical to address this fundamental  representation gap to make cryptography accessible to non-crypto specialist developers.
+In this work, we address this gap for the case of secure multi-party computation.
+
+Secure multi-party computation \cite{yao,gmw} (SMC) is a powerful cryptographic tool that allows a set of mutually distrusting parties to compute a joint function of their secret inputs.
 Examples of these include various machine learning prediction algorithm where one party holds the secret input such as a medical report and the other party holds the ML model for disease prediction.
@@
 To allow for widespread use of SMC, it is critical that SMC protocols are programmable by non-cryptographic experts.
-To cater to this, there have been several efforts developing domain-specific languages that are programmer friendly and compile to a SMC back-end.
+To cater to this, there have been several efforts developing domain-specific languages that are programmer friendly and compile to a SMC protocol.
 Most notable of these domain specific languages and systems include \cite{...}.
 All of these works build on a cryptographic back-end that is either entirely boolean \cite{yao,gmw} or arithmetic \cite{homo}.
-In fact, most of these works use boolean circuits based scheme for completeness.
+In fact, most of these works use boolean circuits based scheme for completeness\footnote{Note that comparisons cannot be expressed in arithmetic circuits}.
 However, most interesting functions require a mix of arithmetic and boolean computations.
 Examples include .........
-As one can expect, expressing these computations as boolean circuits is one of the biggest efficiency bottleneck.
-\divya{Note that comparisons cannot be expressed in arithmetic circuits}
+As one can expect, compiling these programs to boolean circuits is one of the biggest efficiency bottlenecks.
+\divya{Our work addresses this performance bottleneck.}
 
-To address this issue, recently Demmler et al.
-\cite{aby} gave a framework for SMC where the parties can mix arithmetic and boolean computations and demonstrated great performance benefits \divya{on small examples}.
+To address this issue \divya{on the cryptographic side}, recently Demmler et al.
+\cite{aby} gave a cryptographic protocol for SMC where the parties can mix arithmetic and boolean computations and claimed potentially great performance benefits.
 However, using their system requires the programmer to be aware of trade-offs of arithmetic and boolean cryptographic schemes.
-The programmer writes circuits consisting of a mix of arithmetic and boolean gates along with appropriate conversion gates.
-In short, as is also mentioned by the authors themselves, \cite{aby} is not suitable to be used by non-specialist programmers.
+In their framework, the programmer is required to write circuits consisting of a mix of arithmetic and boolean gates along with appropriate conversion gates.
+In short, as is also mentioned by the authors themselves, the current system is not suitable to be used by non-specialist programmers.
 
@@
 In this work, we develop and implement our framework \tool where the programmer writes a high level C-like program (instead of a circuit) to describe the function to be computed.
-Our compiler automatically converts this program to a suitable SMC protocol that uses both arithmetic and boolean secure computation components.
-Our framework can work with any secure implementation of a mix of arithmetic and boolean computations.
+Our compiler automatically converts this program to a suitable SMC protocol that uses both arithmetic and boolean secure computation components as well as the correct conversion protocols whenever required.
+Our compiler is general and can work with any secure implementation of a mix of arithmetic and boolean computations.
 In our work, we focus on semi-honest secure two-party computation and  ABY \cite{aby} provides the suitable cryptographic back-end.
 \divya{We provide a formal type system....
-} We evaluate our system on benchmarks from \cite{} and show that automatically generated protocols match the performance of hand-crafted protocols in most cases (see \ref{sec:eval}).
+} We evaluate our system on various benchmarks from \cite{} and show that automatically generated protocols match the performance of hand-crafted protocols in most cases (see \ref{sec:eval}).
 \divya{Below, we will  give an overview of \tool and describe our contributions.}
@@
 \divya{ideally have small paragraphs about key features}
-1. ease to programming.. refer to section with example in our language; contrast to verilog paper; programmer is completely oblivious of security; looks like a normal C program
+1. ease to programming.. refer to section with example in our language; contrast to verilog paper; programmer is completely oblivious of security; looks like a normal C program;
+also say that user does not write keywords secret and public; these are inferred from the information flow
 

### 2017-10-06 OaklandSubmission/eval.tex
first pass

@@
 \label{sec:eval}
-
+We evaluate \tool for the problem of {\it secure prediction}, where a server has a machine learning model, a client has an input, and after the protocol execution the client learns a label on the input.
+From the protocol execution, the server learns nothing about the client input and the client learns nothing about the server model than what is revealed beyond the computed label.
+This setting has been previously studied in~\cite{shafindss,minionn}
 We first implement benchmarks of ~\cite{shafindss} and~\cite{minionn} in \tool to compare gainst 
@@
 28 GBs of RAM. The \tool compiler is written in Python and compiles each of our benchmarks
-in under a second to ABY code~\cite{aby}. Since most of our benchmarks are related to machine learning, we set up some notation standard to machine learning next.
+in under a second to C++ code that makes calls to the ABY library~\cite{aby}. Since most of our benchmarks are related to machine learning, we set up some notation standard to machine learning next.
 
@@
 In this paper, we focus on machine learning models for {\it classification}.
-A {\it classifier} $C$ uses a trained model to predict a {\it label} $\ell$ for an input data point $x$. For example, given a data point which is a tuple of height and weight of an individual,
+A {\it classifier} $C$ uses a trained {\it model} to {\it predict} a {\it label} $\ell$ for an input data point $x$.
+For example, given a data point which is a tuple of height and weight of an individual,
 a classifier can predict a label ``male'' or ``female''.
@@
 and the classifier $C_w\equiv w^Tx>0$, where $\cdot^T$ denotes the matrix transpose operator.
-This classifier requires $n$ multiplications, $n-1$ additions, and a comparison.
+This classifier requires $d$ multiplications, $d-1$ additions, and a comparison.
 
@@
 
-A decision tree of size $N$ takes as input $x\in\R^n$ and the model consists of a boolean predicate
-assigned to each internal node of the tree.
+A decision tree of size $N$ takes as input $x\in\R^n$ and the model consists of a tree with a boolean predicate
+assigned to each internal node.
 In particular, a node $j\in[N]$ at depth $i$ has a predicate $b_{j}^i\equiv x_i\leq w_i$.
@@
 reduces to polynomial evaluation. E.g., a binary tree with depth one and size $N=3$ can be encoded as the polynomial
-$(1-b(x))\ell_1 + b(x)\ell_2$ where $b(x)$ is the predicate at the root, $\ell_1$ is the label of the left leaf and $\ell_r$ is the label of the right leaf.
+$(1-b(x))\ell_1 + b(x)\ell_2$ where $b(x)$ is the predicate at the root, $\ell_1$ is the label of the left leaf and $\ell_2$ is the label of the right leaf.
 
@@
 Therefore, fully connected layers are easier computation-wise compared
-to convolution layers. Such DNNs and CNNs are computationaly heavy but provide
+to convolution layers. In general, DNNs are computationaly heavy but provide
 much better accuracies on computer vision tasks than the classifiers discussed above.
 
-Finally, there are a class of machine learning models that are much more efficient than
-deep neural nets and provide reasonably good accuracies on standard learning tasks. One such model
+Finally, there are a class of machine learning classifiers that are much more efficient than
+DNNs and provide reasonably good accuracies on standard learning tasks. One such classifer
 is \bonsai~\cite{bonsai}, which is a variant of the decision trees described above.
-In particular, each node $m_k$ evaluates a predicate $\theta_k^TZx > 0$ to decide whether
-to pass $x$ to the left branch or the right branch ($\theta_k$ is a vector and $Z$ is a matrix).
+In particular, each node $j$ evaluates a predicate $\theta_j^TZx > 0$ to decide whether
+to pass $x$ to the left branch or the right branch ($\theta_j$ is a vector and $Z$ is a matrix).
 The predicted value is
 \[
-{\it argmax} \sum_k I_k(x)W_k^TZx\circ f(V_k^TZx) 
+{\it argmax} \sum_j I_j(x)W_j^TZx\circ f(V_j^TZx) 
 \]
-Here, $I_k$ is 1 if the $k^{th}$ node is present on the path traversed by $x$
+Here, $I_j$ is 1 if the $j^{th}$ node is present on the path traversed by $x$
 and is zero otherwise. 
@@
 In the following, we implement these classifiers in \tool and report the time taken for making predictions securely.
+Ideally, the machine learning classifiers are mathematical expressions over Reals that are usually approximated by floating-point operations.
+As is standard, we port the classifers to integer manipulating programs by scaling the models and rounding~\cite{minionn}. These ported classifiers are then implemented in \tool.
 
@@
 time'' and communication ``Prev.Comm'' is the time and total network communication reported in~\cite{shafindss}.
-They remark that these times are obtained by mimicing a network with 40ms round trip time, which is the same as our WAN setting.
+They remark that these times are obtained by simulating a network with 40ms round trip time, which is the same as our WAN setting.
 The total execution time of \tool generated code in the LAN and the WAN setting is reported next followed by the total communication.
@@
 Although, there are no multiplications in these benchmarks, these do have a significant number of comparisons that result in many boolean-and gates (13k/750k denote 13000/75000). 
-Here, again columns 4 and 5 show the results reported in~\cite{shafindss} with 40ms network round trip time.
+Here, again columns 4 and 5 show the results reported in~\cite{shafindss}.
 We observe that \tool generated code has competitive performance on this task.
@@
 In~\cite{wu}, Wu et al. describe specialized protocols for trees and forests (that perfrom better than~\cite{shafindss}) and we report their performance in columns 3 and 4 of Table~\ref{tab:dt}.
+Similar to~\cite{shafindss},~\cite{wu} remarks that these computations are infeasible with Yao-based 2PC.
 Again, we observe that the performance of \tool generated generic 2PC code is competitive with specialized protocols.
-Therefore, we believe that \tool can be a good baseline to compare against and can aid the development of specialized protocols.
+Therefore, we believe that \tool generated code can be a better baseline than Yao-based 2PC when comparing against generic 2PC-based implementations for secure prediction tasks.
 
-\begin{table*}[ht]
+\begin{table*}
 \begin{tabular}{c|c|c|c|c |c|c|c|c|c|c}
@@
 
-\begin{table*}[ht]
+\begin{table*}
 \begin{tabular}{c|c|c|c|c|c |c|c|c|c|c|c}
@@
 
-\begin{table*}[ht]
+\begin{table*}
 \begin{tabular}{c|c|c|c |c|c|c|c|c|c|c}
@@
 %dataset | time | comm | timeL | timeW | comm
-\subsection{Neural networks}
-We evaluate \tool on the DNNs described in \minion~\cite{minionn} and the results are reported in Table~\ref{tab:nn}\footnote{~\cite{minionn} do not report the network round-trip time.}.
+\subsection{Deep neural nets}
+We evaluate \tool on the DNNs described in \minion~\cite{minionn} and the results are reported in Table~\ref{tab:nn}\footnote{\minion does not report the network round-trip time~\cite{minionn}.}.
 Our goal here is to demonstrate that \tool
@@
 except for the {\it argmax} at the end.
-Next, we implement the neural network of Cryptonets~\cite{cryptonets} in \tool.
-This network also uses square as the activation function and has one convolution (with 5 output channels) and one fully connected layer.
-Finally, we compare against\minion on a CNN
-with two convolutions (with 16 output channels) and two fully connected layers.
+Next, we implement the DNN described in Cryptonets~\cite{cryptonets} in \tool.
+This DNN also uses square as the activation function and has one convolution (with 5 output channels) and one fully connected layer.
+Finally, we compare against \minion on a CNN
+with two convolutions (with 16 output channels each) and two fully connected layers.
 In contrast to the previous two DNNs, this DNN uses ReLU for activation and
 has significantly higher number of boolean-and gates.
-Recall, that square activation can be implemented using arithmetic gates but ReLU requires boolean-and gates.
+Recall, that square activation can be implemented entirely using arithmetic gates but ReLU requires boolean-and gates.
 For a complete description of these benchmarks,  the reader is referred to the original references.
@@
 We observe that our performance is competitive with \minion in the LAN and the WAN settings and the lines of code required are still small.
-However, \minion also reports performance results on a neural network with 7 convolution layers.
-\tool does not scale to this network without pipelining and we discuss this network in Section~\ref{sec:pipeeval}.
+We note that \minion and \tool implementations are both based on ABY.
+These results are surprising as \minion has a much more efficient preprocessing phase and has SIMD
+(single instruction multiple data) capabilities that allow \minion to perform
+matrix operations efficiently. Our language and compiler are generic and we have none of these.
+Although we could have used the SIMD capabilities of ABY to improve our performance---our preliminary evaluation suggests that SIMD could lead to a 2X performance improvement---we decided against it to avoid being closely tied to ABY and keeping \tool general, i.e., oblivious to the 2PC backend.
+However, \minion also reports performance results on a bigger DNN with 7 convolution layers.
+This benchmark requires pipelining and we discuss it in Section~\ref{sec:pipeeval}.
 %net | minionT | minionC | timeL | timeW | comm | gates | and | mul | length
@@
 
- \caption{Neural network benchmarks}
+ \caption{DNN benchmarks}
  \label{tab:nn} 
@@
 using the MNIST dataset~\cite{mnist}.
-Each image in this dataset is a greyscale image of digits 0 to 9 of size $28\times 28$.
+Each image in this dataset is a greyscale $28\times 28$ image of digits 0 to 9.
 The first model that the tutorial describes is softmax regression
-which provides an accuracy of 92\%. Since this model is over floating-point,
-as is standard (\cite{secureml,minionn}), we port it to a model over 32-bit integers manually. After our port, the prediction accuracy is still preserved.
-Next, we implement this ported model in \tool and the results are shown in the first row of Table~\ref{tab:tf}.
+that provides an accuracy of 92\%. The classifier here evaluates
+\[
+ \mathit{argmax}_{i} w_i^Tx+b_i
+\]
+Here, $i\in[0,9]\cap\mathbb{Z}$, $x$ is a 784 length vector obtained from the input image,
+each $w_i$ is a 784 length vector, and each $b_i$ is a scalar. 
+We implement this classifer in \tool and the results are shown in the first row of Table~\ref{tab:tf}.
 Since, we are not familiar with any other tools that have used this model as a benchmark,
-the columns ``Prev. Time'' and ``Prev. Comm.'' are missing.
+we only report execution time of \tool generated code.
 The column ``Model size'' is the number of 32 bit integers in the trained model.
 
-The next model in the Tensorflow tutorial is a convolution neural net with two convolutions
-(with 32 output channels) and two fully connected layers with RELU as the activation function.
-This network is both bigger and more accurate than the networks presented in the previous section.
-In particular, it has an accuracy of 99.2\%. Again, since this model is in floating-point,
-we port it to integers while maintaining the accuracy\footnote{The accuracy of the integer model is 0.04\% higher than the floating-point model.}.
-We observe that this CNN can take a minute per prediction in the WAN setting and is the largest
+The next classifer in the Tensorflow tutorial is a convolution neural net with two convolutions
+(with 32 output channels) and two fully connected layers with ReLU as the activation function.
+This DNN is both bigger and more accurate than the DNNs presented in the previous section.
+In particular, it has an accuracy of 99.2\%\footnote{The accuracy of the integer model is 0.04\% higher than the floating-point model.}.
+We observe that this DNN can take a minute per prediction in the WAN setting and is the largest
 benchmark that we have evaluated without pipelining.
@@
 
-Before discussing pipelining, we show \bonsai results on three standard data sets: character  (Chars4k~\cite{campos}, accuracy 74.71\%) recognition, text (USPS~\cite{hull}, accuracy 94.4\%) recognition, and object categorization (WARD~\cite{yang}, accuracy 95.7\%).
+Before discussing pipelining, we show \bonsai results on three standard datasets: character recognition (Chars4k~\cite{campos}, accuracy 74.71\%), text recognition (USPS~\cite{hull}, accuracy 94.4\%), and object categorization (WARD~\cite{yang}, accuracy 95.7\%).
 The \bonsai models are already over integers and no port from floating-point is required.
-We implement the trained models in \tool for all benchmarks in~\cite{bonsai}
-and show a subset of our results in Table~\ref{tab:bonsai}.
-In particular, we do not report the results that do not add any additional insight.
+We implement the trained classifiers in \tool for all benchmarks in~\cite{bonsai}
+and show the represenative results in Table~\ref{tab:bonsai}.
 Out of all benchmarks in~\cite{bonsai}, the dataset WARD requires the largest model.
-The column depth shows the depth of the tree used by \bonsai. The size of \tool
-program grows with the depth of the tree as the \tool implementation requires the
-loop over nodes of the tree to be unrolled. 
+The column ``depth'' in Table~\ref{tab:bonsai} shows the depth of the tree used by \bonsai. The size of \tool
+program grows with the depth of the tree as the straightforward \tool implementation requires a
+loop over the nodes of the tree and this loop needs to be unrolled.
+
+To summarize, we have demonstrated that \tool is expressive enough to implement various
+state-of-the-art classifiers. We discuss scalability next. 
 %dataset | size | gates | and | mul | depth | length | comm | timeL | timeW
@@
 \label{sec:pipeeval}
-The largest benchmark of \minion is a DNN for CIFAR-10 dataset~\cite{cifar}.
-The predictor's task is to label colored ($32\times 32$) images for objects in 10 classes. A secure evaluation of this DNN needs more memory than what is available
+The largest benchmark of \minion~\cite{minionn} is a DNN for CIFAR-10 dataset~\cite{cifar}.
+The classifier's task is to categorize colored ($32\times 32$) images into 10 classes. A secure evaluation of this DNN needs more memory than what is available
 on our machines. Therefore, we use pipelining and divide the computation into seven stages.
@@
 The next four stages together perform a convolution that involves multiplying a $64\times 576$ matrix with a
-$576\times 1024$ matrix. The next stage performs a ReLU and a convolution.
+$576\times 1024$ matrix. The sixth stage performs a ReLU and a convolution.
 The final stage has four convolutions, five ReLUs, and a fully connected layer.
@@
 
-The results in Table~\ref{tab:cifar} show the end-to-end results as well as the statistics for the heaviest stage (the sixth stage).
-The number of gates are in millions, hence the suffix 'm' in the last three columns.
+The results in Table~\ref{tab:cifar} show the end-to-end results as well as the statistics for the sixth stage (the heaviest stage). 
+The number of gates are in millions, hence the suffix `m' in the last three columns.
 For this DNN, \minion takes 544 seconds and communicates 9272 Mb. 
-Like the previous comparisons in Table~\ref{tab:nn}, \tool generated implementations are competitive with \minion here as well. 
-\begin{table*}
+Like the results in Table~\ref{tab:nn}, \tool generated implementations are competitive with \minion here as well.
+Therefore, we believe that (with pipelining) \tool generated generic 2PC implementations can scale to arbitrary sized computations  while maintaining performance
+competitive with existing specialized protocols for secure prediction.
+In particular, for a large enough DNN, \minion would run out of memory
+but a pipelined \tool implementation would still succeed.
+\begin{table}
 \begin{tabular}{c|c|c|c |c|c|c}
-           &  LAN time (s) & WAN time (s) & Comm. (Mb)  & \#And & \#Mul & \#Gates \\
+           &  LAN (s) & WAN (s) & Comm. (Mb)  & \#And & \#Mul & \#Gates \\
 \hline
@@
  \label{tab:cifar} 
-\end{table*}
+\end{table}
 

### 2017-10-07 OaklandSubmission/pipe.tex
pipelining first draft.

\section{Pipelining}
\label{sec:pipe}
\nc{Started high level description. Can go deeper and correct notation as we proceed.}
We now describe our technique of ``pipelining'' that allows us to compile and execute large programs that would otherwise have not fitted in memory (due to the large size of the circuit generated by ABY). We do this securely in a manner such that the underlying backend (such as ABY in this case) does not have to modified in anyway and hence our technique is also compatible with any backend. In more detail, let $P$ be a program in our source language that gets compiled to $P'$ in our target language. In many cases, the ABY circuit generated for $P'$ is very large and cannot fit in memory (in fact, there is an upper limit of $2^{32}$ for the circuit size in ABY and for many natural programs, this upper limit is easily crossed). Without pipelining, we would be unable to compile and execute such a program $P$. In order to overcome this size limit, we do the following: we split the program $P$ into a sequence of smaller programs $Q_1||Q_2||\cdots||Q_k$ (that will be appropriately defined). We compile and execute each $Q_i$ separately, feeding the outputs of one program as state information to the next program. Of course, we have to be careful how we do this so as to not compromise the security of the protocol. More details follow.

Let $P$ be a program that takes as (secret) inputs $x$ and $y$ and produces output $z$ to both parties. Let $P_1||P_2||\cdots||P_k$ be a sequence of programs defined such that the following holds. Define $s_0 = \bot$ (the public empty state). For all $1\leq i\leq k-1$, $P_i$ takes as input $x, y$ and $s_{i-1}$ to produce state information $s_i$. Finally, $P_k$ takes as input $x,y$ and $s_{k-1}$ to produce output $z$. Every program $P_i$ ($1\leq i\leq k$) will be required to be ``sufficiently'' small (as described later) and this will enable our pipelining technique. It is easy to see that any program $P$ can be decomposed as illustrated above into such $P_1||P_2||\cdots||P_k$. Now, if one were to compile $P_i$ into $P'_i$, and execute $P'_i$, the parties can execute $P'_1||P'_2||\cdots||P'_k$ sequentially to obtain $s_1,\cdots,s_{k-1},$ and finally output $z$. However, doing so would enable the parties to learn $s_i$ (for all $1\leq i\leq k-1$), which would completely break the security of the protocol.

To overcome this problem, we define a sequence of new programs $Q_i$ ($1\leq i\leq k$) as follows. Once again, define $s_0 = \bot$. Let all $s_i$ be values in some additive ring $(\mathbb{Z},+)$ (in the ABY implementation this ring would simply be the additive ring $(\mathbb{Z}_{2^{32}},+)$ i.e., the additive ring of integers modulo $2^{32}$). This is without loss of generality true (we will also for simplicity assume that each $s_i$ is a single element in the ring, but this restriction can also be easily removed). Let $r_1,\cdots,r_{k-1}$ be a sequence of random values sampled from the same ring $(\mathbb{Z},+)$ by Party A. Let $Q_1$ be the program that takes as input $x,y, r_1$ (and empty state $s_0$), runs $P_1$ (as defined above) to get $s_1$ and then produces output $o_1 = s_1 + r_1$ {\em only to} Party B\footnote{While the description of our protocol here assumes that the underlying backend supports only one party receiving output, this is only a simplifying assumption, and we can easily modify our protocol in the case where both parties must receive the same output. To do so, we let $Q_1$ output $o_1 = s_1+r_1+t_1$ (where $t_1$ is random and chosen by Party B) and have both parties learn this. We can then appropriately modify the remaining steps as well.}. The output of $Q_1$ for Party A is defined to be $r_1$. Now, every $Q_i$ ($2\leq i\leq k-1$) takes as input $x,y,r_{i-1},o_{i-1},r_i$, runs $P_i$ on inputs $x,y$ and state $o_{i-1}-r_{i-1}$ to get $s_i$ (where $-$ denotes subtraction in the ring $(\mathbb{Z},+)$) and then produces $s_i+r_i$ as output only to Party B. Again, Party A's output for $P_i$ is defined to be $r_i$. $Q_k$ takes as input $x,y,r_{k-1},o_{k-1}$, runs $P_k$ on inputs $x,y$ and state $o_{k-1}-r_{k-1}$ and outputs $z$ to both parties.

Our compiler, given $P$, will first split $P$ into $P_1||P_2||\cdots||P_k$ with the guarantee that the ABY circuits generated for every $Q_i$ defined above is small enough to fit into memory. It will then compile and execute the $k$ programs $Q_1||Q_2||\cdots||Q_k$ sequentially (freeing up memory usage after every execution) to produce final output $z$ to both parties.

### 2017-10-10 OaklandSubmission/intro.tex
intro-2

@@
 Since the introduction of SMC in 1980's, there has been a large body of work \cite{..} that has transformed SMC from a mere theoretical tool to something practically efficient/usable.
+Hence, there have been efforts to make SMC accessible to non-specialist programmers.
+To understand the state-of-the-art, let us consider the following toy example.
 
-Unfortunately, implementing functionalities using SMC protocols requires thorough understanding of cryptography.
+Suppose a developer wants to compute $w^Tx >0$ securely using SMC.
+With the current techniques, he has two options.
+He can code up the functionality in a programmer friendly domain-specific language such as \cite{lambdaps,wysteria,oblivm} that would automatically compile it to an SMC protocol.
+The downside is that all of these works build on a cryptographic back-end that is either entirely boolean \cite{yao,gmw} or entirely arithmetic \cite{homo}.
+The efficiency of the generated SMC protocol is hence, bounded by the possibility of efficiently representing the function in a form that is compatible with the cryptographic back-end used.
+For instance, over $\ell$-bit integers, multiplication requires a boolean circuit of size  $O(\ell^2)$.
+For better efficiency, we would like to evaluate $w^Tx$ using an Arithmetic circuit and the comparison using a boolean circuit.
+In fact, most interesting functions require a mix of arithmetic and boolean computations.
+To address this issue, recently Demmler et al.
+\cite{aby} gave a cryptographic protocol for SMC where the parties can mix arithmetic and boolean computations.
+\footnote{Need to note somewhere that many works combined HE and Yao.
+Look at ABY for references.}
+Hence, the second option is to write the functionality in the system developed in \cite{aby}.
+However, using their system requires the programmer to be aware of complex trade-offs between arithmetic and boolean cryptographic schemes.
+In their framework, the programmer is required to write circuits consisting of  arithmetic and boolean gates along with appropriate conversion gates.
+In short, as is also mentioned by the authors themselves, the current system is not suitable to be used by non-specialist programmers.
+\divya{Change tone...}
+Our work .....
+
+%Unfortunately, implementing functionalities using SMC protocols requires thorough understanding of cryptography.
 To allow for widespread use of SMC, it is critical that SMC protocols are programmable by non-cryptographic experts.
@@
 
-To address this issue \divya{on the cryptographic side}, recently Demmler et al.
+%To address this issue \divya{on the cryptographic side}, recently Demmler et al.
 \cite{aby} gave a cryptographic protocol for SMC where the parties can mix arithmetic and boolean computations and claimed potentially great performance benefits.
@@
 
-%Recent works \cite{aby,secureml,minion} have shown that great performance benefit can be obtained by mixing arithmetic and boolean computations in the back-end.
+%%%%%%%Recent works \cite{aby,secureml,minion} have shown that great performance benefit can be obtained by mixing arithmetic and boolean computations in the back-end.
 Certain computations such as integer multiplication are efficient when implemented using arithmetic SMC \cite{gmw} whereas max(x,0) needs to be computed using a boolean back-end \cite{yao}.
@@
 \divya{We provide a formal type system....
-} We evaluate our system on various benchmarks from \cite{} and show that automatically generated protocols match the performance of hand-crafted protocols in most cases (see \ref{sec:eval}).
+} We evaluate our system on various benchmarks such as logistic regression and convolutional neural network (CNN) for MNIST data \cite{minionn}, naive bayes, decision trees from \cite{shafindss}.
+Evaluations show that protocols generated by our compiler match the performance of hand-crafted protocols in most cases (see \ref{sec:eval}).
 \divya{Below, we will  give an overview of \tool and describe our contributions.}
@@
 \divya{ideally have small paragraphs about key features}
-1. ease to programming.. refer to section with example in our language; contrast to verilog paper; programmer is completely oblivious of security; looks like a normal C program;
-also say that user does not write keywords secret and public; these are inferred from the information flow
+1. ease to programming.. refer to section with example in our language; contrast to verilog paper; programmer is completely oblivious of security; looks like a normal/vanilla/ regular C program;
+Moreover, the developer does not write keywords such as secret and public; 
+\input{fig-ex-aby}
+\input{fig-ex-sml}
 
-2. first programmable system that supports a mix of arithmetic and boolean for efficiency give example of multiplication of integers; linear algebra; logistic regression
+2.
+first programmable system that supports a mix of arithmetic and boolean for efficiency give example of multiplication of integers; linear algebra; logistic regression and evaluate on state of the art ML prediction tasks
 
-3.
-prove correctness, type system guarantees correct termination, progress etc..; In contrast with most systems that require a programmer to label variables as public secret, we automatically infer them.
-leads to more efficiency compared to default secret
+3. prove correctness, type system guarantees correct termination, progress etc..;
 
-4. Scaling to large programs. Describe pipeling here...
+%In contrast with most systems that require a programmer to label variables as public secret, we automatically infer them. leads to more efficiency compared to default secret
 
-5. can serve as a baseline for future works on hand-crafted algorithms for specific applications cite minion etc here...
+If program type-checks then runs correctly for example no array out of bound errors
+code runs correctly in distributed env
+programmer writes one code; compiles to server code and client code
+we prove and outputs identical to stand-alone output
+distributed code produces same output as if the code were run by distributed third party
 
-6.
-Suitable to use multiple crypto backends; We formalize the guarantees needed from crypto backend; it is easy for expert programmers to to extend our framework for example add another SMC backend.
-say more about this later...
-there say that possible to integrate minion's more efficient matrix based preprocessing into our framework and matrix ops as native types and instructions in our language
+4.
+Scaling to large programs.
+Describe pipeling here...
+Like all MPC implementations use an intermediate rep as a circuit and largest compute that can run is given by largest circuit that can fit in memory are restricted by memory of machines.
+This in turn limits the size of functions
+We implement/give/design a generic technique called pipelining that allow us to run arbitrarily large computations without affecting security.
 
-7. experiment results...
+5. Previous papers when comparing with generic 2pc compare with Yao; does not scale etc...
+can serve as a baseline for future works on hand-crafted algorithms for specific applications cite minion etc here...
 
@@
 
-\divya{Useful phrases:} progress in practical efficiency; non-specialist programmers; fundamental representation gap; circuits as an abstraction of computation;
+\divya{Useful phrases:} progress in practical efficiency; non-specialist programmers; fundamental representation gap; circuits as an abstraction of computation; task becomes even more tedious, time-consuming, and error-prone,
+since each secure computation protocol has its own
+representation in which a functionality has to be described, e.g.,
+Arithmetic vs. Boolean circuits.
 
@@
 exception of TASTY
+
+\divya{where to put this..
+6.
+Suitable to use multiple crypto backends; We formalize the guarantees needed from crypto backend; it is easy for expert programmers to to extend our framework for example add another SMC backend.
+say more about this later...
+there say that possible to integrate minion's more efficient matrix based preprocessing into our framework and matrix ops as native types and instructions in our language}
+
 \newpage

### 2017-10-11 OaklandSubmission/intro.tex
1.1

@@
 
-In this work, we develop and implement our framework \tool where the programmer writes a high level C-like program (instead of a circuit) to describe the function to be computed.
-Our compiler automatically converts this program to a suitable SMC protocol that uses both arithmetic and boolean secure computation components as well as the correct conversion protocols whenever required.
+In this work, we develop and implement our framework \tool for secure computation that achieves generality, performance and programmer productivity.
+The programmer writes a high level C-like program (instead of a circuit) to describe the function to be computed.
+Our compiler automatically generates an SMC protocol using a mix of Arithmetic and Boolean compute.
 Our compiler is general and can work with any secure implementation of a mix of arithmetic and boolean computations.
-In our work, we focus on semi-honest secure two-party computation and  ABY \cite{aby} provides the suitable cryptographic back-end.
-\divya{We provide a formal type system....
-} We evaluate our system on various benchmarks such as logistic regression and convolutional neural network (CNN) for MNIST data \cite{minionn}, naive bayes, decision trees from \cite{shafindss}.
-Evaluations show that protocols generated by our compiler match the performance of hand-crafted protocols in most cases (see \ref{sec:eval}).
-\divya{Below, we will  give an overview of \tool and describe our contributions.}
+In our work, we focus on semi-honest secure two-party computation.
+We build on ABY \cite{aby} that provides the suitable cryptographic back-end in this setting.
+We provide a formal type system and prove correctness and security of our compiler.
+We evaluate our system on various benchmarks such as logistic regression and convolutional neural network (CNN) for MNIST data \cite{minionn}, naive bayes, decision trees from \cite{shafindss}.
+Evaluations show that protocols generated by our compiler match the performance of hand-crafted protocols in most cases (see \sectionref{sec:eval}).
+\divya{Below, we will  give an overview of \tool and describe our contributions in more detail.}
 
@@
 
-\subsection{\tool: Overview and Contributions} Write high level overview of programming language and tool..
+\subsection{\tool: Overview and Contributions} 
+\label{sec:contrib}
+Write high level overview of programming language and tool.. \divya{have some preamble}
 
+\tool is a programmer-friendly framework to express the function to be computed and uses a mix of arithmetic and boolean SMC back-ends.
+We illustrate the above toy example of $w^Tx >0$ in our language in \figureref{ex-sml}.
+As is clear from the code, the programmer is not required to write keywords such as {\em secret} or {\em public} and can remain completely oblivious of the cryptographic back-end.
+In contrast, \figureref{ex-aby} illustrates the same example using the framework in ABY \cite{aby}.
+Here, as already mentioned, the programmer needs to write the low level gates like $\mathrm{MULT, ADD, CONS, GT}$, etc.
+The programmer is also required to use the appropriate share conversion gates such as A2Y that converts Arithmetic shares to Yao shares.
+
+Our compiler is aware of cryptographic costs of different representations and inter-conversions.
+Based on these costs, our compiler, given a program in \tool, automatically picks Arithmetic and Boolean representations for different sub-parts of the program and inserts the required inter-conversions as well.
+Our experiments show that automatically generated protocols using \tool match the performance of tailor-made protocols in \cite{minionn,shafindss} (see \sectionref{eval}).
+This shows that  \tool achieves generality, performance and programmer productivity.
+These features allow us to evaluate our system on a variety of benchmarks, especially, state-of-the-art ML prediction algorithms with \emph{real} data and models.
+
+We give a formal {\em type system} for our language and prove useful properties.
+We prove that if a program type checks in our language then it would run correctly, e.g., there would not be any array index out of bounds error.
+The programmer writes a single code describing the inputs from each party and the function to be computed.
+Our compiler automatically generates separate executables for the server and the client that can be executed in the distributed setting.
+We prove the {\em correctness} for this compilation, i.e., the outputs obtained in the distributed execution are identical to the stand-alone execution by a trusted third party.
+We formally reduce the {\em security} of our system to the security of the cryptographic back-end.
+For details on correctness and security theorems see \sectionref{ld-proofs}.
+
+Another issue in adoption of SMC is difficulty in scaling to large programs. 
+%In general, SMC implementations struggle in scaling to large programssize of memory on a system limits the largest compute inside the SMC and hence,. 
+The reason being all SMC implementations use a circuit-like representation as an intermediate representation.
+Hence, the largest compute that can be done securely is upper-bounded by the amount of memory on a machine.
+
+\input{fig-ex-aby}
+\input{fig-ex-sml}
+
+\newpage
 \divya{ideally have small paragraphs about key features}
@@
 Moreover, the developer does not write keywords such as secret and public;
-\input{fig-ex-aby}
-\input{fig-ex-sml}
 

### 2017-10-11 OaklandSubmission/intro.tex
1.1

@@
 \label{sec:contrib}
-Write high level overview of programming language and tool.. \divya{have some preamble}
+%Write high level overview of programming language and tool.. \divya{have some preamble}
 
@@
 We formally reduce the {\em security} of our system to the security of the cryptographic back-end.
-For details on correctness and security theorems see \sectionref{ld-proofs}.
+For details on formal language, correctness and security theorems see \sectionref{ld}.
 
@@
 The reason being all SMC implementations use a circuit-like representation as an intermediate representation.
-Hence, the largest compute that can be done securely is upper-bounded by the amount of memory on a machine.
+Hence, the largest compute that can be done securely is upper-bounded by the largest circuit that can fit in memory on a machine.
+To address this issue, we develop a technique called \emph{pipelining} that allows us to run aribitrarily large programs.
+At a high level, we break the original program into a sequence of small sub-programs and mask the intermediate outputs using secure secret sharing between the two parties.
+We appropriately append the share-reconstruction function to the beginning of the subsequent sub-program.
+Now, we execute each of these sub-programs sequentially in our framework.
+We provide details of pipelining and prove the correctness and security of this transformation in \sectionref{pipe}.
+
+Finally, our framework \tool can serve as a meaningful baseline for future works on hand-crafted SMC protocols for specific functionalities such as \textsc{MiniONN} \cite{minionn}. 
+Prior to our work, garbled circuit based two-party computation is cited as the most practical approach and used as the baseline comparison of performance.
+But, it has been observed multiple times this approach does not scale on realistic examples.
 
@@
 
-\newpage
-\divya{ideally have small paragraphs about key features}
-1. ease to programming.. refer to section with example in our language; contrast to verilog paper; programmer is completely oblivious of security; looks like a normal/vanilla/ regular C program;
-Moreover, the developer does not write keywords such as secret and public;
-
-2.
+%\newpage
+%\divya{ideally have small paragraphs about key features}
+%1. ease to programming.. refer to section with example in our language; contrast to verilog paper; programmer is completely oblivious of security; looks like a normal/vanilla/ regular C program;
+%Moreover, the developer does not write keywords such as secret and public; 
+%
+%
+%2.
 first programmable system that supports a mix of arithmetic and boolean for efficiency give example of multiplication of integers; linear algebra; logistic regression and evaluate on state of the art ML prediction tasks
-
-3. prove correctness, type system guarantees correct termination, progress etc..;
-
-%In contrast with most systems that require a programmer to label variables as public secret, we automatically infer them. leads to more efficiency compared to default secret
-
-If program type-checks then runs correctly for example no array out of bound errors
-code runs correctly in distributed env
-programmer writes one code; compiles to server code and client code
-we prove and outputs identical to stand-alone output
-distributed code produces same output as if the code were run by distributed third party
-
-4.
+%
+%3. prove correctness, type system guarantees correct termination, progress etc..; 
+%
+%%In contrast with most systems that require a programmer to label variables as public secret, we automatically infer them. leads to more efficiency compared to default secret
+%
+%If program type-checks then runs correctly for example no array out of bound errors
+%code runs correctly in distributed env
+%programmer writes one code; compiles to server code and client code
+%we prove and outputs identical to stand-alone output
+%distributed code produces same output as if the code were run by distributed third party
+%
+%
+%4.
 Scaling to large programs.
@@
 This in turn limits the size of functions
-We implement/give/design a generic technique called pipelining that allow us to run arbitrarily large computations without affecting security.
+%We implement/give/design a generic technique called pipelining that allow us to run arbitrarily large computations without affecting security.
+%
+%
+%5. Previous papers when comparing with generic 2pc compare with Yao; does not scale etc...
+%can serve as a baseline for future works on hand-crafted algorithms for specific applications cite minion etc here...
+%
 
-5. Previous papers when comparing with generic 2pc compare with Yao; does not scale etc...
-can serve as a baseline for future works on hand-crafted algorithms for specific applications cite minion etc here...
+\vspace{10pt}
 
-\noindent\textbf{Related Work}
+\noindent\textbf{Related Work} \divya{I think we should say a few words here about the following: minionn, verilog paper, 
 \cite{lambdaps,wysteria}
 
-\divya{Useful phrases:} progress in practical efficiency; non-specialist programmers; fundamental representation gap; circuits as an abstraction of computation; task becomes even more tedious, time-consuming, and error-prone,
-since each secure computation protocol has its own
-representation in which a functionality has to be described, e.g.,
-Arithmetic vs. Boolean circuits.
-
-\divya{useful thoughts} It is widely believed that garbled circuit based 2pc is most practical; but not quite true..
-
-\divya{points that have to be there...} Security; traces are ``oblivious'' of secret values if observables/outputs are same; formal cryptographic security guarantee uses simulation paradigm \cite{canetti2000}
-
-\divya{not sure if we wanna say anything about these} 1.
-all the programs written in our language are memory trace oblivious because array indices are public 2.
-instruction trace obliviousness handled by using a multiplexor or executing both branches..
-plus all loops are public counters.
-Our type system rejects programs that loop on secret variables
-
-\divya{questions:} Does ABY allow secret array indices? or does it allow arrays? thats not so important actually
-
-\divya{Related Work: writing some citations here to be used appropriately.}
-\begin{enumerate}
-\item Secure Computation of MIPS Machine Code %https://eprint.iacr.org/2015/547
-\item TASTY https://eprint.iacr.org/2010/365.pdf
-
-\end{enumerate}
-
-There have been many languages and compilers that compile to cryptographic backends
-exception of TASTY
-
-\divya{where to put this..
-6.
-Suitable to use multiple crypto backends; We formalize the guarantees needed from crypto backend; it is easy for expert programmers to to extend our framework for example add another SMC backend.
-say more about this later...
-there say that possible to integrate minion's more efficient matrix based preprocessing into our framework and matrix ops as native types and instructions in our language}
+}
 

### 2017-10-11 OaklandSubmission/related.tex
1.1

\section{Related work}
\label{sec:related}

\divya{Related Work: writing some citations here to be used appropriately.}
\begin{enumerate}
\item Secure Computation of MIPS Machine Code https://eprint.iacr.org/2015/547
\item TASTY https://eprint.iacr.org/2010/365.pdf
\item ABY cites many papers on first page of intro for using a mix of HE and boolean MPC; but they didnt give much gain because HE is expensive

\end{enumerate}

### 2017-10-11 OaklandSubmission/intro.tex
1.1

@@
 Finally, our framework \tool can serve as a meaningful baseline for future works on hand-crafted SMC protocols for specific functionalities such as \textsc{MiniONN} \cite{minionn}. 
-Prior to our work, garbled circuit based two-party computation is cited as the most practical approach and used as the baseline comparison of performance.
-But, it has been observed multiple times this approach does not scale on realistic examples.
+Prior to our work, garbled circuit based two-party computation is cited as the most practical approach and used as the baseline for comparison of performance.
+But, it has been observed multiple times this approach does not even scale on realistic examples.
 

### 2017-10-11 OaklandSubmission/intro.tex
intro-preamble

@@
 \label{sec:intro}
+Today it is hard for developers to create secure implementations using cryptographic techniques.
+Typical developers lack an understanding of cryptography protocols
+and cannot be expected to implement them correctly without tool support.
+In an ideal world, a developer declares the functionality to be implemented
+in a general purpose programming language and an automatic tool takes care
+of all the cryptography and generates a secure and efficient implementation  of the functionality. 
+This paper provides a language-based solution to make
+secure multiparty computation (SMC) accessible to the software developer community.
 
-Most cryptographic tools rely on circuit as abstractions for any general computation but real-life developers write programs.
+%Most cryptographic tools rely on circuit as abstractions for any general computation but real-life developers write programs.
 It is critical to address this fundamental  representation gap to make cryptography accessible to non-crypto specialist developers.
 In this work, we address this gap for the case of secure multi-party computation.
+
+%Most cryptographic tools are very hard to use non-crypto specialist developers.
+The task becomes even more tedious and error-prone for the case of secure multiparty computation since different protocol implementations require the function to be expressed in different representations such as Boolean or Arithmetic circuits.
+To make cryptography accessible to non-crypto specialist developers, we need to design systems that do crypto automatically and satisfy the properties of ...
 

### 2017-10-12 src/smlCodeGen.py
added public int

@@ -169,6 +169,8 @@ class smlCodeGen:
     def convertExpr(self, expr, circ):
         if isinstance(expr, Ident):
             tmpvar = expr.name
+            if tmpvar in self.publicSet:
+                return
             i = self.getIndex(circ)
             varShares = self.dict[tmpvar]
             if varShares[i] != None and varShares[i] != -1:
@@ -244,6 +246,8 @@ class smlCodeGen:
                 varShares[i] = varname
         else:
             name = lhs.name
+            if name in self.publicSet:
+                return
             varShares = self.dict[name]
             i = self.getIndex(circ)
             if varShares[i] is None:
@@ -257,7 +261,9 @@ class smlCodeGen:
                 self.convertExpr(x.rhs, circ)
                 self.declareLhs(x.lhs, circ)
             elif isinstance(x, ForLoop):
+                self.publicSet.add(x.idname.name)
                 self.convertAll(x.block, circ, lhsSet)
+                self.publicSet.remove(x.idname.name)
             elif isinstance(x, IfElse):
                 self.convertAll(x.expr1, circ, lhsSet)
                 if x.arity == 3:
@@ -294,9 +300,12 @@ class smlCodeGen:
                 circ = self.circ
             if isinstance(node.idname, Ident):
                 name = node.idname.name
+                print(' '*offset + node.data_type.idtype, name + ';', file=self.file)
+                if node.isPublic:
+                    self.publicSet.add(name)
+                    return
                 sname = shareName(name, circ)
                 self.putInDict(name, sname, circ)
-                print(' '*offset + node.data_type.idtype, name + ';', file=self.file)
                 print(' '*offset + 'share *' + sname + ';', file=self.file)
             else:
                 name = node.idname.idname.name + node.idname.ref
@@ -308,6 +317,14 @@ class smlCodeGen:
                 self.makeVec('s_'+circ[0]+'_'+arrname, 'share*', refctx)
 
         elif isinstance(node, Ident): # should always get a circ
+            if node.name in self.publicSet:
+                tmpvar = '_tmp_'+str(self.counter)
+                varname = shareName(tmpvar, circ)
+                self.counter += 1
+                print(' '*offset+'uint32_t', tmpvar, '=', node.name,';', file=self.file)
+                print(' '*offset+'share *',varname,'=',circ+'->PutCONSGate(',tmpvar,',bitlen);', file=self.file)
+                self.putInDict(tmpvar, varname, circ)
+                return varname
             shareList = self.dict[node.name]
[…]

### 2017-10-12 src/smlLexer.py

@@ -7,115 +7,120 @@ import sys
 
 def serializedATN():
     with StringIO() as buf:
-        buf.write("\3\u608b\ua72a\u8133\ub9ed\u417c\u3be7\u7786\u5964\2.")
-        buf.write("\u0113\b\1\4\2\t\2\4\3\t\3\4\4\t\4\4\5\t\5\4\6\t\6\4\7")
+        buf.write("\3\u608b\ua72a\u8133\ub9ed\u417c\u3be7\u7786\u5964\2/")
+        buf.write("\u011c\b\1\4\2\t\2\4\3\t\3\4\4\t\4\4\5\t\5\4\6\t\6\4\7")
         buf.write("\t\7\4\b\t\b\4\t\t\t\4\n\t\n\4\13\t\13\4\f\t\f\4\r\t\r")
         buf.write("\4\16\t\16\4\17\t\17\4\20\t\20\4\21\t\21\4\22\t\22\4\23")
         buf.write("\t\23\4\24\t\24\4\25\t\25\4\26\t\26\4\27\t\27\4\30\t\30")
         buf.write("\4\31\t\31\4\32\t\32\4\33\t\33\4\34\t\34\4\35\t\35\4\36")
         buf.write("\t\36\4\37\t\37\4 \t \4!\t!\4\"\t\"\4#\t#\4$\t$\4%\t%")
         buf.write("\4&\t&\4\'\t\'\4(\t(\4)\t)\4*\t*\4+\t+\4,\t,\4-\t-\4.")
-        buf.write("\t.\4/\t/\4\60\t\60\4\61\t\61\4\62\t\62\3\2\3\2\3\3\3")
-        buf.write("\3\3\4\3\4\3\5\3\5\3\5\3\5\3\5\3\5\3\5\3\5\3\5\3\6\3\6")
-        buf.write("\3\6\3\6\3\6\3\6\3\6\3\6\3\6\3\7\3\7\3\b\3\b\3\t\3\t\3")
-        buf.write("\n\3\n\3\n\3\n\3\13\3\13\3\f\3\f\3\f\3\r\3\r\3\16\3\16")
-        buf.write("\3\17\3\17\3\17\3\17\3\17\3\20\3\20\3\21\3\21\3\22\3\22")
[…]

### 2017-10-12 src/smlParser.py

@@ -7,116 +7,118 @@ import sys
 
 def serializedATN():
     with StringIO() as buf:
-        buf.write("\3\u608b\ua72a\u8133\ub9ed\u417c\u3be7\u7786\u5964\3.")
-        buf.write("\u0102\4\2\t\2\4\3\t\3\4\4\t\4\4\5\t\5\4\6\t\6\4\7\t\7")
+        buf.write("\3\u608b\ua72a\u8133\ub9ed\u417c\u3be7\u7786\u5964\3/")
+        buf.write("\u0105\4\2\t\2\4\3\t\3\4\4\t\4\4\5\t\5\4\6\t\6\4\7\t\7")
         buf.write("\4\b\t\b\4\t\t\t\4\n\t\n\4\13\t\13\4\f\t\f\4\r\t\r\4\16")
         buf.write("\t\16\4\17\t\17\4\20\t\20\4\21\t\21\4\22\t\22\4\23\t\23")
         buf.write("\4\24\t\24\4\25\t\25\3\2\7\2,\n\2\f\2\16\2/\13\2\3\3\3")
         buf.write("\3\3\3\3\3\3\3\3\3\5\3\67\n\3\3\4\3\4\7\4;\n\4\f\4\16")
         buf.write("\4>\13\4\3\4\3\4\3\5\3\5\3\5\3\5\5\5F\n\5\3\6\3\6\3\6")
-        buf.write("\3\6\3\6\3\6\3\6\5\6O\n\6\3\6\5\6R\n\6\3\7\3\7\3\b\3\b")
-        buf.write("\3\b\3\b\3\b\3\b\3\b\3\b\3\b\3\b\3\b\3\b\3\b\3\b\3\b\3")
-        buf.write("\b\3\b\3\b\3\b\3\b\5\bj\n\b\3\t\3\t\3\t\3\n\3\n\3\n\6")
-        buf.write("\nr\n\n\r\n\16\ns\3\13\3\13\3\13\3\f\3\f\3\f\3\f\6\f}")
-        buf.write("\n\f\r\f\16\f~\3\r\3\r\3\r\3\r\3\r\3\r\3\r\3\16\3\16\3")
-        buf.write("\16\3\16\3\16\3\16\5\16\u008e\n\16\3\16\3\16\3\17\3\17")
-        buf.write("\3\17\3\17\3\17\3\17\3\17\5\17\u0099\n\17\3\20\3\20\5")
[…]

### 2017-10-12 examples/cifar/1_pipe_arith_cifar.cpp
cifar files

@@ -141,11 +141,11 @@ share * s_a_tmp_4 = acirc->PutOUTGate( s_a__tmp_5 , ALL);
 auto s_a_c2in_o = make_vector<share*>(576, 1024);
 for(int i = 0; i < 576; i++) {
 	for(int j = 0; j < 1024; j++) {
-		s_a_c2in_o[i][j] = acirc->PutOUTGate(s_a_c2in[i][j], SERVER);
+		s_a_c2in_o[i][j] = acirc->PutOUTGate(s_a_c2in[i][j], CLIENT);
 	}
 }
 party->ExecCircuit();
-if(role == SERVER) {
+if(role == CLIENT) {
 	ofstream fout("c2in", std::ofstream::out | std::ofstream::app);
 	for(int i = 0; i < 576; i++) {
 		for(int j = 0; j < 1024; j++) {

### 2017-10-12 examples/cifar/4_pipe_arith_cifar.cpp

@@ -26,13 +26,13 @@ while(!fin.eof()){
 }
 fin.close();
 }
-if(role == SERVER){
+if(role == CLIENT){
 ifstream fin("c2in");
 while(!fin.eof()){
 	uint32_t i, j, val;
 	fin >> i >> j >> val;
 	c2in[i][j] = val;
-	s_a_c2in[i][j]= acirc->PutINGate(c2in[i][j] ,bitlen,SERVER);
+	s_a_c2in[i][j]= acirc->PutINGate(c2in[i][j] ,bitlen,CLIENT);
 	
 }
 fin.close();
@@ -110,11 +110,11 @@ share * s_a_tmp_2 = acirc->PutOUTGate( s_a__tmp_3 , ALL);
 auto s_a_c2out_o = make_vector<share*>(64, 1024);
 for(int i = 32; i < 48; i++) {
 	for(int j = 0; j < 1024; j++) {
-		s_a_c2out_o[i][j] = acirc->PutOUTGate(s_a_c2out[i][j], SERVER);
+		s_a_c2out_o[i][j] = acirc->PutOUTGate(s_a_c2out[i][j], CLIENT);
 	}
 }
 party->ExecCircuit();
-if(role == SERVER) {
+if(role == CLIENT) {
 	ofstream fout("c2out", std::ofstream::out | std::ofstream::app);
 	for(int i = 32; i < 48; i++) {
 		for(int j = 0; j < 1024; j++) {

### 2017-10-12 examples/cifar/5_pipe_arith_cifar.cpp

@@ -26,13 +26,13 @@ while(!fin.eof()){
 }
 fin.close();
 }
-if(role == SERVER){
+if(role == CLIENT){
 ifstream fin("c2in");
 while(!fin.eof()){
 	uint32_t i, j, val;
 	fin >> i >> j >> val;
 	c2in[i][j] = val;
-	s_a_c2in[i][j]= acirc->PutINGate(c2in[i][j] ,bitlen,SERVER);
+	s_a_c2in[i][j]= acirc->PutINGate(c2in[i][j] ,bitlen,CLIENT);
 	
 }
 fin.close();
@@ -109,11 +109,11 @@ share * s_a_tmp_1 = acirc->PutOUTGate( s_a__tmp_2 , ALL);
 auto s_a_c2out_o = make_vector<share*>(64, 1024);
 for(int i = 48; i < 64; i++) {
 	for(int j = 0; j < 1024; j++) {
-		s_a_c2out_o[i][j] = acirc->PutOUTGate(s_a_c2out[i][j], SERVER);
+		s_a_c2out_o[i][j] = acirc->PutOUTGate(s_a_c2out[i][j], CLIENT);
 	}
 }
 party->ExecCircuit();
-if(role == SERVER) {
+if(role == CLIENT) {
 	ofstream fout("c2out", std::ofstream::out | std::ofstream::app);
 	for(int i = 48; i < 64; i++) {
 		for(int j = 0; j < 1024; j++) {

### 2017-10-12 examples/cifar/6_pipe_arith_cifar.cpp

@@ -40,13 +40,13 @@ while(!fin.eof()){
 }
 fin.close();
 }
-if(role == SERVER){
+if(role == CLIENT){
 ifstream fin("c2out");
 while(!fin.eof()){
 	uint32_t i, j, val;
 	fin >> i >> j >> val;
 	c2out[i][j] = val;
-	s_a_c2out[i][j]= acirc->PutINGate(c2out[i][j] ,bitlen,SERVER);
+	s_a_c2out[i][j]= acirc->PutINGate(c2out[i][j] ,bitlen,CLIENT);
 	
 }
 fin.close();
@@ -263,11 +263,11 @@ share * s_a_tmp_7 = acirc->PutOUTGate( s_a__tmp_8 , ALL);
 auto s_a_c4in_o = make_vector<share*>(576, 256);
 for(int i = 0; i < 576; i++) {
 	for(int j = 0; j < 256; j++) {
-		s_a_c4in_o[i][j] = acirc->PutOUTGate(s_a_c4in[i][j], SERVER);
+		s_a_c4in_o[i][j] = acirc->PutOUTGate(s_a_c4in[i][j], CLIENT);
 	}
 }
 party->ExecCircuit();
-if(role == SERVER) {
+if(role == CLIENT) {
 	ofstream fout("c4in", std::ofstream::out | std::ofstream::app);
 	for(int i = 0; i < 576; i++) {
 		for(int j = 0; j < 256; j++) {

### 2017-10-12 OaklandSubmission/app-stash.tex

\section{Random}

Divya is putting some stash here to be used later...

\divya{Useful phrases:} progress in practical efficiency; non-specialist programmers; fundamental representation gap; circuits as an abstraction of computation; task becomes even more tedious, time-consuming, and error-prone,
since each secure computation protocol has its own
representation in which a functionality has to be described, e.g.,
Arithmetic vs. Boolean circuits.

\divya{useful thoughts} It is widely believed that garbled circuit based 2pc is most practical; but not quite true..

\divya{points that have to be there...} Security; traces are ``oblivious'' of secret values if observables/outputs are same; formal cryptographic security guarantee uses simulation paradigm \cite{canetti2000}

\divya{not sure if we wanna say anything about these} 1. all the programs written in our language are memory trace oblivious because array indices are public 2. instruction trace obliviousness handled by using a multiplexor or executing both branches.. plus all loops are public counters. Our type system rejects programs that loop on secret variables

\divya{questions:} Does ABY allow secret array indices? or does it allow arrays? thats not so important actually

\divya{where to put this.. 6. Suitable to use multiple crypto backends; We formalize the guarantees needed from crypto backend; it is easy for expert programmers to to extend our framework for example add another SMC backend. say more about this later... there say that possible to integrate minion's more efficient matrix based preprocessing into our framework and matrix ops as native types and instructions in our language}

### 2017-10-12 OaklandSubmission/eval.tex
1.1
accuracy
eval 6.1

@@
 \label{sec:eval}
-We evaluate \tool for the problem of {\it secure prediction}, where a server has a machine learning model, a client has an input, and after the protocol execution the client learns a label on the input.
-From the protocol execution, the server learns nothing about the client input and the client learns nothing about the server model than what is revealed beyond the computed label.
+We evaluate \tool for the problem of {\it secure prediction}, where a server has a machine learning model, a client has an input, and after the protocol execution the client learns a label on his input.
+From the protocol execution, the server learns nothing about the client's input and the client learns nothing about the server's model beyond what is revealed from the computed label.
 This setting has been previously studied in~\cite{shafindss,minionn}
-We first implement benchmarks of ~\cite{shafindss} and~\cite{minionn} in \tool to compare gainst 
+We first implement benchmarks of ~\cite{shafindss} and~\cite{minionn} in \tool to compare against 
 previous work.
-Next, we show the generality of \tool by implementing state-of-the-art machine learning models.
+Next, we demonstrate the generality of \tool by implementing state-of-the-art machine learning models.
 Finally, we discuss a model that requires pipelinig.
@@
 28 GBs of RAM. The \tool compiler is written in Python and compiles each of our benchmarks
-in under a second to C++ code that makes calls to the ABY library~\cite{aby}. Since most of our benchmarks are related to machine learning, we set up some notation standard to machine learning next.
+in under a second to C++ code that makes calls to the ABY library~\cite{aby}.
+Since most of our benchmarks are related to machine learning, we set up some notation standard to machine learning and describe our benchmarks next.
 
 \subsection{Benchmarks}
+\divya{notation can be moved somewhere... } \\
+
+\noindent\textbf{Notation.} We use $[N]$ to denote $\{0,1,\dotsc, N-1\}$. \rs{Define $\mathsf{argmax}$.}
+
 In this paper, we focus on machine learning models for {\it classification}.
@@
 For example, given a data point which is a tuple of height and weight of an individual,
-a classifier can predict a label ``male'' or ``female''.
+a classifier can predict a label ``male'' or ``female''. The {\it accuracy} of a classifier
+refers to the fraction of data points that the classifier labels correctly from a given set of {\it test} data points.
 
-A binary linear classifier is one of the simplest classifers. Here, the input data point $x\in\R^d$
-and the model is a vector $w\in\R^d$. The possible labels $\ell\in\{\mathit{true},\mathit{false}\}$
-and the classifier $C_w\equiv w^Tx>0$, where $\cdot^T$ denotes the matrix transpose operator.
+A binary linear classifier is one of the simplest classifiers. Here, the input data point $x\in\R^d$
+and the model is a vector $w\in\R^d$. The possible labels are $\ell\in\{\mathit{true},\mathit{false}\}$
+and the classifier is $C_w\equiv w^Tx>0$, where $\cdot^T$ denotes the matrix transpose operator.
 This classifier requires $d$ multiplications, $d-1$ additions, and a comparison.
 
-A more interesting classifier is Naive Bayes that predicts labels from the set $[n]$.
+A more interesting classifier is Na\"{i}ve Bayes that predicts labels from the set $[n]$.
 Here, the input data point is a {\it feature}
-vector $x=(x_1,x_2,\ldots,x_d)^T$ where each $x_i\in [F]$.
-The model has two matrices: a vector $P$ with $|P|=n$ and $P(i) \equiv \log p(\ell=i)$, where $p(\ell=i)$ is the likelihood that the output label is $i$ for $i\in[n]$.
-The other matrix $T$ has size $n\times d\times F$ and $T(i)(j)(x_j)\equiv\log p(x_j|\ell=i)$, i.e., the log-probability that 
-the $j^{th}$ input feature is $x_j$ (where $j\in [d]$) given that the output label is $i$.
+vector $x=(x_0,x_2,\ldots,x_{d-1})^T$ where each $x_j\in [F]$.
+The model has two matrices: a vector $P$ of length $n$ and for each $i \in [n]$, $P(i)$ is the log-probability that the output label is $i$.
+%$P(i) \equiv \log p(\ell=i)$, where $p(\ell=i)$ is the likelihood that the output label is $i$ for $i\in[n]$.
+The other matrix $T$ has size $n\times d\times F$ and for each $i\in [n], j\in [d], k\in [F]$, $T(i)(j)(k)$ is
+%\equiv\log p(x_j|\ell=i)$, i.e., 
+the log-probability that  the $j^{th}$ input feature is $k$  conditioned on  the output label being $i$.
 The classifier $C_{P,T}(x)$ outputs
 \[
-{\it argmax}_i P(i)+\sum_{j=1}^d T(i)(j)(x_j)
+{\sf argmax}_i P(i)+\sum_{j=0}^{d-1} T(i)(j)(x_j)
 \]
 This classifier requires $ndF$ comparisons and additions.
+\divya{Note that computing $T(i)(j)(x_j)$ requires a secret look-up in matrix $T$  based on secret input $x_j$ resulting in $F$ comparisons (see \sectionref{}).} \cmmt{Is there a +1 etc missing or is this exact?}
 
-A decision tree of size $N$ takes as input $x\in\R^n$ and the model consists of a tree with a boolean predicate
+A decision tree of size $N$ takes as input $x\in\R^d$ and the model consists of a binary tree of depth $d$ with a boolean predicate
 assigned to each internal node.
-In particular, a node $j\in[N]$ at depth $i$ has a predicate $b_{j}^i\equiv x_i\leq w_i$.
-We start evaluating the tree from the root and if the predicate at the current node $b_j^i$ is false (true) then $x$ is passed to the left (right) node $b_{2j+1}^{i+1}$ ($b_{2j+2}^{i+1}$).
+The root note is the $0^{th}$ node and for an internal node $j \in [N]$, the children nodes are $2j+1$ and $2j+2$.
+Each internal node $j\in[N]$ at depth $i$ has a predicate $b_{j}^i\equiv x_i\leq w_i$.
+We start evaluating the tree from the root and if the predicate at the current node $b_j^i$ is false (resp., true) then $x$ is passed to the left (resp., right) child with predicate $b_{2j+1}^{i+1}$ (resp., $b_{2j+2}^{i+1}$).
 This process is repeated till we reach a leaf.
 The leaves of the tree are labels and the output label is the leaf visited by this traversal.
-Such a (binary) decision tree can be encoded as a polynomial and then prediction
+Such a (binary) decision tree can be encoded as a polynomial (of degree linear in depth of the tree) and then the prediction task
 reduces to polynomial evaluation. E.g., a binary tree with depth one and size $N=3$ can be encoded as the polynomial
-$(1-b(x))\ell_1 + b(x)\ell_2$ where $b(x)$ is the predicate at the root, $\ell_1$ is the label of the left leaf and $\ell_2$ is the label of the right leaf.
+$(1-b(x))\ell_0 + b(x)\ell_1$ where $b(x)$ is the predicate at the root, $\ell_0$ is the label of the left leaf and $\ell_1$ is the label of the right leaf.
 
 The next class of classifiers that we discuss are deep neural nets or DNNs. 
-A DNN has multiple layers that each compute a matrix multiplication followed by
-an {\it activation} function. The most common activation functions are 
-squaring $f(x)=x^2$ and ReLU $f(x)=\mathit{max}(x,0)$. 
-If $A$ is a matrix then the output of $f(A)$ is also a matrix that contains $f$  applied
- to each entry of $A$ pointwise.
+A DNN has multiple layers such that each layer computes a matrix multiplication followed by an {\it activation} function $f$.
+The most common activation functions are square $f(x)=x^2$ and rectifier linear unit (ReLU) $f(x)=\mathsf{max}(x,0)$.
+If $A$ is a matrix then the output of $f(A)$ is also a matrix that is obtained by applying $f$ to each entry of $A$ pointwise.
 Given an input vector $x$, the predicted label of a DNN is
 \[
- \mathit{argmax} W_n\cdot f_{N-1}(\ldots f_1(W_1\cdot x)\ldots)
+ \mathsf{argmax}\ W_N\cdot f_{N-1}(\ldots f_1(W_1\cdot x)\ldots)
 \]
-Here, $W_i$'s are matrices that are part of the model and $f_i$'s are the activation functions.
+Here, $f_i$'s are the \divya{public} activation functions and the model consists of matrices $W_i$. The input vector is $x \in \R^d$.
 The operator $\cdot$ denotes a matrix multiplication.
@@
 These layers multiply a matrix with a vector.
-Some neural nets have convolution layers and such DNNs are also called CNNs.
+Some neural nets have convolution layers and such DNNs are also called convolutional neural nets or CNNs.
 For the purpose of this paper, a convolution can be considered as a (heavy) matrix-matrix multiplication.
-The size of matrices manipulated by a convolution layer grows linearly with {\it window size} (typically 9 or 25), the number of {\it output channels} (typically 16, 32, or 64),
-and the size of the matrix input to this layer. 
-Therefore, fully connected layers are easier computation-wise compared
+The size of matrices manipulated by a convolution layer grows linearly with {\it window size} (typically 9 or 25), the number of {\it output channels} (typically 16, 32, or 64), and the size of the matrix input to this layer.
+Therefore, fully connected layers are lighter computation-wise compared
 to convolution layers. In general, DNNs are computationaly heavy but provide
@@
 Finally, there are a class of machine learning classifiers that are much more efficient than
-DNNs and provide reasonably good accuracies on standard learning tasks. One such classifer
-is \bonsai~\cite{bonsai}, which is a variant of the decision trees described above.
-In particular, each node $j$ evaluates a predicate $\theta_j^TZx > 0$ to decide whether
-to pass $x$ to the left branch or the right branch ($\theta_j$ is a vector and $Z$ is a matrix).
+DNNs and provide reasonably good accuracies on standard learning tasks. One such classifier
+is \bonsai~\cite{bonsai} that takes as input $x \in \R^d$ and model consists of a binary tree with $N$ nodes.
+Each node $j$ contains matrices $W_j$ and $V_j$.
+The internal node $j$  evaluates a predicate $\theta_j^TZx > 0$ to decide whether
+to pass $x$ to the left child $2j+1$ or the right child $2j+2$ (where $\theta_j$ is a vector and $Z$ is a matrix).
 The predicted value is
 \[
-{\it argmax} \sum_j I_j(x)W_j^TZx\circ f(V_j^TZx) 
+{\sf argmax} \sum_{j\in [N]} I_j(x)W_j^TZx\circ f(V_j^TZx) 
 \]
-Here, $I_j$ is 1 if the $j^{th}$ node is present on the path traversed by $x$
+Here, $I_j(x)$ is 1 if the $j^{th}$ node is present on the path traversed by $x$
 and is zero otherwise. 
-The operation $\circ$ is a pointwise dot product, $W_k$'s and $V_k$'s are vectors,  and
-the function $f$ is given by $f(x) = x$ if $|x| < 1$ and $\mathrm{signum}(x)$ otherwise.
+The operation $\circ$ is a point-wise dot product, $W_j$'s and $V_j$'s are matrices.
+The activation function $f$ is given by $f(y) = y$ if $-1 < y < 1$ and $\mathrm{sign}(x)$ otherwise.
+The model consists of the matrices $W_j$, $V_j$, $Z$ and vectors $\theta_j$.
+\bonsai can be seen as a variant of decision trees where the prediction is a function of the path traversed from root to the leaf and not just the leaf itself.
 
@@
 The execution time is dominated by multiplications and boolean-ands.
-The number of boolean-and gates (\#And) remains constant for both the benchmnarks as they reflect the number of and gates required for a single 32-bit integer comparison with zero.
+The number of boolean-and gates (\#And) remains constant for both the benchmarks as they reflect the number of and gates required for a single 32-bit integer comparison with zero.
 The number of multiplication gates (\#Mul) is the same as the model size $d$.
@@
 
-The results for Naive Bayes are in Table~\ref{tab:nb}. As before, $n$ denotes the length of the input feature vector $x=(x_1,\ldots,x_n)$ and $F$ is the number of possible values of each $x_i$.
+The results for Na\"{i}ve Bayes are in Table~\ref{tab:nb}. As before, $n$ denotes the length of the input feature vector $x=(x_1,\ldots,x_n)$ and $F$ is the number of possible values of each $x_i$.
 Although, there are no multiplications in these benchmarks, these do have a significant number of comparisons that result in many boolean-and gates (13k/750k denote 13000/75000). 
@@
 This claim is further validated by Table~\ref{tab:dt}.
-In~\cite{wu}, Wu et al. describe specialized protocols for trees and forests (that perfrom better than~\cite{shafindss}) and we report their performance in columns 3 and 4 of Table~\ref{tab:dt}.
+In~\cite{wu}, Wu et al. describe specialized protocols for trees and forests (that perform better than~\cite{shafindss}) and we report their performance in columns 3 and 4 of Table~\ref{tab:dt}.
 Similar to~\cite{shafindss},~\cite{wu} remarks that these computations are infeasible with Yao-based 2PC.
@@
 \[
- \mathit{argmax}_{i} w_i^Tx+b_i
+ \mathsf{argmax}_{i} w_i^Tx+b_i
 \]

### 2017-10-12 OaklandSubmission/ex.tex
init
example

--- 2017-10-02 OaklandSubmission/main.tex
@@
-\documentclass[conference,compsoc]{IEEEtran}
+\section{Motivating example}
+\label{sec:ex}
+%$w^Tx>0$, mnist logistic regression
+Consider a hypothetical cloud provider Bob that wants to provide a service to diagnose
+whether a patient has breast cancer or not. Bob trains a machine learning classification model
+that results in a vector $w$ (say of length 30) and a scalar $b$.
+This model is the intellectual property of Bob and he  wants to keep it secret.
+Given a patient's medical report, in the form of a vector $x$ (also of length 30),
+the classifer predicts that the patient has breast cancer iff $w^Tx>b$.
+However, for this task Bob needs access to $x$, which is private data of a customer.
 
-\usepackage{epsfig,endnotes}
-\usepackage{varwidth}
-\usepackage{algorithm}
-\usepackage{algpseudocode}
-\usepackage{balance}
-\usepackage{color}
-\usepackage{nicefrac}
-\usepackage{amsmath}
-\usepackage{braket}
-\usepackage{bm}
-\usepackage{mathtools}
-\usepackage{multirow}
-\usepackage{bigdelim}
-\usepackage{mathtools}
-\usepackage{amssymb}
-\usepackage{indentfirst}
-\usepackage{booktabs}
-\usepackage{enumitem}
-\usepackage{boxedminipage}
-\usepackage{float}
-\usepackage{graphicx}
-\usepackage{caption}
-\usepackage{cleveref}
-\usepackage{subcaption}
-\usepackage[normalem]{ulem}
-\usepackage{xpatch}
-\usepackage{stmaryrd}
+A potential customer Alice might not want to reveal $x$ to Bob because of privacy concerns.
+And Bob does not want to reveal $w$ and $b$ because then Alice can steal the model.
+SMC can help Alice and Bob compute $w^Tx>b$ securely, such that Alice receives the classifier's
+prediction and Bob learns nothing about $x$. Moreover, Alice does not learn anything more about $w$
+and $b$ than what is revealed by the prediction.
 
-\makeatletter
-\xpatchcmd{\algorithmic}{\itemsep\z@}{\itemsep=0.3ex plus2pt}{}{}
-\makeatother
+To implement this system, Bob can write the code in \tool and this code is shown in Figure~\ref{fig:ex-sml}.
+The expression {\tt input1} reads a value from Bob and {\tt input2} reads from Alice.
+The language has arrays and simple loops. A loop {\tt for i = [0:N]} repeats its
+body {\tt N} times, the loop counter {\tt i} is assigned {\tt 0} in the first iteration,
+and is incremented by one after each iteration. Unbounded {\tt while} loops are problematic
+from a cryptographic standpoint and our language only permits these simple {\tt for} loops. 
+The first loop in Figure~\ref{fig:ex-sml} reads the model $w$ from Bob, the second loop reads
+Alice's medical report $x$, and the third loop computes the dot product $w^Tx$.
+The last {\tt output} statement sends the result of the comparison $w^Tx>b$ to Alice.
+In particular, the ternary {\tt ? :} operator performs a branch and the result is the second (third)
+argument if the first argument is true (false).
 
-\renewcommand{\algorithmicrequire}{\textbf{Input:}}
-\renewcommand{\algorithmicensure}{\textbf{Output:}}
-\newcommand{\crefrangeconjunction}{ to~}
-\newcommand{\minus}{\scalebox{0.75}[1.0]{$-$}}
-\newcommand{\sameer}[1]{{\leavevmode\color{red}{#1}}}
-\setlength{\textfloatsep}{0.3cm}
-\newcommand{\ourprotocol}{Tuffy}
+Our compiler compiles the code described in Figure~\ref{fig:ex-sml} to the C++ code in
+Figure~\ref{fig:ex-aby}. An alternative to using \tool is to directly write this C++ code.
+However, this code is much more complex. 
+There are two phases of this code, circuit generation and circuit execution.
+In the first phase, we create a circuit that represents the functionality that we want to compute.
+And in the second phase, we run the circuit generated by the first phase and get the results.
+The code uses the
+{\tt role} variable to decide whether the code is executed by the {\tt SERVER} Bob or by
+the {\tt CLIENT} Alice. The variables with names that start with {\tt s\_a\_} denote that
+they point to circuit gates that compute arithmetic shares. In particular, {\tt s\_a\_w[i]} 
+is an arithmetic share of {\tt w[i]}.
+We are using arithmetic sharing as the dot product computation is more efficient in arithmetic
+than using boolean sharing.
+After the dot product computation (the second loop), {\tt s\_a\_acc}  points to the gate
+which would compute the arithmetic share of the final value of {\tt acc}.
+The {\tt PutA2Y gate} is a conversion gate that converts an arithmetic share of
+{\tt acc} to a Yao share. We need Yao shares for the comparison operator ({\tt >}).
+This operator cannot be implemented on arithmetic shares and needs a boolean sharing like Yao.
+After the {\tt PutGTGate} for comparison, we generate gates that create Yao shares of zero and one.
+Next, a multiplexer gate {\tt PutMUXGate}} encodes the logic of the {\tt ? :} operator.
+Finally, we add an output gate by {\tt PutOUTGate}.
 
-\newcommand{\bits}{\{0,1\}}
-\newcommand{\Ex}{\mathbb{E}}
-\newcommand{\To}{\rightarrow}
-\newcommand{\R}{\mathbb{R}}
-\newcommand{\N}{\mathbb{N}}
-\newcommand{\Z}{\mathbb{Z}}
-\newcommand{\F}{\mathbb{F}}
-\newcommand{\GF}{\mathsf{GF}}
-
-\usepackage{MnSymbol}
-\newcommand{\maxpr}{\text{\rm max-pr}}
-\newcommand{\cclass}[1]{\mathbf{#1}}
-\renewcommand{\P}{\cclass{P}}
-\newtheorem{problem}{Problem}
-\newtheorem{theorem}{Theorem}
-\newtheorem{conjecture}[theorem]{Conjecture}
-\newtheorem{definition}[theorem]{Definition}
-\newtheorem{lemma}[theorem]{Lemma}
-\newtheorem{proposition}[theorem]{Proposition}
-\newtheorem{corollary}[theorem]{Corollary}
-\newtheorem{claim}[theorem]{Claim}
-\newtheorem{fact}[theorem]{Fact}
-\newtheorem{remk}[theorem]{Remark}
-\newtheorem{apdxlemma}{Lemma}
-
-\newcommand{\etal}{{\it et~al.\ }}
-\newcommand{\ie} {{\it i.e.,\ }}
-\newcommand{\eg} {{\it e.g.,\ }}
-\newcommand{\cf}{{\it cf.,\ }}
-\newcommand{\suchthat}{{\;\; : \;\;}}
-\newcommand{\pr}[1]{\Pr\left[#1\right]}
-
-\newcommand{\class}[1]{\mathbf{#1}}
-\newcommand{\SZK}{\class{SZK}}
-\newcommand{\BPP}{\class{BPP}}
-\newcommand{\NP}{\class{NP}}
-\renewcommand{\P}{\class{P}}
-\newcommand{\negl}{{\mathrm{neg}}}
-
-\newcommand{\Enc}{E}
-\newcommand{\Dec}{D}
-\newcommand{\Gen}{G}
-
-\newcommand{\PK}{\mathit{PK}}
-\newcommand{\SK}{\mathit{SK}}
-\newcommand{\pk}{\mathit{pk}}
-\newcommand{\sk}{\mathit{sk}}
-\newcommand{\MD}[1]{\mathrm{MD{#1}}}
-\newcommand{\SHA}{\mbox{SHA-1}}
-
-\begin{document}
-\title{SecureDNN: Efficient Analytics over Encrypted Data}
-
-% author names and affiliations
-% use a multiple column layout for up to three different
-% affiliations
-% \author{\IEEEauthorblockN{Michael Shell}
-% \IEEEauthorblockA{School of Electrical and\\Computer Engineering\\
-% Georgia Institute of Technology\\
-% Atlanta, Georgia 30332--0250\\
-% Email: http://www.michaelshell.org/contact.html}
-% \and
-% \IEEEauthorblockN{Homer Simpson}
-% \IEEEauthorblockA{Twentieth Century Fox\\
-% Springfield, USA\\
-% Email: homer@thesimpsons.com}
-% \and
-% \IEEEauthorblockN{James Kirk\\ and Montgomery Scott}
-% \IEEEauthorblockA{Starfleet Academy\\
-% San Francisco, California 96678-2391\\
-% Telephone: (800) 555--1212\\
-% Fax: (888) 555--1212}}
-
-\maketitle
-% \begin{abstract}
-% The abstract goes here.
-% \end{abstract}
-\IEEEpeerreviewmaketitle
-
-\graphicspath{{./Images/}}
-\DeclareGraphicsExtensions{.pdf,.jpg,.png}
-
-\input{./Sections/CompleteProtocol.tex}
-\newpage
-\clearpage
-\input{./Sections/ReLU.tex}
-\input{./Sections/Independent.tex}
-\input{./Sections/Introduction.tex}
-
-\newpage
-\bibliographystyle{abbrv}
-\bibliography{./bib}
-\balance
-
-\end{document}
+Next, we execute this circuit using {\tt ExecCircuit} and Alice reads the output result.
+Our compiler automatatically compiles Figure~\ref{fig:ex-sml} to Figure~\ref{fig:ex-aby}.
+In particular, it manages the shares and inserts conversions between different sharing schemes
+wherever necessary. 
+On small examples, we have found that the compiler generated C++ code from a \tool implementation 
+is almost as efficient as manually written C++ code.

### 2017-10-12 OaklandSubmission/eval.tex

@@
 \label{sec:shallow}
-We consider three types of classifiers: linear, Naive Bayes, and decision trees. 
+We consider three types of classifiers: linear, Na\"{i}ve Bayes, and decision trees. 
 These classifiers are trained on the followinng data sets from the UCI  machine learning repository~\cite{uci}:
@@
 
-The results for Na\"{i}ve Bayes are in Table~\ref{tab:nb}. As before, $n$ denotes the length of the input feature vector $x=(x_1,\ldots,x_n)$ and $F$ is the number of possible values of each $x_i$.
-Although, there are no multiplications in these benchmarks, these do have a significant number of comparisons that result in many boolean-and gates (13k/750k denote 13000/75000). 
+The results for Na\"{i}ve Bayes are in Table~\ref{tab:nb}. As before, $d$ denotes the length of the input feature vector $x$ and $F$ is the number of possible values of $x_i$ for each $i \in [d]$.
+Although, there are no multiplications in these benchmarks, these do have a significant number of comparisons (due to secret look-up) that result in many boolean-and gates (13k/750k denote 13000/75000).
 Here, again columns 4 and 5 show the results reported in~\cite{shafindss}.
@@
 \begin{tabular}{c|c|c|c|c|c |c|c|c|c|c|c}
-Dataset & $n$ & $F$ & Prev. time (s) & Prev. Comm (Mb) & LAN time (s) & WAN time (s) & Comm. (Mb)  & \#And & \#Mul & \#Gates  & LOC\\
+Dataset & $d$ & $F$ & Prev. time (s) & Prev. Comm (Mb) & LAN time (s) & WAN time (s) & Comm. (Mb)  & \#And & \#Mul & \#Gates  & LOC\\
 \hline
@@
 
- \caption{Naive Bayes results}
+ \caption{Na\"{i}ve Bayes results}
  \label{tab:nb} 

### 2017-10-12 OaklandSubmission/impl.tex

\section{Implementation}
\label{sec:impl}
\divya{Some pointers by Divya}
Describe the following
\begin{enumerate}
\item python compiler etc
\item array index out of bounds static check implementation
\item pipelining? automatic or manual?
\item secret indexing in arrays; not supported directly but coded as a tree; future work automatic compilation may be

\end{enumerate}

Where do we mention about getting integer models; using unsigned ints for signed compute etc?

### 2017-10-12 OaklandSubmission/ex.tex
minor
overview

@@
 %$w^Tx>0$, mnist logistic regression
-Consider a hypothetical cloud provider Bob that wants to provide a service to diagnose
+Consider a  cloud service provider Bob that wants to provide a service to diagnose
 whether a patient has breast cancer or not. Bob trains a machine learning classification model
@@
 Given a patient's medical report, in the form of a vector $x$ (also of length 30),
-the classifer predicts that the patient has breast cancer iff $w^Tx>b$.
+the classifer predicts that the patient has breast cancer if $w^Tx>b$.
 However, for this task Bob needs access to $x$, which is private data of a customer.
@@
 The last {\tt output} statement sends the result of the comparison $w^Tx>b$ to Alice.
-In particular, the ternary {\tt ? :} operator performs a branch and the result is the second (third)
+In particular, the ternary ``{\tt ? :}" operator performs a branch and the result is the second (third)
 argument if the first argument is true (false).
 
-Our compiler compiles the code described in Figure~\ref{fig:ex-sml} to the C++ code in
+The \tool compiler is fully automatic and compiles the code described in Figure~\ref{fig:ex-sml} to the C++ code that makes calls to the ABY library  in
 Figure~\ref{fig:ex-aby}. An alternative to using \tool is to directly write this C++ code.
-However, this code is much more complex. 
-There are two phases of this code, circuit generation and circuit execution.
-In the first phase, we create a circuit that represents the functionality that we want to compute.
-And in the second phase, we run the circuit generated by the first phase and get the results.
-The code uses the
-{\tt role} variable to decide whether the code is executed by the {\tt SERVER} Bob or by
-the {\tt CLIENT} Alice. The variables with names that start with {\tt s\_a\_} denote that
-they point to circuit gates that compute arithmetic shares. In particular, {\tt s\_a\_w[i]} 
-is an arithmetic share of {\tt w[i]}.
-We are using arithmetic sharing as the dot product computation is more efficient in arithmetic
-than using boolean sharing.
-After the dot product computation (the second loop), {\tt s\_a\_acc}  points to the gate
-which would compute the arithmetic share of the final value of {\tt acc}.
-The {\tt PutA2Y gate} is a conversion gate that converts an arithmetic share of
-{\tt acc} to a Yao share. We need Yao shares for the comparison operator ({\tt >}).
-This operator cannot be implemented on arithmetic shares and needs a boolean sharing like Yao.
-After the {\tt PutGTGate} for comparison, we generate gates that create Yao shares of zero and one.
-Next, a multiplexer gate {\tt PutMUXGate}} encodes the logic of the {\tt ? :} operator.
-Finally, we add an output gate by {\tt PutOUTGate}.
+However, this code is much more complex than the implementation written in \tool.
+Unlike \tool implementations, the ``secret" variables containing private data ({\tt w,x,b,acc}) need to be handled differently compared to public variables such as loop counters ({\tt i}).
+In particular, the former are manipulated by the ABY library and the latter are manipulated like standard  C++ variables.
+Additionally, the C++ code branches on  the special variable
+{\tt role}  to decide whether the enclosing code is executed by the {\tt SERVER} Bob or by
+the {\tt CLIENT} Alice. In Figure~\ref{fig:ex-sml}, the developer does not need to
+manipulate {\tt role} explicitly. In addition to simplifying the implementation, \tool prevents developers from writing buggy code such as\\
+\verb+if(role == SERVER) {...manipulate x...}+
 
-Next, we execute this circuit using {\tt ExecCircuit} and Alice reads the output result.
-Our compiler automatatically compiles Figure~\ref{fig:ex-sml} to Figure~\ref{fig:ex-aby}.
-In particular, it manages the shares and inserts conversions between different sharing schemes
-wherever necessary. 
-On small examples, we have found that the compiler generated C++ code from a \tool implementation 
-is almost as efficient as manually written C++ code.
+Furthermore, the code in Figure~\ref{fig:ex-aby} is too low level.
+In particular, it is first builds a circuit that represents the computation
+$w^Tx>b$ and then executes it via a call to {\tt ExecCircuit}.
+To build a circuit, the code adds multiplication gates ({\tt PutMULGate})
+and addition gates ({\tt PutADDGate}) for the dot product.
+Subsequently, it uses a ``greater than" gate ({\tt PutGTGate})
+and a multiplexer gate ({\tt PutMUXGate}) for the comparison with $b$.
+For efficiency, the operations for dot product need to be performed using 
+arithmetic circuits. However, ABY's arithmetic circuits cannot express
+comparisons and these need  boolean circuits.
+A program that uses both arithmetic circuit and boolean circuits
+requires conversion gates that help  interconvert between arithmetic
+representations and boolean representations ({\tt PutA2YGate}).
+Bigger computations often require multiple conversions and the
+developer effort can quickly become significant.
+
+Moreover, it is difficult to maintain the code in Figure~\ref{fig:ex-aby}.
+For example, if the multiplication is changed to a bitwise-or then,
+in the absence of \tool, an efficiency-conscious developer would need to  use boolean circuits everywhere and would be tasked with removing all arithmetic circuits and the interconversion gates. 
+ With \tool, the developer needs to change only a single character (replace {\tt *} in Figure~\ref{fig:ex-sml} with {\tt |}) to achieve the same effect. 
+On this example and other small examples, we have found that the compiler generated C++ code from a \tool implementation is as efficient as manually written C++ code.
+In particular, for $w^Tx>b$, the compiler automatically generates code that uses arithmetic circuits for dot product and inserts appropriate conversion gates.
+Finally, although this paper uses ABY for evaluation, the  compiler can be easily retargeted to generate code for other cryptography backends.
+
+The \tool compiler provides strong static guarantees. For example, a well typed
+\tool program is guaranteed to execute to termination without errors.
+A \tool program cannot go into non-termination or dereference illegal memory, i.e., no buffer overflows or underflows can happen at run time.
+This guarantees are
+much stronger than what are offered by production compilers such as {\tt gcc}.
+The primary reason that we are able to achieve these guarantees is because
+\tool has been designed to be verifiable and is backed by formal semantics. 
+Furthermore, the compiler output is guaranteed to be a cryptographically secure
+implementation of the functionality declared by the \tool code. 
+Moreover, the \tool compiler also generates a C implementation that can be run natively on a single machine for functional testing.
+
+To summarize, \tool raises the level of abstraction, provides strong static guarantees, and generates efficient code automatically.

### 2017-10-12 OaklandSubmission/ex.tex
edits and comments to Sec 1 and 2.

@@
 and $b$ than what is revealed by the prediction. 
-
+\nc{This is a bit repetitive from the intro, isnt it? Should we have another example perhaps?}
 To implement this system, Bob can write the code in \tool and this code is shown in Figure~\ref{fig:ex-sml}.
@@
 Finally, although this paper uses ABY for evaluation, the  compiler can be easily retargeted to generate code for other cryptography backends.
-
+\nc{I like the above point. Perhaps we can mention something like this earlier in the intro itself?}
 The \tool compiler provides strong static guarantees. For example, a well typed
@@
 A \tool program cannot go into non-termination or dereference illegal memory, i.e., no buffer overflows or underflows can happen at run time.
-This guarantees are
+These guarantees are
 much stronger than what are offered by production compilers such as {\tt gcc}.

### 2017-10-12 OaklandSubmission/eval.tex
minor

@@
 Therefore, fully connected layers are lighter computation-wise compared
-to convolution layers. In general, DNNs are computationaly heavy but provide
+to convolution layers. However, the model size of fully connected layers is larger than those of convolution layers.
+In general, DNNs are computationaly heavy but provide
 much better accuracies on computer vision tasks than the classifiers discussed above.
@@
 The total execution time of \tool generated code in the LAN and the WAN setting is reported next followed by the total communication.
+We observe that \tool generated code has performance competitive to~\cite{shafindss}.
 The execution time is dominated by multiplications and boolean-ands.
@@
 
- \caption{Linear classification results}
+ \caption{Linear classification results. We compare our results (columns 5, 6, 7) with~\cite{shafindss} (columns 3 and 4)}
  \label{tab:lc} 
@@
 
- \caption{Na\"{i}ve Bayes results}
+ \caption{Na\"{i}ve Bayes results. We compare our results (columns 6 , 7, 8) with~\cite{shafindss} (columns 4 and 5)}
  \label{tab:nb} 
@@
 
- \caption{Decision tree benchmarks}
+ \caption{Decision tree benchmarks. We compare our results (columns 5 , 6, 7) with~\cite{wu} (columns 3 and 4)}
  \label{tab:dt} 
@@
 Recall, that square activation can be implemented entirely using arithmetic gates but ReLU requires boolean-and gates.
-For a complete description of these benchmarks,  the reader is referred to the original references.
+For a complete description of these benchmarks and their accuracies,  the reader is referred to the original references.
 
@@
 outperforms SecureML and Cryptonets and columns 2 and 3 show the time taken by \minion. 
+The column ``Model size'' is the number of 32 bit integers in the trained model.
 We observe that our performance is competitive with \minion in the LAN and the WAN settings and the lines of code required are still small.
@@
 \begin{table*}
-\begin{tabular}{c|c|c|c |c|c|c|c|c|c}
-Name  & Prev. time (s) & Prev. Comm (Mb) & LAN time (s) & WAN time (s) & Comm. (Mb)  & \#And & \#Mul & \#Gates & LOC\\
+\begin{tabular}{c|c|c|c |c|c|c|c|c|c|c}
+Name  & Prev. time (s) & Prev. Comm (Mb) & LAN time (s) & WAN time (s) & Comm. (Mb)  & \#And & \#Mul & \#Gates & Model Size & LOC\\
 \hline
-SecureML   &  1.1 & 15.8 & 0.7 & 1.7  & 76   &  2k   & 119k & 366k   & 78\\
+SecureML   &  1.1 & 15.8 & 0.7 & 1.7  & 76   &  2k   & 119k & 366k   & 119k & 78\\
 \hline
-Cryptonets &  1.3 & 47.6 & 0.6 & 1.6  & 7    & 2k    & 108k & 316k & 88\\
+Cryptonets &  1.3 & 47.6 & 0.6 & 1.6  & 7    & 2k    & 108k & 316k & 86k & 88\\
 \hline
-CNN        &  9.4 & 657.5& 5.1 & 11.6 & 501  & 1640k & 667k & 9480k & 154\\
+CNN        &  9.4 & 657.5& 5.1 & 11.6 & 501  & 1640k & 667k & 9480k & 35k & 154\\
 \hline
@@
 
- \caption{DNN benchmarks}
+ \caption{DNN benchmarks. We compare our results (columns 4 , 5, 6) with~\cite{minionn} (columns 2 and 3)}
  \label{tab:nn} 
@@
 \]
-Here, $i\in[0,9]\cap\mathbb{Z}$, $x$ is a 784 length vector obtained from the input image,
+Here, $i\in [10]$, $x$ is a 784 length vector obtained from the input image,
 each $w_i$ is a 784 length vector, and each $b_i$ is a scalar. 
@@
 we only report execution time of \tool generated code.
-The column ``Model size'' is the number of 32 bit integers in the trained model.
 
@@
 program grows with the depth of the tree as the straightforward \tool implementation requires a 
-loop over the nodes of the tree and this loop needs to be unrolled.
+loop for each layer of the tree.
 

### 2017-10-12 OaklandSubmission/ex.tex
minor

--- 2017-10-12 OaklandSubmission/ex.tex
@@
 However, this code is much more complex than the implementation written in \tool.
-Unlike \tool implementations, the ``secret" variables containing private data ({\tt w,x,b,acc}) need to be handled differently compared to public variables such as loop counters ({\tt i}).
+Therefore, it is difficult for Alice and bob to verify its correctness and security.
+Unlike \tool implementations, the ``secret" variables containing private data ({\tt w,x,b,acc}) need to be handled differently compared to ``public" variables such as loop counters ({\tt i}).
 In particular, the former are manipulated by the ABY library and the latter are manipulated like standard  C++ variables.
@@
 For example, if the multiplication is changed to a bitwise-or then,
-in the absence of \tool, an efficiency-conscious developer would need to  use boolean circuits everywhere and would be tasked with removing all arithmetic circuits and the interconversion gates. 
+in the absence of \tool, an efficiency-conscious developer would need to  use boolean circuits everywhere and would be tasked with removing all arithmetic circuits and the interconversion gates (the cost of  conversion overweighs the gains of performing an arithmetic addition instead of a boolean addition).
  With \tool, the developer needs to change only a single character (replace {\tt *} in Figure~\ref{fig:ex-sml} with {\tt |}) to achieve the same effect. 
@@
 Finally, although this paper uses ABY for evaluation, the  compiler can be easily retargeted to generate code for other cryptography backends.
+In particular,
+ABY implements two party computation (2PC), i.e., SMC between two parties
+and this paper focuses on this setting.
 

### 2017-10-12 OaklandSubmission/ex.tex

@@
 Finally, although this paper uses ABY for evaluation, the  compiler can be easily retargeted to generate code for other cryptography backends.
-In particular,
-ABY implements two party computation (2PC), i.e., SMC between two parties
-and this paper focuses on this setting.
+\nc{I like the above point. Perhaps we can mention something like this earlier in the intro itself?}
 

### 2017-10-12 OaklandSubmission/intro.tex

@@
 This paper provides a language-based solution to make
-secure multiparty computation (SMC) accessible to the software developer community.
+secure multiparty computation (SMC) accessible to the software developer community. \nc{Why do we call it SMC? A bit non-standard for crypto audiences isnt it?}
 
@@
 Examples of these include various machine learning prediction algorithm where one party holds the secret input such as a medical report and the other party holds the ML model for disease prediction.
-Since the introduction of SMC in 1980's, there has been a large body of work \cite{..} that has transformed SMC from a mere theoretical tool to something practically efficient/usable.
+Since the introduction of SMC in 1980's, there has been a large body of work \cite{..} that has transformed SMC from a mere theoretical tool to something practically efficient and usable.
 Hence, there have been efforts to make SMC accessible to non-specialist programmers.
@@
 With the current techniques, he has two options.
-He can code up the functionality in a programmer friendly domain-specific language such as \cite{lambdaps,wysteria,oblivm} that would automatically compile it to an SMC protocol.
+\nc{The two options appear quite far apart in the text.
+Can we make bullet points maybe?} He can code up the functionality in a programmer friendly domain-specific language such as \cite{lambdaps,wysteria,oblivm} that would automatically compile it to an SMC protocol.
 The downside is that all of these works build on a cryptographic back-end that is either entirely boolean \cite{yao,gmw} or entirely arithmetic \cite{homo}.
@@
 In short, as is also mentioned by the authors themselves, the current system is not suitable to be used by non-specialist programmers.
-\divya{Change tone...}
+\nc{Also say that the programmer would need to know which functionality is better suited for arithmetic/boolean and also how to split the complete functionality into such parts.} \divya{Change tone...}
 Our work .....
+\nc{Should we also say here that a third option is to hire a cryptographer and build specific crypto protocols for the functionalities?
+:) This seems to be the option we are targetting when we show our comparisons and we say that we beat or match hand-crafted protocols}
 
@@
 
-In this work, we develop and implement our framework \tool for secure computation that achieves generality, performance and programmer productivity.
+In this work, we develop and implement our framework \tool \nc{I still like EasyPC :)} for secure computation that achieves generality, performance and programmer productivity.
 The programmer writes a high level C-like program (instead of a circuit) to describe the function to be computed.
@@
 For details on formal language, correctness and security theorems see \sectionref{ld}.
+\nc{Shall we have paragraph heads or someting to outline the different contributions? Above feels like one big story.}
 
@@
 Hence, the largest compute that can be done securely is upper-bounded by the largest circuit that can fit in memory on a machine.
+\nc{Should we point out here that ABY itself has an explicit limit on circuit size and also give an example of a program that easily surpasses this bound?
+Also, hadnt ABY also said something about this?
+If so, we can mention this.}
 To address this issue, we develop a technique called \emph{pipelining} that allows us to run aribitrarily large programs.

### 2017-10-12 OaklandSubmission/intro.tex
chat-bot

@@
 This paper provides a language-based solution to make
-secure multiparty computation (SMC) accessible to the software developer community. \nc{Why do we call it SMC? A bit non-standard for crypto audiences isnt it?}
+secure multiparty computation (SMC) accessible to the software developer community.
+\nc{Why do we call it SMC?
+A bit non-standard for crypto audiences isnt it?} \divya{I drank my tears and a part of me died when I did this.
+Apparently thats what it is called outside the theory community.}
 
@@
 \nc{The two options appear quite far apart in the text.
-Can we make bullet points maybe?} He can code up the functionality in a programmer friendly domain-specific language such as \cite{lambdaps,wysteria,oblivm} that would automatically compile it to an SMC protocol.
+Can we make bullet points maybe?} \divya{Bullet list would look bad.
+I tried listing both options first and then discussing but its hard to talk about ABY before.} He can code up the functionality in a programmer friendly domain-specific language such as \cite{lambdaps,wysteria,oblivm} that would automatically compile it to an SMC protocol.
 The downside is that all of these works build on a cryptographic back-end that is either entirely boolean \cite{yao,gmw} or entirely arithmetic \cite{homo}.
@@
 \nc{Should we also say here that a third option is to hire a cryptographer and build specific crypto protocols for the functionalities?
-:) This seems to be the option we are targetting when we show our comparisons and we say that we beat or match hand-crafted protocols}
+:) This seems to be the option we are targeting when we show our comparisons and we say that we beat or match hand-crafted protocols} \divya{We can do this!
+wMight have to see the flow.}
 
@@
 
-In this work, we develop and implement our framework \tool \nc{I still like EasyPC :)} for secure computation that achieves generality, performance and programmer productivity.
-The programmer writes a high level C-like program (instead of a circuit) to describe the function to be computed.
+In this work, we develop and implement our framework \tool \nc{I still like EasyPC :)} \divya{I agree that \tool is bad.
+What would be the full form for EasyPC?
+Shouldn't it be EasySC?} for secure computation that achieves generality, performance and programmer productivity.
+The programmer writes a high level \divya{C-like} program (instead of a circuit) to describe the function to be computed.
 Our compiler automatically generates an SMC protocol using a mix of Arithmetic and Boolean compute.
@@
 %Write high level overview of programming language and tool.. \divya{have some preamble}
+\divya{I can add paragraph headings..}
 
 \tool is a programmer-friendly framework to express the function to be computed and uses a mix of arithmetic and boolean SMC back-ends.
-We illustrate the above toy example of $w^Tx >0$ in our language in \figureref{ex-sml}.
+We illustrate the above toy example of $w^Tx >b$ in our language in \figureref{ex-sml}.
 As is clear from the code, the programmer is not required to write keywords such as {\em secret} or {\em public} and can remain completely oblivious of the cryptographic back-end.
@@
 For details on formal language, correctness and security theorems see \sectionref{ld}.
-\nc{Shall we have paragraph heads or someting to outline the different contributions? Above feels like one big story.}
+\nc{Shall we have paragraph heads or something to outline the different contributions? Above feels like one big story.}
 
@@
 But, it has been observed multiple times this approach does not even scale on realistic examples.
-
-\input{fig-ex-aby}
-\input{fig-ex-sml}
 

### 2017-10-13 OaklandSubmission/intro.tex
intro small changes

@@
 Hence, the second option is to write the functionality in the system developed in \cite{aby}.
-However, using their system requires the programmer to be aware of complex trade-offs between arithmetic and boolean cryptographic schemes.
+However, using their system requires the programmer to be aware of complex trade-offs between arithmetic and boolean cryptographic schemes and split the functionality into such parts accordingly.
 In their framework, the programmer is required to write circuits consisting of  arithmetic and boolean gates along with appropriate conversion gates.
-In short, as is also mentioned by the authors themselves, the current system is not suitable to be used by non-specialist programmers.
-\nc{Also say that the programmer would need to know which functionality is better suited for arithmetic/boolean and also how to split the complete functionality into such parts.} \divya{Change tone...}
-Our work .....
+In short, the current system of ABY is not suitable to be used by non-specialist programmers.
+\nc{Also say that the programmer would need to know which functionality is better suited for arithmetic/boolean and also how to split the complete functionality into such parts.}
+Now the third option is to hire a cryptographer to design specialised protocols for the functionalities. 
+In our work, we achieve the best of all options. \divya{add a word like automatically..}
 \nc{Should we also say here that a third option is to hire a cryptographer and build specific crypto protocols for the functionalities?
 :) This seems to be the option we are targeting when we show our comparisons and we say that we beat or match hand-crafted protocols} \divya{We can do this!
-wMight have to see the flow.}
+Might have to see the flow.}
 

### 2017-10-13 OaklandSubmission/intro.tex
related

@@
 secure multiparty computation (SMC) accessible to the software developer community.
-\nc{Why do we call it SMC?
-A bit non-standard for crypto audiences isnt it?} \divya{I drank my tears and a part of me died when I did this.
-Apparently thats what it is called outside the theory community.}
 
@@
 
-Suppose a developer wants to compute $w^Tx >0$ securely using SMC.
-With the current techniques, he has two options.
-\nc{The two options appear quite far apart in the text.
-Can we make bullet points maybe?} \divya{Bullet list would look bad.
-I tried listing both options first and then discussing but its hard to talk about ABY before.} He can code up the functionality in a programmer friendly domain-specific language such as \cite{lambdaps,wysteria,oblivm} that would automatically compile it to an SMC protocol.
+Suppose a developer wants to compute $w^Tx >0$ securely using SMC. With the current techniques, he has two options.  
+\begin{tiret}
+\item He can code up the functionality in a programmer friendly domain-specific language such as \cite{lambdaps,wysteria,oblivm} that would automatically compile it to an SMC protocol.
 The downside is that all of these works build on a cryptographic back-end that is either entirely boolean \cite{yao,gmw} or entirely arithmetic \cite{homo}.
@@
 In fact, most interesting functions require a mix of arithmetic and boolean computations.
-To address this issue, recently Demmler et al.
+\item To address this issue, recently Demmler et al.
 \cite{aby} gave a cryptographic protocol for SMC where the parties can mix arithmetic and boolean computations.
@@
 In short, the current system of ABY is not suitable to be used by non-specialist programmers.
-\nc{Also say that the programmer would need to know which functionality is better suited for arithmetic/boolean and also how to split the complete functionality into such parts.}
-Now the third option is to hire a cryptographer to design specialised protocols for the functionalities. 
-In our work, we achieve the best of all options. \divya{add a word like automatically..}
-\nc{Should we also say here that a third option is to hire a cryptographer and build specific crypto protocols for the functionalities?
-:) This seems to be the option we are targeting when we show our comparisons and we say that we beat or match hand-crafted protocols} \divya{We can do this!
-Might have to see the flow.}
+
+\item Now the third option is to hire a cryptographer to design specialised protocols for the functionalities. 
+\end{tiret}
+In our work, we achieve the best of all options.
 
@@
 
-\tool is a programmer-friendly framework to express the function to be computed and uses a mix of arithmetic and boolean SMC back-ends.
+\textbf{Programmability.} \tool is a programmer-friendly framework to express the function to be computed and uses a mix of arithmetic and boolean SMC back-ends.
 We illustrate the above toy example of $w^Tx >b$ in our language in \figureref{ex-sml}.
@@
 
-Our compiler is aware of cryptographic costs of different representations and inter-conversions.
+\textbf{Performance.} Our compiler is aware of cryptographic costs of different representations and inter-conversions.
 Based on these costs, our compiler, given a program in \tool, automatically picks Arithmetic and Boolean representations for different sub-parts of the program and inserts the required inter-conversions as well.
@@
 
-We give a formal {\em type system} for our language and prove useful properties.
+\textbf{Type system.} We give a formal {\em type system} for our language and prove useful properties.
 We prove that if a program type checks in our language then it would run correctly, e.g., there would not be any array index out of bounds error.
@@
 
-Another issue in adoption of SMC is difficulty in scaling to large programs. 
+\textbf{Pipelining.} Another issue in adoption of SMC is difficulty in scaling to large programs. 
 %In general, SMC implementations struggle in scaling to large programssize of memory on a system limits the largest compute inside the SMC and hence,. 
@@
 
-Finally, our framework \tool can serve as a meaningful baseline for future works on hand-crafted SMC protocols for specific functionalities such as \textsc{MiniONN} \cite{minionn}. 
+\textbf{Baseline.} Finally, our framework \tool can serve as a meaningful baseline for future works on hand-crafted SMC protocols for specific functionalities such as \textsc{MiniONN} \cite{minionn}. 
 Prior to our work, garbled circuit based two-party computation is cited as the most practical approach and used as the baseline for comparison of performance.
@@
 
-\noindent\textbf{Related Work} \divya{I think we should say a few words here about the following: minionn, verilog paper, 
+\noindent\textbf{Related Work} \divya{I think we should say a few words here about the following: minionn, verilog paper, tasty, 
 \cite{lambdaps,wysteria}

### 2017-10-13 OaklandSubmission/related.tex
edits and comments to Sec 1 and 2.
related

@@
 \item Secure Computation of MIPS Machine Code https://eprint.iacr.org/2015/547
-\item TASTY https://eprint.iacr.org/2010/365.pdf
+\item TASTY https://eprint.iacr.org/2010/365.pdf. Does not have language; user has to know what parts in HE and what in Yao and writes something similar to ABY
 \item ABY cites many papers on first page of intro for using a mix of HE and boolean MPC; but they didnt give much gain because HE is expensive
+\nc{I will look into this and write.}
+\item verilog, hardware tool chain paper http://encrypto.de/papers/DDKSSZ15.pdf
+\item automatic choosing between HE and Yao https://eprint.iacr.org/2014/200.pdf 
+I think this one still does not have a language; I don't know how to specify function
 

### 2017-10-13 OaklandSubmission/intro.tex
minor

@@
 \begin{tiret}
-\item He can code up the functionality in a programmer friendly domain-specific language such as \cite{lambdaps,wysteria,oblivm} that would automatically compile it to an SMC protocol.
+\item He can code up the functionality in a programmer friendly domain-specific language such as \cite{lambdaps,wysteria,oblivm,smcl} that would automatically compile it to an SMC protocol.
 The downside is that all of these works build on a cryptographic back-end that is either entirely boolean \cite{yao,gmw} or entirely arithmetic \cite{homo}.

### 2017-10-13 OaklandSubmission/intro.tex
intro related

@@
 In contrast, \figureref{ex-aby} illustrates the same example using the framework in ABY \cite{aby}.
-Here, as already mentioned, the programmer needs to write the low level gates like $\mathrm{MULT, ADD, CONS, GT}$, etc.
-The programmer is also required to use the appropriate share conversion gates such as A2Y that converts Arithmetic shares to Yao shares.
+Here, as already mentioned, the programmer needs to write the low level gates like {\tt MULT, ADD, CONS, GT}, etc.
+The programmer is also required to use the appropriate share conversion gates such as {\tt A2Y} that converts Arithmetic shares to Yao shares.
+Moreover, it is easy to maintain code in \tool and small change to the functionality requires only a small change in the code.
+In contrast, working with ABY, even for a small change a developer might have to re-work the whole circuit for efficiency reasons.
+We describe these examples in detail in \sectionref{ex}.
 
@@
 
-\noindent\textbf{Related Work} \divya{I think we should say a few words here about the following: minionn, verilog paper, tasty, 
-\cite{lambdaps,wysteria}
-
-}
-
-\newpage
+\noindent\textbf{Other Related Work} 
+Before ABY \cite{aby}, to improve efficiency, several works proposed
+to mix secure computation protocols based on homomorphic
+encryption with Yao’s garbled circuits protocol, e.g., \cite{barni,blanton,brickell,franz,huang,valeriaMatrix,valeriaRidge,shropfer11}
+  and developed tools that allowed writing a mixed SMC protocol, e.g., \cite{bogdanov,lone,tasty}. The assignment can
+even be done automatically as was shown in \cite{autoS}. 
+However, as noted by ABY, due to high conversion cost between homomorphic encryption and Yao's garbled circuit, these mixed protocols do not gain much over single protocol.
+Also, none of these works give a high-level developer-friendly language support.
+No prior work has given a formal type system and operational semantics for mixed arithmetic and boolean back-ends.
+In our evaluation, we use benchmarks for the problem of secure prediction where the cloud holds the secret model and the client holds the secret input.
+For this problem various works have considered tailor-made protocols depending on the model considered \cite{shafindss,wu,secureml,minionn}.
+In our evaluation, we compare our performance with these specialized protocols and give competitive run-times.
+We provide a detailed survey of related work in \sectionref{related}.

### 2017-10-13 OaklandSubmission/intro.tex
reference

@@
 to mix secure computation protocols based on homomorphic
-encryption with Yao’s garbled circuits protocol, e.g., \cite{barni,blanton,brickell,franz,huang,valeriaMatrix,valeriaRidge,shropfer11}
+encryption with Yao’s garbled circuits protocol, e.g., \cite{barni,blanton,brickell,franz,huang,valeriaMatrix,valeriaRidge,schropferK11}
   and developed tools that allowed writing a mixed SMC protocol, e.g., \cite{bogdanov,lone,tasty}. The assignment can

### 2017-10-13 OaklandSubmission/related.tex
first pass at related work.

--- 2017-10-11 OaklandSubmission/related.tex
@@
 
-\divya{Related Work: writing some citations here to be used appropriately.}
+%\divya{Related Work: writing some citations here to be used appropriately.}
+%\begin{enumerate}
+%\item Secure Computation of MIPS Machine Code https://eprint.iacr.org/2015/547
+%\item TASTY https://eprint.iacr.org/2010/365.pdf. Does not have language; user has to know what parts in HE and what in Yao and writes something similar to ABY
+%\item ABY cites many papers on first page of intro for using a mix of HE and boolean MPC; but they didnt give much gain because HE is expensive
+%\nc{I will look into this and write.}
+%\item verilog, hardware tool chain paper http://encrypto.de/papers/DDKSSZ15.pdf
+%\item automatic choosing between HE and Yao https://eprint.iacr.org/2014/200.pdf 
+%I think this one still does not have a language; I don't know how to specify function
+
+%\end{enumerate}
+\nc{First pass of the related work.}\\
+\nc{ABY have a list of papers that combine boolean and arithmetic compute (hand crafted protocols). We can cite them in the intro or somewhere else to say all those can be automated.}
+
+\noindent\textbf{Automatic Generation of SMC protocols.}
+There has been vast literature on the automatic generation of SMC protocols.
+These works all provide tools at various levels of abstraction that help a programmer automatically generate cryptographic protocols for various functionalities.
+These levels can be classified as follows:
+\begin{tiret}
+\item {\bf Function Description:} Tools at this level are a language (similar to \tool) where the programmer can specify the functionality to be securely implemented in a high level language and the compiler compiles the functionality to generate a secure computation protocol.
+Such languages do not require the programmer to have expertise in writing cryptographic protocols.
+All such existing languages in literature are built with only one backend: Fairplay's Secure Function Definition Language (SFDL)~\cite{fairplay,fairplaymp} compiles high-level C or Pascal like programs into boolean circuits that are then evaluated using garbled circuits~\cite{yao} for the two-party case and distributed garbled circuits~\cite{bmr} for the multi-party case~\cite{bmr}.
+The Secure Multiparty Computation Language (SMCL)~\cite{smcl} that compiles high-level Java like programs into arithmetic circuits that are then evaluated using the VIFF framework~\cite{viff}.
+The work of Wysteria~\cite{wysteria} provides a way of writing ``mixed-mode'' programs that combine normal or per-party, local computations, with ``secure'' (joint) computations repeatedly in a seamless manner.
+This work compiles programs into boolean circuits and uses Choi {\em et al.}~\cite{choi} implementation of the GMW~\cite{gmw} protocol as the backend.
+The work of Mitchell {\em et al.}~\cite{lambdaps} compiles their language into either a backend that works over arithmetic secret shares or over boolean secret shares using homomorphic encryption~\cite{gentry}.
+Finally, the work of ObliVM~\cite{oblivm} also shows how to incorporate programs that may read/write portions of the secret data, based on other secrets, without revealing access patterns (using techniques from oblivious RAM~\cite{oram1,oram2}).
+Their compute is through boolean circuits and the garbling technique.
+It is to be noted that none of the above works combine arithmetic and boolean computation together within the same program.
+\item {\bf Protocol Description:} Tools at this level of abstraction provide a set of commands that can be put together to describe efficient secure computation protocols.
+These tools are at a high level (such as making a call to execute secure computation of an AND gate or a MULT gate over shared inputs).
+The exact cryptographic protocols that are used as backends to implement these secure computation protocols need not be known to the programmer, but in order to generate efficient protocols for the entire functionality, the programmer must know how the functionality will be broken into components that make calls to various secure computation protocols for sub-functions.
+As an example, the work of ABY \cite{aby} falls into this category.
+While, the programmer need not know how ABY implements boolean or arithmetic computation, he or she must know the type of compute (boolean or arithmetic) that is more efficient for specific sub-computations and also explicitly write calls to convert shares from one scheme to another (boolean secret sharing to arithmetic secret sharing and vice-versa).
+While the work of Kerschbaum {\em et al.}~\cite{kos14} provides a heuristic to automatically select the protocol for various types of sub-computations, one would still need to manually write the program using the share conversion tools of ABY.
+Similarly, the TASTY tool~\cite{tasty} provides a way to mix both boolean and arithmetic circuits (using homomorphic encryption for the arithmetic compute).
+As noted by ABY~\cite{aby}, the performance is much poorer than tools like ABY that combine boolean and arithmetic secure computation techniques due to very high conversion costs between homomorphic encryption and boolean shares as well as the cost of homomorphic encryption itself as the security parameter increases.
+This abstraction requires a fairly involved knowledge of cryptography from the programmer.
+Other examples include the VIFF framework~\cite{viff} for arithmetic computations, Sharemind~\cite{sharemind} (that provides information-theoretic secure 3-party boolean computation), and the Secure Multiparty Computation language~\cite{nsy04,securemc}.
+\item {\bf Protocol Implementation:} Tools at this level of abstraction require the programmer to have a deep understanding of how cryptographic secure computation protocols work and the programmer must provide calls to low-level cryptographic building blocks such as oblivious transfer.
+While the programmer need not know the specific oblivious transfer protocol implemented in the backend, he or she still needs to know how to build a secure computation protocol from such building blocks.
+This abstraction requires the highest knowledge of cryptography from the programmer.
+Examples of such tools include L1~\cite{l1} and Qilin~\cite{qilin}.
+\end{tiret}
+
+Other low level implementations of secure computation protocols look to optimize circuits for depth (such as~\cite{ddkssz15}) or provide protocols for MIPS code~\cite{mips} using techniques from ORAM to hide access patterns~\cite{oram1,oram2}.
+\\\\
+\noindent\textbf{Custom SMC protocols.}
+
+\nc{Need to write this.}
+
 \begin{enumerate}
-\item Secure Computation of MIPS Machine Code https://eprint.iacr.org/2015/547
-\item TASTY https://eprint.iacr.org/2010/365.pdf
-\item ABY cites many papers on first page of intro for using a mix of HE and boolean MPC; but they didnt give much gain because HE is expensive
+
+\item ML prediction: Minionn, Shafi paper, David Wu paper
 

### 2017-10-15 OaklandSubmission/related.tex
more commits to related work section.

@@
 %\end{enumerate}
-\nc{First pass of the related work.}\\
-\nc{ABY have a list of papers that combine boolean and arithmetic compute (hand crafted protocols). We can cite them in the intro or somewhere else to say all those can be automated.}
-
 \noindent\textbf{Automatic Generation of SMC protocols.}
@@
 \\\\
-\noindent\textbf{Custom SMC protocols.}
-
-\nc{Need to write this.}
-
-\begin{enumerate}
-
-\item ML prediction: Minionn, Shafi paper, David Wu paper
-
-\end{enumerate}
+\noindent\textbf{Custom SMC protocols.} Many works have designed specialized, hand-crafted protocols for various secure two-party computation tasks.
+The reason for doing this was to allow for a mix between arithmetic and boolean computations as a secure computation protocol for a pure boolean circuit would have been very inefficient.
+Such hand-crafting of protocols require deep cryptography know-how of designing secure computation protocols with various tools such as homomorphic encryption, garbled circuits, oblivious transfer and so on.
+A sample of such work includes various works on privacy-preserving classification of medical ElectroCardioGram (ECG)~\cite{ecgclassifier}, iris and fingerprint recognition~\cite{irisfingerprint}, remote diagnostics~\cite{remotediagnostics}, private sequence analysis~\cite{sequenceanalysis}, private biometric identification~\cite{biometricidentification}, privacy-preserving matrix factorization and ridge-regression~\cite{matrixfactorization, ridgeregression}, machine-learning classification~\cite{shafindss}, decision trees and forests~\cite{wu}, neural network training and prediction~\cite{secureml,minionn}.
+As seen from our experiments, the protocols generated by \tool give us comparable execution times as many of the above works with minimal programmer overhead.
+\nc{What do we want to say here exactly?}

### 2017-10-15 OaklandSubmission/intro.tex

@@
 We evaluate our system on various benchmarks such as logistic regression and convolutional neural network (CNN) for MNIST data \cite{minionn}, naive bayes, decision trees from \cite{shafindss}.
-Evaluations show that protocols generated by our compiler match the performance of hand-crafted protocols in most cases (see \sectionref{sec:eval}).
-\divya{Below, we will  give an overview of \tool and describe our contributions in more detail.}
+Evaluations show that protocols generated by our compiler match the performance of hand-crafted protocols in most cases (see \sectionref{eval}).
+Below, we will  give an overview of \tool and describe our contributions in more detail.
 
@@
 %Write high level overview of programming language and tool.. \divya{have some preamble}
-\divya{I can add paragraph headings..}
 
@@
 For details on formal language, correctness and security theorems see \sectionref{ld}.
-\nc{Shall we have paragraph heads or something to outline the different contributions? Above feels like one big story.}
 

### 2017-10-15 OaklandSubmission/pipe.tex

@@
 \nc{Started high level description. Can go deeper and correct notation as we proceed.}
-We now describe our technique of ``pipelining'' that allows us to compile and execute large programs that would otherwise have not fitted in memory (due to the large size of the circuit generated by ABY).
+We now describe our technique of ``pipelining'' that allows us to compile and execute large programs that would otherwise have not fit in memory (due to the large size of the circuit generated by ABY).
 We do this securely in a manner such that the underlying backend (such as ABY in this case) does not have to modified in anyway and hence our technique is also compatible with any backend.
@@
 Now, if one were to compile $P_i$ into $P'_i$, and execute $P'_i$, the parties can execute $P'_1||P'_2||\cdots||P'_k$ sequentially to obtain $s_1,\cdots,s_{k-1},$ and finally output $z$.
-However, doing so would enable the parties to learn $s_i$ (for all $1\leq i\leq k-1$), which would completely break the security of the protocol.
+At the $i^{\textrm{\tiny{th}}}$ step, the parties would only need to store information proportional to $x,y,s_i$ and $P'_i$ (which will all be much smaller than $P'$).
+However, if we were to execute the programs as outlined above, it would enable the parties to learn $s_i$ (for all $1\leq i\leq k-1$), which would completely break the security of the protocol.
 
@@
 This is without loss of generality true (we will also for simplicity assume that each $s_i$ is a single element in the ring, but this restriction can also be easily removed).
-Let $r_1,\cdots,r_{k-1}$ be a sequence of random values sampled from the same ring $(\mathbb{Z},+)$ by Party A.
+Let $r_1,\cdots,r_{k-1}$ be a sequence of random values sampled from the same ring $(\mathbb{Z},+)$ by Party A (in our implementation, all $r_i$ values can be generated using a pseudorandom function $F$).
 Let $Q_1$ be the program that takes as input $x,y, r_1$ (and empty state $s_0$), runs $P_1$ (as defined above) to get $s_1$ and then produces output $o_1 = s_1 + r_1$ {\em only to} Party B\footnote{While the description of our protocol here assumes that the underlying backend supports only one party receiving output, this is only a simplifying assumption, and we can easily modify our protocol in the case where both parties must receive the same output.
@@
 We can then appropriately modify the remaining steps as well.}.
-The output of $Q_1$ for Party A is defined to be $r_1$.
+The output of $Q_1$ for Party A in Step 1 is defined to be $r_1$.
 Now, every $Q_i$ ($2\leq i\leq k-1$) takes as input $x,y,r_{i-1},o_{i-1},r_i$, runs $P_i$ on inputs $x,y$ and state $o_{i-1}-r_{i-1}$ to get $s_i$ (where $-$ denotes subtraction in the ring $(\mathbb{Z},+)$) and then produces $s_i+r_i$ as output only to Party B.

### 2017-10-15 OaklandSubmission/intro.tex
started to read the paper draft

@@
 \label{sec:intro}
+\aseem{Is the title too general :). May be we want something more
+  specific on the lines of A compiler for mixed efficient MPC
+  protocols, or something?}
 Today it is hard for developers to create secure implementations using cryptographic techniques.
 Typical developers lack an understanding of cryptography protocols
-and cannot be expected to implement them correctly without tool support.
+and cannot be expected to implement them correctly without tool
+support.\aseem{We might want to bring in efficiency in the first
+  couple of sentences. Else, we are contradicting the previous two
+  sentences below when we say ``He can code up the functionality in a
+  programmer friendly DSL''.}
 In an ideal world, a developer declares the functionality to be implemented
@@
 
-Secure multi-party computation \cite{yao,gmw} (SMC) is a powerful cryptographic tool that allows a set of mutually distrusting parties to compute a joint function of their secret inputs.
-Examples of these include various machine learning prediction algorithm where one party holds the secret input such as a medical report and the other party holds the ML model for disease prediction.
-Since the introduction of SMC in 1980's, there has been a large body of work \cite{..} that has transformed SMC from a mere theoretical tool to something practically efficient and usable.
-Hence, there have been efforts to make SMC accessible to non-specialist programmers.
-To understand the state-of-the-art, let us consider the following toy example.
+\aseem{May be it is in transition still ... but the paragraph below
+  should be the first paragraph of the paper, else it seems strange to
+  introduce SMC after using SMC in the first paragraph? Also, SMC or MPC?}
+Secure multi-party computation \cite{yao,gmw} (SMC) is a powerful
+cryptographic tool that allows a set of mutually distrusting parties
+to compute a joint function of their secret inputs. Examples of these
+include various machine learning prediction algorithm where one party
+holds the secret input such as a medical report and the other party
+holds the ML model for disease prediction. Since the introduction of
+SMC in 1980's, there has been a large body of work \cite{..} that has
+transformed SMC from a mere theoretical tool to something practically
+efficient and usable. Hence, there have been efforts to make SMC
+accessible to non-specialist programmers. To understand the
+state-of-the-art, let us consider the following toy
+example.\aseem{Could we not call it a toy example, and say that this
+  computation is part (sub-computation) of some real ML application or
+  something?}
 
-Suppose a developer wants to compute $w^Tx >0$ securely using SMC. With the current techniques, he has two options.  
+Suppose a developer wants to compute $w^Tx >0$ securely using
+SMC. With the current techniques, he has two options.\aseem{Nice
+  first example.} \aseem{We should
+  use she for the programmer instead of he?}
 \begin{tiret}
-\item He can code up the functionality in a programmer friendly domain-specific language such as \cite{lambdaps,wysteria,oblivm,smcl} that would automatically compile it to an SMC protocol.
-The downside is that all of these works build on a cryptographic back-end that is either entirely boolean \cite{yao,gmw} or entirely arithmetic \cite{homo}.
+\item He can code up the functionality in a programmer friendly
+  domain-specific language such as
+  \cite{lambdaps,wysteria,oblivm,smcl} that would automatically
+  compile it to an SMC protocol. The downside is that all of these
+  works build on a cryptographic back-end that is either entirely
+  boolean \cite{yao,gmw} or entirely arithmetic \cite{homo}\aseem{This
+    homo citation is broken.}.
 The efficiency of the generated SMC protocol is hence, bounded by the possibility of efficiently representing the function in a form that is compatible with the cryptographic back-end used.
@@
 Hence, the second option is to write the functionality in the system developed in \cite{aby}.
-However, using their system requires the programmer to be aware of complex trade-offs between arithmetic and boolean cryptographic schemes and split the functionality into such parts accordingly.
-In their framework, the programmer is required to write circuits consisting of  arithmetic and boolean gates along with appropriate conversion gates.
-In short, the current system of ABY is not suitable to be used by non-specialist programmers.
+However, using their system requires the programmer to be aware of
+complex trade-offs between arithmetic and boolean cryptographic
+schemes and split the functionality into such parts accordingly. In
+their framework, the programmer is required to write circuits
+consisting of  arithmetic and boolean gates along with appropriate
+conversion gates. In short, the current system of ABY is not suitable
+to be used by non-specialist programmers.\aseem{Doesn't their paper
+  itself say that the framework is not programmable and we need
+  a high-level language and a compiler for mixed protocols?}
 
@@
 
-In this work, we develop and implement our framework \tool \nc{I still like EasyPC :)} \divya{I agree that \tool is bad.
-What would be the full form for EasyPC?
-Shouldn't it be EasySC?} for secure computation that achieves generality, performance and programmer productivity.
-The programmer writes a high level \divya{C-like} program (instead of a circuit) to describe the function to be computed.
-Our compiler automatically generates an SMC protocol using a mix of Arithmetic and Boolean compute.
-Our compiler is general and can work with any secure implementation of a mix of arithmetic and boolean computations.
-In our work, we focus on semi-honest secure two-party computation.
-We build on ABY \cite{aby} that provides the suitable cryptographic back-end in this setting.
-We provide a formal type system and prove correctness and security of our compiler.
-We evaluate our system on various benchmarks such as logistic regression and convolutional neural network (CNN) for MNIST data \cite{minionn}, naive bayes, decision trees from \cite{shafindss}.
-Evaluations show that protocols generated by our compiler match the performance of hand-crafted protocols in most cases (see \sectionref{eval}).
-Below, we will  give an overview of \tool and describe our contributions in more detail.
+In this work, we develop and implement our framework \tool \nc{I still
+  like EasyPC :)} \divya{I agree that \tool is bad. What would be the
+  full form for EasyPC? Shouldn't it be EasySC?} for secure
+computation that achieves generality, performance and programmer
+productivity. The programmer writes a high level \divya{C-like}
+program (instead of a circuit) to describe the function to be
+computed. Our compiler automatically generates an SMC protocol using a
+mix of Arithmetic and Boolean compute. Our compiler is general and can
+work with any secure implementation of a mix of arithmetic and boolean
+computations. In our work, we focus on semi-honest secure two-party
+computation. We build on ABY \cite{aby} that provides the suitable
+cryptographic back-end in this setting. We provide a formal type
+system and prove correctness and security of our compiler. We evaluate
+our system on various benchmarks such as logistic regression and
+convolutional neural network (CNN) for MNIST data \cite{minionn},
+naive bayes, decision trees from \cite{shafindss}. Evaluations show
+that protocols generated by our compiler match (\aseem{or better?})
+the performance of hand-crafted protocols in most cases (see
+\sectionref{eval}). Below, we will  give an overview of \tool and
+describe our contributions in more detail.
 
@@
 
-\textbf{Programmability.} \tool is a programmer-friendly framework to express the function to be computed and uses a mix of arithmetic and boolean SMC back-ends.
-We illustrate the above toy example of $w^Tx >b$ in our language in \figureref{ex-sml}.
-As is clear from the code, the programmer is not required to write keywords such as {\em secret} or {\em public} and can remain completely oblivious of the cryptographic back-end.
-In contrast, \figureref{ex-aby} illustrates the same example using the framework in ABY \cite{aby}.
-Here, as already mentioned, the programmer needs to write the low level gates like {\tt MULT, ADD, CONS, GT}, etc.
-The programmer is also required to use the appropriate share conversion gates such as {\tt A2Y} that converts Arithmetic shares to Yao shares.
-Moreover, it is easy to maintain code in \tool and small change to the functionality requires only a small change in the code.
-In contrast, working with ABY, even for a small change a developer might have to re-work the whole circuit for efficiency reasons.
-We describe these examples in detail in \sectionref{ex}.
+\textbf{Programmability.} \tool is a programmer-friendly framework to
+express the function to be computed and uses a mix of arithmetic and
+boolean SMC back-ends. We illustrate the above toy example of $w^Tx
+>b$ in our language in \figureref{ex-sml}.\aseem{Some questions on the
+  example, related to initialization, will discuss in tomorrow's
+  meeting.} As is clear from the code,
+the programmer is not required to write keywords such as {\em secret}
+or {\em public} and can remain completely oblivious of the
+cryptographic back-end. In contrast, \figureref{ex-aby} illustrates
+the same example using the framework in ABY \cite{aby}. Here, as
+already mentioned, the programmer needs to write the low level gates
+like {\tt MULT, ADD, CONS, GT}, etc. The programmer is also required
+to use the appropriate share conversion gates such as {\tt A2Y} that
+converts Arithmetic shares to Yao shares. Moreover, it is easy to
+maintain code in \tool and small change to the functionality requires
+only a small change in the code. In contrast, working with ABY, even
+for a small change a developer might have to re-work the whole circuit
+for efficiency reasons. We describe these examples in detail in
+\sectionref{ex}.
 

### 2017-10-16 OaklandSubmission/intro.tex
minor fix

@@
 
-Suppose a developer wants to compute $w^Tx >0$ securely using
+Suppose a developer wants to compute $w^Tx >b$ securely using
 SMC. With the current techniques, he has two options.\aseem{Nice
@@
 naive bayes, decision trees from \cite{shafindss}. Evaluations show
-that protocols generated by our compiler match (\aseem{or better?})
-the performance of hand-crafted protocols in most cases (see
+that protocols generated by our compiler match
+the performance or outperform hand-crafted protocols in most cases (see
 \sectionref{eval}). Below, we will  give an overview of \tool and

### 2017-10-17 OaklandSubmission/eval.tex
seahorn
minor

@@
 in under a second to C++ code that makes calls to the ABY library~\cite{aby}.
+We use an off-the-shelf solver (SeaHorn~\cite{seahorn}) to check the verification conditions
+required to prove that the array indices remain in bound.
+The solver takes less than a minute on our largest benchmark.
 Since most of our benchmarks are related to machine learning, we set up some notation standard to machine learning and describe our benchmarks next.
@@
 
-\noindent\textbf{Notation.} We use $[N]$ to denote $\{0,1,\dotsc, N-1\}$. \rs{Define $\mathsf{argmax}$.}
+\noindent\textbf{Notation.} We use $[N]$ to denote $\{0,1,\dotsc, N-1\}$. 
+Given a vector $x\in\R^d$, $\mathsf{argmax}\ x = i$ if $x(i) = \mathsf{max}\ \{x_1,\ldots,x_d\}$
 
@@
 %$P(i) \equiv \log p(\ell=i)$, where $p(\ell=i)$ is the likelihood that the output label is $i$ for $i\in[n]$.
-The other matrix $T$ has size $n\times d\times F$ and for each $i\in [n], j\in [d], k\in [F]$, $T(i)(j)(k)$ is
+The other matrix $T$ has size $d\times F\times n$ and for each $j\in [d], k\in [F], i\in [n]$, $T(j)(k)(i)$ is
 %\equiv\log p(x_j|\ell=i)$, i.e., 
@@
 \[
-{\sf argmax}_i P(i)+\sum_{j=0}^{d-1} T(i)(j)(x_j)
+{\sf argmax}\ P+\sum_{j=0}^{d-1} T(j)(x_j)
 \]
@@
 Finally, there are a class of machine learning classifiers that are much more efficient than
-DNNs and provide reasonably good accuracies on standard learning tasks. One such classifier
-is \bonsai~\cite{bonsai} that takes as input $x \in \R^d$ and model consists of a binary tree with $N$ nodes.
+DNNs and provide reasonably good accuracies on standard learning tasks. \bonsai~\cite{bonsai} is a state-of-the art classifier in this class and \tool provides the first 2PC protocol for it.
+\bonsai takes as input $x \in \R^d$ and model consists of a binary tree with $N$ nodes.
 Each node $j$ contains matrices $W_j$ and $V_j$.
@@
 (single instruction multiple data) capabilities that allow \minion to perform
-matrix operations efficiently. Our language and compiler are generic and we have none of these.
-Although we could have used the SIMD capabilities of ABY to improve our performance---our preliminary evaluation suggests that SIMD could lead to a 2X performance improvement---we decided against it to avoid being closely tied to ABY and keeping \tool general, i.e., oblivious to the 2PC backend.
+matrix operations efficiently. Our language and compiler 
+focus on generic 2PC and such optimizations are beyond the scope of this paper.
 However, \minion also reports performance results on a bigger DNN with 7 convolution layers.
@@
 \begin{tabular}{c|c|c|c |c|c|c|c|c|c|c}
-Name  & Prev. time (s) & Prev. Comm (Mb) & LAN time (s) & WAN time (s) & Comm. (Mb)  & \#And & \#Mul & \#Gates & Model Size & LOC\\
+DNN  & Prev. time (s) & Prev. Comm (Mb) & LAN time (s) & WAN time (s) & Comm. (Mb)  & \#And & \#Mul & \#Gates & Model Size & LOC\\
 \hline
@@
 \hline
-Cryptonets &  1.3 & 47.6 & 0.6 & 1.6  & 7    & 2k    & 108k & 316k & 86k & 88\\
+Cryptonets &  1.3 & 47.6 & 0.6 & 1.6  & 70    & 2k    & 108k & 316k & 86k & 88\\
 \hline
@@
 \begin{tabular}{c|c|c|c |c|c|c|c|c|c | c}
-Name       & LAN time (s) & WAN time (s) & Comm. (Mb)  & \#And & \#Mul & \#Gates & Model size & LOC\\
+Classifier       & LAN time (s) & WAN time (s) & Comm. (Mb)  & \#And & \#Mul & \#Gates & Model size & LOC\\
 \hline
@@
 \begin{tabular}{c|c|c|c |c|c|c|c|c|c | c}
-Name       & LAN time (s) & WAN time (s) & Comm. (Mb)  & \#And & \#Mul & \#Gates & depth & LOC\\
+Dataset       & LAN time (s) & WAN time (s) & Comm. (Mb)  & \#And & \#Mul & \#Gates & depth & LOC\\
 \hline
@@
 
- \caption{Pipelining results}
+ \caption{Pipelining results for CIFAR-10. \minion takes 544 seconds and communicates 9272 Mb to evaluate this DNN.}
  \label{tab:cifar} 

### 2017-10-17 OaklandSubmission/ex.tex
intro related

@@
 %$w^Tx>0$, mnist logistic regression
+\divya{add reference to before and later... as mentioned we would run evaluations on secure prediction of different models. Here we describe scenario in detail..}
+
 Consider a  cloud service provider Bob that wants to provide a service to diagnose
@@
 Finally, although this paper uses ABY for evaluation, the  compiler can be easily retargeted to generate code for other cryptography backends.
-\nc{I like the above point. Perhaps we can mention something like this earlier in the intro itself?}
 
@@
 A \tool program cannot go into non-termination or dereference illegal memory, i.e., no buffer overflows or underflows can happen at run time.
-These guarantees are
-much stronger than what are offered by production compilers such as {\tt gcc}.
+Production compilers such as {\tt gcc} provide no such guarantees
+for the code in Figure~\ref{fig:ex-aby}.
 The primary reason that we are able to achieve these guarantees is because

### 2017-10-17 OaklandSubmission/intro.tex

@@
 
-In this work, we develop and implement our framework \tool \nc{I still
-  like EasyPC :)} \divya{I agree that \tool is bad. What would be the
-  full form for EasyPC? Shouldn't it be EasySC?} for secure
+In this work, we develop and implement our framework \tool\footnote{\tool, read as ``easy peasy'', stands for Easy 2 Party Computation} for secure
 computation that achieves generality, performance and programmer

### 2017-10-17 OaklandSubmission/intro.tex

@@
 
-Suppose a developer wants to compute $w^Tx >b$ securely using
+Suppose a developer wants to compute $w^Tx >0$ securely using
 SMC. With the current techniques, he has two options.\aseem{Nice
@@
 naive bayes, decision trees from \cite{shafindss}. Evaluations show
-that protocols generated by our compiler match
-the performance or outperform hand-crafted protocols in most cases (see
+that protocols generated by our compiler match (\aseem{or better?})
+the performance of hand-crafted protocols in most cases (see
 \sectionref{eval}). Below, we will  give an overview of \tool and

### 2017-10-17 OaklandSubmission/ex.tex
polished beginning of intro untill last bullet point

@@
 And Bob does not want to reveal $w$ and $b$ because then Alice can steal the model.
+\divya{Use HIPAA compliance; as it can leak information of patients that was used in training; more compelling argument.
+See Shafi intro.}
 SMC can help Alice and Bob compute $w^Tx>b$ securely, such that Alice receives the classifier's

### 2017-10-17 examples/gradient.cpp
boneh

@@ -1,0 +1,143 @@
+ABYParty *party = new ABYParty(role, address, port, seclvl, bitlen, nthreads, mt_alg, 120000000);
+vector<Sharing*>& sharings = party->GetSharings();
+Circuit* ycirc = sharings[S_YAO]->GetCircuitBuildRoutine();
+Circuit* acirc = sharings[S_ARITH]->GetCircuitBuildRoutine();
+Circuit* bcirc = sharings[S_BOOL]->GetCircuitBuildRoutine();
+auto S1S5 = make_vector<uint32_t>(10, 100);
+auto s_a_S1S5 = make_vector<share*>(10, 100);
+auto S1S6 = make_vector<uint32_t>(10, 100);
+auto s_a_S1S6 = make_vector<share*>(10, 100);
+auto S2S5 = make_vector<uint32_t>(10, 100);
+auto s_a_S2S5 = make_vector<share*>(10, 100);
+auto S2S6 = make_vector<uint32_t>(10, 100);
+auto s_a_S2S6 = make_vector<share*>(10, 100);
+auto S5 = make_vector<uint32_t>(10, 100);
+auto s_a_S5 = make_vector<share*>(10, 100);
+auto S5P = make_vector<uint32_t>(10, 100);
+auto s_a_S5P = make_vector<share*>(10, 100);
+auto S6 = make_vector<uint32_t>(10, 100);
+auto s_a_S6 = make_vector<share*>(10, 100);
+auto S6P = make_vector<uint32_t>(10, 100);
+auto s_a_S6P = make_vector<share*>(10, 100);
+auto S1R = make_vector<uint32_t>(100);
+auto s_a_S1R = make_vector<share*>(100);
+auto S2R = make_vector<uint32_t>(100);
+auto s_a_S2R = make_vector<share*>(100);
+auto R = make_vector<uint32_t>(100);
+auto s_a_R = make_vector<share*>(100);
+auto IP = make_vector<uint32_t>(100);
+auto s_a_IP = make_vector<share*>(100);
+for (uint32_t i = 0; i < 10; i++)
+{
+    for (uint32_t j = 0; j < 100; j++)
+    {
+        if (role == SERVER) {
+            S1S5[i][j] = 50 ;
[…]

### 2017-10-17 OaklandSubmission/intro.tex

--- 2017-10-17 OaklandSubmission/intro.tex
@@
-\section{Introduction}
-\label{sec:intro}
 \aseem{Is the title too general :). May be we want something more
@@
   protocols, or something?}
+
+\section{Introduction}
+\label{sec:intro}
+
 Today it is hard for developers to create secure implementations using cryptographic techniques.
 Typical developers lack an understanding of cryptography protocols
-and cannot be expected to implement them correctly without tool
-support.\aseem{We might want to bring in efficiency in the first
-  couple of sentences. Else, we are contradicting the previous two
-  sentences below when we say ``He can code up the functionality in a
-  programmer friendly DSL''.}
+and cannot be expected to implement them correctly and efficiently without tool
+support.
 In an ideal world, a developer declares the functionality to be implemented
@@
 This paper provides a language-based solution to make
-secure multiparty computation (SMC) accessible to the software developer community.
+secure  computation  accessible to the software developer community.
 
@@
 
-\aseem{May be it is in transition still ... but the paragraph below
-  should be the first paragraph of the paper, else it seems strange to
-  introduce SMC after using SMC in the first paragraph? Also, SMC or MPC?}
-Secure multi-party computation \cite{yao,gmw} (SMC) is a powerful
+Secure multi-party computation (\mpc) \cite{yao,gmw} is a powerful
 cryptographic tool that allows a set of mutually distrusting parties
 to compute a joint function of their secret inputs. Examples of these
-include various machine learning prediction algorithm where one party
-holds the secret input such as a medical report and the other party
-holds the ML model for disease prediction. Since the introduction of
-SMC in 1980's, there has been a large body of work \cite{..} that has
-transformed SMC from a mere theoretical tool to something practically
-efficient and usable. Hence, there have been efforts to make SMC
+include {\em secure prediction} \cite{shafindss,wu,barni,minionn,secureml}. Here,  one party holds the propriety classifier for disease, genomics or spam detection and the other party  
+holds the private input and the second party learns the classification label. Since the introduction of
+\mpc in 1980's, there has been a long sequence of work that has
+transformed \mpc from a mere theoretical tool to something practically
+efficient and usable. Hence, there have been efforts to make \mpc
 accessible to non-specialist programmers. To understand the
-state-of-the-art, let us consider the following toy
-example.\aseem{Could we not call it a toy example, and say that this
-  computation is part (sub-computation) of some real ML application or
-  something?}
+state-of-the-art, let us consider the following 
+example that is the sub-computation in many classification tasks.
 
 Suppose a developer wants to compute $w^Tx >b$ securely using
-SMC. With the current techniques, he has two options.\aseem{Nice
-  first example.} \aseem{We should
+\mpc. With the current techniques, he has three options. \aseem{We should
   use she for the programmer instead of he?}
@@
   works build on a cryptographic back-end that is either entirely
-  boolean \cite{yao,gmw} or entirely arithmetic \cite{homo}\aseem{This
-    homo citation is broken.}.
-The efficiency of the generated SMC protocol is hence, bounded by the possibility of efficiently representing the function in a form that is compatible with the cryptographic back-end used.
+  boolean \cite{yao,gmw} or entirely arithmetic \cite{homo}.
+The efficiency of the generated \mpc protocol is hence, bounded by the possibility of efficiently representing the function in a form that is compatible with the cryptographic back-end used.
 For instance, over $\ell$-bit integers, multiplication requires a boolean circuit of size  $O(\ell^2)$.
-For better efficiency, we would like to evaluate $w^Tx$ using an Arithmetic circuit and the comparison using a boolean circuit.
-In fact, most interesting functions require a mix of arithmetic and boolean computations.
-\item To address this issue, recently Demmler et al.
-\cite{aby} gave a cryptographic protocol for SMC where the parties can mix arithmetic and boolean computations.
-\footnote{Need to note somewhere that many works combined HE and Yao.
-Look at ABY for references.}
+For better efficiency, we would like to evaluate $w^Tx$ using an arithmetic circuit and the comparison using a boolean circuit.
+In fact, most interesting functions require a mix of arithmetic and boolean computations for efficiency.
+
+\item To address this issue of efficient representation, recently Demmler et al. \cite{aby} gave a cryptographic protocol for \mpc where the parties can mix arithmetic and boolean computations.
 Hence, the second option is to write the functionality in the system developed in \cite{aby}.
 However, using their system requires the programmer to be aware of
-complex trade-offs between arithmetic and boolean cryptographic
-schemes and split the functionality into such parts accordingly. In
+trade-offs between arithmetic and boolean cryptographic
+schemes as well as the complex inter-conversion costs, and manually split the functionality into such parts accordingly. In
 their framework, the programmer is required to write circuits
 consisting of  arithmetic and boolean gates along with appropriate
-conversion gates. In short, the current system of ABY is not suitable
-to be used by non-specialist programmers.\aseem{Doesn't their paper
-  itself say that the framework is not programmable and we need
-  a high-level language and a compiler for mixed protocols?}
+conversion gates. In short, as is also mentioned by the authors, the current system of ABY is not suitable
+to be used by non-specialist programmers. \divya{since they say in future they want to automate the splitting.}
 
-\item Now the third option is to hire a cryptographer to design specialised protocols for the functionalities. 
+\item Finally, the third option is to hire a cryptographer to design a specialized efficient protocol for this functionality such as \cite{shafindss,wu,minionn}. 
 \end{tiret}
-In our work, we achieve the best of all options.
+In our work, we achieve the best of all the above options.
 
-%Unfortunately, implementing functionalities using SMC protocols requires thorough understanding of cryptography.
-To allow for widespread use of SMC, it is critical that SMC protocols are programmable by non-cryptographic experts.
+%Unfortunately, implementing functionalities using \mpc protocols requires thorough understanding of cryptography.
+To allow for widespread use of \mpc, it is critical that \mpc protocols are programmable by non-cryptographic experts.
 To cater to this, there have been several efforts developing domain-specific languages that are programmer friendly and compile to a SMC protocol.
@@
 program (instead of a circuit) to describe the function to be
-computed. Our compiler automatically generates an SMC protocol using a
+computed. Our compiler automatically generates an \mpc protocol using a
 mix of Arithmetic and Boolean compute. Our compiler is general and can
@@
 \divya{Say something about what kind of backend we want.
-provides SMC for a mix of arithmetic and boolean circuit with appropriate secure conversion between two types and ABY provides such a framework for semi-honest 2pc.} Our work is compatible with malicious, multiparty as well...
+provides \mpc for a mix of arithmetic and boolean circuit with appropriate secure conversion between two types and ABY provides such a framework for semi-honest 2pc.} Our work is compatible with malicious, multiparty as well...
 etc We give a type system, correctness .....
@@
 express the function to be computed and uses a mix of arithmetic and
-boolean SMC back-ends. We illustrate the above toy example of $w^Tx
->b$ in our language in \figureref{ex-sml}.\aseem{Some questions on the
-  example, related to initialization, will discuss in tomorrow's
-  meeting.} As is clear from the code,
+boolean \mpc back-ends. We illustrate the above toy example of $w^Tx
+>b$ in our language in \figureref{ex-sml}. As is clear from the code,
 the programmer is not required to write keywords such as {\em secret}
@@
 
-\textbf{Pipelining.} Another issue in adoption of SMC is difficulty in scaling to large programs. 
-%In general, SMC implementations struggle in scaling to large programssize of memory on a system limits the largest compute inside the SMC and hence,. 
-The reason being all SMC implementations use a circuit-like representation as an intermediate representation.
+\textbf{Pipelining.} Another issue in adoption of \mpc is difficulty in scaling to large programs. 
+%In general, \mpc implementations struggle in scaling to large programssize of memory on a system limits the largest compute inside the SMC and hence,. 
+The reason being all \mpc implementations use a circuit-like representation as an intermediate representation.
 Hence, the largest compute that can be done securely is upper-bounded by the largest circuit that can fit in memory on a machine.
@@
 If so, we can mention this.}
-To address this issue, we develop a technique called \emph{pipelining} that allows us to run aribitrarily large programs.
+To address this issue, we develop a technique called \emph{pipelining} that allows us to run arbitrarily large programs.
 At a high level, we break the original program into a sequence of small sub-programs and mask the intermediate outputs using secure secret sharing between the two parties.
@@
 
-\textbf{Baseline.} Finally, our framework \tool can serve as a meaningful baseline for future works on hand-crafted SMC protocols for specific functionalities such as \textsc{MiniONN} \cite{minionn}. 
+\textbf{Baseline.} Finally, our framework \tool can serve as a meaningful baseline for future works on hand-crafted \mpc protocols for specific functionalities such as \textsc{MiniONN} \cite{minionn}.
 Prior to our work, garbled circuit based two-party computation is cited as the most practical approach and used as the baseline for comparison of performance.
@@
 encryption with Yao’s garbled circuits protocol, e.g., \cite{barni,blanton,brickell,franz,huang,valeriaMatrix,valeriaRidge,schropferK11}
-  and developed tools that allowed writing a mixed SMC protocol, e.g., \cite{bogdanov,lone,tasty}. The assignment can
+  and developed tools that allowed writing a mixed \mpc protocol, e.g., \cite{bogdanov,lone,tasty}. The assignment can
 even be done automatically as was shown in \cite{autoS}. 

### 2017-10-17 OaklandSubmission/intro.tex
intro contrib

@@
   protocols, or something?}
+
+  \divya{Do we write the intro for \mpc or 2PC?}
 
@@
 
-In this work, we develop and implement our framework \tool\footnote{\tool, read as ``easy peasy'', stands for Easy 2 Party Computation} for secure
-computation that achieves generality, performance and programmer
-productivity. The programmer writes a high level \divya{C-like}
-program (instead of a circuit) to describe the function to be
-computed. Our compiler automatically generates an \mpc protocol using a
-mix of Arithmetic and Boolean compute. Our compiler is general and can
-work with any secure implementation of a mix of arithmetic and boolean
-computations. In our work, we focus on semi-honest secure two-party
-computation. We build on ABY \cite{aby} that provides the suitable
-cryptographic back-end in this setting. We provide a formal type
-system and prove correctness and security of our compiler. We evaluate
-our system on various benchmarks such as logistic regression and
-convolutional neural network (CNN) for MNIST data \cite{minionn},
-naive bayes, decision trees from \cite{shafindss}. Evaluations show
-that protocols generated by our compiler match
-the performance or outperform hand-crafted protocols in most cases (see
-\sectionref{eval}). Below, we will  give an overview of \tool and
-describe our contributions in more detail.
+%\divya{old para...} In this work, we develop and implement our framework \tool\footnote{\tool, read as ``easy peasy'', stands for Easy 2 Party Computation} for secure
+%computation that achieves generality, performance and programmer
+%productivity. The programmer writes a high level \divya{C-like}
+%program (instead of a circuit) to describe the function to be
+%computed. Our compiler automatically generates an \mpc protocol using a
+%mix of Arithmetic and Boolean compute. Our compiler is general and can
+%work with any secure implementation of a mix of arithmetic and boolean
+%computations. In our work, we focus on semi-honest secure two-party
+%computation. We build on ABY \cite{aby} that provides the suitable
+%cryptographic back-end in this setting. We provide a formal type
+%system and prove correctness and security of our compiler. We evaluate
+%our system on various benchmarks such as logistic regression and
+%convolutional neural network (CNN) for MNIST data \cite{minionn},
+%naive bayes, decision trees from \cite{shafindss}. Evaluations show
+%that protocols generated by our compiler match
+%the performance or outperform hand-crafted protocols in most cases (see
+%\sectionref{eval}). Below, we will  give an overview of \tool and
+%describe our contributions in more detail.
 
@@
 %Write high level overview of programming language and tool.. \divya{have some preamble}
+
+In this work, we develop and implement our framework \tool\footnote{\tool, read as ``easy peasy'', stands for Easy 2 Party Computation} for secure
+computation. We provide the first compiler where the developer writes a high-level program and the compiler automatically generates an efficient \mpc protocol that uses a
+mix of Arithmetic and Boolean compute. 
+It is compatible with any implementation of \mpc that allows a mix of arithmetic and boolean computations. 
+We focus on semi-honest secure two-party computation and use ABY \cite{aby} as the suitable cryptographic back-end.
+We provide a formal type system and operational semantics for our language and prove correctness and security of our compiler.
+We evaluate our system on wide range of benchmarks \cite{shafindss,wu,minionn,valeriaMatrix} computing linear classifier, decision trees, deep neural networks, and matrix factorization.
+Our evaluations show that protocols generated by our compiler match
+the performance or outperform hand-crafted protocols in most cases (see
+\sectionref{eval}). \divya{Moreover, unlike previous systems, our system can scale to arbitrarily large computations using {\em pipelining} (see below). What do we say about Yao pipelining?}
+Hence, we show that our framework achieves all three properties of generality, performance and programmer productivity.
+
+\divya{Points.. 1. Prior to our work, there was CBMGC no perf; ABY no prog; 
+3. Programmability allows writing programs generating large circuits... 
+}
+
+Below, we describe the salient features of \tool in more detail.
 

### 2017-10-20 OaklandSubmission/ld.tex
adding formal definitions

\newcommand{\kw}[1]{{\lstinline[basicstyle=\small\color{magenta}]{#1}}}
\newcommand{\ftext}[1]{\text{\small{#1}}}
\newcommand{\cond}[3]{\ensuremath{{{#1}\:?\:{#2}\::{#3}}}}
\newcommand{\for}[4]{\ensuremath{\kw{for}\:{#1}\:\kw{in}\:[{#2}, {#3}]\:\kw{do}\:{#4}}}
\newcommand{\ite}[3]{\ensuremath{\kw{if}({#1}, {#2}, {#3})}}
\newcommand{\loops}[3]{\ensuremath{\kw{while}\:{#1} \leq {#2}\:\kw{do}\:{#3}}}

\section{Formal development}
\label{sec:ld}

In this section we prove the correctness and security of our compiler.

\begin{figure}
  \[
  \begin{array}{rrcl}
    \ftext{Base type} & \sigma &::=& \kw{uint} \mid \kw{bool}\\
    \ftext{Type} & \psi &::=& \sigma \mid \sigma[n]\\
    \ftext{Constant} & c &::=& n \mid \top \mid \bot\\
    \ftext{Expression} & e &::=& c \mid x \mid e_{1} + e_{2} \mid e_{1} > e_{2} \mid \cond{e_{1}}{e_{2}}{e_{3}}\\
    & &\mid& x[e] \mid [\overline{e_{i}}]_{n} \mid \kw{in}_{j}\\
    \ftext{Statement} & s &::=& \psi\:x = e \mid x := e \mid \for{x}{n_{1}}{n_{2}}{s}\\
    & & \mid& x[e_{1}] := e_{2} \mid \ite{e}{s_{1}}{s_{2}} \mid \kw{out}\:e \mid s_{1}; s_{2}\\
    & & \mid& \loops{x}{n}{s}
  \end{array}
  \]
\label{fig:srclang}
\caption{Source language}
\end{figure}

\begin{figure}
  \fbox{$\rho \vdash e \Downarrow v$}
  \[
  \\
  \begin{array}{c}
    \inferrule*[lab={\footnotesize{E-Var}}]
               {
               }
               {
                 \rho \vdash x \Downarrow \rho(x)
               }

               \hspace{0.1cm}

    \inferrule*[lab={\footnotesize{E-Add}}]
               {
                 \forall i \in \{1, 2\}.\:\rho \vdash e_{i} \Downarrow n_{i}
               }
               {
                 \rho \vdash e_{1} + e_{2} \Downarrow n_{1} + n_{2}
               }

               \hspace{0.1cm}

    \inferrule*[lab={\footnotesize{E-Read}}]
               {
                 \rho \vdash x \Downarrow [\overline{c_{i}}]_{n_{1}} \\\\
                 \rho \vdash e \Downarrow n \quad n < n_{1}
               }
               {
                 \rho \vdash x[e] \Downarrow c_{n}
               }
\\\\
    \inferrule*[lab={\footnotesize{E-Arr}}]
               {
                 \forall i \in \{0 \dots n - 1\}.\:\rho \vdash e_{i} \Downarrow c_{i}
               }
               {
                 \rho \vdash [\overline{e_{i}}]_{n} \Downarrow [\overline{c_{i}}]_{n}
               }
               \quad
    \inferrule*[lab={\footnotesize{E-Inp}}]
               {
               }
               {
                 \rho \vdash \kw{in}_{j} \Downarrow c
               }
  \end{array}
  \]
  \\\\
    \fbox{$\rho \vdash c \Downarrow \rho'; O$}
  \[
  \\
  \begin{array}{c}
    \inferrule*[lab={\footnotesize{E-Decl}}]
               {
                 \rho \vdash e \Downarrow v
               }
               {
                 \rho \vdash \psi\:x = e \Downarrow \rho, x \mapsto v; \cdot
               }

               \hspace{0.1cm}

    \inferrule*[lab={\footnotesize{E-LoopT}}]
               {
                 \rho(x) > n
               }
               {
                 \rho \vdash \loops{x}{n}{s} \Downarrow \rho; \cdot
               }

               \\\\

    \inferrule*[lab={\footnotesize{E-LoopI}}]
               {
                 \rho(x) \leq n\\\\
                 \rho \vdash s \Downarrow \rho_{1}; O_{1}\\\\
                 \rho_{2} = [\rho_{1}]_{\mathsf{dom}(\rho)}[x \mapsto \rho_{1}(x) + 1]\\\\
                 \rho_{2} \vdash \loops{x}{n}{s} \Downarrow \rho'; O_{2}
               }
               {
                 \rho \vdash \loops{x}{n}{s} \Downarrow \rho'; O_{1}, O_{2}
               }

               \hspace{0.2cm}

    \inferrule*[lab={\footnotesize{E-If}}]
               {
                 \rho \vdash e \Downarrow c\\\\
                 c = \top \Rightarrow s = s_{1}\\\\
                 c = \bot \Rightarrow s = s_{2}\\\\
                 \rho \vdash s \Downarrow \rho'; O
               }
               {
                 \rho \vdash \ite{e}{s_{1}}{s_{2}} \Downarrow \rho'; O
               }

               \\\\

    \inferrule*[lab={\footnotesize{E-For}}]
               {
                 \rho, x \mapsto n_{1} \vdash \loops{x}{n_{2}}{s} \Downarrow \rho_{1}; O
               }
               {
                 \rho \vdash \for{x}{n_{1}}{n_{2}}{s} \Downarrow \rho_{1} - \{x\}; O
               }

               \hspace{0.2cm}

    \inferrule*[lab={\footnotesize{E-Out}}]
               {
                 \rho \vdash e \Downarrow c
               }
               {
                 \rho \vdash \kw{out}\:e \Downarrow \rho; c
               }

\end{array}
  \]
\label{fig:srcsem}
\caption{Source semantics (selected rules)}
\end{figure}

\newcommand{\lcond}[4]{\ensuremath{{{#2}\:?_{{#1}}\:{#3}\::{#4}}}}
%\newcommand{\for}[4]{\ensuremath{\kw{for}\:{#1}\:\kw{in}\:[{#2}, {#3}]\:\kw{do}\:{#4}}}
%\newcommand{\ite}[3]{\ensuremath{\kw{if}({#1}, {#2}, {#3})}}
%\newcommand{\loops}[3]{\ensuremath{\kw{while}\:{#1} \leq {#2}\:\kw{do}\:{#3}}}

\begin{figure}
  \[
  \begin{array}{rrcl}
    \ftext{Secret label} & m &::=& \mathcal{A} \mid \mathcal{B}\\
    \ftext{Label} & \ell &::=& \mathcal{P} \mid m\\
    \ftext{Type} & \tau &::=& \sigma^{\ell} \mid \sigma^{\ell}[n]\\
    \ftext{Expression} & \widetilde{e} &::=& c \mid x \mid \widetilde{e_{1}} +_{\ell} \widetilde{e_{2}} \mid \widetilde{e_{1}} >_{\ell} \widetilde{e_{2}} \\
    & & \mid & \lcond{\ell}{{\widetilde{e}}}{{\widetilde{e_{1}}}}{{\widetilde{e_{2}}}} \mid x[\widetilde{e}] \mid [\overline{\widetilde{e_{i}}}]_{n} \mid \kw{in}^{m}_{j} \mid \widetilde{e} \rhd m\\
    \ftext{Statement} & \widetilde{s} &::=& \tau\:x = \widetilde{e} \mid x := \widetilde{e} \mid \dots \mid \widetilde{s_{1}}; \widetilde{s_{2}} \mid \dots\\
  \end{array}
  \]
\label{fig:tgtlang}
\caption{Target language}
\end{figure}

\begin{figure}
  \[
  \begin{array}{rrcl}
    \ftext{Wire id} & r &&\\
    \ftext{Circuit gate} & \kappa & ::= & r \mid \kw{in}^{m}_{j} \mid \kw{add}\:\kappa_{1}\:\kappa_{2} \mid \kw{gt}\:\kappa_{1}\:\kappa_{2}\\
    & & \mid & \kw{mux}\:\kappa\:\kappa_{1}\:\kappa_{2} \mid \widetilde{w} \rhd m\\
    \ftext{Base value} & \widetilde{w} & ::= & c \mid \kappa\\
    \ftext{Value} & \widetilde{v} & ::= & \widetilde{w} \mid [\overline{\widetilde{w}_{i}}]_{n}\\
    \ftext{Circuit} & \chi & ::= & \cdot \mid \kw{bind}\:\kappa\:r \mid \kw{out}\:\kappa \mid \kappa_{1}; \kappa_{2}
  \end{array}
  \]
\label{fig:tgtruntime}
\caption{Target runtime}
\end{figure}

\begin{figure}
  \fbox{$\widetilde{\rho} \vdash \widetilde{e} \Downarrow \widetilde{v}$}
  \[
  \\
  \begin{array}{c}
    \inferrule*[lab={\footnotesize{S-Var}}]
               {
               }
               {
                 \widetilde{\rho} \vdash x \Downarrow \widetilde{\rho}(x)
               }

    \inferrule*[lab={\footnotesize{S-PAdd}}]
               {
                 \forall i \in \{1, 2\}.\:\widetilde{\rho} \vdash \widetilde{e_{i}} \Downarrow n_{i}
               }
               {
                 \widetilde{\rho} \vdash \widetilde{e_{1}} +_{\mathcal{P}} \widetilde{e_{2}} \Downarrow n_{1} + n_{2}
               }

    \inferrule*[lab={\footnotesize{S-Read}}]
               {
                 \widetilde{\rho} \vdash x \Downarrow [\overline{\widetilde{w_{i}}}]_{n_{1}} \\\\
                 \widetilde{\rho} \vdash \widetilde{e} \Downarrow n \quad n < n_{1}
               }
               {
                 \widetilde{\rho} \vdash x[\widetilde{e}] \Downarrow \widetilde{w_{n}}
               }\\\\
    \inferrule*[lab={\footnotesize{S-SAdd}}]
               {
                 \forall i \in \{1, 2\}.\:\widetilde{\rho} \vdash \widetilde{e_{i}} \Downarrow \kappa_{i}
               }
               {
                 \widetilde{\rho} \vdash \widetilde{e_{1}} +_{\mathcal{A}} \widetilde{e_{2}} \Downarrow \kw{add}\:\kappa_{1}\:\kappa_{2}
               }

               \hspace{0.3cm}

    \inferrule*[lab={\footnotesize{S-SGt}}]
               {
                 \forall i \in \{1, 2\}.\:\widetilde{\rho} \vdash \widetilde{e_{i}} \Downarrow \kappa_{i}
               }
               {
                 \widetilde{\rho} \vdash \widetilde{e_{1}} >_{\mathcal{B}} \widetilde{e_{2}} \Downarrow \kw{gt}\:\kappa_{1}\:\kappa_{2}
               }\\\\
               \inferrule*[lab={\footnotesize{S-SCond}}]
               {
                 \forall i \in \{1, 2, 3\}.\:\widetilde{\rho} \vdash \widetilde{e_{i}} \Downarrow \kappa_{i}
               }
               {
                 \widetilde{\rho} \vdash \lcond{\mathcal{B}}{\widetilde{e_{1}}}{\widetilde{e_{2}}}{\widetilde{e_{3}}} \Downarrow \kw{mux}\:\kappa_{1}\:\kappa_{2}\:\kappa_{3}
               }
               \hspace{0.2cm}
    \inferrule*[lab={\footnotesize{S-Coerce}}]
               {
                 \widetilde{\rho} \vdash \widetilde{e} \Downarrow \widetilde{w}
               }
               {
                 \widetilde{\rho} \vdash \widetilde{e} \rhd m \Downarrow \widetilde{w} \rhd m
               }
               \\\\
               %% \inferrule*[lab={\footnotesize{S-Arr}}]
               %% {
               %%   \forall i \in \{0 \dots n - 1\}.\:\widetilde{\rho} \vdash \widetilde{e_{i}} \Downarrow \widetilde{w_{i}}
               %% }
               %% {
               %%   \widetilde{\rho} \vdash [\overline{\widetilde{e_{i}}}]_{n} \Downarrow [\overline{\widetilde{w_{i}}}]_{n}
               %% }

               \inferrule*[lab={\footnotesize{S-PCond}}]
               {
                 \widetilde{\rho} \vdash \widetilde{e} \Downarrow c\\\\
                 c = \top \Rightarrow \widetilde{e'} = \widetilde{e_{1}}\\
                 c = \bot \Rightarrow \widetilde{e'} = \widetilde{e_{2}}\\\\
                 \widetilde{\rho} \vdash \widetilde{e'} \Downarrow \widetilde{v}
               }
               {
                 \widetilde{\rho} \vdash \lcond{\mathcal{P}}{\widetilde{e}}{\widetilde{e_{1}}}{\widetilde{e_{2}}} \Downarrow \widetilde{v}
               }
               %\hspace{0.5cm}

    \inferrule*[lab={\footnotesize{S-Inp}}]
               {
               }
               {
                 \widetilde{\rho} \vdash \kw{in}^{m}_{j} \Downarrow \kw{in}^{m}_{j}
               }
  \end{array}
  \]
  \\\\
    \fbox{$\widetilde{\rho} \vdash \widetilde{c} \Downarrow \widetilde{\rho'}; \chi$}
  \[
  \\
  \begin{array}{c}
    \inferrule*[lab={\footnotesize{S-DeclC}}]
               {
                 \widetilde{\rho} \vdash \widetilde{e} \Downarrow \widetilde{v}\\\\
                 \widetilde{v} = c \vee \widetilde{v} = [\overline{c_{i}}]_{n}\\\\
                 \widetilde{\rho'} = \widetilde{\rho}, x \mapsto \widetilde{v}
               }
               {
                 \widetilde{\rho} \vdash \tau\:x = \widetilde{e} \Downarrow \widetilde{\rho'}; \cdot
               }

               \hspace{0.3cm}

    \inferrule*[lab={\footnotesize{S-DeclCkt}}]
               {
                 \widetilde{\rho} \vdash \widetilde{e} \Downarrow \kappa\\\\
                 \mathsf{fresh}\:r\\\\
                 \widetilde{\rho'} = \widetilde{\rho}, x \mapsto r\quad
                 \chi = \kw{bind}\:\kappa\:r
               }
               {
                 \widetilde{\rho} \vdash \tau\:x = \widetilde{e} \Downarrow \widetilde{\rho'}; \chi
               }

\\\\
    \inferrule*[lab={\footnotesize{S-DeclCktA}}]
               {
                 \widetilde{\rho} \vdash \widetilde{e} \Downarrow [\overline{\kappa_{i}}]_{n}\\\\
                 \forall i \in \{0 \dots n - 1\}.\:\mathsf{fresh}\:r_{i}\\\\
                 \widetilde{\rho'} = \widetilde{\rho}, x \mapsto [r_{i}]_{n}\quad
                 \chi = \overline{\kw{bind}\:\kappa_{i}\:r_{i}}
               }
               {
                 \widetilde{\rho} \vdash \tau\:x = \widetilde{e} \Downarrow \widetilde{\rho'}; \chi
               }

               \hspace{0.2cm}

    \inferrule*[lab={\footnotesize{S-Out}}]
               {
                 \widetilde{\rho} \vdash \widetilde{e} \Downarrow \kappa
               }
               {
                 \widetilde{\rho} \vdash \kw{out}\:\widetilde{e} \Downarrow \widetilde{\rho}; \kw{out}\:\kappa
               }

               \\\\
    \inferrule*[lab={\footnotesize{S-If}}]
               {
                 \widetilde{\rho} \vdash \widetilde{e} \Downarrow c\\\\
                 c = \top \Rightarrow \widetilde{s} = \widetilde{s_{1}}\\\\
                 c = \bot \Rightarrow \widetilde{s} = \widetilde{s_{2}}\\\\
                 \widetilde{\rho} \vdash \widetilde{s} \Downarrow \widetilde{\rho'}; \chi
               }
               {
                 \widetilde{\rho} \vdash \ite{\widetilde{e}}{\widetilde{s_{1}}}{\widetilde{s_{2}}} \Downarrow \widetilde{\rho'}; \chi
               }

               \hspace{0.3cm}

    \inferrule*[lab={\footnotesize{S-Write}}]
               {
                 \widetilde{\rho} \vdash x \Downarrow [\overline{\widetilde{w_{i}}}]_{n}\\\\
                 \widetilde{\rho} \vdash \widetilde{e_{1}} \Downarrow n_{1}\quad
                 n_{1} < n\\\\
                 \widetilde{\rho} \vdash \widetilde{e_{2}} \Downarrow \widetilde{w}\\\\
                 \widetilde{\rho'} = \widetilde{\rho}[x \mapsto [\overline{\widetilde{w_{i}}}]_{n}[n_{1} \mapsto \widetilde{w}]]
               }
               {
                 \widetilde{\rho} \vdash x[\widetilde{e_{1}}] := \widetilde{e_{2}} \Downarrow \widetilde{\rho'}; \cdot
               }

\end{array}
  \]
\label{fig:tgtsem}
\caption{Target semantics (selected rules)}
\end{figure}

\begin{figure}
  \fbox{$\widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \kappa \Downarrow b_{1}, b_{2}$}
  \[
  \\
  \begin{array}{c}
    %% \inferrule*[lab={\footnotesize{C-Wire}}]
    %%            {
    %%              b_{1}, b_{2} = \widehat{\rho_{1}}(r), \widehat{\rho_{2}}(r)
    %%            }
    %%            {
    %%              \widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash r \Downarrow b_{1}, b_{2}
    %%            }               
    \inferrule*[lab={\footnotesize{C-In}}]
               {
                 b_{1}, b_{2} = \mathcal{E}_{m}(c)
               }
               {
                 \widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \kw{in}^{m}_{j} \Downarrow b_{1}, b_{2}
               }

               \hspace{0.3cm}

    \inferrule*[lab={\footnotesize{C-Coerce}}]
               {
                 \widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \kappa \Downarrow b_{1}, b_{2}\\\\
                 c = \mathcal{D}_{m_{1}}(b_{1}, b_{2})\\\\
                 b'_{1}, b'_{2} = \mathcal{E}_{m}(c)
               }
               {
                 \widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \kappa \rhd m \Downarrow b'_{1}, b'_{2}
               }
\\\\

    \inferrule*[lab={\footnotesize{C-Add}}]
               {
                 \forall i \in \{1, 2\}.\:\widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \kappa_{i} \Downarrow b_{1i}, b_{2i}\quad
                 n_{i} = \mathcal{D}_{\mathcal{A}}(b_{1i}, b_{2i})\\\\
                 b_{1}, b_{2} = \mathcal{E}_{\mathcal{A}}(n_{1} + n_{2})
               }
               {
                 \widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \kw{add}\:\kappa_{1}\:\kappa_{2} \Downarrow b_{1}, b_{2}
               }

\\\\

    \inferrule*[lab={\footnotesize{C-Gt}}]
               {
                 \forall i \in \{1, 2\}.\:\widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \kappa_{i} \Downarrow b_{1i}, b_{2i}\quad
                 n_{i} = \mathcal{D}_{\mathcal{B}}(b_{1i}, b_{2i})\\\\
                 b_{1}, b_{2} = \mathcal{E}_{\mathcal{B}}(n_{1} > n_{2})
               }
               {
                 \widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \kw{gt}\:\kappa_{1}\:\kappa_{2} \Downarrow b_{1}, b_{2}
               }

\\\\

    \inferrule*[lab={\footnotesize{C-Mux}}]
               {
                 \forall i \in \{1, 2, 3\}.\:\widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \kappa_{i} \Downarrow b_{1i}, b_{2i}\quad
                 c_{i} = \mathcal{D}_{\mathcal{B}}(b_{1i}, b_{2i})\\\\
                 c_{1} = \top \Rightarrow b_{1}, b_{2} = \mathcal{E}_{\mathcal{B}}(c_{1})\quad
                 c_{1} = \bot \Rightarrow b_{1}, b_{2} = \mathcal{E}_{\mathcal{B}}(c_{2})
               }
               {
                 \widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \kw{mux}\:\kappa_{1}\:\kappa_{2}\:\kappa_{3} \Downarrow b_{1}, b_{2}
               }

  \end{array}
  \]
  \\\\
    \fbox{$\widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \chi \Downarrow \widehat{\rho'_{1}}, \widehat{\rho'_{2}}; O$}
  \[
  \\
  \begin{array}{c}
    \inferrule*[lab={\footnotesize{C-Bind}}]
               {
                 \widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \kappa \Downarrow b_{1}, b_{2}\\\\
                 \widehat{\rho'_{1}} = \widehat{\rho_{1}}[r \mapsto b_{1}] \quad
                 \widehat{\rho'_{2}} = \widehat{\rho_{2}}[r \mapsto b_{2}]
               }
               {
                 \widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \kw{bind}\:\kappa\:r \Downarrow \widehat{\rho'_{1}}, \widehat{\rho'_{2}}; \cdot
               }

               \hspace{0.1cm}

    \inferrule*[lab={\footnotesize{C-Out}}]
               {
                 \widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \kappa \Downarrow b_{1}, b_{2}\\\\
                 c = \mathcal{D}_{m}(b_{1}, b_{2})
               }
               {
                 \widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \kw{out}\:\kappa \Downarrow \widehat{\rho_{1}}, \widehat{\rho_{2}}; c
               }
\end{array}
  \]
\label{fig:cktsem}
\caption{Circuit semantics (selected rules)}
\end{figure}

  %% \\
  %% \[
  %% \begin{array}{rrcl}
  %%   \ftext{Value} & v &::=& c \mid [\overline{c_{i}}]_{n}\\
  %%   \ftext{Observation} & O &::=& \cdot \mid c; O\\
  %% \end{array}
  %% \]

### 2017-10-22 OaklandSubmission/ld.tex
formatting
cp
fixing a typo in one of the rules
compilation judgment
fixing nit in a rule
nit
text on source language and semantics

@@
-\newcommand{\kw}[1]{{\lstinline[basicstyle=\small\color{magenta}]{#1}}}
+\newcommand{\kw}[1]{{\lstinline[basicstyle=\small\color{blue}]{#1}}}
 \newcommand{\ftext}[1]{\text{\small{#1}}}
@@
 In this section we prove the correctness and security of our compiler.
+%
+We first formalize the source and target languages. Our source runtime
+semantics is a model of the ideal, trusted third-party semantics, and
+generates observations corresponding to the values revealed to the
+parties.
+%
+The target language semantics (a model of the C++ code generated by
+our compiler implementation) ``computes away'' the public parts of the
+compiled program, generating a secure computation circuit.
+%
+Finally, we formalize the circuit semantics that computes the
+generated circuit, and like the source semantics, emits observations.
+
+We then present the compilation judgments. To prove the correctness
+of our compiler, we prove that it preserves the observations. For
+security of our compiler, we reduce the security argument to the
+security of the cryptographic protocol used to compute the secure
+computation circuit.
+
+We present only selected parts of our formalization for space
+reasons. Full definitions and proofs can be found in the supplementary
+material submitted along with the paper.
 
 \begin{figure}
+  \small
   \[
@@
     \ftext{Expression} & e &::=& c \mid x \mid e_{1} + e_{2} \mid e_{1} > e_{2} \mid \cond{e_{1}}{e_{2}}{e_{3}}\\
-    & &\mid& x[e] \mid [\overline{e_{i}}]_{n} \mid \kw{in}_{j}\\
+    & &\mid& [\overline{e_{i}}]_{n} \mid x[e] \mid \kw{in}_{j}\\
     \ftext{Statement} & s &::=& \psi\:x = e \mid x := e \mid \for{x}{n_{1}}{n_{2}}{s}\\
@@
   \]
+\caption{Source language}
 \label{fig:srclang}
-\caption{Source language}
 \end{figure}
@@
 \begin{figure}
+  \small
   \fbox{$\rho \vdash e \Downarrow v$}
@@
   \]
+\caption{Source semantics (selected rules)}
 \label{fig:srcsem}
-\caption{Source semantics (selected rules)}
 \end{figure}
+
+\subsubsection*{Source language}Our source language
+(Figure~\ref{fig:srclang}) is a simple imperative language. Types
+$\psi$ in the language consist of the base types $\sigma$, and arrays
+of base types $\sigma[n]$, where $n$ is the array length. Though we
+model only one dimensional arrays, our implementation supports higher
+dimensional arrays as well. Expressions in the language include the
+integer constants $n$, \kw{bool} constants $\top$ and $\bot$,
+variables $x$, binary operations $e_{1} + e_{2}$ and $e_{1} > e_{2}$,
+conditionals $\cond{e}{e_{1}}{e_{2}}$, array literals 
+$[\overline{e_{i}}]_{n}$\footnote{We write $\overline{e}$ (and
+  similarly for other symbols) to denote a sequence of expressions.
+The length of the sequence is usually clear from the context.}, and
+array reads $x[e]$. The expression $\kw{in}_{j}$ denotes input from
+party $j$. The statements $s$ in the language comprise of variable
+declarations and assignments ($\psi\:x = e$ and $x := e$ resp.),
+\kw{for} loops, array writes ($x[e_{1}] := e_{2}$), \kw{if}
+statements, and sequence of statements ($s_{1}; s_{2}$). The statement
+$\kw{out}\:e$ denotes revealing the value of $e$ to the
+parties. The \kw{while} statement is an internal syntax that is not
+exposed to the programmer.
+
+The runtime semantics for the source language is shown in
+Figure~\ref{fig:srcsem}. Values $v$, runtime environments $\rho$, and
+observations $O$ are defined as follows:
+
+\vspace{0.2cm}
+$
+\small
+\begin{array}{rrcl}
+    \ftext{Value} & v &::=& c \mid [\overline{c_{i}}]_{n}\\
+    \ftext{Runtime environment} & \rho &::=& \cdot \mid \rho[x \mapsto v]\\
+    \ftext{Observation} & O & ::= & \cdot \mid c, O \\
+\end{array}
+$
+
+\vspace{0.2cm}
+Values consist of constants and array values. Runtime environment
+$\rho$ maps variables to values. Observations are sequences of
+constants.
+
+The judgment $\rho \vdash e \Downarrow v$ denotes the big-step
+evaluation of an expression~$e$ to a value~$v$ under the runtime
+environment~$\rho$. Rule ({\sc{E-Var}}) looks up the value of $x$ in
+the environment. Rule ({\sc{E-Add}}) inductively evaluates $e_{1}$ and
+$e_{2}$ and evaluates to their addition. Rule ({\sc{E-Read}})
+evaluates an array read operation. It first evaluates $x$ to an array
+value $[\overline{c_{i}}]_{n_{1}}$, and $e$ to an \kw{uint} value
+$n$. It then returns $c_{n}$, the $n$-th index value in the array,
+provided $n < n_{1}$, the length of the array. Rule ({\sc{E-Inp}})
+evaluates to some constant $c$ denoting party $j$'s input. We model
+the inputs to be base constants, an array input can be written in the
+language as $[\kw{in}_{j}]_{n}$, which can then evaluate using the
+rule ({\sc{E-Arr}}). The remaining rules are straightforward, and are
+elided for space reasons.
+
+The judgment $\rho \vdash s \Downarrow \rho'; O$ represents the
+big-step evaluation of an statement $s$ under environment $\rho$
+producing a new environment $\rho'$ and observations $O$. Rule
+({\sc{E-Decl}}) evaluates the expression $e$ to $v$, and returns the
+updated environment $\rho[x \mapsto v]$, with empty observations. Rule
+({\sc{E-If}}) evaluates the guard expression, and then evaluates
+either $s_{1}$ or $s_{2}$ accordingly. $\kw{for}$ statements evaluate
+through the internal $\kw{while}$ syntax. Specifically, rule
+({\sc{E-For}}) updates $\rho$ with the initial counter value $n_{1}$,
+evaluates $\loops{x}{n_{2}}{s}$ to $\rho_{1}; O$, and returns
+$\rho_{1} - \{x\}$ (removing $x$ from the environment) and $O$. Rule
+({\sc{E-LoopI}}) shows the inductive case for \kw{while}
+statements, when $\rho(x) \leq n$. The rule evaluates $s$, producing
+$\rho_{1}; O_{1}$. It then restricts $\rho_{1}$ to the domain of
+$\rho$ ($[\rho_{1}]_{\mathsf{dom}(\rho)}$) to remove the variables
+added by $s$, increments the value if $x$, and evaluates the
+\kw{while} statement under this updated environment. Rule
+({\sc{E-LoopT}}) is the termination case for \kw{while} statements,
+when $\rho(x) > n$. Finally, the rule ({\sc{E-Out}}) evaluates the
+expression, and adds it value to the observations.
+
+\subsubsection{Target language}
+
+\begin{theorem}[Correctness of the compiler]
+
+\end{theorem}
 
@@
 \begin{figure}
+  \small
   \[
@@
 \begin{figure}
+  \small
   \fbox{$\widetilde{\rho} \vdash \widetilde{e} \Downarrow \widetilde{v}$}
@@
 
-               \hspace{0.3cm}
+               \hspace{0.5cm}
 
@@
                {
-                 \widetilde{\rho} \vdash \widetilde{e} \Downarrow \kappa\\\\
+                 \widetilde{\rho} \vdash \widetilde{e} \Downarrow \kappa\quad
                  \mathsf{fresh}\:r\\\\
@@
 
-               \hspace{0.3cm}
+               \hspace{0.5cm}
 
-    \inferrule*[lab={\footnotesize{S-Write}}]
+    \inferrule*[lab={\footnotesize{S-WriteCkt}}]
                {
-                 \widetilde{\rho} \vdash x \Downarrow [\overline{\widetilde{w_{i}}}]_{n}\\\\
-                 \widetilde{\rho} \vdash \widetilde{e_{1}} \Downarrow n_{1}\quad
-                 n_{1} < n\\\\
-                 \widetilde{\rho} \vdash \widetilde{e_{2}} \Downarrow \widetilde{w}\\\\
-                 \widetilde{\rho'} = \widetilde{\rho}[x \mapsto [\overline{\widetilde{w_{i}}}]_{n}[n_{1} \mapsto \widetilde{w}]]
+                 \widetilde{\rho} \vdash x \Downarrow [\overline{r_{i}}]_{n}\quad
+                 \widetilde{\rho} \vdash \widetilde{e_{1}} \Downarrow n_{1}\\\\
+                 n_{1} < n\quad
+                 \mathsf{fresh}\:r\quad
+                 \widetilde{\rho} \vdash \widetilde{e_{2}} \Downarrow \kappa\\\\
+                 \widetilde{\rho'} = \widetilde{\rho}[x \mapsto [\overline{r_{i}}]_{n}[n_{1} \mapsto r]]
                }
                {
-                 \widetilde{\rho} \vdash x[\widetilde{e_{1}}] := \widetilde{e_{2}} \Downarrow \widetilde{\rho'}; \cdot
+                 \widetilde{\rho} \vdash x[\widetilde{e_{1}}] := \widetilde{e_{2}} \Downarrow \widetilde{\rho'}; \kw{bind}\:\kappa\:r
                }
@@
 \begin{figure}
+  \small
   \fbox{$\widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \kappa \Downarrow b_{1}, b_{2}$}
@@
                  c_{i} = \mathcal{D}_{\mathcal{B}}(b_{1i}, b_{2i})\\\\
-                 c_{1} = \top \Rightarrow b_{1}, b_{2} = \mathcal{E}_{\mathcal{B}}(c_{1})\quad
-                 c_{1} = \bot \Rightarrow b_{1}, b_{2} = \mathcal{E}_{\mathcal{B}}(c_{2})
+                 c_{1} = \top \Rightarrow b_{1}, b_{2} = \mathcal{E}_{\mathcal{B}}(c_{2})\quad
+                 c_{1} = \bot \Rightarrow b_{1}, b_{2} = \mathcal{E}_{\mathcal{B}}(c_{3})
                }
@@
 
+\begin{figure}
+  \small
+  \fbox{$\Gamma \vdash e : \tau \leadsto \widetilde{e}$}
+  \[
+  \begin{array}{c}
+     \inferrule*[lab={\footnotesize{T-Cons}}]
+               {
+                 \tau = \mathsf{typeof}(c)^{\mathcal{P}}
+               }
+               {
+                 \Gamma \vdash c : \tau \leadsto c 
+               }
+
+     \inferrule*[lab={\footnotesize{T-Add}}]
+               {
+                 \forall i \in \{1,2\}.\:\Gamma \vdash e_{i} : \kw{uint}^{\ell} \leadsto \widetilde{e_{i}}\\\\
+                 \ell = \mathcal{P} \vee \ell = \mathcal{A}
+               }
+               {
+                 \Gamma \vdash e_{1} + e_{2} : \kw{uint}^{\ell} \leadsto \widetilde{e_{1}} +_{\ell} \widetilde{e_{2}}
+               }
+
+    %% \inferrule*[lab={\footnotesize{T-Cons}}]
+    %%            {
+    %%              \tau = \mathsf{typeof}(c)^{\mathcal{P}}
+    %%            }
+    %%            {
+    %%              \Gamma \vdash c : \tau \leadsto c
+    %%            }
+
+    %%  \inferrule*[lab={\footnotesize{T-Var}}]
+    %%            {
+    %%            }
+    %%            {
+    %%              \Gamma \vdash x : \Gamma(x) \leadsto x
+    %%            }
+\\\\
+     \inferrule*[lab={\footnotesize{T-Gt}}]
+               {
+                 \forall i \in \{1,2\}.\:\Gamma \vdash e_{i} : \kw{uint}^{\ell} \leadsto \widetilde{e_{i}}\\\\
+                 \ell = \mathcal{P} \vee \ell = \mathcal{B}
+               }
+               {
+                 \Gamma \vdash e_{1} > e_{2} : \kw{bool}^{\ell} \leadsto \widetilde{e_{1}} >_{\ell} \widetilde{e_{2}}
+               }
+
+     \inferrule*[lab={\footnotesize{T-Read}}]
+               {
+                 \Gamma \vdash x : \sigma^{\ell}[n] \leadsto x\\\\
+                 \Gamma \vdash e : \kw{uint}^{\mathcal{P}} \leadsto \widetilde{e}\\\\
+                 \Gamma \models e < n
+               }
+               {
+                 \Gamma \vdash x[e] : \sigma^{\ell} \leadsto x[\widetilde{e}]
+               }
+
+\\\\
+
+     \inferrule*[lab={\footnotesize{T-Cond}}]
+               {
+                 \Gamma \vdash e : \kw{bool}^{\ell} \leadsto \widetilde{e}\\\\
+                 \forall i \in \{1,2\}.\:\Gamma \vdash e_{i} : \sigma^{\ell'} \leadsto \widetilde{e_{i}}\\\\
+                 \ell = \mathcal{P} \vee (\ell = \mathcal{B} \wedge \ell' =\mathcal{B})
+               }
+               {
+                 \Gamma \vdash \cond{e}{e_{1}}{e_{2}} : \sigma^{\ell'} \leadsto \lcond{\ell}{\widetilde{e}}{\widetilde{e_{1}}}{\widetilde{e_{2}}}
+               }
+
+     \inferrule*[lab={\footnotesize{T-Inp}}]
+               {
+               }
+               {
+                 \Gamma \vdash \kw{in}_{j} : \sigma^{m} \leadsto \kw{in}^{\sigma}_{j}
+               }
+
+\\\\
+
+     \inferrule*[lab={\footnotesize{T-Arr}}]
+               {
+                 \forall i \in \{0 \dots n - 1\}.\:\Gamma \vdash e_{i} : \sigma^{\ell} \leadsto \widetilde{e_{i}}
+               }
+               {
+                 \Gamma \vdash [\overline{e_{i}}]_{n} : \sigma^{\ell}[n] \leadsto [\overline{\widetilde{e_{i}}}]_{n}
+               }
+
+     \inferrule*[lab={\footnotesize{T-Sub}}]
+               {
+                 \Gamma \vdash e : \sigma^{\ell} \leadsto \widetilde{e}
+               }
+               {
+                 \Gamma \vdash e : \sigma^{m} \leadsto \widetilde{e} \rhd m
+               }
+
+  \end{array}
+  \]
+  \\\\
+    \fbox{$\Gamma \vdash s \leadsto \widetilde{s} \mid \Gamma'$}
+  \[
+  \\
+  \begin{array}{c}
+     \inferrule*[lab={\footnotesize{T-Decl}}]
+               {
+                 \psi = \sigma \Rightarrow \tau = \sigma^{\ell}\\\\
+                 \psi = \sigma[n] \Rightarrow \tau = \sigma^{\ell}[n]\\\\
+                 \Gamma \vdash e : \tau \leadsto \widetilde{e}
+               }
+               {
+                 \Gamma \vdash \psi\:x = e \leadsto \tau\:x = \widetilde{e} \mid \Gamma, x:\tau
+               }
+
+     \inferrule*[lab={\footnotesize{T-Assgn}}]
+               {
+                 \Gamma(x) = \sigma^{\ell}\\\\
+                 \Gamma \vdash e : \sigma^{\ell} \leadsto \widetilde{e}
+               }
+               {
+                 \Gamma \vdash x := e \leadsto x = \widetilde{e} \mid \Gamma
+               }
+
+\\\\
+
+     \inferrule*[lab={\footnotesize{T-For}}]
+               {
+                 \Gamma' = \Gamma, x::\kw{uint}^{\mathcal{P}}\\\\
+                 \Gamma' \vdash \loops{x}{n_{2}}{s} \leadsto \loops{x}{n_{2}}{\widetilde{s}} \mid \Gamma'
+               }
+               {
+                 \Gamma \vdash \for{x}{n_{1}}{n_{2}}{s} \leadsto \for{x}{n_{1}}{n_{2}}{\widetilde{s}} \mid \Gamma
+               }
+
+               \\\\
+
+     \inferrule*[lab={\footnotesize{T-Write}}]
+               {
+                 \Gamma \vdash x : \sigma^{\ell}[n] \leadsto x\\\\
+                 \Gamma \vdash e_{1} : \kw{uint}^{\mathcal{P}} \leadsto \widetilde{e_{1}}\\\\
+                 \Gamma \vdash e_{2} : \sigma^{\ell} \leadsto \widetilde{e_{2}}\\\\
+                 \Gamma \models e_{1} < n
+               }
+               {
+                 \Gamma \vdash x[e_{1}] := e_{2} \leadsto x[\widetilde{e_{1}}] := \widetilde{e_{2}} \mid \Gamma
+               }
+
+     \inferrule*[lab={\footnotesize{T-Out}}]
+               {
+                 \Gamma \vdash e : \sigma^{m} \leadsto \widetilde{e}
+               }
+               {
+                 \Gamma \vdash \kw{out}\:e \leadsto \kw{out}\:\widetilde{e} \mid \Gamma
+               }
+
+\\\\
+
+     \inferrule*[lab={\footnotesize{T-If}}]
+               {
+                 \Gamma \vdash e : \kw{bool}^{\mathcal{P}} \leadsto \widetilde{e}\\\\
+                 \forall i \in \{1, 2\}.\:\Gamma \vdash s_{i} \leadsto \widetilde{s_{i}} \mid \_
+               }
+               {
+                 \Gamma \vdash \ite{e}{s_{1}}{s_{2}} \leadsto \ite{\widetilde{e}}{\widetilde{s_{1}}}{\widetilde{s_{2}}}  \mid \Gamma
+               }
+
+     \inferrule*[lab={\footnotesize{T-Seq}}]
+               {
+                 \Gamma \vdash s_{1} \leadsto \widetilde{s_{1}} \mid \Gamma_{1}\\\\
+                 \Gamma_{1} \vdash s_{2} \leadsto \widetilde{s_{2}} \mid \Gamma_{2}
+               }
+               {
+                 \Gamma \vdash s_{1}; s_{2} \leadsto \widetilde{s_{1}}; \widetilde{s_{2}} \mid \Gamma_{2}
+               }
+
+\\\\
+
+     \inferrule*[lab={\footnotesize{T-While}}]
+               {                 
+                 \Gamma(x) = \kw{uint}^{\mathcal{P}}\quad
+                 \Gamma \vdash s \leadsto \widetilde{s} \mid \_\quad
+                 x \notin \mathsf{modifies}(s)
+               }
+               {
+                 \Gamma \vdash \loops{x}{n_{2}}{s} \leadsto \loops{x}{n_{2}}{\widetilde{s}} \mid \Gamma
+               }
+
+  \end{array}
+  \]
+\label{fig:compile}
+\caption{Compilation judgments (selected rules)}
+\end{figure}
+
   %% \\

### 2017-10-22 OaklandSubmission/ld.tex
some text for target language and semantics

@@
 
-\subsubsection{Target language}
-
-\begin{theorem}[Correctness of the compiler]
-
-\end{theorem}
-
 \newcommand{\lcond}[4]{\ensuremath{{{#2}\:?_{{#1}}\:{#3}\::{#4}}}}
@@
   \]
+\caption{Target language}
 \label{fig:tgtlang}
-\caption{Target language}
 \end{figure}
@@
     & & \mid & \kw{mux}\:\kappa\:\kappa_{1}\:\kappa_{2} \mid \widetilde{w} \rhd m\\
-    \ftext{Base value} & \widetilde{w} & ::= & c \mid \kappa\\
-    \ftext{Value} & \widetilde{v} & ::= & \widetilde{w} \mid [\overline{\widetilde{w}_{i}}]_{n}\\
+    \ftext{Base value} & \widetilde{w} & ::= & c \mid r\\
+    \ftext{Value} & \widetilde{v} & ::= & \widetilde{w} \mid \kappa \mid [\overline{\widetilde{w}_{i}}]_{n}\\
+    \ftext{Runtime environment} & \widetilde{\rho} & ::= & \cdot \mid \widetilde{\rho}[x \mapsto \widetilde{v}]\\
     \ftext{Circuit} & \chi & ::= & \cdot \mid \kw{bind}\:\kappa\:r \mid \kw{out}\:\kappa \mid \kappa_{1}; \kappa_{2}
@@
   \]
+\caption{Target runtime}
 \label{fig:tgtruntime}
-\caption{Target runtime}
 \end{figure}
@@
   \]
+\caption{Target semantics (selected rules)}
 \label{fig:tgtsem}
-\caption{Target semantics (selected rules)}
 \end{figure}
+
+\subsubsection*{Target language} Figure~\ref{fig:tgtlang} shows the
+target language of our compiler. The syntax follows that of the source
+languag, except that the types and operators are \emph{labeled}. Below
+we mainly focus on the bits that are different from the source
+language.
+
+Labels $\ell$ consist of secret labels $\mathcal{A}$ and
+$\mathcal{B}$, denoting the arithmetic and boolean shared secrets
+resp., and the public label $\mathcal{P}$. Types $\tau$ are then
+labeled base types $\sigma^{\ell}$ and arrays of labeled base types
+$\sigma^{\ell}[n]$.
+
+Most of the expression forms $\widetilde{e}$ are same as $e$, except
+that the binary operators, and the conditional forms
+are annotated with label $\ell$, denoting how the operators should be
+evaluated ($\mathcal{P}$ for in-clear, and $\mathcal{A}$ and
+$\mathcal{B}$ for using arithmetic or boolean circuits resp.). The
+form $\widetilde{e} \rhd m$ denotes coercing $\widetilde{e}$ to be
+$m$-secret shared. The statements $\widetilde{s}$ are analogous to
+$s$.
+
+The target semantics models the semantics of the code output by
+our compiler (C++ code in our implementation). It computes over the
+public parts of the program, emitting a circuit that can later be
+evaluated using a cryptographic MPC protocol. Crucially, this phase of
+the semantics does not have access to the secrets.
+
+Figure~\ref{fig:tgtruntime} shows the runtime syntax. A wire id range
+$r$ denotes a set of circuit wires that carry the value of a secret
+shared variable in the output circuit. Circuit gates $\kappa$ are
+$r$, input gates, \kw{add}, \kw{gt}, and \kw{mux} gates, and coerce
+gates $\widetilde{w} \rhd m$. Target values $\widetilde{v}$ then
+consist of base values $c$ and $r$, circuit gates $\kappa$, and arrays
+of base values.
+
+Figure~\ref{fig:tgtsem} shows the judgments for the target
+semantics. We first focus on the expression evaluation judgment
+$\widetilde{\rho} \vdash \widetilde{e} \Downarrow
+\widetilde{v}$. Rules ({\sc{S-PAdd}}) and ({\sc{S-SAdd}}) illustrate
+the significance of the operator labels. In particular, the rule
+({\sc{S-PAdd}}) evaluates a public addition $\widetilde{e_{1}}
++_{\mathcal{P}} \widetilde{e_{2}}$ to $n_{1} + n_{2}$, similar to the
+source language. In contrast, the rule
+({\sc{S-SAdd}}) evaluates a secret addition $\widetilde{e_{1}}
++_{\mathcal{A}} \widetilde{e_{2}}$, using an arithmetic add
+gate, to $\kw{add}\:\kappa_{1}\:\kappa_{2}$. Since our compiler
+compiles secret add operations to arithmetic gates, the target
+expressions from our compiler never have $\widetilde{e_{1}}
++_{\mathcal{B}} \widetilde{e_{2}}$ (we present the compilation rules
+later in the section).
+
+\begin{theorem}[Correctness of the compiler]
+
+\end{theorem}
 

### 2017-10-22 OaklandSubmission/ld.tex
finishing the text for languages and semantics

@@
     & & \mid & \kw{mux}\:\kappa\:\kappa_{1}\:\kappa_{2} \mid \widetilde{w} \rhd m\\
-    \ftext{Base value} & \widetilde{w} & ::= & c \mid r\\
-    \ftext{Value} & \widetilde{v} & ::= & \widetilde{w} \mid \kappa \mid [\overline{\widetilde{w}_{i}}]_{n}\\
+    \ftext{Base value} & \widetilde{w} & ::= & c \mid \kappa\\
+    \ftext{Value} & \widetilde{v} & ::= & \widetilde{w} \mid [\overline{\widetilde{w}_{i}}]_{n}\\
     \ftext{Runtime environment} & \widetilde{\rho} & ::= & \cdot \mid \widetilde{\rho}[x \mapsto \widetilde{v}]\\
-    \ftext{Circuit} & \chi & ::= & \cdot \mid \kw{bind}\:\kappa\:r \mid \kw{out}\:\kappa \mid \kappa_{1}; \kappa_{2}
+    \ftext{Circuit} & \chi & ::= & \cdot \mid \kw{bind}\:\kappa\:r \mid \kw{out}\:\kappa \mid \chi_{1}; \chi_{2}
   \end{array}
@@
                {
-                 \widetilde{\rho} \vdash \widetilde{e} \Downarrow \widetilde{w}
+                 \widetilde{\rho} \vdash \widetilde{e} \Downarrow \widetilde{v}
                }
                {
-                 \widetilde{\rho} \vdash \widetilde{e} \rhd m \Downarrow \widetilde{w} \rhd m
+                 \widetilde{\rho} \vdash \widetilde{e} \rhd m \Downarrow \widetilde{v} \rhd m
                }
@@
 gates $\widetilde{w} \rhd m$. Target values $\widetilde{v}$ then
-consist of base values $c$ and $r$, circuit gates $\kappa$, and arrays
-of base values.
+consist of base values $c$ and $\kappa$, and arrays of base values.
 
@@
 +_{\mathcal{B}} \widetilde{e_{2}}$ (we present the compilation rules
-later in the section).
+later in the section). Rules ({\sc{S-PCond}}) and ({\sc{S-SCond}})
+are along the similar lines. Rule ({\sc{S-PCond}}) evaluates a public
+conditional to the values from one of the branches, while the rule
+({\sc{S-SCond}}) evaluates to a \kw{mux} gate that takes input
+circuits from the conditional ($\kappa_{1}$) and both the branches
+($\kappa_{2}$ and $\kappa_{3}$). Rules {\sc{S-Coerce}} and
+{\sc{S-Inp}} evaluate to the coerce and input gates respectively.
 
-\begin{theorem}[Correctness of the compiler]
-
-\end{theorem}
+Statement evaluation $\widetilde{\rho} \vdash \widetilde{s} \Downarrow
+\widetilde{\rho'}; \chi$ evaluates a target statement to produce a new
+environment $\widetilde{\rho'}$, and a circuit $\chi$. $\chi$ (shown
+in Figure~\ref{fig:tgtlang}) is either empty, \kw{bind}-ing of a
+circuit gate $\kappa$ to range $r$, \kw{out} gate, or a sequence of
+circuits. Rules ({\sc{S-DeclC}}), ({\sc{S-DeclCkt}}), and
+({\sc{S-DeclCktArr}}) show the variable declaration cases. Rule
+({\sc{S-DeclC}}) shows the case when $\widetilde{e}$ evaluates to
+$\widetilde{v}$, where $\widetilde{v}$ is a constant or an array of
+constants. In this case, the mapping $x \mapsto \widetilde{v}$ is
+added to the environment, and the resulting circuit is empty. When
+$\widetilde{e}$ evaluates to a circuit $\kappa$, rule
+({\sc{S-DeclCkt}}) picks a fresh $r$, adds the mapping $x \mapsto r$
+to the environment, and outputs the circuit
+$\kw{bind}\:\kappa\:r$. Rule ({\sc{S-DeclCktA}}) is analogous for
+$\widetilde{e}$ evaluating to an array of circuits. The variable
+assignment rules (not shown in the figure) are similar. Rule
+({\sc{S-WriteCkt}}) shows the case for writing to an array, where the
+array contents are secret. Finally, rule ({\sc{S-Out}}) compiles to an
+\kw{out} circuit.
 
@@
   \]
+\caption{Circuit semantics (selected rules)}
 \label{fig:cktsem}
-\caption{Circuit semantics (selected rules)}
 \end{figure}
+
+\subsubsection*{Circuit semantics} The target semantics produces a
+circuit to compute over the secret data using an MPC protocol. With
+our circuit semantics, we model the \emph{functional} aspect of the
+protocol, parametrized by the sharing functions.
+
+During the circuit evaluation, the wire ranges $r$ are
+mapped to (opaque) strings $b$. The semantics of these strings is
+given by pairs of encryption-decryption functions, written as
+$\mathcal{E}_{m}$ and $\mathcal{D}_{m}$ (where $m$ is either
+$\mathcal{A}$ or $\mathcal{B}$). More concretely,
+$\mathcal{E}_{m}(c)$ returns a pair of two string $b_{1}$ and $b_{2}$
+(shares of the two parties), with the property that
+$\mathcal{D}_{m}(b_{1}, b_{2}) = c$. We assume that the underlying MPC
+protocol instantiates $\mathcal{E}_{m}$ and $\mathcal{D}_{m}$
+appropriately.
+
+The circuit semantics is shown in Figure~\ref{fig:cktsem} using the
+judgments $\widehat{\rho_{1}}, \widehat{\rho_{1}} \vdash \kappa
+\Downarrow b_{1}, b_{2}$, and $\widehat{\rho_{1}}, \widehat{\rho_{2}}
+\vdash \chi \Downarrow \widehat{\rho_{1}'}, \widehat{\rho_{2}'}; O$,
+where $\widehat{\rho_{1}}$ and $\widehat{\rho_{2}}$ are the circuit
+environments of the two parties, mapping wire ranges $r$ to strings
+$b$. Rule {\sc{C-Add}} evinces the pattern for evaluating circuit
+gates $\kappa$. To evaluate $\kw{add}\:\kappa_{1}\:\kappa_{2}$, the
+rule first evaluates $\kappa_{1}$ to $(b_{11}, b_{21})$ and
+$\kappa_{2}$ to $(b_{12}, b_{22})$. Shares $(b_{11}, b_{21})$ are then
+combined using $\mathcal{D_{\mathcal{A}}}$ to $n_{1}$, and similarly
+$(b_{12}, b_{22})$ are combined to $n_{2}$. The final output of the
+\kw{add} gate is then $\mathcal{E}_{\mathcal{A}}(n_{1} + n_{2})$. Note
+that this is a functional description of how the \kw{add} gate
+evaluates, of course, concretely $n_{1}$ and $n_{2}$ are not observed
+by the parties. Rule {\sc{C-Coerce}} re-encrypts the shares using the
+scheme for $m$. Coming to the evaluation of circuits, the evaluation
+of \kw{bind} updates the mapping of $r$ in the input environments, and
+the rule {\sc{C-Out}} outputs the clear value $c$ to the observations.
+
+\subsubsection*{Compilation judgments}
+
+\begin{theorem}[Correctness of the compiler]
+
+\end{theorem}
 

### 2017-10-22 OaklandSubmission/ld.tex
text completing s3 except the security theorem

@@
 The target language semantics (a model of the C++ code generated by
-our compiler implementation) ``computes away'' the public parts of the
-compiled program, generating a secure computation circuit.
+our compiler implementation) ``computes away'' the public parts and
+arrays from the compiled program, generating a secure computation
+circuit.
 %
@@
 public parts of the program, emitting a circuit that can later be
-evaluated using a cryptographic MPC protocol. Crucially, this phase of
-the semantics does not have access to the secrets.
+evaluated using a cryptographic MPC protocol. The semantics also
+flattens the circuit so that the circuits are unaware of the array
+structure. Crucially, this phase of the semantics does not have access
+to the secrets.
 
@@
 circuits from the conditional ($\kappa_{1}$) and both the branches
-($\kappa_{2}$ and $\kappa_{3}$). Rules {\sc{S-Coerce}} and
-{\sc{S-Inp}} evaluate to the coerce and input gates respectively.
+($\kappa_{2}$ and $\kappa_{3}$). Rules ({\sc{S-Coerce}}) and
+({\sc{S-Inp}}) evaluate to the coerce and input gates respectively.
 
@@
 
-\subsubsection*{Compilation judgments}
-
-\begin{theorem}[Correctness of the compiler]
-
-\end{theorem}
-
 \begin{figure}
@@
                {
-                 \Gamma' = \Gamma, x::\kw{uint}^{\mathcal{P}}\\\\
+                 \Gamma' = \Gamma, x:\kw{uint}^{\mathcal{P}}\\\\
                  \Gamma' \vdash \loops{x}{n_{2}}{s} \leadsto \loops{x}{n_{2}}{\widetilde{s}} \mid \Gamma'
@@
   \]
+\caption{Compilation judgments (selected rules)}
 \label{fig:compile}
-\caption{Compilation judgments (selected rules)}
 \end{figure}
+
+\subsubsection*{Compilation judgments} We now come to the compilation
+rules shown in Figure~\ref{fig:compile}. We present the rules in a
+declarative style, where the rules are also non-syntax
+directed, and the labels $\ell$ are also chosen
+non-deterministically.
+
+Judgment $\Gamma \vdash e : \tau \leadsto \widetilde{e}$, where
+$\Gamma$ maps variables $x$ to types $\tau$, says that
+under $\Gamma$, $e$ compiles to $\widetilde{e}$ with type $\tau$.
+
+The rule ({\sc{T-Cons}}) assigns the label $\mathcal{P}$ to the
+constants, as the constants are always public. Rule ({\sc{T-Add}})
+compiles an addition to either public addition ($+_{\mathcal{P}}$) or
+secret addition using an arithmetic circuit ($+_{\mathcal{A}}$). As
+our compiler is cryptographic cost aware, it never compiles the
+addition to boolean addition $+_{\mathcal{B}}$. In a similar manner,
+rule ({\sc{T-Gt}}) compiles $e_{1} > e_{2}$ to either public comparison
+or secret comparison using boolean circuits. The rule for conditional
+({\sc{T-Cond}}) has two cases: when the conditional expression $e$ is of
+type $\kw{bool}^{\mathcal{P}}$, both the branches have a base type
+$\sigma^{\ell'}$, for an arbitrary $\ell'$, and the conditional is
+compiled to public conditional, whereas when the conditional
+expression has type $\kw{bool}^{\mathcal{B}}$, $\ell'$ is also
+$\mathcal{B}$, and the conditional is compiled to a secret conditional
+using a boolean circuit. Note that we restrict the type of the branches
+to be of base type. Rule ({\sc{T-Read}}) types an array read. It checks
+that the array index $e$ is public, and uses a static bounds checking
+judgment $\Gamma \models e < n$ to prove that the array index is in
+bounds. Section~\ref{sec:impl} discusses our implementation of bounds
+checking. Rule ({\sc{T-Inp}}) picks a label $m$ for the input. Finally,
+rule ({\sc{T-Sub}}) is the subsumtion rule that coerces an expression of
+type $\sigma^{\ell}$ to an expression of type $\sigma^{m}$ using the
+coerce expression. Importantly, secrets cannot be coerced to public
+values.
+
+Judgment $\Gamma \vdash s : \tau \leadsto \widetilde{s} \mid \Gamma'$
+compiles a statement $s$ resulting in the statement $\widetilde{s}$
+and type environment $\Gamma'$. Rule ({\sc{T-Decl}}) picks a label
+$\ell$, and adds the binding for $x$ to the environment. Rule
+({\sc{T-For}}) adds the loop counter $x$ to $\Gamma$ at type
+$\kw{uint}^{\mathcal{P}}$, and delegates type checking to the
+\kw{while} form. The rule ({\sc{T-Write}}), similar to ({\sc{T-Read}}),
+checks that the index is public, and is in bounds. Rule ({\sc{T-Out}})
+types the expression $e$ at some secret label $m$. Rule ({\sc{T-If}})
+checks that the conditional expression is public, and rule
+({\sc{T-Seq}}) sequences the type environments. Finally, the typing
+rule for the (internal) \kw{while} form ensures that $x$ is mapped in
+$\Gamma$ at type $\kw{uint}^{\mathcal{P}}$, and that $x \notin
+\mathsf{modifies}(s)$--this helps us prove the termination.
+
+\subsubsection*{Correctness theorem} We first define a judgment for
+translation of type environments, $\Gamma \vdash \rho \leadsto
+\widetilde{\rho}; \widehat{\rho_{1}}, \widehat{\rho_{2}}$. The
+judgment is defined inductively on $\Gamma$, and establishes that the
+the static environment $\Gamma$ is consistent with the runtime
+environments $\rho$, $\widetilde{\rho}$, and $\widehat{\rho_{1}}$ and
+$\widehat{\rho_{2}}$. As an example, if $x:\kw{uint}^{\mathcal{P}} \in
+\Gamma$, the judgment checks that $\rho(x) = \widetilde{\rho}(x) =
+n$. We elide the judgment, but it is defined in the supplementary
+material.
+
+With this judgment in hand, the correctness theorem is as follows:
+
+\begin{theorem}[Correctness]
+  $\forall\:\Gamma, s, \widetilde{s}, \Gamma', \rho, \widetilde{\rho}, \widehat{\rho_{1}}, \widehat{\rho_{2}}$, if:
+
+  \begin{enumerate}
+  \item $\Gamma \vdash s \leadsto \widehat{s} \mid \Gamma'$
+  \item $\Gamma \vdash \rho \leadsto \widetilde{\rho}; \widehat{\rho_{1}}, \widehat{\rho_{2}}$
+  \end{enumerate}
+
+  then $\exists \rho', O_{1}, \widetilde{\rho'}, \chi, \widehat{\rho_{1}'}, \widehat{\rho_{1}'}, O_{2}$, s.t.
+
+  \begin{enumerate}
+  \item $\rho \vdash s \Downarrow \rho'; O_{1}$
+  \item $\widetilde{\rho} \vdash \widetilde{s} \Downarrow \widetilde{\rho'}; \chi$
+  \item $\widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \chi \Downarrow \widehat{\rho_{1}'}, \widehat{\rho_{2}'}; O_{2}$
+  \item $\Gamma' \vdash \rho' \leadsto \widetilde{\rho'}; \widehat{\rho_{1}'}, \widehat{\rho_{2}'}$
+  \item $O_{1} = O_{2}$
+  \end{enumerate}
+
+\end{theorem}
+
+The theorem states that if a source statement $s$ is well-typed, and
+compiled to $\widetilde{s}$, then $s$ terminates in the source
+semantics with observations $O_{1}$, $\widetilde{s}$ terminates in the
+target semantics with circuit $\chi$, and $\chi$ terminates in the
+circuit semantics with observations $O_{2}$, where $O_{1} = O_{2}$
+Furthermore, the final environments are also in the compilation
+relation.
+
+\subsubsection*{Security theorem}
 

### 2017-10-22 OaklandSubmission/ld.tex
nits

@@
 \subsubsection*{Correctness theorem} We first define a judgment for
-translation of type environments, $\Gamma \vdash \rho \leadsto
+translation of environments, $\Gamma \vdash \rho \leadsto
 \widetilde{\rho}; \widehat{\rho_{1}}, \widehat{\rho_{2}}$. The
-judgment is defined inductively on $\Gamma$, and establishes that the
+judgment is defined inductively on $\Gamma$, and establishes that
 the static environment $\Gamma$ is consistent with the runtime
@@
 \Gamma$, the judgment checks that $\rho(x) = \widetilde{\rho}(x) =
-n$. We elide the judgment, but it is defined in the supplementary
-material.
+n$, for some $n$. We elide the judgment, but it is defined in the
+supplementary material.
 
@@
 
-  then $\exists \rho', O_{1}, \widetilde{\rho'}, \chi, \widehat{\rho_{1}'}, \widehat{\rho_{1}'}, O_{2}$, s.t.
+  then $\exists \rho', O_{1}, \widetilde{\rho}', \chi, \widehat{\rho_{1}}', \widehat{\rho_{1}}', O_{2}$, s.t.
 
@@
   \item $\rho \vdash s \Downarrow \rho'; O_{1}$
-  \item $\widetilde{\rho} \vdash \widetilde{s} \Downarrow \widetilde{\rho'}; \chi$
-  \item $\widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \chi \Downarrow \widehat{\rho_{1}'}, \widehat{\rho_{2}'}; O_{2}$
-  \item $\Gamma' \vdash \rho' \leadsto \widetilde{\rho'}; \widehat{\rho_{1}'}, \widehat{\rho_{2}'}$
+  \item $\widetilde{\rho} \vdash \widetilde{s} \Downarrow \widetilde{\rho}'; \chi$
+  \item $\widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \chi \Downarrow \widehat{\rho_{1}}', \widehat{\rho_{2}}'; O_{2}$
+  \item $\Gamma' \vdash \rho' \leadsto \widetilde{\rho}'; \widehat{\rho_{1}}', \widehat{\rho_{2}}'$
   \item $O_{1} = O_{2}$

### 2017-10-22 OaklandSubmission/ld.tex
typo

@@
 checking. Rule ({\sc{T-Inp}}) picks a label $m$ for the input. Finally,
-rule ({\sc{T-Sub}}) is the subsumtion rule that coerces an expression of
+rule ({\sc{T-Sub}}) is the subsumption rule that coerces an expression of
 type $\sigma^{\ell}$ to an expression of type $\sigma^{m}$ using the

### 2017-10-23 OaklandSubmission/intro.tex
intro

@@
 \sectionref{eval}). \divya{Moreover, unlike previous systems, our system can scale to arbitrarily large computations using {\em pipelining} (see below). What do we say about Yao pipelining?}
-Hence, we show that our framework achieves all three properties of generality, performance and programmer productivity.
+Hence, we show that our framework achieves all four properties of generality, performance, programmer productivity and scalability.
 
-\divya{Points.. 1. Prior to our work, there was CBMGC no perf; ABY no prog; 
-3. Programmability allows writing programs generating large circuits... 
-}
+Prior to our work, there have been systems that provide only a subset of the above properties and lack in remaining properties.
+For instance, using the framework of CBMC-GC \cite{cbmcgc}, one can write functionalities in high-level language like C but it gives poor performance due to its usage of boolean circuits as underlying representation.
+Similarly, ABY framework can potentially give good performance but is not programmable.
+Technique of pipelining Yao garbled circuits \cite{yao-pipe} allowed scaling to lage circuits but gives poor performance.
+\divya{what about programming?} To summarize, none of the prior frameworks could be used to evaluate the benchmarks such as linear regression, deep neural network prediction etc.
+In fact, it was widely believed that generic 2PC would not scale to the secure prediction and hence, various protocols were hand-crafted for this task \cite{shafindss,wu,secureml,minionn}.
+Our framework contradicts this belief.
+
+In this work, we solve an interesting design problem to give a user friendly language that is expressive enough together with a compiler generating efficient secure 2PC protocols.
+Programmability and performance of our system has enabled us to write large programs whose circuits  exhaust the memory of the system (~32 GB).
+To combat this issue, we give a technique called pipelining (see below) that allows scaling of our  scaling of our framework to arbitrary size computations.
 

### 2017-10-23 OaklandSubmission/ld.tex
another nit
added some high level security theorems etc.

@@
                {
-                 \Gamma \vdash \kw{in}_{j} : \sigma^{m} \leadsto \kw{in}^{\sigma}_{j}
+                 \Gamma \vdash \kw{in}_{j} : \sigma^{m} \leadsto \kw{in}^{m}_{j}
                }
@@
 
-\begin{theorem}[Correctness]
+\begin{theorem}[Correctness]\label{theorem:correctness}
   $\forall\:\Gamma, s, \widetilde{s}, \Gamma', \rho, \widetilde{\rho}, \widehat{\rho_{1}}, \widehat{\rho_{2}}$, if:
@@
 
-\subsubsection*{Security theorem}
-
+\nc{Super high level security theorem}
+\subsubsection*{Security theorem} The protocols we generate provide simulation-based security against a semi-honest adversary, in the framework of ~\cite{gmw,can00,can01}.
+At a very high level, in this framework, parties are modeled as non-uniform interactive turing machines (ITMs), with inputs provided by an environment $\env$.
+An adversary $\adv$, selects and ``corrupts'' one of the parties - however, $\adv$ still follows the protocol specification.
+$\adv$ interacts with the $\env$, which observes the view of the corrupted party.
+At the end of the interaction, $\env$ outputs a single bit.
+Two different interactions are defined: the {\em real world} and an {\em ideal world}.
+In the real interaction, the parties run the protocol $\prot$ in the presence of $\adv$ and $\env$.
+Let $\real_{\prot,\adv,\env}$ denote the binary distribution ensemble describing $\env$'s output in this interaction.
+In the ideal interaction, parties send their inputs to an additional entity, a trusted functionality machine $\F$ that carries the desired computation truthfully.
+Let $\simu$ (the simulator) denote the adversary in this idealized execution, and $\ideal_{\F,\simu,\env}$ the binary distribution ensemble describing $\env$'s output after interacting with adversary $\simu$ and ideal functionality $\F$.
+A protocol $\prot$ is said to {\em securely realize} a functionality $\F$ if for every adversary $\adv$ in the real interaction, there is an adversary $\simu$ in the ideal interaction, such that no environment $\env$, on any input, can tell the real interaction apart from the ideal interaction, except with negligible probability (in the security parameter $\secparam$).
+More precisely, if the two binary distribution ensembles above are computationally indistinguishable.
+\nc{I'm not happy with the theorem statement.}
+\begin{theorem}[Security]\label{theorem:security}
+Let $\F$ be a functionality in our source language with outputs (or observations) $O_1$, that is compiled into a protocol $\prot$ in our target language with outputs $O_2$, then $\prot$ securely realizes $\F$.
+\end{theorem}
+\nc{We should probably connect things to earlier? Theorem 1 and 2 have very different languages.}
+\noindent {\em Proof.} From Theorem \ref{theorem:correctness}, we have that the observations in $\prot$ and $\F$ are the same (i.e.
+$O_2 = O_1$).
+Now, from the simulation security of our backend (\cite{aby}), the proof of the above theorem follows.
+\nc{I dont know how to say more.}
   %% \\

### 2017-10-23 OaklandSubmission/pipe.tex
added some high level security theorems etc.

@@
 \label{sec:pipe}
-\nc{Started high level description. Can go deeper and correct notation as we proceed.}
+
 We now describe our technique of ``pipelining'' that allows us to compile and execute large programs that would otherwise have not fit in memory (due to the large size of the circuit generated by ABY).
@@
 
-Our compiler, given $P$, will first split $P$ into $P_1||P_2||\cdots||P_k$ with the guarantee that the ABY circuits generated for every $Q_i$ defined above is small enough to fit into memory.
-It will then compile and execute the $k$ programs $Q_1||Q_2||\cdots||Q_k$ sequentially (freeing up memory usage after every execution) to produce final output $z$ to both parties.
+Given $P$, we will first split $P$ into $P_1||P_2||\cdots||P_k$ with the guarantee that the ABY circuits generated for every $Q_i$ defined above is small enough to fit into memory.
+We then compile and execute the $k$ programs $Q_1||Q_2||\cdots||Q_k$ sequentially (freeing up memory usage after every execution) to produce final output $z$ to both parties.
+
+\begin{theorem}[Pipelining security]
+Let $P$ be a program in our source language with output $z$, that is split into programs $Q_1||Q_2||\cdots ||Q_k$ (as defined in the pipelining algorithm above) and compiled into protocols $\prot_1||\prot_2\cdots ||\prot_k$ that outputs $r_1||r_2||\cdots||z$ (to Party A) and $o_1||o_2||\cdots||z$ (to Party B), then $\prot$ securely realizes $P$.
+\end{theorem}
+
+\noindent {\em Proof.} From Theorem \ref{theorem:security}, it follows that $\prot_i$ securely realizes $Q_i$ for every $1\leq i\leq k$.
+Hence, we know that the only information learnt by Party A is $r_1||r_2||\cdots||r_{k-1}||z$ and by Party B is $o_1||o_2||\cdots||o_{k-1}||z$.
+From the way we defined $r_i$ and $o_i$ ($1\leq i\leq k-1$), to be secret shares of $s_i$, they are individually uniformly random (in $(\mathbb{Z},+)$ and hence perfectly simulatable.
+From this, the proof of the above theorem follows.

### 2017-10-23 OaklandSubmission/ld.tex
pass over s3 fixing various nits

@@
 arrays from the compiled program, generating a secure computation
-circuit.
+circuit. Crucially, this semantics does not have access to the secret
+inputs--those are processed by the secure computation circuit.
 %
 Finally, we formalize the circuit semantics that computes the
-generated circuit, and like the source semantics, emits observations.
+generated circuit, and like the source semantics, emits observations
+corresponding to the values revealed to the parties.
 
-We then present the compilation judgments. To prove the correctness
+We then present the compilation rules. To prove the correctness
 of our compiler, we prove that it preserves the observations. For
@@
                  \rho_{2} = [\rho_{1}]_{\mathsf{dom}(\rho)}[x \mapsto \rho_{1}(x) + 1]\\\\
-                 \rho_{2} \vdash \loops{x}{n}{s} \Downarrow \rho'; O_{2}
+                 \rho_{2} \vdash \loops{x}{n}{s} \Downarrow \rho_{3}; O_{2}
                }
                {
-                 \rho \vdash \loops{x}{n}{s} \Downarrow \rho'; O_{1}, O_{2}
+                 \rho \vdash \loops{x}{n}{s} \Downarrow \rho_{3}; O_{1}, O_{2}
                }
@@
                  c = \bot \Rightarrow s = s_{2}\\\\
-                 \rho \vdash s \Downarrow \rho'; O
+                 \rho \vdash s \Downarrow \rho_{1}; O
                }
                {
-                 \rho \vdash \ite{e}{s_{1}}{s_{2}} \Downarrow \rho'; O
+                 \rho \vdash \ite{e}{s_{1}}{s_{2}} \Downarrow \rho_{1}; O
                }
@@
 (Figure~\ref{fig:srclang}) is a simple imperative language. Types
-$\psi$ in the language consist of the base types $\sigma$, and arrays
+$\psi$ consist of the base types $\sigma$, and arrays
 of base types $\sigma[n]$, where $n$ is the array length. Though we
@@
 \vspace{0.2cm}
-Values consist of constants and array values. Runtime environment
-$\rho$ maps variables to values. Observations are sequences of
+Values consist of constants and array values, runtime environment
+$\rho$ maps variables to values, and observations are sequences of
 constants.
@@
 the environment. Rule ({\sc{E-Add}}) inductively evaluates $e_{1}$ and
-$e_{2}$ and evaluates to their addition. Rule ({\sc{E-Read}})
+$e_{2}$, and evaluates to their sum. Rule ({\sc{E-Read}})
 evaluates an array read operation. It first evaluates $x$ to an array
-value $[\overline{c_{i}}]_{n_{1}}$, and $e$ to an \kw{uint} value
+value $[\overline{c_{i}}]_{n_{1}}$, and $e$ to a \kw{uint} value
 $n$. It then returns $c_{n}$, the $n$-th index value in the array,
@@
 
-The judgment $\rho \vdash s \Downarrow \rho'; O$ represents the
-big-step evaluation of an statement $s$ under environment $\rho$
-producing a new environment $\rho'$ and observations $O$. Rule
+The judgment $\rho \vdash s \Downarrow \rho_{1}; O$ represents the
+big-step evaluation of a statement $s$ under environment $\rho$
+producing a new environment $\rho_{1}$ and observations $O$. Rule
 ({\sc{E-Decl}}) evaluates the expression $e$ to $v$, and returns the
@@
 either $s_{1}$ or $s_{2}$ accordingly. $\kw{for}$ statements evaluate
-through the internal $\kw{while}$ syntax. Specifically, rule
+through the internal $\kw{while}$ syntax. Specifically, the rule
 ({\sc{E-For}}) updates $\rho$ with the initial counter value $n_{1}$,
 evaluates $\loops{x}{n_{2}}{s}$ to $\rho_{1}; O$, and returns
-$\rho_{1} - \{x\}$ (removing $x$ from the environment) and $O$. Rule
+$\rho_{1} - \{x\}$ (removing $x$ from $\rho_{1}$) and $O$. Rule
 ({\sc{E-LoopI}}) shows the inductive case for \kw{while}
@@
 when $\rho(x) > n$. Finally, the rule ({\sc{E-Out}}) evaluates the
-expression, and adds it value to the observations.
+expression, and adds its value to the observations.
 
@@
                  \widetilde{v} = c \vee \widetilde{v} = [\overline{c_{i}}]_{n}\\\\
-                 \widetilde{\rho'} = \widetilde{\rho}, x \mapsto \widetilde{v}
+                 \widetilde{\rho}_{1} = \widetilde{\rho}, x \mapsto \widetilde{v}
                }
                {
-                 \widetilde{\rho} \vdash \tau\:x = \widetilde{e} \Downarrow \widetilde{\rho'}; \cdot
+                 \widetilde{\rho} \vdash \tau\:x = \widetilde{e} \Downarrow \widetilde{\rho}_{1}; \cdot
                }
@@
                  \mathsf{fresh}\:r\\\\
-                 \widetilde{\rho'} = \widetilde{\rho}, x \mapsto r\quad
+                 \widetilde{\rho}_{1} = \widetilde{\rho}, x \mapsto r\quad
                  \chi = \kw{bind}\:\kappa\:r
@@
                {
-                 \widetilde{\rho} \vdash \tau\:x = \widetilde{e} \Downarrow \widetilde{\rho'}; \chi
+                 \widetilde{\rho} \vdash \tau\:x = \widetilde{e} \Downarrow \widetilde{\rho}_{1}; \chi
                }
@@
                  \forall i \in \{0 \dots n - 1\}.\:\mathsf{fresh}\:r_{i}\\\\
-                 \widetilde{\rho'} = \widetilde{\rho}, x \mapsto [r_{i}]_{n}\quad
+                 \widetilde{\rho}_{1} = \widetilde{\rho}, x \mapsto [r_{i}]_{n}\quad
                  \chi = \overline{\kw{bind}\:\kappa_{i}\:r_{i}}
@@
                  c = \bot \Rightarrow \widetilde{s} = \widetilde{s_{2}}\\\\
-                 \widetilde{\rho} \vdash \widetilde{s} \Downarrow \widetilde{\rho'}; \chi
+                 \widetilde{\rho} \vdash \widetilde{s} \Downarrow \widetilde{\rho}_{1}; \chi
                }
                {
-                 \widetilde{\rho} \vdash \ite{\widetilde{e}}{\widetilde{s_{1}}}{\widetilde{s_{2}}} \Downarrow \widetilde{\rho'}; \chi
+                 \widetilde{\rho} \vdash \ite{\widetilde{e}}{\widetilde{s_{1}}}{\widetilde{s_{2}}} \Downarrow \widetilde{\rho}_{1}; \chi
                }
@@
                  \widetilde{\rho} \vdash \widetilde{e_{2}} \Downarrow \kappa\\\\
-                 \widetilde{\rho'} = \widetilde{\rho}[x \mapsto [\overline{r_{i}}]_{n}[n_{1} \mapsto r]]
+                 \widetilde{\rho}_{1} = \widetilde{\rho}[x \mapsto [\overline{r_{i}}]_{n}[n_{1} \mapsto r]]
                }
                {
-                 \widetilde{\rho} \vdash x[\widetilde{e_{1}}] := \widetilde{e_{2}} \Downarrow \widetilde{\rho'}; \kw{bind}\:\kappa\:r
+                 \widetilde{\rho} \vdash x[\widetilde{e_{1}}] := \widetilde{e_{2}} \Downarrow \widetilde{\rho}_{1}; \kw{bind}\:\kappa\:r
                }
@@
 target language of our compiler. The syntax follows that of the source
-languag, except that the types and operators are \emph{labeled}. Below
+language, except that the types and operators are \emph{labeled}. Below
 we mainly focus on the bits that are different from the source
@@
 evaluated using a cryptographic MPC protocol. The semantics also
-flattens the circuit so that the circuits are unaware of the array
-structure. Crucially, this phase of the semantics does not have access
-to the secrets.
+\emph{flattens} the arrays so that the circuits are unaware of the
+array structure. Crucially, this phase of the semantics does not have
+access to the secrets.
 
@@
 $r$ denotes a set of circuit wires that carry the value of a secret
-shared variable in the output circuit. Circuit gates $\kappa$ are
-$r$, input gates, \kw{add}, \kw{gt}, and \kw{mux} gates, and coerce
-gates $\widetilde{w} \rhd m$. Target values $\widetilde{v}$ then
-consist of base values $c$ and $\kappa$, and arrays of base values.
+shared variable in the output circuit (we will concretely define these
+values later as part of the circuit semantics). Circuit gates $\kappa$
+are $r$, input gates, \kw{add}, \kw{gt}, and \kw{mux} gates, and
+coerce gates $\widetilde{w} \rhd m$. Target values $\widetilde{v}$
+then consist of base values $c$ and $\kappa$, and arrays of base
+values.
 
@@
 ({\sc{S-SAdd}}) evaluates a secret addition $\widetilde{e_{1}}
-+_{\mathcal{A}} \widetilde{e_{2}}$, using an arithmetic add
-gate, to $\kw{add}\:\kappa_{1}\:\kappa_{2}$. Since our compiler
-compiles secret add operations to arithmetic gates, the target
++_{\mathcal{A}} \widetilde{e_{2}}$ to an arithmetic add
+gate $\kw{add}\:\kappa_{1}\:\kappa_{2}$. Foreshadowing on the
+compilation rules to be presented later, the target
 expressions from our compiler never have $\widetilde{e_{1}}
-+_{\mathcal{B}} \widetilde{e_{2}}$ (we present the compilation rules
-later in the section). Rules ({\sc{S-PCond}}) and ({\sc{S-SCond}})
++_{\mathcal{B}} \widetilde{e_{2}}$, as our compiler is aware that $+$
+is more performant using arithmetic circuit, rather than a boolean
+one. Rules ({\sc{S-PCond}}) and ({\sc{S-SCond}})
 are along the similar lines. Rule ({\sc{S-PCond}}) evaluates a public
-conditional to the values from one of the branches, while the rule
+conditional to the value from one of the branches, while the rule
 ({\sc{S-SCond}}) evaluates to a \kw{mux} gate that takes input
 circuits from the conditional ($\kappa_{1}$) and both the branches
-($\kappa_{2}$ and $\kappa_{3}$). Rules ({\sc{S-Coerce}}) and
-({\sc{S-Inp}}) evaluate to the coerce and input gates respectively.
+($\kappa_{2}$ and $\kappa_{3}$). Once again for performance reasons,
+the target expressions from our compiler do not have
+$\lcond{\mathcal{A}}{e_{1}}{e_{2}}{e_{3}}$. Rules
+({\sc{S-Coerce}}) and ({\sc{S-Inp}}) evaluate to the coerce and input
+gates respectively.
 
 Statement evaluation $\widetilde{\rho} \vdash \widetilde{s} \Downarrow
-\widetilde{\rho'}; \chi$ evaluates a target statement to produce a new
-environment $\widetilde{\rho'}$, and a circuit $\chi$. $\chi$ (shown
+\widetilde{\rho}_{1}; \chi$ evaluates a target statement to produce a new
+environment $\widetilde{\rho}_{1}$, and a circuit $\chi$. $\chi$ (shown
 in Figure~\ref{fig:tgtlang}) is either empty, \kw{bind}-ing of a
@@
 circuits. Rules ({\sc{S-DeclC}}), ({\sc{S-DeclCkt}}), and
-({\sc{S-DeclCktArr}}) show the variable declaration cases. Rule
+({\sc{S-DeclCktA}}) show the variable declaration cases. Rule
 ({\sc{S-DeclC}}) shows the case when $\widetilde{e}$ evaluates to
-$\widetilde{v}$, where $\widetilde{v}$ is a constant or an array of
-constants. In this case, the mapping $x \mapsto \widetilde{v}$ is
-added to the environment, and the resulting circuit is empty. When
+$\widetilde{v}$, where $\widetilde{v}$ is either a constant or an
+array of constants. In this case, the mapping $x \mapsto
+\widetilde{v}$ is added to the environment, and the resulting circuit
+is empty. When 
 $\widetilde{e}$ evaluates to a circuit $\kappa$, rule
@@
 During the circuit evaluation, the wire ranges $r$ are
-mapped to (opaque) strings $b$. The semantics of these strings is
+mapped to (random) strings $b$. The semantics of these strings is
 given by pairs of encryption-decryption functions, written as
@@
 The circuit semantics is shown in Figure~\ref{fig:cktsem} using the
-judgments $\widehat{\rho_{1}}, \widehat{\rho_{1}} \vdash \kappa
-\Downarrow b_{1}, b_{2}$, and $\widehat{\rho_{1}}, \widehat{\rho_{2}}
-\vdash \chi \Downarrow \widehat{\rho_{1}'}, \widehat{\rho_{2}'}; O$,
-where $\widehat{\rho_{1}}$ and $\widehat{\rho_{2}}$ are the circuit
+judgments $\widehat{\rho}_{1}, \widehat{\rho}_{2} \vdash \kappa
+\Downarrow b_{1}, b_{2}$, and $\widehat{\rho}_{1}, \widehat{\rho}_{2}
+\vdash \chi \Downarrow \widehat{\rho}'_{1}, \widehat{\rho}'_{2}; O$,
+where $\widehat{\rho}_{1}$ and $\widehat{\rho}_{2}$ are the circuit
 environments of the two parties, mapping wire ranges $r$ to strings
-$b$. Rule {\sc{C-Add}} evinces the pattern for evaluating circuit
+$b$. Rule ({\sc{C-Add}}) evinces the pattern for evaluating circuit
 gates $\kappa$. To evaluate $\kw{add}\:\kappa_{1}\:\kappa_{2}$, the
@@
 evaluates, of course, concretely $n_{1}$ and $n_{2}$ are not observed
-by the parties. Rule {\sc{C-Coerce}} re-encrypts the shares using the
+by the parties. Rule ({\sc{C-Coerce}}) re-encrypts the shares using the
 scheme for $m$. Coming to the evaluation of circuits, the evaluation
 of \kw{bind} updates the mapping of $r$ in the input environments, and
-the rule {\sc{C-Out}} outputs the clear value $c$ to the observations.
+the rule ({\sc{C-Out}}) outputs the clear value $c$ to the
+observations.
 
@@
                  \Gamma \vdash e : \kw{bool}^{\ell} \leadsto \widetilde{e}\\\\
-                 \forall i \in \{1,2\}.\:\Gamma \vdash e_{i} : \sigma^{\ell'} \leadsto \widetilde{e_{i}}\\\\
+                 \forall i \in \{1,2\}.\:\Gamma \vdash e_{i} : \sigma^{\ell_{1}} \leadsto \widetilde{e_{i}}\\\\
                  \ell = \mathcal{P} \vee (\ell = \mathcal{B} \wedge \ell' =\mathcal{B})
@@
                {
-                 \Gamma \vdash \cond{e}{e_{1}}{e_{2}} : \sigma^{\ell'} \leadsto \lcond{\ell}{\widetilde{e}}{\widetilde{e_{1}}}{\widetilde{e_{2}}}
+                 \Gamma \vdash \cond{e}{e_{1}}{e_{2}} : \sigma^{\ell_{1}} \leadsto \lcond{\ell}{\widetilde{e}}{\widetilde{e_{1}}}{\widetilde{e_{2}}}
                }
@@
   \\\\
-    \fbox{$\Gamma \vdash s \leadsto \widetilde{s} \mid \Gamma'$}
+    \fbox{$\Gamma \vdash s \leadsto \widetilde{s} \mid \Gamma_{1}$}
   \[
@@
                {
-                 \Gamma' = \Gamma, x:\kw{uint}^{\mathcal{P}}\\\\
-                 \Gamma' \vdash \loops{x}{n_{2}}{s} \leadsto \loops{x}{n_{2}}{\widetilde{s}} \mid \Gamma'
+                 \Gamma_{1} = \Gamma, x:\kw{uint}^{\mathcal{P}}\\\\
+                 \Gamma_{1} \vdash \loops{x}{n_{2}}{s} \leadsto \loops{x}{n_{2}}{\widetilde{s}} \mid \Gamma_{1}
                }
@@
 rules shown in Figure~\ref{fig:compile}. We present the rules in a
-declarative style, where the rules are also non-syntax
-directed, and the labels $\ell$ are also chosen
-non-deterministically.
+declarative style, where the rules are non-syntax
+directed, and the labels $\ell$ are chosen
+non-deterministically. Section~\ref{sec:impl} describes the inference
+scheme in our implementation.
 
@@
 constants, as the constants are always public. Rule ({\sc{T-Add}})
-compiles an addition to either public addition ($+_{\mathcal{P}}$) or
+compiles an addition to either public addition ($+_{\mathcal{P}}$), or
 secret addition using an arithmetic circuit ($+_{\mathcal{A}}$). As
@@
 addition to boolean addition $+_{\mathcal{B}}$. In a similar manner,
-rule ({\sc{T-Gt}}) compiles $e_{1} > e_{2}$ to either public comparison
+rule ({\sc{T-Gt}}) compiles $e_{1} > e_{2}$ to either public comparison,
 or secret comparison using boolean circuits. The rule for conditional
@@
 type $\kw{bool}^{\mathcal{P}}$, both the branches have a base type
-$\sigma^{\ell'}$, for an arbitrary $\ell'$, and the conditional is
+$\sigma^{\ell_{1}}$, for an arbitrary $\ell_{1}$, and the conditional is
 compiled to public conditional, whereas when the conditional
-expression has type $\kw{bool}^{\mathcal{B}}$, $\ell'$ is also
+expression has type $\kw{bool}^{\mathcal{B}}$, $\ell_{1}$ is also
 $\mathcal{B}$, and the conditional is compiled to a secret conditional
@@
 
-Judgment $\Gamma \vdash s : \tau \leadsto \widetilde{s} \mid \Gamma'$
+Judgment $\Gamma \vdash s : \tau \leadsto \widetilde{s} \mid \Gamma_{1}$
 compiles a statement $s$ resulting in the statement $\widetilde{s}$
-and type environment $\Gamma'$. Rule ({\sc{T-Decl}}) picks a label
+and type environment $\Gamma_{1}$. Rule ({\sc{T-Decl}}) picks a label
 $\ell$, and adds the binding for $x$ to the environment. Rule
@@
 \kw{while} form. The rule ({\sc{T-Write}}), similar to ({\sc{T-Read}}),
-checks that the index is public, and is in bounds. Rule ({\sc{T-Out}})
+checks that the index has type $\kw{uint}^{\mathcal{P}}$, and is in
+bounds. Rule ({\sc{T-Out}})
 types the expression $e$ at some secret label $m$. Rule ({\sc{T-If}})
@@
 $\Gamma$ at type $\kw{uint}^{\mathcal{P}}$, and that $x \notin
-\mathsf{modifies}(s)$--this helps us prove the termination.
+\mathsf{modifies}(s)$--this is necessary for ensuring the termination
+of our semantics.
 
@@
 translation of environments, $\Gamma \vdash \rho \leadsto
-\widetilde{\rho}; \widehat{\rho_{1}}, \widehat{\rho_{2}}$. The
+\widetilde{\rho}; \widehat{\rho}_{1}, \widehat{\rho}_{2}$. The
 judgment is defined inductively on $\Gamma$, and establishes that
 the static environment $\Gamma$ is consistent with the runtime
-environments $\rho$, $\widetilde{\rho}$, and $\widehat{\rho_{1}}$ and
-$\widehat{\rho_{2}}$. As an example, if $x:\kw{uint}^{\mathcal{P}} \in
+environments $\rho$, $\widetilde{\rho}$, and $\widehat{\rho}_{1}$ and
+$\widehat{\rho}_{2}$. As an example, if $x:\kw{uint}^{\mathcal{P}} \in
 \Gamma$, the judgment checks that $\rho(x) = \widetilde{\rho}(x) =
-n$, for some $n$. We elide the judgment, but it is defined in the
-supplementary material.
+n$, for some $n$. We elide the judgment, providing the full definition
+in the supplementary material.
 
@@
 \begin{theorem}[Correctness]\label{theorem:correctness}
-  $\forall\:\Gamma, s, \widetilde{s}, \Gamma', \rho, \widetilde{\rho}, \widehat{\rho_{1}}, \widehat{\rho_{2}}$, if:
+  $\forall\:\Gamma, s, \widetilde{s}, \Gamma_{1}, \rho, \widetilde{\rho}, \widehat{\rho}_{1}, \widehat{\rho}_{2}$, if:
 
   \begin{enumerate}
-  \item $\Gamma \vdash s \leadsto \widehat{s} \mid \Gamma'$
-  \item $\Gamma \vdash \rho \leadsto \widetilde{\rho}; \widehat{\rho_{1}}, \widehat{\rho_{2}}$
+  \item $\Gamma \vdash s \leadsto \widehat{s} \mid \Gamma_{1}$
+  \item $\Gamma \vdash \rho \leadsto \widetilde{\rho}; \widehat{\rho}_{1}, \widehat{\rho}_{2}$
   \end{enumerate}
 
-  then $\exists \rho', O_{1}, \widetilde{\rho}', \chi, \widehat{\rho_{1}}', \widehat{\rho_{1}}', O_{2}$, s.t.
+  then $\exists \rho_{1}, O_{1}, \widetilde{\rho}_{1}, \chi, \widehat{\rho}'_{1}, \widehat{\rho}'_{2}, O_{2}$, s.t.
 
   \begin{enumerate}
-  \item $\rho \vdash s \Downarrow \rho'; O_{1}$
-  \item $\widetilde{\rho} \vdash \widetilde{s} \Downarrow \widetilde{\rho}'; \chi$
-  \item $\widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \chi \Downarrow \widehat{\rho_{1}}', \widehat{\rho_{2}}'; O_{2}$
-  \item $\Gamma' \vdash \rho' \leadsto \widetilde{\rho}'; \widehat{\rho_{1}}', \widehat{\rho_{2}}'$
+  \item $\rho \vdash s \Downarrow \rho_{1}; O_{1}$
+  \item $\widetilde{\rho} \vdash \widetilde{s} \Downarrow \widetilde{\rho}_{1}; \chi$
+  \item $\widehat{\rho}_{1}, \widehat{\rho}_{2} \vdash \chi \Downarrow \widehat{\rho}'_{1}, \widehat{\rho}'_{2}; O_{2}$
+  \item $\Gamma_{1} \vdash \rho_{1} \leadsto \widetilde{\rho}_{1}; \widehat{\rho}'_{1}, \widehat{\rho}'_{2}$
   \item $O_{1} = O_{2}$
@@
 The theorem states that if a source statement $s$ is well-typed, and
-compiled to $\widetilde{s}$, then $s$ terminates in the source
+compiles to $\widetilde{s}$, then $s$ terminates in the source
 semantics with observations $O_{1}$, $\widetilde{s}$ terminates in the

### 2017-10-24 OaklandSubmission/ld.tex
comment on the security theorem

@@
 \nc{Super high level security theorem}
-\subsubsection*{Security theorem} The protocols we generate provide simulation-based security against a semi-honest adversary, in the framework of ~\cite{gmw,can00,can01}.
-At a very high level, in this framework, parties are modeled as non-uniform interactive turing machines (ITMs), with inputs provided by an environment $\env$.
-An adversary $\adv$, selects and ``corrupts'' one of the parties - however, $\adv$ still follows the protocol specification.
-$\adv$ interacts with the $\env$, which observes the view of the corrupted party.
-At the end of the interaction, $\env$ outputs a single bit.
-Two different interactions are defined: the {\em real world} and an {\em ideal world}.
-In the real interaction, the parties run the protocol $\prot$ in the presence of $\adv$ and $\env$.
-Let $\real_{\prot,\adv,\env}$ denote the binary distribution ensemble describing $\env$'s output in this interaction.
-In the ideal interaction, parties send their inputs to an additional entity, a trusted functionality machine $\F$ that carries the desired computation truthfully.
-Let $\simu$ (the simulator) denote the adversary in this idealized execution, and $\ideal_{\F,\simu,\env}$ the binary distribution ensemble describing $\env$'s output after interacting with adversary $\simu$ and ideal functionality $\F$.
-A protocol $\prot$ is said to {\em securely realize} a functionality $\F$ if for every adversary $\adv$ in the real interaction, there is an adversary $\simu$ in the ideal interaction, such that no environment $\env$, on any input, can tell the real interaction apart from the ideal interaction, except with negligible probability (in the security parameter $\secparam$).
-More precisely, if the two binary distribution ensembles above are computationally indistinguishable.
+\subsubsection*{Security theorem} The protocols we generate provide
+simulation-based security against a semi-honest adversary, in the
+framework of ~\cite{gmw,can00,can01}. At a very high level, in this
+framework, parties are modeled as non-uniform interactive turing
+machines (ITMs), with inputs provided by an environment $\env$. An
+adversary $\adv$, selects and ``corrupts'' one of the parties -
+however, $\adv$ still follows the protocol specification. $\adv$
+interacts with the $\env$, which observes the view of the corrupted
+party. At the end of the interaction, $\env$ outputs a single bit. Two
+different interactions are defined: the {\em real world} and an {\em
+  ideal world}. In the real interaction, the parties run the protocol
+$\prot$ in the presence of $\adv$ and $\env$. Let
+$\real_{\prot,\adv,\env}$ denote the binary distribution ensemble
+describing $\env$'s output in this interaction. In the ideal
+interaction, parties send their inputs to an additional entity, a
+trusted functionality machine $\F$ that carries the desired
+computation truthfully. Let $\simu$ (the simulator) denote the
+adversary in this idealized execution, and $\ideal_{\F,\simu,\env}$
+the binary distribution ensemble describing $\env$'s output after
+interacting with adversary $\simu$ and ideal functionality $\F$. A
+protocol $\prot$ is said to {\em securely realize} a functionality
+$\F$ if for every adversary $\adv$ in the real interaction, there is
+an adversary $\simu$ in the ideal interaction, such that no
+environment $\env$, on any input, can tell the real interaction apart
+from the ideal interaction, except with negligible probability (in the
+security parameter $\secparam$). More precisely, if the two binary
+distribution ensembles above are computationally indistinguishable.
 \nc{I'm not happy with the theorem statement.}
@@
 \nc{We should probably connect things to earlier? Theorem 1 and 2 have very different languages.}
-\noindent {\em Proof.} From Theorem \ref{theorem:correctness}, we have that the observations in $\prot$ and $\F$ are the same (i.e.
-$O_2 = O_1$).
-Now, from the simulation security of our backend (\cite{aby}), the proof of the above theorem follows.
-\nc{I dont know how to say more.}
+\noindent {\em Proof.} From Theorem \ref{theorem:correctness}, we have
+that the observations in $\prot$ and $\F$ are the same (i.e. $O_2 =
+O_1$). Now, from the simulation security of our backend (\cite{aby}),
+the proof of the above theorem follows. \nc{I dont know how to say
+  more.}
+
+\aseem{Can we give a construction of the simulator S in the ideal
+  semantics? The simulator S whenever it receives an observable event,
+  outputs it to Z, and whenever A receives a string in the real
+  semantics, S outputs a random string to Z. Now, with theorem 1,
+  observable events are same in the ideal and real semantics, and with
+  the security of the crypto backend, the random strings and the
+  strings in the protocol are indistinguishable. Is it
+  incorrect/naive?
+
+  Also, we should emaphasize here that the ideal semantics is the
+  source semantics from Figure x and the real semantics is the circuit
+  semantics from Figure y. For the target language semantics, we
+  should emphasize that it does not have access to the secrets, so
+  there are no observations there.
+}
+
   %% \\

### 2017-10-24 examples/factor64.cpp
factor64

@@ -1,0 +1,2235 @@
+bitlen = 64;
+ABYParty *party = new ABYParty(role, address, port, seclvl, bitlen, nthreads, mt_alg, 120000000);
+vector<Sharing*>& sharings = party->GetSharings();
+Circuit* ycirc = sharings[S_YAO]->GetCircuitBuildRoutine();
+Circuit* acirc = sharings[S_ARITH]->GetCircuitBuildRoutine();
+Circuit* bcirc = sharings[S_BOOL]->GetCircuitBuildRoutine();
+auto S1 = make_vector<uint64_t>(16384, 24);
+auto s_a_S1 = make_vector<share*>(16384, 24);
+auto S2 = make_vector<uint64_t>(16384, 24);
+auto s_a_S2 = make_vector<share*>(16384, 24);
+auto S = make_vector<uint64_t>(16384, 24);
+auto s_a_S = make_vector<share*>(16384, 24);
+auto outmat = make_vector<uint64_t>(16384, 24);
+auto s_a_outmat = make_vector<share*>(16384, 24);
+auto boolarr = make_vector<uint64_t>(1000000);
+auto s_a_boolarr = make_vector<share*>(1000000);
+auto temp = make_vector<uint64_t>(10);
+auto s_a_temp = make_vector<share*>(10);
+auto dots = make_vector<uint64_t>(15444);
+auto s_a_dots = make_vector<share*>(15444);
+uint64_t boolidx;
+uint64_t b;
+share *s_a_b;
+uint64_t b1;
+share *s_a_b1;
+boolidx = 0;
+for (uint64_t i = 0; i < 16384; i++)
+{
+    for (uint64_t j = 0; j < 24; j++)
+    {
+        if (role == SERVER) {
+            S1[i][j] = 50 ;
+            s_a_S1[i][j] = acirc->PutINGate( S1[i][j] ,bitlen,SERVER);
+        } else {
+            s_a_S1[i][j] = acirc->PutDummyINGate(bitlen);
+        }
+    }
+}
+for (uint64_t i = 0; i < 1000000; i++)
+{
+    if (role == SERVER) {
+        boolarr[i] = 1 ;
+        s_a_boolarr[i] = acirc->PutINGate( boolarr[i] ,bitlen,SERVER);
[…]

### 2017-10-24 examples/factor/1_pipe_arith_cifar.cpp
factor partition

@@ -1,0 +1,823 @@
+bitlen = 64;
+ABYParty *party = new ABYParty(role, address, port, seclvl, bitlen, nthreads, mt_alg, 120000000);
+vector<Sharing*>& sharings = party->GetSharings();
+Circuit* ycirc = sharings[S_YAO]->GetCircuitBuildRoutine();
+Circuit* acirc = sharings[S_ARITH]->GetCircuitBuildRoutine();
+Circuit* bcirc = sharings[S_BOOL]->GetCircuitBuildRoutine();
+auto S1 = make_vector<uint64_t>(16384, 24);
+auto s_a_S1 = make_vector<share*>(16384, 24);
+auto S2 = make_vector<uint64_t>(16384, 24);
+auto s_a_S2 = make_vector<share*>(16384, 24);
+auto S = make_vector<uint64_t>(16384, 24);
+auto s_a_S = make_vector<share*>(16384, 24);
+auto outmat = make_vector<uint64_t>(16384, 24);
+auto s_a_outmat = make_vector<share*>(16384, 24);
+auto boolarr = make_vector<uint64_t>(1000000);
+auto s_a_boolarr = make_vector<share*>(1000000);
+auto temp = make_vector<uint64_t>(10);
+auto s_a_temp = make_vector<share*>(10);
+auto dots = make_vector<uint64_t>(15444);
+auto s_a_dots = make_vector<share*>(15444);
+uint64_t boolidx;
+auto Routmat = make_vector<uint64_t>(16384, 24);
+auto s_a_Routmat = make_vector<share*>(16384, 24);
+auto IRoutmat = make_vector<uint64_t>(16384, 24);
+auto s_a_IRoutmat = make_vector<share*>(16384, 24);
+uint64_t b;
+share *s_a_b;
+uint64_t b1;
+share *s_a_b1;
+boolidx = 0;
+if(role == SERVER){
+ofstream fout("IRoutmat", std::ofstream::out | std::ofstream::app);
+for(int i = 0; i < 16384; i++) {
+	for(int j = 0; j < 24; j++) {
+		fout << i << " " << j << " " <<(IRoutmat[i][j] = rand()) << endl;
+	}
+}
+fout.close();
+}
+for (uint64_t i = 0; i < 16384; i++)
+{
[…]

### 2017-10-24 examples/factor/1_pipe_arith_cifar_debug.cpp

@@ -1,0 +1,1059 @@
+#include <iostream>
+#include <vector>
+using namespace std;
+namespace details {
+template<class T>struct tag{using type=T;};
+template<class Tag>using type=typename Tag::type;
+template<class T, size_t n>
+struct n_dim_vec:tag< std::vector< type< n_dim_vec<T, n-1> > > > {};
+template<class T>
+struct n_dim_vec<T, 0>:tag<T>{};
+template<class T, size_t n>
+using n_dim_vec_t = type<n_dim_vec<T,n>>;
+template <class T, class R=std::vector<T>>
+R make_vector(size_t size) {
+return R(size);
+}
+template<class T, class...Args, class R=n_dim_vec_t<T, sizeof...(Args)+1>>
+R make_vector(size_t top, Args...args){
+return R(top, make_vector<T>(std::forward<Args>(args)...));
+}
+}
+template <class T, class... Args, class R=details::n_dim_vec_t<T, sizeof...(Args)>>
+R make_vector(Args... args)
+{ return details::make_vector<T>(std::forward<Args>(args)...); }
+int main() {
+auto S1 = make_vector<uint32_t>(16384, 24);
+auto S2 = make_vector<uint32_t>(16384, 24);
+auto S = make_vector<uint32_t>(16384, 24);
+auto outmat = make_vector<uint32_t>(16384, 24);
+auto boolarr = make_vector<uint32_t>(1000000);
+auto temp = make_vector<uint32_t>(10);
+auto dots = make_vector<uint32_t>(15444);
+public uint32_t ;
+auto Routmat = make_vector<uint32_t>(16384, 24);
+auto IRoutmat = make_vector<uint32_t>(16384, 24);
+uint32_t b ;
+uint32_t b1 ;
+boolidx = 0;
+for (uint32_t i = 0; i < 16384; i++)
+{
+for (uint32_t j = 0; j < 24; j++)
+{
[…]

### 2017-10-24 examples/factor/2_pipe_arith_cifar.cpp

@@ -1,0 +1,888 @@
+bitlen = 64;
+ABYParty *party = new ABYParty(role, address, port, seclvl, bitlen, nthreads, mt_alg, 120000000);
+vector<Sharing*>& sharings = party->GetSharings();
+Circuit* ycirc = sharings[S_YAO]->GetCircuitBuildRoutine();
+Circuit* acirc = sharings[S_ARITH]->GetCircuitBuildRoutine();
+Circuit* bcirc = sharings[S_BOOL]->GetCircuitBuildRoutine();
+auto outmat = make_vector<uint64_t>(16384, 24);
+auto s_a_outmat = make_vector<share*>(16384, 24);
+auto outmp = make_vector<uint64_t>(16384, 24);
+auto s_a_outmp = make_vector<share*>(16384, 24);
+auto boolarr = make_vector<uint64_t>(1000000);
+auto s_a_boolarr = make_vector<share*>(1000000);
+auto temp = make_vector<uint64_t>(10);
+auto s_a_temp = make_vector<share*>(10);
+auto dots = make_vector<uint64_t>(15444);
+auto s_a_dots = make_vector<share*>(15444);
+uint64_t boolidx;
+auto Routmat = make_vector<uint64_t>(16384, 24);
+auto s_a_Routmat = make_vector<share*>(16384, 24);
+auto IRoutmat = make_vector<uint64_t>(16384, 24);
+auto s_a_IRoutmat = make_vector<share*>(16384, 24);
+auto Routmp = make_vector<uint64_t>(16384, 24);
+auto s_a_Routmp = make_vector<share*>(16384, 24);
+auto IRoutmp = make_vector<uint64_t>(16384, 24);
+auto s_a_IRoutmp = make_vector<share*>(16384, 24);
+uint64_t b;
+share *s_a_b;
+uint64_t b1;
+share *s_a_b1;
+boolidx = 0;
+uint64_t t;
+if(role == SERVER){
+ifstream fin("IRoutmat");
+while(!fin.eof()){
+	uint64_t i, j, val;
+	fin >> i >> j >> val;
+	IRoutmat[i][j] = val;
+}
+fin.close();
+}
+if(role == CLIENT){
+ifstream fin("outmat");
+while(!fin.eof()){
+	uint64_t i, j, val;
[…]

### 2017-10-24 examples/factor/2_pipe_arith_cifar_debug.cpp

@@ -1,0 +1,1129 @@
+#include <iostream>
+#include <vector>
+using namespace std;
+namespace details {
+template<class T>struct tag{using type=T;};
+template<class Tag>using type=typename Tag::type;
+template<class T, size_t n>
+struct n_dim_vec:tag< std::vector< type< n_dim_vec<T, n-1> > > > {};
+template<class T>
+struct n_dim_vec<T, 0>:tag<T>{};
+template<class T, size_t n>
+using n_dim_vec_t = type<n_dim_vec<T,n>>;
+template <class T, class R=std::vector<T>>
+R make_vector(size_t size) {
+return R(size);
+}
+template<class T, class...Args, class R=n_dim_vec_t<T, sizeof...(Args)+1>>
+R make_vector(size_t top, Args...args){
+return R(top, make_vector<T>(std::forward<Args>(args)...));
+}
+}
+template <class T, class... Args, class R=details::n_dim_vec_t<T, sizeof...(Args)>>
+R make_vector(Args... args)
+{ return details::make_vector<T>(std::forward<Args>(args)...); }
+int main() {
+auto outmat = make_vector<uint32_t>(16384, 24);
+auto outmp = make_vector<uint32_t>(16384, 24);
+auto boolarr = make_vector<uint32_t>(1000000);
+auto temp = make_vector<uint32_t>(10);
+auto dots = make_vector<uint32_t>(15444);
+public uint32_t ;
+auto Routmat = make_vector<uint32_t>(16384, 24);
+auto IRoutmat = make_vector<uint32_t>(16384, 24);
+auto Routmp = make_vector<uint32_t>(16384, 24);
+auto IRoutmp = make_vector<uint32_t>(16384, 24);
+uint32_t b ;
+uint32_t b1 ;
+boolidx = 0;
+public uint32_t ;
+for (uint32_t i = 0; i < 1000000; i++)
+{
+boolarr[i] = 1;
[…]

### 2017-10-24 examples/factor/3_pipe_arith_cifar.cpp

@@ -1,0 +1,792 @@
+bitlen = 64;
+ABYParty *party = new ABYParty(role, address, port, seclvl, bitlen, nthreads, mt_alg, 120000000);
+vector<Sharing*>& sharings = party->GetSharings();
+Circuit* ycirc = sharings[S_YAO]->GetCircuitBuildRoutine();
+Circuit* acirc = sharings[S_ARITH]->GetCircuitBuildRoutine();
+Circuit* bcirc = sharings[S_BOOL]->GetCircuitBuildRoutine();
+auto outmp = make_vector<uint64_t>(16384, 24);
+auto s_a_outmp = make_vector<share*>(16384, 24);
+auto boolarr = make_vector<uint64_t>(1000000);
+auto s_a_boolarr = make_vector<share*>(1000000);
+auto temp = make_vector<uint64_t>(10);
+auto s_a_temp = make_vector<share*>(10);
+auto dots = make_vector<uint64_t>(15444);
+auto s_a_dots = make_vector<share*>(15444);
+uint64_t boolidx;
+auto Routmp = make_vector<uint64_t>(16384, 24);
+auto s_a_Routmp = make_vector<share*>(16384, 24);
+auto IRoutmp = make_vector<uint64_t>(16384, 24);
+auto s_a_IRoutmp = make_vector<share*>(16384, 24);
+uint64_t b;
+share *s_a_b;
+uint64_t b1;
+share *s_a_b1;
+boolidx = 0;
+uint64_t t;
+if(role == SERVER){
+ifstream fin("IRoutmp");
+while(!fin.eof()){
+	uint64_t i, j, val;
+	fin >> i >> j >> val;
+	IRoutmp[i][j] = val;
+}
+fin.close();
+}
+if(role == CLIENT){
+ifstream fin("outmp");
+while(!fin.eof()){
+	uint64_t i, j, val;
+	fin >> i >> j >> val;
+	outmp[i][j] = val;
+	s_a_outmp[i][j]= acirc->PutINGate(outmp[i][j] ,bitlen,CLIENT);
+	
+}
+fin.close();
+}
+else {
+ for(int i = 0; i <16384;i++) {
+	for (int j = 0; j <24; j++) {
+		s_a_outmp[i][j] = acirc->PutDummyINGate(bitlen);
+	}
+}
+}
[…]

### 2017-10-24 examples/factor/3_pipe_arith_cifar_debug.cpp

@@ -1,0 +1,1033 @@
+#include <iostream>
+#include <vector>
+using namespace std;
+namespace details {
+template<class T>struct tag{using type=T;};
+template<class Tag>using type=typename Tag::type;
+template<class T, size_t n>
+struct n_dim_vec:tag< std::vector< type< n_dim_vec<T, n-1> > > > {};
+template<class T>
+struct n_dim_vec<T, 0>:tag<T>{};
+template<class T, size_t n>
+using n_dim_vec_t = type<n_dim_vec<T,n>>;
+template <class T, class R=std::vector<T>>
+R make_vector(size_t size) {
+return R(size);
+}
+template<class T, class...Args, class R=n_dim_vec_t<T, sizeof...(Args)+1>>
+R make_vector(size_t top, Args...args){
+return R(top, make_vector<T>(std::forward<Args>(args)...));
+}
+}
+template <class T, class... Args, class R=details::n_dim_vec_t<T, sizeof...(Args)>>
+R make_vector(Args... args)
+{ return details::make_vector<T>(std::forward<Args>(args)...); }
+int main() {
+auto outmp = make_vector<uint32_t>(16384, 24);
+auto boolarr = make_vector<uint32_t>(1000000);
+auto temp = make_vector<uint32_t>(10);
+auto dots = make_vector<uint32_t>(15444);
+public uint32_t ;
+auto Routmp = make_vector<uint32_t>(16384, 24);
+auto IRoutmp = make_vector<uint32_t>(16384, 24);
+uint32_t b ;
+uint32_t b1 ;
+boolidx = 0;
+public uint32_t ;
+for (uint32_t i = 0; i < 1000000; i++)
+{
+boolarr[i] = 1;
+}
+for (uint32_t i = 0; i < 16384; i++)
+{
+for (uint32_t j = 0; j < 24; j++)
+{
+Routmp[i][j] = IRoutmp[i][j];
+}
+}
[…]

### 2017-10-24 examples/factor/1_pipe_arith_cifar.cpp
reverting to 32

@@ -1,31 +1,30 @@
-bitlen = 64;
 ABYParty *party = new ABYParty(role, address, port, seclvl, bitlen, nthreads, mt_alg, 120000000);
 vector<Sharing*>& sharings = party->GetSharings();
 Circuit* ycirc = sharings[S_YAO]->GetCircuitBuildRoutine();
 Circuit* acirc = sharings[S_ARITH]->GetCircuitBuildRoutine();
 Circuit* bcirc = sharings[S_BOOL]->GetCircuitBuildRoutine();
-auto S1 = make_vector<uint64_t>(16384, 24);
+auto S1 = make_vector<uint32_t>(16384, 24);
 auto s_a_S1 = make_vector<share*>(16384, 24);
-auto S2 = make_vector<uint64_t>(16384, 24);
+auto S2 = make_vector<uint32_t>(16384, 24);
 auto s_a_S2 = make_vector<share*>(16384, 24);
-auto S = make_vector<uint64_t>(16384, 24);
+auto S = make_vector<uint32_t>(16384, 24);
 auto s_a_S = make_vector<share*>(16384, 24);
-auto outmat = make_vector<uint64_t>(16384, 24);
+auto outmat = make_vector<uint32_t>(16384, 24);
 auto s_a_outmat = make_vector<share*>(16384, 24);
-auto boolarr = make_vector<uint64_t>(1000000);
+auto boolarr = make_vector<uint32_t>(1000000);
 auto s_a_boolarr = make_vector<share*>(1000000);
-auto temp = make_vector<uint64_t>(10);
+auto temp = make_vector<uint32_t>(10);
 auto s_a_temp = make_vector<share*>(10);
-auto dots = make_vector<uint64_t>(15444);
+auto dots = make_vector<uint32_t>(15444);
 auto s_a_dots = make_vector<share*>(15444);
-uint64_t boolidx;
-auto Routmat = make_vector<uint64_t>(16384, 24);
+uint32_t boolidx;
+auto Routmat = make_vector<uint32_t>(16384, 24);
 auto s_a_Routmat = make_vector<share*>(16384, 24);
-auto IRoutmat = make_vector<uint64_t>(16384, 24);
+auto IRoutmat = make_vector<uint32_t>(16384, 24);
[…]

### 2017-10-24 examples/factor/2_pipe_arith_cifar.cpp

@@ -1,38 +1,37 @@
-bitlen = 64;
 ABYParty *party = new ABYParty(role, address, port, seclvl, bitlen, nthreads, mt_alg, 120000000);
 vector<Sharing*>& sharings = party->GetSharings();
 Circuit* ycirc = sharings[S_YAO]->GetCircuitBuildRoutine();
 Circuit* acirc = sharings[S_ARITH]->GetCircuitBuildRoutine();
 Circuit* bcirc = sharings[S_BOOL]->GetCircuitBuildRoutine();
-auto outmat = make_vector<uint64_t>(16384, 24);
+auto outmat = make_vector<uint32_t>(16384, 24);
 auto s_a_outmat = make_vector<share*>(16384, 24);
-auto outmp = make_vector<uint64_t>(16384, 24);
+auto outmp = make_vector<uint32_t>(16384, 24);
 auto s_a_outmp = make_vector<share*>(16384, 24);
-auto boolarr = make_vector<uint64_t>(1000000);
+auto boolarr = make_vector<uint32_t>(1000000);
 auto s_a_boolarr = make_vector<share*>(1000000);
-auto temp = make_vector<uint64_t>(10);
+auto temp = make_vector<uint32_t>(10);
 auto s_a_temp = make_vector<share*>(10);
-auto dots = make_vector<uint64_t>(15444);
+auto dots = make_vector<uint32_t>(15444);
 auto s_a_dots = make_vector<share*>(15444);
-uint64_t boolidx;
-auto Routmat = make_vector<uint64_t>(16384, 24);
+uint32_t boolidx;
+auto Routmat = make_vector<uint32_t>(16384, 24);
 auto s_a_Routmat = make_vector<share*>(16384, 24);
-auto IRoutmat = make_vector<uint64_t>(16384, 24);
+auto IRoutmat = make_vector<uint32_t>(16384, 24);
 auto s_a_IRoutmat = make_vector<share*>(16384, 24);
-auto Routmp = make_vector<uint64_t>(16384, 24);
+auto Routmp = make_vector<uint32_t>(16384, 24);
 auto s_a_Routmp = make_vector<share*>(16384, 24);
-auto IRoutmp = make_vector<uint64_t>(16384, 24);
[…]

### 2017-10-24 examples/factor/3_pipe_arith_cifar.cpp

@@ -1,32 +1,31 @@
-bitlen = 64;
 ABYParty *party = new ABYParty(role, address, port, seclvl, bitlen, nthreads, mt_alg, 120000000);
 vector<Sharing*>& sharings = party->GetSharings();
 Circuit* ycirc = sharings[S_YAO]->GetCircuitBuildRoutine();
 Circuit* acirc = sharings[S_ARITH]->GetCircuitBuildRoutine();
 Circuit* bcirc = sharings[S_BOOL]->GetCircuitBuildRoutine();
-auto outmp = make_vector<uint64_t>(16384, 24);
+auto outmp = make_vector<uint32_t>(16384, 24);
 auto s_a_outmp = make_vector<share*>(16384, 24);
-auto boolarr = make_vector<uint64_t>(1000000);
+auto boolarr = make_vector<uint32_t>(1000000);
 auto s_a_boolarr = make_vector<share*>(1000000);
-auto temp = make_vector<uint64_t>(10);
+auto temp = make_vector<uint32_t>(10);
 auto s_a_temp = make_vector<share*>(10);
-auto dots = make_vector<uint64_t>(15444);
+auto dots = make_vector<uint32_t>(15444);
 auto s_a_dots = make_vector<share*>(15444);
-uint64_t boolidx;
-auto Routmp = make_vector<uint64_t>(16384, 24);
+uint32_t boolidx;
+auto Routmp = make_vector<uint32_t>(16384, 24);
 auto s_a_Routmp = make_vector<share*>(16384, 24);
-auto IRoutmp = make_vector<uint64_t>(16384, 24);
+auto IRoutmp = make_vector<uint32_t>(16384, 24);
 auto s_a_IRoutmp = make_vector<share*>(16384, 24);
-uint64_t b;
+uint32_t b;
 share *s_a_b;
-uint64_t b1;
+uint32_t b1;
 share *s_a_b1;
 boolidx = 0;
-uint64_t t;
+uint32_t t;
 if(role == SERVER){
 ifstream fin("IRoutmp");
 while(!fin.eof()){
-	uint64_t i, j, val;
+	uint32_t i, j, val;
 	fin >> i >> j >> val;
 	IRoutmp[i][j] = val;
 }
@@ -35,7 +34,7 @@ fin.close();
 if(role == CLIENT){
 ifstream fin("outmp");
[…]

### 2017-10-24 OaklandSubmission/intro.tex
intro first part

@@
 Today it is hard for developers to create secure implementations using cryptographic techniques.
-Typical developers lack an understanding of cryptography protocols
+Typical developers lack an understanding of cryptographic protocols
 and cannot be expected to implement them correctly and efficiently without tool
@@
 
-Secure multi-party computation (\mpc) \cite{yao,gmw} is a powerful
-cryptographic tool that allows a set of mutually distrusting parties
-to compute a joint function of their secret inputs. Examples of these
-include {\em secure prediction} \cite{shafindss,wu,barni,minionn,secureml}. Here,  one party holds the propriety classifier for disease, genomics or spam detection and the other party  
-holds the private input and the second party learns the classification label. Since the introduction of
+Secure two-party computation (\mpc) \cite{yao,gmw} is a powerful
+cryptographic tool that allows two mutually distrusting parties
+to compute a joint function of their secret inputs such as {\em secure prediction} \cite{shafindss,wu,barni,minionn,secureml}.
+Here,  one party holds the propriety classifier for disease, genomics or spam detection and the other party
+holds the private input and the second party learns the classification label such as the presence of the disease. Since its introduction
 \mpc in 1980's, there has been a long sequence of work that has
@@
 Suppose a developer wants to compute $w^Tx >b$ securely using
-\mpc. With the current techniques, he has three options. \aseem{We should
+\mpc, where $w$ and $x$ are two vectors, $b$ is a scalar, and $w^Tx$ denotes the inner product of $w$ and $x$.  With the current techniques, he has three options. \aseem{We should
   use she for the programmer instead of he?}
@@
   domain-specific language such as
-  \cite{lambdaps,wysteria,oblivm,smcl} that would automatically
-  compile it to an SMC protocol. The downside is that all of these
-  works build on a cryptographic back-end that is either entirely
-  boolean \cite{yao,gmw} or entirely arithmetic \cite{homo}.
-The efficiency of the generated \mpc protocol is hence, bounded by the possibility of efficiently representing the function in a form that is compatible with the cryptographic back-end used.
-For instance, over $\ell$-bit integers, multiplication requires a boolean circuit of size  $O(\ell^2)$.
-For better efficiency, we would like to evaluate $w^Tx$ using an arithmetic circuit and the comparison using a boolean circuit.
+  \cite{lambdaps,wysteria,oblivm,smcl,cbmcgc} that would automatically
+  compile it to a \mpc protocol. The downside is that all of these
+  works build on a cryptographic back-end that takes as input a functionality expressed as either a boolean circuit \cite{yao,gmw} or an arithmetic circuit \cite{homo}.
+  The efficiency of the generated \mpc protocol is hence, bounded by the possibility of efficiently representing the functionality in one of these circuit representations.
+For instance, multiplication of two $\ell$-bit integers can be expressed as a boolean circuit of size  $O(\ell^2)$ or an arithmetic circuit with 1 multiplication gate.
+Hence, for better efficiency, we would like to securely compute $w^Tx$ using an arithmetic circuit and the comparison with $b$ using a boolean circuit. \divya{that is not supported by above systems.}
 In fact, most interesting functions require a mix of arithmetic and boolean computations for efficiency.
 
-\item To address this issue of efficient representation, recently Demmler et al. \cite{aby} gave a cryptographic protocol for \mpc where the parties can mix arithmetic and boolean computations.
+\item To address this issue of efficient representation, recently Demmler et al.
+\cite{aby} gave a cryptographic protocol for \mpc where the functionality can be expressed as a circuit with a mix of both arithmetic and boolean parts.
 Hence, the second option is to write the functionality in the system developed in \cite{aby}.
-However, using their system requires the programmer to be aware of
+However, in their framework, the programmer is required to write circuits consisting of  arithmetic and boolean gates along with appropriate
+inter-conversion gates.
+This requires the programmer to be aware of
 trade-offs between arithmetic and boolean cryptographic
-schemes as well as the complex inter-conversion costs, and manually split the functionality into such parts accordingly. In
-their framework, the programmer is required to write circuits
-consisting of  arithmetic and boolean gates along with appropriate
-conversion gates. In short, as is also mentioned by the authors, the current system of ABY is not suitable
-to be used by non-specialist programmers. \divya{since they say in future they want to automate the splitting.}
+schemes as well as the complex inter-conversion costs between arithmetic and boolean, and manually split the functionality into such parts accordingly.  
+In short, as is also mentioned by Demmler et al., the current system  is not suitable
+to be used by \divya{non-cryptography expert developers. }
 

### 2017-10-24 examples/factor/2_pipe_arith_cifar.cpp
revert to 32

@@ -7,8 +7,8 @@ auto outmat = make_vector<uint32_t>(16384, 24);
 auto s_a_outmat = make_vector<share*>(16384, 24);
 auto outmp = make_vector<uint32_t>(16384, 24);
 auto s_a_outmp = make_vector<share*>(16384, 24);
-auto boolarr = make_vector<uint32_t>(1000000);
-auto s_a_boolarr = make_vector<share*>(1000000);
+auto boolarr = make_vector<uint32_t>(200000);
+auto s_a_boolarr = make_vector<share*>(200000);
 auto temp = make_vector<uint32_t>(10);
 auto s_a_temp = make_vector<share*>(10);
 auto dots = make_vector<uint32_t>(15444);
@@ -28,6 +28,8 @@ uint32_t b1;
 share *s_a_b1;
 boolidx = 0;
 uint32_t t;
+uint32_t cas;
+share *s_a_cas;
 if(role == SERVER){
 ifstream fin("IRoutmat");
 while(!fin.eof()){
@@ -64,7 +66,7 @@ for(int i = 0; i < 16384; i++) {
 }
 fout.close();
 }
-for (uint32_t i = 0; i < 1000000; i++)
+for (uint32_t i = 0; i < 200000; i++)
 {
     if (role == SERVER) {
         boolarr[i] = 1 ;
@@ -113,10 +115,11 @@ for (uint32_t i = 0; i < 8192; i++)
     boolidx = boolidx+1;
     for (uint32_t j = 0; j < 24; j++)
     {
+        s_a_cas = create_new_share(s_a_outmat[2*i][j]->get_wires(), acirc );
         share * s_a_tmp_1 = acirc->PutMULGate( s_a_b , s_a_outmat[2*i][j] );
         share * s_a_tmp_2 = acirc->PutMULGate( s_a_b1 , s_a_outmat[2*i+1][j] );
         s_a_outmat[2*i][j] = acirc->PutADDGate( s_a_tmp_1 , s_a_tmp_2 );
-        share * s_a_tmp_3 = acirc->PutMULGate( s_a_b1 , s_a_outmat[2*i][j] );
+        share * s_a_tmp_3 = acirc->PutMULGate( s_a_b1 , s_a_cas );
         share * s_a_tmp_4 = acirc->PutMULGate( s_a_b , s_a_outmat[2*i+1][j] );
[…]

### 2017-10-24 examples/factor/3_pipe_arith_cifar.cpp

@@ -5,8 +5,8 @@ Circuit* acirc = sharings[S_ARITH]->GetCircuitBuildRoutine();
 Circuit* bcirc = sharings[S_BOOL]->GetCircuitBuildRoutine();
 auto outmp = make_vector<uint32_t>(16384, 24);
 auto s_a_outmp = make_vector<share*>(16384, 24);
-auto boolarr = make_vector<uint32_t>(1000000);
-auto s_a_boolarr = make_vector<share*>(1000000);
+auto boolarr = make_vector<uint32_t>(200000);
+auto s_a_boolarr = make_vector<share*>(200000);
 auto temp = make_vector<uint32_t>(10);
 auto s_a_temp = make_vector<share*>(10);
 auto dots = make_vector<uint32_t>(15444);
@@ -22,6 +22,8 @@ uint32_t b1;
 share *s_a_b1;
 boolidx = 0;
 uint32_t t;
+uint32_t cas;
+share *s_a_cas;
 if(role == SERVER){
 ifstream fin("IRoutmp");
 while(!fin.eof()){
@@ -49,7 +51,7 @@ else {
 	}
 }
 }
-for (uint32_t i = 0; i < 1000000; i++)
+for (uint32_t i = 0; i < 200000; i++)
 {
     if (role == SERVER) {
         boolarr[i] = 1 ;
@@ -86,10 +88,11 @@ for (uint32_t i = 0; i < 8192; i++)
     boolidx = boolidx+1;
     for (uint32_t j = 0; j < 24; j++)
     {
+        s_a_cas = create_new_share(s_a_outmp[2*i][j]->get_wires(), acirc );
         share * s_a_tmp_1 = acirc->PutMULGate( s_a_b , s_a_outmp[2*i][j] );
         share * s_a_tmp_2 = acirc->PutMULGate( s_a_b1 , s_a_outmp[2*i+1][j] );
         s_a_outmp[2*i][j] = acirc->PutADDGate( s_a_tmp_1 , s_a_tmp_2 );
-        share * s_a_tmp_3 = acirc->PutMULGate( s_a_b1 , s_a_outmp[2*i][j] );
+        share * s_a_tmp_3 = acirc->PutMULGate( s_a_b1 , s_a_cas );
         share * s_a_tmp_4 = acirc->PutMULGate( s_a_b , s_a_outmp[2*i+1][j] );
[…]

### 2017-10-24 OaklandSubmission/intro.tex

@@
 In short, as is also mentioned by Demmler et al., the current system  is not suitable
-to be used by \divya{non-cryptography expert developers. }
+to be used by non-specialist programmers without security expertise.
 
@@
 computation. We provide the first compiler where the developer writes a high-level program and the compiler automatically generates an efficient \mpc protocol that uses a
-mix of Arithmetic and Boolean compute. 
-It is compatible with any implementation of \mpc that allows a mix of arithmetic and boolean computations. 
+mix of arithmetic and boolean circuits. 
+It is compatible with any implementation of \mpc that allows a mix of arithmetic and boolean circuits. 
 We focus on semi-honest secure two-party computation and use ABY \cite{aby} as the suitable cryptographic back-end.
 We provide a formal type system and operational semantics for our language and prove correctness and security of our compiler.
-We evaluate our system on wide range of benchmarks \cite{shafindss,wu,minionn,valeriaMatrix} computing linear classifier, decision trees, deep neural networks, and matrix factorization.
+We evaluate our system on wide range of benchmarks for secure prediction \cite{shafindss,wu,barni,secureml,minionn,bonsai} using linear classifier, decision trees and deep neural networks, matrix factorization \cite{valeriaMatrix} and training of deep neural network \cite{secureml}.
 Our evaluations show that protocols generated by our compiler match
 the performance or outperform hand-crafted protocols in most cases (see
-\sectionref{eval}). \divya{Moreover, unlike previous systems, our system can scale to arbitrarily large computations using {\em pipelining} (see below). What do we say about Yao pipelining?}
-Hence, we show that our framework achieves all four properties of generality, performance, programmer productivity and scalability.
+\sectionref{eval}).
+\TODO{Add some text to increase enthusiasm about results.} Moreover, unlike previous systems such as \cite{wysteria,aby}, our system can scale to arbitrarily large computations using {\em secure code partitioning} (see below).
+\divya{I think we can skip Yao pipelining here, right?}
+Hence, we show that our framework achieves all four properties of programmability, generality, performance, and scalability.
 

### 2017-10-24 OaklandSubmission/intro.tex
sec 1.1

@@
   \divya{Do we write the intro for \mpc or 2PC?}
+
+  \TODO{While taking a pass on intro, check citation correctness or if we are missing any citation.}
 
@@
 \sectionref{eval}).
-\TODO{Add some text to increase enthusiasm about results.} Moreover, unlike previous systems such as \cite{wysteria,aby}, our system can scale to arbitrarily large computations using {\em secure code partitioning} (see below).
+\TODO{Add some text to increase enthusiasm about results.} Moreover, unlike previous systems such as \cite{wysteria,aby,cbmcgc}, our system can scale to arbitrarily large computations using our technique of {\em secure code partitioning} (see below).
 \divya{I think we can skip Yao pipelining here, right?}
-Hence, we show that our framework achieves all four properties of programmability, generality, performance, and scalability.
+Hence, we show that our framework achieves all four properties of programmability, generality, performance, and scalability (see below for details).
 
-Prior to our work, there have been systems that provide only a subset of the above properties and lack in remaining properties.
-For instance, using the framework of CBMC-GC \cite{cbmcgc}, one can write functionalities in high-level language like C but it gives poor performance due to its usage of boolean circuits as underlying representation.
-Similarly, ABY framework can potentially give good performance but is not programmable.
-Technique of pipelining Yao garbled circuits \cite{yao-pipe} allowed scaling to lage circuits but gives poor performance.
-\divya{what about programming?} To summarize, none of the prior frameworks could be used to evaluate the benchmarks such as linear regression, deep neural network prediction etc.
+Prior to our work, there have been systems that provide only a subset of the above properties.
+For instance, using the framework of CBMC-GC \cite{cbmcgc} allows  functionalities to be expressed in high-level language C but it gives poor performance due to its usage of boolean circuits as underlying representation.
+Similarly, ABY \cite{aby} framework can potentially give good performance (due to mix of boolean and arithmetic circuits) but is not programmable.
+Yao's garbled circuits \cite{yao} work with boolean circuits and are the most widely used 2PC protocols but they  do not scale to large functions due to large memory requirement.
+Huang et al.
+\cite{yao-pipe} gave the technique of pipelining Yao's garbled circuits that enabled scaling to large functions but still gives poor performance due to its use of boolean circuit representation.
+To summarize, none of the prior frameworks have been used to evaluate the benchmarks such as secure prediction using linear regression, decision trees, deep neural networks, etc.
 In fact, it was widely believed that generic 2PC would not scale to the secure prediction and hence, various protocols were hand-crafted for this task \cite{shafindss,wu,secureml,minionn}.
-Our framework contradicts this belief.
+Our framework contradicts this belief (see below).
 
-In this work, we solve an interesting design problem to give a user friendly language that is expressive enough together with a compiler generating efficient secure 2PC protocols.
-Programmability and performance of our system has enabled us to write large programs whose circuits  exhaust the memory of the system (~32 GB).
-To combat this issue, we give a technique called pipelining (see below) that allows scaling of our  scaling of our framework to arbitrary size computations.
+In this work, we solve the design problem to provide a user friendly language that is expressive and can be compiled to efficient secure 2PC protocols.
+Programmability in \tool and performance of generated protocols has enabled us to write large programs whose circuits  exhaust the memory of the system (~32 GB).
+To combat this issue, we give a technique called secure code partitioning (see below) that allows scaling of \tool to arbitrary size computations.
 
@@
 express the function to be computed and uses a mix of arithmetic and
-boolean \mpc back-ends. We illustrate the above toy example of $w^Tx
->b$ in our language in \figureref{ex-sml}. As is clear from the code,
-the programmer is not required to write keywords such as {\em secret}
-or {\em public} and can remain completely oblivious of the
-cryptographic back-end. In contrast, \figureref{ex-aby} illustrates
+boolean \mpc back-ends. We illustrate the above example of $w^Tx
+>b$ in our language in \figureref{ex-sml}. 
+As is clear from the code, the programmer can remain completely oblivious of the cryptographic back-end and underlying circuit representation. 
+Also,  the programmer is not required to write keywords such as {\em secret}
+or {\em public} (that are required by languages like Wysteria \cite{wysteria}). 
+ In contrast, \figureref{ex-aby} illustrates
 the same example using the framework in ABY \cite{aby}. Here, as
-already mentioned, the programmer needs to write the low level gates
-like {\tt MULT, ADD, CONS, GT}, etc. The programmer is also required
-to use the appropriate share conversion gates such as {\tt A2Y} that
-converts Arithmetic shares to Yao shares. Moreover, it is easy to
-maintain code in \tool and small change to the functionality requires
+already mentioned, the programmer needs to create shares and write the low level gates
+like $\mathtt{MUL, ADD, CONS, GT}$, etc \TODO{make these fonts consistent}. The programmer is also required
+to use the appropriate share conversion gates such as $\mathtt{A2Y}$ that
+converts arithmetic shares to Yao-based boolean shares to be used in comparison after the inner-product computation.
+Moreover, it is easy to
+maintain code in \tool and a small change to the functionality requires
 only a small change in the code. In contrast, working with ABY, even
 for a small change a developer might have to re-work the whole circuit
-for efficiency reasons. We describe these examples in detail in
+for efficiency reasons. We elaborate on these examples  in
 \sectionref{ex}.
 
-\textbf{Performance.} Our compiler is aware of cryptographic costs of different representations and inter-conversions.
-Based on these costs, our compiler, given a program in \tool, automatically picks Arithmetic and Boolean representations for different sub-parts of the program and inserts the required inter-conversions as well.
-Our experiments show that automatically generated protocols using \tool match the performance of tailor-made protocols in \cite{minionn,shafindss} (see \sectionref{eval}).
-This shows that  \tool achieves generality, performance and programmer productivity.
-These features allow us to evaluate our system on a variety of benchmarks, especially, state-of-the-art ML prediction algorithms with \emph{real} data and models.
+\textbf{Generality and Performance.} Our compiler is aware of cryptographic costs of different representations and corresponding inter-conversions.
+Based on these costs, our compiler, given a program in \tool, automatically picks arithmetic and boolean representations for different sub-parts of the program and inserts the required inter-conversions as well.
+We demonstrate the generality of our framework by writing a wide range of functionalities in \tool such as linear regression, deep neural networks, decision trees and matrix factorization.
+Our evaluations show that automatically generated protocols using \tool are competitive with tailor-made protocols in \cite{shafindss,wu,barni,minionn,secureml,valeriaMatrix} (see \sectionref{eval}).
+These features allow us to evaluate our system state-of-the-art ML prediction algorithms, e.g., \sectionref{eval} describes evaluation on ML models from a recent ICML paper \cite{bonsai} and Tensorflow \cite{tensorflow}.
 
@@
 
-\textbf{Pipelining.} Another issue in adoption of \mpc is difficulty in scaling to large programs. 
+\textbf{Secure Code Partitioning (or Partitioning in short).} Another issue in adoption of \mpc is difficulty in scaling to large programs. 
 %In general, \mpc implementations struggle in scaling to large programssize of memory on a system limits the largest compute inside the SMC and hence,. 
-The reason being all \mpc implementations use a circuit-like representation as an intermediate representation.
+The reason being most \mpc implementations use a circuit-like representation as an intermediate representation.
 Hence, the largest compute that can be done securely is upper-bounded by the largest circuit that can fit in memory on a machine.
@@
 If so, we can mention this.}
-To address this issue, we develop a technique called \emph{pipelining} that allows us to run arbitrarily large programs.
-At a high level, we break the original program into a sequence of small sub-programs and mask the intermediate outputs using secure secret sharing between the two parties.
+To address this issue, we develop a technique called partitioning that allows us to run arbitrarily large programs.
+At a high level, we break the original program into a sequence of small sub-programs and mask the intermediate outputs using a secure secret sharing between the two parties.
 We appropriately append the share-reconstruction function to the beginning of the subsequent sub-program.
 Now, we execute each of these sub-programs sequentially in our framework.
-We provide details of pipelining and prove the correctness and security of this transformation in \sectionref{pipe}.
+We provide details of partitioning and prove the correctness and security of this transformation in \sectionref{pipe}.
 

### 2017-10-24 OaklandSubmission/eval.tex
working on s6

@@
 \label{sec:eval}
-We evaluate \tool for the problem of {\it secure prediction}, where a server has a machine learning model, a client has an input, and after the protocol execution the client learns a label on his input.
-From the protocol execution, the server learns nothing about the client's input and the client learns nothing about the server's model beyond what is revealed from the computed label.
-This setting has been previously studied in~\cite{shafindss,minionn}
-We first implement benchmarks of ~\cite{shafindss} and~\cite{minionn} in \tool to compare against 
-previous work.
-Next, we demonstrate the generality of \tool by implementing state-of-the-art machine learning models.
-Finally, we discuss a model that requires pipelinig.
-Our results are in two settings, a LAN setting and a cross-continent WAN setting.
-The round trip time for two machines in a LAN setting is 1ms and in WAN setting is 40ms.
-Each machine has a Intel(R) Xeon(R) CPU E5-2673 v3 processor running at 2.40GHz with
-28 GBs of RAM. The \tool compiler is written in Python and compiles each of our benchmarks
-in under a second to C++ code that makes calls to the ABY library~\cite{aby}.
-We use an off-the-shelf solver (SeaHorn~\cite{seahorn}) to check the verification conditions
-required to prove that the array indices remain in bound.
-The solver takes less than a minute on our largest benchmark.
-Since most of our benchmarks are related to machine learning, we set up some notation standard to machine learning and describe our benchmarks next.
+We evaluate \tool on the problem of {\it secure prediction}, where one
+party (the server) has a machine learning model, and the other party
+(the client) has an input. The goal is to compute the output of the
+model on client's input, with the guarantee that the server
+learns nothing about the input, and the client learns nothing
+about the model beyond what is revealed from the output.
+
+We first implement the benchmarks from Bost et al.~\cite{shafindss}
+and \minion~\cite{minionn} (both of which study the same setting),
+and show that the performance of the high-level code written in \tool
+is comparable to their hand-crafted circuits
+(Section~\ref{sec:shallow} and Section~\ref{sec:dnneval}).
+%
+Next, we demonstrate the generality and programmability aspects of
+\tool by implementing state-of-the-art machine learning models from
+Tensorflow~\cite{tensorflow} and \bonsai (\aseem{cite?}).
+%
+Finally, we
+implement a Deep Neural Network (DNN) for CIFAR-10
+dataset~\cite{cifar} from \minion~\cite{minionn} to illustrate the use
+of partitioning in \tool.
+
+We present the numbers for two network settings, a LAN setting and
+a cross-continent WAN setting. The round trip time between the server
+and the client machines in the two settings is 1ms and 40ms resp. Each
+machine has an Intel(R) Xeon(R) CPU E5-2673 v3 processor running at
+2.40GHz with 28 GBs of RAM.
+%
+%% The  \tool compiler is written in Python and
+%% compiles each of our benchmarks in under a second to C++ code that makes calls to the ABY
+%% library~\cite{aby}. We use an off-the-shelf solver
+%% (SeaHorn~\cite{seahorn}) to check the verification conditions required
+%% to prove that the array indices remain in bound. The solver takes less
+%% than a minute on our largest benchmark.
+Since most of our benchmarks
+are related to machine learning, we set up some notation (largely
+standard in machine learning) and describe our benchmarks next.
 
@@
 \noindent\textbf{Notation.} We use $[N]$ to denote $\{0,1,\dotsc, N-1\}$. 
-Given a vector $x\in\R^d$, $\mathsf{argmax}\ x = i$ if $x(i) = \mathsf{max}\ \{x_1,\ldots,x_d\}$
+Given a vector $x\in\R^d$, $\mathsf{argmax}\ x = i$ if $x_i =
+\mathsf{max}\ \{x_1,\ldots,x_d\}$.
 
-In this paper, we focus on machine learning models for {\it classification}.
+We focus on the machine learning models for {\it classification}.
 A {\it classifier} $C$ uses a trained {\it model} to {\it predict} a {\it label} $\ell$ for an input data point $x$.
@@
 \subsection{Deep neural nets}
+\label{sec:dnneval}
 We evaluate \tool on the DNNs described in \minion~\cite{minionn} and the results are reported in Table~\ref{tab:nn}\footnote{\minion does not report the network round-trip time~\cite{minionn}.}.
@@
 \[
- \mathsf{argmax}_{i} w_i^Tx+b_i
+ \mathsf{argmax}\ Wx+b
 \]
 Here, $i\in [10]$, $x$ is a 784 length vector obtained from the input image,
-each $w_i$ is a 784 length vector, and each $b_i$ is a scalar. 
+each $W$ is a $10\times 784$ matrix, and each $b$ is a $784$ length vector. 
 We implement this classifer in \tool and the results are shown in the first row of Table~\ref{tab:tf}.

### 2017-10-24 OaklandSubmission/intro.tex

@@
-\aseem{Is the title too general :). May be we want something more
-  specific on the lines of A compiler for mixed efficient MPC
-  protocols, or something?}
-
-  \divya{Do we write the intro for \mpc or 2PC?}
-
   \TODO{While taking a pass on intro, check citation correctness or if we are missing any citation.}
@@
 
-Below, we describe the salient features of \tool in more detail.
+Below, we describe the salient features of \tool in more detail. \\
 
-\textbf{Programmability.} \tool is a programmer-friendly framework to
+\noindent\textbf{Programmability.} \tool is a programmer-friendly framework to
 express the function to be computed and uses a mix of arithmetic and
@@
 for efficiency reasons. We elaborate on these examples  in
-\sectionref{ex}.
+\sectionref{ex}. \\
 
-\textbf{Generality and Performance.} Our compiler is aware of cryptographic costs of different representations and corresponding inter-conversions.
+\noindent\textbf{Generality and Performance.} Our compiler is aware of cryptographic costs of different representations and corresponding inter-conversions.
 Based on these costs, our compiler, given a program in \tool, automatically picks arithmetic and boolean representations for different sub-parts of the program and inserts the required inter-conversions as well.
@@
 These features allow us to evaluate our system state-of-the-art ML prediction algorithms, e.g., \sectionref{eval} describes evaluation on ML models from a recent ICML paper \cite{bonsai} and Tensorflow \cite{tensorflow}.
+\\
 
-\textbf{Type system.} We give a formal {\em type system} for our language and prove useful properties.
+\noindent\textbf{Performance Baseline.} Most works cite garbled circuit based two-party computation as the most practical approach so far and use it as the baseline for comparison of performance.
+But, it has been shown multiple times that the garbled circuits based approach does not scale for realistic examples.
+Even for small benchmarks (e.g., Na\"{i}ve Bayes) the size of garbled circuit can be huge.
+Hence, comparisons are often done on artificial data sizes and garbled circuits are shown to be 500x slower than hand-crafted protocols \cite{shafindss}.
+This is cited as the motivation for need for specialized protocols for different ML algorithms in secure prediction such as linear regression \cite{shafindss}, decision trees \cite{wu}, deep neural networks \cite{minionn} and matrix factorization \cite{valeriaMatrix}.
+Our work is based on generic \mpc and we show competitive performance with the specialized protocols contrary to the popular belief.
+Hence, we believe \tool based implementations can serve as the performance baseline for future works on specialized protocols.
+\\
+
+%Finally, our framework \tool can serve as a meaningful baseline for future works on hand-crafted \mpc protocols for specific functionalities such as \textsc{MiniONN} \cite{minionn}. 
+%Prior to our work, garbled circuit based two-party computation is cited as the most practical approach and used as the baseline for comparison of performance.
+But, it has been observed multiple times this approach does not even scale on realistic examples.
+
+\noindent\textbf{Static Guarantees.} We give a formal {\em type system} for our language and prove useful properties.
 We prove that if a program type checks in our language then it would run correctly, e.g., there would not be any array index out of bounds error.
@@
 For details on formal language, correctness and security theorems see \sectionref{ld}.
+\\
 
-\textbf{Secure Code Partitioning (or Partitioning in short).} Another issue in adoption of \mpc is difficulty in scaling to large programs. 
+\noindent\textbf{Secure Code Partitioning (or Partitioning in short).} Another issue in adoption of \mpc is difficulty in scaling to large programs. 
 %In general, \mpc implementations struggle in scaling to large programssize of memory on a system limits the largest compute inside the SMC and hence,. 
 The reason being most \mpc implementations use a circuit-like representation as an intermediate representation.
-Hence, the largest compute that can be done securely is upper-bounded by the largest circuit that can fit in memory on a machine.
-\nc{Should we point out here that ABY itself has an explicit limit on circuit size and also give an example of a program that easily surpasses this bound?
-Also, hadnt ABY also said something about this?
-If so, we can mention this.}
-To address this issue, we develop a technique called partitioning that allows us to run arbitrarily large programs.
+Hence, the largest compute that can be done securely is upper-bounded by the largest circuit that can fit in memory on a machine\footnote{Using swap and disk space is feasible but it causes huge slowdown.}.
+To address this issue, we develop a technique called partitioning.
 At a high level, we break the original program into a sequence of small sub-programs and mask the intermediate outputs using a secure secret sharing between the two parties.
@@
 Now, we execute each of these sub-programs sequentially in our framework.
+This allows us to run arbitrarily large programs.
 We provide details of partitioning and prove the correctness and security of this transformation in \sectionref{pipe}.
+\\
 
-\textbf{Baseline.} Finally, our framework \tool can serve as a meaningful baseline for future works on hand-crafted \mpc protocols for specific functionalities such as \textsc{MiniONN} \cite{minionn}.
-Prior to our work, garbled circuit based two-party computation is cited as the most practical approach and used as the baseline for comparison of performance.
-But, it has been observed multiple times this approach does not even scale on realistic examples.
-
-%\newpage
-%\divya{ideally have small paragraphs about key features}
-%1. ease to programming.. refer to section with example in our language; contrast to verilog paper; programmer is completely oblivious of security; looks like a normal/vanilla/ regular C program;
-%Moreover, the developer does not write keywords such as secret and public; 
-%
-%
-%2.
-first programmable system that supports a mix of arithmetic and boolean for efficiency give example of multiplication of integers; linear algebra; logistic regression and evaluate on state of the art ML prediction tasks
-%
-%3. prove correctness, type system guarantees correct termination, progress etc..; 
-%
-%%In contrast with most systems that require a programmer to label variables as public secret, we automatically infer them. leads to more efficiency compared to default secret
-%
-%If program type-checks then runs correctly for example no array out of bound errors
-%code runs correctly in distributed env
-%programmer writes one code; compiles to server code and client code
-%we prove and outputs identical to stand-alone output
-%distributed code produces same output as if the code were run by distributed third party
-%
-%
-%4.
-Scaling to large programs.
-Describe pipeling here...
-Like all MPC implementations use an intermediate rep as a circuit and largest compute that can run is given by largest circuit that can fit in memory are restricted by memory of machines.
-This in turn limits the size of functions
-%We implement/give/design a generic technique called pipelining that allow us to run arbitrarily large computations without affecting security.
-%
-%
-%5. Previous papers when comparing with generic 2pc compare with Yao; does not scale etc...
-%can serve as a baseline for future works on hand-crafted algorithms for specific applications cite minion etc here...
-%
-
-\vspace{10pt}
-
-\noindent\textbf{Other Related Work} 
+\noindent\textbf{Other Related Work.} 
 Before ABY \cite{aby}, to improve efficiency, several works proposed
 to mix secure computation protocols based on homomorphic
-encryption with Yao’s garbled circuits protocol, e.g., \cite{barni,blanton,brickell,franz,huang,valeriaMatrix,valeriaRidge,schropferK11}
-  and developed tools that allowed writing a mixed \mpc protocol, e.g., \cite{bogdanov,lone,tasty}. The assignment can
+encryption and Yao's garbled circuits protocol, e.g., \cite{barni,blanton,brickell,franz,huang,valeriaMatrix,valeriaRidge,schropferK11}
+  and developed tools that allowed writing a mixed \mpc protocol, e.g., \cite{bogdanov,lone,tasty}. The breaking up of the function into homomorphic encryption and garbled circuits can
 even be done automatically as was shown in \cite{autoS}. 
-However, as noted by ABY, due to high conversion cost between homomorphic encryption and Yao's garbled circuit, these mixed protocols do not gain much over single protocol.
+However, as was noted by ABY, due to high conversion cost between homomorphic encryption and Yao's garbled circuit, these mixed protocols do not gain much performance over a single protocol.
 Also, none of these works give a high-level developer-friendly language support.
 No prior work has given a formal type system and operational semantics for mixed arithmetic and boolean back-ends.
-In our evaluation, we use benchmarks for the problem of secure prediction where the cloud holds the secret model and the client holds the secret input.
-For this problem various works have considered tailor-made protocols depending on the model considered \cite{shafindss,wu,secureml,minionn}.
-In our evaluation, we compare our performance with these specialized protocols and give competitive run-times.
 We provide a detailed survey of related work in \sectionref{related}.

### 2017-10-24 OaklandSubmission/eval.tex
intro pass

--- 2017-10-17 OaklandSubmission/eval.tex
@@
 This improvement is significant as~\cite{shafindss} uses custom designed protocols and we are executing a generic two party computation (2PC).
-Moreover,~\cite{shafindss} remarks that generic 2PC did not scale to the smallest of their Naive Bayes classifiers and they had to scale down the prediction task and there a Yao-based implementation was 500 times slower.
+Moreover,~\cite{shafindss} remarks that generic 2PC did not scale to the smallest of their Na\"{i}ve Bayes classifiers and they had to scale down the prediction task and there a Yao-based implementation was 500 times slower.
 We have shown that by using \tool, generic 2PC can scale to real prediction tasks and provide performance competitive with specialized protocols.

### 2017-10-24 OaklandSubmission/ld.tex
nit
pass over section 3 and 4.

@@
 computation circuit.
-
+\nc{Should we also say that security relies on correctness? (they go together)}
 We present only selected parts of our formalization for space
@@
 elided for space reasons.
-
+\nc{fancy PL folks language :)}
 The judgment $\rho \vdash s \Downarrow \rho_{1}; O$ represents the
@@
 \subsubsection*{Circuit semantics} The target semantics produces a
-circuit to compute over the secret data using an MPC protocol. With
+circuit to compute over the secret data using an MPC protocol. \nc{We use 2PC and MPC interchangeably. I guess its fine?}With
 our circuit semantics, we model the \emph{functional} aspect of the
@@
 mapped to (random) strings $b$. The semantics of these strings is
-given by pairs of encryption-decryption functions, written as
+given by pairs of encryption-decryption functions\nc{perhaps call this encode-decode functions since its technically not encryption-decryption?
+In fact more appropriate might be share-reconstruct, but that would require changing notation so lets not do that.}, written as
 $\mathcal{E}_{m}$ and $\mathcal{D}_{m}$ (where $m$ is either
@@
                {
-                 \Gamma_{1} = \Gamma, x:\kw{uint}^{\mathcal{P}}\\\\
-                 \Gamma_{1} \vdash \loops{x}{n_{2}}{s} \leadsto \loops{x}{n_{2}}{\widetilde{s}} \mid \Gamma_{1}
+                 \Gamma, x:\kw{uint}^{\mathcal{P}} \vdash \loops{x}{n_{2}}{s} \leadsto \loops{x}{n_{2}}{\widetilde{s}} \mid \_
                }
@@
 
-  \begin{enumerate}
+  \begin{enumerate}[label=(\alph*)]
   \item $\rho \vdash s \Downarrow \rho_{1}; O_{1}$
@@
 
-\nc{Super high level security theorem}
+\nc{Should we point out that the proof of Theorem 1 is provided elsewhere?}
 \subsubsection*{Security theorem} The protocols we generate provide
@@
 distribution ensembles above are computationally indistinguishable.
-\nc{I'm not happy with the theorem statement.}
+
+We shall assume a cryptographic MPC backend that securely implements any circuit $\chi$ that is output by our compiler.
+In more detail, this means that for every source program $s$, let $\chi$ be the circuit output by our compiler (as in Theorem \ref{theorem:correctness}).
+We assume that there exists a two-party secure computation protocol $\prot$ that securely realizes the functionality $\chi$ and we will call the corresponding simulator for the protocol as $\simu_{\tiny{2pc}}$ (that runs on input $\chi$ and the output produced by $\chi$ on the parties inputs).
+We note that the work of \cite{aby} provides such a protocol $\prot$ and simulator $\simu_{\tiny{2pc}}$ for all circuits $\chi$ output by our compiler.
+We are now ready to state and prove our security theorem.
+
 \begin{theorem}[Security]\label{theorem:security}
-Let $\F$ be a functionality in our source language with outputs (or observations) $O_1$, that is compiled into a protocol $\prot$ in our target language with outputs $O_2$, then $\prot$ securely realizes $\F$.
+Let $s$ be any functionality or program in our source language with outputs (or observations) $O_1$, that is compiled into a circuit $\chi$ (as defined in Theorem \ref{theorem:correctness}).
+Let protocol $\prot$ be the two-party secure computation protocol that securely realizes $\chi$ (as defined above).
+Then, $\prot$ securely realizes $s$.
 \end{theorem}
-\nc{We should probably connect things to earlier? Theorem 1 and 2 have very different languages.}
-\noindent {\em Proof.} From Theorem \ref{theorem:correctness}, we have
-that the observations in $\prot$ and $\F$ are the same (i.e. $O_2 =
-O_1$). Now, from the simulation security of our backend (\cite{aby}),
-the proof of the above theorem follows. \nc{I dont know how to say
-  more.}
-
+
+\noindent {\em Proof.} Our simulator $\simu$ simply runs our compiler on program $s$ to obtain $\chi$ and upon receiving output $O_1$, executes $\simu_{\tiny{2pc}}$ on $\chi$ and $O_1$.
+Let $O_2$ be the outputs (or observations) of $\chi$.
+First, from Theorem \ref{theorem:correctness}, we have that the observations in $s$ and $\chi$ are the same (i.e.
+$O_2 = O_1$).
+Now, from the security of $\prot$, we have that the simulated view output by $\simu_{\tiny{2pc}}$ (with circuit $\chi$ and output $O_2$) is indistinguishable from the real view of the $\prot$ (when executed on circuit $\chi$ and when the output is $O_2$).
+Combining these two statements, the proof of the theorem follows.
 \aseem{Can we give a construction of the simulator S in the ideal
@@
 
+\nc{OK now?}
   %% \\

### 2017-10-24 OaklandSubmission/eval.tex
6.1

@@
 \tool by implementing state-of-the-art machine learning models from
-Tensorflow~\cite{tensorflow} and \bonsai (\aseem{cite?}).
+Tensorflow~\cite{tensorflow} and \bonsai~\cite{bonsai}. Indeed, we
+provide the first 2PC implementation of \bonsai.
 %
@@
 \subsection{Benchmarks}
-\divya{notation can be moved somewhere... } \\
+%\divya{notation can be moved somewhere... } \\
 
-\noindent\textbf{Notation.} We use $[N]$ to denote $\{0,1,\dotsc, N-1\}$. 
-Given a vector $x\in\R^d$, $\mathsf{argmax}\ x = i$ if $x_i =
-\mathsf{max}\ \{x_1,\ldots,x_d\}$.
+\noindent(We use $[N]$ to denote $\{0,1,\dotsc, N-1\}$. 
+Further, given a vector $x\in\R^d$, we say $\mathsf{argmax}\ x = i$ if
+$x_i = \mathsf{max}\ \{x_1,\ldots,x_d\}$.)
 
 We focus on the machine learning models for {\it classification}.
-A {\it classifier} $C$ uses a trained {\it model} to {\it predict} a {\it label} $\ell$ for an input data point $x$.
-For example, given a data point which is a tuple of height and weight of an individual,
-a classifier can predict a label ``male'' or ``female''. The {\it accuracy} of a classifier
-refers to the fraction of data points that the classifier labels correctly from a given set of {\it test} data points.
+A classifier $C$ uses a trained model to {\it predict} a
+label $\ell$ for an input data point $x$. For example, given a
+data point which is a tuple of height and weight of an individual,
+a classifier can predict a label ``male'' or ``female''. The {\it
+  accuracy} of a classifier refers to the fraction of data points that
+the classifier labels correctly from a given set of test data
+points.
 
-A binary linear classifier is one of the simplest classifiers. Here, the input data point $x\in\R^d$
-and the model is a vector $w\in\R^d$. The possible labels are $\ell\in\{\mathit{true},\mathit{false}\}$
-and the classifier is $C_w\equiv w^Tx>0$, where $\cdot^T$ denotes the matrix transpose operator.
-This classifier requires $d$ multiplications, $d-1$ additions, and a comparison.
+\subsubsection*{Standard classifiers}
+A binary linear classifier is one of the simplest classifiers. Here,
+the input is a data point $x\in\R^d$,
+and the model is a vector $w\in\R^d$. The possible labels are
+$\ell\in\{\mathit{true},\mathit{false}\}$ and the classifier is
+$C_w\equiv w^Tx>0$, where $\cdot^T$ denotes the matrix transpose
+operator. This classifier requires $d$ multiplications, $d-1$
+additions, and a comparison.
 
-A more interesting classifier is Na\"{i}ve Bayes that predicts labels from the set $[n]$.
+A more interesting classifier is Na\"{i}ve Bayes that predicts labels
+from the set $[n]$.
 Here, the input data point is a {\it feature}
 vector $x=(x_0,x_2,\ldots,x_{d-1})^T$ where each $x_j\in [F]$.
-The model has two matrices: a vector $P$ of length $n$ and for each $i \in [n]$, $P(i)$ is the log-probability that the output label is $i$.
-%$P(i) \equiv \log p(\ell=i)$, where $p(\ell=i)$ is the likelihood that the output label is $i$ for $i\in[n]$.
-The other matrix $T$ has size $d\times F\times n$ and for each $j\in [d], k\in [F], i\in [n]$, $T(j)(k)(i)$ is
+The model has two matrices: a vector $P$ of length $n$ s.t. for each
+$i
+\in [n]$, $P(i)$ is the log-probability that the output label is $i$.
+%$P(i) \equiv \log p(\ell=i)$, where $p(\ell=i)$ is the likelihood
+%that the output label is $i$ for $i\in[n]$.
+The other matrix $T$ has size $d\times F\times n$ and for each $j\in
+[d], k\in [F], i\in [n]$, the entry $T(j)(k)(i)$ is 
 %\equiv\log p(x_j|\ell=i)$, i.e.,
@@
 \]
-This classifier requires $ndF$ comparisons and additions.
-\divya{Note that computing $T(i)(j)(x_j)$ requires a secret look-up in matrix $T$  based on secret input $x_j$ resulting in $F$ comparisons (see \sectionref{}).} \cmmt{Is there a +1 etc missing or is this exact?}
+This classifier requires $ndF$ comparisons and additions. \divya{Note
+  that computing $T(i)(j)(x_j)$ requires a secret look-up in matrix
+  $T$  based on secret input $x_j$ resulting in $F$ comparisons (see
+  \sectionref{}).} \cmmt{Is there a +1 etc missing or is this exact?}
 
-A decision tree of size $N$ takes as input $x\in\R^d$ and the model consists of a binary tree of depth $d$ with a boolean predicate
-assigned to each internal node.
-The root note is the $0^{th}$ node and for an internal node $j \in [N]$, the children nodes are $2j+1$ and $2j+2$.
-Each internal node $j\in[N]$ at depth $i$ has a predicate $b_{j}^i\equiv x_i\leq w_i$.
-We start evaluating the tree from the root and if the predicate at the current node $b_j^i$ is false (resp., true) then $x$ is passed to the left (resp., right) child with predicate $b_{2j+1}^{i+1}$ (resp., $b_{2j+2}^{i+1}$).
-This process is repeated till we reach a leaf.
-The leaves of the tree are labels and the output label is the leaf visited by this traversal.
-Such a (binary) decision tree can be encoded as a polynomial (of degree linear in depth of the tree) and then the prediction task
-reduces to polynomial evaluation. E.g., a binary tree with depth one and size $N=3$ can be encoded as the polynomial
-$(1-b(x))\ell_0 + b(x)\ell_1$ where $b(x)$ is the predicate at the root, $\ell_0$ is the label of the left leaf and $\ell_1$ is the label of the right leaf.
+A decision tree of size $N$ takes as input an $x\in\R^d$, and the
+model consists of a binary tree of depth $d$ with a boolean predicate
+assigned to each internal node. The root note is the $0^{th}$ node and
+for an internal node $j \in [N]$, the children nodes are $2j+1$ and
+$2j+2$. Each internal node $j\in[N]$ at depth $i$ has a predicate
+$b_{j}^i\equiv x_i\leq w_{j}$. We start evaluating the tree from the
+root and if the predicate at the current node $b_j^i$ is false (resp.,
+true) then $x$ is passed to the left (resp., right) child with
+predicate $b_{2j+1}^{i+1}$ (resp., $b_{2j+2}^{i+1}$). This process is
+repeated till we reach a leaf. The leaves of the tree are labels and
+the output label is the leaf visited by this traversal.
+Such a (binary) decision tree can be encoded as a polynomial (of
+degree linear in depth of the tree) and then the prediction task
+reduces to polynomial evaluation. E.g., a binary tree with depth one
+and size $N=3$ can be encoded as the polynomial
+$(1-b(x))\ell_0 + b(x)\ell_1$ where $b(x)$ is the predicate at the
+root, $\ell_0$ is the label of the left leaf and $\ell_1$ is the label
+of the right leaf.
 
-The next class of classifiers that we discuss are deep neural nets or DNNs. 
-A DNN has multiple layers such that each layer computes a matrix multiplication followed by an {\it activation} function $f$.
-The most common activation functions are square $f(x)=x^2$ and rectifier linear unit (ReLU) $f(x)=\mathsf{max}(x,0)$.
-If $A$ is a matrix then the output of $f(A)$ is also a matrix that is obtained by applying $f$ to each entry of $A$ pointwise.
+\subsubsection*{Deep neural nets}
+The next class of classifiers that we benchmark are deep neural nets
+or DNNs. A DNN has multiple layers such that each layer computes a
+matrix
+multiplication followed by an {\it activation} function $f$. The most
+common activation functions are square $f(x)=x^2$ and rectifier linear
+unit (ReLU) $f(x)=\mathsf{max}(x,0)$.
+If $A$ is a matrix then the output of $f(A)$ is also a matrix that is
+obtained by applying $f$ to each entry of $A$ pointwise.
 Given an input vector $x$, the predicted label of a DNN is
@@
 \]
-Here, $f_i$'s are the \divya{public} activation functions and the model consists of matrices $W_i$. The input vector is $x \in \R^d$.
-The operator $\cdot$ denotes a matrix multiplication.
-Neural nets usually have one or more fully connected layers.
-These layers multiply a matrix with a vector.
-Some neural nets have convolution layers and such DNNs are also called convolutional neural nets or CNNs.
-For the purpose of this paper, a convolution can be considered as a (heavy) matrix-matrix multiplication.
-The size of matrices manipulated by a convolution layer grows linearly with {\it window size} (typically 9 or 25), the number of {\it output channels} (typically 16, 32, or 64), and the size of the matrix input to this layer.
+Here, $f_i$'s are the (public) activation functions, the model
+consists of matrices $W_i$,  $x \in \R^d$ is the input vector, and the
+operator $\cdot$ denotes a matrix multiplication.
+Neural nets usually have one or more fully connected layers, each of
+which multiplies a matrix with a vector.
+Some neural nets have convolution layers and such DNNs are also called
+Convolutional Neural Nets or CNNs.
+For the purpose of this paper, a convolution can be considered as a
+(heavy) matrix-matrix multiplication. The size of matrices manipulated
+by a convolution layer grows linearly with {\it window size}
+(typically 9 or 25), the number of {\it output channels} (typically
+16, 32, or 64), and the size of the matrix input to this layer.
 Therefore, fully connected layers are lighter computation-wise compared
-to convolution layers. However, the model size of fully connected layers is larger than those of convolution layers.
+to convolution layers. However, the model size of fully connected
+layers is larger than those of convolution layers.
 In general, DNNs are computationaly heavy but provide
-much better accuracies on computer vision tasks than the classifiers discussed above.
+much better accuracies on computer vision tasks than the classifiers
+discussed above.
 
-Finally, there are a class of machine learning classifiers that are much more efficient than
-DNNs and provide reasonably good accuracies on standard learning tasks. \bonsai~\cite{bonsai} is a state-of-the art classifier in this class and \tool provides the first 2PC protocol for it.
-\bonsai takes as input $x \in \R^d$ and model consists of a binary tree with $N$ nodes.
-Each node $j$ contains matrices $W_j$ and $V_j$.
-The internal node $j$  evaluates a predicate $\theta_j^TZx > 0$ to decide whether
-to pass $x$ to the left child $2j+1$ or the right child $2j+2$ (where $\theta_j$ is a vector and $Z$ is a matrix).
+\subsubsection*{State-of-the-art classifiers}
+Finally, there are a class of machine learning classifiers that are
+much more efficient than
+DNNs and provide reasonably good accuracies on standard learning
+tasks. \bonsai~\cite{bonsai} is a state-of-the art classifier in this
+class and \tool provides the first 2PC protocol for it.
+\bonsai takes as input $x \in \R^d$, and its model consists of a
+binary tree with $N$ nodes, and a matrix $Z$. Each node $j$ contains
+matrices $W_j$ and $V_j$, and a vector $\theta_{j}$. The internal node
+$j$  evaluates a predicate $\theta_j^TZx > 0$ to decide whether
+to pass $x$ to the left child $2j+1$ or the right child $2j+2$.
 The predicted value is
 \[
-{\sf argmax} \sum_{j\in [N]} I_j(x)W_j^TZx\circ f(V_j^TZx) 
+{\sf argmax} \sum_{j = 0}^{N - 1} I_j(x)W_j^TZx\circ f(V_j^TZx) 
 \]
@@
 and is zero otherwise. 
-The operation $\circ$ is a point-wise dot product, $W_j$'s and $V_j$'s are matrices.
-The activation function $f$ is given by $f(y) = y$ if $-1 < y < 1$ and $\mathrm{sign}(x)$ otherwise.
-The model consists of the matrices $W_j$, $V_j$, $Z$ and vectors $\theta_j$.
-\bonsai can be seen as a variant of decision trees where the prediction is a function of the path traversed from root to the leaf and not just the leaf itself.
+The operation $\circ$ is a pointwise dot product, $W_j$'s and $V_j$'s
+are matrices. The activation function $f$ is given by $f(y) = y$ if
+$-1 < y < 1$ and $\mathrm{sign}(y)$ otherwise. \bonsai can be
+seen as a variant of decision trees where the prediction is a function
+of the path traversed from root to the leaf and not just the leaf
+itself.
 
-In the following, we implement these classifiers in \tool and report the time taken for making predictions securely.
-Ideally, the machine learning classifiers are mathematical expressions over Reals that are usually approximated by floating-point operations.
-As is standard, we port the classifers to integer manipulating programs by scaling the models and rounding~\cite{minionn}. These ported classifiers are then implemented in \tool.
+In the following, we implement these classifiers in \tool and report
+the time taken for making secure predictions. Ideally, the machine
+learning classifiers are mathematical expressions over $\R$ that are
+usually approximated by floating-point operations. 
+As is standard, we port the classifers to integer manipulating
+programs by scaling the models and rounding~\cite{minionn}. These
+ported classifiers are then implemented in \tool.
 

### 2017-10-24 OaklandSubmission/eval.tex
adding a note

@@
 The predicted value is
+
+\aseem{I had a hard time parsing this. (1) Should we use $\cdot$ for
+  matrix multiplication as we did for DNNs? (2) Precedence of $\circ$
+  over scalar multiplication (or is it associative so doesn't matter?)
+(3) The argument to $f$ is a scalar (right?), and so is its output,
+so then is $\circ$ defined OR is it applying $f$ pointwise to a vector?}
+
 \[
@@
 \]
+
 Here, $I_j(x)$ is 1 if the $j^{th}$ node is present on the path traversed by $x$

### 2017-10-24 OaklandSubmission/eval.tex
checkpointing at the beginning of 6.2

@@
 
-\subsection{Standard models}
+\subsection{Standard classifiers}
 \label{sec:shallow}
-We consider three types of classifiers: linear, Na\"{i}ve Bayes, and decision trees. 
-These classifiers are trained on the followinng data sets from the UCI  machine learning repository~\cite{uci}:
+We train the three standard classifiers, linear, Na\"{i}ve Bayes, and
+decision trees, on the following data sets from the UCI
+machine learning repository~\cite{uci}:
  the Wisconsin Breast Cancer data set, 
-Credit Approval data set, Audiology (Standardized) data set, Nursery data set, and
-ECG (electrocardiogram) classification data from~\cite{barni}.
+Credit Approval data set, Audiology (Standardized) data set, Nursery
+data set, and ECG (electrocardiogram) classification data
+from~\cite{barni}.
 
-The results for linear classification are in Table~\ref{tab:lc}.
-The input and model are both vectors of length $d$.
-The time in columns ``Prev.
-time'' and communication ``Prev.Comm'' is the time and total network communication reported in~\cite{shafindss}.
-They remark that these times are obtained by simulating a network with 40ms round trip time, which is the same as our WAN setting.
-The total execution time of \tool generated code in the LAN and the WAN setting is reported next followed by the total communication.
-We observe that \tool generated code has performance competitive to~\cite{shafindss}.
+The results for linear classification are in Table~\ref{tab:lc}. The
+input and model are both vectors of length $d$. The time in columns
+``Prev. time'' and communication ``Prev. comm'' is the time and total
+network communication reported by Bost et al.~\cite{shafindss}. They
+report these numbers over a simulated network with 40ms
+round trip
+time, which is same as our WAN setting. The total execution time
+of \tool generated code in the LAN and the WAN setting is reported
+next, followed by the total communication.
+We observe that \tool generated code has performance competitive
+to~\cite{shafindss}.
 The execution time is dominated by multiplications and boolean-ands.
-The number of boolean-and gates (\#And) remains constant for both the benchmarks as they reflect the number of and gates required for a single 32-bit integer comparison with zero.
-The number of multiplication gates (\#Mul) is the same as the model size $d$.
-The last column shows the lines of code of \tool program, which is also independent of the model size.
+The number of boolean-and gates (\#And) remains constant for both the
+benchmarks as they reflect the number of and gates required for a
+single 32-bit integer comparison with zero. The number of
+multiplication gates (\#Mul) is the same as the model size $d$. The
+last column shows the lines of code of \tool program, which is also
+independent of the model size.
 
@@
 \begin{tabular}{c|c|c|c|c |c|c|c|c|c|c}
-Dataset & $d$  & Prev. time (s) & Prev. Comm (kb) & LAN time (s) & WAN time (s) & Comm. (kb)  & \#And & \#Mul & \#Gates & LOC\\
+Dataset & $d$  & Prev. time (s) & Prev. comm (kb) & LAN time (s) & WAN time (s) & Comm. (kb)  & \#And & \#Mul & \#Gates & LOC\\
 \hline

### 2017-10-24 OaklandSubmission/eval.tex
some more in 6.2

@@
 
-The results for linear classification are in Table~\ref{tab:lc}. The
+\subsubsection*{Linear classifier}The results for linear
+classification are in Table~\ref{tab:lc}. The
 input and model are both vectors of length $d$. The time in columns
@@
 next, followed by the total communication.
-We observe that \tool generated code has performance competitive
-to~\cite{shafindss}.
-The execution time is dominated by multiplications and boolean-ands.
+We observe that the \tool code performance matches that of Bost
+et al. On the other hand, the programmer effort in \tool is just 20
+lines (last column in the table) of code in the source language, as
+opposed to writing low-level circuits. \aseem{They write circuits,
+  right?} The execution time is dominated by multiplications and
+boolean-ands.
 The number of boolean-and gates (\#And) remains constant for both the
 benchmarks as they reflect the number of and gates required for a
-single 32-bit integer comparison with zero. The number of
-multiplication gates (\#Mul) is the same as the model size $d$. The
-last column shows the lines of code of \tool program, which is also
-independent of the model size.
+single 32-bit integer comparison with zero, while the number of
+multiplication gates (\#Mul) is the same as the model size $d$.
+%% The
+%% last column shows the lines of code of \tool program, which is also
+%% independent of the model size.
 
-The results for Na\"{i}ve Bayes are in Table~\ref{tab:nb}. As before, $d$ denotes the length of the input feature vector $x$ and $F$ is the number of possible values of $x_i$ for each $i \in [d]$.
+\subsubsection*{Na\"{i}ve Bayes} The results for Na\"{i}ve Bayes are
+in Table~\ref{tab:nb}. As before, $d$ denotes the length of the input
+feature vector $x$ and $F$ is the number of possible values of $x_i$
+for each $i \in [d]$.
 Although, there are no multiplications in these benchmarks, these do have a significant number of comparisons (due to secret look-up) that result in many boolean-and gates (13k/750k denote 13000/75000).

### 2017-10-25 examples/factor/64_1_pipe_arith_cifar.cpp
factgor 64

@@ -1,0 +1,861 @@
+bitlen =64;
+ABYParty *party = new ABYParty(role, address, port, seclvl, bitlen, nthreads, mt_alg, 120000000);
+vector<Sharing*>& sharings = party->GetSharings();
+Circuit* ycirc = sharings[S_YAO]->GetCircuitBuildRoutine();
+Circuit* acirc = sharings[S_ARITH]->GetCircuitBuildRoutine();
+Circuit* bcirc = sharings[S_BOOL]->GetCircuitBuildRoutine();
+auto S1 = make_vector<uint64_t>(16384, 24);
+auto s_a_S1 = make_vector<share*>(16384, 24);
+auto S2 = make_vector<uint64_t>(16384, 24);
+auto s_a_S2 = make_vector<share*>(16384, 24);
+auto S = make_vector<uint64_t>(16384, 24);
+auto s_a_S = make_vector<share*>(16384, 24);
+auto outmat = make_vector<uint64_t>(16384, 24);
+auto s_a_outmat = make_vector<share*>(16384, 24);
+auto boolarr = make_vector<uint64_t>(200000);
+auto s_a_boolarr = make_vector<share*>(200000);
+auto temp = make_vector<uint64_t>(10);
+auto s_a_temp = make_vector<share*>(10);
+auto dots = make_vector<uint64_t>(15444);
+auto s_a_dots = make_vector<share*>(15444);
+uint64_t boolidx;
+auto Routmat = make_vector<uint64_t>(16384, 24);
+auto s_a_Routmat = make_vector<share*>(16384, 24);
+auto IRoutmat = make_vector<uint64_t>(16384, 24);
+auto s_a_IRoutmat = make_vector<share*>(16384, 24);
+uint64_t b;
+share *s_a_b;
+uint64_t b1;
+share *s_a_b1;
+boolidx = 0;
+uint64_t cas;
+share *s_a_cas;
+if(role == SERVER){
+ofstream fout("IRoutmat", std::ofstream::out | std::ofstream::app);
+for(int i = 0; i < 16384; i++) {
+	for(int j = 0; j < 24; j++) {
+		fout << i << " " << j << " " <<(IRoutmat[i][j] = rand()) << endl;
+	}
+}
+fout.close();
+}
[…]

### 2017-10-25 examples/factor/64_2_pipe_arith_cifar.cpp

@@ -1,0 +1,927 @@
+bitlen = 64;
+ABYParty *party = new ABYParty(role, address, port, seclvl, bitlen, nthreads, mt_alg, 120000000);
+vector<Sharing*>& sharings = party->GetSharings();
+Circuit* ycirc = sharings[S_YAO]->GetCircuitBuildRoutine();
+Circuit* acirc = sharings[S_ARITH]->GetCircuitBuildRoutine();
+Circuit* bcirc = sharings[S_BOOL]->GetCircuitBuildRoutine();
+auto outmat = make_vector<uint64_t>(16384, 24);
+auto s_a_outmat = make_vector<share*>(16384, 24);
+auto outmp = make_vector<uint64_t>(16384, 24);
+auto s_a_outmp = make_vector<share*>(16384, 24);
+auto boolarr = make_vector<uint64_t>(200000);
+auto s_a_boolarr = make_vector<share*>(200000);
+auto temp = make_vector<uint64_t>(10);
+auto s_a_temp = make_vector<share*>(10);
+auto dots = make_vector<uint64_t>(15444);
+auto s_a_dots = make_vector<share*>(15444);
+uint64_t boolidx;
+auto Routmat = make_vector<uint64_t>(16384, 24);
+auto s_a_Routmat = make_vector<share*>(16384, 24);
+auto IRoutmat = make_vector<uint64_t>(16384, 24);
+auto s_a_IRoutmat = make_vector<share*>(16384, 24);
+auto Routmp = make_vector<uint64_t>(16384, 24);
+auto s_a_Routmp = make_vector<share*>(16384, 24);
+auto IRoutmp = make_vector<uint64_t>(16384, 24);
+auto s_a_IRoutmp = make_vector<share*>(16384, 24);
+uint64_t b;
+share *s_a_b;
+uint64_t b1;
+share *s_a_b1;
+boolidx = 0;
+uint64_t t;
+uint64_t cas;
+share *s_a_cas;
+if(role == SERVER){
+ifstream fin("IRoutmat");
+while(!fin.eof()){
+	uint64_t i, j, val;
+	fin >> i >> j >> val;
+	IRoutmat[i][j] = val;
+}
+fin.close();
+}
+if(role == CLIENT){
+ifstream fin("outmat");
+while(!fin.eof()){
[…]

### 2017-10-25 examples/factor/64_3_pipe_arith_cifar.cpp

@@ -1,0 +1,831 @@
+bitlen=64;
+ABYParty *party = new ABYParty(role, address, port, seclvl, bitlen, nthreads, mt_alg, 120000000);
+vector<Sharing*>& sharings = party->GetSharings();
+Circuit* ycirc = sharings[S_YAO]->GetCircuitBuildRoutine();
+Circuit* acirc = sharings[S_ARITH]->GetCircuitBuildRoutine();
+Circuit* bcirc = sharings[S_BOOL]->GetCircuitBuildRoutine();
+auto outmp = make_vector<uint64_t>(16384, 24);
+auto s_a_outmp = make_vector<share*>(16384, 24);
+auto boolarr = make_vector<uint64_t>(200000);
+auto s_a_boolarr = make_vector<share*>(200000);
+auto temp = make_vector<uint64_t>(10);
+auto s_a_temp = make_vector<share*>(10);
+auto dots = make_vector<uint64_t>(15444);
+auto s_a_dots = make_vector<share*>(15444);
+uint64_t boolidx;
+auto Routmp = make_vector<uint64_t>(16384, 24);
+auto s_a_Routmp = make_vector<share*>(16384, 24);
+auto IRoutmp = make_vector<uint64_t>(16384, 24);
+auto s_a_IRoutmp = make_vector<share*>(16384, 24);
+uint64_t b;
+share *s_a_b;
+uint64_t b1;
+share *s_a_b1;
+boolidx = 0;
+uint64_t t;
+uint64_t cas;
+share *s_a_cas;
+if(role == SERVER){
+ifstream fin("IRoutmp");
+while(!fin.eof()){
+	uint64_t i, j, val;
+	fin >> i >> j >> val;
+	IRoutmp[i][j] = val;
+}
+fin.close();
+}
+if(role == CLIENT){
+ifstream fin("outmp");
+while(!fin.eof()){
+	uint64_t i, j, val;
+	fin >> i >> j >> val;
+	outmp[i][j] = val;
+	s_a_outmp[i][j]= acirc->PutINGate(outmp[i][j] ,bitlen,CLIENT);
+	
+}
+fin.close();
+}
+else {
+ for(int i = 0; i <16384;i++) {
+	for (int j = 0; j <24; j++) {
+		s_a_outmp[i][j] = acirc->PutDummyINGate(bitlen);
+	}
+}
+}
[…]

### 2017-10-25 OaklandSubmission/impl.tex
first draft of implementation

@@
 \label{sec:impl}
-\divya{Some pointers by Divya}
-Describe the following
-\begin{enumerate}
-\item python compiler etc
-\item array index out of bounds static check implementation
-\item pipelining? automatic or manual?
-\item secret indexing in arrays; not supported directly but coded as a tree; future work automatic compilation may be
+The  \tool compiler is written in Python and
+compiles each of our benchmarks in under a second to C++ code that makes calls to the ABY
+library~\cite{aby}. We use the default options of ABY except 
+for the maximum number of gates. Our benchmarks require us to set this value to a
+hundred million.
+ABY provides facilities to use multiple threads\footnote{The protocol has an offline phase and an online phase.
+Although multiple threads can be used in the offline phase, the online phase is single threaded.}.
+We use at most four threads in our evaluation.
 
-\end{enumerate}
+We use an off-the-shelf solver
+(SeaHorn~\cite{seahorn}) to check the verification conditions required
+to prove that the array indices remain in bound ($\Gamma\vDash e < n$ in {\sc T-Read}, Figure~\ref{fig:compile}). The solver takes less
+than a minute on our largest benchmark.
+The type label is assigned conservatively in  {\sc T-Decl}.
+Only the variables that govern the control flow, i.e., variables
+in if-conditions and for-loops are assigned public labels.
+All other variables are assigned arithmetic labels.
+In the future, we would like to explore more intelligent strategies for label assignment.
 
-Where do we mention about getting integer models; using unsigned ints for signed compute etc?
+Some of our benchamrks require accessing arrays at secret location.
+For example, consider the expression $A[x]$ where $A$ is an array of size 2 and $x$ is shared.
+\tool enforces the array indices to be public. The developer expresses this functionality in \tool
+as $x > 0 ? A[1] : A[0]$. In general, for an array of size $2^n$, we need to write $2^{n-1}$ multiplexers
+in \tool. This design choice of making secret indexing inconvinient to implement was deliberate.
+Accessing arrays at secret locations is a powerful but expensive primitive
+and we want the developers to perform this operation sparingly.
+
+The secure code partitioning is not fully automatic.
+in particular, the developer needs to decompose a program $P$ into $P_1||\ldots||P_k$ manually (Section~\ref{sec:pipe}). Then \tool generates $Q_1||\ldots||Q_k$ automatically.
+Automating the decomposition step requires an analysis that can 
+statically estimate the resource usage of a \tool program and we leave the construction of such
+an analysis as future work.
+
+The compilation rules of Figure~\ref{fig:compile} can introduce redundant coercions from arithmetic to
+bool and vice versa.
+The implemntation of \tool performs several optimizations to reduce these.
+For example, on each coercion, \tool memoizes the pair of arithmetic share and boolean share involved in the coercion.
+\tool invalidates such pairs when the variables corresponding to the shares are overwritten by assignments.
+In subsequent coercions, \tool can reuse valid pairs (if available) instead of recomputing them afresh.

### 2017-10-25 OaklandSubmission/ld.tex
ld

@@
 \label{sec:ld}
+\TODO{Later, in implementation we can write that we use ABY for compiling intermediate to target also.}
 
@@
 
+\divya{
+In this section we prove the correctness and security of our compiler.
+We formalize the source, the intermediate and the circuit languages. 
+Our source runtime semantics is a model of the ideal, trusted third-party semantics, and
+generates observations corresponding to the values revealed to the
+parties.
+%
+We present the compilation rules that type check a program in the source language and compiles it to a program in the intermediate language.
+%
+We evaluate the program in our intermediate language (a model of the C++ code generated by our compiler implementation) to a circuit by ``computing away'' the public parts and the arrays.
+Crucially, this step does not have access to the secret inputs--those are processed by the secure computation circuit.
+%
+Finally, we formalize the distributed circuit semantics (of the cryptographic back-end) that computes the generated circuit in the distributed setting, and like  the source semantics, emits observations
+corresponding to the values revealed to the parties.
+
+To prove the correctness of our compiler, we prove that it preserves the observations in source semantics and distributed circuit semantics (see \theoremref{correctness}). 
+For the security of our compiler, we rely on the correctness guarantee and  reduce the security to the
+security of the underlying cryptographic protocol used to compute the generated circuit (see \theoremref{security}).
+}
+
+\TODO{Do we want to write things in logical order or order in paper?}
+
 \begin{figure}
@@
                }
-               \quad
+               \hspace{0.1cm}
     \inferrule*[lab={\footnotesize{E-Inp}}]
@@
   \\\\
-    \fbox{$\rho \vdash c \Downarrow \rho'; O$}
+    \fbox{$\rho \vdash s \Downarrow \rho'; O$}
   \[
@@
 model only one dimensional arrays, our implementation supports higher
-dimensional arrays as well. Expressions in the language include the
+dimensional arrays as well. Expressions $e$ in the language include the
 integer constants $n$, \kw{bool} constants $\top$ and $\bot$,
@@
     \ftext{Value} & v &::=& c \mid [\overline{c_{i}}]_{n}\\
-    \ftext{Runtime environment} & \rho &::=& \cdot \mid \rho[x \mapsto v]\\
+    \ftext{Runtime environment} & \rho &::=& \cdot \mid \rho,x \mapsto v\\
     \ftext{Observation} & O & ::= & \cdot \mid c, O \\
@@
 \vspace{0.2cm}
-Values consist of constants and array values, runtime environment
+Values consist of constants and array of constants, runtime environment
 $\rho$ maps variables to values, and observations are sequences of
@@
 elided for space reasons.
+\divya{Does conditionals $\cond{e}{e_{1}}{e_{2}}$ have anything interesting? just checking!}
 \nc{fancy PL folks language :)}
+
 The judgment $\rho \vdash s \Downarrow \rho_{1}; O$ represents the

### 2017-10-25 OaklandSubmission/eval.tex
6.2

@@
 
-\subsection{Benchmarks}
+\subsection{Benchmarks description}
 %\divya{notation can be moved somewhere... } \\
@@
 classification are in Table~\ref{tab:lc}. The
-input and model are both vectors of length $d$. The time in columns
-``Prev. time'' and communication ``Prev. comm'' is the time and total
-network communication reported by Bost et al.~\cite{shafindss}. They
-report these numbers over a simulated network with 40ms
-round trip
-time, which is same as our WAN setting. The total execution time
+input and the model are both vectors of length $d$. The columns
+``Prev. time'' and ``Prev. comm'' show the time and the total
+network communication reported by Bost et al.~\cite{shafindss} for a
+network setting with 40ms
+round trip time, which is same as our WAN setting. The total execution
+time
 of \tool generated code in the LAN and the WAN setting is reported
@@
 We observe that the \tool code performance matches that of Bost
-et al. On the other hand, the programmer effort in \tool is just 20
-lines (last column in the table) of code in the source language, as
+et al., while the programmer effort in \tool is just 20
+lines (last column in the table) of high-level code in the \tool
+source language, as
 opposed to writing low-level circuits. \aseem{They write circuits,
-  right?} The execution time is dominated by multiplications and
+  right?} Our execution time is dominated by the multiplications and
 boolean-ands.
 The number of boolean-and gates (\#And) remains constant for both the
-benchmarks as they reflect the number of and gates required for a
+benchmarks as they reflect the number of gates required for a
 single 32-bit integer comparison with zero, while the number of
@@
 for each $i \in [d]$.
-Although, there are no multiplications in these benchmarks, these do have a significant number of comparisons (due to secret look-up) that result in many boolean-and gates (13k/750k denote 13000/75000).
-Here, again columns 4 and 5 show the results reported in~\cite{shafindss}.
-We observe that \tool generated code has competitive performance on this task.
-This improvement is significant as~\cite{shafindss} uses custom designed protocols and we are executing a generic two party computation (2PC).
-Moreover,~\cite{shafindss} remarks that generic 2PC did not scale to the smallest of their Na\"{i}ve Bayes classifiers and they had to scale down the prediction task and there a Yao-based implementation was 500 times slower.
-We have shown that by using \tool, generic 2PC can scale to real prediction tasks and provide performance competitive with specialized protocols.
+Although, there are no multiplications in these benchmarks, these do
+have a significant number of comparisons (due to array look-ups using
+secret indices) that
+result in many boolean-and gates. %(13k/750k denote 13000/75000). 
+As before columns ``Prev. time'' and ``Prev. comm.'' (columns 4 and 5)
+show the numbers reported
+by Bost et al.~\cite{shafindss}. We observe that \tool generated code
+has better performance, despite using a generic 2PC,
+as opposed to custom designed protocols developed by Bost et
+al. Moreover, they remark that in their setup, generic Yao-based 2PC
+did not scale to the smallest of their Na\"{i}ve Bayes classifiers, so
+they had to scale down the prediction task, and even then their
+implementation was 500x slower. Whereas, we show that by using a
+cryptographic-cost aware compiler, we can scale generic 2PC to real
+prediction tasks, and get performance competitive to or better than the
+specialized protocols. \aseem{This last claim is a bit
+  tricky. Their claim was specifically about Yao-based. Is our
+  output protocol also Yao-based? If not, then we are not comparing
+  apples to apples.}
 
-This claim is further validated by Table~\ref{tab:dt}.
-In~\cite{wu}, Wu et al. describe specialized protocols for trees and forests (that perform better than~\cite{shafindss}) and we report their performance in columns 3 and 4 of Table~\ref{tab:dt}.
-Similar to~\cite{shafindss},~\cite{wu} remarks that these computations are infeasible with Yao-based 2PC.
-Again, we observe that the performance of \tool generated generic 2PC code is competitive with specialized protocols.
-Therefore, we believe that \tool generated code can be a better baseline than Yao-based 2PC when comparing against generic 2PC-based implementations for secure prediction tasks.
+\subsubsection*{Decision trees}
+Our decision trees numbers in Table~\ref{tab:dt} further validate
+this claim. We compare with Wu et al.~\cite{wu} (columns 3 and 4) who
+describe specialized protocols for trees and forests (that perform
+better than~\cite{shafindss}).
+Similar to Bost et al.~\cite{shafindss}, they also remarks that these
+computations are infeasible with Yao-based 2PC.
+\aseem{Same comment as before, if our output protocols are not
+  Yao-based, then it is an unfair comparison.} Again, we observe that
+the performance of \tool generated generic 2PC
+code is competitive with specialized protocols. Therefore, we believe
+that \tool generated code can be a better baseline than Yao-based 2PC
+when comparing against generic 2PC-based implementations for secure
+prediction tasks. \aseem{I don't understand this last sentence. Why is
+  ours a better baseline?}
 

### 2017-10-25 OaklandSubmission/ld.tex

@@
 security of the underlying cryptographic protocol used to compute the generated circuit (see \theoremref{security}).
+We present only selected parts of our formalization for space
+reasons. \TODO{Full definitions and proofs can be found in the supplementary
+material submitted along with the paper.}
 }
@@
                \hspace{0.1cm}
-
+               \inferrule*[lab={\footnotesize{E-Cond}}]
+               {
+                 \rho \vdash e \Downarrow c\\\\
+                 c = \top \Rightarrow e' = e_{1}\\\\
+                 c = \bot \Rightarrow e' = e_{2}\\\\
+                 \rho \vdash e \Downarrow c'
+               }
+               {
+                 \rho \vdash \cond{e}{e_{1}}{e_{2}} \Downarrow c'
+               }
+
+\\\\
 	 \inferrule*[lab={\footnotesize{E-Read}}]
@@
                }
-\\\\
+
+               \hspace{0.1cm}
     \inferrule*[lab={\footnotesize{E-Arr}}]
                {
-                 \forall i \in \{0 \dots n - 1\}.\:\rho \vdash e_{i} \Downarrow c_{i}
+                 \forall i \in [n].\:\rho \vdash e_{i} \Downarrow c_{i}
                }
@@
 \subsubsection*{Source language}Our source language
-(Figure~\ref{fig:srclang}) is a simple imperative language. Types
+(\figureref{srclang}) is a simple imperative language. Types
 $\psi$ consist of the base types $\sigma$, and arrays
@@
 
-The runtime semantics for the source language is shown in
-Figure~\ref{fig:srcsem}. Values $v$, runtime environments $\rho$, and
+\noindent\textbf{Source Semantics.} The runtime semantics for the source language is shown in
+\figureref{srcsem}. \TODO{Add something about claer text stand alone exec.} Values $v$, runtime environments $\rho$, and
 observations $O$ are defined as follows:
@@
 elided for space reasons.
-\divya{Does conditionals $\cond{e}{e_{1}}{e_{2}}$ have anything interesting? just checking!}
-\nc{fancy PL folks language :)}
 
@@
 ({\sc{E-Decl}}) evaluates the expression $e$ to $v$, and returns the
-updated environment $\rho[x \mapsto v]$, with empty observations. Rule
+updated environment $\rho,x \mapsto v$, with empty observations. Rule
 ({\sc{E-If}}) evaluates the guard expression, and then evaluates
-either $s_{1}$ or $s_{2}$ accordingly. $\kw{for}$ statements evaluate
+either $s_{1}$ or $s_{2}$ accordingly. The $\kw{for}$ statements evaluate
 through the internal $\kw{while}$ syntax. Specifically, the rule
-({\sc{E-For}}) updates $\rho$ with the initial counter value $n_{1}$,
+({\sc{E-For}}) appends $\rho$ with $x\mapsto n_{1}$,
 evaluates $\loops{x}{n_{2}}{s}$ to $\rho_{1}; O$, and returns
@@
 $\rho$ ($[\rho_{1}]_{\mathsf{dom}(\rho)}$) to remove the variables
-added by $s$, increments the value if $x$, and evaluates the
+added by $s$, increments the value of $x$, and evaluates the
 \kw{while} statement under this updated environment. Rule
@@
   \]
-\caption{Target language}
-\label{fig:tgtlang}
+\caption{Intermediate Language}
+\label{fig:interlang}
 \end{figure}
@@
   \begin{array}{rrcl}
-    \ftext{Wire id} & r &&\\
-    \ftext{Circuit gate} & \kappa & ::= & r \mid \kw{in}^{m}_{j} \mid \kw{add}\:\kappa_{1}\:\kappa_{2} \mid \kw{gt}\:\kappa_{1}\:\kappa_{2}\\
-    & & \mid & \kw{mux}\:\kappa\:\kappa_{1}\:\kappa_{2} \mid \widetilde{w} \rhd m\\
-    \ftext{Base value} & \widetilde{w} & ::= & c \mid \kappa\\
-    \ftext{Value} & \widetilde{v} & ::= & \widetilde{w} \mid [\overline{\widetilde{w}_{i}}]_{n}\\
-    \ftext{Runtime environment} & \widetilde{\rho} & ::= & \cdot \mid \widetilde{\rho}[x \mapsto \widetilde{v}]\\
-    \ftext{Circuit} & \chi & ::= & \cdot \mid \kw{bind}\:\kappa\:r \mid \kw{out}\:\kappa \mid \chi_{1}; \chi_{2}
+    \ftext{Wire id} & \wire &&\\
+    \ftext{Circuit gate} & \gate & ::= & \wire \mid \kw{in}^{m}_{j} \mid \kw{add}\:\gate_1\:\gate_{2} \mid \kw{gt}\:\gate_{1}\:\gate_{2}\\
+    & & \mid & \kw{mux}\:\gate\:\gate_{1}\:\gate_{2} \mid \widetilde{w} \rhd m\\
+    \ftext{Base value} & \bval & ::= & c \mid \gate\\
+    \ftext{Value} & \val & ::= & \bval \mid [\overline{\bval_{i}}]_{n}\\
+    \ftext{Runtime environment} & \widetilde{\rho} & ::= & \cdot \mid \widetilde{\rho},x \mapsto \val \\
+    \ftext{Circuit} & \crct & ::= & \cdot \mid \kw{bind}\:\gate\:\wire \mid \kw{out}\:\gate \mid \crct_{1}; \crct_{2}
   \end{array}
   \]
-\caption{Target runtime}
-\label{fig:tgtruntime}
+\caption{Circuits   \divya{Is environ misplaced?}}
+\label{fig:circuits}
 \end{figure}
 
-\begin{figure}
-  \small
-  \fbox{$\widetilde{\rho} \vdash \widetilde{e} \Downarrow \widetilde{v}$}
-  \[
-  \\
-  \begin{array}{c}
-    \inferrule*[lab={\footnotesize{S-Var}}]
-               {
-               }
-               {
-                 \widetilde{\rho} \vdash x \Downarrow \widetilde{\rho}(x)
-               }
-
-    \inferrule*[lab={\footnotesize{S-PAdd}}]
-               {
-                 \forall i \in \{1, 2\}.\:\widetilde{\rho} \vdash \widetilde{e_{i}} \Downarrow n_{i}
-               }
-               {
-                 \widetilde{\rho} \vdash \widetilde{e_{1}} +_{\mathcal{P}} \widetilde{e_{2}} \Downarrow n_{1} + n_{2}
-               }
-
-    \inferrule*[lab={\footnotesize{S-Read}}]
-               {
-                 \widetilde{\rho} \vdash x \Downarrow [\overline{\widetilde{w_{i}}}]_{n_{1}} \\\\
-                 \widetilde{\rho} \vdash \widetilde{e} \Downarrow n \quad n < n_{1}
-               }
-               {
-                 \widetilde{\rho} \vdash x[\widetilde{e}] \Downarrow \widetilde{w_{n}}
-               }\\\\
-    \inferrule*[lab={\footnotesize{S-SAdd}}]
-               {
-                 \forall i \in \{1, 2\}.\:\widetilde{\rho} \vdash \widetilde{e_{i}} \Downarrow \kappa_{i}
-               }
-               {
-                 \widetilde{\rho} \vdash \widetilde{e_{1}} +_{\mathcal{A}} \widetilde{e_{2}} \Downarrow \kw{add}\:\kappa_{1}\:\kappa_{2}
-               }
-
-               \hspace{0.3cm}
-
-    \inferrule*[lab={\footnotesize{S-SGt}}]
-               {
-                 \forall i \in \{1, 2\}.\:\widetilde{\rho} \vdash \widetilde{e_{i}} \Downarrow \kappa_{i}
-               }
-               {
-                 \widetilde{\rho} \vdash \widetilde{e_{1}} >_{\mathcal{B}} \widetilde{e_{2}} \Downarrow \kw{gt}\:\kappa_{1}\:\kappa_{2}
-               }\\\\
-               \inferrule*[lab={\footnotesize{S-SCond}}]
-               {
-                 \forall i \in \{1, 2, 3\}.\:\widetilde{\rho} \vdash \widetilde{e_{i}} \Downarrow \kappa_{i}
-               }
-               {
-                 \widetilde{\rho} \vdash \lcond{\mathcal{B}}{\widetilde{e_{1}}}{\widetilde{e_{2}}}{\widetilde{e_{3}}} \Downarrow \kw{mux}\:\kappa_{1}\:\kappa_{2}\:\kappa_{3}
-               }
-               \hspace{0.2cm}
-    \inferrule*[lab={\footnotesize{S-Coerce}}]
-               {
-                 \widetilde{\rho} \vdash \widetilde{e} \Downarrow \widetilde{v}
-               }
-               {
-                 \widetilde{\rho} \vdash \widetilde{e} \rhd m \Downarrow \widetilde{v} \rhd m
-               }
-               \\\\
-               %% \inferrule*[lab={\footnotesize{S-Arr}}]
-               %% {
-               %%   \forall i \in \{0 \dots n - 1\}.\:\widetilde{\rho} \vdash \widetilde{e_{i}} \Downarrow \widetilde{w_{i}}
-               %% }
-               %% {
-               %%   \widetilde{\rho} \vdash [\overline{\widetilde{e_{i}}}]_{n} \Downarrow [\overline{\widetilde{w_{i}}}]_{n}
-               %% }
-
-               \inferrule*[lab={\footnotesize{S-PCond}}]
-               {
-                 \widetilde{\rho} \vdash \widetilde{e} \Downarrow c\\\\
-                 c = \top \Rightarrow \widetilde{e'} = \widetilde{e_{1}}\\
-                 c = \bot \Rightarrow \widetilde{e'} = \widetilde{e_{2}}\\\\
-                 \widetilde{\rho} \vdash \widetilde{e'} \Downarrow \widetilde{v}
-               }
-               {
-                 \widetilde{\rho} \vdash \lcond{\mathcal{P}}{\widetilde{e}}{\widetilde{e_{1}}}{\widetilde{e_{2}}} \Downarrow \widetilde{v}
-               }
-               %\hspace{0.5cm}
-
-    \inferrule*[lab={\footnotesize{S-Inp}}]
-               {
-               }
-               {
-                 \widetilde{\rho} \vdash \kw{in}^{m}_{j} \Downarrow \kw{in}^{m}_{j}
-               }
-  \end{array}
-  \]
-  \\\\
-    \fbox{$\widetilde{\rho} \vdash \widetilde{c} \Downarrow \widetilde{\rho'}; \chi$}
-  \[
-  \\
-  \begin{array}{c}
-    \inferrule*[lab={\footnotesize{S-DeclC}}]
-               {
-                 \widetilde{\rho} \vdash \widetilde{e} \Downarrow \widetilde{v}\\\\
-                 \widetilde{v} = c \vee \widetilde{v} = [\overline{c_{i}}]_{n}\\\\
-                 \widetilde{\rho}_{1} = \widetilde{\rho}, x \mapsto \widetilde{v}
-               }
-               {
-                 \widetilde{\rho} \vdash \tau\:x = \widetilde{e} \Downarrow \widetilde{\rho}_{1}; \cdot
-               }
-
-               \hspace{0.5cm}
-
-    \inferrule*[lab={\footnotesize{S-DeclCkt}}]
-               {
-                 \widetilde{\rho} \vdash \widetilde{e} \Downarrow \kappa\quad
-                 \mathsf{fresh}\:r\\\\
-                 \widetilde{\rho}_{1} = \widetilde{\rho}, x \mapsto r\quad
-                 \chi = \kw{bind}\:\kappa\:r
-               }
-               {
-                 \widetilde{\rho} \vdash \tau\:x = \widetilde{e} \Downarrow \widetilde{\rho}_{1}; \chi
-               }
-
-\\\\
-    \inferrule*[lab={\footnotesize{S-DeclCktA}}]
-               {
-                 \widetilde{\rho} \vdash \widetilde{e} \Downarrow [\overline{\kappa_{i}}]_{n}\\\\
-                 \forall i \in \{0 \dots n - 1\}.\:\mathsf{fresh}\:r_{i}\\\\
-                 \widetilde{\rho}_{1} = \widetilde{\rho}, x \mapsto [r_{i}]_{n}\quad
-                 \chi = \overline{\kw{bind}\:\kappa_{i}\:r_{i}}
-               }
-               {
-                 \widetilde{\rho} \vdash \tau\:x = \widetilde{e} \Downarrow \widetilde{\rho}_{1}; \chi
-               }
-
-               \hspace{0.2cm}
-
-    \inferrule*[lab={\footnotesize{S-Out}}]
-               {
-                 \widetilde{\rho} \vdash \widetilde{e} \Downarrow \kappa
-               }
-               {
-                 \widetilde{\rho} \vdash \kw{out}\:\widetilde{e} \Downarrow \widetilde{\rho}; \kw{out}\:\kappa
-               }
-
-               \\\\
-    \inferrule*[lab={\footnotesize{S-If}}]
-               {
-                 \widetilde{\rho} \vdash \widetilde{e} \Downarrow c\\\\
-                 c = \top \Rightarrow \widetilde{s} = \widetilde{s_{1}}\\\\
-                 c = \bot \Rightarrow \widetilde{s} = \widetilde{s_{2}}\\\\
-                 \widetilde{\rho} \vdash \widetilde{s} \Downarrow \widetilde{\rho}_{1}; \chi
-               }
-               {
-                 \widetilde{\rho} \vdash \ite{\widetilde{e}}{\widetilde{s_{1}}}{\widetilde{s_{2}}} \Downarrow \widetilde{\rho}_{1}; \chi
-               }
-
-               \hspace{0.5cm}
-
-    \inferrule*[lab={\footnotesize{S-WriteCkt}}]
-               {
-                 \widetilde{\rho} \vdash x \Downarrow [\overline{r_{i}}]_{n}\quad
-                 \widetilde{\rho} \vdash \widetilde{e_{1}} \Downarrow n_{1}\\\\
-                 n_{1} < n\quad
-                 \mathsf{fresh}\:r\quad
-                 \widetilde{\rho} \vdash \widetilde{e_{2}} \Downarrow \kappa\\\\
-                 \widetilde{\rho}_{1} = \widetilde{\rho}[x \mapsto [\overline{r_{i}}]_{n}[n_{1} \mapsto r]]
-               }
-               {
-                 \widetilde{\rho} \vdash x[\widetilde{e_{1}}] := \widetilde{e_{2}} \Downarrow \widetilde{\rho}_{1}; \kw{bind}\:\kappa\:r
-               }
-
-\end{array}
-  \]
-\caption{Target semantics (selected rules)}
-\label{fig:tgtsem}
-\end{figure}
-
-\subsubsection*{Target language} Figure~\ref{fig:tgtlang} shows the
-target language of our compiler. The syntax follows that of the source
+\subsubsection*{Intermediate language} \figureref{interlang} shows the
+intermediate language of our compiler. The syntax follows that of the source
 language, except that the types and operators are \emph{labeled}. Below
@@
 $\mathcal{B}$ for using arithmetic or boolean circuits resp.). The
-form $\widetilde{e} \rhd m$ denotes coercing $\widetilde{e}$ to be
+form $\widetilde{e} \rhd m$ denotes coercing \divya{the value of} $\widetilde{e}$ to be
 $m$-secret shared. The statements $\widetilde{s}$ are analogous to
@@
 
-The target semantics models the semantics of the code output by
-our compiler (C++ code in our implementation). It computes over the
-public parts of the program, emitting a circuit that can later be
-evaluated using a cryptographic MPC protocol. The semantics also
-\emph{flattens} the arrays so that the circuits are unaware of the
-array structure. Crucially, this phase of the semantics does not have
-access to the secrets.
-
-Figure~\ref{fig:tgtruntime} shows the runtime syntax. A wire id range
-$r$ denotes a set of circuit wires that carry the value of a secret
+\aseem{The target semantics models the semantics of the code output by
+our compiler (C++ code in our implementation).}
+A program in the intermediate language is evaluated to a circuit that can be executed in the distributed setting later.
+The evaluation to a circuit computes over the public parts of the program and also \emph{flattens} the arrays so that the circuits are unaware of the array structure.
+Crucially, this phase of the semantics does not have access to the secrets.
+Below, we first provide the language for the circuits used followed by our evaluation rules.
+
+\subsubsection*{Circuits}
+\figureref{circuits} shows the circuits syntax. A wire id range
+$\wire$ denotes a set of circuit wires that carry the value of a secret
 shared variable in the output circuit (we will concretely define these
-values later as part of the circuit semantics). Circuit gates $\kappa$
-are $r$, input gates, \kw{add}, \kw{gt}, and \kw{mux} gates, and
-coerce gates $\widetilde{w} \rhd m$. Target values $\widetilde{v}$
-then consist of base values $c$ and $\kappa$, and arrays of base
-values.
-
-Figure~\ref{fig:tgtsem} shows the judgments for the target
-semantics. We first focus on the expression evaluation judgment
+values later as part of the circuit semantics). Circuit gates $\gate$
+are $\wire$, input gates, \kw{add}, \kw{gt}, and \kw{mux} gates, and
+coerce gates $\bval \rhd m$. \divya{Gates are not just gates but a circuit?} Target values $\val$
+then consist of base values $c$ and $\gate$, and arrays of base
+values. \\
+
+\divya{Start here....}
+\noindent\textbf{Evaluation to Circuit.} \figureref{eval-inter-circ} shows the judgments for the evaluation of the intermediate language to a circuit.
+We first focus on the expression evaluation judgment
 $\widetilde{\rho} \vdash \widetilde{e} \Downarrow
-\widetilde{v}$. Rules ({\sc{S-PAdd}}) and ({\sc{S-SAdd}}) illustrate
+\val$. 
+%
+Rules ({\sc{S-PAdd}}) and ({\sc{S-SAdd}}) illustrate
 the significance of the operator labels. In particular, the rule
@@
 \kw{out} circuit.
+
+\begin{figure}
+  \small
+  \fbox{$\widetilde{\rho} \vdash \widetilde{e} \Downarrow \widetilde{v}$}
+  \[
+  \\
+  \begin{array}{c}
+    \inferrule*[lab={\footnotesize{S-Var}}]
+               {
+               }
+               {
+                 \widetilde{\rho} \vdash x \Downarrow \widetilde{\rho}(x)
+               }
+
+    \inferrule*[lab={\footnotesize{S-PAdd}}]
+               {
+                 \forall i \in \{1, 2\}.\:\widetilde{\rho} \vdash \widetilde{e_{i}} \Downarrow n_{i}
+               }
+               {
+                 \widetilde{\rho} \vdash \widetilde{e_{1}} +_{\mathcal{P}} \widetilde{e_{2}} \Downarrow n_{1} + n_{2}
+               }
+
+    \inferrule*[lab={\footnotesize{S-Read}}]
+               {
+                 \widetilde{\rho} \vdash x \Downarrow [\overline{\bval_i}]_{n_{1}} \\\\
+                 \widetilde{\rho} \vdash \widetilde{e} \Downarrow n \quad n < n_{1}
+               }
+               {
+                 \widetilde{\rho} \vdash x[\widetilde{e}] \Downarrow \bval_n
+               }\\\\
+    \inferrule*[lab={\footnotesize{S-SAdd}}]
+               {
+                 \forall i \in \{1, 2\}.\:\widetilde{\rho} \vdash \widetilde{e_{i}} \Downarrow \kappa_{i}
+               }
+               {
+                 \widetilde{\rho} \vdash \widetilde{e_{1}} +_{\mathcal{A}} \widetilde{e_{2}} \Downarrow \kw{add}\:\kappa_{1}\:\kappa_{2}
+               }
+
+               \hspace{0.3cm}
+
+    \inferrule*[lab={\footnotesize{S-SGt}}]
+               {
+                 \forall i \in \{1, 2\}.\:\widetilde{\rho} \vdash \widetilde{e_{i}} \Downarrow \kappa_{i}
+               }
+               {
+                 \widetilde{\rho} \vdash \widetilde{e_{1}} >_{\mathcal{B}} \widetilde{e_{2}} \Downarrow \kw{gt}\:\kappa_{1}\:\kappa_{2}
+               }\\\\
+               \inferrule*[lab={\footnotesize{S-SCond}}]
+               {
+                 \forall i \in \{1, 2, 3\}.\:\widetilde{\rho} \vdash \widetilde{e_{i}} \Downarrow \kappa_{i}
+               }
+               {
+                 \widetilde{\rho} \vdash \lcond{\mathcal{B}}{\widetilde{e_{1}}}{\widetilde{e_{2}}}{\widetilde{e_{3}}} \Downarrow \kw{mux}\:\kappa_{1}\:\kappa_{2}\:\kappa_{3}
+               }
+               \hspace{0.2cm}
+    \inferrule*[lab={\footnotesize{S-Coerce}}]
+               {
+                 \widetilde{\rho} \vdash \widetilde{e} \Downarrow \widetilde{v}
+               }
+               {
+                 \widetilde{\rho} \vdash \widetilde{e} \rhd m \Downarrow \widetilde{v} \rhd m
+               }
+               \\\\
+               %% \inferrule*[lab={\footnotesize{S-Arr}}]
+               %% {
+               %%   \forall i \in \{0 \dots n - 1\}.\:\widetilde{\rho} \vdash \widetilde{e_{i}} \Downarrow \widetilde{w_{i}}
+               %% }
+               %% {
+               %%   \widetilde{\rho} \vdash [\overline{\widetilde{e_{i}}}]_{n} \Downarrow [\overline{\widetilde{w_{i}}}]_{n}
+               %% }
+
+               \inferrule*[lab={\footnotesize{S-PCond}}]
+               {
+                 \widetilde{\rho} \vdash \widetilde{e} \Downarrow c\\\\
+                 c = \top \Rightarrow \widetilde{e'} = \widetilde{e_{1}}\\
+                 c = \bot \Rightarrow \widetilde{e'} = \widetilde{e_{2}}\\\\
+                 \widetilde{\rho} \vdash \widetilde{e'} \Downarrow \widetilde{v}
+               }
+               {
+                 \widetilde{\rho} \vdash \lcond{\mathcal{P}}{\widetilde{e}}{\widetilde{e_{1}}}{\widetilde{e_{2}}} \Downarrow \widetilde{v}
+               }
+               %\hspace{0.5cm}
+
+    \inferrule*[lab={\footnotesize{S-Inp}}]
+               {
+               }
+               {
+                 \widetilde{\rho} \vdash \kw{in}^{m}_{j} \Downarrow \kw{in}^{m}_{j}
+               }
+  \end{array}
+  \]
+  \\\\
+    \fbox{$\widetilde{\rho} \vdash \widetilde{s} \Downarrow \widetilde{\rho'}; \chi$}
+  \[
+  \\
+  \begin{array}{c}
+    \inferrule*[lab={\footnotesize{S-DeclC}}]
+               {
+                 \widetilde{\rho} \vdash \widetilde{e} \Downarrow \widetilde{v}\\\\
+                 \widetilde{v} = c \vee \widetilde{v} = [\overline{c_{i}}]_{n}\\\\
+                 \widetilde{\rho}_{1} = \widetilde{\rho}, x \mapsto \widetilde{v}
+               }
+               {
+                 \widetilde{\rho} \vdash \tau\:x = \widetilde{e} \Downarrow \widetilde{\rho}_{1}; \cdot
+               }
+
+               \hspace{0.5cm}
+
+    \inferrule*[lab={\footnotesize{S-DeclCkt}}]
+               {
+                 \widetilde{\rho} \vdash \widetilde{e} \Downarrow \kappa\quad
+                 \mathsf{fresh}\:r\\\\
+                 \widetilde{\rho}_{1} = \widetilde{\rho}, x \mapsto r\quad
+                 \chi = \kw{bind}\:\kappa\:r
+               }
+               {
+                 \widetilde{\rho} \vdash \tau\:x = \widetilde{e} \Downarrow \widetilde{\rho}_{1}; \chi
+               }
+
+\\\\
+    \inferrule*[lab={\footnotesize{S-DeclCktA}}]
+               {
+                 \widetilde{\rho} \vdash \widetilde{e} \Downarrow [\overline{\kappa_{i}}]_{n}\\\\
+                 \forall i \in \{0 \dots n - 1\}.\:\mathsf{fresh}\:r_{i}\\\\
+                 \widetilde{\rho}_{1} = \widetilde{\rho}, x \mapsto [r_{i}]_{n}\quad
+                 \chi = \overline{\kw{bind}\:\kappa_{i}\:r_{i}}
+               }
+               {
+                 \widetilde{\rho} \vdash \tau\:x = \widetilde{e} \Downarrow \widetilde{\rho}_{1}; \chi
+               }
+
+               \hspace{0.2cm}
+
+    \inferrule*[lab={\footnotesize{S-Out}}]
+               {
+                 \widetilde{\rho} \vdash \widetilde{e} \Downarrow \kappa
+               }
+               {
+                 \widetilde{\rho} \vdash \kw{out}\:\widetilde{e} \Downarrow \widetilde{\rho}; \kw{out}\:\kappa
+               }
+
+               \\\\
+    \inferrule*[lab={\footnotesize{S-If}}]
+               {
+                 \widetilde{\rho} \vdash \widetilde{e} \Downarrow c\\\\
+                 c = \top \Rightarrow \widetilde{s} = \widetilde{s_{1}}\\\\
+                 c = \bot \Rightarrow \widetilde{s} = \widetilde{s_{2}}\\\\
+                 \widetilde{\rho} \vdash \widetilde{s} \Downarrow \widetilde{\rho}_{1}; \chi
+               }
+               {
+                 \widetilde{\rho} \vdash \ite{\widetilde{e}}{\widetilde{s_{1}}}{\widetilde{s_{2}}} \Downarrow \widetilde{\rho}_{1}; \chi
+               }
+
+               \hspace{0.5cm}
+
+    \inferrule*[lab={\footnotesize{S-WriteCkt}}]
+               {
+                 \widetilde{\rho} \vdash x \Downarrow [\overline{r_{i}}]_{n}\quad
+                 \widetilde{\rho} \vdash \widetilde{e_{1}} \Downarrow n_{1}\\\\
+                 n_{1} < n\quad
+                 \mathsf{fresh}\:r\quad
+                 \widetilde{\rho} \vdash \widetilde{e_{2}} \Downarrow \kappa\\\\
+                 \widetilde{\rho}_{1} = \widetilde{\rho}[x \mapsto [\overline{r_{i}}]_{n}[n_{1} \mapsto r]]
+               }
+               {
+                 \widetilde{\rho} \vdash x[\widetilde{e_{1}}] := \widetilde{e_{2}} \Downarrow \widetilde{\rho}_{1}; \kw{bind}\:\kappa\:r
+               }
+
+\end{array}
+  \]
+\caption{Evaluation of Intermediate Language to Circuit (selected rules)}
+\label{fig:eval-inter-circ}
+\end{figure}
 

### 2017-10-25 OaklandSubmission/eval.tex
6.4 and 6.5

@@
 \label{sec:dnneval}
-We evaluate \tool on the DNNs described in \minion~\cite{minionn} and the results are reported in Table~\ref{tab:nn}\footnote{\minion does not report the network round-trip time~\cite{minionn}.}.
+We evaluate \tool on the DNNs described in \minion~\cite{minionn} and
+report the results in Table~\ref{tab:nn}\footnote{\minion does
+  not report the network round-trip time.}.
 Our goal here is to demonstrate that \tool
-can provide performance competitive to specialized protocols for DNNs described in~\cite{secureml,cryptonets,minionn}. 
+can provide performance competitive to specialized protocols for DNNs
+described in~\cite{secureml,cryptonets,minionn}.  
 The first benchmark  is the DNN described in SecureML~\cite{secureml}.
 It has three fully connected layers with square as the activation function.
-This DNN is light-weight as it requires only arithmetic computations (additions and multiplications)
+This DNN is light-weight as it requires only arithmetic computations
+(additions and multiplications) 
 except for the {\it argmax} at the end.
 Next, we implement the DNN described in Cryptonets~\cite{cryptonets} in \tool.
-This DNN also uses square as the activation function and has one convolution (with 5 output channels) and one fully connected layer.
+This DNN also uses square as the activation function and has one
+convolution (with 5 output channels) and one fully connected layer. 
 Finally, we compare against \minion on a CNN
-with two convolutions (with 16 output channels each) and two fully connected layers.
+with two convolutions (with 16 output channels each) and two fully
+connected layers. 
 In contrast to the previous two DNNs, this DNN uses ReLU for activation and
 has significantly higher number of boolean-and gates.
-Recall, that square activation can be implemented entirely using arithmetic gates but ReLU requires boolean-and gates.
-For a complete description of these benchmarks and their accuracies,  the reader is referred to the original references.
+Recall, that square activation can be implemented entirely using
+arithmetic gates but ReLU requires boolean-and gates.  For a complete
+description of these benchmarks and their accuracies,  the reader is
+referred to the original references.
 
@@
 The column ``Model size'' is the number of 32 bit integers in the trained model.
-We observe that our performance is competitive with \minion in the LAN and the WAN settings and the lines of code required are still small.
+We observe that our performance is competitive with \minion in the LAN
+and the WAN settings and the lines of code required are still small. 
 We note that \minion and \tool implementations are both based on ABY.
-These results are surprising as \minion has a much more efficient preprocessing phase and has SIMD
+These results are surprising as \minion has a much more efficient
+preprocessing phase and has SIMD 
 (single instruction multiple data) capabilities that allow \minion to perform
@@
 focus on generic 2PC and such optimizations are beyond the scope of this paper.
-However, \minion also reports performance results on a bigger DNN with 7 convolution layers.
+However, \minion also reports performance results on a bigger DNN with
+7 convolution layers. 
 This benchmark requires pipelining and we discuss it in Section~\ref{sec:pipeeval}.
@@
 
-\subsection{Practical models}
+\subsection{State-of-the-art classifiers}
 Tensorflow~\cite{tensorflow} is a standard machine learning toolkit.
@@
 using the MNIST dataset~\cite{mnist}.
-Each image in this dataset is a greyscale $28\times 28$ image of digits 0 to 9.
-The first model that the tutorial describes is softmax regression
-that provides an accuracy of 92\%. The classifier here evaluates
+Each image in this dataset is a greyscale $28\times 28$ image of
+digits 0 to 9.
+The first model that the tutorial describes is a softmax regression
+that provides an accuracy of 92\%. The classifier evaluates
 \[
@@
 Here, $i\in [10]$, $x$ is a 784 length vector obtained from the input image,
-each $W$ is a $10\times 784$ matrix, and each $b$ is a $784$ length vector. 
-We implement this classifer in \tool and the results are shown in the first row of Table~\ref{tab:tf}.
-Since, we are not familiar with any other tools that have used this model as a benchmark,
-we only report execution time of \tool generated code.
+each $W$ is a $10\times 784$ matrix, and each $b$ is a $784$ length
+vector. \aseem{Questions on this expression: (1) where is $i$?
+  (2) Should $b$ be 10 length vector? (3) Again, should be use $\cdot$
+  for matrix multiplication
+as we did earlier to be consistent?}
+We implement this classifer in \tool and present the results in the
+first row of Table~\ref{tab:tf}.
+Since, we are not aware of any other tools that has used this
+model as a benchmark, we only report numbers for \tool.
 
@@
 This DNN is both bigger and more accurate than the DNNs presented in the previous section.
-In particular, it has an accuracy of 99.2\%\footnote{The accuracy of the integer model is 0.04\% higher than the floating-point model.}.
-We observe that this DNN can take a minute per prediction in the WAN setting and is the largest
-benchmark that we have evaluated without pipelining.
+In particular, it has an accuracy of 99.2\%\footnote{The accuracy of
+  the integer model is 0.04\% higher than the floating-point
+  model.}. \aseem{Is this footnote necessary? As a reader, I won't
+  miss it if it was not there.}
+We observe that this DNN can take a minute per prediction in the WAN
+setting and is the largest benchmark that we have evaluated without
+partitioning.
 
@@
 
-Before discussing pipelining, we show \bonsai results on three standard datasets: character recognition (Chars4k~\cite{campos}, accuracy 74.71\%), text recognition (USPS~\cite{hull}, accuracy 94.4\%), and object categorization (WARD~\cite{yang}, accuracy 95.7\%).
-The \bonsai models are already over integers and no port from floating-point is required.
-We implement the trained classifiers in \tool for all benchmarks in~\cite{bonsai}
+We next present \bonsai~\cite{bonsai} results on three
+standard datasets: character recognition (Chars4k~\cite{campos},
+accuracy 74.71\%), text recognition (USPS~\cite{hull}, accuracy
+94.4\%), and object categorization (WARD~\cite{yang}, accuracy
+95.7\%). 
+The \bonsai models are already over integers and no porting from
+floating-point is required.
+We implement the trained classifiers in \tool for all the benchmarks
+from~\cite{bonsai},
 and show the represenative results in Table~\ref{tab:bonsai}.
-Out of all benchmarks in~\cite{bonsai}, the dataset WARD requires the largest model.
-The column ``depth'' in Table~\ref{tab:bonsai} shows the depth of the tree used by \bonsai. The size of \tool
-program grows with the depth of the tree as the straightforward \tool implementation requires a 
-loop for each layer of the tree.
+Out of all the benchmarks from~\cite{bonsai}, the dataset WARD
+requires the largest model. 
+The column ``depth''  shows the depth of the
+tree used by \bonsai. The size of \tool 
+program grows with the depth of the tree, as the straightforward \tool
+implementation requires a loop for each layer of the
+tree. \aseem{Should we remark here that once we have functions in the
+  language, LOC would come down, perhaps in the footnote?}
 
-To summarize, we have demonstrated that \tool is expressive enough to implement various
-state-of-the-art classifiers. We discuss scalability next. 
+To summarize, by providing first 2PC implementations of
+state-of-the-art classifiers, we have demonstrated the expressiveness
+of \tool. We discuss scalability next.
 %dataset | size | gates | and | mul | depth | length | comm | timeL | timeW
@@
 
-\subsection{Pipelining}
+\subsection{Partitioning}
 \label{sec:pipeeval}
-The largest benchmark of \minion~\cite{minionn} is a DNN for CIFAR-10 dataset~\cite{cifar}.
-The classifier's task is to categorize colored ($32\times 32$) images into 10 classes. A secure evaluation of this DNN needs more memory than what is available
-on our machines. Therefore, we use pipelining and divide the computation into seven stages.
-The first step does a convolution with 64 output channels and a ReLU activation.
-The next four stages together perform a convolution that involves multiplying a $64\times 576$ matrix with a
-$576\times 1024$ matrix. The sixth stage performs a ReLU and a convolution.
-The final stage has four convolutions, five ReLUs, and a fully connected layer.
-The total number of lines of \tool code for this benchamrk is 336 lines.
+The largest benchmark of \minion~\cite{minionn} is a DNN for CIFAR-10
+dataset~\cite{cifar}. 
+The classifier's task is to categorize colored ($32\times 32$) images
+into 10 classes. A secure evaluation of this DNN needs more memory
+than what is available
+on our machines \aseem{Is there limitation machine RAM or ABY
+  limit?}. Therefore, we use partitioning and divide the
+computation into seven stages. 
+The first step does a convolution with 64 output channels and a ReLU
+activation. The next four stages together perform a convolution that
+involves multiplying a $64\times 576$ matrix with a 
+$576\times 1024$ matrix. The sixth stage performs a ReLU and a
+convolution. The final stage has four convolutions, five ReLUs, and a
+fully connected layer. The total number of lines of \tool code for
+this benchamrk is 336 lines. \aseem{Would it be better if we had
+  functions, for loop expressions etc.? If so, we should add a
+  footnote.}
 
-The results in Table~\ref{tab:cifar} show the end-to-end results as well as the statistics for the sixth stage (the heaviest stage). 
-The number of gates are in millions, hence the suffix `m' in the last three columns.
-For this DNN, \minion takes 544 seconds and communicates 9272 Mb. 
-Like the results in Table~\ref{tab:nn}, \tool generated implementations are competitive with \minion here as well.
-Therefore, we believe that (with pipelining) \tool generated generic 2PC implementations can scale to arbitrary sized computations  while maintaining performance
-competitive with existing specialized protocols for secure prediction.
-In particular, for a large enough DNN, \minion would run out of memory
-but a pipelined \tool implementation would still succeed.
+Table~\ref{tab:cifar} shows the end-to-end numbers as
+well as the numbers for the sixth (the heaviest) stage. 
+The number of gates are in millions, hence the suffix `m' in the last
+three columns.
+For this DNN, the specialized \minion protocol takes 544 seconds and
+communicates 9272 Mb.
+As with Table~\ref{tab:nn}, \tool generated generic 2PC protocol is
+competitive with \minion here as well. Therefore,
+we believe that with partitioning, \tool can scale to arbitrary sized
+computations while maintaining performance
+competitive with existing specialized protocols. In particular, for a
+large enough DNN, \minion would run
+out of memory but a pipelined \tool implementation would still
+succeed. \aseem{Last sentence is speculative? Should we say \minion
+  would or could or might?}
+
 \begin{table}
@@
 
- \caption{Pipelining results for CIFAR-10. \minion takes 544 seconds and communicates 9272 Mb to evaluate this DNN.}
+ \caption{Partitioning results for CIFAR-10. \minion takes 544 seconds and communicates 9272 Mb to evaluate this DNN.}
  \label{tab:cifar} 

### 2017-10-25 OaklandSubmission/eval.tex
finishing pass on 6

@@
 \label{sec:dnneval}
-We evaluate \tool on the DNNs described in \minion~\cite{minionn} and
-report the results in Table~\ref{tab:nn}\footnote{\minion does
-  not report the network round-trip time.}.
-Our goal here is to demonstrate that \tool
-can provide performance competitive to specialized protocols for DNNs
-described in~\cite{secureml,cryptonets,minionn}.  
-The first benchmark  is the DNN described in SecureML~\cite{secureml}.
+We evaluate \tool on the DNNs described in SecureML~\cite{secureml},
+Cryptonets~\cite{cryptonets}, and CNN from \minion~\cite{minionn}. For
+comparison, we consider their implementations from
+\minion~\cite{minionn}, which outperforms their previous
+implementations. Table~\ref{tab:nn} shows the
+results\footnote{\minion does not report the network round-trip time.}.
+We note that for each of these DNNs, \minion provides a
+specialized protocol, while \tool uses a generic 2PC protocol
+(auto) generated from high-level code.
+
+The first benchmark  is the DNN described in
+SecureML~\cite{secureml}.
 It has three fully connected layers with square as the activation function.
@@
 (additions and multiplications) 
-except for the {\it argmax} at the end.
-Next, we implement the DNN described in Cryptonets~\cite{cryptonets} in \tool.
+except for the $\mathsf{argmax}$ at the end.
+Next, we implement the DNN described in Cryptonets~\cite{cryptonets}
+in \tool.
 This DNN also uses square as the activation function and has one
 convolution (with 5 output channels) and one fully connected layer. 
-Finally, we compare against \minion on a CNN
-with two convolutions (with 16 output channels each) and two fully
-connected layers. 
-In contrast to the previous two DNNs, this DNN uses ReLU for activation and
+Finally, we implement CNN from \minion, that has
+two convolutions (with 16 output channels each) and two fully
+connected layers.  In contrast to the previous two DNNs, it uses
+ReLU for activation and
 has significantly higher number of boolean-and gates.
@@
 arithmetic gates but ReLU requires boolean-and gates.  For a complete
-description of these benchmarks and their accuracies,  the reader is
-referred to the original references.
+description of these benchmarks and their accuracies, we refer the
+reader to the original references.
 
-For the benchmarks in Table~\ref{tab:nn}, \minion
-outperforms SecureML and Cryptonets and columns 2 and 3 show the time taken by \minion. 
-The column ``Model size'' is the number of 32 bit integers in the trained model.
-We observe that our performance is competitive with \minion in the LAN
-and the WAN settings and the lines of code required are still small. 
+In Table~\ref{tab:nn}, columns 2 and 3 show the time taken by \minion.
+The column ``Model size'' is the number of 32 bit integers in the
+trained model.
+We observe that our performance is competitive with specialized
+\minion protocols, for both LAN and the WAN settings. Further, lines
+of \tool source code required is still small. 
 We note that \minion and \tool implementations are both based on ABY.
@@
 preprocessing phase and has SIMD 
-(single instruction multiple data) capabilities that allow \minion to perform
-matrix operations efficiently. Our language and compiler 
-focus on generic 2PC and such optimizations are beyond the scope of this paper.
-However, \minion also reports performance results on a bigger DNN with
-7 convolution layers. 
-This benchmark requires pipelining and we discuss it in Section~\ref{sec:pipeeval}.
+(single instruction multiple data) capabilities that allow it to perform
+matrix operations efficiently. In contrast, \tool focuses on generic
+2PC, and such optimizations are part of our future work.
+\minion also reports performance results on a bigger DNN with 7
+convolution layers. In \tool, this benchmark requires partitioning and
+we discuss it in Section~\ref{sec:pipeeval}.
 %net | minionT | minionC | timeL | timeW | comm | gates | and | mul | length
@@
 tree. \aseem{Should we remark here that once we have functions in the
-  language, LOC would come down, perhaps in the footnote?}
+  language, LOC would come down, perhaps in the footnote? Else a PL
+  reviewer might scoff at it :).}
 
@@
 than what is available
-on our machines \aseem{Is there limitation machine RAM or ABY
+on our machines \aseem{Is the limitation machine RAM or ABY
   limit?}. Therefore, we use partitioning and divide the

### 2017-10-25 OaklandSubmission/eval.tex
nits

@@
 prediction tasks, and get performance competitive to or better than the
-specialized protocols. \aseem{This last claim is a bit
-  tricky. Their claim was specifically about Yao-based. Is our
-  output protocol also Yao-based? If not, then we are not comparing
-  apples to apples.}
+specialized protocols.
 
@@
 computations are infeasible with Yao-based 2PC.
-\aseem{Same comment as before, if our output protocols are not
-  Yao-based, then it is an unfair comparison.} Again, we observe that
+Again, we observe that
 the performance of \tool generated generic 2PC

### 2017-10-25 OaklandSubmission/impl.tex
pass over sec 5.

@@
 library~\cite{aby}. We use the default options of ABY except 
-for the maximum number of gates. Our benchmarks require us to set this value to a
-hundred million.
-ABY provides facilities to use multiple threads\footnote{The protocol has an offline phase and an online phase.
+for the maximum number of gates. \nc{What are these options? Do we want to spell it out?} Our benchmarks require us to set this value to a
+hundred million.\nc{I dont fully follow this.
+Why require us to set it to 100 million?
+Are we saying that because of the size of data etc and type of protocol generated by our benchmark, we limit the max gates to 100 million (when ABY supports up to 4 billion) so that we dont run into memory issues?} ABY provides facilities to use multiple threads\footnote{The protocol has an offline phase and an online phase.
 Although multiple threads can be used in the offline phase, the online phase is single threaded.}.
-We use at most four threads in our evaluation.
+We use at most four threads in our evaluation.\nc{ABY did say that more than 4 threads didnt help them, right? Should we say that this is why we use 4 threads in our evaluations?}
 
@@
 as $x > 0 ? A[1] : A[0]$. In general, for an array of size $2^n$, we need to write $2^{n-1}$ multiplexers
-in \tool. This design choice of making secret indexing inconvinient to implement was deliberate.
+in \tool. This design choice of making secret indexing inconvenient to implement was deliberate.
 Accessing arrays at secret locations is a powerful but expensive primitive
 and we want the developers to perform this operation sparingly.
+\nc{Should we say here that if we did not do it this way, then we would either be insecure or would have to use something like ObliVM that will hide locations accessed?}
 
@@
 bool and vice versa.
-The implemntation of \tool performs several optimizations to reduce these.
-For example, on each coercion, \tool memoizes the pair of arithmetic share and boolean share involved in the coercion.
+The implementation of \tool performs several optimizations to reduce these.
+For example, on each coercion, \tool memorizes the pair of arithmetic share and boolean share involved in the coercion.
 \tool invalidates such pairs when the variables corresponding to the shares are overwritten by assignments.
 In subsequent coercions, \tool can reuse valid pairs (if available) instead of recomputing them afresh.
+
+\divya{
+\begin{enumerate}
+\item add all the expressions that our language supports such as mult, bit-wise and and anything else.
+\item Say that we compile to source to intermediate and use inter to circuit of ABY.. We can decide if we want to say this.
+\end{enumerate}
+}

### 2017-10-25 OaklandSubmission/ld.tex
ld first para

@@
 \newcommand{\kw}[1]{{\lstinline[basicstyle=\small\color{blue}]{#1}}}
+%\newcommand{\kw}[1]{{\ensuremath{\mathtt{#1}}}}
 \newcommand{\ftext}[1]{\text{\small{#1}}}
@@
 
-In this section we prove the correctness and security of our compiler.
+%In this section we prove the correctness and security of our compiler.
+%%
+%We first formalize the source and target languages. Our source runtime
+%semantics is a model of the ideal, trusted third-party semantics, and
+%generates observations corresponding to the values revealed to the
+%parties.
+%%
+%The target language semantics (a model of the C++ code generated by
+%our compiler implementation) ``computes away'' the public parts and
+%arrays from the compiled program, generating a secure computation
+%circuit. Crucially, this semantics does not have access to the secret
+%inputs--those are processed by the secure computation circuit.
+%%
+%Finally, we formalize the circuit semantics that computes the
+%generated circuit, and like the source semantics, emits observations
+%corresponding to the values revealed to the parties.
 %
-We first formalize the source and target languages. Our source runtime
-semantics is a model of the ideal, trusted third-party semantics, and
+%We then present the compilation rules. To prove the correctness
+%of our compiler, we prove that it preserves the observations. For
+%security of our compiler, we reduce the security argument to the
+%security of the cryptographic protocol used to compute the secure
+%computation circuit.
+%\nc{Should we also say that security relies on correctness? (they go together)}
+%We present only selected parts of our formalization for space
+%reasons. Full definitions and proofs can be found in the supplementary
+%material submitted along with the paper.
+
+In this section we prove correctness and security of our compiler \tool.
+We formalize our source, intermediate and circuit languages. 
+An example of a program in our source language is provided in \figureref{ex-sml}.
+%Our source runtime semantics is a model of the ideal, trusted third-party semantics, and
+%generates observations corresponding to the values revealed to the
+%parties.
+Our source runtime semantics describe the standard execution semantics of the source programs and
 generates observations corresponding to the values revealed to the
@@
 %
-The target language semantics (a model of the C++ code generated by
-our compiler implementation) ``computes away'' the public parts and
-arrays from the compiled program, generating a secure computation
-circuit. Crucially, this semantics does not have access to the secret
-inputs--those are processed by the secure computation circuit.
+We present the compilation rules that type check a program in the source language and generates a program in the intermediate language.
+This intermediate language models programs such as \figureref{ex-aby} generated by our compiler implementation.
 %
-Finally, we formalize the circuit semantics that computes the
-generated circuit, and like the source semantics, emits observations
-corresponding to the values revealed to the parties.
-
-We then present the compilation rules. To prove the correctness
-of our compiler, we prove that it preserves the observations. For
-security of our compiler, we reduce the security argument to the
-security of the cryptographic protocol used to compute the secure
-computation circuit.
-\nc{Should we also say that security relies on correctness? (they go together)}
+We evaluate the program in our intermediate language  to a circuit by ``computing away'' the public parts and the arrays.
+Crucially, this step does not have access to the secret inputs--those are processed by the distributed circuit semantics that model the \mpc back-end.
+%
+Here, the generated circuit is evaluated in the distributed setting, where the parties run an interactive protocol.
+This step, like  the source semantics, emits observations corresponding to the values revealed to the parties.
+%Finally, we formalize the distributed circuit semantics (of the cryptographic back-end) that computes the generated circuit in the distributed setting, and like  the source semantics, emits observationscorresponding to the values revealed to the parties.
+
+To prove the correctness of \tool, we prove that  the observations in source semantics and distributed circuit semantics are identical (see \theoremref{correctness}). 
+We combine this correctness and the security of the \mpc back-end to argue security of the implementations generated by \tool (see \theoremref{security}).
+%For the security of our compiler, we rely on the correctness guarantee and  reduce the security to the security of the underlying cryptographic protocol used to compute the generated circuit (see \theoremref{security}).
 We present only selected parts of our formalization for space
-reasons. Full definitions and proofs can be found in the supplementary
-material submitted along with the paper.
-
-\divya{
-In this section we prove the correctness and security of our compiler.
-We formalize the source, the intermediate and the circuit languages. 
-Our source runtime semantics is a model of the ideal, trusted third-party semantics, and
-generates observations corresponding to the values revealed to the
-parties.
-%
-We present the compilation rules that type check a program in the source language and compiles it to a program in the intermediate language.
-%
-We evaluate the program in our intermediate language (a model of the C++ code generated by our compiler implementation) to a circuit by ``computing away'' the public parts and the arrays.
-Crucially, this step does not have access to the secret inputs--those are processed by the secure computation circuit.
-%
-Finally, we formalize the distributed circuit semantics (of the cryptographic back-end) that computes the generated circuit in the distributed setting, and like  the source semantics, emits observations
-corresponding to the values revealed to the parties.
-
-To prove the correctness of our compiler, we prove that it preserves the observations in source semantics and distributed circuit semantics (see \theoremref{correctness}). 
-For the security of our compiler, we rely on the correctness guarantee and  reduce the security to the
-security of the underlying cryptographic protocol used to compute the generated circuit (see \theoremref{security}).
-We present only selected parts of our formalization for space
-reasons. \TODO{Full definitions and proofs can be found in the supplementary
-material submitted along with the paper.}
-}
+reasons. Full definitions and proofs can be found \cite{anonymized}.
 
@@
                  c = \bot \Rightarrow e' = e_{2}\\\\
-                 \rho \vdash e \Downarrow c'
+                 \rho \vdash e' \Downarrow c'
                }

### 2017-10-25 OaklandSubmission/ld.tex
ld

@@
     \ftext{Circuit gate} & \gate & ::= & \wire \mid \kw{in}^{m}_{j} \mid \kw{add}\:\gate_1\:\gate_{2} \mid \kw{gt}\:\gate_{1}\:\gate_{2}\\
-    & & \mid & \kw{mux}\:\gate\:\gate_{1}\:\gate_{2} \mid \widetilde{w} \rhd m\\
+    & & \mid & \kw{mux}\:\gate\:\gate_{1}\:\gate_{2} \mid \bval \rhd m\\
     \ftext{Base value} & \bval & ::= & c \mid \gate\\

### 2017-10-25 OaklandSubmission/intro.tex
cp for 1.0

@@
 
-Today it is hard for developers to create secure implementations using cryptographic techniques.
-Typical developers lack an understanding of cryptographic protocols
-and cannot be expected to implement them correctly and efficiently without tool
-support.
-In an ideal world, a developer declares the functionality to be implemented
-in a general purpose programming language and an automatic tool takes care
-of all the cryptography and generates a secure and efficient implementation  of the functionality. 
-This paper provides a language-based solution to make
-secure  computation  accessible to the software developer community.
+Today it is hard for the developers to program secure applications
+using cryptographic techniques. Typical developers lack a deep
+understanding of the cryptographic protocols, and therefore cannot be
+expected to use them correctly and efficiently on their own.
+Ideally, a developer would declare the functionality in a general
+purpose, high-level programming language and a
+tool, e.g. a compiler, would generate an efficient protocol that
+implements the functionality securely, while hiding the cryptography
+behind-the-scenes. This paper presents such a framework for Secure
+Two-party Computation (\mpc).
 
@@
 
-Secure two-party computation (\mpc) \cite{yao,gmw} is a powerful
-cryptographic tool that allows two mutually distrusting parties
-to compute a joint function of their secret inputs such as {\em secure prediction} \cite{shafindss,wu,barni,minionn,secureml}.
-Here,  one party holds the propriety classifier for disease, genomics or spam detection and the other party
-holds the private input and the second party learns the classification label such as the presence of the disease. Since its introduction
-\mpc in 1980's, there has been a long sequence of work that has
-transformed \mpc from a mere theoretical tool to something practically
-efficient and usable. Hence, there have been efforts to make \mpc
-accessible to non-specialist programmers. To understand the
-state-of-the-art, let us consider the following 
-example that is the sub-computation in many classification tasks.
+\mpc~(\cite{yao,gmw}) is a powerful
+cryptographic technique that allows two mutually distrusting parties
+to compute a joint function of their secret inputs in a way that both
+the parties learn nothing about the inputs of each other beyond what
+is revealed by their (possibly different) outputs. For example, \mpc
+can be used for {\em secure
+prediction}~(\cite{shafindss,wu,barni,minionn,secureml}),
+where one party (the server) holds a proprietary classifier to predict
+a label (e.g. a disease, genomics, or spam detection), and the other
+party (the client) holds a private input that it wants to run the
+classifier on. Using \mpc guarantees that the server learns nothing
+about the client's input or output, and that the client learns nothing
+about the server model, beyond what is revelaed by the output
+label.
 
-Suppose a developer wants to compute $w^Tx >b$ securely using
-\mpc, where $w$ and $x$ are two vectors, $b$ is a scalar, and $w^Tx$ denotes the inner product of $w$ and $x$.  With the current techniques, he has three options. \aseem{We should
-  use she for the programmer instead of he?}
-\begin{tiret}
-\item He can code up the functionality in a programmer friendly
-  domain-specific language such as
-  \cite{lambdaps,wysteria,oblivm,smcl,cbmcgc} that would automatically
-  compile it to a \mpc protocol. The downside is that all of these
-  works build on a cryptographic back-end that takes as input a functionality expressed as either a boolean circuit \cite{yao,gmw} or an arithmetic circuit \cite{homo}.
-  The efficiency of the generated \mpc protocol is hence, bounded by the possibility of efficiently representing the functionality in one of these circuit representations.
-For instance, multiplication of two $\ell$-bit integers can be expressed as a boolean circuit of size  $O(\ell^2)$ or an arithmetic circuit with 1 multiplication gate.
-Hence, for better efficiency, we would like to securely compute $w^Tx$ using an arithmetic circuit and the comparison with $b$ using a boolean circuit. \divya{that is not supported by above systems.}
-In fact, most interesting functions require a mix of arithmetic and boolean computations for efficiency.
+Since its introduction in 1980s, there has been a long sequence
+of work that has transformed \mpc from a mere theoretical concept to
+something practically efficient and usable. To understand the
+state-of-the-art, let us consider the following example used by many
+secure prediction algorithms. Suppose Alice wants to write
+a \mpc protocol to securely
+compute $w^Tx >b$. Here $w$ and $x$ are two vectors (inputs of the two
+parties), $b$ is a scalar, $\cdot^{T}$ is the matrix transpose
+operator, and $w^Tx$ denotes the inner product of
+$w^{T}$ and $x$.  With the current state-of-the-art, Alice has three
+options:
 
-\item To address this issue of efficient representation, recently Demmler et al.
-\cite{aby} gave a cryptographic protocol for \mpc where the functionality can be expressed as a circuit with a mix of both arithmetic and boolean parts.
-Hence, the second option is to write the functionality in the system developed in \cite{aby}.
-However, in their framework, the programmer is required to write circuits consisting of  arithmetic and boolean gates along with appropriate
-inter-conversion gates.
-This requires the programmer to be aware of
-trade-offs between arithmetic and boolean cryptographic
-schemes as well as the complex inter-conversion costs between arithmetic and boolean, and manually split the functionality into such parts accordingly.  
-In short, as is also mentioned by Demmler et al., the current system  is not suitable
-to be used by non-specialist programmers without security expertise.
+%\begin{tiret}
+%\item
 
-\item Finally, the third option is to hire a cryptographer to design a specialized efficient protocol for this functionality such as \cite{shafindss,wu,minionn}. 
-\end{tiret}
-In our work, we achieve the best of all the above options.
+She can program the computation in one of the several programmer
+friendly, domain-specific languages (such
+as~\cite{lambdaps,wysteria,oblivm,smcl,cbmcgc}) that would
+automatically compile it to a \mpc protocol. However, all of these
+frameworks use cryptographic backends that take as input
+the computation expressed as either a boolean
+circuit~(\cite{yao,gmw}) or an arithmetic circuit~(\cite{gentry}).
+%
+The efficiency of the generated \mpc protocol is thus bounded by the
+efficiency of representing the computation in one of
+these representations. For instance, multiplication of two
+$\ell$-bit integers can be either expressed as a boolean circuit of
+size $O(\ell^2)$, or as an arithmetic circuit with 1 multiplication
+gate.
+%
+Hence, for better efficiency, Alice would ideally like to compute
+$w^Tx$ using an arithmetic circuit, and the comparison with $b$ using
+a boolean circuit.
+%
+Unfortunately, none of the above frameworks support combinations of
+arithmetic and boolean circuits, and using different tools for
+different parts of the computation is cumbersome and error-prone.
+
+%% In fact, most interesting functions require a mix of arithmetic and
+%% boolean computations for efficiency.
+
+%\item
+
+Alternatively, Alice can use a tool such as ABY (Demmlet et
+al.~\cite{aby}) that allows
+the computation to be expressed as a combination of arithmetic and
+boolean circuits. However, here the programming interface is quite
+low-level -- the programmer is required to first manually split the
+computation into arithmetic and boolean components, and then write the
+circuits for all the componenets manually, including the appropriate
+inter-conversion gates between them. Clearly, writing correct
+and efficient protocols in such a framework is beyond an average
+programmer who does not understand the various trade-offs between
+arithmetic and boolean circuits, and even for an expert cryptographer,
+writing large computations in such a framework can be tedious and
+error-prone (a sentiment also echoed by Demmler et al.~\cite{aby}).
+
+%\item
+The third option for Alice is to earn a PhD in cryptography and design
+specialized, efficient protocols (similar to
+~\cite{shafindss,wu,minionn}) for her tasks.
+%\end{tiret}
+%In our work, we achieve the best of all the above options.
 

### 2017-10-25 OaklandSubmission/eval.tex

@@
 \label{sec:eval}
-We evaluate \tool on the problem of {\it secure prediction}, where one
+We evaluate \tool on a variety of problems that can be thought to fall under the umbrella of \nc{saying it like this so that it does not seem we are evaluating on one problem.} {\it secure prediction}, where one
 party (the server) has a machine learning model, and the other party
@@
 
-We first implement the benchmarks from Bost et al.~\cite{shafindss}
+To begin, we first implement the benchmarks from Bost et al.~\cite{shafindss}
 and \minion~\cite{minionn} (both of which study the same setting),
@@
 Further, given a vector $x\in\R^d$, we say $\mathsf{argmax}\ x = i$ if
-$x_i = \mathsf{max}\ \{x_1,\ldots,x_d\}$.)
+$x_i = \mathsf{max}\ \{x_1,\ldots,x_d\}$.) \nc{Why is notation in brackets?}
 
@@
 data point which is a tuple of height and weight of an individual,
-a classifier can predict a label ``male'' or ``female''. The {\it
+a classifier might be able to predict a label as ``male'' or ``female''\nc{Is this less sexist?}. The {\it
   accuracy} of a classifier refers to the fraction of data points that
@@
 \subsubsection*{Standard classifiers}
-A binary linear classifier is one of the simplest classifiers. Here,
+A {\em binary linear classifier} is one of the simplest classifiers. Here,
 the input is a data point $x\in\R^d$,
@@
 $C_w\equiv w^Tx>0$, where $\cdot^T$ denotes the matrix transpose
-operator. This classifier requires $d$ multiplications, $d-1$
+operator. \nc{it might be obvious, but should we say $x > y$ returns 1 if $x>y$ and 0 otherwise?} This classifier requires $d$ multiplications, $d-1$
 additions, and a comparison.
 
-A more interesting classifier is Na\"{i}ve Bayes that predicts labels
+A more interesting classifier is {\em Na\"{i}ve Bayes} that predicts labels
 from the set $[n]$.
@@
 \]
-This classifier requires $ndF$ comparisons and additions. \divya{Note
-  that computing $T(i)(j)(x_j)$ requires a secret look-up in matrix
-  $T$  based on secret input $x_j$ resulting in $F$ comparisons (see
-  \sectionref{}).} \cmmt{Is there a +1 etc missing or is this exact?}
+This classifier requires $ndF$ comparisons and additions.
+Note here  that computing $T(i)(j)(x_j)$ requires a secret look-up in matrix  $T$  based on secret input $x_j$ resulting in $F$ comparisons using multiplexers (see
+  Section~\ref{sec:impl}).
+\nc{Made DG's comment as part of regular text - worth point out}.
+\cmmt{Is there a +1 etc missing or is this exact?} \nc{By our calculation in Sec 5, it seems that this should be $F/2$ multiplexers}
 
-A decision tree of size $N$ takes as input an $x\in\R^d$, and the
+A {\em decision tree} of size $N$ takes as input an $x\in\R^d$, and the
 model consists of a binary tree of depth $d$ with a boolean predicate
@@
 so then is $\circ$ defined OR is it applying $f$ pointwise to a vector?}
-
+\nc{(1) I agree we should do that to be consistent with before.
+(2) I assumed $f()$ is computed to give a vector somehow and $I_j(x)W_j^TZx$ also gives a vector somehow and then we do $\circ$; I might be wrong.
+(3) Hmm..
+I thought it was a vector, but I dont know how.
+(4) Can we specify somewhere the dimensions of these matrices?}
 \[
@@
 data set, and ECG (electrocardiogram) classification data
-from~\cite{barni}.
+from~\cite{barni}. \nc{Say here that the training is done on cleartext data?}
 
@@
 opposed to writing low-level circuits. \aseem{They write circuits,
-  right?} Our execution time is dominated by the multiplications and
+  right?} \nc{Actually, they have a specific protocols for dot product and comparison and also say that their developer effort is minimal as they have to write two function calls to these protocols :).} Our execution time is dominated by the multiplications and
 boolean-ands.
@@
 \subsubsection*{Decision trees}
-Our decision trees numbers in Table~\ref{tab:dt} further validate
-this claim. We compare with Wu et al.~\cite{wu} (columns 3 and 4) who
+Our decision tree numbers in Table~\ref{tab:dt} further validate
+this claim. We compare with the recent work of Wu et al.~\cite{wu} (columns 3 and 4) who
 describe specialized protocols for trees and forests (that perform
@@
 prediction tasks. \aseem{I don't understand this last sentence. Why is
-  ours a better baseline?}
+  ours a better baseline?} \nc{This last line is more a comment for a ``conclusion'' and is a bit out of place here.
+I guess we want to say that till recently if generic 2PC using Yao was not feasible, then people had to come up with specialized protocols and didnt have any benchmark or had a poor benchmark (of Yao) to compare their 2PC with.
+Moving forward, our language generated 2PC can be the baseline benchmark for all 2PC protocols.}
 
@@
 \begin{tabular}{c|c|c|c|c |c|c|c|c|c|c}
-Dataset & $d$  & Prev. time (s) & Prev. comm (kb) & LAN time (s) & WAN time (s) & Comm. (kb)  & \#And & \#Mul & \#Gates & LOC\\
+Dataset & $d$  & Prev. time (s) & Prev. Comm (kb) & LAN time (s) & WAN time (s) & Comm. (kb)  & \#And & \#Mul & \#Gates & LOC\\
 \hline

### 2017-10-25 OaklandSubmission/intro.tex
nit

@@
 The third option for Alice is to earn a PhD in cryptography and design
-specialized, efficient protocols (similar to
+specialized, efficient \mpc protocols (similar to
 ~\cite{shafindss,wu,minionn}) for her tasks.

### 2017-10-25 OaklandSubmission/intro.tex
comment

@@
 
+\aseem{Add to DSLs: Fairplay, TASTY, VIFF, secrec (sharemind), what
+about Elaine's languages? PCF?}
 She can program the computation in one of the several programmer

### 2017-10-26 OaklandSubmission/ld.tex
till compilation rules

@@
 \TODO{Later, in implementation we can write that we use ABY for compiling intermediate to target also.}
+
+\TODO{Use mult op instead of add?
+arguments of using arithmetic are clearer there; though we also use arithmetic always for add also.
+Or write both in exp and write rule for mul; how about bit-wise And?
+shifts?
+Say that more exp written in implementation}
 
@@
 
-\TODO{Do we want to write things in logical order or order in paper?}
+%\TODO{Do we want to write things in logical order or order in paper?}
 
@@
     \ftext{Constant} & c &::=& n \mid \top \mid \bot\\
-    \ftext{Expression} & e &::=& c \mid x \mid e_{1} + e_{2} \mid e_{1} > e_{2} \mid \cond{e_{1}}{e_{2}}{e_{3}}\\
+    \ftext{Expression} & e &::=& c \mid x \mid e_{1} \mulop e_{2} \mid e_{1} > e_{2} \mid \cond{e_{1}}{e_{2}}{e_{3}}\\
     & &\mid& [\overline{e_{i}}]_{n} \mid x[e] \mid \kw{in}_{j}\\
@@
 
-    \inferrule*[lab={\footnotesize{E-Add}}]
+    \inferrule*[lab={\footnotesize{E-Mult}}]
                {
@@
                {
-                 \rho \vdash e_{1} + e_{2} \Downarrow n_{1} + n_{2}
+                 \rho \vdash e_{1} \mulop e_{2} \Downarrow n_{1} \mulop n_{2}
                }
@@
 
-\subsubsection*{Source language}Our source language
-(\figureref{srclang}) is a simple imperative language. Types
+\subsubsection*{Source language} Our source language
+ is a simple imperative language. \divya{We present the core of our language in \figureref{srclang} but our implementation is more expressive (see \sectionref{impl}).} Types
 $\psi$ consist of the base types $\sigma$, and arrays
@@
 integer constants $n$, \kw{bool} constants $\top$ and $\bot$,
-variables $x$, binary operations $e_{1} + e_{2}$ and $e_{1} > e_{2}$,
+variables $x$, binary operations $e_{1} \mulop e_{2}$ and $e_{1} > e_{2}$,
 conditionals $\cond{e}{e_{1}}{e_{2}}$, array literals 
@@
 parties. The \kw{while} statement is an internal syntax that is not
-exposed to the programmer.
+exposed to the programmer. \\
 
@@
 environment~$\rho$. Rule ({\sc{E-Var}}) looks up the value of $x$ in
-the environment. Rule ({\sc{E-Add}}) inductively evaluates $e_{1}$ and
-$e_{2}$, and evaluates to their sum. Rule ({\sc{E-Read}})
+the environment. Rule ({\sc{E-Mult}}) inductively evaluates $e_{1}$ and
+$e_{2}$, and evaluates to their product. Rule ({\sc{E-Read}})
 evaluates an array read operation. It first evaluates $x$ to an array
@@
     \ftext{Type} & \tau &::=& \sigma^{\ell} \mid \sigma^{\ell}[n]\\
-    \ftext{Expression} & \widetilde{e} &::=& c \mid x \mid \widetilde{e_{1}} +_{\ell} \widetilde{e_{2}} \mid \widetilde{e_{1}} >_{\ell} \widetilde{e_{2}} \\
+    \ftext{Expression} & \widetilde{e} &::=& c \mid x \mid \widetilde{e_{1}} \mulop_{\ell} \widetilde{e_{2}} \mid \widetilde{e_{1}} >_{\ell} \widetilde{e_{2}} \\
     & & \mid & \lcond{\ell}{{\widetilde{e}}}{{\widetilde{e_{1}}}}{{\widetilde{e_{2}}}} \mid x[\widetilde{e}] \mid [\overline{\widetilde{e_{i}}}]_{n} \mid \kw{in}^{m}_{j} \mid \widetilde{e} \rhd m\\
@@
 \end{figure}
+
+\subsubsection*{Intermediate language} \figureref{interlang} shows the
+intermediate language of our compiler. The syntax follows that of the source
+language, except that the types and operators are \emph{labeled}. Below
+we mainly focus on the bits that are different from the source
+language.
+
+Labels $\ell$ consist of secret labels $\mathcal{A}$ and
+$\mathcal{B}$, denoting the arithmetic and boolean shared secrets
+resp., and the public label $\mathcal{P}$. Types $\tau$ are then
+labeled base types $\sigma^{\ell}$ and arrays of labeled base types
+$\sigma^{\ell}[n]$.
+
+Most of the expression forms $\widetilde{e}$ are same as $e$, except
+that the binary operators, and the conditional forms
+are annotated with label $\ell$, denoting how the operators should be
+evaluated ($\mathcal{P}$ for in-clear, and $\mathcal{A}$ and
+$\mathcal{B}$ for using arithmetic or boolean circuits resp.). The
+form $\widetilde{e} \rhd m$ denotes coercing the value of $\widetilde{e}$ to be
+$m$-secret shared. The statements $\widetilde{s}$ are analogous to
+$s$.
+
+\begin{figure}[t]
+  \small
+  \fbox{$\Gamma \vdash e : \tau \leadsto \widetilde{e}$}
+  \[
+  \begin{array}{c}
+     \inferrule*[lab={\footnotesize{T-Cons}}]
+               {
+                 \tau = \mathsf{typeof}(c)^{\mathcal{P}}
+               }
+               {
+                 \Gamma \vdash c : \tau \leadsto c 
+               }
+
+                \inferrule*[lab={\footnotesize{T-Inp}}]
+               {
+               }
+               {
+                 \Gamma \vdash \kw{in}_{j} : \sigma^{m} \leadsto \kw{in}^{m}_{j}
+               }
+
+    %% \inferrule*[lab={\footnotesize{T-Cons}}]
+    %%            {
+    %%              \tau = \mathsf{typeof}(c)^{\mathcal{P}}
+    %%            }
+    %%            {
+    %%              \Gamma \vdash c : \tau \leadsto c
+    %%            }
+
+    %%  \inferrule*[lab={\footnotesize{T-Var}}]
+    %%            {
+    %%            }
+    %%            {
+    %%              \Gamma \vdash x : \Gamma(x) \leadsto x
+    %%            }
+\\\\
+
+	  \inferrule*[lab={\footnotesize{T-Mult}}]
+               {
+                 \forall i \in \{1,2\}.\:\Gamma \vdash e_{i} : \kw{uint}^{\ell} \leadsto \widetilde{e_{i}}\\\\
+                 \left(\ell = \mathcal{P}\right) \vee \left(\ell = \mathcal{A}\right)
+               }
+               {
+                 \Gamma \vdash e_{1} \mulop e_{2} : \kw{uint}^{\ell} \leadsto \widetilde{e_{1}} \mulop_{\ell} \widetilde{e_{2}}
+               }
+
+     \inferrule*[lab={\footnotesize{T-Gt}}]
+               {
+                 \forall i \in \{1,2\}.\:\Gamma \vdash e_{i} : \kw{uint}^{\ell} \leadsto \widetilde{e_{i}}\\\\
+                 \left(\ell = \mathcal{P}\right) \vee \left(\ell = \mathcal{B}\right)
+               }
+               {
+                 \Gamma \vdash e_{1} > e_{2} : \kw{bool}^{\ell} \leadsto \widetilde{e_{1}} >_{\ell} \widetilde{e_{2}}
+               }
+
+\\\\
+
+	\inferrule*[lab={\footnotesize{T-Read}}]
+               {
+                 \Gamma \vdash x : \sigma^{\ell}[n] \leadsto x\\\\
+                 \Gamma \vdash e : \kw{uint}^{\mathcal{P}} \leadsto \widetilde{e}\\\\
+                 \Gamma \models e < n
+               }
+               {
+                 \Gamma \vdash x[e] : \sigma^{\ell} \leadsto x[\widetilde{e}]
+               }
+
+     \inferrule*[lab={\footnotesize{T-Cond}}]
+               {
+                 \Gamma \vdash e : \kw{bool}^{\ell} \leadsto \widetilde{e}\\\\
+                 \forall i \in \{1,2\}.\:\Gamma \vdash e_{i} : \sigma^{\ell'} \leadsto \widetilde{e_{i}}\\\\
+                 \ell = \mathcal{P} \vee (\ell = \mathcal{B} \wedge \ell' =\mathcal{B})
+               }
+               {
+                 \Gamma \vdash \cond{e}{e_{1}}{e_{2}} : \sigma^{\ell'} \leadsto \lcond{\ell}{\widetilde{e}}{\widetilde{e_{1}}}{\widetilde{e_{2}}}
+               }
+
+\\\\
+
+     \inferrule*[lab={\footnotesize{T-Arr}}]
+               {
+                 \forall i \in [n].\:\Gamma \vdash e_{i} : \sigma^{\ell} \leadsto \widetilde{e_{i}}
+               }
+               {
+                 \Gamma \vdash [\overline{e_{i}}]_{n} : \sigma^{\ell}[n] \leadsto [\overline{\widetilde{e_{i}}}]_{n}
+               }
+
+     \inferrule*[lab={\footnotesize{T-Sub}}]
+               {
+                 \Gamma \vdash e : \sigma^{\ell} \leadsto \widetilde{e}
+               }
+               {
+                 \Gamma \vdash e : \sigma^{m} \leadsto \widetilde{e} \rhd m
+               }
+
+  \end{array}
+  \]
+  \\\\
+    \fbox{$\Gamma \vdash s \leadsto \widetilde{s} \mid \Gamma_{1}$}
+  \[
+  \\
+  \begin{array}{c}
+     \inferrule*[lab={\footnotesize{T-Decl}}]
+               {
+                 \tau = \psi^\ell \\\\
+                 \Gamma \vdash e : \tau \leadsto \widetilde{e}
+               }
+               {
+                 \Gamma \vdash \psi\:x = e \leadsto \tau\:x = \widetilde{e} \mid \Gamma, x:\tau
+               }
+
+     \inferrule*[lab={\footnotesize{T-Assgn}}]
+               {
+                 \Gamma(x) = \sigma^{\ell}\\\\
+                 \Gamma \vdash e : \sigma^{\ell} \leadsto \widetilde{e}
+               }
+               {
+                 \Gamma \vdash x := e \leadsto x = \widetilde{e} \mid \Gamma
+               }
+
+\\\\
+
+     \inferrule*[lab={\footnotesize{T-For}}]
+               {
+                 \Gamma, x:\kw{uint}^{\mathcal{P}} \vdash \loops{x}{n_{2}}{s} \leadsto \loops{x}{n_{2}}{\widetilde{s}} \mid \_
+               }
+               {
+                 \Gamma \vdash \for{x}{n_{1}}{n_{2}}{s} \leadsto \for{x}{n_{1}}{n_{2}}{\widetilde{s}} \mid \Gamma
+               }
+
+               \\\\
+
+     \inferrule*[lab={\footnotesize{T-Write}}]
+               {
+                 \Gamma \vdash x : \sigma^{\ell}[n] \leadsto x\\\\
+                 \Gamma \vdash e_{1} : \kw{uint}^{\mathcal{P}} \leadsto \widetilde{e_{1}}\\\\
+                 \Gamma \vdash e_{2} : \sigma^{\ell} \leadsto \widetilde{e_{2}}\\\\
+                 \Gamma \models e_{1} < n
+               }
+               {
+                 \Gamma \vdash x[e_{1}] := e_{2} \leadsto x[\widetilde{e_{1}}] := \widetilde{e_{2}} \mid \Gamma
+               }
+
+     \inferrule*[lab={\footnotesize{T-Out}}]
+               {
+                 \Gamma \vdash e : \sigma^{m} \leadsto \widetilde{e}
+               }
+               {
+                 \Gamma \vdash \kw{out}\:e \leadsto \kw{out}\:\widetilde{e} \mid \Gamma
+               }
+
+\\\\
+
+     \inferrule*[lab={\footnotesize{T-If}}]
+               {
+                 \Gamma \vdash e : \kw{bool}^{\mathcal{P}} \leadsto \widetilde{e}\\\\
+                 \forall i \in \{1, 2\}.\:\Gamma \vdash s_{i} \leadsto \widetilde{s_{i}} \mid \_
+               }
+               {
+                 \Gamma \vdash \ite{e}{s_{1}}{s_{2}} \leadsto \ite{\widetilde{e}}{\widetilde{s_{1}}}{\widetilde{s_{2}}}  \mid \Gamma
+               }
+
+     \inferrule*[lab={\footnotesize{T-Seq}}]
+               {
+                 \Gamma \vdash s_{1} \leadsto \widetilde{s_{1}} \mid \Gamma_{1}\\\\
+                 \Gamma_{1} \vdash s_{2} \leadsto \widetilde{s_{2}} \mid \Gamma_{2}
+               }
+               {
+                 \Gamma \vdash s_{1}; s_{2} \leadsto \widetilde{s_{1}}; \widetilde{s_{2}} \mid \Gamma_{2}
+               }
+
+\\\\
+
+     \inferrule*[lab={\footnotesize{T-While}}]
+               {                 
+                 \Gamma(x) = \kw{uint}^{\mathcal{P}}\quad
+                 \Gamma \vdash s \leadsto \widetilde{s} \mid \_\quad
+                 x \notin \mathsf{modifies}(s)
+               }
+               {
+                 \Gamma \vdash \loops{x}{n_{2}}{s} \leadsto \loops{x}{n_{2}}{\widetilde{s}} \mid \Gamma
+               }
+
+  \end{array}
+  \]
+\caption{Compilation judgments (selected rules)}
+\label{fig:compile}
+\end{figure}
+
+\subsubsection*{Source to Intermediate Compilation judgments} We provide the compilation
+rules in \figureref{compile}. \divya{PL jargon?} We present the rules in a
+declarative style, where the rules are non-syntax
+directed, and the labels $\ell$ are chosen
+non-deterministically. \sectionref{impl} describes the label  inference
+scheme in our implementation.
+
+Judgment $\Gamma \vdash e : \tau \leadsto \widetilde{e}$, where
+$\Gamma$ maps variables $x$ to types $\tau$, says that
+under $\Gamma$, $e$ (in source language) compiles to $\widetilde{e}$ (in intermediate language) with type $\tau$.
+
+The rule ({\sc{T-Cons}}) assigns the label $\mathcal{P}$ to the
+constants, as the constants are always public. Rule ({\sc{T-Mult}})
+compiles a multiplication to either a public multiplication  ($\mulop_{\mathcal{P}}$), or a
+secret arithmetic multiplication ($\mulop_{\mathcal{A}}$). As
+our compiler is cryptographic cost aware, it never compiles the
+multiplication to boolean multiplication $\mulop_{\mathcal{B}}$. In a similar manner,
+rule ({\sc{T-Gt}}) compiles $e_{1} > e_{2}$ to either public comparison,
+or secret boolean comparison. The rule for conditional
+({\sc{T-Cond}}) has two cases: when the conditional expression $e$ is of
+type $\kw{bool}^{\mathcal{P}}$, both the branches have a base type
+$\sigma^{\ell'}$, for an arbitrary $\ell'$, and the conditional is
+compiled to a public conditional, whereas when the conditional
+expression has type $\kw{bool}^{\mathcal{B}}$, $\ell'$ is also
+$\mathcal{B}$, and the conditional is compiled to a secret conditional
+using a boolean circuit. Note that we restrict the type of the branches
+to be of base type. 
+Rule ({\sc{T-Read}}) type checks an array read. It checks
+that the array index $e$ is public, and uses a static bounds checking
+judgment $\Gamma \models e < n$ to prove that the array index is in
+bounds. Section~\ref{sec:impl} discusses our implementation of bounds
+checking. Rule ({\sc{T-Inp}}) picks a label $m$ for the input. Finally,
+rule ({\sc{T-Sub}}) is the subsumption rule that coerces an expression of
+type $\sigma^{\ell}$ to an expression of type $\sigma^{m}$ using the
+coerce expression. It is important for security that the secrets cannot be coerced to public
+values.
+
+Judgment $\Gamma \vdash s : \tau \leadsto \widetilde{s} \mid \Gamma_{1}$
+compiles a statement $s$ resulting in the statement $\widetilde{s}$
+and type environment $\Gamma_{1}$. Rule ({\sc{T-Decl}}) picks a label
+$\ell$, and adds the binding for $x$ to the environment. If $\psi = \sigma$, $\psi^\ell = \sigma^\ell$, else if $\psi = \sigma[n]$, $\psi^\ell = \sigma^\ell[n]$. 
+Rule ({\sc{T-Assign}}) looks up the type of $x$ in $\Gamma$ and compiles $e$ to $\tilde{e}$ of same type.
+Note that in this rule we restrict the type of variable $x$ to be of base type.
+Rule ({\sc{T-For}}) adds the loop counter $x$ to $\Gamma$ at type
+$\kw{uint}^{\mathcal{P}}$, and delegates type checking to the
+\kw{while} form. The rule ({\sc{T-Write}}), similar to ({\sc{T-Read}}),
+checks that the index has type $\kw{uint}^{\mathcal{P}}$, and is in
+bounds. Rule ({\sc{T-Out}})
+types the expression $e$ at some secret label $m$. Rule ({\sc{T-If}})
+checks that the conditional expression is public, and rule
+({\sc{T-Seq}}) sequences the type environments. Finally, the typing
+rule for the (internal) \kw{while} form ensures that $x$ is mapped in
+$\Gamma$ at type $\kw{uint}^{\mathcal{P}}$, and that $x \notin
+\mathsf{modifies}(s)$--this is necessary for ensuring the termination
+of our semantics.
+
+\aseem{The target semantics models the semantics of the code output by
+our compiler (C++ code in our implementation).}
+A program in the intermediate language is evaluated to a circuit that can be executed in the distributed setting later.
+The evaluation to a circuit computes over the public parts of the program and also \emph{flattens} the arrays so that the circuits are unaware of the array structure.
+Crucially, this phase of the semantics does not have access to the secrets.
+Below, we first provide the language for the circuits used followed by our evaluation rules.
 
@@
   \]
-\caption{Circuits   \divya{Is environ misplaced?}}
+\caption{Circuits   \divya{Is environ misplaced? Can we merge base value and circuit gate?}}
 \label{fig:circuits}
 \end{figure}
-
-\subsubsection*{Intermediate language} \figureref{interlang} shows the
-intermediate language of our compiler. The syntax follows that of the source
-language, except that the types and operators are \emph{labeled}. Below
-we mainly focus on the bits that are different from the source
-language.
-
-Labels $\ell$ consist of secret labels $\mathcal{A}$ and
-$\mathcal{B}$, denoting the arithmetic and boolean shared secrets
-resp., and the public label $\mathcal{P}$. Types $\tau$ are then
-labeled base types $\sigma^{\ell}$ and arrays of labeled base types
-$\sigma^{\ell}[n]$.
-
-Most of the expression forms $\widetilde{e}$ are same as $e$, except
-that the binary operators, and the conditional forms
-are annotated with label $\ell$, denoting how the operators should be
-evaluated ($\mathcal{P}$ for in-clear, and $\mathcal{A}$ and
-$\mathcal{B}$ for using arithmetic or boolean circuits resp.). The
-form $\widetilde{e} \rhd m$ denotes coercing \divya{the value of} $\widetilde{e}$ to be
-$m$-secret shared. The statements $\widetilde{s}$ are analogous to
-$s$.
-
-\aseem{The target semantics models the semantics of the code output by
-our compiler (C++ code in our implementation).}
-A program in the intermediate language is evaluated to a circuit that can be executed in the distributed setting later.
-The evaluation to a circuit computes over the public parts of the program and also \emph{flattens} the arrays so that the circuits are unaware of the array structure.
-Crucially, this phase of the semantics does not have access to the secrets.
-Below, we first provide the language for the circuits used followed by our evaluation rules.
 
@@
 
-\begin{figure}
-  \small
-  \fbox{$\Gamma \vdash e : \tau \leadsto \widetilde{e}$}
-  \[
-  \begin{array}{c}
-     \inferrule*[lab={\footnotesize{T-Cons}}]
-               {
-                 \tau = \mathsf{typeof}(c)^{\mathcal{P}}
-               }
-               {
-                 \Gamma \vdash c : \tau \leadsto c 
-               }
-
-     \inferrule*[lab={\footnotesize{T-Add}}]
-               {
-                 \forall i \in \{1,2\}.\:\Gamma \vdash e_{i} : \kw{uint}^{\ell} \leadsto \widetilde{e_{i}}\\\\
-                 \ell = \mathcal{P} \vee \ell = \mathcal{A}
-               }
-               {
-                 \Gamma \vdash e_{1} + e_{2} : \kw{uint}^{\ell} \leadsto \widetilde{e_{1}} +_{\ell} \widetilde{e_{2}}
-               }
-
-    %% \inferrule*[lab={\footnotesize{T-Cons}}]
-    %%            {
-    %%              \tau = \mathsf{typeof}(c)^{\mathcal{P}}
-    %%            }
-    %%            {
-    %%              \Gamma \vdash c : \tau \leadsto c
-    %%            }
-
-    %%  \inferrule*[lab={\footnotesize{T-Var}}]
-    %%            {
-    %%            }
-    %%            {
-    %%              \Gamma \vdash x : \Gamma(x) \leadsto x
-    %%            }
-\\\\
-     \inferrule*[lab={\footnotesize{T-Gt}}]
-               {
-                 \forall i \in \{1,2\}.\:\Gamma \vdash e_{i} : \kw{uint}^{\ell} \leadsto \widetilde{e_{i}}\\\\
-                 \ell = \mathcal{P} \vee \ell = \mathcal{B}
-               }
-               {
-                 \Gamma \vdash e_{1} > e_{2} : \kw{bool}^{\ell} \leadsto \widetilde{e_{1}} >_{\ell} \widetilde{e_{2}}
-               }
-
-     \inferrule*[lab={\footnotesize{T-Read}}]
-               {
-                 \Gamma \vdash x : \sigma^{\ell}[n] \leadsto x\\\\
-                 \Gamma \vdash e : \kw{uint}^{\mathcal{P}} \leadsto \widetilde{e}\\\\
-                 \Gamma \models e < n
-               }
-               {
-                 \Gamma \vdash x[e] : \sigma^{\ell} \leadsto x[\widetilde{e}]
-               }
-
-\\\\
-
-     \inferrule*[lab={\footnotesize{T-Cond}}]
-               {
-                 \Gamma \vdash e : \kw{bool}^{\ell} \leadsto \widetilde{e}\\\\
-                 \forall i \in \{1,2\}.\:\Gamma \vdash e_{i} : \sigma^{\ell_{1}} \leadsto \widetilde{e_{i}}\\\\
-                 \ell = \mathcal{P} \vee (\ell = \mathcal{B} \wedge \ell' =\mathcal{B})
-               }
-               {
-                 \Gamma \vdash \cond{e}{e_{1}}{e_{2}} : \sigma^{\ell_{1}} \leadsto \lcond{\ell}{\widetilde{e}}{\widetilde{e_{1}}}{\widetilde{e_{2}}}
-               }
-
-     \inferrule*[lab={\footnotesize{T-Inp}}]
-               {
-               }
-               {
-                 \Gamma \vdash \kw{in}_{j} : \sigma^{m} \leadsto \kw{in}^{m}_{j}
-               }
-
-\\\\
-
-     \inferrule*[lab={\footnotesize{T-Arr}}]
-               {
-                 \forall i \in \{0 \dots n - 1\}.\:\Gamma \vdash e_{i} : \sigma^{\ell} \leadsto \widetilde{e_{i}}
-               }
-               {
-                 \Gamma \vdash [\overline{e_{i}}]_{n} : \sigma^{\ell}[n] \leadsto [\overline{\widetilde{e_{i}}}]_{n}
-               }
-
-     \inferrule*[lab={\footnotesize{T-Sub}}]
-               {
-                 \Gamma \vdash e : \sigma^{\ell} \leadsto \widetilde{e}
-               }
-               {
-                 \Gamma \vdash e : \sigma^{m} \leadsto \widetilde{e} \rhd m
-               }
-
-  \end{array}
-  \]
-  \\\\
-    \fbox{$\Gamma \vdash s \leadsto \widetilde{s} \mid \Gamma_{1}$}
-  \[
-  \\
-  \begin{array}{c}
-     \inferrule*[lab={\footnotesize{T-Decl}}]
-               {
-                 \psi = \sigma \Rightarrow \tau = \sigma^{\ell}\\\\
-                 \psi = \sigma[n] \Rightarrow \tau = \sigma^{\ell}[n]\\\\
-                 \Gamma \vdash e : \tau \leadsto \widetilde{e}
-               }
-               {
-                 \Gamma \vdash \psi\:x = e \leadsto \tau\:x = \widetilde{e} \mid \Gamma, x:\tau
-               }
-
-     \inferrule*[lab={\footnotesize{T-Assgn}}]
-               {
-                 \Gamma(x) = \sigma^{\ell}\\\\
-                 \Gamma \vdash e : \sigma^{\ell} \leadsto \widetilde{e}
-               }
-               {
-                 \Gamma \vdash x := e \leadsto x = \widetilde{e} \mid \Gamma
-               }
-
-\\\\
-
-     \inferrule*[lab={\footnotesize{T-For}}]
-               {
-                 \Gamma, x:\kw{uint}^{\mathcal{P}} \vdash \loops{x}{n_{2}}{s} \leadsto \loops{x}{n_{2}}{\widetilde{s}} \mid \_
-               }
-               {
-                 \Gamma \vdash \for{x}{n_{1}}{n_{2}}{s} \leadsto \for{x}{n_{1}}{n_{2}}{\widetilde{s}} \mid \Gamma
-               }
-
-               \\\\
-
-     \inferrule*[lab={\footnotesize{T-Write}}]
-               {
-                 \Gamma \vdash x : \sigma^{\ell}[n] \leadsto x\\\\
-                 \Gamma \vdash e_{1} : \kw{uint}^{\mathcal{P}} \leadsto \widetilde{e_{1}}\\\\
-                 \Gamma \vdash e_{2} : \sigma^{\ell} \leadsto \widetilde{e_{2}}\\\\
-                 \Gamma \models e_{1} < n
-               }
-               {
-                 \Gamma \vdash x[e_{1}] := e_{2} \leadsto x[\widetilde{e_{1}}] := \widetilde{e_{2}} \mid \Gamma
-               }
-
-     \inferrule*[lab={\footnotesize{T-Out}}]
-               {
-                 \Gamma \vdash e : \sigma^{m} \leadsto \widetilde{e}
-               }
-               {
-                 \Gamma \vdash \kw{out}\:e \leadsto \kw{out}\:\widetilde{e} \mid \Gamma
-               }
-
-\\\\
-
-     \inferrule*[lab={\footnotesize{T-If}}]
-               {
-                 \Gamma \vdash e : \kw{bool}^{\mathcal{P}} \leadsto \widetilde{e}\\\\
-                 \forall i \in \{1, 2\}.\:\Gamma \vdash s_{i} \leadsto \widetilde{s_{i}} \mid \_
-               }
-               {
-                 \Gamma \vdash \ite{e}{s_{1}}{s_{2}} \leadsto \ite{\widetilde{e}}{\widetilde{s_{1}}}{\widetilde{s_{2}}}  \mid \Gamma
-               }
-
-     \inferrule*[lab={\footnotesize{T-Seq}}]
-               {
-                 \Gamma \vdash s_{1} \leadsto \widetilde{s_{1}} \mid \Gamma_{1}\\\\
-                 \Gamma_{1} \vdash s_{2} \leadsto \widetilde{s_{2}} \mid \Gamma_{2}
-               }
-               {
-                 \Gamma \vdash s_{1}; s_{2} \leadsto \widetilde{s_{1}}; \widetilde{s_{2}} \mid \Gamma_{2}
-               }
-
-\\\\
-
-     \inferrule*[lab={\footnotesize{T-While}}]
-               {                 
-                 \Gamma(x) = \kw{uint}^{\mathcal{P}}\quad
-                 \Gamma \vdash s \leadsto \widetilde{s} \mid \_\quad
-                 x \notin \mathsf{modifies}(s)
-               }
-               {
-                 \Gamma \vdash \loops{x}{n_{2}}{s} \leadsto \loops{x}{n_{2}}{\widetilde{s}} \mid \Gamma
-               }
-
-  \end{array}
-  \]
-\caption{Compilation judgments (selected rules)}
-\label{fig:compile}
-\end{figure}
-
-\subsubsection*{Compilation judgments} We now come to the compilation
-rules shown in Figure~\ref{fig:compile}. We present the rules in a
-declarative style, where the rules are non-syntax
-directed, and the labels $\ell$ are chosen
-non-deterministically. Section~\ref{sec:impl} describes the inference
-scheme in our implementation.
-
-Judgment $\Gamma \vdash e : \tau \leadsto \widetilde{e}$, where
-$\Gamma$ maps variables $x$ to types $\tau$, says that
-under $\Gamma$, $e$ compiles to $\widetilde{e}$ with type $\tau$.
-
-The rule ({\sc{T-Cons}}) assigns the label $\mathcal{P}$ to the
-constants, as the constants are always public. Rule ({\sc{T-Add}})
-compiles an addition to either public addition ($+_{\mathcal{P}}$), or
-secret addition using an arithmetic circuit ($+_{\mathcal{A}}$). As
-our compiler is cryptographic cost aware, it never compiles the
-addition to boolean addition $+_{\mathcal{B}}$. In a similar manner,
-rule ({\sc{T-Gt}}) compiles $e_{1} > e_{2}$ to either public comparison,
-or secret comparison using boolean circuits. The rule for conditional
-({\sc{T-Cond}}) has two cases: when the conditional expression $e$ is of
-type $\kw{bool}^{\mathcal{P}}$, both the branches have a base type
-$\sigma^{\ell_{1}}$, for an arbitrary $\ell_{1}$, and the conditional is
-compiled to public conditional, whereas when the conditional
-expression has type $\kw{bool}^{\mathcal{B}}$, $\ell_{1}$ is also
-$\mathcal{B}$, and the conditional is compiled to a secret conditional
-using a boolean circuit. Note that we restrict the type of the branches
-to be of base type. Rule ({\sc{T-Read}}) types an array read. It checks
-that the array index $e$ is public, and uses a static bounds checking
-judgment $\Gamma \models e < n$ to prove that the array index is in
-bounds. Section~\ref{sec:impl} discusses our implementation of bounds
-checking. Rule ({\sc{T-Inp}}) picks a label $m$ for the input. Finally,
-rule ({\sc{T-Sub}}) is the subsumption rule that coerces an expression of
-type $\sigma^{\ell}$ to an expression of type $\sigma^{m}$ using the
-coerce expression. Importantly, secrets cannot be coerced to public
-values.
-
-Judgment $\Gamma \vdash s : \tau \leadsto \widetilde{s} \mid \Gamma_{1}$
-compiles a statement $s$ resulting in the statement $\widetilde{s}$
-and type environment $\Gamma_{1}$. Rule ({\sc{T-Decl}}) picks a label
-$\ell$, and adds the binding for $x$ to the environment. Rule
-({\sc{T-For}}) adds the loop counter $x$ to $\Gamma$ at type
-$\kw{uint}^{\mathcal{P}}$, and delegates type checking to the
-\kw{while} form. The rule ({\sc{T-Write}}), similar to ({\sc{T-Read}}),
-checks that the index has type $\kw{uint}^{\mathcal{P}}$, and is in
-bounds. Rule ({\sc{T-Out}})
-types the expression $e$ at some secret label $m$. Rule ({\sc{T-If}})
-checks that the conditional expression is public, and rule
-({\sc{T-Seq}}) sequences the type environments. Finally, the typing
-rule for the (internal) \kw{while} form ensures that $x$ is mapped in
-$\Gamma$ at type $\kw{uint}^{\mathcal{P}}$, and that $x \notin
-\mathsf{modifies}(s)$--this is necessary for ensuring the termination
-of our semantics.
-
 \subsubsection*{Correctness theorem} We first define a judgment for

### 2017-10-26 OaklandSubmission/impl.tex

--- 2017-10-25 OaklandSubmission/impl.tex
@@
 compiles each of our benchmarks in under a second to C++ code that makes calls to the ABY
-library~\cite{aby}. We use the default options of ABY except 
-for the maximum number of gates. Our benchmarks require us to set this value to a
-hundred million.
+library~\cite{aby}.
+We set the security paramenter to 128 and use OT extension-based arithmetic multiplication triples generation.
 ABY provides facilities to use multiple threads\footnote{The protocol has an offline phase and an online phase.
@@
 We use at most four threads in our evaluation.
+
+The current language implementation supports the following operators:
+addition, subtraction, multiplication, division by powers of two, left shift,
+logical and arithmetic right shifts, bitwise-(and, or, xor), unary negation,
+bitwise negation. Because of its high cost, general division is not supported natively by \tool.
+However, it can be implemented in about 30 lines.
 
@@
 \tool enforces the array indices to be public. The developer expresses this functionality in \tool
-as $x > 0 ? A[1] : A[0]$. In general, for an array of size $2^n$, we need to write $2^{n-1}$ multiplexers
-in \tool. This design choice of making secret indexing inconvinient to implement was deliberate.
-Accessing arrays at secret locations is a powerful but expensive primitive
-and we want the developers to perform this operation sparingly.
-
-The secure code partitioning is not fully automatic.
-in particular, the developer needs to decompose a program $P$ into $P_1||\ldots||P_k$ manually (Section~\ref{sec:pipe}). Then \tool generates $Q_1||\ldots||Q_k$ automatically.
+as $x > 0 ? A[1] : A[0]$. In general, a secret access to an array of size $2^n$ requires $2^{n}-1$ multiplexers in \tool. Secure code partitioning currently requires a manual step.
+In particular, the developer needs to decompose a program $P$ into $P_1||\ldots||P_k$ manually (Section~\ref{sec:pipe}). Then \tool generates $Q_1||\ldots||Q_k$ automatically.
 Automating the decomposition step requires an analysis that can 
@@
 bool and vice versa.
-The implemntation of \tool performs several optimizations to reduce these.
-For example, on each coercion, \tool memoizes the pair of arithmetic share and boolean share involved in the coercion.
+The \tool compiler performs several optimizations to reduce these.
+For example, on each coercion, \tool memorizes the pair of arithmetic share and boolean share involved in the coercion.
 \tool invalidates such pairs when the variables corresponding to the shares are overwritten by assignments.
 In subsequent coercions, \tool can reuse valid pairs (if available) instead of recomputing them afresh.
+
+\divya{
+\begin{enumerate}
+\item add all the expressions that our language supports such as mult, bit-wise and and anything else.
+\item Say that we compile to source to intermediate and use inter to circuit of ABY.. We can decide if we want to say this.
+\end{enumerate}
+}

### 2017-10-26 OaklandSubmission/ld.tex

@@
 
-\aseem{The target semantics models the semantics of the code output by
-our compiler (C++ code in our implementation).}
-A program in the intermediate language is evaluated to a circuit that can be executed in the distributed setting later.
-The evaluation to a circuit computes over the public parts of the program and also \emph{flattens} the arrays so that the circuits are unaware of the array structure.
+As mentioned earlier, the intermediate language models the code such as in \figureref{ex-aby} output by our compiler. 
+Next, a program in the intermediate language is evaluated to a circuit that can be executed in the distributed setting later.
+The evaluation to a circuit computes away the public parts of the program and also \emph{flattens} the arrays so that the circuits are unaware of the array structure.
 Crucially, this phase of the semantics does not have access to the secrets.
@@
     \ftext{Wire id} & \wire &&\\
-    \ftext{Circuit gate} & \gate & ::= & \wire \mid \kw{in}^{m}_{j} \mid \kw{add}\:\gate_1\:\gate_{2} \mid \kw{gt}\:\gate_{1}\:\gate_{2}\\
+    \ftext{Circuit gate} & \gate & ::= & \wire \mid \kw{in}^{m}_{j} \mid \kw{mult}\:\gate_1\:\gate_{2} \mid \kw{gt}\:\gate_{1}\:\gate_{2}\\
     & & \mid & \kw{mux}\:\gate\:\gate_{1}\:\gate_{2} \mid \bval \rhd m\\
@@
 values later as part of the circuit semantics). Circuit gates $\gate$
-are $\wire$, input gates, \kw{add}, \kw{gt}, and \kw{mux} gates, and
+are $\wire$, input gates, \kw{mult}, \kw{gt}, and \kw{mux} gates, and
 coerce gates $\bval \rhd m$. \divya{Gates are not just gates but a circuit?} Target values $\val$

### 2017-10-26 OaklandSubmission/train.tex

\subsection{Matrix factorization}
\tool, as a language, is not tied to secure prediction and can express more general computations.
To demonstrate the expressiveness of \tool, we use it to implement secure matrix factorization~\cite{valeriaMatrix}. Abstractly, given a sparse matrix $\mathcal{M}$ of dimensions
$n\times m$ and $M$ non-zero entries, the goal is to generate a matrix $U$ of dimension $n\times d$ and a matrix
$V$ of dimension $d\times m$ such that $\mathcal{M}\approx UV$. This operator is useful in recommender systems.
In particular, Nikolaenko et al.~\cite{valeriaMatrix} shows how to implement a movie recommender system which does not require users to reveal their data in the clear, i.e., the ratings the users have assigned to movies. The implementation is a two party computation of an iterative algorithm for matrix factorization (Algorithm 1 in~\cite{valeriaMatrix}).
This algorithm is based on gradient descent and iteratively converges to a local minima.
We implement this algorithm in \tool.

To ensure that the algorithm converges to the right local minima, Nikolaenko et al. require
36 bits of precision. Since ABY supports either 32-bit or 64-bit integers, our \tool implementation
manipulates 64-bit variables. For $\mathcal{M}$ Nikolaenko et al. consider $n=940$ users, $m=40$ most popular movies, and $M=14683$ ratings from the MovieLens dataset. The time reported in~\cite{valeriaMatrix}
for one iteration is 2.9 hours\footnote{~\cite{valeriaMatrix} does not report the network round-trip time.}. This compuation is large enough that we partition each iteration
into three stages. The first stage involves a Batcher~\cite{Batcher} sorting network followed by a linear pass.
The second stage involves sorting and gradient computations and is the heaviest stage.
The third  stage is similar to the first stage. The results are presented in Table~\ref{tab:factor}.

\begin{table}
\begin{tabular}{c|c|c|c |c|c| c}
  Stage         &  LAN (s) & WAN (s) & Comm. (Mb)  & depth & \#Gates & LOC\\
\hline
1    &  175       & 662        & 29816       & 16370    & 33m    & 500  \\
\hline
2    &  193        & 1095        & 31945        & 30916    & 37m & 516 \\
\hline
3    &  178        & 627        & 29810        & 16369    & 32m  & 478  \\
\hline
Total    &  546      & 2384        & 91571        & --    & 102m & 1494 \\
\end{tabular}

 \caption{Partitioning results for matrix factorization. The time reported by~\cite{valeriaMatrix} for this computation is about 10440 seconds.}
 \label{tab:factor} 
\end{table}

We observe that in the LAN setting, we are about 19 times faster than~\cite{valeriaMatrix} and in the WAN
setting we are about 4 times faster. The main source of these significant speedups is that, unlike~\cite{valeriaMatrix}, \tool does not need to convert the functionality into boolean circuits. 
However, this benchmark requires more lines of code than the previous benchmarks.
It is inconvinient to write Batcher's sort in an iterative language (about 450 lines of \tool).
The size of a recursive implementation would be much slower.
In the future, we would like to add support for functions and recursion.
However, the current programmer effort seems miniscule compared to the mammoth implementation effort
put in by Nikolaenko et al. (Section 5 of ~\cite{valeriaMatrix}) to scale a boolean circuit based
crypto-backend to this benchmark.

### 2017-10-26 OaklandSubmission/impl.tex
impl

@@
 ABY provides facilities to use multiple threads\footnote{The protocol has an offline phase and an online phase.
-Although multiple threads can be used in the offline phase, the online phase is single threaded.}.
-We use at most four threads in our evaluation.
+Although multiple threads can be used in the offline phase, the online phase is single threaded.} and we use at most four threads in our evaluation.
 
@@
 logical and arithmetic right shifts, bitwise-(and, or, xor), unary negation,
-bitwise negation. Because of its high cost, general division is not supported natively by \tool.
-However, it can be implemented in about 30 lines.
+bitwise-negation, logical-(not, and, or, xor), and comparisons (less than, greater than, equality).
+Because of their high cost, integral division and floating-point operators are not supported natively by \tool.
+However, we have implemented integral division in 30 lines of \tool and floating-point support in
+ABY is under active development~\cite{ddkssz15}.
+
+Some of our benchamrks require accessing arrays at secret location.
+For example, consider the expression $A[x]$ where $A$ is an array of size 2 and $x$ is shared.
+\tool enforces the array indices to be public. The developer expresses this functionality in \tool
+as $x > 0 ? A[1] : A[0]$. In general, a secret access to an array of size $2^n$ requires $2^{n}-1$ multiplexers in \tool.
+
+ Secure code partitioning currently requires a manual step.
+In particular, the developer needs to decompose a program $P$ into $P_1||\ldots||P_k$ manually (Section~\ref{sec:pipe}). Then \tool generates $Q_1||\ldots||Q_k$ automatically.
+Automating the decomposition step requires an analysis that can 
+statically estimate the resource usage of a \tool program and we leave the construction of such
+an analysis as future work.
 
@@
 in if-conditions and for-loops are assigned public labels.
-All other variables are assigned arithmetic labels.
+All other variables are assigned arithmetic labels (and can later be coerced to boolean).
 In the future, we would like to explore more intelligent strategies for label assignment.
 
-Some of our benchamrks require accessing arrays at secret location.
-For example, consider the expression $A[x]$ where $A$ is an array of size 2 and $x$ is shared.
-\tool enforces the array indices to be public. The developer expresses this functionality in \tool
-as $x > 0 ? A[1] : A[0]$. In general, a secret access to an array of size $2^n$ requires $2^{n}-1$ multiplexers in \tool. Secure code partitioning currently requires a manual step.
-In particular, the developer needs to decompose a program $P$ into $P_1||\ldots||P_k$ manually (Section~\ref{sec:pipe}). Then \tool generates $Q_1||\ldots||Q_k$ automatically.
-Automating the decomposition step requires an analysis that can 
-statically estimate the resource usage of a \tool program and we leave the construction of such
-an analysis as future work.
-
 The compilation rules of Figure~\ref{fig:compile} can introduce redundant coercions from arithmetic to
-bool and vice versa.
+boolean  and vice versa.
 The \tool compiler performs several optimizations to reduce these.
@@
 \tool invalidates such pairs when the variables corresponding to the shares are overwritten by assignments.
-In subsequent coercions, \tool can reuse valid pairs (if available) instead of recomputing them afresh.
+In subsequent coercions, \tool  reuses valid pairs (if available) instead of recomputing them afresh.
 

### 2017-10-26 OaklandSubmission/intro.tex

--- 2017-10-25 OaklandSubmission/intro.tex
@@
 operator, and $w^Tx$ denotes the inner product of
-$w^{T}$ and $x$.  With the current state-of-the-art, Alice has three
+$w^{T}$ and $x$.  With the current state-of-the-art, Alice has following
 options:
@@
 friendly, domain-specific languages (such
-as~\cite{lambdaps,wysteria,oblivm,smcl,cbmcgc}) that would
+as Fairplay~\cite{fairplay}, Wysteria~\cite{wysteria},
+ObliVM~\cite{oblivm}, CBMC-GC~\cite{cbmcgc}, SMCL~\cite{smcl},
+Sharemind~\cite{sharemind}, \cite{lambdaps} etc.) that would
 automatically compile it to a \mpc protocol. However, all of these
@@
 writing large computations in such a framework can be tedious and
-error-prone (a sentiment also echoed by Demmler et al.~\cite{aby}).
+error-prone (a sentiment echoed by Demmler et al.~\cite{aby} themselves).
 
 %\item
-The third option for Alice is to earn a PhD in cryptography and design
-specialized, efficient protocols (similar to
-~\cite{shafindss,wu,minionn}) for her tasks.
+A third option for Alice is to earn a PhD in cryptography, and design
+and implement specialized, efficient \mpc protocols (similar
+to~\cite{shafindss,wu,minionn}) for her tasks.
 %\end{tiret}
@@
 
-\subsection{\tool: Overview and Contributions} 
+\subsubsection*{\tool}
 \label{sec:contrib}
@@
 
-In this work, we develop and implement our framework \tool\footnote{\tool, read as ``easy peasy'', stands for Easy 2 Party Computation} for secure
-computation. We provide the first compiler where the developer writes a high-level program and the compiler automatically generates an efficient \mpc protocol that uses a
-mix of arithmetic and boolean circuits. 
-It is compatible with any implementation of \mpc that allows a mix of arithmetic and boolean circuits. 
-We focus on semi-honest secure two-party computation and use ABY \cite{aby} as the suitable cryptographic back-end.
-We provide a formal type system and operational semantics for our language and prove correctness and security of our compiler.
-We evaluate our system on wide range of benchmarks for secure prediction \cite{shafindss,wu,barni,secureml,minionn,bonsai} using linear classifier, decision trees and deep neural networks, matrix factorization \cite{valeriaMatrix} and training of deep neural network \cite{secureml}.
-Our evaluations show that protocols generated by our compiler match
-the performance or outperform hand-crafted protocols in most cases (see
-\sectionref{eval}).
-\TODO{Add some text to increase enthusiasm about results.} Moreover, unlike previous systems such as \cite{wysteria,aby,cbmcgc}, our system can scale to arbitrarily large computations using our technique of {\em secure code partitioning} (see below).
-\divya{I think we can skip Yao pipelining here, right?}
-Hence, we show that our framework achieves all four properties of programmability, generality, performance, and scalability (see below for details).
+This paper presents \tool\footnote{Read as ``easy peasy'',
+stands for Easy 2 Party Computation}, a new \mpc framework.
+\tool is the first
+``cryptographic-cost aware'' framework that generates efficient and
+scalable \mpc protocols, comprising combinations of arithmetic and
+boolean circuits, from high-level programs devoid of any
+cryptography. Below we describe the salient features of \tool.
 
-Prior to our work, there have been systems that provide only a subset of the above properties.
-For instance, using the framework of CBMC-GC \cite{cbmcgc} allows  functionalities to be expressed in high-level language C but it gives poor performance due to its usage of boolean circuits as underlying representation.
-Similarly, ABY \cite{aby} framework can potentially give good performance (due to mix of boolean and arithmetic circuits) but is not programmable.
-Yao's garbled circuits \cite{yao} work with boolean circuits and are the most widely used 2PC protocols but they  do not scale to large functions due to large memory requirement.
-Huang et al.
-\cite{yao-pipe} gave the technique of pipelining Yao's garbled circuits that enabled scaling to large functions but still gives poor performance due to its use of boolean circuit representation.
-To summarize, none of the prior frameworks have been used to evaluate the benchmarks such as secure prediction using linear regression, decision trees, deep neural networks, etc.
-In fact, it was widely believed that generic 2PC would not scale to the secure prediction and hence, various protocols were hand-crafted for this task \cite{shafindss,wu,secureml,minionn}.
-Our framework contradicts this belief (see below).
+\subsubsection*{Programmability} \tool source programs are ideal
+functionalities that describe ``what'' computation needs to be done,
+rather than ``how'' to do it. In particular, the programmer writes the
+high-level computation without thinking about the underlying
+cryptographic details. For example, an \tool source
+program for $w^Tx >b$ is as follows:
 
-In this work, we solve the design problem to provide a user friendly language that is expressive and can be compiled to efficient secure 2PC protocols.
-Programmability in \tool and performance of generated protocols has enabled us to write large programs whose circuits  exhaust the memory of the system (~32 GB).
-To combat this issue, we give a technique called secure code partitioning (see below) that allows scaling of \tool to arbitrary size computations.
+\vspace{0.2cm}
+\begin{lstlisting}[language=C]
+uint w[30] = input1(); uint b = input1();
+uint x[30] = input2();
+uint acc = 0;
+for i in [0:30] { acc =  acc + (w[i] * x[i]); }
+output((acc > b) ? 1 : 0)
+\end{lstlisting}
 
-Below, we describe the salient features of \tool in more detail. \\
+\vspace{0.2cm}
+We note that the program is quite similar to what a programmer might
+write in a common programming language. The simplicity of the
+language comes with its usual benefits: it is easily accessible to the
+developers, there are fewer avenues for making mistakes, and it is
+easy to maintain and modify the programs. Needless to say, frameworks
+that expose low-level circuit APIs to the programmer do not enjoy
+these benefits.
 
-\noindent\textbf{Programmability.} \tool is a programmer-friendly framework to
-express the function to be computed and uses a mix of arithmetic and
-boolean \mpc back-ends. We illustrate the above example of $w^Tx
->b$ in our language in \figureref{ex-sml}. 
-As is clear from the code, the programmer can remain completely oblivious of the cryptographic back-end and underlying circuit representation. 
-Also,  the programmer is not required to write keywords such as {\em secret}
-or {\em public} (that are required by languages like Wysteria \cite{wysteria}). 
- In contrast, \figureref{ex-aby} illustrates
-the same example using the framework in ABY \cite{aby}. Here, as
-already mentioned, the programmer needs to create shares and write the low level gates
-like $\mathtt{MUL, ADD, CONS, GT}$, etc \TODO{make these fonts consistent}. The programmer is also required
-to use the appropriate share conversion gates such as $\mathtt{A2Y}$ that
-converts arithmetic shares to Yao-based boolean shares to be used in comparison after the inner-product computation.
-Moreover, it is easy to
-maintain code in \tool and a small change to the functionality requires
-only a small change in the code. In contrast, working with ABY, even
-for a small change a developer might have to re-work the whole circuit
-for efficiency reasons. We elaborate on these examples  in
-\sectionref{ex}. \\
+\subsubsection*{Cryptogaphic-cost awareness} \mpc compiler is aware
+of the cryptographic costs of arithmetic and boolean representations
+of the source language operators, and their corresponding
+inter-conversion gates. Based on these costs, the compiler, given a
+source program, automatically picks arithmetic or boolean
+representations for different parts of the program, and generates the
+corresponding circuits along with the required inter-conversion
+gates. The outcome is an effcient \mpc protocol using a combination of
+arithmetic and boolean circuits, while the programmer remains
+oblivious of all the low-level cryptographic details. Our evaluation
+shows that the performance of the
+protocols automatically generated by the \tool compiler is comparable
+to or better than the custom, specialized protocols developed in the
+previous work. Indeed, \tool is the first such cryptographic-cost
+aware compiler.
 
-\noindent\textbf{Generality and Performance.} Our compiler is aware of cryptographic costs of different representations and corresponding inter-conversions.
-Based on these costs, our compiler, given a program in \tool, automatically picks arithmetic and boolean representations for different sub-parts of the program and inserts the required inter-conversions as well.
-We demonstrate the generality of our framework by writing a wide range of functionalities in \tool such as linear regression, deep neural networks, decision trees and matrix factorization.
-Our evaluations show that automatically generated protocols using \tool are competitive with tailor-made protocols in \cite{shafindss,wu,barni,minionn,secureml,valeriaMatrix} (see \sectionref{eval}).
-These features allow us to evaluate our system state-of-the-art ML prediction algorithms, e.g., \sectionref{eval} describes evaluation on ML models from a recent ICML paper \cite{bonsai} and Tensorflow \cite{tensorflow}.
-\\
+\subsubsection*{Scalability (secure code partitioning)} \mpc tools
+often do not scale to large programs. The reason is that
+most \mpc implementations use a circuit-like representation as an
+intermediate language. Hence, the largest compute that can be done
+securely is upper-bounded by the largest circuit that can fit in the
+memory on a machine\footnote{Using swap and
+disk space is feasible but it causes huge slowdown.}. This is a
+show-stopper for applications like secure machine learning, secure
+prediction, etc. that operate on large amounts of data. \aseem{Can we
+cite something here to support this claim?}
+\tool addresses the scalability concern using a novel technique that
+we call secure code partitioning (or partitioning in short). At
+a high level, we break the original program into a sequence of small
+sub-programs, which are then sequentially processed by our
+framework, while appropriately threading the intermediate outputs
+along. While this
+addresses the scalability concern (the circuit
+sizes of the sub-programs are now small enough to fit in the memory),
+we still have to address
+the security risk of revealing the intermediate outputs. To address
+this, \tool (a) masks the intermediate outputs using a secure secret
+sharing between the two parties, and (b) appends the
+share-reconstruction function to the beginning of the subsequent
+sub-program. As we show in our evaluation, partitioning allows us to
+program large applications in \tool.
 
-\noindent\textbf{Performance Baseline.} Most works cite garbled circuit based two-party computation as the most practical approach so far and use it as the baseline for comparison of performance.
-But, it has been shown multiple times that the garbled circuits based approach does not scale for realistic examples.
-Even for small benchmarks (e.g., Na\"{i}ve Bayes) the size of garbled circuit can be huge.
-Hence, comparisons are often done on artificial data sizes and garbled circuits are shown to be 500x slower than hand-crafted protocols \cite{shafindss}.
-This is cited as the motivation for need for specialized protocols for different ML algorithms in secure prediction such as linear regression \cite{shafindss}, decision trees \cite{wu}, deep neural networks \cite{minionn} and matrix factorization \cite{valeriaMatrix}.
-Our work is based on generic \mpc and we show competitive performance with the specialized protocols contrary to the popular belief.
-Hence, we believe \tool based implementations can serve as the performance baseline for future works on specialized protocols.
-\\
+%% We provide details of partitioning and prove the
+%% correctness and security of this transformation
+%% in \sectionref{pipe}
+
+\subsubsection*{Formal guarantees} We prove a formal correctness and
+security theorem for our compiler. The correctness theorem relates the
+``ideal'' semantics (the trusted-third party semantics) of a source
+program and the ``protocol'' semantics (the distributed \mpc
+semantics that relies on circuit evaluation) of the corresponding
+compiled program. Both these semantics are instrumented
+with \emph{observations} that record the outputs. We prove that a
+well-typed source program terminates successfully in the source
+semantics. In addition, we prove that for a well-typed source program,
+its observations in the source semantics are same as the observations
+of its corresponding compiled program in the protocol
+semantics. {\aseem{Had a hard time writing this last sentence,
+suggestions welcome.}}
+For the security theorem, we formally reduce the security of our
+compiler to the security of the back-end cryptographic
+protocol.
+
+%% an intermdiate language (that is the output of our
+%% compiler), and a circuit language for arithmetic and boolean
+%% circuits.
+
+%% We give a formal {\em type
+%% system} for our language and prove useful properties. We prove that if
+%% a program type checks in our language then it would run correctly,
+%% e.g., there would not be any array index out of bounds error. The
+%% programmer writes a single code describing the inputs from each party
+%% and the function to be computed. Our compiler automatically generates
+%% separate executables for the server and the client that can be
+%% executed in the distributed setting. We prove the {\em correctness}
+%% for this compilation, i.e., the outputs obtained in the distributed
+%% execution are identical to the stand-alone execution by a trusted
+%% third party. For details on formal
+%% language, correctness and security theorems see \sectionref{ld}.
+
+%% We demonstrate the generality of our framework by writing a wide range
+%% of functionalities in \tool such as linear regression, deep neural
+%% networks, decision trees and matrix factorization. 
+%% Our evaluations show that automatically generated protocols
+%% using \tool are competitive with tailor-made protocols
+%% in \cite{shafindss,wu,barni,minionn,secureml,valeriaMatrix}
+%% (see \sectionref{eval}). These features allow us to evaluate our
+%% system state-of-the-art ML prediction algorithms,
+%% e.g., \sectionref{eval} describes evaluation on ML models from a
+%% recent ICML paper \cite{bonsai} and Tensorflow \cite{tensorflow}. \\
+
+%% \tool is a programmer-friendly
+%% framework to
+%% express the function to be computed and uses a mix of arithmetic and
+%% boolean \mpc back-ends. We illustrate the above example of $w^Tx
+%% >b$ in our language in \figureref{ex-sml}. 
+%% As is clear from the code, the programmer can remain completely
+%% oblivious of the cryptographic back-end and underlying circuit
+%% representation.
+%% Also,  the programmer is not required to write keywords such as {\em
+%% secret}
+%% or {\em public} (that are required by languages like
+%% Wysteria \cite{wysteria}).
+%%  In contrast, \figureref{ex-aby} illustrates
+%% the same example using the framework in ABY \cite{aby}. Here, as
+%% already mentioned, the programmer needs to create shares and write the
+%% low level gates
+%% like $\mathtt{MUL, ADD, CONS, GT}$, etc \TODO{make these fonts
+%% consistent}. The programmer is also required
+%% to use the appropriate share conversion gates such as $\mathtt{A2Y}$ that
+%% converts arithmetic shares to Yao-based boolean shares to be used in
+%% comparison after the inner-product computation.
+%% Moreover, it is easy to
+%% maintain code in \tool and a small change to the functionality requires
+%% only a small change in the code. In contrast, working with ABY, even
+%% for a small change a developer might have to re-work the whole circuit
+%% for efficiency reasons. We elaborate on these examples  in
+%% \sectionref{ex}. \\
+
+{\aseem{Not sure about this baseline paragraph. Did not touch it.}} 
+\subsubsection*{Performance Baseline} Most works cite garbled
+circuit based two-party computation as the most practical approach so
+far and use it as the baseline for comparison of performance. But, it
+has been shown multiple times that the garbled circuits based approach
+does not scale for realistic examples. Even for small benchmarks
+(e.g., Na\"{i}ve Bayes) the size of garbled circuit can be
+huge. Hence, comparisons are often done on artificial data sizes and
+garbled circuits are shown to be 500x slower than hand-crafted
+protocols \cite{shafindss}. This is cited as the motivation for need
+for specialized protocols for different ML algorithms in secure
+prediction such as linear regression \cite{shafindss}, decision
+trees \cite{wu}, deep neural networks \cite{minionn} and matrix
+factorization \cite{valeriaMatrix}. Our work is based on generic \mpc
+and we show competitive performance with the specialized protocols
+contrary to the popular belief. Hence, we believe \tool based
+implementations can serve as the performance baseline for future works
+on specialized protocols.
 
@@
 
-\noindent\textbf{Static Guarantees.} We give a formal {\em type system} for our language and prove useful properties.
-We prove that if a program type checks in our language then it would run correctly, e.g., there would not be any array index out of bounds error.
-The programmer writes a single code describing the inputs from each party and the function to be computed.
-Our compiler automatically generates separate executables for the server and the client that can be executed in the distributed setting.
-We prove the {\em correctness} for this compilation, i.e., the outputs obtained in the distributed execution are identical to the stand-alone execution by a trusted third party.
-We formally reduce the {\em security} of our system to the security of the cryptographic back-end.
-For details on formal language, correctness and security theorems see \sectionref{ld}.
-\\
+%% We have implemented \tool using ABY~\cite{aby} as the crypto backend.
+%% We evaluate our system on wide range of benchmarks for secure
+%% prediction \cite{shafindss,wu,barni,secureml,minionn,bonsai} using
+%% linear classifier, decision trees and deep neural networks, matrix
+%% factorization \cite{valeriaMatrix} and training of deep neural
+%% network \cite{secureml}.
 
-\noindent\textbf{Secure Code Partitioning (or Partitioning in short).} Another issue in adoption of \mpc is difficulty in scaling to large programs. 
-%In general, \mpc implementations struggle in scaling to large programssize of memory on a system limits the largest compute inside the SMC and hence,. 
-The reason being most \mpc implementations use a circuit-like representation as an intermediate representation.
-Hence, the largest compute that can be done securely is upper-bounded by the largest circuit that can fit in memory on a machine\footnote{Using swap and disk space is feasible but it causes huge slowdown.}.
-To address this issue, we develop a technique called partitioning.
-At a high level, we break the original program into a sequence of small sub-programs and mask the intermediate outputs using a secure secret sharing between the two parties.
-We appropriately append the share-reconstruction function to the beginning of the subsequent sub-program.
-Now, we execute each of these sub-programs sequentially in our framework.
-This allows us to run arbitrarily large programs.
-We provide details of partitioning and prove the correctness and security of this transformation in \sectionref{pipe}.
-\\
+%% for secure
+%% computation. We provide the first compiler where the developer writes
+%% a high-level program and the compiler automatically generates an
+%% efficient \mpc protocol that uses a
+%% mix of arithmetic and boolean circuits. 
+%% It is compatible with any implementation of \mpc that allows a mix of
+%% arithmetic and boolean circuits.
+%% We focus on semi-honest secure two-party computation and use
+%% ABY \cite{aby} as the suitable cryptographic back-end.
+%% We provide a formal type system and operational semantics for our
+%% language and prove correctness and security of our compiler.
+%% Our evaluations show that protocols generated
+%% by our compiler match
+%% the performance or outperform hand-crafted protocols in most cases (see
+%% \sectionref{eval}). \TODO{Add some text to increase enthusiasm about
+%% results.} Moreover, unlike previous systems such
+%% as \cite{wysteria,aby,cbmcgc}, our system can scale to arbitrarily
+%% large computations using our technique of {\em secure code
+%% partitioning} (see below). \divya{I think we can skip Yao pipelining
+%% here, right?}
+%% Hence, we show that our framework achieves all four properties of
+%% programmability, generality, performance, and scalability (see below
+%% for details).
+
+%% Prior to our work, there have been systems that provide only a subset of the above properties.
+For instance, using the framework of CBMC-GC \cite{cbmcgc} allows  functionalities to be expressed in high-level language C but it gives poor performance due to its usage of boolean circuits as underlying representation.
+Similarly, ABY \cite{aby} framework can potentially give good performance (due to mix of boolean and arithmetic circuits) but is not programmable.
+Yao's garbled circuits \cite{yao} work with boolean circuits and are the most widely used 2PC protocols but they  do not scale to large functions due to large memory requirement.
+%% Huang et al.
+\cite{yao-pipe} gave the technique of pipelining Yao's garbled circuits that enabled scaling to large functions but still gives poor performance due to its use of boolean circuit representation.
+%% To summarize, none of the prior frameworks have been used to evaluate the benchmarks such as secure prediction using linear regression, decision trees, deep neural networks, etc.
+In fact, it was widely believed that generic 2PC would not scale to the secure prediction and hence, various protocols were hand-crafted for this task \cite{shafindss,wu,secureml,minionn}.
+Our framework contradicts this belief (see below).
+
+%% In this work, we solve the design problem to provide a user friendly language that is expressive and can be compiled to efficient secure 2PC protocols.
+%% Programmability in \tool and performance of generated protocols has enabled us to write large programs whose circuits  exhaust the memory of the system (~32 GB).
+To combat this issue, we give a technique called secure code partitioning (see below) that allows scaling of \tool to arbitrary size computations.
+
+%% Below, we describe the salient features of \tool in more detail. \\
+
+\aseem{Checkpointing here. Two more paragraphs to come, one for
+summary of implementation and evaluation, and one for related work.}
 

### 2017-10-26 OaklandSubmission/conclude.tex
pass over sec 6 complete.

--- 2017-10-25 OaklandSubmission/eval.tex
@@
-\section{Evaluation}
-\label{sec:eval}
-We evaluate \tool on the problem of {\it secure prediction}, where one
-party (the server) has a machine learning model, and the other party
-(the client) has an input. The goal is to compute the output of the
-model on client's input, with the guarantee that the server
-learns nothing about the input, and the client learns nothing
-about the model beyond what is revealed from the output.
+\section{Conclusions}
+\label{sec:conclude}
 
-We first implement the benchmarks from Bost et al.~\cite{shafindss}
-and \minion~\cite{minionn} (both of which study the same setting),
-and show that the performance of the high-level code written in \tool
-is comparable to their hand-crafted circuits
-(Section~\ref{sec:shallow} and Section~\ref{sec:dnneval}).
-%
-Next, we demonstrate the generality and programmability aspects of
-\tool by implementing state-of-the-art machine learning models from
-Tensorflow~\cite{tensorflow} and \bonsai~\cite{bonsai}. Indeed, we
-provide the first 2PC implementation of \bonsai.
-%
-Finally, we
-implement a Deep Neural Network (DNN) for CIFAR-10
-dataset~\cite{cifar} from \minion~\cite{minionn} to illustrate the use
-of partitioning in \tool.
-
-We present the numbers for two network settings, a LAN setting and
-a cross-continent WAN setting. The round trip time between the server
-and the client machines in the two settings is 1ms and 40ms resp. Each
-machine has an Intel(R) Xeon(R) CPU E5-2673 v3 processor running at
-2.40GHz with 28 GBs of RAM.
-%
-%% The  \tool compiler is written in Python and
-%% compiles each of our benchmarks in under a second to C++ code that makes calls to the ABY
-%% library~\cite{aby}. We use an off-the-shelf solver
-%% (SeaHorn~\cite{seahorn}) to check the verification conditions required
-%% to prove that the array indices remain in bound. The solver takes less
-%% than a minute on our largest benchmark.
-Since most of our benchmarks
-are related to machine learning, we set up some notation (largely
-standard in machine learning) and describe our benchmarks next.
-
-\subsection{Benchmarks description}
-%\divya{notation can be moved somewhere... } \\
-
-\noindent(We use $[N]$ to denote $\{0,1,\dotsc, N-1\}$. 
-Further, given a vector $x\in\R^d$, we say $\mathsf{argmax}\ x = i$ if
-$x_i = \mathsf{max}\ \{x_1,\ldots,x_d\}$.)
-
-We focus on the machine learning models for {\it classification}.
-A classifier $C$ uses a trained model to {\it predict} a
-label $\ell$ for an input data point $x$. For example, given a
-data point which is a tuple of height and weight of an individual,
-a classifier can predict a label ``male'' or ``female''. The {\it
-  accuracy} of a classifier refers to the fraction of data points that
-the classifier labels correctly from a given set of test data
-points.
-
-\subsubsection*{Standard classifiers}
-A binary linear classifier is one of the simplest classifiers. Here,
-the input is a data point $x\in\R^d$,
-and the model is a vector $w\in\R^d$. The possible labels are
-$\ell\in\{\mathit{true},\mathit{false}\}$ and the classifier is
-$C_w\equiv w^Tx>0$, where $\cdot^T$ denotes the matrix transpose
-operator. This classifier requires $d$ multiplications, $d-1$
-additions, and a comparison.
-
-A more interesting classifier is Na\"{i}ve Bayes that predicts labels
-from the set $[n]$.
-Here, the input data point is a {\it feature}
-vector $x=(x_0,x_2,\ldots,x_{d-1})^T$ where each $x_j\in [F]$.
-The model has two matrices: a vector $P$ of length $n$ s.t. for each
-$i
-\in [n]$, $P(i)$ is the log-probability that the output label is $i$.
-%$P(i) \equiv \log p(\ell=i)$, where $p(\ell=i)$ is the likelihood
-%that the output label is $i$ for $i\in[n]$.
-The other matrix $T$ has size $d\times F\times n$ and for each $j\in
-[d], k\in [F], i\in [n]$, the entry $T(j)(k)(i)$ is 
-%\equiv\log p(x_j|\ell=i)$, i.e.,
-the log-probability that  the $j^{th}$ input feature is $k$  conditioned on  the output label being $i$.
-The classifier $C_{P,T}(x)$ outputs
-\[
-{\sf argmax}\ P+\sum_{j=0}^{d-1} T(j)(x_j)
-\]
-This classifier requires $ndF$ comparisons and additions. \divya{Note
-  that computing $T(i)(j)(x_j)$ requires a secret look-up in matrix
-  $T$  based on secret input $x_j$ resulting in $F$ comparisons (see
-  \sectionref{}).} \cmmt{Is there a +1 etc missing or is this exact?}
-
-A decision tree of size $N$ takes as input an $x\in\R^d$, and the
-model consists of a binary tree of depth $d$ with a boolean predicate
-assigned to each internal node. The root note is the $0^{th}$ node and
-for an internal node $j \in [N]$, the children nodes are $2j+1$ and
-$2j+2$. Each internal node $j\in[N]$ at depth $i$ has a predicate
-$b_{j}^i\equiv x_i\leq w_{j}$. We start evaluating the tree from the
-root and if the predicate at the current node $b_j^i$ is false (resp.,
-true) then $x$ is passed to the left (resp., right) child with
-predicate $b_{2j+1}^{i+1}$ (resp., $b_{2j+2}^{i+1}$). This process is
-repeated till we reach a leaf. The leaves of the tree are labels and
-the output label is the leaf visited by this traversal.
-Such a (binary) decision tree can be encoded as a polynomial (of
-degree linear in depth of the tree) and then the prediction task
-reduces to polynomial evaluation. E.g., a binary tree with depth one
-and size $N=3$ can be encoded as the polynomial
-$(1-b(x))\ell_0 + b(x)\ell_1$ where $b(x)$ is the predicate at the
-root, $\ell_0$ is the label of the left leaf and $\ell_1$ is the label
-of the right leaf.
-
-\subsubsection*{Deep neural nets}
-The next class of classifiers that we benchmark are deep neural nets
-or DNNs. A DNN has multiple layers such that each layer computes a
-matrix
-multiplication followed by an {\it activation} function $f$. The most
-common activation functions are square $f(x)=x^2$ and rectifier linear
-unit (ReLU) $f(x)=\mathsf{max}(x,0)$.
-If $A$ is a matrix then the output of $f(A)$ is also a matrix that is
-obtained by applying $f$ to each entry of $A$ pointwise.
-Given an input vector $x$, the predicted label of a DNN is
-\[
- \mathsf{argmax}\ W_N\cdot f_{N-1}(\ldots f_1(W_1\cdot x)\ldots)
-\]
-Here, $f_i$'s are the (public) activation functions, the model
-consists of matrices $W_i$,  $x \in \R^d$ is the input vector, and the
-operator $\cdot$ denotes a matrix multiplication.
-Neural nets usually have one or more fully connected layers, each of
-which multiplies a matrix with a vector.
-Some neural nets have convolution layers and such DNNs are also called
-Convolutional Neural Nets or CNNs.
-For the purpose of this paper, a convolution can be considered as a
-(heavy) matrix-matrix multiplication. The size of matrices manipulated
-by a convolution layer grows linearly with {\it window size}
-(typically 9 or 25), the number of {\it output channels} (typically
-16, 32, or 64), and the size of the matrix input to this layer.
-Therefore, fully connected layers are lighter computation-wise compared
-to convolution layers. However, the model size of fully connected
-layers is larger than those of convolution layers.
-In general, DNNs are computationaly heavy but provide
-much better accuracies on computer vision tasks than the classifiers
-discussed above.
-
-\subsubsection*{State-of-the-art classifiers}
-Finally, there are a class of machine learning classifiers that are
-much more efficient than
-DNNs and provide reasonably good accuracies on standard learning
-tasks. \bonsai~\cite{bonsai} is a state-of-the art classifier in this
-class and \tool provides the first 2PC protocol for it.
-\bonsai takes as input $x \in \R^d$, and its model consists of a
-binary tree with $N$ nodes, and a matrix $Z$. Each node $j$ contains
-matrices $W_j$ and $V_j$, and a vector $\theta_{j}$. The internal node
-$j$  evaluates a predicate $\theta_j^TZx > 0$ to decide whether
-to pass $x$ to the left child $2j+1$ or the right child $2j+2$.
-The predicted value is
-
-\aseem{I had a hard time parsing this. (1) Should we use $\cdot$ for
-  matrix multiplication as we did for DNNs? (2) Precedence of $\circ$
-  over scalar multiplication (or is it associative so doesn't matter?)
-(3) The argument to $f$ is a scalar (right?), and so is its output,
-so then is $\circ$ defined OR is it applying $f$ pointwise to a vector?}
-
-\[
-{\sf argmax} \sum_{j = 0}^{N - 1} I_j(x)W_j^TZx\circ f(V_j^TZx) 
-\]
-
-Here, $I_j(x)$ is 1 if the $j^{th}$ node is present on the path traversed by $x$
-and is zero otherwise. 
-The operation $\circ$ is a pointwise dot product, $W_j$'s and $V_j$'s
-are matrices. The activation function $f$ is given by $f(y) = y$ if
-$-1 < y < 1$ and $\mathrm{sign}(y)$ otherwise. \bonsai can be
-seen as a variant of decision trees where the prediction is a function
-of the path traversed from root to the leaf and not just the leaf
-itself.
-
-In the following, we implement these classifiers in \tool and report
-the time taken for making secure predictions. Ideally, the machine
-learning classifiers are mathematical expressions over $\R$ that are
-usually approximated by floating-point operations. 
-As is standard, we port the classifers to integer manipulating
-programs by scaling the models and rounding~\cite{minionn}. These
-ported classifiers are then implemented in \tool.
-
-\subsection{Standard classifiers}
-\label{sec:shallow}
-We train the three standard classifiers, linear, Na\"{i}ve Bayes, and
-decision trees, on the following data sets from the UCI
-machine learning repository~\cite{uci}:
- the Wisconsin Breast Cancer data set, 
-Credit Approval data set, Audiology (Standardized) data set, Nursery
-data set, and ECG (electrocardiogram) classification data
-from~\cite{barni}.
-
-\subsubsection*{Linear classifier}The results for linear
-classification are in Table~\ref{tab:lc}. The
-input and the model are both vectors of length $d$. The columns
-``Prev. time'' and ``Prev. comm'' show the time and the total
-network communication reported by Bost et al.~\cite{shafindss} for a
-network setting with 40ms
-round trip time, which is same as our WAN setting. The total execution
-time
-of \tool generated code in the LAN and the WAN setting is reported
-next, followed by the total communication.
-We observe that the \tool code performance matches that of Bost
-et al., while the programmer effort in \tool is just 20
-lines (last column in the table) of high-level code in the \tool
-source language, as
-opposed to writing low-level circuits. \aseem{They write circuits,
-  right?} Our execution time is dominated by the multiplications and
-boolean-ands.
-The number of boolean-and gates (\#And) remains constant for both the
-benchmarks as they reflect the number of gates required for a
-single 32-bit integer comparison with zero, while the number of
-multiplication gates (\#Mul) is the same as the model size $d$.
-%% The
-%% last column shows the lines of code of \tool program, which is also
-%% independent of the model size.
-
-\subsubsection*{Na\"{i}ve Bayes} The results for Na\"{i}ve Bayes are
-in Table~\ref{tab:nb}. As before, $d$ denotes the length of the input
-feature vector $x$ and $F$ is the number of possible values of $x_i$
-for each $i \in [d]$.
-Although, there are no multiplications in these benchmarks, these do
-have a significant number of comparisons (due to array look-ups using
-secret indices) that
-result in many boolean-and gates. %(13k/750k denote 13000/75000). 
-As before columns ``Prev. time'' and ``Prev. comm.'' (columns 4 and 5)
-show the numbers reported
-by Bost et al.~\cite{shafindss}. We observe that \tool generated code
-has better performance, despite using a generic 2PC,
-as opposed to custom designed protocols developed by Bost et
-al. Moreover, they remark that in their setup, generic Yao-based 2PC
-did not scale to the smallest of their Na\"{i}ve Bayes classifiers, so
-they had to scale down the prediction task, and even then their
-implementation was 500x slower. Whereas, we show that by using a
-cryptographic-cost aware compiler, we can scale generic 2PC to real
-prediction tasks, and get performance competitive to or better than the
-specialized protocols.
-
-\subsubsection*{Decision trees}
-Our decision trees numbers in Table~\ref{tab:dt} further validate
-this claim. We compare with Wu et al.~\cite{wu} (columns 3 and 4) who
-describe specialized protocols for trees and forests (that perform
-better than~\cite{shafindss}).
-Similar to Bost et al.~\cite{shafindss}, they also remarks that these
-computations are infeasible with Yao-based 2PC.
-Again, we observe that
-the performance of \tool generated generic 2PC
-code is competitive with specialized protocols. Therefore, we believe
+Therefore, we believe
 that \tool generated code can be a better baseline than Yao-based 2PC
@@
 prediction tasks. \aseem{I don't understand this last sentence. Why is
-  ours a better baseline?}
-
-\begin{table*}
-\begin{tabular}{c|c|c|c|c |c|c|c|c|c|c}
-Dataset & $d$  & Prev. time (s) & Prev. comm (kb) & LAN time (s) & WAN time (s) & Comm. (kb)  & \#And & \#Mul & \#Gates & LOC\\
-\hline
-Breast cancer & 30 & 0.3 & 36 & 0.1 & 0.3 & 25 & 95 & 30 & 727 & 20\\
-\hline
-Credit & 47 & 0.3 & 41 & 0.1 & 0.3 & 36 & 95 & 47 & 795 & 20\\
-\hline
-\end{tabular}
-
- \caption{Linear classification results. We compare our results (columns 5, 6, 7) with~\cite{shafindss} (columns 3 and 4)}
- \label{tab:lc} 
-\end{table*}
-
-\begin{table*}
-\begin{tabular}{c|c|c|c|c|c |c|c|c|c|c|c}
-Dataset & $d$ & $F$ & Prev. time (s) & Prev. Comm (Mb) & LAN time (s) & WAN time (s) & Comm. (Mb)  & \#And & \#Mul & \#Gates  & LOC\\
-\hline
-Nursery & 5 & 9 & 1.5 & 0.2 & 0.1 & 0.4 & 0.6 & 13k & 0 & 73k & 50\\
-\hline
-Audiology & 24 & 70 & 3.9 & 2.0 & 1.5 & 2.9 & 37 & 750k & 0 & 4219k & 50\\
-\hline
-\end{tabular}
-
- \caption{Na\"{i}ve Bayes results. We compare our results (columns 6 , 7, 8) with~\cite{shafindss} (columns 4 and 5)}
- \label{tab:nb} 
-\end{table*}
-
-\begin{table*}
-\begin{tabular}{c|c|c|c |c|c|c|c|c|c|c}
-Dataset  & $N$ & Prev. time (s) & Prev. Comm (kb) & LAN time (s) & WAN time (s) & Comm. (kb)  & \#And & \#Mul & \#Gates & LOC\\
-\hline
-Nursery & 4 & 0.3 & 102 & 0.1 & 0.3 & 32 & 504 & 3 & 3324 & 20\\
-\hline
-ECG &  6 & 0.4 & 102 & 0.1 & 0.4 & 49 & 756 & 5 & 5002 & 20\\
-\hline
-\end{tabular}
-
- \caption{Decision tree benchmarks. We compare our results (columns 5 , 6, 7) with~\cite{wu} (columns 3 and 4)}
- \label{tab:dt} 
-\end{table*}
-
-%dataset  | size | gates | and | mul | depth | length
-
-%dataset | time | comm | timeL | timeW | comm
-\subsection{Deep neural nets}
-\label{sec:dnneval}
-We evaluate \tool on the DNNs described in SecureML~\cite{secureml},
-Cryptonets~\cite{cryptonets}, and CNN from \minion~\cite{minionn}. For
-comparison, we consider their implementations from
-\minion~\cite{minionn}, which outperforms their previous
-implementations. Table~\ref{tab:nn} shows the
-results\footnote{\minion does not report the network round-trip time.}.
-We note that for each of these DNNs, \minion provides a
-specialized protocol, while \tool uses a generic 2PC protocol
-(auto) generated from high-level code.
-
-The first benchmark  is the DNN described in
-SecureML~\cite{secureml}.
-It has three fully connected layers with square as the activation function.
-This DNN is light-weight as it requires only arithmetic computations
-(additions and multiplications) 
-except for the $\mathsf{argmax}$ at the end.
-Next, we implement the DNN described in Cryptonets~\cite{cryptonets}
-in \tool.
-This DNN also uses square as the activation function and has one
-convolution (with 5 output channels) and one fully connected layer. 
-Finally, we implement CNN from \minion, that has
-two convolutions (with 16 output channels each) and two fully
-connected layers.  In contrast to the previous two DNNs, it uses
-ReLU for activation and
-has significantly higher number of boolean-and gates.
-Recall, that square activation can be implemented entirely using
-arithmetic gates but ReLU requires boolean-and gates.  For a complete
-description of these benchmarks and their accuracies, we refer the
-reader to the original references.
-
-In Table~\ref{tab:nn}, columns 2 and 3 show the time taken by \minion.
-The column ``Model size'' is the number of 32 bit integers in the
-trained model.
-We observe that our performance is competitive with specialized
-\minion protocols, for both LAN and the WAN settings. Further, lines
-of \tool source code required is still small. 
-We note that \minion and \tool implementations are both based on ABY.
-These results are surprising as \minion has a much more efficient
-preprocessing phase and has SIMD 
-(single instruction multiple data) capabilities that allow it to perform
-matrix operations efficiently. In contrast, \tool focuses on generic
-2PC, and such optimizations are part of our future work.
-\minion also reports performance results on a bigger DNN with 7
-convolution layers. In \tool, this benchmark requires partitioning and
-we discuss it in Section~\ref{sec:pipeeval}.
-%net | minionT | minionC | timeL | timeW | comm | gates | and | mul | length
-
-\begin{table*}
-\begin{tabular}{c|c|c|c |c|c|c|c|c|c|c}
-DNN  & Prev. time (s) & Prev. Comm (Mb) & LAN time (s) & WAN time (s) & Comm. (Mb)  & \#And & \#Mul & \#Gates & Model Size & LOC\\
-\hline
-SecureML   &  1.1 & 15.8 & 0.7 & 1.7  & 76   &  2k   & 119k & 366k   & 119k & 78\\
-\hline
-Cryptonets &  1.3 & 47.6 & 0.6 & 1.6  & 70    & 2k    & 108k & 316k & 86k & 88\\
-\hline
-CNN        &  9.4 & 657.5& 5.1 & 11.6 & 501  & 1640k & 667k & 9480k & 35k & 154\\
-\hline
-\end{tabular}
-
- \caption{DNN benchmarks. We compare our results (columns 4 , 5, 6) with~\cite{minionn} (columns 2 and 3)}
- \label{tab:nn} 
-\end{table*}
-
-\subsection{State-of-the-art classifiers}
-Tensorflow~\cite{tensorflow} is a standard machine learning toolkit.
-Its introductory tutorial describes two prediction models for handwritten digit recognition
-using the MNIST dataset~\cite{mnist}.
-Each image in this dataset is a greyscale $28\times 28$ image of
-digits 0 to 9.
-The first model that the tutorial describes is a softmax regression
-that provides an accuracy of 92\%. The classifier evaluates
-\[
- \mathsf{argmax}\ Wx+b
-\]
-Here, $i\in [10]$, $x$ is a 784 length vector obtained from the input image,
-each $W$ is a $10\times 784$ matrix, and each $b$ is a $784$ length
-vector. \aseem{Questions on this expression: (1) where is $i$?
-  (2) Should $b$ be 10 length vector? (3) Again, should be use $\cdot$
-  for matrix multiplication
-as we did earlier to be consistent?}
-We implement this classifer in \tool and present the results in the
-first row of Table~\ref{tab:tf}.
-Since, we are not aware of any other tools that has used this
-model as a benchmark, we only report numbers for \tool.
-
-The next classifer in the Tensorflow tutorial is a convolution neural net with two convolutions
-(with 32 output channels) and two fully connected layers with ReLU as the activation function.
-This DNN is both bigger and more accurate than the DNNs presented in the previous section.
-In particular, it has an accuracy of 99.2\%\footnote{The accuracy of
-  the integer model is 0.04\% higher than the floating-point
-  model.}. \aseem{Is this footnote necessary? As a reader, I won't
-  miss it if it was not there.}
-We observe that this DNN can take a minute per prediction in the WAN
-setting and is the largest benchmark that we have evaluated without
-partitioning.
-
-% benchmark | size | gates | and | mul | depth  | length | comm | timeL | timeW
-
-\begin{table*}
-\begin{tabular}{c|c|c|c |c|c|c|c|c|c | c}
-Classifier       & LAN time (s) & WAN time (s) & Comm. (Mb)  & \#And & \#Mul & \#Gates & Model size & LOC\\
-\hline
-Regression &  0.1         & 0.7         & 5            & 2k    & 8k    &  35k    & 8k   & 38\\
-\hline
-CNN        &  30.5        & 60.3        & 2955         & 6082k & 4163k &  42104k & 3226k& 172\\
-\hline
-\end{tabular}
-
- \caption{Tensorflow tutorial benchmarks}
- \label{tab:tf} 
-\end{table*}
-
-We next present \bonsai~\cite{bonsai} results on three
-standard datasets: character recognition (Chars4k~\cite{campos},
-accuracy 74.71\%), text recognition (USPS~\cite{hull}, accuracy
-94.4\%), and object categorization (WARD~\cite{yang}, accuracy
-95.7\%). 
-The \bonsai models are already over integers and no porting from
-floating-point is required.
-We implement the trained classifiers in \tool for all the benchmarks
-from~\cite{bonsai},
-and show the represenative results in Table~\ref{tab:bonsai}.
-Out of all the benchmarks from~\cite{bonsai}, the dataset WARD
-requires the largest model. 
-The column ``depth''  shows the depth of the
-tree used by \bonsai. The size of \tool 
-program grows with the depth of the tree, as the straightforward \tool
-implementation requires a loop for each layer of the
-tree. \aseem{Should we remark here that once we have functions in the
-  language, LOC would come down, perhaps in the footnote? Else a PL
-  reviewer might scoff at it :).}
-
-To summarize, by providing first 2PC implementations of
-state-of-the-art classifiers, we have demonstrated the expressiveness
-of \tool. We discuss scalability next.
-%dataset | size | gates | and | mul | depth | length | comm | timeL | timeW
-
-\begin{table*}
-\begin{tabular}{c|c|c|c |c|c|c|c|c|c | c}
-Dataset       & LAN time (s) & WAN time (s) & Comm. (Mb)  & \#And & \#Mul & \#Gates & depth & LOC\\
-\hline
-Chars4k    &  0.1         & 0.7         & 2            & 18k    & 3k    &  85k     & 1   & 89\\
-\hline
-USPS       &  0.2         & 0.9         & 4            & 62k    & 2k    &  285k    & 2   & 156\\
-\hline
-WARD       &  0.3         & 1.1         & 9            & 106k    & 8k    &  506k    & 3   & 283\\
-\hline
-\end{tabular}
-
- \caption{Bonsai benchmarks}
- \label{tab:bonsai} 
-\end{table*}
-
-\subsection{Partitioning}
-\label{sec:pipeeval}
-The largest benchmark of \minion~\cite{minionn} is a DNN for CIFAR-10
-dataset~\cite{cifar}. 
-The classifier's task is to categorize colored ($32\times 32$) images
-into 10 classes. A secure evaluation of this DNN needs more memory
-than what is available
-on our machines \aseem{Is the limitation machine RAM or ABY
-  limit?}. Therefore, we use partitioning and divide the
-computation into seven stages. 
-The first step does a convolution with 64 output channels and a ReLU
-activation. The next four stages together perform a convolution that
-involves multiplying a $64\times 576$ matrix with a 
-$576\times 1024$ matrix. The sixth stage performs a ReLU and a
-convolution. The final stage has four convolutions, five ReLUs, and a
-fully connected layer. The total number of lines of \tool code for
-this benchamrk is 336 lines. \aseem{Would it be better if we had
-  functions, for loop expressions etc.? If so, we should add a
-  footnote.}
-
-Table~\ref{tab:cifar} shows the end-to-end numbers as
-well as the numbers for the sixth (the heaviest) stage. 
-The number of gates are in millions, hence the suffix `m' in the last
-three columns.
-For this DNN, the specialized \minion protocol takes 544 seconds and
-communicates 9272 Mb.
-As with Table~\ref{tab:nn}, \tool generated generic 2PC protocol is
-competitive with \minion here as well. Therefore,
-we believe that with partitioning, \tool can scale to arbitrary sized
-computations while maintaining performance
-competitive with existing specialized protocols. In particular, for a
-large enough DNN, \minion would run
-out of memory but a pipelined \tool implementation would still
-succeed. \aseem{Last sentence is speculative? Should we say \minion
-  would or could or might?}
-
-\begin{table}
-\begin{tabular}{c|c|c|c |c|c|c}
-           &  LAN (s) & WAN (s) & Comm. (Mb)  & \#And & \#Mul & \#Gates \\
-\hline
-Total      &  265.6       & 647.5        & 40683       & 21m    & 61m    &  337m  \\
-\hline
-Stage 6    &  55.2        & 122.6        & 6744        & 12m    & 10m   &  98m  \\
-\hline
-\end{tabular}
-
- \caption{Partitioning results for CIFAR-10. \minion takes 544 seconds and communicates 9272 Mb to evaluate this DNN.}
- \label{tab:cifar} 
-\end{table}
-
-%stage | time | comm | gates | mul | add | depth
+  ours a better baseline?} \nc{This last line is more a comment for a ``conclusion'' and is a bit out of place here.
+I guess we want to say that till recently if generic 2PC using Yao was not feasible, then people had to come up with specialized protocols and didnt have any benchmark or had a poor benchmark (of Yao) to compare their 2PC with.
+Moving forward, our language generated 2PC can be the baseline benchmark for all 2PC protocols.}

### 2017-10-26 OaklandSubmission/eval.tex
pass over sec 6 complete.

@@
 \label{sec:eval}
-We evaluate \tool on a variety of problems that can be thought to fall under the umbrella of \nc{saying it like this so that it does not seem we are evaluating on one problem.} {\it secure prediction}, where one
+We evaluate \tool on a variety of problems that can fall under the umbrella of {\it secure prediction}, where one
 party (the server) has a machine learning model, and the other party
@@
 
-\noindent(We use $[N]$ to denote $\{0,1,\dotsc, N-1\}$. 
-Further, given a vector $x\in\R^d$, we say $\mathsf{argmax}\ x = i$ if
-$x_i = \mathsf{max}\ \{x_1,\ldots,x_d\}$.) \nc{Why is notation in brackets?}
-
+We use $[N]$ to denote $\{0,1,\dotsc, N-1\}$.
+Further, given a vector $x\in\R^d$, we say $\mathsf{argmax}\ x = i$ if $x_i = \mathsf{max}\ \{x_1,\ldots,x_d\}$.
 We focus on the machine learning models for {\it classification}.
@@
 data point which is a tuple of height and weight of an individual,
-a classifier might be able to predict a label as ``male'' or ``female''\nc{Is this less sexist?}. The {\it
+a classifier might be able to predict a label as ``male'' or ``female''. The {\it
   accuracy} of a classifier refers to the fraction of data points that
@@
 $\ell\in\{\mathit{true},\mathit{false}\}$ and the classifier is
-$C_w\equiv w^Tx>0$, where $\cdot^T$ denotes the matrix transpose
-operator. \nc{it might be obvious, but should we say $x > y$ returns 1 if $x>y$ and 0 otherwise?} This classifier requires $d$ multiplications, $d-1$
-additions, and a comparison.
+$C_w\equiv w^T\matmul x>0$, where $w^T$ denotes the transpose
+of matrix $w$, $\matmul$ denotes matrix multiplication (with a vector in this case) and $a>b$ returns $1$ if $a>b$ and $0$ otherwise. This classifier requires $d$ multiplications, $d-1$
+additions, and a single comparison.
 
@@
 This classifier requires $ndF$ comparisons and additions.
-Note here  that computing $T(i)(j)(x_j)$ requires a secret look-up in matrix  $T$  based on secret input $x_j$ resulting in $F$ comparisons using multiplexers (see
+Note here  that computing $T(i)(j)(x_j)$ requires a secret look-up in matrix  $T$  based on secret input $x_j$ resulting in $F-1$ comparisons using multiplexers (see
   Section~\ref{sec:impl}).
-\nc{Made DG's comment as part of regular text - worth point out}.
+%\nc{Made DG's comment as part of regular text - worth point out}.
 \cmmt{Is there a +1 etc missing or is this exact?} \nc{By our calculation in Sec 5, it seems that this should be $F/2$ multiplexers}
@@
 \[
- \mathsf{argmax}\ W_N\cdot f_{N-1}(\ldots f_1(W_1\cdot x)\ldots)
+ \mathsf{argmax}\ W_N\matmul f_{N-1}(\ldots f_1(W_1\matmul x)\ldots)
 \]
@@
 consists of matrices $W_i$,  $x \in \R^d$ is the input vector, and the
-operator $\cdot$ denotes a matrix multiplication.
+operator $\matmul$ denotes a matrix multiplication.
 Neural nets usually have one or more fully connected layers, each of
@@
 
-\aseem{I had a hard time parsing this. (1) Should we use $\cdot$ for
-  matrix multiplication as we did for DNNs? (2) Precedence of $\circ$
-  over scalar multiplication (or is it associative so doesn't matter?)
-(3) The argument to $f$ is a scalar (right?), and so is its output,
-so then is $\circ$ defined OR is it applying $f$ pointwise to a vector?}
-\nc{(1) I agree we should do that to be consistent with before.
+%\aseem{I had a hard time parsing this. (1) Should we use $\cdot$ for
+%  matrix multiplication as we did for DNNs? (2) Precedence of $\circ$
+%  over scalar multiplication (or is it associative so doesn't matter?)
+%(3) The argument to $f$ is a scalar (right?), and so is its output,
+%so then is $\circ$ defined OR is it applying $f$ pointwise to a vector?}
+%\nc{(1) I agree we should do that to be consistent with before.
 (2) I assumed $f()$ is computed to give a vector somehow and $I_j(x)W_j^TZx$ also gives a vector somehow and then we do $\circ$; I might be wrong.
@@
 \[
-{\sf argmax} \sum_{j = 0}^{N - 1} I_j(x)W_j^TZx\circ f(V_j^TZx) 
+{\sf argmax} \sum_{j = 0}^{N - 1} I_j(x)[ (W_j^T\matmul Z\matmul x) \circ (f(V_j^T\matmul Z\matmul x))] 
 \]
@@
 The operation $\circ$ is a pointwise dot product, $W_j$'s and $V_j$'s
-are matrices. The activation function $f$ is given by $f(y) = y$ if
-$-1 < y < 1$ and $\mathrm{sign}(y)$ otherwise. \bonsai can be
+are matrices of appropriate dimension. The activation function $f$ is given by $f(y) = y$ if
+$-1 < y < 1$ and $\mathrm{sign}(y)$ otherwise ($f$ on vectors and matrices are defined point-wise on every element of the vector/matrix). \bonsai can be
 seen as a variant of decision trees where the prediction is a function
@@
 \label{sec:shallow}
-We train the three standard classifiers, linear, Na\"{i}ve Bayes, and
+We train (on cleartext data) the three standard classifiers, linear, Na\"{i}ve Bayes, and
 decision trees, on the following data sets from the UCI
@@
 data set, and ECG (electrocardiogram) classification data
-from~\cite{barni}. \nc{Say here that the training is done on cleartext data?}
+from~\cite{barni}.
 
@@
 next, followed by the total communication.
-We observe that the \tool code performance matches that of Bost
-et al., while the programmer effort in \tool is just 20
+We observe that the \tool code performance matches the hand-crafted protocol of Bost
+et al., and the programmer effort in \tool is just 20
 lines (last column in the table) of high-level code in the \tool
-source language, as
-opposed to writing low-level circuits. \aseem{They write circuits,
-  right?} \nc{Actually, they have a specific protocols for dot product and comparison and also say that their developer effort is minimal as they have to write two function calls to these protocols :).} Our execution time is dominated by the multiplications and
+source language. %\aseem{They write circuits,
+ % right?} \nc{Actually, they have a specific protocols for dot product and comparison and also say that their developer effort is minimal as they have to write two function calls to these protocols :).}
+ Our execution time is dominated by the multiplications and
 boolean-ands.
@@
 the performance of \tool generated generic 2PC
-code is competitive with specialized protocols. Therefore, we believe
-that \tool generated code can be a better baseline than Yao-based 2PC
-when comparing against generic 2PC-based implementations for secure
-prediction tasks. \aseem{I don't understand this last sentence. Why is
-  ours a better baseline?} \nc{This last line is more a comment for a ``conclusion'' and is a bit out of place here.
-I guess we want to say that till recently if generic 2PC using Yao was not feasible, then people had to come up with specialized protocols and didnt have any benchmark or had a poor benchmark (of Yao) to compare their 2PC with.
-Moving forward, our language generated 2PC can be the baseline benchmark for all 2PC protocols.}
+code is competitive with specialized protocols.
 
@@
 \begin{tabular}{c|c|c|c|c |c|c|c|c|c|c}
-Dataset & $d$  & Prev. time (s) & Prev. Comm (kb) & LAN time (s) & WAN time (s) & Comm. (kb)  & \#And & \#Mul & \#Gates & LOC\\
+Dataset & $d$  & Prev. time (s) & Prev. Comm (kb) & LAN (s) & WAN (s) & Comm. (kb)  & \#And & \#Mul & \#Gates & LOC\\
 \hline
@@
 \begin{tabular}{c|c|c|c|c|c |c|c|c|c|c|c}
-Dataset & $d$ & $F$ & Prev. time (s) & Prev. Comm (Mb) & LAN time (s) & WAN time (s) & Comm. (Mb)  & \#And & \#Mul & \#Gates  & LOC\\
+Dataset & $d$ & $F$ & Prev. time (s) & Prev. Comm (Mb) & LAN (s) & WAN (s) & Comm. (Mb)  & \#And & \#Mul & \#Gates  & LOC\\
 \hline
@@
 implementations. Table~\ref{tab:nn} shows the
-results\footnote{\minion does not report the network round-trip time.}.
+results\footnote{\minion does not report the network round-trip time nor the bit-length of their inputs (we use $32$-bit inputs).}.
 We note that for each of these DNNs, \minion provides a
@@
 (additions and multiplications) 
-except for the $\mathsf{argmax}$ at the end.
+except for the $\mathsf{argmax}$ at the end of the network.
 Next, we implement the DNN described in Cryptonets~\cite{cryptonets}
@@
 \begin{tabular}{c|c|c|c |c|c|c|c|c|c|c}
-DNN  & Prev. time (s) & Prev. Comm (Mb) & LAN time (s) & WAN time (s) & Comm. (Mb)  & \#And & \#Mul & \#Gates & Model Size & LOC\\
+DNN  & Prev. time (s) & Prev. Comm (Mb) & LAN (s) & WAN (s) & Comm. (Mb)  & \#And & \#Mul & \#Gates & Model Size & LOC\\
 \hline
@@
 \[
- \mathsf{argmax}\ Wx+b
+ \mathsf{argmax}\ W\matmul x+b
 \]
-Here, $i\in [10]$, $x$ is a 784 length vector obtained from the input image,
-each $W$ is a $10\times 784$ matrix, and each $b$ is a $784$ length
+Here, $x$ is a 784 length vector obtained from the input image,
+each $W$ is a $10\times 784$ matrix, and each $b$ is a $10$ length
 vector. \aseem{Questions on this expression: (1) where is $i$?
@@
   for matrix multiplication
-as we did earlier to be consistent?}
+as we did earlier to be consistent?} \nc{I think I have fixed this.
+OK now?
+$\mathsf{argmax}$ throughout this section has implicit use of a dimension and I have left it like that.
+We can make it explicit if others feel the need.}
 We implement this classifer in \tool and present the results in the
@@
   model.}. \aseem{Is this footnote necessary? As a reader, I won't
-  miss it if it was not there.}
+  miss it if it was not there.}\nc{I agree. Happy to remove.}
 We observe that this DNN can take a minute per prediction in the WAN
@@
 % benchmark | size | gates | and | mul | depth  | length | comm | timeL | timeW
-
-\begin{table*}
+\setlength\tabcolsep{2.5pt}
+\begin{table}
 \begin{tabular}{c|c|c|c |c|c|c|c|c|c | c}
-Classifier       & LAN time (s) & WAN time (s) & Comm. (Mb)  & \#And & \#Mul & \#Gates & Model size & LOC\\
+Classifier       & \thead{LAN  \\ (s)} & \thead{WAN \\(s)} & \thead{Comm. \\(Mb)}  & \#And & \#Mul & \#Gates & \thead{Model \\size} & LOC\\
 \hline
@@
  \label{tab:tf} 
-\end{table*}
+\end{table}
 
@@
 implementation requires a loop for each layer of the
-tree. \aseem{Should we remark here that once we have functions in the
+tree\footnote{We remark here that our current language does not support functions (which we leave for future work) and with this support, LOC would be even lower.}.\aseem{Should we remark here that once we have functions in the
   language, LOC would come down, perhaps in the footnote? Else a PL
-  reviewer might scoff at it :).}
+  reviewer might scoff at it :).}\nc{added a footnote.}
 
@@
 
-\begin{table*}
+\setlength\tabcolsep{4pt}
+\begin{table}
 \begin{tabular}{c|c|c|c |c|c|c|c|c|c | c}
-Dataset       & LAN time (s) & WAN time (s) & Comm. (Mb)  & \#And & \#Mul & \#Gates & depth & LOC\\
+Dataset       & \thead{LAN \\(s)} & \thead{WAN \\(s)} & \thead{Comm. \\ (Mb)}  & \thead{\#\\And} & \thead{\#\\Mul} & \thead{\#\\Gates} & depth & LOC\\
 \hline
@@
  \label{tab:bonsai} 
-\end{table*}
-
+\end{table}
+\setlength\tabcolsep{6pt}
 \subsection{Partitioning}
@@
 on our machines \aseem{Is the limitation machine RAM or ABY
-  limit?}. Therefore, we use partitioning and divide the
+  limit?}\nc{ABY I think. But ABY's limitation is because of RAM I think}. Therefore, we use partitioning and divide the
 computation into seven stages. 
@@
 succeed. \aseem{Last sentence is speculative? Should we say \minion
-  would or could or might?}
+  would or could or might?}\nc{I think \minion definitely would, but we could say could to be nice.}
 
@@
 \begin{tabular}{c|c|c|c |c|c|c}
-           &  LAN (s) & WAN (s) & Comm. (Mb)  & \#And & \#Mul & \#Gates \\
+           &  LAN (s) & WAN (s) & \thead{Comm. \\ (Mb)}  & \#And & \#Mul & \#Gates \\
 \hline

### 2017-10-26 OaklandSubmission/train.tex
pass over sec 6 complete.

@@
 \begin{tabular}{c|c|c|c |c|c| c}
-  Stage         &  LAN (s) & WAN (s) & Comm. (Mb)  & depth & \#Gates & LOC\\
+  Stage         &  LAN (s) & WAN (s) & \thead{Comm. \\ (Mb)}  & depth & \#Gates & LOC\\
 \hline
@@
 In the future, we would like to add support for functions and recursion.
+\nc{Should we say this in the implementation section or something and not bring it up in the experiments section?
+Also, do reviewers sometimes say ``Why didnt you do that also?'' Anyway to prevent that?}
 However, the current programmer effort seems miniscule compared to the mammoth implementation effort

### 2017-10-26 OaklandSubmission/intro.tex

@@
 
-\subsubsection*{Cryptogaphic-cost awareness} \mpc compiler is aware
+\subsubsection*{Cryptogaphic-cost awareness} \tool compiler is aware
 of the cryptographic costs of arithmetic and boolean representations

### 2017-10-26 OaklandSubmission/impl.tex
impl

@@
 \label{sec:impl}
+We discuss some  implementation details of \tool.
 The  \tool compiler is written in Python and
@@
 
-Some of our benchamrks require accessing arrays at secret location.
+Some of our benchamrks require accessing arrays at secret locations.
 For example, consider the expression $A[x]$ where $A$ is an array of size 2 and $x$ is shared.
-\tool enforces the array indices to be public. The developer expresses this functionality in \tool
+\tool enforces the array indices to be public as accessing arrays at secret locations is expensive.
+However, the developer can still express this functionality in \tool
 as $x > 0 ? A[1] : A[0]$. In general, a secret access to an array of size $2^n$ requires $2^{n}-1$ multiplexers in \tool.
 
- Secure code partitioning currently requires a manual step.
-In particular, the developer needs to decompose a program $P$ into $P_1||\ldots||P_k$ manually (Section~\ref{sec:pipe}). Then \tool generates $Q_1||\ldots||Q_k$ automatically.
-Automating the decomposition step requires an analysis that can 
-statically estimate the resource usage of a \tool program and we leave the construction of such
-an analysis as future work.
-
-We use an off-the-shelf solver
-(SeaHorn~\cite{seahorn}) to check the verification conditions required
-to prove that the array indices remain in bound ($\Gamma\vDash e < n$ in {\sc T-Read}, Figure~\ref{fig:compile}). The solver takes less
-than a minute on our largest benchmark.
-The type label is assigned conservatively in  {\sc T-Decl}.
+We provide implementation details of compilation rules {\sc T-Read} and {\sc T-Decl} (Figure~\ref{fig:compile}).
+In particular, the check $\Gamma\vDash e < n$ in {\sc T-Read}
+which prove that the array indices are appropriately bounded
+is discharged by an off-the-shelf solver (SeaHorn~\cite{seahorn}).
+The solver takes less than a minute on our largest benchmark.
+Our implementation of {\sc T-Decl} assigns type labels conservatively.
 Only the variables that govern the control flow, i.e., variables
-in if-conditions and for-loops are assigned public labels.
+in the predicates of if-conditions and loop counters of for-loops are assigned public labels.
 All other variables are  assigned arithmetic labels (and can later be coerced to boolean).
@@
 
-The compilation rules of Figure~\ref{fig:compile} can introduce redundant coercions from arithmetic to
+Finally, the compilation rules in Figure~\ref{fig:compile} can introduce redundant coercions from arithmetic to
 boolean  and vice versa.
-The \tool compiler performs several optimizations to reduce these.
-For example, on each coercion, \tool memorizes the pair of arithmetic share and boolean share involved in the coercion.
+The  compiler performs several optimizations to reduce these.
+For example, on each coercion, \tool stores the pair of arithmetic share and boolean share involved in the coercion.
 \tool invalidates such pairs when the variables corresponding to the shares are overwritten by assignments.
 In subsequent coercions, \tool  reuses valid pairs (if available) instead of recomputing them afresh.
-
-\divya{
-\begin{enumerate}
-\item add all the expressions that our language supports such as mult, bit-wise and and anything else.
-\item Say that we compile to source to intermediate and use inter to circuit of ABY.. We can decide if we want to say this.
-\end{enumerate}
-}

### 2017-10-26 OaklandSubmission/pipe.tex
pass over section 3 and 4.
impl

@@
 From this, the proof of the above theorem follows.
+\nc{should we write more for this simulator?}
+
+The current implementation of secure code partitioning requires a manual step.
+In particular, the developer needs to decompose a program $P$ into 
+the sequence of small programs $P_1||\ldots||P_k$ manually.
+Then \tool generates the sequence of secure programs $Q_1||\ldots||Q_k$ automatically.
+Automating the decomposition step requires an analysis that can 
+statically estimate the resource usage of a \tool program and we leave the construction of such
+an analysis as future work.

### 2017-10-26 OaklandSubmission/eval.tex
formatting in 6.
eval

@@
 and show that the performance of the high-level code written in \tool
-is comparable to their hand-crafted circuits
+is comparable to their hand-crafted protocols
 (Section~\ref{sec:shallow} and Section~\ref{sec:dnneval}).
@@
 %
-Finally, we
+We
 implement a Deep Neural Network (DNN) for CIFAR-10
 dataset~\cite{cifar} from \minion~\cite{minionn} to illustrate the use
-of partitioning in \tool.
+of partitioning in \tool. Finally, we evaluate \tool on the task of matrix factorization~\cite{valeriaMatrix}.
 
@@
 a cross-continent WAN setting. The round trip time between the server
-and the client machines in the two settings is 1ms and 40ms resp. Each
+and the client machines in the two settings is 1ms and 40ms respectively. Each
 machine has an Intel(R) Xeon(R) CPU E5-2673 v3 processor running at
@@
 label $\ell$ for an input data point $x$. For example, given a
-data point which is a tuple of height and weight of an individual,
-a classifier might be able to predict a label as ``male'' or ``female''. The {\it
+data point which is a tuple of humidity and temperature 
+a classifier might be able to predict a label as ``will rain'' or ``will not rain''. The {\it
   accuracy} of a classifier refers to the fraction of data points that
@@
 the input is a data point $x\in\R^d$,
-and the model is a vector $w\in\R^d$. The possible labels are
+and the model is a vector $w\in\R^d$.
+ The possible labels are
 $\ell\in\{\mathit{true},\mathit{false}\}$ and the classifier is
 $C_w\equiv w^T\matmul x>0$, where $w^T$ denotes the transpose
-of matrix $w$, $\matmul$ denotes matrix multiplication (with a vector in this case) and $a>b$ returns $1$ if $a>b$ and $0$ otherwise. This classifier requires $d$ multiplications, $d-1$
+of matrix $w$, $\matmul$ denotes matrix multiplication (with a vector in this case) and $a>b$ returns $1$ if $a>b$ and $0$ otherwise. 
+The {\it model size} of this classifier, i.e., the number of parameters in a model is $d$.
+This classifier requires $d$ multiplications, $d-1$
 additions, and a single comparison.
@@
 \]
+The additions in the expression above sum $n$-dimensional vectors.
+The model size of this classifier is $|P|+|T|$.
 This classifier requires $ndF$ comparisons and additions.
@@
 and is zero otherwise. 
-The operation $\circ$ is a pointwise dot product, $W_j$'s and $V_j$'s
-are matrices of appropriate dimension. The activation function $f$ is given by $f(y) = y$ if
+The operation $\circ$ is a pointwise multiplication of two vectors, $W_j$'s and $V_j$'s
+are matrices of appropriate dimensions. The activation function $f$ is given by $f(y) = y$ if
 $-1 < y < 1$ and $\mathrm{sign}(y)$ otherwise ($f$ on vectors and matrices are defined point-wise on every element of the vector/matrix). \bonsai can be
@@
 benchmarks as they reflect the number of gates required for a
-single 32-bit integer comparison with zero, while the number of
+single  comparison, while the number of
 multiplication gates (\#Mul) is the same as the model size $d$.
@@
 code is competitive with specialized protocols. 
-
+\setlength\tabcolsep{8pt}
 \begin{table*}
@@
 \end{table*}
-
+\setlength\tabcolsep{7pt}
 \begin{table*}
@@
 We evaluate \tool on the DNNs described in SecureML~\cite{secureml},
-Cryptonets~\cite{cryptonets}, and CNN from \minion~\cite{minionn}. For
+Cryptonets~\cite{cryptonets}, and the CNN from \minion~\cite{minionn}. For
 comparison, we consider their implementations from
@@
 The first benchmark  is the DNN described in
-SecureML~\cite{secureml}.
+SecureML~\cite{secureml} (Figure 10 in~\cite{minionn}).
 It has three fully connected layers with square as the activation function.
@@
 except for the $\mathsf{argmax}$ at the end of the network.
-Next, we implement the DNN described in Cryptonets~\cite{cryptonets}
+Next, we implement the DNN described in Cryptonets~\cite{cryptonets} (Figure 11 in~\cite{minionn})
 in \tool.
@@
 convolution (with 5 output channels) and one fully connected layer. 
-Finally, we implement CNN from \minion, that has
+Finally, we implement CNN from \minion (Figure 12 in~\cite{minionn}), that has
 two convolutions (with 16 output channels each) and two fully
@@
 In Table~\ref{tab:nn}, columns 2 and 3 show the time taken by \minion.
-The column ``Model size'' is the number of 32 bit integers in the
+The column ``Model size'' is the number of parameters in the
 trained model.
@@
 %net | minionT | minionC | timeL | timeW | comm | gates | and | mul | length 
-
+\setlength\tabcolsep{6.5pt}
 \begin{table*}
@@
 each $W$ is a $10\times 784$ matrix, and each $b$ is a $10$ length
-vector. \aseem{Questions on this expression: (1) where is $i$?
-  (2) Should $b$ be 10 length vector? (3) Again, should be use $\cdot$
-  for matrix multiplication
-as we did earlier to be consistent?} \nc{I think I have fixed this.
-OK now?
-$\mathsf{argmax}$ throughout this section has implicit use of a dimension and I have left it like that.
-We can make it explicit if others feel the need.}
+vector.
 We implement this classifer in \tool and present the results in the
-first row of Table~\ref{tab:tf}.
-Since, we are not aware of any other tools that has used this
-model as a benchmark, we only report numbers for \tool.
+first row of Table~\ref{tab:tf}. We note that a pure boolean circuit for this functionality cannot fit in the memory of our machines.
 
@@
 This DNN is both bigger and more accurate than the DNNs presented in the previous section.
-In particular, it has an accuracy of 99.2\%\footnote{The accuracy of
-  the integer model is 0.04\% higher than the floating-point
-  model.}. \aseem{Is this footnote necessary? As a reader, I won't
-  miss it if it was not there.}\nc{I agree. Happy to remove.}
+In particular, it has an accuracy of 99.2\%.
+Since, we are not aware of any other tools that has used this
+model as a benchmark, we only report numbers for \tool.
 We observe that this DNN can take a minute per prediction in the WAN
@@
 implementation requires a loop for each layer of the
-tree\footnote{We remark here that our current language does not support functions (which we leave for future work) and with this support, LOC would be even lower.}.\aseem{Should we remark here that once we have functions in the
-  language, LOC would come down, perhaps in the footnote? Else a PL
-  reviewer might scoff at it :).}\nc{added a footnote.}
+tree\footnote{We remark here that our current language does not support functions (which we leave for future work) and with this support, LOC would be even lower.}.
 
@@
 
-\setlength\tabcolsep{4pt}
+\setlength\tabcolsep{3.5pt}
 \begin{table}
 \begin{tabular}{c|c|c|c |c|c|c|c|c|c | c}
-Dataset       & \thead{LAN \\(s)} & \thead{WAN \\(s)} & \thead{Comm. \\ (Mb)}  & \thead{\#\\And} & \thead{\#\\Mul} & \thead{\#\\Gates} & depth & LOC\\
+Dataset       & \thead{LAN \\(s)} & \thead{WAN \\(s)} & \thead{Comm. \\ (Mb)}  & \#And & \#Mul & \#Gates & depth & LOC\\
 \hline
@@
 than what is available
-on our machines \aseem{Is the limitation machine RAM or ABY
-  limit?}\nc{ABY I think. But ABY's limitation is because of RAM I think}. Therefore, we use partitioning and divide the
+on our machines. Therefore, we use partitioning and divide the
 computation into seven stages. 
@@
 fully connected layer. The total number of lines of \tool code for
-this benchamrk is 336 lines. \aseem{Would it be better if we had
-  functions, for loop expressions etc.? If so, we should add a
-  footnote.}
+this benchamrk is 336 lines.
 
@@
 competitive with existing specialized protocols. In particular, for a
-large enough DNN, \minion would run
+large enough DNN, \minion could run
 out of memory but a pipelined \tool implementation would still
-succeed. \aseem{Last sentence is speculative? Should we say \minion
-  would or could or might?}\nc{I think \minion definitely would, but we could say could to be nice.}
-
+succeed. 
 \begin{table}

### 2017-10-26 OaklandSubmission/impl.tex
S5

--- 2017-10-26 OaklandSubmission/impl.tex
@@
 The  \tool compiler is written in Python and
-compiles each of our benchmarks in under a second to C++ code that makes calls to the ABY
-library~\cite{aby}.
-We set the security paramenter to 128 and use OT extension-based arithmetic multiplication triples generation.
-ABY provides facilities to use multiple threads\footnote{The protocol has an offline phase and an online phase.
-Although multiple threads can be used in the offline phase, the online phase is single threaded.} and we use at most four threads in our evaluation.
+compiles each of our benchmarks in under a second to C++ code that
+makes calls to the ABY
+library~\cite{aby}. We use the default ABY configuration for our
+experiment \footnote{We set the security parameter to 128 and use OT
+extension-based arithmetic multiplication triples generation.}. ABY
+provides multi-threading support (for the offline phase of the
+protocol), we leverage the support and use at most four threads in our
+evaluation.
 
-The current language implementation supports the following operators:
-addition, subtraction, multiplication, division by powers of two, left shift,
-logical and arithmetic right shifts, bitwise-(and, or, xor), unary negation,
-bitwise-negation, logical-(not, and, or, xor), and comparisons (less than, greater than, equality).
-Because of their high cost, integral division and floating-point operators are not supported natively by \tool.
-However, we have implemented integral division in 30 lines of \tool and floating-point support in
-ABY is under active development~\cite{ddkssz15}.
+Our current implementation supports the following operators:
+addition, subtraction, multiplication, division by powers of two, left
+shift,
+logical and arithmetic right shifts, bitwise-(and, or, xor), unary
+negation, 
+bitwise-negation, logical-(not, and, or, xor), and comparisons (less
+than, greater than, equality). 
+Because of their high cost, integral division and floating-point
+operators are not supported natively by \tool. 
+However, we have implemented integral division in 30 lines of \tool,
+while the floating-point support in ABY is under active
+development~\cite{ddkssz15}.
 
-Some of our benchamrks require accessing arrays at secret location.
-For example, consider the expression $A[x]$ where $A$ is an array of size 2 and $x$ is shared.
-\tool enforces the array indices to be public. The developer expresses this functionality in \tool
-as $x > 0 ? A[1] : A[0]$. In general, a secret access to an array of size $2^n$ requires $2^{n}-1$ multiplexers in \tool.
+Some of our benchamrks require accessing arrays at secret
+indices. While \tool enforces the array indices to be
+public, secret indices can be encoded in \tool using multiplexers.
+For example, consider the expression $A[x]$ where $A$ is an array of
+size 2 and $x$ is secret-shared. The developer can expresse
+this functionality in \tool as $\cond{x > 0}{A[1]}{A[0]}$. In general,
+a secret access to an array of
+size $2^n$ requires $2^{n}-1$ multiplexers in \tool.
 
  Secure code partitioning currently requires a manual step.
-In particular, the developer needs to decompose a program $P$ into $P_1||\ldots||P_k$ manually (Section~\ref{sec:pipe}). Then \tool generates $Q_1||\ldots||Q_k$ automatically.
+In particular, the developer needs to decompose a program $P$ into
+$P_1||\ldots||P_k$ manually (Section~\ref{sec:pipe}). Then \tool
+generates $Q_1||\ldots||Q_k$ (the sub-programs with appropriate
+wrappers) automatically.
 Automating the decomposition step requires an analysis that can 
-statically estimate the resource usage of a \tool program and we leave the construction of such
-an analysis as future work.
+statically estimate the resource usage of an \tool program. We leave
+the construction of such an analysis as future work.
 
 We use an off-the-shelf solver
-(SeaHorn~\cite{seahorn}) to check the verification conditions required
-to prove that the array indices remain in bound ($\Gamma\vDash e < n$ in {\sc T-Read}, Figure~\ref{fig:compile}). The solver takes less
+(SeaHorn~\cite{seahorn}) to bound check the array indices
+($\Gamma\vDash e < n$
+in ({\sc {T-Read}}) and ({\sc{T-Write}})),
+Figure~\ref{fig:compile}). We take the \tool source program and
+translate it as an input to the solver. The solver takes less
 than a minute on our largest benchmark.
-The type label is assigned conservatively in  {\sc T-Decl}.
-Only the variables that govern the control flow, i.e., variables
-in if-conditions and for-loops are assigned public labels.
-All other variables are assigned arithmetic labels (and can later be coerced to boolean).
-In the future, we would like to explore more intelligent strategies for label assignment.
 
-The compilation rules of Figure~\ref{fig:compile} can introduce redundant coercions from arithmetic to
+Our implementation assigns the type labels (rule~{\sc{T-Decl}})
+conservatively. Only the variables that govern the control flow, i.e.,
+variables in \kw{if}-conditions and \kw{for}-loop counters are
+assigned public labels.
+All other variables are assigned arithmetic labels (that can later be
+coerced to boolean).
+We leave a more sophisticates type inference procedure for future work.
+
+The compilation rules of Figure~\ref{fig:compile} can introduce
+repeated coercions from arithmetic to
 boolean and vice versa.
-The \tool compiler performs several optimizations to reduce these.
-For example, on each coercion, \tool memorizes the pair of arithmetic share and boolean share involved in the coercion.
-\tool invalidates such pairs when the variables corresponding to the shares are overwritten by assignments.
-In subsequent coercions, \tool  reuses valid pairs (if available) instead of recomputing them afresh.
+The \tool implementation performs several optimizations to reduce
+these coercions.
+For example, on each coercion, \tool memorizes the pair of arithmetic
+share and boolean share involved in the coercion. 
+\tool invalidates such pairs when the variables corresponding to the
+shares are overwritten by assignments. 
+In subsequent coercions, \tool  reuses valid pairs (if available)
+instead of recomputing them afresh.
 

### 2017-10-26 OaklandSubmission/impl.tex
nit

@@
 translate it as an input to the solver. The solver takes less
-than a minute on our largest benchmark.
+than a minute on our largest benchmark to verify that all the array
+accesses are in-bounds.
 
@@
 instead of recomputing them afresh.
-
-\divya{
-\begin{enumerate}
-\item add all the expressions that our language supports such as mult, bit-wise and and anything else.
-\item Say that we compile to source to intermediate and use inter to circuit of ABY.. We can decide if we want to say this.
-\end{enumerate}
-}

### 2017-10-26 OaklandSubmission/related.tex

@@
 These levels can be classified as follows:
-\begin{tiret}
-\item {\bf Function Description:} Tools at this level are a language (similar to \tool) where the programmer can specify the functionality to be securely implemented in a high level language and the compiler compiles the functionality to generate a secure computation protocol.
+\\\\
+\noindent{\bf Function Description:} Tools at this level are a language (similar to \tool) where the programmer can specify the functionality to be securely implemented in a high level language and the compiler compiles the functionality to generate a secure computation protocol.
 Such languages do not require the programmer to have expertise in writing cryptographic protocols.
@@
 It is to be noted that none of the above works combine arithmetic and boolean computation together within the same program.
-\item {\bf Protocol Description:} Tools at this level of abstraction provide a set of commands that can be put together to describe efficient secure computation protocols.
+\\\\
+\noindent{\bf Protocol Description:} Tools at this level of abstraction provide a set of commands that can be put together to describe efficient secure computation protocols.
 These tools are at a high level (such as making a call to execute secure computation of an AND gate or a MULT gate over shared inputs).
@@
 Other examples include the VIFF framework~\cite{viff} for arithmetic computations, Sharemind~\cite{sharemind} (that provides information-theoretic secure 3-party boolean computation), and the Secure Multiparty Computation language~\cite{nsy04,securemc}.
-\item {\bf Protocol Implementation:} Tools at this level of abstraction require the programmer to have a deep understanding of how cryptographic secure computation protocols work and the programmer must provide calls to low-level cryptographic building blocks such as oblivious transfer.
+\\\\
+\noindent{\bf Protocol Implementation:} Tools at this level of abstraction require the programmer to have a deep understanding of how cryptographic secure computation protocols work and the programmer must provide calls to low-level cryptographic building blocks such as oblivious transfer.
 While the programmer need not know the specific oblivious transfer protocol implemented in the backend, he or she still needs to know how to build a secure computation protocol from such building blocks.
@@
 Examples of such tools include L1~\cite{l1} and Qilin~\cite{qilin}.
-\end{tiret}
 
@@
 \nc{What do we want to say here exactly?}
+
+\divya{We need to cite this paper \cite{yao-pipe} for yao pipelining. This technique was also used in \cite{valeriaMatrix,oblivm}}

### 2017-10-26 OaklandSubmission/train.tex

--- 2017-10-26 OaklandSubmission/train.tex
@@
 $V$ of dimension $d\times m$ such that $\mathcal{M}\approx UV$. This operator is useful in recommender systems.
-In particular, Nikolaenko et al.~\cite{valeriaMatrix} shows how to implement a movie recommender system which does not require users to reveal their data in the clear, i.e., the ratings the users have assigned to movies.
-The implementation is a two party computation of an iterative algorithm for matrix factorization (Algorithm 1 in~\cite{valeriaMatrix}).
+In particular, Nikolaenko et al.~\cite{valeriaMatrix} shows how to implement a movie recommender system which does not require users to reveal their data in the clear, i.e., the ratings the users have assigned to movies are kept secret.
+The implementation is a two party computation of an iterative algorithm for matrix factorization (Algorithm~1 in~\cite{valeriaMatrix}).
 This algorithm is based on gradient descent and iteratively converges to a local minima.
@@
 manipulates 64-bit variables.
-For $\mathcal{M}$ Nikolaenko et al.
+For the matrix $\mathcal{M}$ of  user data, Nikolaenko et al.
 consider $n=940$ users, $m=40$ most popular movies, and $M=14683$ ratings from the MovieLens dataset.
@@
 setting we are about 4 times faster. The main source of these significant speedups is that, unlike~\cite{valeriaMatrix}, \tool does not need to convert the functionality into boolean circuits. 
-However, this benchmark requires more lines of code than the previous benchmarks.
-It is inconvinient to write Batcher's sort in an iterative language (about 450 lines of \tool).
-The size of a recursive implementation would be much slower.
-In the future, we would like to add support for functions and recursion.
+However, this benchmark requires more lines of code than the previous benchmarks
+because of  Batcher's sort (450 lines of \tool code in each stage).
 However, the current programmer effort seems miniscule compared to the mammoth implementation effort

### 2017-10-26 OaklandSubmission/impl.tex
minor

@@
 extension-based arithmetic multiplication triples generation.}. ABY
-provides multi-threading support (for the offline phase of the
-protocol), we leverage the support and use at most four threads in our
+provides multi-threading support (for the offline phase of the \mpc
+protocol); we leverage the support and use at most four threads in our
 evaluation.
@@
 
-Secure code partitioning currently requires a manual step.
-In particular, the developer needs to decompose a program $P$ into
-$P_1||\ldots||P_k$ manually (Section~\ref{sec:pipe}). Then \tool
-generates $Q_1||\ldots||Q_k$ (the sub-programs with appropriate
-wrappers) automatically.
-Automating the decomposition step requires an analysis that can 
-statically estimate the resource usage of an \tool program. We leave
-the construction of such an analysis as future work.
+%Moved this to Section 4 since P_i and Q_i are ill defined here
+%Secure code partitioning currently requires a manual step.
+%In particular, the developer needs to decompose a program $P$ into
+%$P_1||\ldots||P_k$ manually (Section~\ref{sec:pipe}). Then \tool
+%generates $Q_1||\ldots||Q_k$ (the sub-programs with appropriate
+%wrappers) automatically.
+%Automating the decomposition step requires an analysis that can 
+%statically estimate the resource usage of an \tool program. We leave
+%the construction of such an analysis as future work.
 
@@
 boolean and vice versa.
-The \tool implementation performs several optimizations to reduce
-these coercions.
-For example, on each coercion, \tool memorizes the pair of arithmetic
+The \tool implementation performs  the standard ``common subexpression elimination"
+optimization to reduce these coercions.
+On each coercion, \tool memorizes the pair of arithmetic
 share and boolean share involved in the coercion. 
@@
 In subsequent coercions, \tool  reuses valid pairs (if available)
-instead of recomputing them afresh.
+instead of inserting code to recompute them afresh.

### 2017-10-26 OaklandSubmission/ld.tex
fig 8
circuit compilation

@@
     \ftext{Value} & \val & ::= & \bval \mid [\overline{\bval_{i}}]_{n}\\
-    \ftext{Runtime environment} & \widetilde{\rho} & ::= & \cdot \mid \widetilde{\rho},x \mapsto \val \\
+    %\ftext{Runtime environment} & \widetilde{\rho} & ::= & \cdot \mid \widetilde{\rho},x \mapsto \val \\
     \ftext{Circuit} & \crct & ::= & \cdot \mid \kw{bind}\:\gate\:\wire \mid \kw{out}\:\gate \mid \crct_{1}; \crct_{2}
@@
   \]
-\caption{Circuits   \divya{Is environ misplaced? Can we merge base value and circuit gate?}}
+\caption{Circuits   \divya{Can we merge base value and circuit gate?}}
 \label{fig:circuits}
@@
 
-\divya{Start here....}
-\noindent\textbf{Evaluation to Circuit.} \figureref{eval-inter-circ} shows the judgments for the evaluation of the intermediate language to a circuit.
-We first focus on the expression evaluation judgment
-$\widetilde{\rho} \vdash \widetilde{e} \Downarrow
-\val$. 
-%
-Rules ({\sc{S-PAdd}}) and ({\sc{S-SAdd}}) illustrate
-the significance of the operator labels. In particular, the rule
-({\sc{S-PAdd}}) evaluates a public addition $\widetilde{e_{1}}
-+_{\mathcal{P}} \widetilde{e_{2}}$ to $n_{1} + n_{2}$, similar to the
-source language. In contrast, the rule
-({\sc{S-SAdd}}) evaluates a secret addition $\widetilde{e_{1}}
-+_{\mathcal{A}} \widetilde{e_{2}}$ to an arithmetic add
-gate $\kw{add}\:\kappa_{1}\:\kappa_{2}$. Foreshadowing on the
-compilation rules to be presented later, the target
-expressions from our compiler never have $\widetilde{e_{1}}
-+_{\mathcal{B}} \widetilde{e_{2}}$, as our compiler is aware that $+$
-is more performant using arithmetic circuit, rather than a boolean
-one. Rules ({\sc{S-PCond}}) and ({\sc{S-SCond}})
-are along the similar lines. Rule ({\sc{S-PCond}}) evaluates a public
-conditional to the value from one of the branches, while the rule
-({\sc{S-SCond}}) evaluates to a \kw{mux} gate that takes input
-circuits from the conditional ($\kappa_{1}$) and both the branches
-($\kappa_{2}$ and $\kappa_{3}$). Once again for performance reasons,
-the target expressions from our compiler do not have
-$\lcond{\mathcal{A}}{e_{1}}{e_{2}}{e_{3}}$. Rules
-({\sc{S-Coerce}}) and ({\sc{S-Inp}}) evaluate to the coerce and input
-gates respectively.
-
-Statement evaluation $\widetilde{\rho} \vdash \widetilde{s} \Downarrow
-\widetilde{\rho}_{1}; \chi$ evaluates a target statement to produce a new
-environment $\widetilde{\rho}_{1}$, and a circuit $\chi$. $\chi$ (shown
-in Figure~\ref{fig:tgtlang}) is either empty, \kw{bind}-ing of a
-circuit gate $\kappa$ to range $r$, \kw{out} gate, or a sequence of
-circuits. Rules ({\sc{S-DeclC}}), ({\sc{S-DeclCkt}}), and
-({\sc{S-DeclCktA}}) show the variable declaration cases. Rule
-({\sc{S-DeclC}}) shows the case when $\widetilde{e}$ evaluates to
-$\widetilde{v}$, where $\widetilde{v}$ is either a constant or an
-array of constants. In this case, the mapping $x \mapsto
-\widetilde{v}$ is added to the environment, and the resulting circuit
-is empty. When 
-$\widetilde{e}$ evaluates to a circuit $\kappa$, rule
-({\sc{S-DeclCkt}}) picks a fresh $r$, adds the mapping $x \mapsto r$
-to the environment, and outputs the circuit
-$\kw{bind}\:\kappa\:r$. Rule ({\sc{S-DeclCktA}}) is analogous for
-$\widetilde{e}$ evaluating to an array of circuits. The variable
-assignment rules (not shown in the figure) are similar. Rule
-({\sc{S-WriteCkt}}) shows the case for writing to an array, where the
-array contents are secret. Finally, rule ({\sc{S-Out}}) compiles to an
-\kw{out} circuit.
-
-\begin{figure}
+\begin{figure}[t]
   \small
-  \fbox{$\widetilde{\rho} \vdash \widetilde{e} \Downarrow \widetilde{v}$}
+  \fbox{$\widetilde{\rho} \vdash \widetilde{e} \Downarrow \val$}
   \[
@@
 
-    \inferrule*[lab={\footnotesize{S-PAdd}}]
+    \inferrule*[lab={\footnotesize{S-PMult}}]
                {
@@
                {
-                 \widetilde{\rho} \vdash \widetilde{e_{1}} +_{\mathcal{P}} \widetilde{e_{2}} \Downarrow n_{1} + n_{2}
+                 \widetilde{\rho} \vdash \widetilde{e_{1}} \mulop_{\mathcal{P}} \widetilde{e_{2}} \Downarrow n_{1} \mulop n_{2}
                }
@@
                {
-                 \widetilde{\rho} \vdash x \Downarrow [\overline{\bval_i}]_{n_{1}} \\\\
+                 \widetilde{\rho} \vdash x \Downarrow [\overline{\wire_i}]_{n_{1}} \\\\
                  \widetilde{\rho} \vdash \widetilde{e} \Downarrow n \quad n < n_{1}
@@
                }\\\\
-    \inferrule*[lab={\footnotesize{S-SAdd}}]
-               {
-                 \forall i \in \{1, 2\}.\:\widetilde{\rho} \vdash \widetilde{e_{i}} \Downarrow \kappa_{i}
-               }
-               {
-                 \widetilde{\rho} \vdash \widetilde{e_{1}} +_{\mathcal{A}} \widetilde{e_{2}} \Downarrow \kw{add}\:\kappa_{1}\:\kappa_{2}
+    \inferrule*[lab={\footnotesize{S-SMult}}]
+               {
+                 \forall i \in \{1, 2\}.\:\widetilde{\rho} \vdash \widetilde{e_{i}} \Downarrow \gate_{i}
+               }
+               {
+                 \widetilde{\rho} \vdash \widetilde{e_{1}} \mulop_{\mathcal{A}} \widetilde{e_{2}} \Downarrow \kw{mult}\:\gate_{1}\:\gate_{2}
                }
@@
                {
-                 \forall i \in \{1, 2\}.\:\widetilde{\rho} \vdash \widetilde{e_{i}} \Downarrow \kappa_{i}
-               }
-               {
-                 \widetilde{\rho} \vdash \widetilde{e_{1}} >_{\mathcal{B}} \widetilde{e_{2}} \Downarrow \kw{gt}\:\kappa_{1}\:\kappa_{2}
+                 \forall i \in \{1, 2\}.\:\widetilde{\rho} \vdash \widetilde{e_{i}} \Downarrow \gate_{i}
+               }
+               {
+                 \widetilde{\rho} \vdash \widetilde{e_{1}} >_{\mathcal{B}} \widetilde{e_{2}} \Downarrow \kw{gt}\:\gate_{1}\:\gate_{2}
                }\\\\
@@
                {
-                 \forall i \in \{1, 2, 3\}.\:\widetilde{\rho} \vdash \widetilde{e_{i}} \Downarrow \kappa_{i}
-               }
-               {
-                 \widetilde{\rho} \vdash \lcond{\mathcal{B}}{\widetilde{e_{1}}}{\widetilde{e_{2}}}{\widetilde{e_{3}}} \Downarrow \kw{mux}\:\kappa_{1}\:\kappa_{2}\:\kappa_{3}
+                 \forall i \in \{1, 2, 3\}.\:\widetilde{\rho} \vdash \widetilde{e_{i}} \Downarrow \gate_{i}
+               }
+               {
+                 \widetilde{\rho} \vdash \lcond{\mathcal{B}}{\widetilde{e_{1}}}{\widetilde{e_{2}}}{\widetilde{e_{3}}} \Downarrow \kw{mux}\:\gate_{1}\:\gate_{2}\:\gate_{3}
                }
@@
                {
-                 \widetilde{\rho} \vdash \widetilde{e} \Downarrow \widetilde{v}
-               }
-               {
-                 \widetilde{\rho} \vdash \widetilde{e} \rhd m \Downarrow \widetilde{v} \rhd m
+                 \widetilde{\rho} \vdash \widetilde{e} \Downarrow \val
+               }
+               {
+                 \widetilde{\rho} \vdash \widetilde{e} \rhd m \Downarrow \val \rhd m
                }
@@
                  c = \bot \Rightarrow \widetilde{e'} = \widetilde{e_{2}}\\\\
-                 \widetilde{\rho} \vdash \widetilde{e'} \Downarrow \widetilde{v}
-               }
-               {
-                 \widetilde{\rho} \vdash \lcond{\mathcal{P}}{\widetilde{e}}{\widetilde{e_{1}}}{\widetilde{e_{2}}} \Downarrow \widetilde{v}
+                 \widetilde{\rho} \vdash \widetilde{e'} \Downarrow \val
+               }
+               {
+                 \widetilde{\rho} \vdash \lcond{\mathcal{P}}{\widetilde{e}}{\widetilde{e_{1}}}{\widetilde{e_{2}}} \Downarrow \val
                }
@@
   \\\\
-    \fbox{$\widetilde{\rho} \vdash \widetilde{s} \Downarrow \widetilde{\rho'}; \chi$}
+    \fbox{$\widetilde{\rho} \vdash \widetilde{s} \Downarrow \widetilde{\rho_1}; \crct$}
   \[
@@
                {
-                 \widetilde{\rho} \vdash \widetilde{e} \Downarrow \widetilde{v}\\\\
-                 \widetilde{v} = c \vee \widetilde{v} = [\overline{c_{i}}]_{n}\\\\
-                 \widetilde{\rho}_{1} = \widetilde{\rho}, x \mapsto \widetilde{v}
+                 \widetilde{\rho} \vdash \widetilde{e} \Downarrow \val\\\\
+                 \left(\val = c\right) \vee \left(\val = [\overline{c_{i}}]_{n}\right)\\\\
+                 \widetilde{\rho}_{1} = \widetilde{\rho}, x \mapsto \val
                }
@@
                {
-                 \widetilde{\rho} \vdash \widetilde{e} \Downarrow \kappa\quad
-                 \mathsf{fresh}\:r\\\\
-                 \widetilde{\rho}_{1} = \widetilde{\rho}, x \mapsto r\quad
-                 \chi = \kw{bind}\:\kappa\:r
+                 \widetilde{\rho} \vdash \widetilde{e} \Downarrow \gate\quad
+                 \mathsf{fresh}\:\wire\\\\
+                 \widetilde{\rho}_{1} = \widetilde{\rho}, x \mapsto \wire\quad
+                 \crct = \kw{bind}\:\gate\:\wire
+               }
+               {
+                 \widetilde{\rho} \vdash \tau\:x = \widetilde{e} \Downarrow \widetilde{\rho}_{1}; \crct
+               }
+
+\\\\
+    \inferrule*[lab={\footnotesize{S-DeclCktA}}]
+               {
+                 \widetilde{\rho} \vdash \widetilde{e} \Downarrow [\overline{\gate_{i}}]_{n}\\\\
+                 \forall i \in [n].\:\mathsf{fresh}\:\wire_{i}\\\\
+                 \widetilde{\rho}_{1} = \widetilde{\rho}, x \mapsto [\overline{\wire_{i}}]_{n}\quad
+                 \crct = \overline{\kw{bind}\:\gate_{i}\:\wire_{i}}
                }
@@
 
-\\\\
-    \inferrule*[lab={\footnotesize{S-DeclCktA}}]
-               {
-                 \widetilde{\rho} \vdash \widetilde{e} \Downarrow [\overline{\kappa_{i}}]_{n}\\\\
-                 \forall i \in \{0 \dots n - 1\}.\:\mathsf{fresh}\:r_{i}\\\\
-                 \widetilde{\rho}_{1} = \widetilde{\rho}, x \mapsto [r_{i}]_{n}\quad
-                 \chi = \overline{\kw{bind}\:\kappa_{i}\:r_{i}}
-               }
-               {
-                 \widetilde{\rho} \vdash \tau\:x = \widetilde{e} \Downarrow \widetilde{\rho}_{1}; \chi
-               }
-
                \hspace{0.2cm}
@@
                {
-                 \widetilde{\rho} \vdash \widetilde{e} \Downarrow \kappa
-               }
-               {
-                 \widetilde{\rho} \vdash \kw{out}\:\widetilde{e} \Downarrow \widetilde{\rho}; \kw{out}\:\kappa
+                 \widetilde{\rho} \vdash \widetilde{e} \Downarrow \gate
+               }
+               {
+                 \widetilde{\rho} \vdash \kw{out}\:\widetilde{e} \Downarrow \widetilde{\rho}; \kw{out}\:\gate
                }
@@
                  c = \bot \Rightarrow \widetilde{s} = \widetilde{s_{2}}\\\\
-                 \widetilde{\rho} \vdash \widetilde{s} \Downarrow \widetilde{\rho}_{1}; \chi
-               }
-               {
-                 \widetilde{\rho} \vdash \ite{\widetilde{e}}{\widetilde{s_{1}}}{\widetilde{s_{2}}} \Downarrow \widetilde{\rho}_{1}; \chi
+                 \widetilde{\rho} \vdash \widetilde{s} \Downarrow \widetilde{\rho}_{1}; \crct
+               }
+               {
+                 \widetilde{\rho} \vdash \ite{\widetilde{e}}{\widetilde{s_{1}}}{\widetilde{s_{2}}} \Downarrow \widetilde{\rho}_{1}; \crct
                }
@@
                {
-                 \widetilde{\rho} \vdash x \Downarrow [\overline{r_{i}}]_{n}\quad
+                 \widetilde{\rho} \vdash x \Downarrow [\overline{\wire_{i}}]_{n}\quad
                  \widetilde{\rho} \vdash \widetilde{e_{1}} \Downarrow n_{1}\\\\
                  n_{1} < n\quad
-                 \mathsf{fresh}\:r\quad
-                 \widetilde{\rho} \vdash \widetilde{e_{2}} \Downarrow \kappa\\\\
-                 \widetilde{\rho}_{1} = \widetilde{\rho}[x \mapsto [\overline{r_{i}}]_{n}[n_{1} \mapsto r]]
-               }
-               {
-                 \widetilde{\rho} \vdash x[\widetilde{e_{1}}] := \widetilde{e_{2}} \Downarrow \widetilde{\rho}_{1}; \kw{bind}\:\kappa\:r
+                 \mathsf{fresh}\:\wire\quad
+                 \widetilde{\rho} \vdash \widetilde{e_{2}} \Downarrow \gate\\\\
+                 \widetilde{\rho}_{1} = \widetilde{\rho}[x \mapsto ([\overline{\wire_{i}}]_{n}[n_{1} \mapsto \wire])]
+               }
+               {
+                 \widetilde{\rho} \vdash x[\widetilde{e_{1}}] := \widetilde{e_{2}} \Downarrow \widetilde{\rho}_{1}; \kw{bind}\:\gate\:\wire
                }
@@
 
-\begin{figure}
+\subsubsection*{Evaluation to Circuit} \figureref{eval-inter-circ} shows the judgments for the evaluation of the intermediate language to a circuit. 
+The runtime environment is defined as follows:
+
+\vspace{0.2cm}
+$
 \small
-  \fbox{$\widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \kappa \Downarrow b_{1}, b_{2}$}
-  \[
-  \\
-  \begin{array}{c}
-    %% \inferrule*[lab={\footnotesize{C-Wire}}]
-    %%            {
-    %%              b_{1}, b_{2} = \widehat{\rho_{1}}(r), \widehat{\rho_{2}}(r)
-    %%            }
-    %%            {
-    %%              \widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash r \Downarrow b_{1}, b_{2}
-    %%            }               
-    \inferrule*[lab={\footnotesize{C-In}}]
-               {
-                 b_{1}, b_{2} = \mathcal{E}_{m}(c)
-               }
-               {
-                 \widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \kw{in}^{m}_{j} \Downarrow b_{1}, b_{2}
-               }
-
-               \hspace{0.3cm}
-
-    \inferrule*[lab={\footnotesize{C-Coerce}}]
-               {
-                 \widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \kappa \Downarrow b_{1}, b_{2}\\\\
-                 c = \mathcal{D}_{m_{1}}(b_{1}, b_{2})\\\\
-                 b'_{1}, b'_{2} = \mathcal{E}_{m}(c)
-               }
-               {
-                 \widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \kappa \rhd m \Downarrow b'_{1}, b'_{2}
-               }
-\\\\
-
-    \inferrule*[lab={\footnotesize{C-Add}}]
-               {
-                 \forall i \in \{1, 2\}.\:\widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \kappa_{i} \Downarrow b_{1i}, b_{2i}\quad
-                 n_{i} = \mathcal{D}_{\mathcal{A}}(b_{1i}, b_{2i})\\\\
-                 b_{1}, b_{2} = \mathcal{E}_{\mathcal{A}}(n_{1} + n_{2})
-               }
-               {
-                 \widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \kw{add}\:\kappa_{1}\:\kappa_{2} \Downarrow b_{1}, b_{2}
-               }
-
-\\\\
-
-    \inferrule*[lab={\footnotesize{C-Gt}}]
-               {
-                 \forall i \in \{1, 2\}.\:\widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \kappa_{i} \Downarrow b_{1i}, b_{2i}\quad
-                 n_{i} = \mathcal{D}_{\mathcal{B}}(b_{1i}, b_{2i})\\\\
-                 b_{1}, b_{2} = \mathcal{E}_{\mathcal{B}}(n_{1} > n_{2})
-               }
-               {
-                 \widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \kw{gt}\:\kappa_{1}\:\kappa_{2} \Downarrow b_{1}, b_{2}
-               }
-
-\\\\
-
-    \inferrule*[lab={\footnotesize{C-Mux}}]
-               {
-                 \forall i \in \{1, 2, 3\}.\:\widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \kappa_{i} \Downarrow b_{1i}, b_{2i}\quad
-                 c_{i} = \mathcal{D}_{\mathcal{B}}(b_{1i}, b_{2i})\\\\
-                 c_{1} = \top \Rightarrow b_{1}, b_{2} = \mathcal{E}_{\mathcal{B}}(c_{2})\quad
-                 c_{1} = \bot \Rightarrow b_{1}, b_{2} = \mathcal{E}_{\mathcal{B}}(c_{3})
-               }
-               {
-                 \widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \kw{mux}\:\kappa_{1}\:\kappa_{2}\:\kappa_{3} \Downarrow b_{1}, b_{2}
-               }
-
-  \end{array}
-  \]
-  \\\\
-    \fbox{$\widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \chi \Downarrow \widehat{\rho'_{1}}, \widehat{\rho'_{2}}; O$}
-  \[
-  \\
-  \begin{array}{c}
-    \inferrule*[lab={\footnotesize{C-Bind}}]
-               {
-                 \widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \kappa \Downarrow b_{1}, b_{2}\\\\
-                 \widehat{\rho'_{1}} = \widehat{\rho_{1}}[r \mapsto b_{1}] \quad
-                 \widehat{\rho'_{2}} = \widehat{\rho_{2}}[r \mapsto b_{2}]
-               }
-               {
-                 \widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \kw{bind}\:\kappa\:r \Downarrow \widehat{\rho'_{1}}, \widehat{\rho'_{2}}; \cdot
-               }
-
-               \hspace{0.1cm}
-
-    \inferrule*[lab={\footnotesize{C-Out}}]
-               {
-                 \widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \kappa \Downarrow b_{1}, b_{2}\\\\
-                 c = \mathcal{D}_{m}(b_{1}, b_{2})
-               }
-               {
-                 \widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \kw{out}\:\kappa \Downarrow \widehat{\rho_{1}}, \widehat{\rho_{2}}; c
-               }
-\end{array}
-  \]
-\caption{Circuit semantics (selected rules)}
-\label{fig:cktsem}
-\end{figure}
-
+\begin{array}{rrcl}
+    \ftext{Runtime environment} & \widetilde\rho &::=& \cdot \mid \rho,x \mapsto \val\\
+\end{array}
+$
+\vspace{0.2cm}
+
+We first focus on the expression evaluation judgment
+$\widetilde{\rho} \vdash \widetilde{e} \Downarrow\val$. 
+%
+Rules ({\sc{S-PMult}}) and ({\sc{S-SMult}}) illustrate
+the significance of the operator labels. In particular, the rule
+({\sc{S-PMult}}) evaluates a public multiplication $\widetilde{e_{1}}
+\mulop_{\mathcal{P}} \widetilde{e_{2}}$ to $n_{1} \mulop n_{2}$, similar to the
+source language. In contrast, the rule
+({\sc{S-SMult}}) evaluates a secret multiplication $\widetilde{e_{1}}
+\mulop_{\mathcal{A}} \widetilde{e_{2}}$ to an arithmetic multiplication
+gate $\kw{mult}\:\gate_{1}\:\gate_{2}$. As mentioned above, the intermediate language
+expressions generated by our compiler never have $\widetilde{e_{1}}
+\mulop_{\mathcal{B}} \widetilde{e_{2}}$, as our compiler is aware that $\mulop$
+is more performant using an arithmetic circuit compared to a boolean
+one. Rules ({\sc{S-PCond}}) and ({\sc{S-SCond}})
+are along the similar lines. Rule ({\sc{S-PCond}}) evaluates a public
+conditional to the value from one of the branches, while the rule
+({\sc{S-SCond}}) evaluates to a \kw{mux} gate that takes input
+circuits from the guard ($\gate_{1}$) and both the branches
+($\gate_{2}$ and $\gate_{3}$). Once again for performance reasons,
+the target expressions from our compiler do not have
+$\lcond{\mathcal{A}}{e_{1}}{e_{2}}{e_{3}}$. Rules
+({\sc{S-Coerce}}) and ({\sc{S-Inp}}) evaluate to the coerce and input
+gates respectively.
+
+Statement evaluation $\widetilde{\rho} \vdash \widetilde{s} \Downarrow
+\widetilde{\rho}_{1}; \chi$ evaluates statements $\widetilde{s}$ to produce a new
+environment $\widetilde{\rho}_{1}$, and a circuit $\crct$. A circuit $\crct$ (shown
+in \figureref{circuits}) is either empty, \kw{bind}-ing of a
+circuit gate $\gate$ to wire $\wire$, \kw{out} gate, or a sequence of
+circuits. Rules ({\sc{S-DeclC}}), ({\sc{S-DeclCkt}}), and
+({\sc{S-DeclCktA}}) show the variable declaration cases. Rule
+({\sc{S-DeclC}}) shows the case when $\widetilde{e}$ evaluates to
+$\val$, where $\val$ is either a constant or an
+array of constants. In this case, the mapping $x \mapsto
+\val$ is added to the environment, and the resulting circuit
+is empty. When 
+$\widetilde{e}$ evaluates to a circuit $\gate$, rule
+({\sc{S-DeclCkt}}) picks a fresh wire $\wire$, adds the mapping $x \mapsto \wire$
+to the environment $\widetilde{\rho}$, and outputs the circuit
+$\kw{bind}\:\gate\:\wire$. Rule ({\sc{S-DeclCktA}}) is analogous for
+$\widetilde{e}$ evaluating to an array of circuits. The variable
+assignment rules (not shown in the figure) are similar. Rule
+({\sc{S-WriteCkt}}) shows the case for writing to an array, where the
+array contents are secret. Finally, rule ({\sc{S-Out}}) compiles to an
+\kw{out} circuit.
+
+\divya{Start here...}
 \subsubsection*{Circuit semantics} The target semantics produces a
@@
 observations.
+
+\begin{figure}
+  \small
+  \fbox{$\widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \kappa \Downarrow b_{1}, b_{2}$}
+  \[
+  \\
+  \begin{array}{c}
+    %% \inferrule*[lab={\footnotesize{C-Wire}}]
+    %%            {
+    %%              b_{1}, b_{2} = \widehat{\rho_{1}}(r), \widehat{\rho_{2}}(r)
+    %%            }
+    %%            {
+    %%              \widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash r \Downarrow b_{1}, b_{2}
+    %%            }               
+    \inferrule*[lab={\footnotesize{C-In}}]
+               {
+                 b_{1}, b_{2} = \mathcal{E}_{m}(c)
+               }
+               {
+                 \widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \kw{in}^{m}_{j} \Downarrow b_{1}, b_{2}
+               }
+
+               \hspace{0.3cm}
+
+    \inferrule*[lab={\footnotesize{C-Coerce}}]
+               {
+                 \widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \kappa \Downarrow b_{1}, b_{2}\\\\
+                 c = \mathcal{D}_{m_{1}}(b_{1}, b_{2})\\\\
+                 b'_{1}, b'_{2} = \mathcal{E}_{m}(c)
+               }
+               {
+                 \widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \kappa \rhd m \Downarrow b'_{1}, b'_{2}
+               }
+\\\\
+
+    \inferrule*[lab={\footnotesize{C-Add}}]
+               {
+                 \forall i \in \{1, 2\}.\:\widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \kappa_{i} \Downarrow b_{1i}, b_{2i}\quad
+                 n_{i} = \mathcal{D}_{\mathcal{A}}(b_{1i}, b_{2i})\\\\
+                 b_{1}, b_{2} = \mathcal{E}_{\mathcal{A}}(n_{1} + n_{2})
+               }
+               {
+                 \widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \kw{add}\:\kappa_{1}\:\kappa_{2} \Downarrow b_{1}, b_{2}
+               }
+
+\\\\
+
+    \inferrule*[lab={\footnotesize{C-Gt}}]
+               {
+                 \forall i \in \{1, 2\}.\:\widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \kappa_{i} \Downarrow b_{1i}, b_{2i}\quad
+                 n_{i} = \mathcal{D}_{\mathcal{B}}(b_{1i}, b_{2i})\\\\
+                 b_{1}, b_{2} = \mathcal{E}_{\mathcal{B}}(n_{1} > n_{2})
+               }
+               {
+                 \widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \kw{gt}\:\kappa_{1}\:\kappa_{2} \Downarrow b_{1}, b_{2}
+               }
+
+\\\\
+
+    \inferrule*[lab={\footnotesize{C-Mux}}]
+               {
+                 \forall i \in \{1, 2, 3\}.\:\widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \kappa_{i} \Downarrow b_{1i}, b_{2i}\quad
+                 c_{i} = \mathcal{D}_{\mathcal{B}}(b_{1i}, b_{2i})\\\\
+                 c_{1} = \top \Rightarrow b_{1}, b_{2} = \mathcal{E}_{\mathcal{B}}(c_{2})\quad
+                 c_{1} = \bot \Rightarrow b_{1}, b_{2} = \mathcal{E}_{\mathcal{B}}(c_{3})
+               }
+               {
+                 \widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \kw{mux}\:\kappa_{1}\:\kappa_{2}\:\kappa_{3} \Downarrow b_{1}, b_{2}
+               }
+
+  \end{array}
+  \]
+  \\\\
+    \fbox{$\widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \chi \Downarrow \widehat{\rho'_{1}}, \widehat{\rho'_{2}}; O$}
+  \[
+  \\
+  \begin{array}{c}
+    \inferrule*[lab={\footnotesize{C-Bind}}]
+               {
+                 \widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \kappa \Downarrow b_{1}, b_{2}\\\\
+                 \widehat{\rho'_{1}} = \widehat{\rho_{1}}[r \mapsto b_{1}] \quad
+                 \widehat{\rho'_{2}} = \widehat{\rho_{2}}[r \mapsto b_{2}]
+               }
+               {
+                 \widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \kw{bind}\:\kappa\:r \Downarrow \widehat{\rho'_{1}}, \widehat{\rho'_{2}}; \cdot
+               }
+
+               \hspace{0.1cm}
+
+    \inferrule*[lab={\footnotesize{C-Out}}]
+               {
+                 \widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \kappa \Downarrow b_{1}, b_{2}\\\\
+                 c = \mathcal{D}_{m}(b_{1}, b_{2})
+               }
+               {
+                 \widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \kw{out}\:\kappa \Downarrow \widehat{\rho_{1}}, \widehat{\rho_{2}}; c
+               }
+\end{array}
+  \]
+\caption{Circuit semantics (selected rules)}
+\label{fig:cktsem}
+\end{figure}
 

### 2017-10-26 OaklandSubmission/ld.tex
small change

@@
 \label{sec:ld}
-\TODO{Later, in implementation we can write that we use ABY for compiling intermediate to target also.}
-
-\TODO{Use mult op instead of add?
-arguments of using arithmetic are clearer there; though we also use arithmetic always for add also.
-Or write both in exp and write rule for mul; how about bit-wise And?
-shifts?
-Say that more exp written in implementation}
 
@@
 \divya{Start here...}
-\subsubsection*{Circuit semantics} The target semantics produces a
-circuit to compute over the secret data using an MPC protocol. \nc{We use 2PC and MPC interchangeably. I guess its fine?}With
-our circuit semantics, we model the \emph{functional} aspect of the
-protocol, parametrized by the sharing functions.
+\subsubsection*{Circuit semantics} Evaluating a program in the intermediate language produces a
+circuit over the secret data to be computed using an \mpc protocol in the distributed setting. With
+our circuit semantics, we model the \emph{functional} aspect of a \mpc protocol, parametrized by the sharing functions.
 

### 2017-10-26 OaklandSubmission/eval.tex

@@
 $C_w\equiv w^T\matmul x>0$, where $w^T$ denotes the transpose
-of matrix $w$, $\matmul$ denotes matrix multiplication (with a vector in this case) and $a>b$ returns $1$ if $a>b$ and $0$ otherwise. 
-The {\it model size} of this classifier, i.e., the number of parameters in a model is $d$.
+of matrix $w$, $\matmul$ denotes matrix multiplication (with a vector in this case) and $a>b$ returns $\mathit{true}$ if $a>b$ and $\mathit{false}$ otherwise. 
+The {\it model size} of this classifier, i.e., the number of parameters in the model is $|w|$, i.e., $d$.
 This classifier requires $d$ multiplications, $d-1$
@@
 The model size of this classifier is $|P|+|T|$.
-This classifier requires $ndF$ comparisons and additions.
-Note here  that computing $T(i)(j)(x_j)$ requires a secret look-up in matrix  $T$  based on secret input $x_j$ resulting in $F-1$ comparisons using multiplexers (see
+This classifier requires $\Theta(ndF)$ comparisons and additions.
+Note here  that computing $T(j)(x_j)(i)$ requires a secret look-up in matrix  $T$  based on secret input $x_j$ resulting in $F-1$ comparisons using multiplexers (see
   Section~\ref{sec:impl}).
@@
 are matrices of appropriate dimensions. The activation function $f$ is given by $f(y) = y$ if
-$-1 < y < 1$ and $\mathrm{sign}(y)$ otherwise ($f$ on vectors and matrices are defined point-wise on every element of the vector/matrix). \bonsai can be
+$-1 < y < 1$ and $\mathrm{sign}(y)$ otherwise ($f$ on vectors and matrices is defined point-wise on every element of the vector/matrix). \bonsai can be
 seen as a variant of decision trees where the prediction is a function
@@
 We implement this classifer in \tool and present the results in the
-first row of Table~\ref{tab:tf}. We note that a pure boolean circuit for this functionality cannot fit in the memory of our machines.
+first row of Table~\ref{tab:tf}. We note that a pure boolean circuit for this simple functionality does not fit in the memory of our machines.
 
@@
 large enough DNN, \minion could run
-out of memory but a pipelined \tool implementation would still
+out of memory but an appropriately partitioned \tool implementation would still
 succeed. 

### 2017-10-27 OaklandSubmission/intro.tex
some edits in the intro

--- 2017-10-26 OaklandSubmission/intro.tex
@@
-  \TODO{While taking a pass on intro, check citation correctness or if we are missing any citation.}
+  %% \TODO{While taking a pass on intro, check citation correctness or if we are missing any citation.}
 
@@
 
-Today it is hard for the developers to program secure applications
-using cryptographic techniques. Typical developers lack a deep
-understanding of the cryptographic protocols, and therefore cannot be
-expected to use them correctly and efficiently on their own.
-Ideally, a developer would declare the functionality in a general
-purpose, high-level programming language and a
-tool, e.g. a compiler, would generate an efficient protocol that
-implements the functionality securely, while hiding the cryptography
-behind-the-scenes. This paper presents such a framework for Secure
-Two-party Computation (\mpc).
+%% \aseem{The intuition of this paragraph is conveyed in rest of the
+%% intro already?}
+
+%% Today it is hard for the developers to program secure applications
+%% using cryptographic techniques. Typical developers lack a deep
+%% understanding of the cryptographic protocols, and therefore cannot be
+%% expected to use them correctly and efficiently on their own.
+%% Ideally, a developer would declare the functionality in a general
+%% purpose, high-level programming language and a
+%% tool, e.g. a compiler, would generate an efficient protocol that
+%% implements the functionality securely, while hiding the cryptography
+%% behind-the-scenes. This paper presents such a framework for Secure
+%% Two-party Computation (\mpc).
 
@@
 
-\mpc~(\cite{yao,gmw}) is a powerful
+Secure Two-party Computation (\mpc)~(\cite{yao,gmw}) is a powerful
 cryptographic technique that allows two mutually distrusting parties
@@
 something practically efficient and usable. To understand the
-state-of-the-art, let us consider the following example used by many
+state-of-the-art, let us consider the following example underlying many
 secure prediction algorithms. Suppose Alice wants to write
 a \mpc protocol to securely
-compute $w^Tx >b$. Here $w$ and $x$ are two vectors (inputs of the two
-parties), $b$ is a scalar, $\cdot^{T}$ is the matrix transpose
-operator, and $w^Tx$ denotes the inner product of
-$w^{T}$ and $x$.  With the current state-of-the-art, Alice has following
-options:
+compute $w^Tx >b$. Here $w$ (a vector) and $b$ (a scalar) constitute
+the server model, and $x$ is the client's input vector. Further,
+$\cdot^{T}$ is the matrix transpose operator, and $w^Tx$ denotes the
+inner product of  $w^{T}$ and $x$. With the current state-of-the-art,
+Alice has following options:
 
@@
 The efficiency of the generated \mpc protocol is thus bounded by the
-efficiency of representing the computation in one of
+efficiency of representing the computation in \emph{one} of
 these representations. For instance, multiplication of two
@@
 ``cryptographic-cost aware'' framework that generates efficient and
-scalable \mpc protocols, comprising combinations of arithmetic and
-boolean circuits, from high-level programs devoid of any
-cryptography. Below we describe the salient features of \tool.
+scalable \mpc protocols from high-level programs devoid of any
+cryptographic details. The generated protocols comprise combinations
+of arithmetic and boolean circuits and have performance comparable to
+or better than the custom, specialized
+protocols from previous work. Below we describe the
+salient features of \tool.
 
-\subsubsection*{Programmability} \tool source programs are ideal
+\subsubsection*{Ease of programming} \tool source programs are ideal
 functionalities that describe ``what'' computation needs to be done,
@@
 \vspace{0.2cm}
-We note that the program is quite similar to what a programmer might
-write in a common programming language. The simplicity of the
+The program is quite similar to what a programmer might
+write in C++ or Java. The simplicity of the
 language comes with its usual benefits: it is easily accessible to the
-developers, there are fewer avenues for making mistakes, and it is
-easy to maintain and modify the programs. Needless to say, frameworks
-that expose low-level circuit APIs to the programmer do not enjoy
-these benefits.
+developers, there are fewer avenues for making mistakes, developers
+don't bear the burden of getting cryptographic details right, and it
+is easy to maintain and modify the programs. Needless to say,
+frameworks that expose low-level circuit APIs to the programmer do not
+enjoy these benefits.
 
-\subsubsection*{Cryptogaphic-cost awareness} \mpc compiler is aware
+\subsubsection*{Cryptogaphic-cost aware compiler} The \tool compiler
+compiles a source program to a hybrid computation consisting
+of \emph{public} and \emph{secret} parts. In the example above, for
+instance, \tool compiler realizes that the array index~\ls{i} is
+public, and hence, the array accesses need not be compiled as
+multiplexers. Further, within the secret parts, \tool compiler is
+aware
 of the cryptographic costs of arithmetic and boolean representations
-of the source language operators, and their corresponding
-inter-conversion gates. Based on these costs, the compiler, given a
-source program, automatically picks arithmetic or boolean
-representations for different parts of the program, and generates the
+of the source language operators. Based on these costs, the compiler
+automatically picks arithmetic or boolean
+representations for different sub-parts, and generates the
 corresponding circuits along with the required inter-conversion
-gates. The outcome is an effcient \mpc protocol using a combination of
-arithmetic and boolean circuits, while the programmer remains
-oblivious of all the low-level cryptographic details. Our evaluation
-shows that the performance of the
-protocols automatically generated by the \tool compiler is comparable
+gates. The outcome is an effcient \mpc protocol combining arithmetic
+and boolean circuits, while the programmer remains
+oblivious of all these cryptographic details. Our evaluation
+shows that the performance of these automatically generated protocols
+is comparable
 to or better than the custom, specialized protocols developed in the
@@
 securely is upper-bounded by the largest circuit that can fit in the
-memory on a machine\footnote{Using swap and
+machine memory\footnote{Using swap and
 disk space is feasible but it causes huge slowdown.}. This is a
@@
 a high level, we break the original program into a sequence of small
-sub-programs, which are then sequentially processed by our
-framework, while appropriately threading the intermediate outputs
+sub-programs, which are then sequentially processed by \tool, while
+appropriately threading the intermediate outputs
 along. While this
@@
 
-\subsubsection*{Formal guarantees} We prove a formal correctness and
-security theorem for our compiler. The correctness theorem relates the
+\subsubsection*{Formal guarantees} We prove formal correctness and
+security theorems for our compiler. The correctness theorem proves a
+forward simluation between the
 ``ideal'' semantics (the trusted-third party semantics) of a source
@@
 semantics that relies on circuit evaluation) of the corresponding
-compiled program. Both these semantics are instrumented
-with \emph{observations} that record the outputs. We prove that a
-well-typed source program terminates successfully in the source
-semantics. In addition, we prove that for a well-typed source program,
-its observations in the source semantics are same as the observations
-of its corresponding compiled program in the protocol
-semantics. {\aseem{Had a hard time writing this last sentence,
-suggestions welcome.}}
-For the security theorem, we formally reduce the security of our
-compiler to the security of the back-end cryptographic
-protocol.
+compiled program. The simulation
+theorem not only guarantees that a
+well-typed source program terminates successfully in the ideal
+semantics (e.g. there are no array index out-of-bounds errors), but
+also that the observable outputs in the two semantics are the same. For
+the security theorem, we formally reduce the security
+of our compiler to the security of the back-end cryptographic
+protocol. We also prove a formal security theorem for our partitioning
+scheme.
 
@@
 
-{\aseem{Not sure about this baseline paragraph. Did not touch it.}} 
-\subsubsection*{Performance Baseline} Most works cite garbled
-circuit based two-party computation as the most practical approach so
-far and use it as the baseline for comparison of performance. But, it
-has been shown multiple times that the garbled circuits based approach
-does not scale for realistic examples. Even for small benchmarks
-(e.g., Na\"{i}ve Bayes) the size of garbled circuit can be
-huge. Hence, comparisons are often done on artificial data sizes and
-garbled circuits are shown to be 500x slower than hand-crafted
-protocols \cite{shafindss}. This is cited as the motivation for need
-for specialized protocols for different ML algorithms in secure
-prediction such as linear regression \cite{shafindss}, decision
-trees \cite{wu}, deep neural networks \cite{minionn} and matrix
-factorization \cite{valeriaMatrix}. Our work is based on generic \mpc
-and we show competitive performance with the specialized protocols
-contrary to the popular belief. Hence, we believe \tool based
-implementations can serve as the performance baseline for future works
-on specialized protocols.
+%% {\aseem{Not sure about this baseline paragraph. Did not touch it.}} 
+%% \subsubsection*{Performance Baseline} Most works cite garbled
+%% circuit based two-party computation as the most practical approach so
+%% far and use it as the baseline for comparison of performance. But, it
+%% has been shown multiple times that the garbled circuits based approach
+%% does not scale for realistic examples. Even for small benchmarks
+%% (e.g., Na\"{i}ve Bayes) the size of garbled circuit can be
+%% huge. Hence, comparisons are often done on artificial data sizes and
+%% garbled circuits are shown to be 500x slower than hand-crafted
+%% protocols \cite{shafindss}. This is cited as the motivation for need
+%% for specialized protocols for different ML algorithms in secure
+%% prediction such as linear regression \cite{shafindss}, decision
+%% trees \cite{wu}, deep neural networks \cite{minionn} and matrix
+%% factorization \cite{valeriaMatrix}. Our work is based on generic \mpc
+%% and we show competitive performance with the specialized protocols
+%% contrary to the popular belief. Hence, we believe \tool based
+%% implementations can serve as the performance baseline for future works
+%% on specialized protocols.
 
@@
 
-\aseem{Checkpointing here. Two more paragraphs to come, one for
-summary of implementation and evaluation, and one for related work.}
+\aseem{Checkpointing here. Three more paragraphs to come, one for
+summary of implementation and evaluation, one for related work, and
+one for baseline argument.}
 

### 2017-10-27 OaklandSubmission/intro.tex
some nits in the partitioning intro para

@@
 we still have to address
-the security risk of revealing the intermediate outputs. To address
-this, \tool (a) masks the intermediate outputs using a secure secret
-sharing between the two parties, and (b) appends the
-share-reconstruction function to the beginning of the subsequent
-sub-program. As we show in our evaluation, partitioning allows us to
-program large applications in \tool.
+the security risk of revealing the intermediate outputs. \tool comes
+to the rescue -- it automatically inserts the required instrumentation
+to ensure security in the presence of these intermediate outputs. As
+we show in our
+evaluation, partitioning allows us to program large applications
+in \tool.
 

### 2017-10-27 OaklandSubmission/impl.tex
partitioning

@@
 (SeaHorn~\cite{seahorn}) to bound check the array indices
-($\Gamma\vDash e < n$
+($P\vDash e < n$
 in ({\sc {T-Read}}) and ({\sc{T-Write}})),
@@
 The \tool implementation performs  the standard ``common subexpression elimination"
-optimization to reduce these coercions.
+optimization~\cite{dragonbook} to reduce these coercions.
 On each coercion, \tool memorizes the pair of arithmetic

### 2017-10-27 OaklandSubmission/pipe.tex
partitioning

--- 2017-10-15 OaklandSubmission/pipe.tex
@@
-\section{Pipelining}
+\section{Secure code partitioning}
 \label{sec:pipe}
-\nc{Started high level description. Can go deeper and correct notation as we proceed.}
-We now describe our technique of ``pipelining'' that allows us to compile and execute large programs that would otherwise have not fit in memory (due to the large size of the circuit generated by ABY).
-We do this securely in a manner such that the underlying backend (such as ABY in this case) does not have to modified in anyway and hence our technique is also compatible with any backend.
-In more detail, let $P$ be a program in our source language that gets compiled to $P'$ in our target language.
-In many cases, the ABY circuit generated for $P'$ is very large and cannot fit in memory (in fact, there is an upper limit of $2^{32}$ for the circuit size in ABY and for many natural programs, this upper limit is easily crossed).
-Without pipelining, we would be unable to compile and execute such a program $P$.
-In order to overcome this size limit, we do the following: we split the program $P$ into a sequence of smaller programs $Q_1||Q_2||\cdots||Q_k$ (that will be appropriately defined).
-We compile and execute each $Q_i$ separately, feeding the outputs of one program as state information to the next program.
-Of course, we have to be careful how we do this so as to not compromise the security of the protocol.
+
+We  describe ``secure code partitioning'' that allows us to execute {\it large} programs, i.e., programs with a large number of operations.
+In more detail, let $P$ be a program in our source language that generates a circuit $\crct$.
+For a large enough $P$, the circuit $\crct$  can be larger than the memory size\footnote{In fact, there is an upper limit of $2^{32}-1$ gates for the circuit size in ABY but for most machines the memory limit is hit first} and fail to execute.
+Partitioning enables us to
+execute such programs via a source to source transformation that is oblivious to the underlying \mpc backend.
+Partitioning decomposes the program $P$ into a sequence of smaller \tool programs $Q_1||Q_2||\cdots||Q_k$ (defined below).
+We compile and execute each $Q_i$ sequentially, feeding the outputs of $Q_i$ as state information to $Q_{i+1}$.
+We prove that partioning is correct ($P$ and $Q_1||Q_2||\cdots||Q_k$ compute the same functionality) and secure (execution of $Q_1||Q_2||\cdots||Q_k$ does not reveal any more information than $P$).
 More details follow.
 
-Let $P$ be a program that takes as (secret) inputs $x$ and $y$ and produces output $z$ to both parties.
-Let $P_1||P_2||\cdots||P_k$ be a sequence of programs defined such that the following holds.
-Define $s_0 = \bot$ (the public empty state).
-For all $1\leq i\leq k-1$, $P_i$ takes as input $x, y$ and $s_{i-1}$ to produce state information $s_i$.
-Finally, $P_k$ takes as input $x,y$ and $s_{k-1}$ to produce output $z$.
-Every program $P_i$ ($1\leq i\leq k$) will be required to be ``sufficiently'' small (as described later) and this will enable our pipelining technique.
-It is easy to see that any program $P$ can be decomposed as illustrated above into such $P_1||P_2||\cdots||P_k$.
-Now, if one were to compile $P_i$ into $P'_i$, and execute $P'_i$, the parties can execute $P'_1||P'_2||\cdots||P'_k$ sequentially to obtain $s_1,\cdots,s_{k-1},$ and finally output $z$.
-At the $i^{\textrm{\tiny{th}}}$ step, the parties would only need to store information proportional to $x,y,s_i$ and $P'_i$ (which will all be much smaller than $P'$).
-However, if we were to execute the programs as outlined above, it would enable the parties to learn $s_i$ (for all $1\leq i\leq k-1$), which would completely break the security of the protocol.
+Let $P$ be a program that takes (secret) inputs $x$ and $y$ and produces an output $z$ to both parties (Alice and Bob).
+Let $P_1||P_2||\cdots||P_k$ be a sequence of programs such that the following holds: Define $s_0 = \bot$ (the public empty state).
+For all $1\leq i\leq k-1$, $P_i$ takes inputs $x, y$ and $s_{i-1}$ and outputs  state information $s_i$.
+Finally, $P_k$ takes inputs $x,y$ and $s_{k-1}$ to  output $z$.
+Each program $P_i$ ($1\leq i\leq k$) is required to be ``sufficiently'' small (as described later).
+It is possible to decompose any program $P$  into such $P_1||P_2||\cdots||P_k$.
+If \tool generates circuit $\crct_i$ from $P_i$, the parties can execute $\crct_1||\crct_2||\cdots||\crct_k$ sequentially (in a distributed setting)  to obtain $s_1,\cdots,s_{k-1},$ and finally output $z$.
+At the $i^{\textrm{\tiny{th}}}$ step, the parties only need to store information proportional to $x,y,s_{i-1}$ and $\crct_i$ (which is much smaller than $\crct$).
+However, this execution enables the parties to learn $s_i$ (for all $1\leq i\leq k-1$), which completely breaks the security.
 
@@
 Once again, define $s_0 = \bot$.
-Let all $s_i$ be values in some additive ring $(\mathbb{Z},+)$ (in the ABY implementation this ring would simply be the additive ring $(\mathbb{Z}_{2^{32}},+)$ i.e., the additive ring of integers modulo $2^{32}$).
-This is without loss of generality true (we will also for simplicity assume that each $s_i$ is a single element in the ring, but this restriction can also be easily removed).
-Let $r_1,\cdots,r_{k-1}$ be a sequence of random values sampled from the same ring $(\mathbb{Z},+)$ by Party A (in our implementation, all $r_i$ values can be generated using a pseudorandom function $F$).
-Let $Q_1$ be the program that takes as input $x,y, r_1$ (and empty state $s_0$), runs $P_1$ (as defined above) to get $s_1$ and then produces output $o_1 = s_1 + r_1$ {\em only to} Party B\footnote{While the description of our protocol here assumes that the underlying backend supports only one party receiving output, this is only a simplifying assumption, and we can easily modify our protocol in the case where both parties must receive the same output.
-To do so, we let $Q_1$ output $o_1 = s_1+r_1+t_1$ (where $t_1$ is random and chosen by Party B) and have both parties learn this.
+Without loss of generality, let all $s_i$ be values in some additive ring $(\mathbb{Z},+)$ (e.g., the additive ring $(\mathbb{Z}_{2^{64}},+)$ i.e., the additive ring of integers modulo $2^{64}$).
+Let $r_1,\cdots,r_{k-1}$ be a sequence of random values sampled from the same ring $(\mathbb{Z},+)$ by Alice (in our implementation, all $r_i$ values are generated by a pseudorandom function).
+Let $Q_1$ be the program that takes as inputs $x,y, r_1$ (and empty state $s_0$), runs $P_1$ (as defined above) to get $s_1$ and then outputs $o_1 = s_1 + r_1$ {\em only to} Bob\footnote{While the description of our protocol here assumes that the underlying backend supports only one party receiving output, this is only a simplifying assumption, and we can easily modify our protocol in the case where both parties must receive the same output.
+To do so, we let $Q_1$ output $o_1 = s_1+r_1+t_1$ (where $t_1$ is random and chosen by Bob) and have both parties learn this.
 We can then appropriately modify the remaining steps as well.}.
-The output of $Q_1$ for Party A in Step 1 is defined to be $r_1$.
-Now, every $Q_i$ ($2\leq i\leq k-1$) takes as input $x,y,r_{i-1},o_{i-1},r_i$, runs $P_i$ on inputs $x,y$ and state $o_{i-1}-r_{i-1}$ to get $s_i$ (where $-$ denotes subtraction in the ring $(\mathbb{Z},+)$) and then produces $s_i+r_i$ as output only to Party B.
-Again, Party A's output for $P_i$ is defined to be $r_i$.
-$Q_k$ takes as input $x,y,r_{k-1},o_{k-1}$, runs $P_k$ on inputs $x,y$ and state $o_{k-1}-r_{k-1}$ and outputs $z$ to both parties.
+Alice's output from $Q_1$ is $r_1$.
+Next, every $Q_i$ ($2\leq i\leq k-1$) takes as inputs $x,y,r_{i-1},o_{i-1},r_i$, runs $P_i$ on inputs $x,y$ and state $o_{i-1}-r_{i-1}$ (where $-$ denotes subtraction in the ring $(\mathbb{Z},+)$) and then outputs $s_i+r_i$ only to Bob.
+Again, Alice's output from $P_i$ is $r_i$.
+The last program $Q_k$ takes inputs $x,y,r_{k-1},o_{k-1}$, runs $P_k$ on inputs $x,y$ and state $o_{k-1}-r_{k-1}$ and outputs $z$ to both parties.
 
-Our compiler, given $P$, will first split $P$ into $P_1||P_2||\cdots||P_k$ with the guarantee that the ABY circuits generated for every $Q_i$ defined above is small enough to fit into memory.
-It will then compile and execute the $k$ programs $Q_1||Q_2||\cdots||Q_k$ sequentially (freeing up memory usage after every execution) to produce final output $z$ to both parties.
+Given $P$, partitioning first decomposes $P$ into $P_1||P_2||\cdots||P_k$ with the guarantee that the circuits generated for every $Q_i$ defined above are small enough to fit in the machine memory.
+We then compile and execute the $k$ programs $Q_1||Q_2||\cdots||Q_k$ sequentially (freeing up memory usage after  execution of each $Q_i$) to finally output $z$ to both parties.
+
+\begin{theorem}[Correctness and security of partitioning]
+Let $P$ be a program in our source language with output $z$, that is decomposed into programs $Q_1||Q_2||\cdots ||Q_k$ (as defined in the partioning procedure above) and compiled into the protocol $\prot_1||\prot_2\cdots ||\prot_k$ that outputs $r_1||r_2||\cdots||z$ (to Alice) and $o_1||o_2||\cdots||z$ (to Bob), then $\prot$ securely realizes $P$.
+\end{theorem}
+
+\noindent {\em Proof.} From Theorem \ref{theorem:security}, it follows that $\prot_i$ securely realizes $Q_i$ for every $1\leq i\leq k$.
+Hence, we know that the only information learnt by Alice is $r_1||r_2||\cdots||r_{k-1}||z$ and by Bob is $o_1||o_2||\cdots||o_{k-1}||z$.
+Since $r_i$ and $o_i$ ($1\leq i\leq k-1$) are individually uniformly random (in $(\mathbb{Z},+)$, the proof follows.
+
+Our current implementation of secure code partitioning requires a manual step.
+In particular, the developer needs to decompose the large program $P$ into 
+the sequence of small programs $P_1||\ldots||P_k$ manually.
+Then \tool generates $Q_1||\ldots||Q_k$ automatically.
+Automating this decomposition step requires an analysis that can 
+statically estimate the resource usage of a \tool program and we leave the construction of such
+an analysis as future work.

### 2017-10-27 OaklandSubmission/intro.tex
intro

@@
 show-stopper for applications like secure machine learning, secure
-prediction, etc. that operate on large amounts of data. \aseem{Can we
-cite something here to support this claim?}
+prediction, etc. that operate on large amounts of data.
+%\aseem{Can we
+%cite something here to support this claim?}
 \tool addresses the scalability concern using a novel technique that
@@
 to the rescue -- it automatically inserts the required instrumentation
-to ensure security in the presence of these intermediate outputs. As
+to ensure security of these intermediate outputs. As
 we show in our
@@
 compiled program. The simulation
-theorem not only guarantees that a
-well-typed source program terminates successfully in the ideal
-semantics (e.g. there are no array index out-of-bounds errors), but
-also that the observable outputs in the two semantics are the same. For
+guarantees that for a well-typed source program, the two semantics
+successfully terminate
+(e.g. there are no array index out-of-bounds errors), and that the
+observable outputs in the two semantics are
+identical. For
 the security theorem, we formally reduce the security
@@
 scheme.
+
+We have implemented the \tool toolchain using ABY~\cite{aby} as the
+cryptographic back-end. We evaluate \tool by implementing a wide range
+of secure prediction benchmarks including linear and na\"{i}ve bayes
+classifiers, decision trees, deep neural networks, state-of-the-art
+classifiers such as Tensorflow~\cite{tensorflow}
+and \bonsai~\cite{bonsai}, and also the matrix factorization example
+from Nikolaenko et al.~\cite{valeriaMatrix}. Our results demonstrate
+three key points. First, \tool makes it convenient for general
+programmers to write \mpc protocols. E.g. we provide the first \mpc
+implementation of \bonsai~\cite{bonsai}, and it was programmed at a
+high-level in \tool by a
+cryptography novice. Second, the performance of the protocols
+generated by \tool are comparable to or better than (up to 19x in some
+cases) their state-of-the-art specialized implementations. Finally, we
+show that with partitioning, we are able to implement applications
+that require more than 100 million gates.
+
+The rest of the paper is structured as follows. We formalize \tool and
+prove its correctnss and security theorems in
+Section~\ref{sec:ld}. Section~\ref{sec:pipe} details our partitioning
+scheme and formally proves its security. Section~\ref{sec:impl}
+and~\ref{sec:eval} present details of \tool implementation and
+evaluation. Finally, Section~\ref{sec:related} talks about the related
+work. We start with an overview of \tool.
+
+%% Most works cite garbled
+%% circuit based two-party computation as the most practical approach so
+%% far and use it as the baseline for comparison of performance. But, it
+%% has been shown multiple times that the garbled circuits based approach
+%% does not scale for realistic examples. Even for small benchmarks
+%% (e.g., Na\"{i}ve Bayes) the size of garbled circuit can be
+%% huge. Hence, comparisons are often done on artificial data sizes and
+%% garbled circuits are shown to be 500x slower than hand-crafted
+%% protocols \cite{shafindss}. This is cited as the motivation for need
+%% for specialized protocols for different ML algorithms in secure
+%% prediction such as linear regression \cite{shafindss}, decision
+%% trees \cite{wu}, deep neural networks \cite{minionn} and matrix
+%% factorization \cite{valeriaMatrix}. Our work is based on generic \mpc
+%% and we show competitive performance with the specialized protocols
+%% contrary to the popular belief. Hence, we believe \tool based
+%% implementations can serve as the performance baseline for future works
+%% on specialized protocols.
 
@@
 %% {\aseem{Not sure about this baseline paragraph. Did not touch it.}}
-%% \subsubsection*{Performance Baseline} Most works cite garbled
-%% circuit based two-party computation as the most practical approach so
-%% far and use it as the baseline for comparison of performance. But, it
-%% has been shown multiple times that the garbled circuits based approach
-%% does not scale for realistic examples. Even for small benchmarks
-%% (e.g., Na\"{i}ve Bayes) the size of garbled circuit can be
-%% huge. Hence, comparisons are often done on artificial data sizes and
-%% garbled circuits are shown to be 500x slower than hand-crafted
-%% protocols \cite{shafindss}. This is cited as the motivation for need
-%% for specialized protocols for different ML algorithms in secure
-%% prediction such as linear regression \cite{shafindss}, decision
-%% trees \cite{wu}, deep neural networks \cite{minionn} and matrix
-%% factorization \cite{valeriaMatrix}. Our work is based on generic \mpc
-%% and we show competitive performance with the specialized protocols
-%% contrary to the popular belief. Hence, we believe \tool based
-%% implementations can serve as the performance baseline for future works
-%% on specialized protocols.
 
@@
 
-\aseem{Checkpointing here. Three more paragraphs to come, one for
-summary of implementation and evaluation, one for related work, and
-one for baseline argument.}
+%% \aseem{Checkpointing here. Three more paragraphs to come, one for
+%% summary of implementation and evaluation, one for related work, and
+%% one for baseline argument.}
 
-\noindent\textbf{Other Related Work.} 
-Before ABY \cite{aby}, to improve efficiency, several works proposed
-to mix secure computation protocols based on homomorphic
-encryption and Yao's garbled circuits protocol, e.g., \cite{barni,blanton,brickell,franz,huang,valeriaMatrix,valeriaRidge,schropferK11}
-  and developed tools that allowed writing a mixed \mpc protocol, e.g., \cite{bogdanov,lone,tasty}. The breaking up of the function into homomorphic encryption and garbled circuits can
-even be done automatically as was shown in \cite{autoS}. 
-However, as was noted by ABY, due to high conversion cost between homomorphic encryption and Yao's garbled circuit, these mixed protocols do not gain much performance over a single protocol.
+%% \noindent\textbf{Other Related Work.} 
+%% Before ABY \cite{aby}, to improve efficiency, several works proposed
+%% to mix secure computation protocols based on homomorphic
+%% encryption and Yao's garbled circuits protocol, e.g., \cite{barni,blanton,brickell,franz,huang,valeriaMatrix,valeriaRidge,schropferK11}
+%%   and developed tools that allowed writing a mixed \mpc protocol, e.g., \cite{bogdanov,lone,tasty}. The breaking up of the function into homomorphic encryption and garbled circuits can
+%% even be done automatically as was shown in \cite{autoS}. 
+%% However, as was noted by ABY, due to high conversion cost between homomorphic encryption and Yao's garbled circuit, these mixed protocols do not gain much performance over a single protocol.
 Also, none of these works give a high-level developer-friendly language support.
-No prior work has given a formal type system and operational semantics for mixed arithmetic and boolean back-ends.
-We provide a detailed survey of related work in \sectionref{related}.
+%% No prior work has given a formal type system and operational semantics for mixed arithmetic and boolean back-ends.
+%% We provide a detailed survey of related work in \sectionref{related}.

### 2017-10-27 OaklandSubmission/intro.tex
nit
intro

@@
 
-Since its introduction in 1980s, there has been a long sequence
-of work that has transformed \mpc from a mere theoretical concept to
-something practically efficient and usable. To understand the
-state-of-the-art, let us consider the following example underlying many
-secure prediction algorithms. Suppose Alice wants to write
+Our goal is to build a framework that enables programmers, who are not
+expert in cryptography, to easily program efficient and scalable \mpc
+protocols. Since its introduction in 1980s, there has been a long
+sequence of work along the same lines. To understand the
+state-of-the-art, let us consider an example underlying many
+secure prediction algorithms. Say Alice wants to write
 a \mpc protocol to securely
@@
 $\cdot^{T}$ is the matrix transpose operator, and $w^Tx$ denotes the
-inner product of  $w^{T}$ and $x$. With the current state-of-the-art,
-Alice has following options:
+inner product of  $w^{T}$ and $x$. Alice has following options.
 
@@
 This paper presents \tool\footnote{Read as ``easy peasy'',
-stands for Easy 2 Party Computation}, a new \mpc framework.
+stands for Easy 2 Party Computation.}, a new \mpc framework.
 \tool is the first
@@
 
-\subsubsection*{Ease of programming} \tool source programs are ideal
-functionalities that describe ``what'' computation needs to be done,
-rather than ``how'' to do it. In particular, the programmer writes the
-high-level computation without thinking about the underlying
-cryptographic details. For example, an \tool source
-program for $w^Tx >b$ is as follows:
-
-\vspace{0.2cm}
+\begin{figure}
 \begin{lstlisting}[language=C]
@@
 \end{lstlisting}
+\caption{\tool code for $w^Tx >b$}
+\label{fig:introexample}
+\end{figure}
 
-\vspace{0.2cm}
+\subsubsection*{Ease of programming} \tool source programs are ideal
+functionalities that describe ``what'' computation needs to be done,
+rather than ``how'' to do it. In particular, the programmer writes the
+high-level computation without thinking about the underlying
+cryptographic details. For example, Figure~\ref{fig:introexample}
+shows an \tool source
+program for $w^Tx >b$.
+%
+%\vspace{0.2cm}
+%
+%\vspace{0.2cm}
 The program is quite similar to what a programmer might
@@
 instance, \tool compiler realizes that the array index~\ls{i} is
-public, and hence, the array accesses need not be compiled as
-multiplexers. Further, within the secret parts, \tool compiler is
+public, and hence, the array accesses need not be compiled
+obliviously. Further, within the secret parts, \tool compiler is
 aware
@@
 
-We have implemented the \tool toolchain using ABY~\cite{aby} as the
+\subsubsection*{Evaluation} We have implemented the \tool toolchain
+using ABY~\cite{aby} as the
 cryptographic back-end. We evaluate \tool by implementing a wide range
@@
 classifiers, decision trees, deep neural networks, state-of-the-art
-classifiers such as Tensorflow~\cite{tensorflow}
+classifiers from Tensorflow~\cite{tensorflow}
 and \bonsai~\cite{bonsai}, and also the matrix factorization example
@@
 programmers to write \mpc protocols. E.g. we provide the first \mpc
-implementation of \bonsai~\cite{bonsai}, and it was programmed at a
-high-level in \tool by a
-cryptography novice. Second, the performance of the protocols
+implementation of \bonsai~\cite{bonsai}, and it was programmed in the
+high-level \tool source language by a non-cryptographer. Second, the
+performance
+of the protocols
 generated by \tool are comparable to or better than (up to 19x in some
@@
 show that with partitioning, we are able to implement applications
-that require more than 100 million gates.
+that require more than 300 million gates.
+
+\subsubsection*{Related Work}
+Before ABY~\cite{aby}, several works have proposed
+combining secure computation protocols based on homomorphic
+encryption and Yao's garbled circuits
+(e.g. \cite{barni,blanton,brickell,franz,huang,valeriaMatrix,valeriaRidge,schropferK11}),
+  and some have also developed tools that allow writing such combinations
+  (e.g. \cite{bogdanov,lone,tasty}). Kerschbaum et al.~\cite{autoS}
+  even provide a scheme to automatically select optimal protocols
+  for sub-compuations. However, as Demmler et al.~\cite{aby} note, due
+  to high conversion cost between
+homomorphic encryption and Yao's garbled circuits, these combined
+protocols do not gain much performance over a single
+protocol. Additionally, these prior works neither have a
+developer-friendly language, nor do they formalize their type system
+and operational semantics, and show them to be consistent.
+We provide a detailed survey of related work in \sectionref{related}.
 
@@
 Section~\ref{sec:ld}. Section~\ref{sec:pipe} details our partitioning
-scheme and formally proves its security. Section~\ref{sec:impl}
-and~\ref{sec:eval} present details of \tool implementation and
-evaluation. Finally, Section~\ref{sec:related} talks about the related
-work. We start with an overview of \tool.
+scheme and proves its security. Section~\ref{sec:impl}
+presents the implementation details, and
+Section~\ref{sec:eval} present a comprehensive evaluation of \tool. We
+begin with an overview.
 
@@
 %% one for baseline argument.}
-
-%% \noindent\textbf{Other Related Work.} 
-%% Before ABY \cite{aby}, to improve efficiency, several works proposed
-%% to mix secure computation protocols based on homomorphic
-%% encryption and Yao's garbled circuits protocol, e.g., \cite{barni,blanton,brickell,franz,huang,valeriaMatrix,valeriaRidge,schropferK11}
-%%   and developed tools that allowed writing a mixed \mpc protocol, e.g., \cite{bogdanov,lone,tasty}. The breaking up of the function into homomorphic encryption and garbled circuits can
-%% even be done automatically as was shown in \cite{autoS}. 
-%% However, as was noted by ABY, due to high conversion cost between homomorphic encryption and Yao's garbled circuit, these mixed protocols do not gain much performance over a single protocol.
-Also, none of these works give a high-level developer-friendly language support.
-%% No prior work has given a formal type system and operational semantics for mixed arithmetic and boolean back-ends.
-%% We provide a detailed survey of related work in \sectionref{related}.

### 2017-10-27 OaklandSubmission/ld.tex
ld.tex

@@
 
-\divya{Start here...}
 \subsubsection*{Circuit semantics} Evaluating a program in the intermediate language produces a
@@
 mapped to (random) strings $b$. The semantics of these strings is
-given by pairs of encryption-decryption functions\nc{perhaps call this encode-decode functions since its technically not encryption-decryption?
-In fact more appropriate might be share-reconstruct, but that would require changing notation so lets not do that.}, written as
+given by pairs of encode-decode algorithms, written as
 $\mathcal{E}_{m}$ and $\mathcal{D}_{m}$ (where $m$ is either
-$\mathcal{A}$ or $\mathcal{B}$). More concretely,
+$\mathcal{A}$ or $\mathcal{B}$). 
+More concretely,
 $\mathcal{E}_{m}(c)$ returns a pair of two string $b_{1}$ and $b_{2}$
-(shares of the two parties), with the property that
-$\mathcal{D}_{m}(b_{1}, b_{2}) = c$. We assume that the underlying MPC
+ with the property that
+$\mathcal{D}_{m}(b_{1}, b_{2}) = c$. We assume that the underlying \mpc
 protocol instantiates $\mathcal{E}_{m}$ and $\mathcal{D}_{m}$
 appropriately.
-
-The circuit semantics is shown in Figure~\ref{fig:cktsem} using the
-judgments $\widehat{\rho}_{1}, \widehat{\rho}_{2} \vdash \kappa
+For ABY protocol \cite{aby}, algorithms $(\mathcal{E}_{\mathcal{A}},\mathcal{D}_{\mathcal{A}})$ (resp.
+$(\mathcal{E}_{\mathcal{B}},\mathcal{D}_{\mathcal{B}})$) correspond to the arithmetic (resp.
+boolean) sharing and reconstruction algorithms.
+
+\figureref{cktsem} gives the judgements for evaluation of circuit by the two parties using the \mpc protocol. The circuit environment for party $j$ is defined as follows:
+
+\vspace{0.2cm}
+$
+\small
+\begin{array}{rrcl}
+    \ftext{Circuit environment} & \widehat\rho_j &::=& \cdot \mid \widehat\rho_j,\wire \mapsto b_j\\
+\end{array}
+$
+\vspace{0.2cm}
+
+\divya{strings $b$ are not formally defined.}
+
+We give the 
+judgments $\widehat{\rho}_{1}, \widehat{\rho}_{2} \vdash \gate
 \Downarrow b_{1}, b_{2}$, and $\widehat{\rho}_{1}, \widehat{\rho}_{2}
 \vdash \chi \Downarrow \widehat{\rho}'_{1}, \widehat{\rho}'_{2}; O$,
-where $\widehat{\rho}_{1}$ and $\widehat{\rho}_{2}$ are the circuit
-environments of the two parties, mapping wire ranges $r$ to strings
-$b$. Rule ({\sc{C-Add}}) evinces the pattern for evaluating circuit
-gates $\kappa$. To evaluate $\kw{add}\:\kappa_{1}\:\kappa_{2}$, the
-rule first evaluates $\kappa_{1}$ to $(b_{11}, b_{21})$ and
-$\kappa_{2}$ to $(b_{12}, b_{22})$. Shares $(b_{11}, b_{21})$ are then
+where $O$ are the observations (similar to source semantics). 
+Rule ({\sc{C-In}}) is used to evaluate the input gate $\kw{in}^{m}_{j}$ that creates the $m$ type shares of the value $c$ input by party $j$.
+Rule ({\sc{C-Mult}}) evinces the pattern for evaluating circuit
+gates $\gate$. To evaluate $\kw{mult}\:\gate_{1}\:\gate_{2}$, the
+rule first evaluates $\gate_{1}$ to $(b_{11}, b_{21})$ and
+$\gate_{2}$ to $(b_{12}, b_{22})$. Shares $(b_{11}, b_{21})$ are then
 combined using $\mathcal{D_{\mathcal{A}}}$ to $n_{1}$, and similarly
 $(b_{12}, b_{22})$ are combined to $n_{2}$. The final output of the
-\kw{add} gate is then $\mathcal{E}_{\mathcal{A}}(n_{1} + n_{2})$. Note
-that this is a functional description of how the \kw{add} gate
+\kw{mult} gate is then $\mathcal{E}_{\mathcal{A}}(n_{1} \mulop n_{2})$. Note
+that this is a functional description of how the \kw{mult} gate
 evaluates, of course, concretely $n_{1}$ and $n_{2}$ are not observed
-by the parties. Rule ({\sc{C-Coerce}}) re-encrypts the shares using the
+by the parties. Rule ({\sc{C-Coerce}}) creates the shares using the
 scheme for $m$. Coming to the evaluation of circuits, the evaluation
-of \kw{bind} updates the mapping of $r$ in the input environments, and
+of \kw{bind} updates the mapping of $\wire$ in the input environments, and
 the rule ({\sc{C-Out}}) outputs the clear value $c$ to the
@@
   \small
-  \fbox{$\widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \kappa \Downarrow b_{1}, b_{2}$}
+  \fbox{$\widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \gate \Downarrow b_{1}, b_{2}$}
   \[
@@
                {
-                 b_{1}, b_{2} = \mathcal{E}_{m}(c)
+                 (b_{1}, b_{2}) = \mathcal{E}_{m}(c)
                }
@@
                {
-                 \widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \kappa \Downarrow b_{1}, b_{2}\\\\
+                 \widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \gate \Downarrow b_{1}, b_{2}\\\\
                  c = \mathcal{D}_{m_{1}}(b_{1}, b_{2})\\\\
-                 b'_{1}, b'_{2} = \mathcal{E}_{m}(c)
-               }
-               {
-                 \widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \kappa \rhd m \Downarrow b'_{1}, b'_{2}
-               }
-\\\\
-
-    \inferrule*[lab={\footnotesize{C-Add}}]
-               {
-                 \forall i \in \{1, 2\}.\:\widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \kappa_{i} \Downarrow b_{1i}, b_{2i}\quad
+                 (b'_{1}, b'_{2}) = \mathcal{E}_{m}(c)
+               }
+               {
+                 \widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \gate \rhd m \Downarrow b'_{1}, b'_{2}
+               }
+\\\\
+
+    \inferrule*[lab={\footnotesize{C-Mult}}]
+               {
+                 \forall i \in \{1, 2\}.\:\widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \gate_{i} \Downarrow b_{1i}, b_{2i}\quad
                  n_{i} = \mathcal{D}_{\mathcal{A}}(b_{1i}, b_{2i})\\\\
-                 b_{1}, b_{2} = \mathcal{E}_{\mathcal{A}}(n_{1} + n_{2})
-               }
-               {
-                 \widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \kw{add}\:\kappa_{1}\:\kappa_{2} \Downarrow b_{1}, b_{2}
+                 (b_{1}, b_{2}) = \mathcal{E}_{\mathcal{A}}(n_{1} \mulop n_{2})
+               }
+               {
+                 \widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \kw{mult}\:\gate_{1}\:\gate_{2} \Downarrow b_{1}, b_{2}
                }
@@
                {
-                 \forall i \in \{1, 2\}.\:\widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \kappa_{i} \Downarrow b_{1i}, b_{2i}\quad
+                 \forall i \in \{1, 2\}.\:\widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \gate_{i} \Downarrow b_{1i}, b_{2i}\quad
                  n_{i} = \mathcal{D}_{\mathcal{B}}(b_{1i}, b_{2i})\\\\
-                 b_{1}, b_{2} = \mathcal{E}_{\mathcal{B}}(n_{1} > n_{2})
-               }
-               {
-                 \widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \kw{gt}\:\kappa_{1}\:\kappa_{2} \Downarrow b_{1}, b_{2}
+                 (b_{1}, b_{2}) = \mathcal{E}_{\mathcal{B}}(n_{1} > n_{2})
+               }
+               {
+                 \widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \kw{gt}\:\gate_{1}\:\gate_{2} \Downarrow b_{1}, b_{2}
                }
@@
                {
-                 \forall i \in \{1, 2, 3\}.\:\widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \kappa_{i} \Downarrow b_{1i}, b_{2i}\quad
+                 \forall i \in \{1, 2, 3\}.\:\widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \gate_{i} \Downarrow b_{1i}, b_{2i}\quad
                  c_{i} = \mathcal{D}_{\mathcal{B}}(b_{1i}, b_{2i})\\\\
-                 c_{1} = \top \Rightarrow b_{1}, b_{2} = \mathcal{E}_{\mathcal{B}}(c_{2})\quad
-                 c_{1} = \bot \Rightarrow b_{1}, b_{2} = \mathcal{E}_{\mathcal{B}}(c_{3})
-               }
-               {
-                 \widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \kw{mux}\:\kappa_{1}\:\kappa_{2}\:\kappa_{3} \Downarrow b_{1}, b_{2}
+                 (c_{1} = \top) \Rightarrow \left((b_{1}, b_{2}) = \mathcal{E}_{\mathcal{B}}(c_{2})\right)\hspace{0.1cm}
+                 (c_{1} = \bot) \Rightarrow \left((b_{1}, b_{2}) = \mathcal{E}_{\mathcal{B}}(c_{3})\right)
+               }
+               {
+                 \widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \kw{mux}\:\gate_{1}\:\gate_{2}\:\gate_{3} \Downarrow b_{1}, b_{2}
                }
@@
   \\\\
-    \fbox{$\widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \chi \Downarrow \widehat{\rho'_{1}}, \widehat{\rho'_{2}}; O$}
+    \fbox{$\widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \crct \Downarrow \widehat{\rho'_{1}}, \widehat{\rho'_{2}}; O$}
   \[
@@
                {
-                 \widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \kappa \Downarrow b_{1}, b_{2}\\\\
-                 \widehat{\rho'_{1}} = \widehat{\rho_{1}}[r \mapsto b_{1}] \quad
-                 \widehat{\rho'_{2}} = \widehat{\rho_{2}}[r \mapsto b_{2}]
-               }
-               {
-                 \widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \kw{bind}\:\kappa\:r \Downarrow \widehat{\rho'_{1}}, \widehat{\rho'_{2}}; \cdot
+                 \widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \gate \Downarrow b_{1}, b_{2}\\\\
+                 \widehat{\rho'_{1}} = \widehat{\rho_{1}}[\wire \mapsto b_{1}] \quad
+                 \widehat{\rho'_{2}} = \widehat{\rho_{2}}[\wire \mapsto b_{2}]
+               }
+               {
+                 \widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \kw{bind}\:\gate\:\wire \Downarrow \widehat{\rho'_{1}}, \widehat{\rho'_{2}}; \cdot
                }
@@
                {
-                 \widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \kappa \Downarrow b_{1}, b_{2}\\\\
+                 \widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \gate \Downarrow b_{1}, b_{2}\\\\
                  c = \mathcal{D}_{m}(b_{1}, b_{2})
@@
                {
-                 \widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \kw{out}\:\kappa \Downarrow \widehat{\rho_{1}}, \widehat{\rho_{2}}; c
+                 \widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \kw{out}\:\gate \Downarrow \widehat{\rho_{1}}, \widehat{\rho_{2}}; c
                }
@@
 \Gamma$, the judgment checks that $\rho(x) = \widetilde{\rho}(x) =
-n$, for some $n$. We elide the judgment, providing the full definition
+n$, for some $n$.
+\divya{This judgement is not clear.
+If I am correct, it says that for public values $\rho$ and $\widetilde{\rho}$ are consistent and for secret values $\rho$ has values and $\widehat{rho}_j$ contain shares.} We elide the judgment, providing the full definition
 in the supplementary material.
@@
   \begin{enumerate}
-  \item $\Gamma \vdash s \leadsto \widehat{s} \mid \Gamma_{1}$
+  \item $\Gamma \vdash s \leadsto \widetilde{s} \mid \Gamma_{1}$
   \item $\Gamma \vdash \rho \leadsto \widetilde{\rho}; \widehat{\rho}_{1}, \widehat{\rho}_{2}$
@@
 The theorem states that if a source statement $s$ is well-typed, and
-compiles to $\widetilde{s}$, then $s$ terminates in the source
-semantics with observations $O_{1}$, $\widetilde{s}$ terminates in the
-target semantics with circuit $\chi$, and $\chi$ terminates in the
+compiles to $\widetilde{s}$ in intermediate language, then $s$ terminates in the source
+semantics with observations $O_{1}$, $\widetilde{s}$ evaluates to circuit $\crct$, and $\crct$ terminates in the
 circuit semantics with observations $O_{2}$, where $O_{1} = O_{2}$
 Furthermore, the final environments are also in the compilation
-relation.
-
-\nc{Should we point out that the proof of Theorem 1 is provided elsewhere?}
-\subsubsection*{Security theorem} The protocols we generate provide
-simulation-based security against a semi-honest adversary, in the
-framework of ~\cite{gmw,can00,can01}. At a very high level, in this
-framework, parties are modeled as non-uniform interactive turing
-machines (ITMs), with inputs provided by an environment $\env$. An
-adversary $\adv$, selects and ``corrupts'' one of the parties -
-however, $\adv$ still follows the protocol specification. $\adv$
-interacts with the $\env$, which observes the view of the corrupted
-party. At the end of the interaction, $\env$ outputs a single bit. Two
-different interactions are defined: the {\em real world} and an {\em
-  ideal world}. In the real interaction, the parties run the protocol
-$\prot$ in the presence of $\adv$ and $\env$. Let
-$\real_{\prot,\adv,\env}$ denote the binary distribution ensemble
-describing $\env$'s output in this interaction. In the ideal
-interaction, parties send their inputs to an additional entity, a
-trusted functionality machine $\F$ that carries the desired
-computation truthfully. Let $\simu$ (the simulator) denote the
-adversary in this idealized execution, and $\ideal_{\F,\simu,\env}$
-the binary distribution ensemble describing $\env$'s output after
-interacting with adversary $\simu$ and ideal functionality $\F$. A
-protocol $\prot$ is said to {\em securely realize} a functionality
-$\F$ if for every adversary $\adv$ in the real interaction, there is
-an adversary $\simu$ in the ideal interaction, such that no
-environment $\env$, on any input, can tell the real interaction apart
-from the ideal interaction, except with negligible probability (in the
-security parameter $\secparam$). More precisely, if the two binary
-distribution ensembles above are computationally indistinguishable.
-
-We shall assume a cryptographic MPC backend that securely implements any circuit $\chi$ that is output by our compiler.
-In more detail, this means that for every source program $s$, let $\chi$ be the circuit output by our compiler (as in Theorem \ref{theorem:correctness}).
-We assume that there exists a two-party secure computation protocol $\prot$ that securely realizes the functionality $\chi$ and we will call the corresponding simulator for the protocol as $\simu_{\tiny{2pc}}$ (that runs on input $\chi$ and the output produced by $\chi$ on the parties inputs).
-We note that the work of \cite{aby} provides such a protocol $\prot$ and simulator $\simu_{\tiny{2pc}}$ for all circuits $\chi$ output by our compiler.
-We are now ready to state and prove our security theorem.
-
-\begin{theorem}[Security]\label{theorem:security}
-Let $s$ be any functionality or program in our source language with outputs (or observations) $O_1$, that is compiled into a circuit $\chi$ (as defined in Theorem \ref{theorem:correctness}).
-Let protocol $\prot$ be the two-party secure computation protocol that securely realizes $\chi$ (as defined above).
-Then, $\prot$ securely realizes $s$.
-\end{theorem}
-
-\noindent {\em Proof.} Our simulator $\simu$ simply runs our compiler on program $s$ to obtain $\chi$ and upon receiving output $O_1$, executes $\simu_{\tiny{2pc}}$ on $\chi$ and $O_1$.
-Let $O_2$ be the outputs (or observations) of $\chi$.
-First, from Theorem \ref{theorem:correctness}, we have that the observations in $s$ and $\chi$ are the same (i.e.
-$O_2 = O_1$).
-Now, from the security of $\prot$, we have that the simulated view output by $\simu_{\tiny{2pc}}$ (with circuit $\chi$ and output $O_2$) is indistinguishable from the real view of the $\prot$ (when executed on circuit $\chi$ and when the output is $O_2$).
-Combining these two statements, the proof of the theorem follows.
-\aseem{Can we give a construction of the simulator S in the ideal
-  semantics? The simulator S whenever it receives an observable event,
-  outputs it to Z, and whenever A receives a string in the real
-  semantics, S outputs a random string to Z. Now, with theorem 1,
-  observable events are same in the ideal and real semantics, and with
-  the security of the crypto backend, the random strings and the
-  strings in the protocol are indistinguishable. Is it
-  incorrect/naive?
-
-  Also, we should emaphasize here that the ideal semantics is the
-  source semantics from Figure x and the real semantics is the circuit
-  semantics from Figure y. For the target language semantics, we
-  should emphasize that it does not have access to the secrets, so
-  there are no observations there.
-}
-
-\nc{OK now?}
-  %% \\
-  %% \[
-  %% \begin{array}{rrcl}
-  %%   \ftext{Value} & v &::=& c \mid [\overline{c_{i}}]_{n}\\
-  %%   \ftext{Observation} & O &::=& \cdot \mid c; O\\
-  %% \end{array}
-  %% \]
+relation. \divya{Should we call the judgement between environs as compilation relation?}
+Due to space reasons, we provide the proof in \cite{anonymized}.

### 2017-10-27 OaklandSubmission/security.tex
ld.tex

\divya{Start here......}
\subsubsection*{Security theorem} The protocols we generate provide
simulation-based security against a semi-honest adversary, in the
framework of ~\cite{gmw,can00,can01}. At a very high level, in this
framework, parties are modeled as non-uniform interactive turing
machines (ITMs), with inputs provided by an environment $\env$. An
adversary $\adv$, selects and ``corrupts'' one of the parties -
however, $\adv$ still follows the protocol specification. $\adv$
interacts with the $\env$, which observes the view of the corrupted
party. At the end of the interaction, $\env$ outputs a single bit. Two
different interactions are defined: the {\em real world} and an {\em
  ideal world}. In the real interaction, the parties run the protocol
$\prot$ in the presence of $\adv$ and $\env$. Let
$\real_{\prot,\adv,\env}$ denote the binary distribution ensemble
describing $\env$'s output in this interaction. In the ideal
interaction, parties send their inputs to an additional entity, a
trusted functionality machine $\F$ that carries the desired
computation truthfully. Let $\simu$ (the simulator) denote the
adversary in this idealized execution, and $\ideal_{\F,\simu,\env}$
the binary distribution ensemble describing $\env$'s output after
interacting with adversary $\simu$ and ideal functionality $\F$. A
protocol $\prot$ is said to {\em securely realize} a functionality
$\F$ if for every adversary $\adv$ in the real interaction, there is
an adversary $\simu$ in the ideal interaction, such that no
environment $\env$, on any input, can tell the real interaction apart
from the ideal interaction, except with negligible probability (in the
security parameter $\secparam$). More precisely, if the two binary
distribution ensembles above are computationally indistinguishable.

We shall assume a cryptographic MPC backend that securely implements any circuit $\chi$ that is output by our compiler. In more detail, this means that for every source program $s$, let $\chi$ be the circuit output by our compiler (as in Theorem \ref{theorem:correctness}). We assume that there exists a two-party secure computation protocol $\prot$ that securely realizes the functionality $\chi$ and we will call the corresponding simulator for the protocol as $\simu_{\tiny{2pc}}$ (that runs on input $\chi$ and the output produced by $\chi$ on the parties inputs). We note that the work of \cite{aby} provides such a protocol $\prot$ and simulator $\simu_{\tiny{2pc}}$ for all circuits $\chi$ output by our compiler. We are now ready to state and prove our security theorem.

\begin{theorem}[Security]\label{theorem:security}
Let $s$ be any functionality or program in our source language with outputs (or observations) $O_1$, that is compiled into a circuit $\chi$ (as defined in Theorem \ref{theorem:correctness}). Let protocol $\prot$ be the two-party secure computation protocol that securely realizes $\chi$ (as defined above). Then, $\prot$ securely realizes $s$.  
\end{theorem}

\noindent {\em Proof.} Our simulator $\simu$ simply runs our compiler on program $s$ to obtain $\chi$ and upon receiving output $O_1$, executes $\simu_{\tiny{2pc}}$ on $\chi$ and $O_1$. Let $O_2$ be the outputs (or observations) of $\chi$. First, from Theorem \ref{theorem:correctness}, we have that the observations in $s$ and $\chi$ are the same (i.e. $O_2 = O_1$). Now, from the security of $\prot$, we have that the simulated view output by $\simu_{\tiny{2pc}}$ (with circuit $\chi$ and output $O_2$) is indistinguishable from the real view of the $\prot$ (when executed on circuit $\chi$ and when the output is $O_2$). Combining these two statements, the proof of the theorem follows. 
\aseem{Can we give a construction of the simulator S in the ideal
  semantics? The simulator S whenever it receives an observable event,
  outputs it to Z, and whenever A receives a string in the real
  semantics, S outputs a random string to Z. Now, with theorem 1,
  observable events are same in the ideal and real semantics, and with
  the security of the crypto backend, the random strings and the
  strings in the protocol are indistinguishable. Is it
  incorrect/naive?

  Also, we should emaphasize here that the ideal semantics is the
  source semantics from Figure x and the real semantics is the circuit
  semantics from Figure y. For the target language semantics, we
  should emphasize that it does not have access to the secrets, so
  there are no observations there.
}

\nc{OK now?}
  %% \\
  %% \[
  %% \begin{array}{rrcl}
  %%   \ftext{Value} & v &::=& c \mid [\overline{c_{i}}]_{n}\\
  %%   \ftext{Observation} & O &::=& \cdot \mid c; O\\
  %% \end{array}
  %% \]

### 2017-10-27 OaklandSubmission/intro.tex
kerschbaum

@@
 (e.g. \cite{barni,blanton,brickell,franz,huang,valeriaMatrix,valeriaRidge,schropferK11}),
-  and some have also developed tools that allow writing such combinations
-  (e.g. \cite{bogdanov,lone,tasty}). Kerschbaum et al.~\cite{autoS}
-  even provide a scheme to automatically select optimal protocols
-  for sub-compuations. However, as Demmler et al.~\cite{aby} note, due
+and some have also developed tools that allow writing such
+combinations (e.g. \cite{bogdanov,lone,tasty}). However, as Demmler et
+al.~\cite{aby} note, due
  to high conversion cost between
@@
 protocols do not gain much performance over a single
-protocol. Additionally, these prior works neither have a
-developer-friendly language, nor do they formalize their type system
-and operational semantics, and show them to be consistent.
-We provide a detailed survey of related work in \sectionref{related}.
+protocol. Additionally, these prior works neither have
+a developer-friendly language, nor do they formalize their type
+system and operational semantics, and show them to be
+consistent. Kerschbaum et al.~\cite{autoS}
+provide a scheme to automatically select optimal protocols
+for sub-compuations, but they consider a toy language consisting only
+of arithmetic expressions. We provide a detailed survey of related
+work in \sectionref{related}.
 

### 2017-10-27 OaklandSubmission/main.tex
bib
minor
init
1.1
chat-bot
change title
cp for 1.0
intro
abstract first pass

@@
 
-\usepackage{epsfig,endnotes}
-\usepackage{varwidth}
-\usepackage{algorithm}
-\usepackage{algpseudocode}
-\usepackage{balance}
-\usepackage{color}
-\usepackage{nicefrac}
-\usepackage{amsmath}
-\usepackage{braket}
-\usepackage{bm}
-\usepackage{mathtools}
-\usepackage{multirow}
-\usepackage{bigdelim}
-\usepackage{mathtools}
-\usepackage{amssymb}
-\usepackage{indentfirst}
-\usepackage{booktabs}
-\usepackage{enumitem}
-\usepackage{boxedminipage}
-\usepackage{float}
-\usepackage{graphicx}
-\usepackage{caption}
-\usepackage{cleveref}
-\usepackage{subcaption}
-\usepackage[normalem]{ulem}
-\usepackage{xpatch}
-\usepackage{stmaryrd}
-
-\makeatletter
-\xpatchcmd{\algorithmic}{\itemsep\z@}{\itemsep=0.3ex plus2pt}{}{}
-\makeatother
-
-\renewcommand{\algorithmicrequire}{\textbf{Input:}}
-\renewcommand{\algorithmicensure}{\textbf{Output:}}
-\newcommand{\crefrangeconjunction}{ to~}
-\newcommand{\minus}{\scalebox{0.75}[1.0]{$-$}}
-\newcommand{\sameer}[1]{{\leavevmode\color{red}{#1}}}
-\setlength{\textfloatsep}{0.3cm}
-\newcommand{\ourprotocol}{Tuffy}
-
-\newcommand{\bits}{\{0,1\}}
-\newcommand{\Ex}{\mathbb{E}}
-\newcommand{\To}{\rightarrow}
-\newcommand{\R}{\mathbb{R}}
-\newcommand{\N}{\mathbb{N}}
-\newcommand{\Z}{\mathbb{Z}}
-\newcommand{\F}{\mathbb{F}}
-\newcommand{\GF}{\mathsf{GF}}
-
-\usepackage{MnSymbol}
-\newcommand{\maxpr}{\text{\rm max-pr}}
-\newcommand{\cclass}[1]{\mathbf{#1}}
-\renewcommand{\P}{\cclass{P}}
-\newtheorem{problem}{Problem}
-\newtheorem{theorem}{Theorem}
-\newtheorem{conjecture}[theorem]{Conjecture}
-\newtheorem{definition}[theorem]{Definition}
-\newtheorem{lemma}[theorem]{Lemma}
-\newtheorem{proposition}[theorem]{Proposition}
-\newtheorem{corollary}[theorem]{Corollary}
-\newtheorem{claim}[theorem]{Claim}
-\newtheorem{fact}[theorem]{Fact}
-\newtheorem{remk}[theorem]{Remark}
-\newtheorem{apdxlemma}{Lemma}
-
-\newcommand{\etal}{{\it et~al.\ }}
-\newcommand{\ie} {{\it i.e.,\ }}
-\newcommand{\eg} {{\it e.g.,\ }}
-\newcommand{\cf}{{\it cf.,\ }}
-\newcommand{\suchthat}{{\;\; : \;\;}}
-\newcommand{\pr}[1]{\Pr\left[#1\right]}
-
-\newcommand{\class}[1]{\mathbf{#1}}
-\newcommand{\SZK}{\class{SZK}}
-\newcommand{\BPP}{\class{BPP}}
-\newcommand{\NP}{\class{NP}}
-\renewcommand{\P}{\class{P}}
-\newcommand{\negl}{{\mathrm{neg}}}
-
-\newcommand{\Enc}{E}
-\newcommand{\Dec}{D}
-\newcommand{\Gen}{G}
-
-\newcommand{\PK}{\mathit{PK}}
-\newcommand{\SK}{\mathit{SK}}
-\newcommand{\pk}{\mathit{pk}}
-\newcommand{\sk}{\mathit{sk}}
-\newcommand{\MD}[1]{\mathrm{MD{#1}}}
-\newcommand{\SHA}{\mbox{SHA-1}}
+\input{headers}
 
 \begin{document}
-\title{SecureDNN: Efficient Analytics over Encrypted Data}
+\title{\tool: Programmable, Efficient, and Scalable Secure Two-Party Computation}
 
@@
 \maketitle
-% \begin{abstract}
-% The abstract goes here.
-% \end{abstract}
+\begin{abstract}
+Secure two-party computation (\mpc) allows two mutually distrusting parties to compute any function on their joint inputs, without revealing anything other than the output of the function to one another.
+While the application and potential impact of secure computation is virtually limitless, writing an efficient and scalable secure protocol is by no means an easy feat.
+Prior work in designing such protocols, suffered from either a) a lack of programmability (e.g.
+requiring the programmer to write cryptography-aware circuits for the computation); b) a lack of efficiency for several natural computations (such as machine learning applications); or c) required cryptographers to design hand-crafted protocols for various applications.
+In this work, we present \tool: a new ``cryptographic-cost aware'' \mpc framework that generates efficient and scalable \mpc protocols from high-level programs that contain absolutely no cryptographic details.
+Programs in our language are easy-to-write, with lines of code that are comparable to their insecure implementations.
+Furthermore, the \mpc protocols generated by our framework outperform or match almost all recent works on hand-crafted secure two-party computation protocols (sometimes up to a 19x factor!).
+\end{abstract}
 \IEEEpeerreviewmaketitle
@@
 
-\input{./Sections/CompleteProtocol.tex}
-\newpage
-\clearpage
-\input{./Sections/ReLU.tex}
-\input{./Sections/Independent.tex}
-\input{./Sections/Introduction.tex}
+\input{intro}
+%\input{fig-ex-sml}
+\input{fig-ex-aby}
+
+\input{ex}
+\input{ld}
+\input{pipe}
+\input{impl}
+
+\input{eval}
+\input{train}
+\input{related}
+\input{conclude}
 
@@
 \bibliographystyle{abbrv}
-\bibliography{./bib}
+\bibliography{bib}
 \balance
 
+\appendix
+\input{app-stash}
+
 \end{document}

### 2017-10-27 OaklandSubmission/security.tex

@@
-\divya{Start here......}
 \subsubsection*{Security theorem} The protocols we generate provide
@@
 however, $\adv$ still follows the protocol specification. $\adv$
-interacts with the $\env$, which observes the view of the corrupted
-party. At the end of the interaction, $\env$ outputs a single bit. Two
+interacts with $\env$ that observes the view of the corrupted
+party. At the end of the interaction, $\env$ outputs a single bit based on the output of the honest party and the view of the adversary. Two
 different interactions are defined: the {\em real world} and an {\em
@@
 $\prot$ in the presence of $\adv$ and $\env$. Let
-$\real_{\prot,\adv,\env}$ denote the binary distribution ensemble
-describing $\env$'s output in this interaction. In the ideal
-interaction, parties send their inputs to an additional entity, a
-trusted functionality machine $\F$ that carries the desired
+$\real_{\prot,\adv,\env}$ denote the  distribution ensemble
+describing $\env$'s output in this interaction. 
+In the ideal interaction, parties send their inputs to  a
+trusted functionality that performs the desired
 computation truthfully. Let $\simu$ (the simulator) denote the
-adversary in this idealized execution, and $\ideal_{\F,\simu,\env}$
-the binary distribution ensemble describing $\env$'s output after
-interacting with adversary $\simu$ and ideal functionality $\F$. A
+adversary in this ideal execution, and $\ideal_{\F,\simu,\env}$
+the distribution ensemble describing $\env$'s output after
+interacting with the ideal adversary $\simu$. A
 protocol $\prot$ is said to {\em securely realize} a functionality
@@
 from the ideal interaction, except with negligible probability (in the
-security parameter $\secparam$). More precisely, if the two binary
-distribution ensembles above are computationally indistinguishable.
+security parameter $\secparam$). More precisely, the above two
+distribution ensembles  are computationally indistinguishable.
 
-We shall assume a cryptographic MPC backend that securely implements any circuit $\chi$ that is output by our compiler.
-In more detail, this means that for every source program $s$, let $\chi$ be the circuit output by our compiler (as in Theorem \ref{theorem:correctness}).
-We assume that there exists a two-party secure computation protocol $\prot$ that securely realizes the functionality $\chi$ and we will call the corresponding simulator for the protocol as $\simu_{\tiny{2pc}}$ (that runs on input $\chi$ and the output produced by $\chi$ on the parties inputs).
-We note that the work of \cite{aby} provides such a protocol $\prot$ and simulator $\simu_{\tiny{2pc}}$ for all circuits $\chi$ output by our compiler.
+We shall assume a cryptographic \mpc backend that securely implements any circuit $\crct$ that is output by our compiler (see \figureref{circuits}).
+In more detail, this means that for any well-typed source program $s$, let $\crct$ be the circuit output by our compiler (as in \theoremref{correctness}).
+We assume that there exists a \mpc protocol $\prot$ that securely realizes the functionality $\crct$ and let $\simu_{\tiny{2pc}}$ be the corresponding simulator  (that runs on $\crct$, input of the corrupt party and the output obtained from trusted functionality for $\crct$).
+We note that the work of \cite{aby} provides such a protocol $\prot$ and simulator $\simu_{\tiny{2pc}}$ for all circuits $\crct$ output by our compiler\footnote{\divya{We note that \cite{aby} provides a \mpc protocol for a more general .....}}.
 We are now ready to state and prove our security theorem.
@@
 
-\noindent {\em Proof.} Our simulator $\simu$ simply runs our compiler on program $s$ to obtain $\chi$ and upon receiving output $O_1$, executes $\simu_{\tiny{2pc}}$ on $\chi$ and $O_1$.
+\noindent {\em Proof.} 
+Our simulator $\simu$ simply runs our compiler on program $s$ to obtain $\crct$.
+It is crucial that this compilation to circuits does not require the secret inputs of the parties.
+Next, $\simu$ sends the input of the corrupt party to the trusted functionality of $s$ to obtain outputs $O_1$.
+Note that $O_1$ is same as the observations in the source semantics.
+By \theoremref{correctness}, these outputs $O_1$ are identical to outputs (or observations) $O_2$ of $\crct$ under circuit semantics. 
+Next, $\simu$ runs $\simu_{\tiny{2pc}}$ on $\crct$, input of the corrupt party and $O_2$. 
+From the security of $\prot$, we have that the simulated view output by $\simu_{\tiny{2pc}}$ is indistinguishable from the real view of the $\prot$ for $\crct$ between the two parties.
+Hence, the security follows.
+
+% and upon receiving output $O_1$, executes $\simu_{\tiny{2pc}}$ on $\chi$ and $O_1$.
 Let $O_2$ be the outputs (or observations) of $\chi$.
@@
 Combining these two statements, the proof of the theorem follows.
-\aseem{Can we give a construction of the simulator S in the ideal
-  semantics? The simulator S whenever it receives an observable event,
-  outputs it to Z, and whenever A receives a string in the real
-  semantics, S outputs a random string to Z. Now, with theorem 1,
-  observable events are same in the ideal and real semantics, and with
-  the security of the crypto backend, the random strings and the
-  strings in the protocol are indistinguishable. Is it
-  incorrect/naive?
-
+\aseem{
   Also, we should emaphasize here that the ideal semantics is the

### 2017-10-27 OaklandSubmission/main.tex
security

@@
 \maketitle
-\begin{abstract}
-Secure two-party computation (\mpc) allows two mutually distrusting parties to compute any function on their joint inputs, without revealing anything other than the output of the function to one another.
-While the application and potential impact of secure computation is virtually limitless, writing an efficient and scalable secure protocol is by no means an easy feat.
-Prior work in designing such protocols, suffered from either a) a lack of programmability (e.g.
-requiring the programmer to write cryptography-aware circuits for the computation); b) a lack of efficiency for several natural computations (such as machine learning applications); or c) required cryptographers to design hand-crafted protocols for various applications.
-In this work, we present \tool: a new ``cryptographic-cost aware'' \mpc framework that generates efficient and scalable \mpc protocols from high-level programs that contain absolutely no cryptographic details.
-Programs in our language are easy-to-write, with lines of code that are comparable to their insecure implementations.
-Furthermore, the \mpc protocols generated by our framework outperform or match almost all recent works on hand-crafted secure two-party computation protocols (sometimes up to a 19x factor!).
-\end{abstract}
+% \begin{abstract}
+% The abstract goes here.
+% \end{abstract}
 \IEEEpeerreviewmaketitle
@@
 \input{ld}
+\input{security}
 \input{pipe}

### 2017-10-27 OaklandSubmission/intro.tex
minor

--- 2017-10-27 OaklandSubmission/intro.tex
@@
 \caption{\tool code for $w^Tx >b$}
-\label{fig:introexample}
+\label{fig:ex-sml}
 \end{figure}
@@
 high-level computation without thinking about the underlying
-cryptographic details. For example, Figure~\ref{fig:introexample}
+cryptographic details. For example, Figure~\ref{fig:ex-sml}
 shows an \tool source
@@
 (e.g. \cite{barni,blanton,brickell,franz,huang,valeriaMatrix,valeriaRidge,schropferK11}),
-  and some have also developed tools that allow writing such combinations
-  (e.g. \cite{bogdanov,lone,tasty}). Kerschbaum et al.~\cite{autoS}
-  even provide a scheme to automatically select optimal protocols
-  for sub-compuations. However, as Demmler et al.~\cite{aby} note, due
+and some have also developed tools that allow writing such
+combinations (e.g. \cite{bogdanov,lone,tasty,autoS}). However, as Demmler et
+al.~\cite{aby} note, due
  to high conversion cost between
@@
 protocols do not gain much performance over a single
-protocol. Additionally, these prior works neither have a
-developer-friendly language, nor do they formalize their type system
-and operational semantics, and show them to be consistent.
-We provide a detailed survey of related work in \sectionref{related}.
+protocol. 
+Additionally, these prior works provide informal languages
+or libraries and lack formal semantics.
+%Can be added to related work
+% Kerschbaum et al.~\cite{autoS}
+%provide a scheme to automatically assign homomorphic encryption or
+%garbled circuits to each operators in a computation that has been expressed as a sequence of dyadic operations.
+%They conjecture that the problem is NP hard and gave a linear programming based solution
+%as well as a quadratic time greedy heuristic. \tool  uses a simpler constant time heuristic
+%to assign arithmetic secret sharing or garbled circuits to sub-computations
+%in a richer language with control flow statements such as loops and branches.
+ We provide a detailed survey of related
+work in \sectionref{related}.
 

### 2017-10-27 OaklandSubmission/related.tex
edited related work

@@
 %\end{enumerate}
-\noindent\textbf{Automatic Generation of SMC protocols.}
-There has been vast literature on the automatic generation of SMC protocols.
-These works all provide tools at various levels of abstraction that help a programmer automatically generate cryptographic protocols for various functionalities.
+\noindent\textbf{Generating \mpc protocols.}
+There has been vast literature on the generation of \mpc protocols.
+These works all provide tools at various levels of abstraction that help a programmer generate cryptographic protocols for various functionalities.
 These levels can be classified as follows:
@@
 All such existing languages in literature are built with only one backend: Fairplay's Secure Function Definition Language (SFDL)~\cite{fairplay,fairplaymp} compiles high-level C or Pascal like programs into boolean circuits that are then evaluated using garbled circuits~\cite{yao} for the two-party case and distributed garbled circuits~\cite{bmr} for the multi-party case~\cite{bmr}.
-The Secure Multiparty Computation Language (SMCL)~\cite{smcl} that compiles high-level Java like programs into arithmetic circuits that are then evaluated using the VIFF framework~\cite{viff}.
+The Secure Multiparty Computation Language (SMCL)~\cite{smcl} that compiles high-level Java like programs into arithmetic circuits that are then evaluated using the VIFF framework~\cite{viff} is another example.
 The work of Wysteria~\cite{wysteria} provides a way of writing ``mixed-mode'' programs that combine normal or per-party, local computations, with ``secure'' (joint) computations repeatedly in a seamless manner.
@@
 While, the programmer need not know how ABY implements boolean or arithmetic computation, he or she must know the type of compute (boolean or arithmetic) that is more efficient for specific sub-computations and also explicitly write calls to convert shares from one scheme to another (boolean secret sharing to arithmetic secret sharing and vice-versa).
-While the work of Kerschbaum {\em et al.}~\cite{kos14} provides a heuristic to automatically select the protocol for various types of sub-computations, one would still need to manually write the program using the share conversion tools of ABY.
-Similarly, the TASTY tool~\cite{tasty} provides a way to mix both boolean and arithmetic circuits (using homomorphic encryption for the arithmetic compute).
+The work of Kerschbaum et al.~\cite{autoS} provides a scheme to automatically assign homomorphic encryption or garbled circuits to each operators in a computation that is expressed as a sequence of dyadic operations.
+They conjecture that the problem is NP hard and gave a linear programming based solution
+as well as a quadratic time greedy heuristic.
 As noted by ABY~\cite{aby}, the performance is much poorer than tools like ABY that combine boolean and arithmetic secure computation techniques due to very high conversion costs between homomorphic encryption and boolean shares as well as the cost of homomorphic encryption itself as the security parameter increases.
 This abstraction requires a fairly involved knowledge of cryptography from the programmer.
+%On the other hand \tool uses a simpler constant time heuristic to assign arithmetic secret sharing or garbled circuits to sub-computations in a richer language with control flow statements such as loops and branches.
+%While the work of Kerschbaum {\em et al.}~\cite{kos14} provides a heuristic to automatically select the protocol for various types of sub-computations, one would still need to manually write the program using the share conversion tools of ABY.
+Similarly, the TASTY tool~\cite{tasty} provides a way to mix both boolean and arithmetic circuits (using homomorphic encryption for the arithmetic compute).
 Other examples include the VIFF framework~\cite{viff} for arithmetic computations, Sharemind~\cite{sharemind} (that provides information-theoretic secure 3-party boolean computation), and the Secure Multiparty Computation language~\cite{nsy04,securemc}.
@@
 \noindent{\bf Protocol Implementation:} Tools at this level of abstraction require the programmer to have a deep understanding of how cryptographic secure computation protocols work and the programmer must provide calls to low-level cryptographic building blocks such as oblivious transfer.
-While the programmer need not know the specific oblivious transfer protocol implemented in the backend, he or she still needs to know how to build a secure computation protocol from such building blocks.
+For example while the programmer need not know the specific oblivious transfer protocol implemented in the backend, he or she still needs to know how to build a secure computation protocol from such building blocks.
 This abstraction requires the highest knowledge of cryptography from the programmer.
-Examples of such tools include L1~\cite{l1} and Qilin~\cite{qilin}.
+Examples of such tools include L1~\cite{lone} and Qilin~\cite{qilin}.
 
@@
 \\\\
-\noindent\textbf{Custom SMC protocols.} Many works have designed specialized, hand-crafted protocols for various secure two-party computation tasks.
+\noindent\textbf{Custom \mpc protocols.} Many works have designed specialized, hand-crafted protocols for various secure two-party computation tasks.
 The reason for doing this was to allow for a mix between arithmetic and boolean computations as a secure computation protocol for a pure boolean circuit would have been very inefficient.
 Such hand-crafting of protocols require deep cryptography know-how of designing secure computation protocols with various tools such as homomorphic encryption, garbled circuits, oblivious transfer and so on.
-A sample of such work includes various works on privacy-preserving classification of medical ElectroCardioGram (ECG)~\cite{ecgclassifier}, iris and fingerprint recognition~\cite{irisfingerprint}, remote diagnostics~\cite{remotediagnostics}, private sequence analysis~\cite{sequenceanalysis}, private biometric identification~\cite{biometricidentification}, privacy-preserving matrix factorization and ridge-regression~\cite{matrixfactorization, ridgeregression}, machine-learning classification~\cite{shafindss}, decision trees and forests~\cite{wu}, neural network training and prediction~\cite{secureml,minionn}.
-As seen from our experiments, the protocols generated by \tool give us comparable execution times as many of the above works with minimal programmer overhead.
-\nc{What do we want to say here exactly?}
+A sample of such work includes various works on privacy-preserving classification of medical ElectroCardioGram (ECG)~\cite{barni}, iris and fingerprint recognition~\cite{blanton}, remote diagnostics~\cite{brickell}, private sequence analysis~\cite{franz}, private biometric identification~\cite{huang}, privacy-preserving matrix factorization and ridge-regression~\cite{valeriaMatrix, valeriaRidge}, machine-learning classification~\cite{shafindss}, decision trees and forests~\cite{wu}, neural network training and prediction~\cite{secureml,minionn}.
+As seen from our experiments, the protocols generated by \tool give us better or comparable execution times as many of the above works with minimal programmer overhead.
+%\nc{What do we want to say here exactly?}
 
-\divya{We need to cite this paper \cite{yao-pipe} for yao pipelining. This technique was also used in \cite{valeriaMatrix,oblivm}}
+\divya{We need to cite this paper \cite{yao-pipe} for yao pipelining.
+This technique was also used in \cite{valeriaMatrix,oblivm}}\nc{We could just cite this in secure partitioning if we want to say that partitioning is similar in spirit to yao pipelining?
+Otherwise, is out of place here; isnt it?}

### 2017-10-28 OaklandSubmission/intro.tex
minor edits to intro and abstract.

@@
 Our goal is to build a framework that enables programmers, who are not
-expert in cryptography, to easily program efficient and scalable \mpc
+experts in cryptography, to easily program efficient and scalable \mpc
 protocols. Since its introduction in 1980s, there has been a long
@@
 %
-Hence, for better efficiency, Alice would ideally like to compute
+For better efficiency, Alice would ideally like to compute
 $w^Tx$ using an arithmetic circuit, and the comparison with $b$ using
@@
 or better than the custom, specialized
-protocols from previous work. Below we describe the
+protocols from previous work (up to a factor of 19x in our experiments, depending on the functionality) \nc{I added this but I guess we are saying it many times, but might be nice to highlight perf early no?}.
+Below we describe the
 salient features of \tool.
@@
 is comparable
-to or better than the custom, specialized protocols developed in the
-previous work. Indeed, \tool is the first such cryptographic-cost
+to or better than the custom, specialized protocols developed in previous works. Indeed, \tool is the first such cryptographic-cost
 aware compiler.
@@
 security theorems for our compiler. The correctness theorem proves a
-forward simluation between the
+forward simulation between the
 ``ideal'' semantics (the trusted-third party semantics) of a source
@@
 of our compiler to the security of the back-end cryptographic
-protocol. We also prove a formal security theorem for our partitioning
+protocol (by also invoking our correctness theorem). We also prove a formal security theorem for our partitioning
 scheme.
@@
 The rest of the paper is structured as follows. We formalize \tool and
-prove its correctnss and security theorems in
+prove its correctness and security theorems in
 Section~\ref{sec:ld}. Section~\ref{sec:pipe} details our partitioning
@@
 Section~\ref{sec:eval} present a comprehensive evaluation of \tool. We
-begin with an overview.
+begin with an overview of \tool through an example in Section~\ref{sec:ex}. \nc{provided a pointer to sec 2 as well.}
 

### 2017-10-28 OaklandSubmission/main.tex
minor edits to intro and abstract.

--- 2017-10-27 OaklandSubmission/main.tex
@@
 Programs in our language are easy-to-write, with lines of code that are comparable to their insecure implementations.
-Furthermore, the \mpc protocols generated by our framework outperform or match almost all recent works on hand-crafted secure two-party computation protocols (sometimes up to a 19x factor!).
+Furthermore, the \mpc protocols generated by our framework match or outperform almost all recent works that provide hand-crafted \mpc protocols for various complex functionalities (sometimes up to a 19x factor!).
 \end{abstract}

### 2017-10-28 OaklandSubmission/intro.tex
comments

@@
 or better than the custom, specialized
-protocols from previous work (up to a factor of 19x in our experiments, depending on the functionality) \nc{I added this but I guess we are saying it many times, but might be nice to highlight perf early no?}.
-Below we describe the
-salient features of \tool.
+protocols from previous work (up to a factor of 19x in our
+experiments, depending on the functionality) \nc{I added this but I
+  guess we are saying it many times, but might be nice to highlight
+  perf early no?} \aseem{My rationale was to save the numbers for
+  Evaluation summary and remain high-level here.} Below we describe
+the salient features of \tool.
 
@@
 Section~\ref{sec:eval} present a comprehensive evaluation of \tool. We
-begin with an overview of \tool through an example in Section~\ref{sec:ex}. \nc{provided a pointer to sec 2 as well.}
+begin with an overview of \tool through an example in
+Section~\ref{sec:ex}. \nc{provided a pointer to sec 2 as well.}
+\aseem{Sure :), I did not add it because Sec. 2 was just next line, so
+  I thought it did not need a pointer.}
 

### 2017-10-28 OaklandSubmission/intro.tex
nit
intro minor edit

--- 2017-10-28 OaklandSubmission/intro.tex
@@
 or better than the custom, specialized
-protocols from previous work (up to a factor of 19x in our experiments, depending on the functionality) \nc{I added this but I guess we are saying it many times, but might be nice to highlight perf early no?}.
-Below we describe the
-salient features of \tool.
+protocols from previous work %(up to a factor of 19x in our
+%experiments, depending on the functionality) \nc{I added this but I
+%  guess we are saying it many times, but might be nice to highlight
+%  perf early no?} \aseem{My rationale was to save the numbers for
+%  Evaluation summary and remain high-level here.} 
+Below we describe
+the salient features of \tool.
 
@@
 and boolean circuits, while the programmer remains
-oblivious of all these cryptographic details. Our evaluation
-shows that the performance of these automatically generated protocols
-is comparable
-to or better than the custom, specialized protocols developed in previous works. Indeed, \tool is the first such cryptographic-cost
-aware compiler.
+oblivious of all these cryptographic details. Indeed, \tool is the
+first such cryptographic-cost aware compiler.
 
@@
 Section~\ref{sec:eval} present a comprehensive evaluation of \tool. We
-begin with an overview of \tool through an example in Section~\ref{sec:ex}. \nc{provided a pointer to sec 2 as well.}
+begin with an overview of \tool through an example in
+Section~\ref{sec:ex}. %\nc{provided a pointer to sec 2 as well.}
+%\aseem{Sure :), I did not add it because Sec. 2 was just next line, so
+%  I thought it did not need a pointer.}
 

### 2017-10-28 OaklandSubmission/intro.tex
minor

@@
 cryptographic technique that allows two mutually distrusting parties
-to compute a joint function of their secret inputs in a way that both
+to compute a publicly known joint function of their secret inputs in a way that both
 the parties learn nothing about the inputs of each other beyond what
@@
 developers, there are fewer avenues for making mistakes, developers
-don't bear the burden of getting cryptographic details right, and it
+don't bear the burden of getting cryptographic details right, 
+code optimizations can be left to the compiler,
+and it
 is easy to maintain and modify the programs. Needless to say,

### 2017-10-28 OaklandSubmission/main.tex
minor

--- 2017-10-27 OaklandSubmission/main.tex
@@
 \maketitle
-% \begin{abstract}
-% The abstract goes here.
-% \end{abstract}
+\begin{abstract}
+Secure two-party computation (\mpc) allows two mutually distrusting parties to compute any function on their joint inputs, without revealing anything other than the output of the function to one another.
+%While the application and potential impact of secure computation is virtually limitless, writing an efficient and scalable secure protocol is by no means an easy feat.
+ Prior work in designing such protocols, suffered from either a) a lack of programmability (e.g.
+requiring the programmer to write cryptography-aware circuits for the computation); b) a lack of efficiency for several natural computations (such as machine learning classification tasks); or c) required cryptographers to design hand-crafted protocols for various applications.
+In this work, we present \tool: a new ``cryptographic-cost aware'' \mpc framework that generates efficient and scalable \mpc protocols from high-level programs that
+%contain absolutely no cryptographic details. Programs in our language 
+are easy-to-write, with lines of code that are comparable to their insecure implementations.
+Furthermore, the \mpc protocols generated by our framework match or outperform recent works that provide hand-crafted \mpc protocols for various functionalities such as classification and matrix factorization (up to a 19x factor!).
+\end{abstract}
 \IEEEpeerreviewmaketitle

### 2017-10-28 OaklandSubmission/intro.tex
nit

@@
 about the client's input or output, and that the client learns nothing
-about the server model, beyond what is revelaed by the output
+about the server classifier, beyond what is revelaed by the output
 label.
@@
 compute $w^Tx >b$. Here $w$ (a vector) and $b$ (a scalar) constitute
-the server model, and $x$ is the client's input vector. Further,
+the server classifier, and $x$ is the client's input vector. Further,
 $\cdot^{T}$ is the matrix transpose operator, and $w^Tx$ denotes the
@@
 for i in [0:30] { acc =  acc + (w[i] * x[i]); }
-output((acc > b) ? 1 : 0)
+output((acc > b) ? 1 : 0, 2) //only to party 2
 \end{lstlisting}

### 2017-10-28 OaklandSubmission/eval.tex
table one column
removed reference to and , mult gates

@@
  % right?} \nc{Actually, they have a specific protocols for dot product and comparison and also say that their developer effort is minimal as they have to write two function calls to these protocols :).}
- Our execution time is dominated by the multiplications and
-boolean-ands.
-The number of boolean-and gates (\#And) remains constant for both the
-benchmarks as they reflect the number of gates required for a
-single  comparison, while the number of
-multiplication gates (\#Mul) is the same as the model size $d$.
+% Our execution time is dominated by the multiplications and
+%boolean-ands.
+%The number of boolean-and gates (\#And) remains constant for both the
+%benchmarks as they reflect the number of gates required for a
+%single  comparison, while the number of
+%multiplication gates (\#Mul) is the same as the model size $d$.
 %% The
@@
 %% independent of the model size.
+
+%\setlength\tabcolsep{7pt}
+%\begin{table*}
+%\begin{tabular}{c|c|c|c|c |c|c|c|c|c|c}
+%Dataset & $d$  & Prev. time (s) & Prev. Comm (kb) & LAN (s) & WAN (s) & Comm. (kb)  & \#And & \#Mul & \#Gates & LOC\\
+%\hline
+%Breast cancer & 30 & 0.3 & 36 & 0.1 & 0.3 & 25 & 95 & 30 & 727 & 20\\
+%\hline
+%Credit & 47 & 0.3 & 41 & 0.1 & 0.3 & 36 & 95 & 47 & 795 & 20\\
+%\hline
+%\end{tabular}
+
+\setlength\tabcolsep{2.5pt}
+\begin{table}
+\begin{tabular}{|c|c|c|c|c|c|c|c|c|}
+\hline
+Dataset & $d$  & \thead{Prev. \\ time\\ (s)} & \thead{Prev. \\ Comm\\ (kb)} & \thead{LAN  \\ (s)} & \thead{WAN\\ (s)} & \thead{Comm. \\(kb)}  &  \#Gates & LOC\\
+\hline
+Breast cancer & 30 & 0.3 & 36 & 0.1 & 0.3 & 25  & 727 & 20\\
+\hline
+Credit & 47 & 0.3 & 41 & 0.1 & 0.3 & 36  & 795 & 20\\
+\hline
+\end{tabular}
+ \caption{Linear classification results. We compare our results (columns 5, 6, 7) with~\cite{shafindss} (columns 3 and 4)}
+ \label{tab:lc} 
+\end{table}
 
@@
 
+\setlength\tabcolsep{2.5pt}
+\begin{table}
+\begin{tabular}{|c|c|c|c|c|c |c|c|c|c|}
+\hline
+Dataset & $d$ & $F$ & \thead{Prev. \\ time\\ (s)} & \thead{Prev. \\ Comm\\ (kb)} & \thead{LAN  \\ (s)} & \thead{WAN\\ (s)} & \thead{Comm. \\(kb)} & \#Gates  & LOC\\
+\hline
+Nursery & 5 & 9 & 1.5 & 0.2 & 0.1 & 0.4 & 0.6  & 73k & 50\\
+\hline
+Audiology & 24 & 70 & 3.9 & 2.0 & 1.5 & 2.9 & 37  & 4219k & 50\\
+\hline
+\end{tabular}
+\caption{Na\"{i}ve Bayes results. We compare our results (columns 6 , 7, 8) with~\cite{shafindss} (columns 4 and 5)}
+ \label{tab:nb} 
+\end{table}
+
 \subsubsection*{Decision trees}
@@
 code is competitive with specialized protocols.
-\setlength\tabcolsep{8pt}
-\begin{table*}
-\begin{tabular}{c|c|c|c|c |c|c|c|c|c|c}
-Dataset & $d$  & Prev. time (s) & Prev. Comm (kb) & LAN (s) & WAN (s) & Comm. (kb)  & \#And & \#Mul & \#Gates & LOC\\
+
+%\begin{table*}
+%\begin{tabular}{|c|c|c|c |c|c|c|c|c|c|c|}
+%\hline
+%Dataset  & $N$ & \thead{Prev. \\ time\\ (s)} & \thead{Prev. \\ Comm\\ (kb)} & \thead{LAN  \\ (s)} & \thead{WAN\\ (s)} & \thead{Comm. \\(kb)}& \#And & \#Mul &\#Gates & LOC\\
+%\hline
+%Nursery & 4 & 0.3 & 102 & 0.1 & 0.3 & 32 & 504 & 3 & 3324 & 20\\
+%\hline
+%ECG &  6 & 0.4 & 102 & 0.1 & 0.4 & 49 & 756 & 5 & 5002 & 20\\
+%\hline
+%\end{tabular}
+% \caption{Decision tree benchmarks. We compare our results (columns 5 , 6, 7) with~\cite{wu} (columns 3 and 4)}
+% \label{tab:dt} 
+%\end{table*}
+
+\begin{table}
+\begin{tabular}{|c|c|c|c|c|c|c|c|c|}
 \hline
-Breast cancer & 30 & 0.3 & 36 & 0.1 & 0.3 & 25 & 95 & 30 & 727 & 20\\
+Dataset  & $N$ & \thead{Prev. \\ time\\ (s)} & \thead{Prev. \\ Comm\\ (kb)} & \thead{LAN  \\ (s)} & \thead{WAN\\ (s)} & \thead{Comm. \\(kb)}& \#Gates & LOC\\
 \hline
-Credit & 47 & 0.3 & 41 & 0.1 & 0.3 & 36 & 95 & 47 & 795 & 20\\
+Nursery & 4 & 0.3 & 102 & 0.1 & 0.3 & 32 & 3324 & 20\\
+\hline
+ECG &  6 & 0.4 & 102 & 0.1 & 0.4 & 49  & 5002 & 20\\
 \hline
 \end{tabular}
-
- \caption{Linear classification results. We compare our results (columns 5, 6, 7) with~\cite{shafindss} (columns 3 and 4)}
- \label{tab:lc} 
-\end{table*}
-\setlength\tabcolsep{7pt}
-\begin{table*}
-\begin{tabular}{c|c|c|c|c|c |c|c|c|c|c|c}
-Dataset & $d$ & $F$ & Prev. time (s) & Prev. Comm (Mb) & LAN (s) & WAN (s) & Comm. (Mb)  & \#And & \#Mul & \#Gates  & LOC\\
-\hline
-Nursery & 5 & 9 & 1.5 & 0.2 & 0.1 & 0.4 & 0.6 & 13k & 0 & 73k & 50\\
-\hline
-Audiology & 24 & 70 & 3.9 & 2.0 & 1.5 & 2.9 & 37 & 750k & 0 & 4219k & 50\\
-\hline
-\end{tabular}
-
- \caption{Na\"{i}ve Bayes results. We compare our results (columns 6 , 7, 8) with~\cite{shafindss} (columns 4 and 5)}
- \label{tab:nb} 
-\end{table*}
-
-\begin{table*}
-\begin{tabular}{c|c|c|c |c|c|c|c|c|c|c}
-Dataset  & $N$ & Prev. time (s) & Prev. Comm (kb) & LAN time (s) & WAN time (s) & Comm. (kb)  & \#And & \#Mul & \#Gates & LOC\\
-\hline
-Nursery & 4 & 0.3 & 102 & 0.1 & 0.3 & 32 & 504 & 3 & 3324 & 20\\
-\hline
-ECG &  6 & 0.4 & 102 & 0.1 & 0.4 & 49 & 756 & 5 & 5002 & 20\\
-\hline
-\end{tabular}
-
 \caption{Decision tree benchmarks. We compare our results (columns 5 , 6, 7) with~\cite{wu} (columns 3 and 4)}
  \label{tab:dt} 
-\end{table*}
+\end{table}
 
@@
 %net | minionT | minionC | timeL | timeW | comm | gates | and | mul | length
-\setlength\tabcolsep{6.5pt}
-\begin{table*}
-\begin{tabular}{c|c|c|c |c|c|c|c|c|c|c}
-DNN  & Prev. time (s) & Prev. Comm (Mb) & LAN (s) & WAN (s) & Comm. (Mb)  & \#And & \#Mul & \#Gates & Model Size & LOC\\
+
+%\setlength\tabcolsep{6.5pt}
+%\begin{table*}
+%\begin{tabular}{c|c|c|c |c|c|c|c|c|c|c}
+%DNN  & Prev. time (s) & Prev. Comm (Mb) & LAN (s) & WAN (s) & Comm. (Mb)  & \#And & \#Mul & \#Gates & Model Size & LOC\\
+%\hline
+%SecureML   &  1.1 & 15.8 & 0.7 & 1.7  & 76   &  2k   & 119k & 366k   & 119k & 78\\
+%\hline
+%Cryptonets &  1.3 & 47.6 & 0.6 & 1.6  & 70    & 2k    & 108k & 316k & 86k & 88\\
+%\hline
+%CNN        &  9.4 & 657.5& 5.1 & 11.6 & 501  & 1640k & 667k & 9480k & 35k & 154\\
+%\hline
+%\end{tabular}
+%
+% \caption{DNN benchmarks. We compare our results (columns 4 , 5, 6) with~\cite{minionn} (columns 2 and 3)}
+% \label{tab:nn} 
+%\end{table*}
+
+\setlength\tabcolsep{2.5pt}
+\begin{table}
+\begin{tabular}{|c|c|c|c |c|c|c|c|c|}
 \hline
-SecureML   &  1.1 & 15.8 & 0.7 & 1.7  & 76   &  2k   & 119k & 366k   & 119k & 78\\
+DNN  & \thead{Prev. \\ time\\ (s)} & \thead{Prev. \\ Comm\\ (kb)} & \thead{LAN  \\ (s)} & \thead{WAN\\ (s)} & \thead{Comm. \\(kb)} & \#Gates & \thead{Model\\ Size} & LOC\\
 \hline
-Cryptonets &  1.3 & 47.6 & 0.6 & 1.6  & 70    & 2k    & 108k & 316k & 86k & 88\\
+SecureML   &  1.1 & 15.8 & 0.7 & 1.7  & 76   &   366k   & 119k & 78\\
 \hline
-CNN        &  9.4 & 657.5& 5.1 & 11.6 & 501  & 1640k & 667k & 9480k & 35k & 154\\
+Cryptonets &  1.3 & 47.6 & 0.6 & 1.6  & 70    &  316k & 86k & 88\\
+\hline
+CNN        &  9.4 & 657.5& 5.1 & 11.6 & 501  & 9480k & 35k & 154\\
 \hline
@@
  \label{tab:nn} 
-\end{table*}
+\end{table}
 
@@
 \begin{table}
-\begin{tabular}{c|c|c|c |c|c|c|c|c|c | c}
+\begin{tabular}{|c|c|c|c |c|c|c|c|c|c | c}
+\hline
 Classifier       & \thead{LAN  \\ (s)} & \thead{WAN \\(s)} & \thead{Comm. \\(Mb)}  & \#And & \#Mul & \#Gates & \thead{Model \\size} & LOC\\
@@
 \begin{table}
-\begin{tabular}{c|c|c|c |c|c|c|c|c|c | c}
+\begin{tabular}{|c|c|c|c |c|c|c|c|c|c | c|}
+\hline
 Dataset       & \thead{LAN \\(s)} & \thead{WAN \\(s)} & \thead{Comm. \\ (Mb)}  & \#And & \#Mul & \#Gates & depth & LOC\\
@@
 \begin{table}
-\begin{tabular}{c|c|c|c |c|c|c}
+\begin{tabular}{|c|c|c|c |c|c|c|}
+\hline
            &  LAN (s) & WAN (s) & \thead{Comm. \\ (Mb)}  & \#And & \#Mul & \#Gates \\

### 2017-10-28 OaklandSubmission/ex.tex
intro
toolchain figure
wip in S2

--- 2017-10-12 OaklandSubmission/ex.tex
@@
-\section{Motivating example}
+\section{\tool Overview}
 \label{sec:ex}
 %$w^Tx>0$, mnist logistic regression
-Consider a  cloud service provider Bob that wants to provide a service to diagnose
-whether a patient has breast cancer or not. Bob trains a machine learning classification model
-that results in a vector $w$ (say of length 30) and a scalar $b$.
-This model is the intellectual property of Bob and he  wants to keep it secret.
-Given a patient's medical report, in the form of a vector $x$ (also of length 30),
-the classifer predicts that the patient has breast cancer if $w^Tx>b$.
-However, for this task Bob needs access to $x$, which is private data of a customer.
+%% \divya{add reference to before and later... as mentioned we would run evaluations on secure prediction of different models. Here we describe scenario in detail..}
 
-A potential customer Alice might not want to reveal $x$ to Bob because of privacy concerns.
-And Bob does not want to reveal $w$ and $b$ because then Alice can steal the model.
-SMC can help Alice and Bob compute $w^Tx>b$ securely, such that Alice receives the classifier's
-prediction and Bob learns nothing about $x$. Moreover, Alice does not learn anything more about $w$
-and $b$ than what is revealed by the prediction. 
-\nc{This is a bit repetitive from the intro, isnt it? Should we have another example perhaps?}
-To implement this system, Bob can write the code in \tool and this code is shown in Figure~\ref{fig:ex-sml}.
-The expression {\tt input1} reads a value from Bob and {\tt input2} reads from Alice.
-The language has arrays and simple loops. A loop {\tt for i = [0:N]} repeats its
-body {\tt N} times, the loop counter {\tt i} is assigned {\tt 0} in the first iteration,
-and is incremented by one after each iteration. Unbounded {\tt while} loops are problematic
-from a cryptographic standpoint and our language only permits these simple {\tt for} loops. 
-The first loop in Figure~\ref{fig:ex-sml} reads the model $w$ from Bob, the second loop reads
-Alice's medical report $x$, and the third loop computes the dot product $w^Tx$.
-The last {\tt output} statement sends the result of the comparison $w^Tx>b$ to Alice.
-In particular, the ternary ``{\tt ? :}" operator performs a branch and the result is the second (third)
-argument if the first argument is true (false).
+\begin{figure}
+  %\vspace{-20pt}
+  %\begin{center}
+  \includegraphics[width=0.45\textwidth]{toolchain}
+  %\end{center}
+  %\vspace{-20pt}
+\caption{\tool toolchain}
+\label{fig:toolchain}
+\end{figure}
 
-The \tool compiler is fully automatic and compiles the code described in Figure~\ref{fig:ex-sml} to the C++ code that makes calls to the ABY library  in
-Figure~\ref{fig:ex-aby}. An alternative to using \tool is to directly write this C++ code.
-However, this code is much more complex than the implementation written in \tool.
-Therefore, it is difficult for Alice and bob to verify its correctness and security.
-Unlike \tool implementations, the ``secret" variables containing private data ({\tt w,x,b,acc}) need to be handled differently compared to ``public" variables such as loop counters ({\tt i}).
+Figure~\ref{fig:toolchain} shows an overview of the \tool
+toolchain. We give a brief overview of each of these phases below.
+
+\subsubsection*{Source language}
+Consider the example $w^Tx >b$ from Section~\ref{sec:intro}, where
+$w$ and $b$ constitute the server's input (a classifier) and $x$ is
+the client's input vector. Figure~\ref{fig:ex-sml} shows \tool code
+for this example. The code first reads the inputs of the two parties
+using the \ls{input1} and \ls{input2} expressions. It then uses a
+\ls{for} loop to compute \ls{acc}, the dot product of \ls{w} and
+\ls{x}. Finally, the code outputs the result of comparing \ls{acc}
+with \ls{b} only to the client.
+
+\tool source language is a simple, imperative language that enables
+the programmers to express \mpc computations in terms of their
+``ideal'' functionalities, without dealing with any cryptographic
+details. The languages provides multi-dimensional arrays, conditional
+expressions (the ternary $?\::$ operator), \ls{for}~loops,
+\ls{if}~statements, and special syntax for input/output.
+
+\input{fig-ex-aby}
+
+\subsubsection*{\tool compiler}
+\tool compiler takes as input a source program and produces a C++
+program as output. Figure~\ref{fig:ex-aby} shows the output code for
+the example in Figure~\ref{fig:ex-sml}. The output program contains
+contains party-specific code for inputs and outputs
+(\ls{role == SERVER} and \ls{role == CLIENT}), and common code for the
+computation.
+
+The compiler splits the input
+program into \emph{public} and \emph{secret} components. The public
+components translate into regular C++ code, while the secret
+components translate into API calls into our crypto back-end
+(ABY). For instance, for the code in Figure~\ref{fig:ex-sml}, the \tool
+compiler realizes that the array index \ls{i} in the dot product loop
+is public, and hence the array accesses need not be compiled
+obliviously. Therefore, it compiles the \ls{for}~loop into a C++
+\ls{for}~loop that will be executed in-clear
+(line~\ref{line:dotproductloop}).
+
+Within the secret components, the \tool compiler is ``cryptographic
+cost-aware'', and appropriately picks either arithmetic or boolean
+circuit representations for different sub-componenets. For example,
+the compiler realizes that the dot product computation is more
+efficient in the arithmetic representation, and therefore it builds
+the corresponding circuit using the arithmetic circuit builder
+\ls{acirc} (lines~\ref{line:dotmulgate} and~\ref{line:dotaddgate}). On
+the other hand, since the comparison with \ls{b}, and the conditional
+expression computation are more efficient in the boolean
+representation, the \tool compiler uses the Yao circuit builder
+\ls{ycirc} to build the corresponding circuits
+(lines~\ref{line:condyaobegin} to~\ref{line:condyaoend})\footnote{As
+  we mention in
+  Section~\ref{sec:impl}, in our usage of ABY, boolean is synonymous
+  with Yao.}.
+
+Importantly, conversions between the arithmetic and boolean
+represenation require share-conversions. The \tool compiler also
+instruments these conversion gates accordingly. For example, in
+lines~\ref{line:convacc} and~\ref{line:convb}, the compiler converts
+\ls{a_acc} and \ls{a_b} to boolean representation, before they are
+input to the comparison and multiplexer circuits.
+
+\subsubsection*{Circuit generation and evaluation} The next step in
+\tool is to
+compile the output C++ code and execute it. Doing so reduces away the
+public parts of the program, including the array accesses, and
+generates a \mpc circuit comprising of arithmetic and boolean gates,
+with appropriate conversion gates. The circuit is then evaluated using
+a \mpc protocol.
+
+%% Consider a  cloud service provider Bob that wants to provide a service to diagnose
+%% whether a patient has breast cancer or not. Bob trains a machine learning classification model
+%% that results in a vector $w$ (say of length 30) and a scalar $b$.
+%% This model is the intellectual property of Bob and he  wants to keep it secret.
+%% Given a patient's medical report, in the form of a vector $x$ (also of length 30),
+%% the classifer predicts that the patient has breast cancer if $w^Tx>b$.
+%% However, for this task Bob needs access to $x$, which is private data of a customer.
+
+%% A potential customer Alice might not want to reveal $x$ to Bob because of privacy concerns.
+%% And Bob does not want to reveal $w$ and $b$ because then Alice can steal the model.
+\divya{Use HIPAA compliance; as it can leak information of patients that was used in training; more compelling argument.
+See Shafi intro.}
+%% SMC can help Alice and Bob compute $w^Tx>b$ securely, such that Alice receives the classifier's
+%% prediction and Bob learns nothing about $x$. Moreover, Alice does not learn anything more about $w$
+%% and $b$ than what is revealed by the prediction. 
+%% \nc{This is a bit repetitive from the intro, isnt it? Should we have another example perhaps?}
+%% To implement this system, Bob can write the code in \tool and this code is shown in Figure~\ref{fig:ex-sml}.
+%% The expression {\tt input1} reads a value from Bob and {\tt input2} reads from Alice.
+%% The language has arrays and simple loops. A loop {\tt for i = [0:N]} repeats its
+%% body {\tt N} times, the loop counter {\tt i} is assigned {\tt 0} in the first iteration,
+%% and is incremented by one after each iteration. Unbounded {\tt while} loops are problematic
+%% from a cryptographic standpoint and our language only permits these simple {\tt for} loops. 
+%% The first loop in Figure~\ref{fig:ex-sml} reads the model $w$ from Bob, the second loop reads
+%% Alice's medical report $x$, and the third loop computes the dot product $w^Tx$.
+%% The last {\tt output} statement sends the result of the comparison $w^Tx>b$ to Alice.
+%% In particular, the ternary ``{\tt ? :}" operator performs a branch and the result is the second (third)
+%% argument if the first argument is true (false).
+
+%% The \tool compiler is fully automatic and compiles the code described in Figure~\ref{fig:ex-sml} to the C++ code that makes calls to the ABY library  in
+%% Figure~\ref{fig:ex-aby}. An alternative to using \tool is to directly write this C++ code.
+%% However, this code is much more complex than the implementation written in \tool.
+%% Therefore, it is difficult for Alice and bob to verify its correctness and security.
+%% Unlike \tool implementations, the ``secret" variables containing private data ({\tt w,x,b,acc}) need to be handled differently compared to ``public" variables such as loop counters ({\tt i}).
 In particular, the former are manipulated by the ABY library and the latter are manipulated like standard  C++ variables.
-Additionally, the C++ code branches on  the special variable
-{\tt role}  to decide whether the enclosing code is executed by the {\tt SERVER} Bob or by
-the {\tt CLIENT} Alice. In Figure~\ref{fig:ex-sml}, the developer does not need to
-manipulate {\tt role} explicitly. In addition to simplifying the implementation, \tool prevents developers from writing buggy code such as\\
-\verb+if(role == SERVER) {...manipulate x...}+
+%% Additionally, the C++ code branches on  the special variable
+%% {\tt role}  to decide whether the enclosing code is executed by the {\tt SERVER} Bob or by
+%% the {\tt CLIENT} Alice. In Figure~\ref{fig:ex-sml}, the developer does not need to
+%% manipulate {\tt role} explicitly. In addition to simplifying the implementation, \tool prevents developers from writing buggy code such as\\
+%% \verb+if(role == SERVER) {...manipulate x...}+
 
-Furthermore, the code in Figure~\ref{fig:ex-aby} is too low level.
-In particular, it is first builds a circuit that represents the computation
-$w^Tx>b$ and then executes it via a call to {\tt ExecCircuit}.
-To build a circuit, the code adds multiplication gates ({\tt PutMULGate})
-and addition gates ({\tt PutADDGate}) for the dot product.
-Subsequently, it uses a ``greater than" gate ({\tt PutGTGate})
-and a multiplexer gate ({\tt PutMUXGate}) for the comparison with $b$.
-For efficiency, the operations for dot product need to be performed using 
-arithmetic circuits. However, ABY's arithmetic circuits cannot express
-comparisons and these need  boolean circuits.
-A program that uses both arithmetic circuit and boolean circuits
-requires conversion gates that help  interconvert between arithmetic
-representations and boolean representations ({\tt PutA2YGate}).
-Bigger computations often require multiple conversions and the
-developer effort can quickly become significant.
+%% Furthermore, the code in Figure~\ref{fig:ex-aby} is too low level.
+%% In particular, it is first builds a circuit that represents the computation
+%% $w^Tx>b$ and then executes it via a call to {\tt ExecCircuit}.
+%% To build a circuit, the code adds multiplication gates ({\tt PutMULGate})
+%% and addition gates ({\tt PutADDGate}) for the dot product.
+%% Subsequently, it uses a ``greater than" gate ({\tt PutGTGate})
+%% and a multiplexer gate ({\tt PutMUXGate}) for the comparison with $b$.
+%% For efficiency, the operations for dot product need to be performed using 
+%% arithmetic circuits. However, ABY's arithmetic circuits cannot express
+%% comparisons and these need  boolean circuits.
+%% A program that uses both arithmetic circuit and boolean circuits
+%% requires conversion gates that help  interconvert between arithmetic
+%% representations and boolean representations ({\tt PutA2YGate}).
+%% Bigger computations often require multiple conversions and the
+%% developer effort can quickly become significant.
 
-Moreover, it is difficult to maintain the code in Figure~\ref{fig:ex-aby}.
-For example, if the multiplication is changed to a bitwise-or then,
-in the absence of \tool, an efficiency-conscious developer would need to  use boolean circuits everywhere and would be tasked with removing all arithmetic circuits and the interconversion gates (the cost of  conversion overweighs the gains of performing an arithmetic addition instead of a boolean addition).
- With \tool, the developer needs to change only a single character (replace {\tt *} in Figure~\ref{fig:ex-sml} with {\tt |}) to achieve the same effect. 
-On this example and other small examples, we have found that the compiler generated C++ code from a \tool implementation is as efficient as manually written C++ code.
+%% Moreover, it is difficult to maintain the code in Figure~\ref{fig:ex-aby}.
+%% For example, if the multiplication is changed to a bitwise-or then,
+%% in the absence of \tool, an efficiency-conscious developer would need to  use boolean circuits everywhere and would be tasked with removing all arithmetic circuits and the interconversion gates (the cost of  conversion overweighs the gains of performing an arithmetic addition instead of a boolean addition).
+%%  With \tool, the developer needs to change only a single character (replace {\tt *} in Figure~\ref{fig:ex-sml} with {\tt |}) to achieve the same effect. 
+%% On this example and other small examples, we have found that the compiler generated C++ code from a \tool implementation is as efficient as manually written C++ code.
 In particular, for $w^Tx>b$, the compiler automatically generates code that uses arithmetic circuits for dot product and inserts appropriate conversion gates.
 Finally, although this paper uses ABY for evaluation, the  compiler can be easily retargeted to generate code for other cryptography backends.
-\nc{I like the above point. Perhaps we can mention something like this earlier in the intro itself?}
 
-The \tool compiler provides strong static guarantees. For example, a well typed
-\tool program is guaranteed to execute to termination without errors.
+%% The \tool compiler provides strong static guarantees. For example, a well typed
+%% \tool program is guaranteed to execute to termination without errors.
 A \tool program cannot go into non-termination or dereference illegal memory, i.e., no buffer overflows or underflows can happen at run time.
-These guarantees are
-much stronger than what are offered by production compilers such as {\tt gcc}.
-The primary reason that we are able to achieve these guarantees is because
-\tool has been designed to be verifiable and is backed by formal semantics. 
-Furthermore, the compiler output is guaranteed to be a cryptographically secure
-implementation of the functionality declared by the \tool code. 
-Moreover, the \tool compiler also generates a C implementation that can be run natively on a single machine for functional testing.
+Production compilers such as {\tt gcc} provide no such guarantees
+%% for the code in Figure~\ref{fig:ex-aby}.
+%% The primary reason that we are able to achieve these guarantees is because
+%% \tool has been designed to be verifiable and is backed by formal semantics. 
+%% Furthermore, the compiler output is guaranteed to be a cryptographically secure
+%% implementation of the functionality declared by the \tool code. 
+%% Moreover, the \tool compiler also generates a C implementation that can be run natively on a single machine for functional testing.
 
-To summarize, \tool raises the level of abstraction, provides strong static guarantees, and generates efficient code automatically.
+%% To summarize, \tool raises the level of abstraction, provides strong static guarantees, and generates efficient code automatically.

### 2017-10-28 OaklandSubmission/security.tex

@@
 Combining these two statements, the proof of the theorem follows.
-\aseem{
-  Also, we should emaphasize here that the ideal semantics is the
-  source semantics from Figure x and the real semantics is the circuit
-  semantics from Figure y. For the target language semantics, we
-  should emphasize that it does not have access to the secrets, so
-  there are no observations there.
-}
-
-\nc{OK now?}
   %% \\

### 2017-10-28 OaklandSubmission/related.tex
related

@@
 
+\tool falls into the category of frameworks that compile high level languages to \mpc protocols. We discuss other such frameworks next.
+%Existing \mpc tools work with either a high or a low level desciption of the functionalities to be implemented securely.
  %\end{enumerate}
-\noindent\textbf{Generating \mpc protocols.}
-There has been vast literature on the generation of \mpc protocols.
+%\noindent\textbf{Generating \mpc protocols.}
+%There has been vast literature on the generation of \mpc protocols.
 These works all provide tools at various levels of abstraction that help a programmer generate cryptographic protocols for various functionalities.
 These levels can be classified as follows:
-\\\\
-\noindent{\bf Function Description:} Tools at this level are a language (similar to \tool) where the programmer can specify the functionality to be securely implemented in a high level language and the compiler compiles the functionality to generate a secure computation protocol.
-Such languages do not require the programmer to have expertise in writing cryptographic protocols.
-All such existing languages in literature are built with only one backend: Fairplay's Secure Function Definition Language (SFDL)~\cite{fairplay,fairplaymp} compiles high-level C or Pascal like programs into boolean circuits that are then evaluated using garbled circuits~\cite{yao} for the two-party case and distributed garbled circuits~\cite{bmr} for the multi-party case~\cite{bmr}.
-The Secure Multiparty Computation Language (SMCL)~\cite{smcl} that compiles high-level Java like programs into arithmetic circuits that are then evaluated using the VIFF framework~\cite{viff} is another example.
-The work of Wysteria~\cite{wysteria} provides a way of writing ``mixed-mode'' programs that combine normal or per-party, local computations, with ``secure'' (joint) computations repeatedly in a seamless manner.
-This work compiles programs into boolean circuits and uses Choi {\em et al.}~\cite{choi} implementation of the GMW~\cite{gmw} protocol as the backend.
-The work of Mitchell {\em et al.}~\cite{lambdaps} compiles their language into either a backend that works over arithmetic secret shares or over boolean secret shares using homomorphic encryption~\cite{gentry}.
-Finally, the work of ObliVM~\cite{oblivm} also shows how to incorporate programs that may read/write portions of the secret data, based on other secrets, without revealing access patterns (using techniques from oblivious RAM~\cite{oram1,oram2}).
-Their compute is through boolean circuits and the garbling technique.
-It is to be noted that none of the above works combine arithmetic and boolean computation together within the same program.
-\\\\
-\noindent{\bf Protocol Description:} Tools at this level of abstraction provide a set of commands that can be put together to describe efficient secure computation protocols.
-These tools are at a high level (such as making a call to execute secure computation of an AND gate or a MULT gate over shared inputs).
-The exact cryptographic protocols that are used as backends to implement these secure computation protocols need not be known to the programmer, but in order to generate efficient protocols for the entire functionality, the programmer must know how the functionality will be broken into components that make calls to various secure computation protocols for sub-functions.
-As an example, the work of ABY \cite{aby} falls into this category.
-While, the programmer need not know how ABY implements boolean or arithmetic computation, he or she must know the type of compute (boolean or arithmetic) that is more efficient for specific sub-computations and also explicitly write calls to convert shares from one scheme to another (boolean secret sharing to arithmetic secret sharing and vice-versa).
-The work of Kerschbaum et al.~\cite{autoS} provides a scheme to automatically assign homomorphic encryption or garbled circuits to each operators in a computation that is expressed as a sequence of dyadic operations.
+%Examples of tools in the former category include the following: 
+Fairplay's Secure Function Definition Language (SFDL)~\cite{fairplay,fairplaymp} and CBMC-GC~\cite{cbmcgc} compile C or Pascal like programs into boolean circuits that are then evaluated using garbled circuits~\cite{yao}.
+ObliVM~\cite{oblivm} protects access patterns using an oblivious RAM~\cite{oram1,oram2} and also uses garbled circuits for compute.
+In Secure Multiparty Computation Language (SMCL)~\cite{smcl},  Java like programs are compiled into arithmetic circuits that are then evaluated using the VIFF framework~\cite{viff}.
+Wysteria~\cite{wysteria} compiles  programs into boolean circuits and uses a GMW-based backend~\cite{choi,gmw}.
+ Mitchell {\em et al.}~\cite{lambdaps} allow the user to select between Shamir's secret sharing~\cite{sss} and fully homomorphic encryption~\cite{gentry}.
+All these tools use either a arithmetic backend or a boolean backend and do not require share conversions.
+
+Next, we discuss tools that expose libraries which  developers can use to describe \mpc protocols. 
+% The exact cryptographic protocols that are used by the library  need not be known to the programmer. However, in order 
+To generate efficient protocols for a functionality, the programmer must break the functionality into components and call the appropriate library functions.
+For example,  ABY \cite{aby} falls in this category.
+The TASTY tool~\cite{tasty} allows mixing homomorphic encryption based arithmetic computations
+and garbled circuits based boolean computations and the interconversions between the two are inserted by the programmer explicitly.
+% While, the programmer need not know how ABY implements boolean or arithmetic computation, he or she must know the type of compute (boolean or arithmetic) that is more efficient for specific sub-computations and also explicitly write calls to convert shares from one scheme to another (boolean secret sharing to arithmetic secret sharing and vice-versa).
+The work of Kerschbaum et al.~\cite{autoS} provides a scheme to automatically assign homomorphic encryption or garbled circuits to each operator in a computation that is expressed as a sequence of dyadic operations.
 They conjecture that the problem is NP hard and gave a linear programming based solution
-as well as a quadratic time greedy heuristic.
-As noted by ABY~\cite{aby}, the performance is much poorer than tools like ABY that combine boolean and arithmetic secure computation techniques due to very high conversion costs between homomorphic encryption and boolean shares as well as the cost of homomorphic encryption itself as the security parameter increases.
-This abstraction requires a fairly involved knowledge of cryptography from the programmer.
+as well as a quadratic time greedy heuristic. Their performance is much poorer than ABY due to the high 
+costs associated with homomorphic encryption~\cite{aby}.
+%conversion costs between homomorphic encryption and boolean shares~\cite{aby}.
 %On the other hand \tool uses a simpler constant time heuristic to assign arithmetic secret sharing or garbled circuits to sub-computations in a richer language with control flow statements such as loops and branches.
@@
 Similarly, the TASTY tool~\cite{tasty} provides a way to mix both boolean and arithmetic circuits (using homomorphic encryption for the arithmetic compute).
-Other examples include the VIFF framework~\cite{viff} for arithmetic computations, Sharemind~\cite{sharemind} (that provides information-theoretic secure 3-party boolean computation), and the Secure Multiparty Computation language~\cite{nsy04,securemc}.
-\\\\
-\noindent{\bf Protocol Implementation:} Tools at this level of abstraction require the programmer to have a deep understanding of how cryptographic secure computation protocols work and the programmer must provide calls to low-level cryptographic building blocks such as oblivious transfer.
+Other examples include the VIFF framework~\cite{viff} for arithmetic computations and Sharemind~\cite{sharemind} (that provides secure 3-party boolean computation). 
+%\\\\
+%\noindent{\bf Protocol Implementation:} Tools at this level of abstraction require the programmer to have a deep understanding of how cryptographic secure computation protocols work and the programmer must provide calls to low-level cryptographic building blocks such as oblivious transfer.
 For example while the programmer need not know the specific oblivious transfer protocol implemented in the backend, he or she still needs to know how to build a secure computation protocol from such building blocks.
@@
 
-Other low level implementations of secure computation protocols look to optimize circuits for depth (such as~\cite{ddkssz15}) or provide protocols for MIPS code~\cite{mips} using techniques from ORAM to hide access patterns~\cite{oram1,oram2}.
-\\\\
-\noindent\textbf{Custom \mpc protocols.} Many works have designed specialized, hand-crafted protocols for various secure two-party computation tasks.
-The reason for doing this was to allow for a mix between arithmetic and boolean computations as a secure computation protocol for a pure boolean circuit would have been very inefficient.
-Such hand-crafting of protocols require deep cryptography know-how of designing secure computation protocols with various tools such as homomorphic encryption, garbled circuits, oblivious transfer and so on.
-A sample of such work includes various works on privacy-preserving classification of medical ElectroCardioGram (ECG)~\cite{barni}, iris and fingerprint recognition~\cite{blanton}, remote diagnostics~\cite{brickell}, private sequence analysis~\cite{franz}, private biometric identification~\cite{huang}, privacy-preserving matrix factorization and ridge-regression~\cite{valeriaMatrix, valeriaRidge}, machine-learning classification~\cite{shafindss}, decision trees and forests~\cite{wu}, neural network training and prediction~\cite{secureml,minionn}.
-As seen from our experiments, the protocols generated by \tool give us better or comparable execution times as many of the above works with minimal programmer overhead.
+The implementations of \mpc backends  have made  tremendous progress in the last decade.
+For example, the circuits can be optimized for depth~\cite{ddkssz15,cbmcgcdepth}, large garbled circuits can be pipelined~\cite{yao-pipe}, and oblivious RAM~\cite{oram1,oram2} can be used to hide access patterns of MIPS code~\cite{mips}.
+
+%\\\\
+%\noindent\textbf{Custom \mpc protocols.} 
+Many works have designed specialized, hand-crafted protocols for various \mpc tasks. 
+%The reason for doing this was to allow for a mix between arithmetic and boolean computations as a secure computation protocol for a pure boolean circuit would have been very inefficient. 
+This process require deep knowledge of cryptography
+to ensure that the protocols are secure.
+Examples include privacy-preserving classification of medical ElectroCardioGram (ECG)~\cite{barni}, iris and fingerprint recognition~\cite{blanton}, remote diagnostics~\cite{brickell}, private sequence analysis~\cite{franz}, private biometric identification~\cite{huang}, privacy-preserving matrix factorization and ridge-regression~\cite{valeriaMatrix, valeriaRidge}, machine-learning classification~\cite{shafindss}, decision trees and forests~\cite{wu}, neural network training~\cite{secureml} and prediction~\cite{minionn}.
+% As seen from our experiments, the protocols generated by \tool give us better or comparable execution times as many of the above works with minimal programmer overhead.
 %\nc{What do we want to say here exactly?}
-
-\divya{We need to cite this paper \cite{yao-pipe} for yao pipelining.
-This technique was also used in \cite{valeriaMatrix,oblivm}}\nc{We could just cite this in secure partitioning if we want to say that partitioning is similar in spirit to yao pipelining?
-Otherwise, is out of place here; isnt it?}

### 2017-10-28 OaklandSubmission/impl.tex
minor
pipe impl

@@
 makes calls to the ABY
-library~\cite{aby}. We use the default ABY configuration for our
+library~\cite{aby}. 
+ABY provides a GMW-based~\cite{gmw} and a  Yao-based~\cite{yao} protocol for boolean computations.
+Although \tool can generate code for both, we have observed better performance when using the latter
+and use it in our evaluation.
+We use the default ABY configuration for our
 experiment \footnote{We set the security parameter to 128 and use OT
@@
 
-Our current implementation supports the following operators:
+\tool programs can have the following operators:
 addition, subtraction, multiplication, division by powers of two, left
@@
 
-Some of our benchamrks require accessing arrays at secret
+Some of our benchmarks require accessing arrays at secret
 indices. While \tool enforces the array indices to be
@@
 For example, consider the expression $A[x]$ where $A$ is an array of
-size 2 and $x$ is secret-shared. The developer can expresse
+size 2 and $x$ is secret-shared. The developer can express
 this functionality in \tool as $\cond{x > 0}{A[1]}{A[0]}$. In general,
 a secret access to an array of
-size $2^n$ requires $2^{n}-1$ multiplexers in \tool.
+size $n$ requires ${n}-1$ multiplexers in \tool.
 

### 2017-10-28 OaklandSubmission/pipe.tex
pipe impl

@@
 In more detail, let $P$ be a program in our source language that generates a circuit $\crct$.
-For a large enough $P$, the circuit $\crct$  can be larger than the memory size\footnote{In fact, there is an upper limit of $2^{32}-1$ gates for the circuit size in ABY but for most machines the memory limit is hit first} and fail to execute.
+For a large enough $P$, the circuit $\crct$  can be larger than the memory size\footnote{In fact, there is an upper limit of $2^{32}-1$ gates for the circuit size in ABY but for most machines the memory limit is hit first.} and fail to execute.
 Partitioning enables us to
@@
 
-Let $P$ be a program that takes (secret) inputs $x$ and $y$ and produces an output $z$ to both parties (Alice and Bob).
+Let $P$ be a program that takes (secret) inputs $x$ from Alice and $y$ from Bob and produces an output $z$ to both parties.
 Let $P_1||P_2||\cdots||P_k$ be a sequence of programs such that the following holds: Define $s_0 = \bot$ (the public empty state).
@@
 Once again, define $s_0 = \bot$.
-Without loss of generality, let all $s_i$ be values in some additive ring $(\mathbb{Z},+)$ (e.g., the additive ring $(\mathbb{Z}_{2^{64}},+)$ i.e., the additive ring of integers modulo $2^{64}$).
+Without loss of generality, let all $s_i$ be values in some additive ring $(\mathbb{Z},+)$ (e.g., the additive ring $(\mathbb{Z}_{2^{64}},+)$, i.e., the additive ring of integers modulo $2^{64}$).
 Let $r_1,\cdots,r_{k-1}$ be a sequence of random values sampled from the same ring $(\mathbb{Z},+)$ by Alice (in our implementation, all $r_i$ values are generated by a pseudorandom function).
-Let $Q_1$ be the program that takes as inputs $x,y, r_1$ (and empty state $s_0$), runs $P_1$ (as defined above) to get $s_1$ and then outputs $o_1 = s_1 + r_1$ {\em only to} Bob\footnote{While the description of our protocol here assumes that the underlying backend supports only one party receiving output, this is only a simplifying assumption, and we can easily modify our protocol in the case where both parties must receive the same output.
-To do so, we let $Q_1$ output $o_1 = s_1+r_1+t_1$ (where $t_1$ is random and chosen by Bob) and have both parties learn this.
+Let $Q_1$ be the program that takes as input $x,r_1$ from Alice and $y$ from Bob (and empty state $s_0$), and runs $P_1$ (as defined above) to compute $s_1$ and then outputs $o_1 = s_1 + r_1$ {\em only to} Bob\footnote{While the description of our protocol here assumes that the underlying backend supports only one party receiving output, this is only a simplifying assumption, and we can easily modify our protocol in the case where both parties must receive the same output.
+To do so, we modify $Q_1$ to output $o_1 = s_1+r_1+t_1$ to both parties (where $t_1$ is random and chosen by Bob).
 We can then appropriately modify the remaining steps as well.}.
 Alice's output from $Q_1$ is $r_1$.
-Next, every $Q_i$ ($2\leq i\leq k-1$) takes as inputs $x,y,r_{i-1},o_{i-1},r_i$, runs $P_i$ on inputs $x,y$ and state $o_{i-1}-r_{i-1}$ (where $-$ denotes subtraction in the ring $(\mathbb{Z},+)$) and then outputs $s_i+r_i$ only to Bob.
+Next, every $Q_i$ ($2\leq i\leq k-1$) takes as inputs $x,r_{i-1},r_i$ from Alice and $y,o_{i-1}$ from Bob, runs $P_i$ on inputs $x,y$ and state $o_{i-1}-r_{i-1}$ (where $-$ denotes subtraction in the ring $(\mathbb{Z},+)$) and then outputs $s_i+r_i$ only to Bob.
 Again, Alice's output from $P_i$ is $r_i$.
@@
 \begin{theorem}[Correctness and security of partitioning]
-Let $P$ be a program in our source language with output $z$, that is decomposed into programs $Q_1||Q_2||\cdots ||Q_k$ (as defined in the partioning procedure above) and compiled into the protocol $\prot_1||\prot_2\cdots ||\prot_k$ that outputs $r_1||r_2||\cdots||z$ (to Alice) and $o_1||o_2||\cdots||z$ (to Bob), then $\prot$ securely realizes $P$.
+Let $P$ be a program in our source language with output $z$, that is decomposed into programs $Q_1||Q_2||\cdots ||Q_k$ (as defined in the partioning procedure above) and compiled into the protocol $\prot = \prot_1||\prot_2\cdots ||\prot_k$ that outputs $r_1||r_2||\cdots||z$ (to Alice) and $o_1||o_2||\cdots||z$ (to Bob), then $\prot$ securely realizes $P$.
 \end{theorem}
 
-\noindent {\em Proof.} From Theorem \ref{theorem:security}, it follows that $\prot_i$ securely realizes $Q_i$ for every $1\leq i\leq k$.
+\noindent {\em Proof.} From \theoremref{security}, it follows that $\prot_i$ securely realizes $Q_i$ for every $1\leq i\leq k$.
 Hence, we know that the only information learnt by Alice is $r_1||r_2||\cdots||r_{k-1}||z$ and by Bob is $o_1||o_2||\cdots||o_{k-1}||z$.
-Since $r_i$ and $o_i$ ($1\leq i\leq k-1$) are individually uniformly random (in $(\mathbb{Z},+)$, the proof follows.
+Since $r_i$ and $o_i$ ($1\leq i\leq k-1$) are individually uniformly random (in $(\mathbb{Z},+)$, outputs received by the adversary can be easily simulated given the final output $z$.
 

### 2017-10-28 OaklandSubmission/impl.tex
comment

@@
 bitwise-negation, logical-(not, and, or, xor), and comparisons (less
-than, greater than, equality). 
+than, greater than, equality).  \TODO{Say something about our compiler being crypto cost aware; some ops always in either A or B and add can be in both}
 Because of their high cost, integral division and floating-point

### 2017-10-28 OaklandSubmission/intro.tex
minor

@@
 and implement specialized, efficient \mpc protocols (similar
-to~\cite{shafindss,wu,minionn}) for her tasks.
+to~\cite{shafindss,wu,minionn,valeriaMatrix}) for her tasks.
 %\end{tiret}
@@
 
-\subsubsection*{\tool}
-\label{sec:contrib}
+%\subsubsection*{\tool}
+%\label{sec:contrib}
 %Write high level overview of programming language and tool.. \divya{have some preamble}
@@
 or better than the custom, specialized
-protocols from previous work %(up to a factor of 19x in our
+protocols from previous work~\cite{shafindss,wu,minionn,secureml,cryptonets,valeriaMatrix}.
+These papers (and others) cite the inefficiency of generic \mpc as the major motivation
+behind the design of specialized protocols. Using \tool, we empirically  demonstrate
+the surprising fact that generic \mpc implementations are much more efficient than what they were believed to be. 
+ %(up to a factor of 19x in our
 %experiments, depending on the functionality) \nc{I added this but I

### 2017-10-28 OaklandSubmission/ex.tex

@@
 program as output. Figure~\ref{fig:ex-aby} shows the output code for
-the example in Figure~\ref{fig:ex-sml}. The output program contains
+the example in Figure~\ref{fig:ex-sml}. The output program
 contains party-specific code for inputs and outputs
@@
 a \mpc protocol.
+
+We can now concretely see the advantages of \tool. Unarguably, it is
+easier for a developer to program, and get right, the code in
+Figure~\ref{fig:ex-sml}, rather than the code in
+Figure~\ref{fig:ex-aby} (as we mentioned in Section~\ref{sec:intro},
+other high-level languages are limited in their use of a single,
+arithmetic or boolean, crypto back-end). \tool also enables the
+programmer to easily modify their code, while the compiler takes care
+of efficiency. For example, consider in Figure~\ref{fig:ex-sml} a
+change from multiplication to bitwise OR in the \ls{for} loop. It
+turns out that once that's the case, it is more efficient to do both
+the addition and bitwise OR using boolean circuits (if the addition is
+still done using arithmetic, the conversion cost starts to take
+over). In \tool, the programmer simply needs to change one operator in
+the source code, and the compiler generates efficient code that uses
+boolean addition. Whereas, if the programmer was writing ABY
+code, she either has to sacrifice performance, or she would have to
+revisit many parts of the circuit and change them.
+
+To summarize, \tool raises the level of abstraction for the
+programmers, and generates efficient crypto protocols automatically,
+while its metatheory provides strong correctness and security
+guarantees.
 
@@
 %% Moreover, the \tool compiler also generates a C implementation that can be run natively on a single machine for functional testing.
-
-%% To summarize, \tool raises the level of abstraction, provides strong static guarantees, and generates efficient code automatically.

### 2017-10-28 OaklandSubmission/ex.tex
pass over sec 2

@@
 ``ideal'' functionalities, without dealing with any cryptographic
-details. The languages provides multi-dimensional arrays, conditional
+details. The language provides multi-dimensional arrays, conditional
 expressions (the ternary $?\::$ operator), \ls{for}~loops,
@@
 Importantly, conversions between the arithmetic and boolean
-represenation require share-conversions. The \tool compiler also
+representations require share-conversions. The \tool compiler also
 instruments these conversion gates accordingly. For example, in
@@
 To summarize, \tool raises the level of abstraction for the
-programmers, and generates efficient crypto protocols automatically,
+programmer, and generates efficient crypto protocols automatically,
 while its metatheory provides strong correctness and security

### 2017-10-29 OaklandSubmission/ex.tex
minor

@@
 $w$ and $b$ constitute the server's input (a classifier) and $x$ is
-the client's input vector. Figure~\ref{fig:ex-sml} shows \tool code
-for this example. The code first reads the inputs of the two parties
-using the \ls{input1} and \ls{input2} expressions. It then uses a
+the client's input vector. Figure~\ref{fig:ex-sml} shows the \tool code
+for this example. The code first reads the server's (client's) input using
+ \ls{input1} (\ls{input2}). It then uses a
 \ls{for} loop to compute \ls{acc}, the dot product of \ls{w} and
 \ls{x}. Finally, the code outputs the result of comparing \ls{acc}
-with \ls{b} only to the client.
+with \ls{b} only to the client using \ls{output2}.
 
@@
 components translate into API calls into our crypto back-end
-(ABY). For instance, for the code in Figure~\ref{fig:ex-sml}, the \tool
+(ABY). E.g., for the code in Figure~\ref{fig:ex-sml}, the \tool
 compiler realizes that the array index \ls{i} in the dot product loop
@@
 \tool is to
-compile the output C++ code and execute it. Doing so reduces away the
+compile the output C++ code and execute it. Doing so evaluates away the
 public parts of the program, including the array accesses, and
@@
 of efficiency. For example, consider in Figure~\ref{fig:ex-sml} a
-change from multiplication to bitwise OR in the \ls{for} loop. It
+change from multiplication to bitwise-or in the \ls{for} loop. It
 turns out that once that's the case, it is more efficient to do both
-the addition and bitwise OR using boolean circuits (if the addition is
+the addition and bitwise-or using boolean circuits (if the addition is
 still done using arithmetic, the conversion cost starts to take
@@
 while its metatheory provides strong correctness and security
-guarantees.
+guarantees. We provide a formal description of the \tool language next.
 

### 2017-10-29 OaklandSubmission/intro.tex
minor

@@
 for i in [0:30] { acc =  acc + (w[i] * x[i]); }
-output((acc > b) ? 1 : 0, 2) //only to party 2
+output2((acc > b) ? 1 : 0) //only to party 2
 \end{lstlisting}
@@
 instance, \tool compiler realizes that the array index~\ls{i} is
-public, and hence, the array accesses need not be compiled
-obliviously. Further, within the secret parts, \tool compiler is
+public, and generates non-cryptographic code for it.\rs{check this}
+Further, within the secret parts, \tool compiler is
 aware
@@
 \subsubsection*{Scalability (secure code partitioning)} \mpc tools
-often do not scale to large programs. The reason is that
+often do not scale to large functionalities. The reason is that
 most \mpc implementations use a circuit-like representation as an
@@
 the security risk of revealing the intermediate outputs. \tool comes
-to the rescue -- it automatically inserts the required instrumentation
+to the rescue; it automatically inserts the required instrumentation
 to ensure security of these intermediate outputs. As
@@
 Additionally, these prior works provide informal languages
-or libraries and lack formal semantics.
+or libraries that lack formal semantics and static guarantees.
 %Can be added to related work

### 2017-10-29 OaklandSubmission/intro.tex
nit

@@
 instance, \tool compiler realizes that the array index~\ls{i} is
-public, and generates non-cryptographic code for it.\rs{check this}
+public, and generates non-cryptographic code for the array accesses.
 Further, within the secret parts, \tool compiler is

### 2017-10-29 OaklandSubmission/intro.tex
formal guarantees

@@
 \subsubsection*{Formal guarantees} We prove formal correctness and
-security theorems for our compiler. The correctness theorem proves a
-forward simulation between the
-``ideal'' semantics (the trusted-third party semantics) of a source
+security theorems for our compiler. The correctness theorem relates the
+``trusted third party'' semantics of a source
 program and the ``protocol'' semantics (the distributed \mpc
 semantics that relies on circuit evaluation) of the corresponding
-compiled program. The simulation
+compiled program. The theorem
 guarantees that for a well-typed source program, the two semantics
@@
 the security theorem, we formally reduce the security
-of our compiler to the security of the back-end cryptographic
-protocol (by also invoking our correctness theorem). We also prove a formal security theorem for our partitioning
+of our compiler to the security of the  \mpc back-end
+ (by also invoking our correctness theorem). We also prove a formal security theorem for our partitioning
 scheme.

### 2017-10-29 OaklandSubmission/pipe.tex
notational changes in S4

@@
 
-We  describe ``secure code partitioning'' that allows us to execute {\it large} programs, i.e., programs with a large number of operations.
-In more detail, let $P$ be a program in our source language that generates a circuit $\crct$.
-For a large enough $P$, the circuit $\crct$  can be larger than the memory size\footnote{In fact, there is an upper limit of $2^{32}-1$ gates for the circuit size in ABY but for most machines the memory limit is hit first.} and fail to execute.
-Partitioning enables us to
-execute such programs via a source to source transformation that is oblivious to the underlying \mpc backend.
-Partitioning decomposes the program $P$ into a sequence of smaller \tool programs $Q_1||Q_2||\cdots||Q_k$ (defined below).
-We compile and execute each $Q_i$ sequentially, feeding the outputs of $Q_i$ as state information to $Q_{i+1}$.
-We prove that partioning is correct ($P$ and $Q_1||Q_2||\cdots||Q_k$ compute the same functionality) and secure (execution of $Q_1||Q_2||\cdots||Q_k$ does not reveal any more information than $P$).
-More details follow.
+\newcommand{\prog}{s}
+\newcommand{\seq}{||}
+\newcommand{\state}{st}
 
-Let $P$ be a program that takes (secret) inputs $x$ from Alice and $y$ from Bob and produces an output $z$ to both parties.
-Let $P_1||P_2||\cdots||P_k$ be a sequence of programs such that the following holds: Define $s_0 = \bot$ (the public empty state).
-For all $1\leq i\leq k-1$, $P_i$ takes inputs $x, y$ and $s_{i-1}$ and outputs  state information $s_i$.
-Finally, $P_k$ takes inputs $x,y$ and $s_{k-1}$ to  output $z$.
-Each program $P_i$ ($1\leq i\leq k$) is required to be ``sufficiently'' small (as described later).
-It is possible to decompose any program $P$  into such $P_1||P_2||\cdots||P_k$.
-If \tool generates circuit $\crct_i$ from $P_i$, the parties can execute $\crct_1||\crct_2||\cdots||\crct_k$ sequentially (in a distributed setting)  to obtain $s_1,\cdots,s_{k-1},$ and finally output $z$.
-At the $i^{\textrm{\tiny{th}}}$ step, the parties only need to store information proportional to $x,y,s_{i-1}$ and $\crct_i$ (which is much smaller than $\crct$).
-However, this execution enables the parties to learn $s_i$ (for all $1\leq i\leq k-1$), which completely breaks the security.
+We  now describe our ``secure code partitioning'' technique that
+allows \tool to execute programs that require large circuits.
+Let $\prog$ be a program in our source language that
+generates a circuit $\crct$. For some programs, the circuit $\crct$
+can be larger than
+the memory size\footnote{In fact, there is
+  an upper limit of $2^{32}-1$ gates for the circuit size in ABY but
+  for most machines the memory limit is hit first.} and fail to
+execute. Partitioning enables us to 
+execute such programs via a source to source transformation that is
+oblivious to the underlying \mpc backend. Partitioning decomposes the
+program $\prog$ into a sequence of smaller \tool programs
+$\prog_1\seq\prog_2\seq\dots\seq\prog_k$ (as defined below), such
+that the circuit size
+requirement for each of the $\prog_i$ itself is manageable. We compile
+and execute each $\prog_i$ sequentially, feeding the outputs of
+$\prog_i$ as state
+information to $\prog_{i+1}$. We prove our partitioning scheme to be
+correct ($\prog$ and $\prog_1\seq\prog_2\seq\dots\seq\prog_k$ compute
+the same functionality) and secure (execution of
+$\prog_1\seq\prog_2\seq\dots\seq\prog_k$ does
+not reveal any more information than $s$). More details follow.
 
-To overcome this problem, we define a sequence of new programs $Q_i$ ($1\leq i\leq k$) as follows.
-Once again, define $s_0 = \bot$.
-Without loss of generality, let all $s_i$ be values in some additive ring $(\mathbb{Z},+)$ (e.g., the additive ring $(\mathbb{Z}_{2^{64}},+)$, i.e., the additive ring of integers modulo $2^{64}$).
-Let $r_1,\cdots,r_{k-1}$ be a sequence of random values sampled from the same ring $(\mathbb{Z},+)$ by Alice (in our implementation, all $r_i$ values are generated by a pseudorandom function).
-Let $Q_1$ be the program that takes as input $x,r_1$ from Alice and $y$ from Bob (and empty state $s_0$), and runs $P_1$ (as defined above) to compute $s_1$ and then outputs $o_1 = s_1 + r_1$ {\em only to} Bob\footnote{While the description of our protocol here assumes that the underlying backend supports only one party receiving output, this is only a simplifying assumption, and we can easily modify our protocol in the case where both parties must receive the same output.
-To do so, we modify $Q_1$ to output $o_1 = s_1+r_1+t_1$ to both parties (where $t_1$ is random and chosen by Bob).
-We can then appropriately modify the remaining steps as well.}.
-Alice's output from $Q_1$ is $r_1$.
-Next, every $Q_i$ ($2\leq i\leq k-1$) takes as inputs $x,r_{i-1},r_i$ from Alice and $y,o_{i-1}$ from Bob, runs $P_i$ on inputs $x,y$ and state $o_{i-1}-r_{i-1}$ (where $-$ denotes subtraction in the ring $(\mathbb{Z},+)$) and then outputs $s_i+r_i$ only to Bob.
-Again, Alice's output from $P_i$ is $r_i$.
-The last program $Q_k$ takes inputs $x,y,r_{k-1},o_{k-1}$, runs $P_k$ on inputs $x,y$ and state $o_{k-1}-r_{k-1}$ and outputs $z$ to both parties.
+Let $\prog$ be a program that takes (secret) inputs $x$ from Alice and
+$y$ from Bob and produces an output $z$ to both parties. Let
+$\prog_1\seq\prog_2\seq\dots\seq\prog_k$ be a sequence of programs
+such that the following holds. Define $\state_0 = \bot$ (the public empty
+state). For all
+$1\leq i\leq k-1$, $P_i$ takes inputs $x, y$ and $\state_{i-1}$ and
+outputs state $\state_i$. Finally, $\prog_k$ takes inputs $x, y$ and
+$\state_{k-1}$ to  output $z$. It is
+possible to decompose any program $\prog$  into such
+$\prog_1\seq\prog_2\seq\dots\seq\prog_k$. If \tool generates circuit
+$\crct_i$ from
+$\prog_i$, the parties can execute
+$\crct_1, \crct_2, \dots \crct_k$
+sequentially (in a distributed setting)  to obtain
+$\state_1,\dots,\state_{k-1},$ and finally output $z$. At the
+$i^{\textrm{\tiny{th}}}$ step, the parties only need to store
+information proportional to $x,y,\state_{i-1}$ and $\crct_i$ (which is
+much smaller than $\crct$). However, this execution enables the
+parties to learn $\state_i$ (for all $1\leq i\leq k-1$), which
+completely breaks the security.
 
-Given $P$, partitioning first decomposes $P$ into $P_1||P_2||\cdots||P_k$ with the guarantee that the circuits generated for every $Q_i$ defined above are small enough to fit in the machine memory.
-We then compile and execute the $k$ programs $Q_1||Q_2||\cdots||Q_k$ sequentially (freeing up memory usage after  execution of each $Q_i$) to finally output $z$ to both parties.
+To overcome this problem, we define a sequence of new programs
+$\prog'_i$ ($1\leq i\leq k$) as follows. Once again, define $\state_0
+= \bot$. Without
+loss of generality, let all $\state_i$ be values in some additive ring
+$(\mathbb{Z},+)$ (e.g., the additive ring $(\mathbb{Z}_{2^{64}},+)$,
+i.e., the additive ring of integers modulo $2^{64}$).
+Let $r_1,\cdots,r_{k-1}$ be a sequence of random values sampled from
+the same ring $(\mathbb{Z},+)$ by Alice (in our implementation, all
+$r_i$ values are generated by a pseudorandom function). Let $\prog'_1$
+be the program that takes as input $x,r_1$ from Alice and $y$ from Bob
+(and empty state $\state_0$), and runs $\prog_1$ (as defined above) to
+compute $\state_1$ and then outputs $o_1 = \state_1 + r_1$ {\em only
+  to} Bob\footnote{While the description of our protocol here assumes
+  that
+  the underlying backend supports only one party receiving output,
+  this is only a simplifying assumption, and we can easily modify our
+  protocol in the case where both parties must receive the same
+  output. To do so, we modify $\prog'_1$ to output $o_1 = s_1+r_1+t_1$ to
+  both parties (where $t_1$ is random and chosen by Bob). We can then
+  appropriately modify the remaining steps as well.}. Alice's output
+from $\prog'_1$ is $r_1$. Next, every $\prog'_i$ ($2\leq i\leq k-1$)
+takes as inputs $x,r_{i-1},r_i$ from Alice and $y,o_{i-1}$ from Bob,
+runs $\prog_i$
+on inputs $x,y$ and state $o_{i-1}-r_{i-1}$ (where $-$ denotes
+subtraction in the ring $(\mathbb{Z},+)$) and then outputs $s_i+r_i$
+only to Bob. Again, Alice's output from $\prog_i$ is $r_i$. The last
+program $\prog'_k$ takes inputs $x,y,r_{k-1},o_{k-1}$, runs $\prog_k$
+on inputs $x,y$ and state $o_{k-1}-r_{k-1}$ and outputs $z$ to both
+parties.
+
+Given $\prog$, partitioning first decomposes $\prog$ into
+$\prog_1\seq\prog_2\seq\dots\seq\prog_k$ with the guarantee that the
+circuits generated for every $\prog'_i$, as defined above, are small
+enough to fit in the machine memory. We then compile and execute the
+$k$ programs $\prog'_1\seq\prog'_2\seq\dots\seq\prog'_k$ sequentially
+(freeing up memory usage after execution of each $\prog'_i$) to
+finally output $z$ to both parties.
 
 \begin{theorem}[Correctness and security of partitioning]
-Let $P$ be a program in our source language with output $z$, that is decomposed into programs $Q_1||Q_2||\cdots ||Q_k$ (as defined in the partioning procedure above) and compiled into the protocol $\prot = \prot_1||\prot_2\cdots ||\prot_k$ that outputs $r_1||r_2||\cdots||z$ (to Alice) and $o_1||o_2||\cdots||z$ (to Bob), then $\prot$ securely realizes $P$.
+Let $\prog$ be a program in our source language with output $z$, that
+is decomposed into programs
+$\prog'_1\seq\prog'_2\seq\dots\seq\prog'_k$ (as defined in
+the partioning procedure above) and compiled into the protocol $\prot
+= \prot_1\seq\prot_2\dots\seq\prot_k$ that outputs $r_1\seq
+r_2\seq\dots\seq z$
+(to Alice) and $o_1\seq o_2\seq\dots||z$ (to Bob), then $\prot$ securely
+realizes $P$.
 \end{theorem}
 
-\noindent {\em Proof.} From \theoremref{security}, it follows that $\prot_i$ securely realizes $Q_i$ for every $1\leq i\leq k$.
-Hence, we know that the only information learnt by Alice is $r_1||r_2||\cdots||r_{k-1}||z$ and by Bob is $o_1||o_2||\cdots||o_{k-1}||z$.
-Since $r_i$ and $o_i$ ($1\leq i\leq k-1$) are individually uniformly random (in $(\mathbb{Z},+)$, outputs received by the adversary can be easily simulated given the final output $z$.
+\noindent {\em Proof.} From \theoremref{security}, it follows that
+$\prot_i$ securely realizes $\prog'_i$ for every $1\leq i\leq
+k$. Hence, we know that the only information learnt by Alice is
+$r_1||r_2||\cdots||r_{k-1}||z$ and by Bob is
+$o_1||o_2||\cdots||o_{k-1}||z$. Since $r_i$ and $o_i$ ($1\leq i\leq
+k-1$) are individually uniformly random (in $(\mathbb{Z},+)$, outputs
+received by the adversary can be easily simulated given the final
+output $z$.
 
 Our current implementation of secure code partitioning requires a manual step.
-In particular, the developer needs to decompose the large program $P$ into 
-the sequence of small programs $P_1||\ldots||P_k$ manually.
-Then \tool generates $Q_1||\ldots||Q_k$ automatically.
+In particular, the developer needs to decompose the large program $\prog$ into 
+the sequence of small programs $\prog_1\seq\dots\seq\prog_k$ manually.
+Then \tool generates $\prog'_1\seq\dots\seq\prog'_k$ automatically.
 Automating this decomposition step requires an analysis that can 
-statically estimate the resource usage of a \tool program and we leave the construction of such
-an analysis as future work.
+statically estimate the resource usage of a \tool program and we leave
+the construction of such an analysis as future work.

### 2017-10-29 OaklandSubmission/ld.tex
temp

@@
-\newcommand{\kw}[1]{{\lstinline[basicstyle=\small\color{blue}]{#1}}}
-%\newcommand{\kw}[1]{{\ensuremath{\mathtt{#1}}}}
+%\newcommand{\kw}[1]{{\lstinline[basicstyle=\small\color{blue}]{#1}}}
+\newcommand{\kw}[1]{{\ensuremath{\mathtt{#1}}}}
 \newcommand{\ftext}[1]{\text{\small{#1}}}
@@
 
-In this section we prove correctness and security of our compiler \tool.
+In this section we prove correctness and security of our \tool compiler.
 We formalize our source, intermediate and circuit languages. 
@@
 %parties.
-Our source runtime semantics describe the standard execution semantics of the source programs and
-generates observations corresponding to the values revealed to the
+Our source runtime semantics describe the ``trusted third party" execution semantics of the source programs and
+generate observations corresponding to the values revealed to the
 parties.
 %
-We present the compilation rules that type check a program in the source language and generates a program in the intermediate language.
+We present the compilation rules that type check a program in the source language and generate a program in the intermediate language.
 This intermediate language models programs such as \figureref{ex-aby} generated by our compiler implementation.
 %
-We evaluate the program in our intermediate language  to a circuit by ``computing away'' the public parts and the arrays.
-Crucially, this step does not have access to the secret inputs--those are processed by the distributed circuit semantics that model the \mpc back-end.
+We evaluate the program in our intermediate language  to a circuit by ``evaluating away'' the public parts and the arrays.
+Crucially, this step does not have access to the secret inputs; those are processed by the distributed circuit semantics that model the \mpc back-end.
 %
-Here, the generated circuit is evaluated in the distributed setting, where the parties run an interactive protocol.
+Evaluation in the distributed setting involves the parties running an interactive protocol.
 This step, like  the source semantics, emits observations corresponding to the values revealed to the parties.
@@
 To prove the correctness of \tool, we prove that  the observations in source semantics and distributed circuit semantics are identical (see \theoremref{correctness}). 
-We combine this correctness and the security of the \mpc back-end to argue security of the implementations generated by \tool (see \theoremref{security}).
+We combine this correctness guarantee and the security of the \mpc back-end to argue security of the implementations generated by \tool (see \theoremref{security}).
 %For the security of our compiler, we rely on the correctness guarantee and  reduce the security to the security of the underlying cryptographic protocol used to compute the generated circuit (see \theoremref{security}).
 We present only selected parts of our formalization for space
-reasons. Full definitions and proofs can be found \cite{anonymized}.
+reasons. Full definitions and proofs can be found in~\cite{anonymized}.
 
@@
   \]
-\caption{Source language}
+\caption{Source language syntax}
 \label{fig:srclang}
@@
 \subsubsection*{Source language} Our source language
- is a simple imperative language. \divya{We present the core of our language in \figureref{srclang} but our implementation is more expressive (see \sectionref{impl}).} Types
+ is a simple imperative language and we show a subset of the supported operators in \figureref{srclang}. Types
 $\psi$ consist of the base types $\sigma$, and arrays
 of base types $\sigma[n]$, where $n$ is the array length. Though we
-model only one dimensional arrays, our implementation supports higher
-dimensional arrays as well. Expressions $e$ in the language include the
+model only one dimensional arrays, our implementation supports
+ multi-dimensional arrays as well. Expressions $e$ in the language include the
  integer constants $n$, \kw{bool} constants $\top$ and $\bot$,
@@
 
-\noindent\textbf{Source Semantics.} The runtime semantics for the source language is shown in
-\figureref{srcsem}. \TODO{Add something about claer text stand alone exec.} Values $v$, runtime environments $\rho$, and
+\noindent\textbf{Source semantics.} The runtime semantics for the source language is shown in
+\figureref{srcsem}. These semantics show how a ``trusted third party" computes the outputs when given the inputs of both the parties. 
+ Values $v$, runtime environments $\rho$, and
 observations $O$ are defined as follows:
@@
 provided $n < n_{1}$, the length of the array. Rule ({\sc{E-Inp}})
-evaluates to some constant $c$ denoting party $j$'s input. We model
-the inputs to be base constants, an array input can be written in the
+evaluates to some constant $c$ denoting party $j$'s input. 
+%We model the inputs to be base constants, a
+An array input can be written in the
 language as $[\kw{in}_{j}]_{n}$, which can then evaluate using the
@@
   \]
-\caption{Intermediate Language}
+\caption{Intermediate language syntax}
 \label{fig:interlang}
@@
 intermediate language of our compiler. The syntax follows that of the source
-language, except that the types and operators are \emph{labeled}. Below
-we mainly focus on the bits that are different from the source
-language.
-
-Labels $\ell$ consist of secret labels $\mathcal{A}$ and
-$\mathcal{B}$, denoting the arithmetic and boolean shared secrets
-resp., and the public label $\mathcal{P}$. Types $\tau$ are then
+language, except that the types and operators are \emph{labeled}. 
+%Below we mainly focus on the bits that are different from the source
+%language.
+A label $\ell$ can be the  public label $\mathcal{P}$ or  one of the secret labels $\mathcal{A}$ or
+$\mathcal{B}$, which denote arithmetic and boolean respectively.
+Types $\tau$ are then
 labeled base types $\sigma^{\ell}$ and arrays of labeled base types
@@
 that the binary operators, and the conditional forms
-are annotated with label $\ell$, denoting how the operators should be
-evaluated ($\mathcal{P}$ for in-clear, and $\mathcal{A}$ and
-$\mathcal{B}$ for using arithmetic or boolean circuits resp.). The
+are annotated with labels $\ell$, denoting how the operators should be
+evaluated ($\mathcal{P}$ for in-clear, $\mathcal{A}$ for arithmetic circuits, and
+$\mathcal{B}$ for boolean circuits). The
 form $\widetilde{e} \rhd m$ denotes coercing the value of $\widetilde{e}$ to be
-$m$-secret shared. The statements $\widetilde{s}$ are analogous to
-$s$.
+$m$-secret shared, which is useful for share-conversions between boolean and arithmetic and also for sharing public values. 
+%The statements $\widetilde{s}$ are analogous to $s$.
 
@@
                  \Gamma \vdash e : \kw{uint}^{\mathcal{P}} \leadsto \widetilde{e}\\\\
-                 \Gamma \models e < n
+                  \models e < n
                }
@@
                  \Gamma \vdash e_{2} : \sigma^{\ell} \leadsto \widetilde{e_{2}}\\\\
-                 \Gamma \models e_{1} < n
+                  \models e_{1} < n
                }
@@
 
-\subsubsection*{Source to Intermediate Compilation judgments} We provide the compilation
-rules in \figureref{compile}. \divya{PL jargon?} We present the rules in a
+\subsubsection*{Source to intermediate compilation judgments} We provide the compilation
+rules in \figureref{compile}. We present the rules in a
 declarative style, where the rules are non-syntax
@@
 non-deterministically. \sectionref{impl} describes the label  inference
-scheme in our implementation.
-
-Judgment $\Gamma \vdash e : \tau \leadsto \widetilde{e}$, where
+scheme of our implementation.
+
+The judgment $\Gamma \vdash e : \tau \leadsto \widetilde{e}$, where
 $\Gamma$ maps variables $x$ to types $\tau$, says that
-under $\Gamma$, $e$ (in source language) compiles to $\widetilde{e}$ (in intermediate language) with type $\tau$.
-
+under $\Gamma$, $e$ (in source language) compiles to $\widetilde{e}$ (in the intermediate language) and $\widetilde{e}$  has type $\tau$.
 The rule ({\sc{T-Cons}}) assigns the label $\mathcal{P}$ to the
@@
 rule ({\sc{T-Gt}}) compiles $e_{1} > e_{2}$ to either public comparison,
-or secret boolean comparison. The rule for conditional
+or secret boolean comparison $>_\mathcal{B}$ (and never $>_{\mathcal{A}}$). The rule for conditional
 ({\sc{T-Cond}}) has two cases: when the conditional expression $e$ is of
@@
 that the array index $e$ is public, and uses a static bounds checking
-judgment $\Gamma \models e < n$ to prove that the array index is in
-bounds. Section~\ref{sec:impl} discusses our implementation of bounds
-checking. Rule ({\sc{T-Inp}}) picks a label $m$ for the input. Finally,
-rule ({\sc{T-Sub}}) is the subsumption rule that coerces an expression of
+judgment $ \models e < n$ to prove that the array index is in
+bounds. Section~\ref{sec:impl} discusses our implementation of this check. Rule ({\sc{T-Inp}}) picks a label $m$ for the input. Finally,
+the rule ({\sc{T-Sub}}) is the subsumption rule that coerces an expression of
 type $\sigma^{\ell}$ to an expression of type $\sigma^{m}$ using the
-coerce expression. It is important for security that the secrets cannot be coerced to public
-values.
+coerce expression. It is important for security that the secrets cannot be coerced to public values and indeed ({\sc{T-Sub}}) does not permit it.
 

### 2017-10-29 OaklandSubmission/intro.tex
minor intro change

@@
 
+Since its introduction in 1980s, there has been a long
+sequence of work that has transformed 2PC from a mere theoretical tool to
+something practically efficient and usable.
 Our goal is to build a framework that enables programmers, who are not
 experts in cryptography, to easily program efficient and scalable \mpc
-protocols. Since its introduction in 1980s, there has been a long
-sequence of work along the same lines. To understand the
+protocols.  To understand the
 state-of-the-art, let us consider an example underlying many
@@
 the computation expressed as either a boolean
-circuit~(\cite{yao,gmw}) or an arithmetic circuit~(\cite{gentry}).
+circuit~(\cite{yao,gmw}) or an arithmetic circuit~(\cite{sss,viff,gentry}). \TODO{check!}
 %
@@
 the computation to be expressed as a combination of arithmetic and
-boolean circuits. However, here the programming interface is quite
+boolean circuits. However, here, the programming interface is quite
 low-level -- the programmer is required to first manually split the
 computation into arithmetic and boolean components, and then write the
-circuits for all the componenets manually, including the appropriate
+circuits for all the components manually, including the appropriate
 inter-conversion gates between them. Clearly, writing correct
@@
 or better than the custom, specialized
-protocols from previous work~\cite{shafindss,wu,minionn,secureml,cryptonets,valeriaMatrix}.
+protocols from previous works~\cite{shafindss,wu,minionn,secureml,cryptonets,valeriaMatrix}.
 These papers (and others) cite the inefficiency of generic \mpc as the major motivation

### 2017-10-29 OaklandSubmission/ld.tex
till correctness

@@
 that the binary operators, and the conditional forms
-are annotated with labels $\ell$, denoting how the operators should be
-evaluated ($\mathcal{P}$ for in-clear, $\mathcal{A}$ for arithmetic circuits, and
-$\mathcal{B}$ for boolean circuits). The
-form $\widetilde{e} \rhd m$ denotes coercing the value of $\widetilde{e}$ to be
-$m$-secret shared, which is useful for share-conversions between boolean and arithmetic and also for sharing public values. 
+are annotated with labels $\ell$.
+Looking ahead, the label determines how the operators are
+evaluated: $\mathcal{P}$-labeled operators are evaluated in-clear, $\mathcal{A}$-labeled operators generate arithmetic circuits, and
+$\mathcal{B}$-labeled operators generate boolean circuits. The
+form $\widetilde{e} \rhd m$ denotes coercing $\widetilde{e}$ to the label $m$, which generates conversions to $m$. 
 %The statements $\widetilde{s}$ are analogous to $s$.
@@
 The evaluation to a circuit computes away the public parts of the program and also \emph{flattens} the arrays so that the circuits are unaware of the array structure.
-Crucially, this phase of the semantics does not have access to the secrets.
+Crucially, this phase of the semantics does not have access to the secret inputs.
 Below, we first provide the language for the circuits used followed by our evaluation rules.
@@
     \ftext{Circuit gate} & \gate & ::= & \wire \mid \kw{in}^{m}_{j} \mid \kw{mult}\:\gate_1\:\gate_{2} \mid \kw{gt}\:\gate_{1}\:\gate_{2}\\
-    & & \mid & \kw{mux}\:\gate\:\gate_{1}\:\gate_{2} \mid \bval \rhd m\\
-    \ftext{Base value} & \bval & ::= & c \mid \gate\\
-    \ftext{Value} & \val & ::= & \bval \mid [\overline{\bval_{i}}]_{n}\\
+    & & \mid & \kw{mux}\:\gate\:\gate_{1}\:\gate_{2} \mid \gate \rhd m \mid c\\
+    \ftext{Sub-circuit} & \val & ::= & \gate \mid [\overline{\gate_{i}}]_{n}\\
     %\ftext{Runtime environment} & \widetilde{\rho} & ::= & \cdot \mid \widetilde{\rho},x \mapsto \val \\
@@
   \]
-\caption{Circuits   \divya{Can we merge base value and circuit gate?}}
+\caption{Circuits  syntax}
 \label{fig:circuits}
 \end{figure}
-
-\subsubsection*{Circuits}
-\figureref{circuits} shows the circuits syntax. A wire id range
-$\wire$ denotes a set of circuit wires that carry the value of a secret
-shared variable in the output circuit (we will concretely define these
-values later as part of the circuit semantics). Circuit gates $\gate$
-are $\wire$, input gates, \kw{mult}, \kw{gt}, and \kw{mux} gates, and
-coerce gates $\bval \rhd m$. \divya{Gates are not just gates but a circuit?} Target values $\val$
-then consist of base values $c$ and $\gate$, and arrays of base
-values. \\
 
@@
                {
-                 \widetilde{\rho} \vdash x[\widetilde{e}] \Downarrow \bval_n
+                 \widetilde{\rho} \vdash x[\widetilde{e}] \Downarrow \wire_n
                }\\\\
@@
 
-\subsubsection*{Evaluation to Circuit} \figureref{eval-inter-circ} shows the judgments for the evaluation of the intermediate language to a circuit. 
-The runtime environment is defined as follows:
+\subsubsection*{Evaluation to Circuits}
+\figureref{circuits} shows the syntax of circuits. A wire id range
+$\wire$ denotes a set of circuit wires that carry the runtime value of a
+ variable with a secret label (we will concretely define these runtime
+values later as part of the circuit semantics). Circuit gates $\gate$
+are $\wire$, input gates, \kw{mult}, \kw{gt}, and \kw{mux} gates, 
+coerce gates $\gate \rhd m$, and constant gates.  Sub-circuits $\val$ 
+(generated from $\widetilde e$)
+then consist of gates and arrays of gates.
+
+ \figureref{eval-inter-circ} shows the judgments for the evaluation of the intermediate language to a circuit. 
+The runtime environment maps variables to sub-circuits and is defined as follows:
 
@@
 \begin{array}{rrcl}
-    \ftext{Runtime environment} & \widetilde\rho &::=& \cdot \mid \rho,x \mapsto \val\\
+    \ftext{Runtime environment} & \widetilde\rho &::=& \cdot \mid \widetilde\rho,x \mapsto \val\\
 \end{array}
@@
 \mulop_{\mathcal{P}} \widetilde{e_{2}}$ to $n_{1} \mulop n_{2}$, similar to the
-source language. In contrast, the rule
+source semantics of \figureref{srcsem}. In contrast, the rule
 ({\sc{S-SMult}}) evaluates a secret multiplication $\widetilde{e_{1}}
@@
 is more performant using an arithmetic circuit compared to a boolean
-one. Rules ({\sc{S-PCond}}) and ({\sc{S-SCond}})
-are along the similar lines. Rule ({\sc{S-PCond}}) evaluates a public
+one~\cite{aby}. Rules ({\sc{S-PCond}}) and ({\sc{S-SCond}})
+are along similar lines. Rule ({\sc{S-PCond}}) evaluates a public
 conditional to the value from one of the branches, while the rule
-({\sc{S-SCond}}) evaluates to a \kw{mux} gate that takes input
+({\sc{S-SCond}}) evaluates to a multiplexer \kw{mux} gate that takes input
 circuits from the guard ($\gate_{1}$) and both the branches
@@
 $\lcond{\mathcal{A}}{e_{1}}{e_{2}}{e_{3}}$. Rules
-({\sc{S-Coerce}}) and ({\sc{S-Inp}}) evaluate to the coerce and input
+({\sc{S-Coerce}}) and ({\sc{S-Inp}}) evaluate to coerce and input
 gates respectively.
@@
 Statement evaluation $\widetilde{\rho} \vdash \widetilde{s} \Downarrow
-\widetilde{\rho}_{1}; \chi$ evaluates statements $\widetilde{s}$ to produce a new
+\widetilde{\rho}_{1}; \chi$ evaluates a statement $\widetilde{s}$ under 
+the environment $\widetilde{\rho}$  to produce a new
 environment $\widetilde{\rho}_{1}$, and a circuit $\crct$. A circuit $\crct$ (shown
@@
 \subsubsection*{Circuit semantics} Evaluating a program in the intermediate language produces a
-circuit over the secret data to be computed using an \mpc protocol in the distributed setting. With
-our circuit semantics, we model the \emph{functional} aspect of a \mpc protocol, parametrized by the sharing functions.
+circuit over the secret data to be computed using a \mpc protocol in the distributed setting. With
+our runtime circuit semantics, we model the \emph{functional} aspect of a \mpc protocol, parametrized by cryptographic encoding and decoding functions.
 
@@
 More concretely,
-$\mathcal{E}_{m}(c)$ returns a pair of two string $b_{1}$ and $b_{2}$
+$\mathcal{E}_{m}(c)$ returns a pair of two strings $b_{1}$ and $b_{2}$
  with the property that
-$\mathcal{D}_{m}(b_{1}, b_{2}) = c$. We assume that the underlying \mpc
+$\mathcal{D}_{m}(b_{1}, b_{2}) = c$. 
+The string $b_i$ denotes the $i^{th}$ party's {\it share} of $c$.
+We assume that the underlying \mpc
 protocol instantiates $\mathcal{E}_{m}$ and $\mathcal{D}_{m}$
@@
 
-\figureref{cktsem} gives the judgements for evaluation of circuit by the two parties using the \mpc protocol. The circuit environment for party $j$ is defined as follows:
+\figureref{cktsem} gives the judgements for evaluation of circuits by the two parties using the \mpc protocol. The circuit environment  is a map from wires to shares and is defined as follows:
 
@@
 \begin{array}{rrcl}
-    \ftext{Circuit environment} & \widehat\rho_j &::=& \cdot \mid \widehat\rho_j,\wire \mapsto b_j\\
+    \ftext{Circuit environment} & \widehat\rho &::=& \cdot \mid \widehat\rho,\wire \mapsto b\\
 \end{array}
@@
 
-\divya{strings $b$ are not formally defined.}
-
+We use $\widehat\rho_j$ to denote the circuit environment for party $j.$ 
 We give the 
@@
 evaluates, of course, concretely $n_{1}$ and $n_{2}$ are not observed
-by the parties. Rule ({\sc{C-Coerce}}) creates the shares using the
-scheme for $m$. Coming to the evaluation of circuits, the evaluation
+by the parties. Rule ({\sc{C-Coerce}}) creates the shares using  $\mathcal{E}_m$. The evaluation
 of \kw{bind} updates the mapping of $\wire$ in the input environments, and
@@
   \]
-\caption{Circuit semantics (selected rules)}
+\caption{Circuit semantics in a distributed runtime(selected rules)}
 \label{fig:cktsem}
@@
 
-\subsubsection*{Correctness theorem} We first define a judgment for
-translation of environments, $\Gamma \vdash \rho \leadsto
-\widetilde{\rho}; \widehat{\rho}_{1}, \widehat{\rho}_{2}$. The
-judgment is defined inductively on $\Gamma$, and establishes that
-the static environment $\Gamma$ is consistent with the runtime
-environments $\rho$, $\widetilde{\rho}$, and $\widehat{\rho}_{1}$ and
-$\widehat{\rho}_{2}$. As an example, if $x:\kw{uint}^{\mathcal{P}} \in
-\Gamma$, the judgment checks that $\rho(x) = \widetilde{\rho}(x) =
-n$, for some $n$.
-\divya{This judgement is not clear.
-If I am correct, it says that for public values $\rho$ and $\widetilde{\rho}$ are consistent and for secret values $\rho$ has values and $\widehat{rho}_j$ contain shares.} We elide the judgment, providing the full definition
-in the supplementary material.
-
-With this judgment in hand, the correctness theorem is as follows:
+%\subsubsection*{Correctness theorem}
+Finally, the correctness theorem is as follows (the environments on the left of $\vdash$ are empty and we elide the environments on the right of $\vdash$ for brevity):
 
 \begin{theorem}[Correctness]\label{theorem:correctness}
-  $\forall\:\Gamma, s, \widetilde{s}, \Gamma_{1}, \rho, \widetilde{\rho}, \widehat{\rho}_{1}, \widehat{\rho}_{2}$, if:
-
-  \begin{enumerate}
-  \item $\Gamma \vdash s \leadsto \widetilde{s} \mid \Gamma_{1}$
-  \item $\Gamma \vdash \rho \leadsto \widetilde{\rho}; \widehat{\rho}_{1}, \widehat{\rho}_{2}$
-  \end{enumerate}
-
-  then $\exists \rho_{1}, O_{1}, \widetilde{\rho}_{1}, \chi, \widehat{\rho}'_{1}, \widehat{\rho}'_{2}, O_{2}$, s.t.
-
-  \begin{enumerate}[label=(\alph*)]
-  \item $\rho \vdash s \Downarrow \rho_{1}; O_{1}$
-  \item $\widetilde{\rho} \vdash \widetilde{s} \Downarrow \widetilde{\rho}_{1}; \chi$
-  \item $\widehat{\rho}_{1}, \widehat{\rho}_{2} \vdash \chi \Downarrow \widehat{\rho}'_{1}, \widehat{\rho}'_{2}; O_{2}$
-  \item $\Gamma_{1} \vdash \rho_{1} \leadsto \widetilde{\rho}_{1}; \widehat{\rho}'_{1}, \widehat{\rho}'_{2}$
-  \item $O_{1} = O_{2}$
-  \end{enumerate}
-
+ $\forall\: s, \widetilde{s}$, if $\vdash s \leadsto \widetilde{s} \mid \_$  then $\exists  O , \chi$, s.t.
+$\vdash s \Downarrow \_; O$, $ \vdash \widetilde{s} \Downarrow \_; \chi$, and $ \vdash \chi \Downarrow \_, \_; O$
 \end{theorem}
@@
 compiles to $\widetilde{s}$ in intermediate language, then $s$ terminates in the source
-semantics with observations $O_{1}$, $\widetilde{s}$ evaluates to circuit $\crct$, and $\crct$ terminates in the
-circuit semantics with observations $O_{2}$, where $O_{1} = O_{2}$
-Furthermore, the final environments are also in the compilation
-relation. \divya{Should we call the judgement between environs as compilation relation?}
-Due to space reasons, we provide the proof in \cite{anonymized}.
+semantics with observations $O$, $\widetilde{s}$ evaluates to circuit $\crct$, and $\crct$ terminates in the
+circuit semantics with the same observations $O$. 
+Therefore, well typed programs always terminate (array indices are always in bounds, there are no unbounded loops, etc.) and the protocol produces the same outputs as the source semantics.
+% We first define a judgment for
+%translation of environments, $\Gamma \vdash \rho \leadsto
+%\widetilde{\rho}; \widehat{\rho}_{1}, \widehat{\rho}_{2}$. The
+%judgment is defined inductively on $\Gamma$, and establishes that
+%the static environment $\Gamma$ is consistent with the runtime
+%environments $\rho$, $\widetilde{\rho}$, and $\widehat{\rho}_{1}$ and
+%$\widehat{\rho}_{2}$. 
+%As an example, if $x:\kw{uint}^{\mathcal{P}} \in
+%\Gamma$, the judgment checks that $\rho(x) = \widetilde{\rho}(x) =
+%n$, for some $n$. On the other hand, if $x:\kw{uint}^{m} \in
+%\Gamma$ then the judgement checks that $\mathcal{D}_m( \widehat{\rho}_{1}(\widetilde{\rho}(x)),\widehat{\rho}_{2}(\widetilde{\rho}(x)))=\rho(x)$.
+%%We elide the judgment, providing the full definition
+%%in the supplementary material.
+%With this judgment,
+%the correctness theorem is as follows:
+%
+%\begin{theorem}[Correctness]\label{theorem:correctness}
+%  $\forall\:\Gamma, s, \widetilde{s}, \Gamma_{1}, \rho, \widetilde{\rho}, \widehat{\rho}_{1}, \widehat{\rho}_{2}$, if:
+%
+%  \begin{enumerate}
+%  \item $\Gamma \vdash s \leadsto \widetilde{s} \mid \Gamma_{1}$
+%  \item $\Gamma \vdash \rho \leadsto \widetilde{\rho}; \widehat{\rho}_{1}, \widehat{\rho}_{2}$
+%  \end{enumerate}
+%
+%  then $\exists \rho_{1}, O_{1}, \widetilde{\rho}_{1}, \chi, \widehat{\rho}'_{1}, \widehat{\rho}'_{2}, O_{2}$, s.t.
+%
+%  \begin{enumerate}[label=(\alph*)]
+%  \item $\rho \vdash s \Downarrow \rho_{1}; O_{1}$
+%  \item $\widetilde{\rho} \vdash \widetilde{s} \Downarrow \widetilde{\rho}_{1}; \chi$
+%  \item $\widehat{\rho}_{1}, \widehat{\rho}_{2} \vdash \chi \Downarrow \widehat{\rho}'_{1}, \widehat{\rho}'_{2}; O_{2}$
+%  \item $\Gamma_{1} \vdash \rho_{1} \leadsto \widetilde{\rho}_{1}; \widehat{\rho}'_{1}, \widehat{\rho}'_{2}$
+%  \item $O_{1} = O_{2}$
+%  \end{enumerate}
+%  
+%\end{theorem}
+%
+%The theorem states that if a source statement $s$ is well-typed, and
+%compiles to $\widetilde{s}$ in intermediate language, then $s$ terminates in the source
+%semantics with observations $O_{1}$, $\widetilde{s}$ evaluates to circuit $\crct$, and $\crct$ terminates in the
+%circuit semantics with observations $O_{2}$, where $O_{1} = O_{2}$
+%%Furthermore, the final environments are also in the compilation
+%%relation. \divya{Should we call the judgement between environs as compilation relation?}
+%Due to space reasons, we provide the proof in \cite{anonymized}.

### 2017-10-29 OaklandSubmission/security.tex
minor

@@
 We assume that there exists a \mpc protocol $\prot$ that securely realizes the functionality $\crct$ and let $\simu_{\tiny{2pc}}$ be the corresponding simulator  (that runs on $\crct$, input of the corrupt party and the output obtained from trusted functionality for $\crct$).
-We note that the work of \cite{aby} provides such a protocol $\prot$ and simulator $\simu_{\tiny{2pc}}$ for all circuits $\crct$ output by our compiler\footnote{\divya{We note that \cite{aby} provides a \mpc protocol for a more general .....}}.
+We note that the work of \cite{aby} provides such a protocol $\prot$ and simulator $\simu_{\tiny{2pc}}$ for all circuits $\crct$ output by our compiler.
 We are now ready to state and prove our security theorem.

### 2017-10-29 OaklandSubmission/eval.tex
minor

@@
 A more interesting classifier is {\em Na\"{i}ve Bayes} that predicts labels
-from the set $[n]$.
+from the set $[n]$~\cite{shafindss}.
 Here, the input data point is a {\it feature}
 vector $x=(x_0,x_2,\ldots,x_{d-1})^T$ where each $x_j\in [F]$.
-The model has two matrices: a vector $P$ of length $n$ s.t. for each
-$i
-\in [n]$, $P(i)$ is the log-probability that the output label is $i$.
+%The model has two matrices: a vector $P$ of length $n$ s.t. for each
+%$i
+%\in [n]$, $P(i)$ is the log-probability that the output label is $i$.
 %$P(i) \equiv \log p(\ell=i)$, where $p(\ell=i)$ is the likelihood
 %that the output label is $i$ for $i\in[n]$.
-The other matrix $T$ has size $d\times F\times n$ and for each $j\in
-[d], k\in [F], i\in [n]$, the entry $T(j)(k)(i)$ is 
+%The other matrix $T$ has size $d\times F\times n$ and for each $j\in
+%[d], k\in [F], i\in [n]$, the entry $T(j)(k)(i)$ is 
 %\equiv\log p(x_j|\ell=i)$, i.e.,
-the log-probability that  the $j^{th}$ input feature is $k$  conditioned on  the output label being $i$.
-The classifier $C_{P,T}(x)$ outputs
-\[
-{\sf argmax}\ P+\sum_{j=0}^{d-1} T(j)(x_j)
-\]
-The additions in the expression above sum $n$-dimensional vectors.
-The model size of this classifier is $|P|+|T|$.
+%the log-probability that  the $j^{th}$ input feature is $k$  conditioned on  the output label being $i$.
+%The classifier $C_{P,T}(x)$ outputs
+%\[
+%{\sf argmax}\ P+\sum_{j=0}^{d-1} T(j)(x_j)
+%\]
+%The additions in the expression above sum $n$-dimensional vectors.
+%The model size of this classifier is $|P|+|T|$.
 This classifier requires $\Theta(ndF)$ comparisons and additions.
-Note here  that computing $T(j)(x_j)(i)$ requires a secret look-up in matrix  $T$  based on secret input $x_j$ resulting in $F-1$ comparisons using multiplexers (see
-  Section~\ref{sec:impl}).
+%Note here  that computing $T(j)(x_j)(i)$ requires a secret look-up in matrix  $T$  based on secret input $x_j$ resulting in $F-1$ comparisons using multiplexers (see
+  %Section~\ref{sec:impl}).
 %\nc{Made DG's comment as part of regular text - worth point out}.
 \cmmt{Is there a +1 etc missing or is this exact?} \nc{By our calculation in Sec 5, it seems that this should be $F/2$ multiplexers}
-
-A {\em decision tree} of size $N$ takes as input an $x\in\R^d$, and the
-model consists of a binary tree of depth $d$ with a boolean predicate
-assigned to each internal node. The root note is the $0^{th}$ node and
-for an internal node $j \in [N]$, the children nodes are $2j+1$ and
-$2j+2$. Each internal node $j\in[N]$ at depth $i$ has a predicate
-$b_{j}^i\equiv x_i\leq w_{j}$. We start evaluating the tree from the
-root and if the predicate at the current node $b_j^i$ is false (resp.,
-true) then $x$ is passed to the left (resp., right) child with
-predicate $b_{2j+1}^{i+1}$ (resp., $b_{2j+2}^{i+1}$). This process is
-repeated till we reach a leaf. The leaves of the tree are labels and
-the output label is the leaf visited by this traversal.
-Such a (binary) decision tree can be encoded as a polynomial (of
-degree linear in depth of the tree) and then the prediction task
-reduces to polynomial evaluation. E.g., a binary tree with depth one
-and size $N=3$ can be encoded as the polynomial
-$(1-b(x))\ell_0 + b(x)\ell_1$ where $b(x)$ is the predicate at the
-root, $\ell_0$ is the label of the left leaf and $\ell_1$ is the label
-of the right leaf.
+%
+A {\em decision tree} of size $N$ takes as input an $x\in\R^d$
+%,
+ and 
+%the
+%model consists of a binary tree of depth $d$ with a boolean predicate
+%assigned to each internal node.
+% The root note is the $0^{th}$ node and
+%for an internal node $j \in [N]$, the children nodes are $2j+1$ and
+%$2j+2$. Each internal node $j\in[N]$ at depth $i$ has a predicate
+%$b_{j}^i\equiv x_i\leq w_{j}$. We start evaluating the tree from the
+%root and if the predicate at the current node $b_j^i$ is false (resp.,
+%true) then $x$ is passed to the left (resp., right) child with
+%predicate $b_{2j+1}^{i+1}$ (resp., $b_{2j+2}^{i+1}$). This process is
+%repeated till we reach a leaf. The leaves of the tree are labels and
+%the output label is the leaf visited by this traversal.
+%Such a (binary) decision tree 
+%and can be encoded as a polynomial (of
+%degree linear in depth of the tree) and then
+ the prediction task
+reduces to polynomial evaluation~\cite{shafindss}.
+% E.g., a binary tree with depth one
+%and size $N=3$ can be encoded as the polynomial
+%$(1-b(x))\ell_0 + b(x)\ell_1$ where $b(x)$ is the predicate at the
+%root, $\ell_0$ is the label of the left leaf and $\ell_1$ is the label
+%of the right leaf.
 
@@
 that provides an accuracy of 92\%. The classifier evaluates
-\[
- \mathsf{argmax}\ W\matmul x+b
-\]
+$ \mathsf{argmax}\ W\matmul x+b$.
 Here, $x$ is a 784 length vector obtained from the input image,
@@
 \label{sec:pipeeval}
+We evaluate partitioning next.
+\subsubsection*{CIFAR-10}
 The largest benchmark of \minion~\cite{minionn} is a DNN for CIFAR-10

### 2017-10-29 OaklandSubmission/eval.tex
eval

@@
 \label{sec:shallow}
-We train (on cleartext data) the three standard classifiers, linear, Na\"{i}ve Bayes, and
-decision trees, on the following data sets from the UCI
+We evaluate the three standard classifiers, linear, Na\"{i}ve Bayes, and
+decision trees, from~\cite{shafindss} on the following data sets from the UCI
 machine learning repository~\cite{uci}:

### 2017-10-30 OaklandSubmission/intro.tex
till correctness
nits

@@
 successfully terminate
-(e.g. there are no array index out-of-bounds errors), and that the
-observable outputs in the two semantics are
-identical. For
+(e.g. there are no array index out-of-bounds errors), with identical
+observable outputs. For
 the security theorem, we formally reduce the security
 of our compiler to the security of the  \mpc back-end
- (by also invoking our correctness theorem). We also prove a formal security theorem for our partitioning
-scheme.
+ (by also invoking our correctness theorem). We also prove a formal security theorem against semi-honest (or, ``honest but curious") adversaries for our partitioning
+scheme (Section~\ref{sec:ld} and Section~\ref{sec:pipe}).
 
-\subsubsection*{Evaluation} We have implemented the \tool toolchain
+\subsubsection*{Evaluation} We have implemented \tool
 using ABY~\cite{aby} as the
@@
 cases) their state-of-the-art specialized implementations. Finally, we
-show that with partitioning, we are able to implement applications
-that require more than 300 million gates.
+show that partitioning enables us to implement applications
+that require more than 300 million gates (Section~\ref{sec:impl} and
+Section~\ref{sec:eval}).
 
@@
 
-The rest of the paper is structured as follows. We formalize \tool and
-prove its correctness and security theorems in
-Section~\ref{sec:ld}. Section~\ref{sec:pipe} details our partitioning
-scheme and proves its security. Section~\ref{sec:impl}
-presents the implementation details, and
-Section~\ref{sec:eval} present a comprehensive evaluation of \tool. We
-begin with an overview of \tool through an example in
-Section~\ref{sec:ex}. %\nc{provided a pointer to sec 2 as well.}
+%% The rest of the paper is structured as follows. We formalize \tool and
+%% prove its correctness and security theorems in
+%% Section~\ref{sec:ld}. Section~\ref{sec:pipe} details our partitioning
+%% scheme and proves its security. Section~\ref{sec:impl}
+%% presents the implementation details, and
+%% Section~\ref{sec:eval} present a comprehensive evaluation of \tool. We
+%% begin with an overview of \tool through an example in
+%% Section~\ref{sec:ex}. %\nc{provided a pointer to sec 2 as well.}
 %\aseem{Sure :), I did not add it because Sec. 2 was just next line, so

### 2017-10-30 OaklandSubmission/conclude.tex
cutting

@@
-\section{Conclusions}
+\section{Conclusion}
 \label{sec:conclude}
-
-Therefore, we believe
-that \tool generated code can be a better baseline than Yao-based 2PC
-when comparing against generic 2PC-based implementations for secure
-prediction tasks. \aseem{I don't understand this last sentence. Why is
-  ours a better baseline?} \nc{This last line is more a comment for a ``conclusion'' and is a bit out of place here.
-I guess we want to say that till recently if generic 2PC using Yao was not feasible, then people had to come up with specialized protocols and didnt have any benchmark or had a poor benchmark (of Yao) to compare their 2PC with.
-Moving forward, our language generated 2PC can be the baseline benchmark for all 2PC protocols.}
+\tool is the first
+``cryptographic-cost aware'' framework that generates efficient and
+scalable \mpc protocols from high-level programs.
+The generated protocols comprise combinations
+of arithmetic and boolean circuits and have performance comparable to
+or better than the custom, specialized
+protocols from previous works

### 2017-10-30 OaklandSubmission/eval.tex
cutting

--- 2017-10-29 OaklandSubmission/eval.tex
@@
 data point which is a tuple of humidity and temperature 
-a classifier might be able to predict a label as ``will rain'' or ``will not rain''. The {\it
+a classifier might be able to predict a label ``will rain'' or ``will not rain''.
+The {\it model size} of a classifier is the number of parameters in the model.
+For example, the model size of the classifier in~\figureref{ex-sml} is $|w|+1=31$.
+The {\it
   accuracy} of a classifier refers to the fraction of data points that
@@
 $\ell\in\{\mathit{true},\mathit{false}\}$ and the classifier is
-$C_w\equiv w^T\matmul x>0$, where $w^T$ denotes the transpose
-of matrix $w$, $\matmul$ denotes matrix multiplication (with a vector in this case) and $a>b$ returns $\mathit{true}$ if $a>b$ and $\mathit{false}$ otherwise. 
-The {\it model size} of this classifier, i.e., the number of parameters in the model is $|w|$, i.e., $d$.
-This classifier requires $d$ multiplications, $d-1$
-additions, and a single comparison.
-
+$C_w\equiv w^Tx>0$.
+%, where $w^T$ denotes the transpose
+%of matrix $w$, $\matmul$ denotes matrix multiplication (with a vector in this case) and $a>b$ returns %$\mathit{true}$ if $a>b$ and $\mathit{false}$ otherwise. 
+%The {\it model size} of this classifier, i.e., the number of parameters in the model is $|w|$, i.e., $d$.
+%This classifier requires $d$ multiplications, $d-1$
+%additions, and a single comparison.
+%
 A more interesting classifier is {\em Na\"{i}ve Bayes} that predicts labels
@@
 are matrices of appropriate dimensions. The activation function $f$ is given by $f(y) = y$ if
-$-1 < y < 1$ and $\mathrm{sign}(y)$ otherwise ($f$ on vectors and matrices is defined point-wise on every element of the vector/matrix). \bonsai can be
-seen as a variant of decision trees where the prediction is a function
-of the path traversed from root to the leaf and not just the leaf
-itself.
+$-1 < y < 1$ and $\mathrm{sign}(y)$ otherwise ($f$ on vectors and matrices is defined point-wise on every element of the vector/matrix).
+%\bonsai can be
+%seen as a variant of decision trees where the prediction is a function
+%of the path traversed from root to the leaf and not just the leaf
+%itself.
 
@@
 
-\subsection{Standard classifiers}
-\label{sec:shallow}
-We train (on cleartext data) the three standard classifiers, linear, Na\"{i}ve Bayes, and
-decision trees, on the following data sets from the UCI
+\subsection{Secure prediction}
+\noindent\textbf{Standard classifiers}
+%\label{sec:shallow}
+We evaluate the three standard classifiers, linear, Na\"{i}ve Bayes, and
+decision trees, from~\cite{shafindss} on the following data sets from the UCI
 machine learning repository~\cite{uci}:
@@
 
-\subsubsection*{Linear classifier}The results for linear
-classification are in Table~\ref{tab:lc}. The
-input and the model are both vectors of length $d$. The columns
-``Prev. time'' and ``Prev. comm'' show the time and the total
+The results for linear
+classification are in Table~\ref{tab:lc}.
+The input and the model are both vectors of length $d$. 
+The columns ``Prev. time'' and ``Prev. comm'' show the time and the total
 network communication reported by Bost et al.~\cite{shafindss} for a
@@
 
-\subsubsection*{Na\"{i}ve Bayes} The results for Na\"{i}ve Bayes are
-in Table~\ref{tab:nb}. As before, $d$ denotes the length of the input
-feature vector $x$ and $F$ is the number of possible values of $x_i$
-for each $i \in [d]$.
-Although, there are no multiplications in these benchmarks, these do
-have a significant number of comparisons (due to array look-ups using
-secret indices) that
-result in many boolean-and gates. %(13k/750k denote 13000/75000). 
-As before columns ``Prev. time'' and ``Prev. comm.'' (columns 4 and 5)
-show the numbers reported
-by Bost et al.~\cite{shafindss}. We observe that \tool generated code
-has better performance, despite using a generic 2PC,
-as opposed to custom designed protocols developed by Bost et
-al. Moreover, they remark that in their setup, generic Yao-based 2PC
-did not scale to the smallest of their Na\"{i}ve Bayes classifiers, so
-they had to scale down the prediction task, and even then their
-implementation was 500x slower. Whereas, we show that by using a
-cryptographic-cost aware compiler, we can scale generic 2PC to real
-prediction tasks, and get performance competitive to or better than the
-specialized protocols.
+%\subsubsection*{Na\"{i}ve Bayes}
 
@@
 
-\subsubsection*{Decision trees}
-Our decision tree numbers in Table~\ref{tab:dt} further validate
-this claim. We compare with the recent work of Wu et al.~\cite{wu} (columns 3 and 4) who
-describe specialized protocols for trees and forests (that perform
-better than~\cite{shafindss}).
-Similar to Bost et al.~\cite{shafindss}, they also remarks that these
-computations are infeasible with Yao-based 2PC.
-Again, we observe that
-the performance of \tool generated generic 2PC
-code is competitive with specialized protocols.
+%\subsubsection*{Decision trees}
 
@@
 %dataset | time | comm | timeL | timeW | comm
-\subsection{Deep neural nets}
-\label{sec:dnneval}
+%Our decision tree numbers in Table~\ref{tab:dt} further validate
+%this claim. 
+%We compare with the recent work of Wu et al.~\cite{wu} (columns 3 and 4) who
+%describe specialized protocols for trees and forests (that perform
+%better than~\cite{shafindss}).
+%Similar to Bost et al.~\cite{shafindss}, they also remarks that these
+%computations are infeasible with Yao-based 2PC.
+%Again, we observe that
+%the performance of \tool generated generic 2PC
+%code is competitive with specialized protocols. 
+The results for Na\"{i}ve Bayes are
+in Table~\ref{tab:nb}. As before, $d$ denotes the length of the input
+feature vector $x$ and $F$ is the number of possible values of $x_i$
+for each $i \in [d]$.
+%Although, there are no multiplications in these benchmarks, these do
+%have a significant number of comparisons (due to array look-ups using
+%secret indices) that
+%result in many boolean-and gates. %(13k/750k denote 13000/75000). 
+As before columns ``Prev. time'' and ``Prev. comm.'' (columns 4 and 5)
+show the numbers reported
+by Bost et al.~\cite{shafindss}. We observe that \tool generated code
+has better performance, despite using a generic 2PC,
+as opposed to custom designed protocols developed by Bost et
+al. Moreover, they remark that in their setup, generic Yao-based 2PC
+did not scale to the smallest of their Na\"{i}ve Bayes classifiers, so
+they had to scale down the prediction task, and even then their
+implementation was 500x slower. Whereas, we show that by using a
+cryptographic-cost aware compiler, we can scale generic 2PC to real
+prediction tasks, and get performance competitive to or better than the
+specialized protocols. Table~\ref{tab:dt} 
+compares against the more recent work of~\cite{wu}
+on decision trees and further validates this claim.
+\\\\
+\noindent\textbf{Deep neural nets}
 We evaluate \tool on the DNNs described in SecureML~\cite{secureml},
@@
 
-In Table~\ref{tab:nn}, columns 2 and 3 show the time taken by \minion.
-The column ``Model size'' is the number of parameters in the
-trained model.
-We observe that our performance is competitive with specialized
-\minion protocols, for both LAN and the WAN settings. Further, lines
-of \tool source code required is still small. 
-We note that \minion and \tool implementations are both based on ABY.
-These results are surprising as \minion has a much more efficient
-preprocessing phase and has SIMD 
-(single instruction multiple data) capabilities that allow it to perform
-matrix operations efficiently. In contrast, \tool focuses on generic
-2PC, and such optimizations are part of our future work.
-\minion also reports performance results on a bigger DNN with 7
-convolution layers. In \tool, this benchmark requires partitioning and
-we discuss it in Section~\ref{sec:pipeeval}.
 %net | minionT | minionC | timeL | timeW | comm | gates | and | mul | length
@@
 
-\subsection{State-of-the-art classifiers}
+In Table~\ref{tab:nn}, columns 2 and 3 show the time taken by \minion.
+The column ``Model size'' is the number of parameters in the
+trained model.
+We observe that our performance is competitive with specialized
+\minion protocols, for both LAN and the WAN settings. Further, lines
+of \tool source code required is still small. 
+We note that \minion and \tool implementations are both based on ABY.
+These results are surprising as \minion has a much more efficient
+preprocessing phase and has SIMD 
+(single instruction multiple data) capabilities that allow it to perform
+matrix operations efficiently. In contrast, \tool focuses on generic
+2PC, and such optimizations are part of our future work.
+\minion also reports performance results on a bigger DNN with 7
+convolution layers. In \tool, this benchmark requires partitioning and
+we discuss it in Section~\ref{sec:pipeeval}.
+\\\\
+\noindent\textbf{State-of-the-art classifiers}
 Tensorflow~\cite{tensorflow} is a standard machine learning toolkit.
@@
 \setlength\tabcolsep{6pt}
-\subsection{Partitioning}
+\subsection{Secure code partitioning}
 \label{sec:pipeeval}
-We evaluate partitioning next.
-\subsubsection*{CIFAR-10}
+\noindent\textbf{CIFAR-10}
 The largest benchmark of \minion~\cite{minionn} is a DNN for CIFAR-10
@@
 
-Table~\ref{tab:cifar} shows the end-to-end numbers as
-well as the numbers for the sixth (the heaviest) stage. 
-The number of gates are in millions, hence the suffix `m' in the last
-three columns.
-For this DNN, the specialized \minion protocol takes 544 seconds and
-communicates 9272 Mb.
-As with Table~\ref{tab:nn}, \tool generated generic 2PC protocol is
-competitive with \minion here as well. Therefore,
-we believe that with partitioning, \tool can scale to arbitrary sized
-computations while maintaining performance
-competitive with existing specialized protocols. In particular, for a
-large enough DNN, \minion could run
-out of memory but an appropriately partitioned \tool implementation would still
-succeed. 
 \begin{table}
@@
 
+Table~\ref{tab:cifar} shows the end-to-end numbers as
+well as the numbers for the sixth (the heaviest) stage. 
+The number of gates are in millions, hence the suffix `m' in the last
+three columns.
+For this DNN, the specialized \minion protocol takes 544 seconds and
+communicates 9272 Mb.
+As with Table~\ref{tab:nn}, \tool generated generic 2PC protocol is
+competitive with \minion here as well. Therefore,
+we believe that with partitioning, \tool can scale to arbitrary sized
+computations while maintaining performance
+competitive with existing specialized protocols. In particular, for a
+large enough DNN, \minion could run
+out of memory but an appropriately partitioned \tool implementation would still
+succeed.
+\\\\
+\noindent\textbf{Matrix factorization}
+\tool, as a language, is not tied to secure prediction and can express more general computations.
+To demonstrate the expressiveness of \tool, we use it to implement secure matrix factorization~\cite{valeriaMatrix}. Abstractly, given a sparse matrix $\mathcal{M}$ of dimensions
+$n\times m$ and $M$ non-zero entries, the goal is to generate a matrix $U$ of dimension $n\times d$ and a matrix
+$V$ of dimension $d\times m$ such that $\mathcal{M}\approx UV$. This operator is useful in recommender systems.
+In particular, Nikolaenko et al.~\cite{valeriaMatrix} shows how to implement a movie recommender system which does not require users to reveal their data in the clear, i.e., the ratings the users have assigned to movies are kept secret.
+The implementation is a two party computation of an iterative algorithm for matrix factorization (Algorithm~1 in~\cite{valeriaMatrix}).
+This algorithm is based on gradient descent and iteratively converges to a local minima.
+We implement this algorithm in \tool.
+
+To ensure that the algorithm converges to the right local minima, Nikolaenko et al. require
+36 bits of precision. Since ABY supports either 32-bit or 64-bit integers, our \tool implementation
+manipulates 64-bit variables.
+For the matrix $\mathcal{M}$ of  user data, Nikolaenko et al.
+consider $n=940$ users, $m=40$ most popular movies, and $M=14683$ ratings from the MovieLens dataset.
+The time reported in~\cite{valeriaMatrix}
+for one iteration is 2.9 hours\footnote{~\cite{valeriaMatrix} does not report the network round-trip time.}. This compuation is large enough that we partition each iteration
+into three stages. The first stage involves a Batcher~\cite{Batcher} sorting network followed by a linear pass.
+The second stage involves sorting and gradient computations and is the heaviest stage.
+The third  stage is similar to the first stage. The results are presented in Table~\ref{tab:factor}.
+
+\begin{table}
+\begin{tabular}{|c|c|c|c |c|c| c|}
+\hline
+  Stage         &  LAN (s) & WAN (s) & \thead{Comm. \\ (Mb)}  & depth & \#Gates & LOC\\
+\hline
+1    &  175       & 662        & 29816       & 16370    & 33m    & 500  \\
+\hline
+2    &  193        & 1095        & 31945        & 30916    & 37m & 516 \\
+\hline
+3    &  178        & 627        & 29810        & 16369    & 32m  & 478  \\
+\hline
+Total    &  546      & 2384        & 91571        & --    & 102m & 1494 \\
+\hline
+\end{tabular}
+
+ \caption{Partitioning results for matrix factorization. The time reported by~\cite{valeriaMatrix} for this computation is 10440 seconds.}
+ \label{tab:factor} 
+\end{table}
+
+We observe that in the LAN setting, we are about 19 times faster than~\cite{valeriaMatrix} and in the WAN
+setting we are about 4 times faster. The main source of these significant speedups is that, unlike~\cite{valeriaMatrix}, \tool does not need to convert the functionality into boolean circuits. 
+However, this benchmark requires more lines of code than the previous benchmarks
+because of  Batcher's sort (450 lines of \tool code in each stage).
+However, the current programmer effort seems miniscule compared to the mammoth implementation effort
+put in by Nikolaenko et al. (Section 5 of ~\cite{valeriaMatrix}) to scale a boolean circuit based
+crypto-backend to this benchmark.
+
 %stage | time | comm | gates | mul | add | depth

### 2017-10-30 OaklandSubmission/impl.tex
cutting

--- 2017-10-28 OaklandSubmission/impl.tex
@@
 (SeaHorn~\cite{seahorn}) to bound check the array indices
-($P\vDash e < n$
+($\models e < n$
 in ({\sc {T-Read}}) and ({\sc{T-Write}})),
@@
 boolean and vice versa.
-The \tool implementation performs  the standard ``common subexpression elimination"
-optimization~\cite{dragonbook} to reduce these coercions.
+Since \tool is aware of the cryptographic costs associated with these coercions,
+it tries to minimize them using several optimizations.
+For instance, if boolean shares of a variable are available in scope and the variable is involved in
+an addition then \tool performs the addition in boolean rather than first coercing boolean to arithmetic
+and then performing an arithmetic addition. However, multiplication is always performed in arithmetic
+as the cost of a boolean multiplication is much higher than the cost of coercing from boolean to arithmetic, performing an arithmetic multiplication, and then coercing from arithmetic to boolean~\cite{aby}.
+Other optimizations include the standard ``common subexpression elimination"
+optimization~\cite{dragonbook}.
 On each coercion, \tool memorizes the pair of arithmetic

### 2017-10-30 OaklandSubmission/app-stash.tex
eval 6.1
intro related
minor

@@
 
-\divya{points that have to be there...} Security; traces are ``oblivious'' of secret values if observables/outputs are same; formal cryptographic security guarantee uses simulation paradigm \cite{canetti2000}
+\divya{points that have to be there...} Security; traces are ``oblivious'' of secret values if observables/outputs are same; formal cryptographic security guarantee uses simulation paradigm
 
@@
 
+\divya{references:} Lets cite CCS paper for Minionn and not eprint
+
+\divya{experiments:} Should we do matrix factorization?
+
 \divya{where to put this..

### 2017-10-30 OaklandSubmission/conclude.tex
minor

@@
 \tool is the first
-``cryptographic-cost aware'' framework that generates efficient and
+cryptographic-cost aware framework that generates efficient and
 scalable \mpc protocols from high-level programs.
@@
 or better than the custom, specialized
-protocols from previous works
+protocols from previous works.

### 2017-10-30 OaklandSubmission/eval.tex
minor

@@
 and show that the performance of the high-level code written in \tool
-is comparable to their hand-crafted protocols
-(Section~\ref{sec:shallow} and Section~\ref{sec:dnneval}).
+is comparable to their hand-crafted protocols.
 %
@@
 implement a Deep Neural Network (DNN) for CIFAR-10
-dataset~\cite{cifar} from \minion~\cite{minionn} to illustrate the use
-of partitioning in \tool. Finally, we evaluate \tool on the task of matrix factorization~\cite{valeriaMatrix}.
+dataset~\cite{cifar} from \minion~\cite{minionn} and matrix factorization~\cite{valeriaMatrix}
+to illustrate the use
+of partitioning in \tool.
+%Finally, we evaluate \tool on the task of matrix factorization~\cite{valeriaMatrix}.
 
@@
 We use $[N]$ to denote $\{0,1,\dotsc, N-1\}$.
-Further, given a vector $x\in\R^d$, we say $\mathsf{argmax}\ x = i$ if $x_i = \mathsf{max}\ \{x_1,\ldots,x_d\}$.
+Further, given a vector $x\in\R^d$, we say $\mathsf{argmax}\ x = i$ if $x_i = \mathsf{max}\ \{x_0,\ldots,x_{d-1}\}$.
 We focus on the machine learning models for {\it classification}.
@@
 data point which is a tuple of humidity and temperature 
-a classifier might be able to predict a label ``will rain'' or ``will not rain''.
+a classifier can predict a label ``will rain'' or ``will not rain''.
 The {\it model size} of a classifier is the number of parameters in the model.
@@
  the prediction task
-reduces to polynomial evaluation~\cite{shafindss}.
+reduces to evaluation of a $d$-degree polynomial~\cite{shafindss}.
 % E.g., a binary tree with depth one
@@
 \end{tabular}
- \caption{Linear classification results. We compare our results (columns 5, 6, 7) with~\cite{shafindss} (columns 3 and 4)}
+ \caption{Linear classification results. We compare \tool (LAN, WAN, Comm)
+ with~\cite{shafindss} (Prev. Time, Prev. Comm).}
  \label{tab:lc} 
@@
 \end{tabular}
-\caption{Na\"{i}ve Bayes results. We compare our results (columns 6 , 7, 8) with~\cite{shafindss} (columns 4 and 5)}
+\caption{Na\"{i}ve Bayes results. We compare \tool (LAN, WAN, Comm)
+ with~\cite{shafindss} (Prev. Time, Prev. Comm).}
  \label{tab:nb} 
@@
 
-\begin{table}
+\begin{table}[t]
 \begin{tabular}{|c|c|c|c|c|c|c|c|c|}
@@
 \end{tabular}
-\caption{Decision tree benchmarks. We compare our results (columns 5 , 6, 7) with~\cite{wu} (columns 3 and 4)}
+\caption{Decision tree benchmarks.We compare \tool (LAN, WAN, Comm)
+ with~\cite{wu} (Prev. Time, Prev. Comm).}
  \label{tab:dt} 
@@
 did not scale to the smallest of their Na\"{i}ve Bayes classifiers, so
-they had to scale down the prediction task, and even then their
-implementation was 500x slower. Whereas, we show that by using a
+they had to scale down the prediction task, and even then Yao-based \mpc
+was 500x slower. Whereas, we show that by using a
 cryptographic-cost aware compiler, we can scale generic 2PC to real
@@
 It has three fully connected layers with square as the activation function.
-This DNN is light-weight as it requires only arithmetic computations
-(additions and multiplications) 
-except for the $\mathsf{argmax}$ at the end of the network.
+%This DNN is light-weight as it requires only arithmetic computations
+%(additions and multiplications) 
+%except for the $\mathsf{argmax}$ at the end of the network.
 Next, we implement the DNN described in Cryptonets~\cite{cryptonets} (Figure 11 in~\cite{minionn})
@@
 
- \caption{DNN benchmarks. We compare our results (columns 4 , 5, 6) with~\cite{minionn} (columns 2 and 3)}
+ \caption{DNN benchmarks. We compare \tool (LAN, WAN, Comm)
+ with~\cite{minionn} (Prev. Time, Prev. Comm).}
  \label{tab:nn} 
@@
 
-In Table~\ref{tab:nn}, columns 2 and 3 show the time taken by \minion.
-The column ``Model size'' is the number of parameters in the
+In Table~\ref{tab:nn}, 
+the column ``Model size'' is the number of parameters in the
 trained model.
 We observe that our performance is competitive with specialized
-\minion protocols, for both LAN and the WAN settings. Further, lines
+\minion protocols, for both the LAN and the WAN settings. Further, lines
 of \tool source code required is still small. 
@@
 In particular, it has an accuracy of 99.2\%.
-Since, we are not aware of any other tools that has used this
+Since, we are not aware of any other tools that have used this
 model as a benchmark, we only report numbers for \tool.
@@
 \end{table}
+\setlength\tabcolsep{3.5pt}
+\begin{table}
+\begin{tabular}{|c|c|c|c |c|c|c|c|c|c | c|}
+\hline
+Dataset       & \thead{LAN \\(s)} & \thead{WAN \\(s)} & \thead{Comm. \\ (Mb)}  & \#And & \#Mul & \#Gates & depth & LOC\\
+\hline
+Chars4k    &  0.1         & 0.7         & 2            & 18k    & 3k    &  85k     & 1   & 89\\
+\hline
+USPS       &  0.2         & 0.9         & 4            & 62k    & 2k    &  285k    & 2   & 156\\
+\hline
+WARD       &  0.3         & 1.1         & 9            & 106k    & 8k    &  506k    & 3   & 283\\
+\hline
+\end{tabular}
+
+ \caption{Bonsai benchmarks}
+ \label{tab:bonsai} 
+\end{table}
 
@@
 implementation requires a loop for each layer of the
-tree\footnote{We remark here that our current language does not support functions (which we leave for future work) and with this support, LOC would be even lower.}.
+tree\footnote{We remark here that our current language does not support functions (which we leave for future work) and with this support, LOC would be lower and independent of depth.}.
 
@@
 
-\setlength\tabcolsep{3.5pt}
-\begin{table}
-\begin{tabular}{|c|c|c|c |c|c|c|c|c|c | c|}
-\hline
-Dataset       & \thead{LAN \\(s)} & \thead{WAN \\(s)} & \thead{Comm. \\ (Mb)}  & \#And & \#Mul & \#Gates & depth & LOC\\
-\hline
-Chars4k    &  0.1         & 0.7         & 2            & 18k    & 3k    &  85k     & 1   & 89\\
-\hline
-USPS       &  0.2         & 0.9         & 4            & 62k    & 2k    &  285k    & 2   & 156\\
-\hline
-WARD       &  0.3         & 1.1         & 9            & 106k    & 8k    &  506k    & 3   & 283\\
-\hline
-\end{tabular}
-
- \caption{Bonsai benchmarks}
- \label{tab:bonsai} 
-\end{table}
 \setlength\tabcolsep{6pt}
@@
 fully connected layer. The total number of lines of \tool code for
-this benchamrk is 336 lines.
+this benchmark is 336 lines.
 
@@
 
+\begin{table}[t]
+\begin{tabular}{|c|c|c|c |c|c| c|}
+\hline
+  Stage         &  LAN (s) & WAN (s) & \thead{Comm. \\ (Mb)}  & depth & \#Gates & LOC\\
+\hline
+1    &  175       & 662        & 29816       & 16370    & 33m    & 500  \\
+\hline
+2    &  193        & 1095        & 31945        & 30916    & 37m & 516 \\
+\hline
+3    &  178        & 627        & 29810        & 16369    & 32m  & 478  \\
+\hline
+Total    &  546      & 2384        & 91571        & --    & 102m & 1494 \\
+\hline
+\end{tabular}
+
+ \caption{Partitioning results for matrix factorization. The time reported by~\cite{valeriaMatrix} for this computation is 10440 seconds.}
+ \label{tab:factor} 
+\end{table}
 Table~\ref{tab:cifar} shows the end-to-end numbers as
-well as the numbers for the sixth (the heaviest) stage. 
+well as the numbers for the sixth stage, which is the heaviest. 
 The number of gates are in millions, hence the suffix `m' in the last
@@
 
-\begin{table}
-\begin{tabular}{|c|c|c|c |c|c| c|}
-\hline
-  Stage         &  LAN (s) & WAN (s) & \thead{Comm. \\ (Mb)}  & depth & \#Gates & LOC\\
-\hline
-1    &  175       & 662        & 29816       & 16370    & 33m    & 500  \\
-\hline
-2    &  193        & 1095        & 31945        & 30916    & 37m & 516 \\
-\hline
-3    &  178        & 627        & 29810        & 16369    & 32m  & 478  \\
-\hline
-Total    &  546      & 2384        & 91571        & --    & 102m & 1494 \\
-\hline
-\end{tabular}
-
- \caption{Partitioning results for matrix factorization. The time reported by~\cite{valeriaMatrix} for this computation is 10440 seconds.}
- \label{tab:factor} 
-\end{table}
-
 We observe that in the LAN setting, we are about 19 times faster than~\cite{valeriaMatrix} and in the WAN
@@
 put in by Nikolaenko et al. (Section 5 of ~\cite{valeriaMatrix}) to scale a boolean circuit based
-crypto-backend to this benchmark.
+backend to this benchmark.
 

### 2017-10-30 OaklandSubmission/ex.tex
super minor
minor

@@
 the client's input vector. Figure~\ref{fig:ex-sml} shows the \tool code
-for this example. The code first reads the server's (client's) input using
- \ls{input1} (\ls{input2}). It then uses a
+for this example. The code first reads the server's (resp. client's) input using
+ \ls{input1} (resp. \ls{input2}). It then uses a
 \ls{for} loop to compute \ls{acc}, the dot product of \ls{w} and
@@
 components translate into API calls into our crypto back-end
-(ABY). E.g., for the code in Figure~\ref{fig:ex-sml}, the \tool
+(ABY). E.g., in Figure~\ref{fig:ex-sml}, the \tool
 compiler realizes that the array index \ls{i} in the dot product loop
-is public, and hence the array accesses need not be compiled
-obliviously. Therefore, it compiles the \ls{for}~loop into a C++
+is public, and hence the access locations need not be hidden.
+Therefore, it compiles the \ls{for}~loop into a C++
 \ls{for}~loop that will be executed in-clear
@@
 \ls{ycirc} to build the corresponding circuits
-(lines~\ref{line:condyaobegin} to~\ref{line:condyaoend})\footnote{As
-  we mention in
-  Section~\ref{sec:impl}, in our usage of ABY, boolean is synonymous
-  with Yao.}.
+(lines~\ref{line:condyaobegin} to~\ref{line:condyaoend})\footnote{In our usage of ABY, boolean is synonymous
+  with Yao (Section~\ref{sec:impl}).}.
 
-Importantly, conversions between the arithmetic and boolean
-representations require share-conversions. The \tool compiler also
+Using both arithmetic and boolean representations requires conversions between them. The \tool compiler also
 instruments these conversion gates accordingly. For example, in
@@
 while its metatheory provides strong correctness and security
-guarantees. We provide a formal description of the \tool language next.
-
+guarantees. 
 %% Consider a  cloud service provider Bob that wants to provide a service to diagnose

### 2017-10-30 OaklandSubmission/intro.tex
minor

@@
 
-%% Today it is hard for the developers to program secure applications
-%% using cryptographic techniques. Typical developers lack a deep
-%% understanding of the cryptographic protocols, and therefore cannot be
-%% expected to use them correctly and efficiently on their own.
-%% Ideally, a developer would declare the functionality in a general
-%% purpose, high-level programming language and a
-%% tool, e.g. a compiler, would generate an efficient protocol that
-%% implements the functionality securely, while hiding the cryptography
-%% behind-the-scenes. This paper presents such a framework for Secure
-%% Two-party Computation (\mpc).
+ Today it is hard for the developers to program secure applications
+ using cryptographic techniques. Typical developers lack a deep
+ understanding of the cryptographic protocols, and therefore cannot be
+ expected to use them correctly and efficiently on their own.
+ Ideally, a developer would declare the functionality in a general
+ purpose, high-level programming language and a
+ tool, e.g. a compiler, would generate an efficient protocol that
+ implements the functionality securely, while hiding the cryptography
+ behind-the-scenes.
 
+ This paper presents such a framework for Secure
+ Two-party Computation (\mpc),
+%
 %Most cryptographic tools rely on circuit as abstractions for any general computation but real-life developers write programs.
@@
 In this work, we address this gap for the case of secure multi-party computation.
-
+%
 %Most cryptographic tools are very hard to use non-crypto specialist developers.
@@
 To make cryptography accessible to non-crypto specialist developers, we need to design systems that do crypto automatically and satisfy the properties of ...
-
-Secure Two-party Computation (\mpc)~(\cite{yao,gmw}) is a powerful
+%
+%Secure Two-party Computation (\mpc)~(\cite{yao,gmw}) is
+a powerful
 cryptographic technique that allows two mutually distrusting parties
@@
 
-Since its introduction in 1980s, there has been a long
-sequence of work that has transformed 2PC from a mere theoretical tool to
-something practically efficient and usable.
-Our goal is to build a framework that enables programmers, who are not
-experts in cryptography, to easily program efficient and scalable \mpc
-protocols.  To understand the
+%Since its introduction in 1980s, there has been a long
+%sequence of work that has transformed 2PC from a mere theoretical tool to
+%something practically efficient and usable.
+%Our goal is to build a framework that enables programmers, who are not
+%experts in cryptography, to easily program efficient and scalable \mpc
+%protocols. 
+To understand the
 state-of-the-art, let us consider an example underlying many
@@
 This paper presents \tool\footnote{Read as ``easy peasy'',
-stands for Easy 2 Party Computation.}, a new \mpc framework.
-\tool is the first
+stands for Easy 2 Party Computation.}, the first
 ``cryptographic-cost aware'' framework that generates efficient and
 scalable \mpc protocols from high-level programs devoid of any
-cryptographic details. The generated protocols comprise combinations
-of arithmetic and boolean circuits and have performance comparable to
+cryptographic details. The generated protocols use combinations of
+arithmetic and boolean circuits and have performance comparable to
 or better than the custom, specialized
@@
 the security theorem, we formally reduce the security
-of our compiler to the security of the  \mpc back-end
- (by also invoking our correctness theorem). We also prove a formal security theorem against semi-honest (or, ``honest but curious") adversaries for our partitioning
+of the compiler output, against semi-honest (or, ``honest but curious") adversaries,
+ to the security of the  \mpc back-end. 
+ We also prove a formal security theorem against semi-honest  adversaries for our partitioning
 scheme (Section~\ref{sec:ld} and Section~\ref{sec:pipe}).

### 2017-10-30 OaklandSubmission/ld.tex
minor

@@
 
-\subsubsection*{Source to intermediate compilation judgments} We provide the compilation
+\subsubsection*{Source to intermediate compilation} We provide the compilation
 rules in \figureref{compile}. We present the rules in a
@@
 As mentioned earlier, the intermediate language models the code such as in \figureref{ex-aby} output by our compiler. 
-Next, a program in the intermediate language is evaluated to a circuit that can be executed in the distributed setting later.
+Next, a program in the intermediate language is evaluated to a circuit that can be executed in the distributed runtime later.
 The evaluation to a circuit computes away the public parts of the program and also \emph{flattens} the arrays so that the circuits are unaware of the array structure.
 Crucially, this phase of the semantics does not have access to the secret inputs.
-Below, we first provide the language for the circuits used followed by our evaluation rules.
+Below, we first provide the language for the circuits followed by the evaluation rules.
 

### 2017-10-30 OaklandSubmission/pipe.tex
minor

@@
 subtraction in the ring $(\mathbb{Z},+)$) and then outputs $s_i+r_i$
-only to Bob. Again, Alice's output from $\prog_i$ is $r_i$. The last
+only to Bob and $r_i$ only to Alice. The last
 program $\prog'_k$ takes inputs $x,y,r_{k-1},o_{k-1}$, runs $\prog_k$

### 2017-10-30 OaklandSubmission/related.tex
minor

@@
  Mitchell {\em et al.}~\cite{lambdaps} allow the user to select between Shamir's secret sharing~\cite{sss} and fully homomorphic encryption~\cite{gentry}.
-All these tools use either a arithmetic backend or a boolean backend and do not require share conversions.
+Unlike \tool, all these tools use either a arithmetic backend or a boolean backend but not a combination of both.
 
@@
 The work of Kerschbaum et al.~\cite{autoS} provides a scheme to automatically assign homomorphic encryption or garbled circuits to each operator in a computation that is expressed as a sequence of dyadic operations.
-They conjecture that the problem is NP hard and gave a linear programming based solution
-as well as a quadratic time greedy heuristic. Their performance is much poorer than ABY due to the high 
-costs associated with homomorphic encryption~\cite{aby}.
+They conjecture that the problem is NP hard and gave a linear programming based solution as well as a quadratic time greedy heuristic.
+These techniques are not directly applicable to \tool programs because of for-loops and if-conditions.
+However, we are exploring whether these ideas can be extended to yield a better type inference.
+%Their performance is much poorer than ABY due to the high 
+%costs associated with homomorphic encryption~\cite{aby}.
 %conversion costs between homomorphic encryption and boolean shares~\cite{aby}.
@@
 For example, the circuits can be optimized for depth~\cite{ddkssz15,cbmcgcdepth}, large garbled circuits can be pipelined~\cite{yao-pipe}, and oblivious RAM~\cite{oram1,oram2} can be used to hide access patterns of MIPS code~\cite{mips}.
+Incorporating these backends would only improve the performance and scalability of \tool implementations.
 

### 2017-10-30 OaklandSubmission/pipe.tex
pipe

@@
 \newcommand{\prog}{s}
+\newcommand{\progn}{t}
 \newcommand{\seq}{||}
-\newcommand{\state}{st}
+\newcommand{\state}{q}
 
@@
 the same ring $(\mathbb{Z},+)$ by Alice (in our implementation, all
-$r_i$ values are generated by a pseudorandom function). Let $\prog'_1$
+$r_i$ values are generated by a pseudorandom function). Let $\progn_1$
 be the program that takes as input $x,r_1$ from Alice and $y$ from Bob
@@
   protocol in the case where both parties must receive the same
-  output. To do so, we modify $\prog'_1$ to output $o_1 = s_1+r_1+t_1$ to
-  both parties (where $t_1$ is random and chosen by Bob). We can then
+  output. To do so, we modify $\progn_1$ to output $o_1 = s_1+r_1+r'_1$ to
+  both parties (where $r'_1$ is random and chosen by Bob). We can then
   appropriately modify the remaining steps as well.}. Alice's output
-from $\prog'_1$ is $r_1$. Next, every $\prog'_i$ ($2\leq i\leq k-1$)
+from $\progn_1$ is $r_1$. Next, every $\progn_i$ ($2\leq i\leq k-1$)
 takes as inputs $x,r_{i-1},r_i$ from Alice and $y,o_{i-1}$ from Bob,
@@
 only to Bob and $r_i$ only to Alice. The last
-program $\prog'_k$ takes inputs $x,y,r_{k-1},o_{k-1}$, runs $\prog_k$
+program $\progn_k$ takes inputs $x,y,r_{k-1},o_{k-1}$, runs $\prog_k$
 on inputs $x,y$ and state $o_{k-1}-r_{k-1}$ and outputs $z$ to both
@@
 enough to fit in the machine memory. We then compile and execute the
-$k$ programs $\prog'_1\seq\prog'_2\seq\dots\seq\prog'_k$ sequentially
-(freeing up memory usage after execution of each $\prog'_i$) to
+$k$ programs $\progn_1,\progn_2,\ldots,\progn_k$ sequentially
+(freeing up memory usage after execution of each $\progn_i$) to
 finally output $z$ to both parties.
+
+\begin{theorem}[Correctness and security of partitioning]
+If $\prog_1||\prog_2||\ldots||\prog_k$ is a decomposition of a program $\prog$, then there exists a sequence of programs $\progn_1, \progn_2, \ldots, \progn_k$ and protocols $\prot_1, \prot_2, \ldots, \prot_k$ such that $\prot_i$ securely realizes $\progn_i$ and  $\prot
+= \prot_1\seq\prot_2\dots\seq\prot_k$ securely realizes $s$.
+\end{theorem}
+
+\noindent {\em Proof.} Let $\progn_1, \ldots, \progn_k$ be the sequence of programs as defined above corresponding to the decomposition $\prog = \prog_1||\prog_2||\ldots||\prog_k$.
+\divya{Do we say that $\progn_i$ is a well-typed?} For every $1\leq i\leq k$, let $\prot_i$ be the \mpc protocol output by our framework for $\progn_i$.
+Then, by \theoremref{correctness},\theoremref{security}, $\prot_i$ securely realizes $\progn_i$.
+That is, for every $1\leq i \leq k-1$, the protocol $\prot_i$ provides observations $..$ to Alice and $...$ to Bob.
 

### 2017-10-30 OaklandSubmission/intro.tex
intro minor edits

@@
  using cryptographic techniques. Typical developers lack a deep
- understanding of the cryptographic protocols, and therefore cannot be
+ understanding of cryptographic protocols, and therefore cannot be
  expected to use them correctly and efficiently on their own.
@@
 about the client's input or output, and that the client learns nothing
-about the server classifier, beyond what is revelaed by the output
+about the classifier, beyond what is revelaed by the output
 label.
@@
 these representations. For instance, multiplication of two
-$\ell$-bit integers can be either expressed as a boolean circuit of
+$\ell$-bit integers can either be expressed as a boolean circuit of
 size $O(\ell^2)$, or as an arithmetic circuit with 1 multiplication
@@
 protocols from previous works~\cite{shafindss,wu,minionn,secureml,cryptonets,valeriaMatrix}.
-These papers (and others) cite the inefficiency of generic \mpc as the major motivation
+In fact, these papers (and others) cite the inefficiency of generic \mpc as the major motivation
 behind the design of specialized protocols. Using \tool, we empirically  demonstrate
@@
 we call secure code partitioning (or partitioning in short). At
-a high level, we break the original program into a sequence of small
+a high level, we decompose the original program into a sequence of small
 sub-programs, which are then sequentially processed by \tool, while

### 2017-10-30 OaklandSubmission/pipe.tex

@@
 program $\prog$ into a sequence of smaller \tool programs
-$\prog_1\seq\prog_2\seq\dots\seq\prog_k$ (as defined below), such
+$\progn_1,\progn_2,\ldots,\progn_k$ (as defined below) such
 that the circuit size
-requirement for each of the $\prog_i$ itself is manageable. We compile
-and execute each $\prog_i$ sequentially, feeding the outputs of
-$\prog_i$ as state
-information to $\prog_{i+1}$. We prove our partitioning scheme to be
-correct ($\prog$ and $\prog_1\seq\prog_2\seq\dots\seq\prog_k$ compute
-the same functionality) and secure (execution of
-$\prog_1\seq\prog_2\seq\dots\seq\prog_k$ does
-not reveal any more information than $s$). More details follow.
+requirement for each of the $\progn_i$ itself is manageable. We compile
+and execute each $\progn_i$ sequentially, feeding the outputs of
+$\progn_i$ as state
+information to $\progn_{i+1}$. We prove our partitioning scheme to be
+correct ($\prog$ and sequential execution of $\progn_1,\progn_2,\ldots,\progn_k$ compute
+the same functionality) and secure (sequential execution of
+$\progn_1,\progn_2,\ldots,\progn_k$ does
+not reveal any more information than $\prog$). More details follow.
 
@@
 $y$ from Bob and produces an output $z$ to both parties. Let
-$\prog_1\seq\prog_2\seq\dots\seq\prog_k$ be a sequence of programs
+$\prog_1\seq\prog_2\seq\dots\seq\prog_k$ be a decomposition of $\prog$
 such that the following holds. Define $\state_0 = \bot$ (the public empty
 state). For all
-$1\leq i\leq k-1$, $P_i$ takes inputs $x, y$ and $\state_{i-1}$ and
+$1\leq i\leq k-1$, $\prog_i$ takes inputs $x, y$ and $\state_{i-1}$ and
 outputs state $\state_i$. Finally, $\prog_k$ takes inputs $x, y$ and
@@
 To overcome this problem, we define a sequence of new programs
-$\prog'_i$ ($1\leq i\leq k$) as follows. Once again, define $\state_0
+$\progn_i$ ($1\leq i\leq k$) as follows. Once again, define $\state_0
 = \bot$. Without
@@
 runs $\prog_i$
-on inputs $x,y$ and state $o_{i-1}-r_{i-1}$ (where $-$ denotes
-subtraction in the ring $(\mathbb{Z},+)$) and then outputs $s_i+r_i$
-only to Bob and $r_i$ only to Alice. The last
+on inputs $x,y$ and state $\state_{i-1} = (o_{i-1}-r_{i-1})$ (where $-$ denotes
+subtraction in the ring $(\mathbb{Z},+)$) and then outputs $\state_i+r_i$
+to Bob and $r_i$ to Alice. The last
 program $\progn_k$ takes inputs $x,y,r_{k-1},o_{k-1}$, runs $\prog_k$
-on inputs $x,y$ and state $o_{k-1}-r_{k-1}$ and outputs $z$ to both
+on inputs $x,y$ and state $\state_{k-1} = (o_{k-1}-r_{k-1})$ and outputs $z$ to both
 parties.
@@
 $\prog_1\seq\prog_2\seq\dots\seq\prog_k$ with the guarantee that the
-circuits generated for every $\prog'_i$, as defined above, are small
+circuits generated for every $\progn_i$, as defined above, are small
 enough to fit in the machine memory. We then compile and execute the
@@
 \begin{theorem}[Correctness and security of partitioning]
-If $\prog_1||\prog_2||\ldots||\prog_k$ is a decomposition of a program $\prog$, then there exists a sequence of programs $\progn_1, \progn_2, \ldots, \progn_k$ and protocols $\prot_1, \prot_2, \ldots, \prot_k$ such that $\prot_i$ securely realizes $\progn_i$ and  $\prot
+If $\prog_1||\prog_2||\ldots||\prog_k$ is a decomposition of a program $\prog$, then there exists a sequence of programs $\progn_1, \progn_2, \ldots, \progn_k$ and protocols $\prot_1, \prot_2, \ldots, \prot_k$ such that for all $i$, $\prot_i$ securely realizes $\progn_i$ and  $\prot
 = \prot_1\seq\prot_2\dots\seq\prot_k$ securely realizes $s$.
@@
 \noindent {\em Proof.} Let $\progn_1, \ldots, \progn_k$ be the sequence of programs as defined above corresponding to the decomposition $\prog = \prog_1||\prog_2||\ldots||\prog_k$.
-\divya{Do we say that $\progn_i$ is a well-typed?} For every $1\leq i\leq k$, let $\prot_i$ be the \mpc protocol output by our framework for $\progn_i$.
-Then, by \theoremref{correctness},\theoremref{security}, $\prot_i$ securely realizes $\progn_i$.
-That is, for every $1\leq i \leq k-1$, the protocol $\prot_i$ provides observations $..$ to Alice and $...$ to Bob.
+For every $1\leq i\leq k$, let $\prot_i$ be the \mpc protocol output by our framework for $\progn_i$.
+Our construction for programs $\progn_i$ ensures that if $\prog$ is well-typed, then for each $1\leq i \leq k$, $\progn_i$ is well-typed.
+Then, by \theoremref{security}, $\prot_i$ securely realizes $\progn_i$.
+That is, for every $1\leq i \leq k-1$, the protocol $\prot_i$ provides observations $r_i$ to Alice and $o_i$ to Bob.
+And, protocol $\prot_k$ provides observation $z$ to both Alice and Bob.
+Finally, since $r_i$ and $o_i$ ($1\leq i\leq k-1$) are individually uniformly random (in $(\mathbb{Z},+)$, outputs received by the adversary can be simulated given the final output $z$.
 
-\begin{theorem}[Correctness and security of partitioning]
-Let $\prog$ be a program in our source language with output $z$, that
-is decomposed into programs
-$\prog'_1\seq\prog'_2\seq\dots\seq\prog'_k$ (as defined in
-the partioning procedure above) and compiled into the protocol $\prot
-= \prot_1\seq\prot_2\dots\seq\prot_k$ that outputs $r_1\seq
-r_2\seq\dots\seq z$
-(to Alice) and $o_1\seq o_2\seq\dots||z$ (to Bob), then $\prot$ securely
-realizes $P$.
-\end{theorem}
-
-\noindent {\em Proof.} From \theoremref{security}, it follows that
-$\prot_i$ securely realizes $\prog'_i$ for every $1\leq i\leq
-k$. Hence, we know that the only information learnt by Alice is
-$r_1||r_2||\cdots||r_{k-1}||z$ and by Bob is
-$o_1||o_2||\cdots||o_{k-1}||z$. Since $r_i$ and $o_i$ ($1\leq i\leq
-k-1$) are individually uniformly random (in $(\mathbb{Z},+)$, outputs
-received by the adversary can be easily simulated given the final
-output $z$.
+%\begin{theorem}[Correctness and security of partitioning]
+%Let $\prog$ be a program in our source language with output $z$, that
+%is decomposed into programs
+%$\prog'_1\seq\prog'_2\seq\dots\seq\prog'_k$ (as defined in
+%the partioning procedure above) and compiled into the protocol $\prot
+%= \prot_1\seq\prot_2\dots\seq\prot_k$ that outputs $r_1\seq
+%r_2\seq\dots\seq z$ 
+%(to Alice) and $o_1\seq o_2\seq\dots||z$ (to Bob), then $\prot$ securely
+%realizes $P$.
+%\end{theorem}
+%
+%\noindent {\em Proof.} From \theoremref{security}, it follows that
+%$\prot_i$ securely realizes $\prog'_i$ for every $1\leq i\leq
+%k$. Hence, we know that the only information learnt by Alice is
+%$r_1||r_2||\cdots||r_{k-1}||z$ and by Bob is
+%$o_1||o_2||\cdots||o_{k-1}||z$. Since $r_i$ and $o_i$ ($1\leq i\leq
+%k-1$) are individually uniformly random (in $(\mathbb{Z},+)$, outputs
+%received by the adversary can be easily simulated given the final
+%output $z$.
 
@@
 the sequence of small programs $\prog_1\seq\dots\seq\prog_k$ manually.
-Then \tool generates $\prog'_1\seq\dots\seq\prog'_k$ automatically.
+Then \tool generates sequence of programs $\progn_1,\dots\progn_k$ automatically.
 Automating this decomposition step requires an analysis that can 

### 2017-10-30 OaklandSubmission/ex.tex
minor

@@
  \ls{input1} (resp. \ls{input2}). It then uses a
-\ls{for} loop to compute \ls{acc}, the dot product of \ls{w} and
+\ls{for} loop to compute \ls{acc}, the inner product of \ls{w} and
 \ls{x}. Finally, the code outputs the result of comparing \ls{acc}
@@
 
-Using both arithmetic and boolean representations requires conversions between them. The \tool compiler also
+Using both arithmetic and boolean representations requires conversions between them.
+The \tool compiler also
 instruments these conversion gates accordingly. For example, in
@@
 We can now concretely see the advantages of \tool. Unarguably, it is
-easier for a developer to program, and get right, the code in
+easier for a developer to program and get right the code in
 Figure~\ref{fig:ex-sml}, rather than the code in
-Figure~\ref{fig:ex-aby} (as we mentioned in Section~\ref{sec:intro},
-other high-level languages are limited in their use of a single,
-arithmetic or boolean, crypto back-end). \tool also enables the
+Figure~\ref{fig:ex-aby}.
+% (as we mentioned in Section~\ref{sec:intro},
+%other high-level languages are limited in their use of a single,
+%arithmetic or boolean, crypto back-end). 
+\tool also enables the
 programmer to easily modify their code, while the compiler takes care

### 2017-10-30 OaklandSubmission/intro.tex
minor

--- 2017-10-30 OaklandSubmission/intro.tex
@@
 the computation expressed as either a boolean
-circuit~(\cite{yao,gmw}) or an arithmetic circuit~(\cite{sss,viff,gentry}). \TODO{check!}
+circuit~(\cite{yao,gmw}) or an arithmetic circuit~(\cite{sss,viff,gentry}). 
 %
@@
 the security theorem, we formally reduce the security
-of the compiler output, against semi-honest (or, ``honest but curious") adversaries,
+of our scheme, against semi-honest (or, ``honest but curious") adversaries,
  to the security of the  \mpc back-end. 
@@
 of the protocols
-generated by \tool are comparable to or better than (up to 19x in some
-cases) their state-of-the-art specialized implementations. Finally, we
+generated by \tool are comparable to or better than (up to 19x) their state-of-the-art specialized implementations. Finally, we
 show that partitioning enables us to implement applications

### 2017-10-30 OaklandSubmission/ex.tex
minor

@@
 components translate into regular C++ code, while the secret
-components translate into API calls into our crypto back-end
+components translate into API calls into our \mpc back-end
 (ABY). E.g., in Figure~\ref{fig:ex-sml}, the \tool
@@
 To summarize, \tool raises the level of abstraction for the
-programmer, and generates efficient crypto protocols automatically,
+programmer, and generates efficient \mpc protocols automatically,
 while its metatheory provides strong correctness and security

### 2017-10-30 OaklandSubmission/conclude.tex
conclusions draft.

@@
 \label{sec:conclude}
-\tool is the first
+We presented \tool, the first
 cryptographic-cost aware framework that generates efficient and
@@
 protocols from previous works.
+Several areas of improvement remain: a) While the backend that we use~\cite{aby} only provides security against a semi-honest adversary, our compiler is oblivious to the ``type of adversary'' for the underlying backend cryptographic protocol.
+In particular, if our cryptographic backend were maliciously secure, then Theorem \ref{theorem:security} would provide us a framework with malicious security.
+Furthermore, modifying our code partitioning to provide malicious security is fairly easy (one would instead provide {\em signed} shares of intermediate states in the decomposed programs to obtain malicious security).
+b) Our compiler converts programs into circuits before executing a \mpc and is efficient for the applications we present.
+However, one could consider executing programs that exploit secret-dependent memory accesses (such as binary search) to obtain better performance (as in \cite{oblivm}).
+It would be interesting to see if our techniques can be combined with that of \cite{oblivm}.
+\nc{Too much?
+Should we simply say only about the malicious security and stop?}

### 2017-10-30 OaklandSubmission/pipe.tex
nits in S4

@@
 
-Given $\prog$, partitioning first decomposes $\prog$ into
-$\prog_1\seq\prog_2\seq\dots\seq\prog_k$ with the guarantee that the
-circuits generated for every $\progn_i$, as defined above, are small
-enough to fit in the machine memory. We then compile and execute the
-$k$ programs $\progn_1,\progn_2,\ldots,\progn_k$ sequentially
-(freeing up memory usage after execution of each $\progn_i$) to
-finally output $z$ to both parties.
+Thus, given a decomposition of $\prog$ into
+$\prog_1\seq\prog_2\seq\dots\seq\prog_k$, we can use the construction
+above to generate programs $\progn_1,\progn_2,\ldots,\progn_k$, that
+can be sequentially executed, using the unmodified underlying \mpc
+backend. The following theorem state the correctness and security of
+the scheme.
 
@@
 
-\noindent {\em Proof.} Let $\progn_1, \ldots, \progn_k$ be the sequence of programs as defined above corresponding to the decomposition $\prog = \prog_1||\prog_2||\ldots||\prog_k$.
-For every $1\leq i\leq k$, let $\prot_i$ be the \mpc protocol output by our framework for $\progn_i$.
-Our construction for programs $\progn_i$ ensures that if $\prog$ is well-typed, then for each $1\leq i \leq k$, $\progn_i$ is well-typed.
-Then, by \theoremref{security}, $\prot_i$ securely realizes $\progn_i$.
-That is, for every $1\leq i \leq k-1$, the protocol $\prot_i$ provides observations $r_i$ to Alice and $o_i$ to Bob.
-And, protocol $\prot_k$ provides observation $z$ to both Alice and Bob.
-Finally, since $r_i$ and $o_i$ ($1\leq i\leq k-1$) are individually uniformly random (in $(\mathbb{Z},+)$, outputs received by the adversary can be simulated given the final output $z$.
+\noindent {\em Proof.} Let $\progn_1, \ldots, \progn_k$ be the
+sequence of programs as defined above corresponding to the
+decomposition $\prog = \prog_1||\prog_2||\ldots||\prog_k$.  For every
+$1\leq i\leq k$, let $\prot_i$ be the \mpc protocol output by our
+framework for $\progn_i$. Our construction for programs $\progn_i$
+ensures that if $\prog$ is well-typed, then for each $1\leq i \leq k$,
+$\progn_i$ is well-typed. Then, by \theoremref{security}, $\prot_i$
+securely realizes $\progn_i$. That is, for every $1\leq i \leq k-1$,
+the protocol $\prot_i$ provides observations $r_i$ to Alice and $o_i$
+to Bob. And, protocol $\prot_k$ provides observation $z$ to both Alice
+and Bob. Finally, since $r_i$ and $o_i$ ($1\leq i\leq k-1$) are
+individually uniformly random (in $(\mathbb{Z},+)$, outputs received
+by the adversary can be simulated given the final output $z$.
 
@@
 
-Our current implementation of secure code partitioning requires a manual step.
-In particular, the developer needs to decompose the large program $\prog$ into 
-the sequence of small programs $\prog_1\seq\dots\seq\prog_k$ manually.
-Then \tool generates sequence of programs $\progn_1,\dots\progn_k$ automatically.
-Automating this decomposition step requires an analysis that can 
+\subsubsection*{Managing circuit sizes using partitioning}
+We use partitioning for programs that require large
+circuits. Specifically, the programmer decomposes the program
+$\prog$ into a sequence of small programs
+$\prog_1\seq\dots\seq\prog_k$ manually. And then, \tool generates
+sequence of programs $\progn_1,\dots\progn_k$ automatically. We then
+compile and execute the $k$ programs
+$\progn_1,\progn_2,\ldots,\progn_k$ sequentially, freeing up memory
+usage after execution of each $\progn_i$.
+Automating the manual decomposition step requires an analysis that can 
 statically estimate the resource usage of a \tool program and we leave

### 2017-10-30 OaklandSubmission/ex.tex

@@
 (ABY). E.g., in Figure~\ref{fig:ex-sml}, the \tool
-compiler realizes that the array index \ls{i} in the dot product loop
+compiler realizes that the array index \ls{i} in the inner product loop
 is public, and hence the access locations need not be hidden.
@@
 circuit representations for different sub-componenets. For example,
-the compiler realizes that the dot product computation is more
+the compiler realizes that the inner product computation is more
 efficient in the arithmetic representation, and therefore it builds

### 2017-10-30 OaklandSubmission/main.tex
small change
wip in S2
minor

--- 2017-10-27 OaklandSubmission/main.tex
@@
 \maketitle
-% \begin{abstract}
-% The abstract goes here.
-% \end{abstract}
+\begin{abstract}
+Secure two-party computation (\mpc) allows two mutually distrusting parties to compute any function on their joint inputs, without revealing anything other than the output of the function to one another.
+%While the application and potential impact of secure computation is virtually limitless, writing an efficient and scalable secure protocol is by no means an easy feat.
+% Prior work in designing such protocols, suffered from either a) a lack of programmability (e.g.
+requiring the programmer to write cryptography-aware circuits for the computation); b) a lack of efficiency for several natural computations (such as machine learning classification tasks); or c) required cryptographers to design hand-crafted protocols for various applications.
+ In this work, we present \tool: a 
+%new ``cryptographic-cost aware'' 
+\mpc framework that generates efficient and scalable \mpc protocols from high-level programs that 
+%contain absolutely no cryptographic details. Programs in our language 
+are easy-to-write and are statically verified. 
+%Furthermore, 
+The \mpc protocols generated by our framework match or outperform recent works that provide hand-crafted \mpc protocols for various functionalities such as classification and matrix factorization (up to a 19x factor!).
+\end{abstract}
 \IEEEpeerreviewmaketitle
@@
 %\input{fig-ex-sml}
-\input{fig-ex-aby}
+%\input{fig-ex-aby}
 
@@
 
-\newpage
+%\newpage
 \bibliographystyle{abbrv}
 \bibliography{bib}
-\balance
+%\balance
 

### 2017-10-30 OaklandSubmission/main.tex
abstract

@@
 \begin{abstract}
-Secure two-party computation (\mpc) allows two mutually distrusting parties to compute any function on their joint inputs, without revealing anything other than the output of the function to one another.
+  %% Secure two-party computation (\mpc) allows two mutually distrusting
+  %% parties to compute any function on their joint inputs, without
+  %% revealing anything other than the output of the function to one
+  %% another.
+
   %While the application and potential impact of secure computation is virtually limitless, writing an efficient and scalable secure protocol is by no means an easy feat.
@@
 requiring the programmer to write cryptography-aware circuits for the computation); b) a lack of efficiency for several natural computations (such as machine learning classification tasks); or c) required cryptographers to design hand-crafted protocols for various applications.
- In this work, we present \tool: a 
+We present \tool: a 
 %new ``cryptographic-cost aware'' 
-\mpc framework that generates efficient and scalable \mpc protocols from high-level programs that 
-%contain absolutely no cryptographic details. Programs in our language 
-are easy-to-write and are statically verified. 
+secure two-party computation (\mpc) framework that generates efficient
+and scalable
+\mpc protocols from high-level, easy-to-write programs.
+%% that 
+%% %contain absolutely no cryptographic details. Programs in our language 
+%% are easy-to-write and are statically verified. 
 %Furthermore,
-The \mpc protocols generated by our framework match or outperform recent works that provide hand-crafted \mpc protocols for various functionalities such as classification and matrix factorization (up to a 19x factor!).
+\tool provides formal correctness and security guarantees, and we
+empirically demonstrate that the
+protocols generated by our framework match or outperform (up to
+19x in some cases)
+recent works that provide hand-crafted protocols for various
+functionalities such as classification and matrix factorization
+%% (up to
+%% a 19x factor!).
 \end{abstract}

### 2017-10-30 OaklandSubmission/main.tex
nit

@@
 recent works that provide hand-crafted protocols for various
-functionalities such as classification and matrix factorization
+functionalities such as secure prediction and matrix factorization.
 %% (up to

### 2017-10-30 OaklandSubmission/main.tex
abstract

--- 2017-10-30 OaklandSubmission/main.tex
@@
 \mpc protocols from high-level, easy-to-write programs.
+\tool provides formal correctness and security guarantees while maintaining performance and scalability.
 %% that 
@@
 %Furthermore,
-\tool provides formal correctness and security guarantees, and we
+All previous works generate protocols that use either arithmetic or boolean circuits exclusively.
+Our compiler is the first to generate protocols that combine both arithmetic and boolean  for better performance.
+%We give the first compiler that generates protocols which combine both arithmetic and boolean circuits
+%for better performance compared to protocols that use either arithmetic or boolean circuits exclusively.
+We
 empirically demonstrate that the
 protocols generated by our framework match or outperform (up to
-19x in some cases)
+19x)
 recent works that provide hand-crafted protocols for various
-functionalities such as classification and matrix factorization
+functionalities such as secure neural net prediction (Liu et al. CCS'17) and matrix factorization (Nikolaenko et al. CCS'13).
+%\tool provides formal correctness and security guarantees.
 %% (up to

### 2017-10-31 OaklandSubmission/main.tex
minor

@@
 secure two-party computation (\mpc) framework that generates efficient
-and scalable
 \mpc protocols from high-level, easy-to-write programs.
@@
 %Furthermore,
-All previous works generate protocols that use either arithmetic or boolean circuits exclusively.
-Our compiler is the first to generate protocols that combine both arithmetic and boolean  for better performance.
+Previous language frameworks such as Wysteria, ObliVM, CBMC-GC, etc.
+generate protocols that use either arithmetic or boolean circuits exclusively.
+Our compiler is the first to generate protocols that combine both arithmetic sharing and garbled circuits  for better performance.
 %We give the first compiler that generates protocols which combine both arithmetic and boolean circuits
@@
 recent works that provide hand-crafted protocols for various
-functionalities such as secure neural net prediction (Liu et al. CCS'17) and matrix factorization (Nikolaenko et al. CCS'13).
+functionalities such as secure  prediction  and matrix factorization.
 %\tool provides formal correctness and security guarantees.

### 2017-10-31 OaklandSubmission/conclude.tex
trim

--- 2017-10-30 OaklandSubmission/conclude.tex
@@
 protocols from previous works.
+The \mpc backed we use provides security against semi-honest or passive adversaries.
+%our compiler is oblivious to the ``type of adversary'' for the underlying \mpc backend.
+Given a \mpc backend secure against malicious or active adversaries, Theorem \ref{theorem:security} 
+can be extended to obtain a framework with malicious security. Furthermore,  
+partitioning can be modified to provide malicious security by {\it signing} the shares of the intermediate states.
+However, we are not aware of a maliciously secure \mpc backend that works with combinations of
+arithmetic and boolean circuits.
+
+% Several areas of improvement remain: a) While the backend that we use~\cite{aby} only provides security against a semi-honest adversary, our compiler is oblivious to the ``type of adversary'' for the underlying backend cryptographic protocol.
+In particular, if our cryptographic backend were maliciously secure, then Theorem \ref{theorem:security} would provide us a framework with malicious security.
+Furthermore, modifying our code partitioning to provide malicious security is fairly easy (one would instead provide {\em signed} shares of intermediate states in the decomposed programs to obtain malicious security).
+b) Our compiler converts programs into circuits before executing a \mpc and is efficient for the applications we present.
+However, one could consider executing programs that exploit secret-dependent memory accesses (such as binary search) to obtain better performance (as in \cite{oblivm}).
+It would be interesting to see if our techniques can be combined with that of \cite{oblivm}.
+\nc{Too much?
+Should we simply say only about the malicious security and stop?}

### 2017-10-31 OaklandSubmission/ld.tex
reaching until the security theorem

@@
 In this section we prove correctness and security of our \tool compiler.
-We formalize our source, intermediate and circuit languages. 
-An example of a program in our source language is provided in \figureref{ex-sml}.
+%We formalize our source, intermediate and circuit languages. 
+%An example of a program in our source language is provided in \figureref{ex-sml}.
 %Our source runtime semantics is a model of the ideal, trusted third-party semantics, and
@@
 %parties.
-Our source runtime semantics describe the ``trusted third party" execution semantics of the source programs and
-generate observations corresponding to the values revealed to the
+We first formalize our source language (an example program being
+\figureref{ex-sml}), and its runtime semantics.
+This semantics describes the ``trusted third party"
+execution semantics of the source programs and
+generates observations corresponding to the values revealed to the
 parties.
 %
-We present the compilation rules that type check a program in the source language and generate a program in the intermediate language.
-This intermediate language models programs such as \figureref{ex-aby} generated by our compiler implementation.
+We then present the compilation rules that type check a program in the
+source language and generate a program in the intermediate
+language (an example program being \figureref{ex-aby}).
 %
-We evaluate the program in our intermediate language  to a circuit by ``evaluating away'' the public parts and the arrays.
-Crucially, this step does not have access to the secret inputs; those are processed by the distributed circuit semantics that model the \mpc back-end.
+Next, we present the runtime semantics of our intermediate language
+that evaluates to a circuit by ``evaluating away'' the public parts
+and the arrays. Crucially, this step does not have access to the secret
+inputs; those are processed by our distributed circuit semantics that
+model the \mpc back-end.
 %
-Evaluation in the distributed setting involves the parties running an interactive protocol.
-This step, like  the source semantics, emits observations corresponding to the values revealed to the parties.
+Evaluation in this distributed setting involves the parties running an
+interactive protocol. This step, like  the source semantics, emits
+observations corresponding to the values revealed to the parties.
 %Finally, we formalize the distributed circuit semantics (of the cryptographic back-end) that computes the generated circuit in the distributed setting, and like  the source semantics, emits observationscorresponding to the values revealed to the parties.
 
-To prove the correctness of \tool, we prove that  the observations in source semantics and distributed circuit semantics are identical (see \theoremref{correctness}). 
-We combine this correctness guarantee and the security of the \mpc back-end to argue security of the implementations generated by \tool (see \theoremref{security}).
+To prove the correctness of \tool, we prove that  the observations in
+source semantics and the distributed circuit semantics are identical
+(\theoremref{correctness}).
+We combine this correctness theorem and the security of the \mpc
+back-end to prove security of the protocols generated by \tool
+(\theoremref{security}).
 %For the security of our compiler, we rely on the correctness guarantee and  reduce the security to the security of the underlying cryptographic protocol used to compute the generated circuit (see \theoremref{security}).
-We present only selected parts of our formalization for space
-reasons. Full definitions and proofs can be found in~\cite{anonymized}.
+We present only selected parts of our formalization. Full definitions
+and proofs can be found in~\cite{anonymized}.
 
@@
                  \rho \vdash e \Downarrow c\\\\
-                 c = \top \Rightarrow e' = e_{1}\\\\
-                 c = \bot \Rightarrow e' = e_{2}\\\\
-                 \rho \vdash e' \Downarrow c'
-               }
-               {
-                 \rho \vdash \cond{e}{e_{1}}{e_{2}} \Downarrow c'
+                 c = \top \Rightarrow e_{3} = e_{1}\\\\
+                 c = \bot \Rightarrow e_{3} = e_{2}\\\\
+                 \rho \vdash e_{3} \Downarrow c_{3}
+               }
+               {
+                 \rho \vdash \cond{e}{e_{1}}{e_{2}} \Downarrow c_{3}
                }
@@
   \\\\
-    \fbox{$\rho \vdash s \Downarrow \rho'; O$}
+    \fbox{$\rho \vdash s \Downarrow \rho_{1}; O$}
   \[
@@
 \subsubsection*{Source language} Our source language
- is a simple imperative language and we show a subset of the supported operators in \figureref{srclang}. Types
+ is a simple imperative language shown in \figureref{srclang}. Types
 $\psi$ consist of the base types $\sigma$, and arrays
@@
  integer constants $n$, \kw{bool} constants $\top$ and $\bot$,
-variables $x$, binary operations $e_{1} \mulop e_{2}$ and $e_{1} > e_{2}$,
-conditionals $\cond{e}{e_{1}}{e_{2}}$, array literals 
+variables $x$, binary operations $e_{1} \mulop e_{2}$ and $e_{1} >
+e_{2}$ (we support several other
+operators in the implementation, detailed in Section~\ref{sec:impl}),
+conditionals $\cond{e}{e_{1}}{e_{2}}$,
+array literals
 $[\overline{e_{i}}]_{n}$\footnote{We write $\overline{e}$ (and
@@
 
-\vspace{0.2cm}
+\vspace{0.1cm}
 $
@@
 
-\vspace{0.2cm}
+\vspace{0.1cm}
 Values consist of constants and array of constants, runtime environment
@@
 the environment. Rule ({\sc{E-Mult}}) inductively evaluates $e_{1}$ and
-$e_{2}$, and evaluates to their product. Rule ({\sc{E-Read}})
+$e_{2}$, and returns their product. Rule ({\sc{E-Read}})
 evaluates an array read operation. It first evaluates $x$ to an array
@@
 language as $[\kw{in}_{j}]_{n}$, which can then evaluate using the
-rule ({\sc{E-Arr}}). The remaining rules are straightforward, and are
-elided for space reasons.
+rule ({\sc{E-Arr}}) \footnote{The notation $\forall i \in [n]$ is read
+  as $\forall i \in \{0\dots n - 1\}.$}. The remaining rules are
+straightforward, and are elided for space reasons.
 
 The judgment $\rho \vdash s \Downarrow \rho_{1}; O$ represents the
-big-step evaluation of a statement $s$ under environment $\rho$
+big-step evaluation of a statement $s$ under environment $\rho$,
 producing a new environment $\rho_{1}$ and observations $O$. Rule
@@
 \kw{while} statement under this updated environment. Rule
-({\sc{E-LoopT}}) is the termination case for \kw{while} statements,
+({\sc{E-LoopT}}) is the termination case for \kw{while},
 when $\rho(x) > n$. Finally, the rule ({\sc{E-Out}}) evaluates the
@@
     \ftext{Type} & \tau &::=& \sigma^{\ell} \mid \sigma^{\ell}[n]\\
-    \ftext{Expression} & \widetilde{e} &::=& c \mid x \mid \widetilde{e_{1}} \mulop_{\ell} \widetilde{e_{2}} \mid \widetilde{e_{1}} >_{\ell} \widetilde{e_{2}} \\
-    & & \mid & \lcond{\ell}{{\widetilde{e}}}{{\widetilde{e_{1}}}}{{\widetilde{e_{2}}}} \mid x[\widetilde{e}] \mid [\overline{\widetilde{e_{i}}}]_{n} \mid \kw{in}^{m}_{j} \mid \widetilde{e} \rhd m\\
-    \ftext{Statement} & \widetilde{s} &::=& \tau\:x = \widetilde{e} \mid x := \widetilde{e} \mid \dots \mid \widetilde{s_{1}}; \widetilde{s_{2}} \mid \dots\\
+    \ftext{Expression} & \widetilde{e} &::=& c \mid x \mid \widetilde{e}_{1} \mulop_{\ell} \widetilde{e}_{2} \mid \widetilde{e}_{1} >_{\ell} \widetilde{e}_{2} \mid x[\widetilde{e}] \mid [\,\overline{\widetilde{e}}_{i}]_{n} \\
+    & & \mid & \lcond{\ell}{{\widetilde{e}}}{{\widetilde{e}}_{1}}{{\widetilde{e}}_{2}} \mid \kw{in}^{m}_{j} \mid \langle \ell \rhd m \rangle\:\widetilde{e}\\
+    \ftext{Statement} & \widetilde{s} &::=& \tau\:x = \widetilde{e} \mid x := \widetilde{e} \mid \dots \mid \widetilde{s}_{1}; \widetilde{s}_{2} \mid \dots\\
   \end{array}
@@
 $\mathcal{B}$-labeled operators generate boolean circuits. The
-form $\widetilde{e} \rhd m$ denotes coercing $\widetilde{e}$ to the label $m$, which generates conversions to $m$. 
+form $\langle \ell \rhd m \rangle\:\widetilde{e}$ denotes coercing
+$\widetilde{e}$ from label $\ell$ to label $m$.
 %The statements $\widetilde{s}$ are analogous to $s$.
@@
   \begin{array}{c}
-     \inferrule*[lab={\footnotesize{T-Cons}}]
-               {
-                 \tau = \mathsf{typeof}(c)^{\mathcal{P}}
-               }
-               {
-                 \Gamma \vdash c : \tau \leadsto c 
-               }
-
+     \inferrule*[lab={\footnotesize{T-UInt}}]
+               {
+               }
+               {
+                 \Gamma \vdash n : \kw{uint}^{\mathcal{P}} \leadsto n
+               }
+               \hspace{0.1cm}
+     \inferrule*[lab={\footnotesize{T-Bool}}]
+               {
+                 c = \top \vee c = \bot
+               }
+               {
+                 \Gamma \vdash c : \kw{bool}^{\mathcal{P}} \leadsto c 
+               }
+
+               \hspace{0.1cm}
                 \inferrule*[lab={\footnotesize{T-Inp}}]
@@
                {
-                 \forall i \in \{1,2\}.\:\Gamma \vdash e_{i} : \kw{uint}^{\ell} \leadsto \widetilde{e_{i}}\\\\
+                 \forall i \in \{1,2\}.\:\Gamma \vdash e_{i} : \kw{uint}^{\ell} \leadsto \widetilde{e}_{i}\\\\
                  \left(\ell = \mathcal{P}\right) \vee \left(\ell = \mathcal{A}\right)
@@
                {
-                 \Gamma \vdash e_{1} \mulop e_{2} : \kw{uint}^{\ell} \leadsto \widetilde{e_{1}} \mulop_{\ell} \widetilde{e_{2}}
+                 \Gamma \vdash e_{1} \mulop e_{2} : \kw{uint}^{\ell} \leadsto \widetilde{e}_{1} \mulop_{\ell} \widetilde{e}_{2}
                }
@@
                {
-                 \forall i \in \{1,2\}.\:\Gamma \vdash e_{i} : \kw{uint}^{\ell} \leadsto \widetilde{e_{i}}\\\\
+                 \forall i \in \{1,2\}.\:\Gamma \vdash e_{i} : \kw{uint}^{\ell} \leadsto \widetilde{e}_{i}\\\\
                  \left(\ell = \mathcal{P}\right) \vee \left(\ell = \mathcal{B}\right)
@@
                {
-                 \Gamma \vdash e_{1} > e_{2} : \kw{bool}^{\ell} \leadsto \widetilde{e_{1}} >_{\ell} \widetilde{e_{2}}
+                 \Gamma \vdash e_{1} > e_{2} : \kw{bool}^{\ell} \leadsto \widetilde{e}_{1} >_{\ell} \widetilde{e}_{2}
                }
@@
                  \Gamma \vdash e : \kw{bool}^{\ell} \leadsto \widetilde{e}\\\\
-                 \forall i \in \{1,2\}.\:\Gamma \vdash e_{i} : \sigma^{\ell'} \leadsto \widetilde{e_{i}}\\\\
-                 \ell = \mathcal{P} \vee (\ell = \mathcal{B} \wedge \ell' =\mathcal{B})
-               }
-               {
-                 \Gamma \vdash \cond{e}{e_{1}}{e_{2}} : \sigma^{\ell'} \leadsto \lcond{\ell}{\widetilde{e}}{\widetilde{e_{1}}}{\widetilde{e_{2}}}
+                 \forall i \in \{1,2\}.\:\Gamma \vdash e_{i} : \sigma^{\ell_{1}} \leadsto \widetilde{e}_{i}\\\\
+                 \ell = \mathcal{P} \vee (\ell = \mathcal{B} \wedge \ell_{1} =\mathcal{B})
+               }
+               {
+                 \Gamma \vdash \cond{e}{e_{1}}{e_{2}} : \sigma^{\ell_{1}} \leadsto \lcond{\ell}{\widetilde{e}}{\widetilde{e}_{1}}{\widetilde{e}_{2}}
                }
@@
                {
-                 \forall i \in [n].\:\Gamma \vdash e_{i} : \sigma^{\ell} \leadsto \widetilde{e_{i}}
-               }
-               {
-                 \Gamma \vdash [\overline{e_{i}}]_{n} : \sigma^{\ell}[n] \leadsto [\overline{\widetilde{e_{i}}}]_{n}
+                 \forall i \in [n].\:\Gamma \vdash e_{i} : \sigma^{\ell} \leadsto \widetilde{e}_{i}
+               }
+               {
+                 \Gamma \vdash [\overline{e_{i}}]_{n} : \sigma^{\ell}[n] \leadsto [\,\overline{\widetilde{e}}_{i}]_{n}
                }
@@
                {
-                 \Gamma \vdash e : \sigma^{m} \leadsto \widetilde{e} \rhd m
+                 \Gamma \vdash e : \sigma^{m} \leadsto \langle \ell \rhd m \rangle\:\widetilde{e}
                }
@@
                {
-                 \tau = \psi^\ell \\\\
-                 \Gamma \vdash e : \tau \leadsto \widetilde{e}
-               }
-               {
-                 \Gamma \vdash \psi\:x = e \leadsto \tau\:x = \widetilde{e} \mid \Gamma, x:\tau
+                 \Gamma \vdash e : \psi^{\ell} \leadsto \widetilde{e}
+               }
+               {
+                 \Gamma \vdash \psi\:x = e \leadsto \psi^{\ell}\:x = \widetilde{e} \mid \Gamma, x:\tau
                }
@@
                {
-                 \Gamma \vdash x := e \leadsto x = \widetilde{e} \mid \Gamma
+                 \Gamma \vdash x := e \leadsto x := \widetilde{e} \mid \Gamma
                }
@@
                  \Gamma \vdash e : \kw{bool}^{\mathcal{P}} \leadsto \widetilde{e}\\\\
-                 \forall i \in \{1, 2\}.\:\Gamma \vdash s_{i} \leadsto \widetilde{s_{i}} \mid \_
-               }
-               {
-                 \Gamma \vdash \ite{e}{s_{1}}{s_{2}} \leadsto \ite{\widetilde{e}}{\widetilde{s_{1}}}{\widetilde{s_{2}}}  \mid \Gamma
+                 \forall i \in \{1, 2\}.\:\Gamma \vdash s_{i} \leadsto \widetilde{s}_{i} \mid \_
+               }
+               {
+                 \Gamma \vdash \ite{e}{s_{1}}{s_{2}} \leadsto \ite{\widetilde{e}}{\widetilde{s}_{1}}{\widetilde{s}_{2}}  \mid \Gamma
                }
@@
                {
-                 \Gamma \vdash s_{1} \leadsto \widetilde{s_{1}} \mid \Gamma_{1}\\\\
-                 \Gamma_{1} \vdash s_{2} \leadsto \widetilde{s_{2}} \mid \Gamma_{2}
-               }
-               {
-                 \Gamma \vdash s_{1}; s_{2} \leadsto \widetilde{s_{1}}; \widetilde{s_{2}} \mid \Gamma_{2}
+                 \Gamma \vdash s_{1} \leadsto \widetilde{s}_{1} \mid \Gamma_{1}\\\\
+                 \Gamma_{1} \vdash s_{2} \leadsto \widetilde{s}_{2} \mid \Gamma_{2}
+               }
+               {
+                 \Gamma \vdash s_{1}; s_{2} \leadsto \widetilde{s}_{1}; \widetilde{s}_{2} \mid \Gamma_{2}
                }
@@
 non-deterministically. \sectionref{impl} describes the label  inference
-scheme of our implementation.
+scheme in our implementation.
 
@@
 $\Gamma$ maps variables $x$ to types $\tau$, says that
-under $\Gamma$, $e$ (in source language) compiles to $\widetilde{e}$ (in the intermediate language) and $\widetilde{e}$  has type $\tau$.
-The rule ({\sc{T-Cons}}) assigns the label $\mathcal{P}$ to the
-constants, as the constants are always public. Rule ({\sc{T-Mult}})
+under $\Gamma$, $e$ (in source language) compiles to $\widetilde{e}$
+(in the intermediate language), where $\widetilde{e}$ has type $\tau$.
+Rules ({\sc{T-UInt}}) and ({\sc{T-Bool}}) assigns the label
+$\mathcal{P}$ to the constants, as the constants are always
+public. Rule ({\sc{T-Mult}})
 compiles a multiplication to either a public multiplication  ($\mulop_{\mathcal{P}}$), or a
@@
 rule ({\sc{T-Gt}}) compiles $e_{1} > e_{2}$ to either public comparison,
-or secret boolean comparison $>_\mathcal{B}$ (and never $>_{\mathcal{A}}$). The rule for conditional
+or secret boolean comparison $>_\mathcal{B}$ (and never
+$>_{\mathcal{A}}$).
+%
+The rule for conditional
 ({\sc{T-Cond}}) has two cases: when the conditional expression $e$ is of
 type $\kw{bool}^{\mathcal{P}}$, both the branches have a base type
-$\sigma^{\ell'}$, for an arbitrary $\ell'$, and the conditional is
+$\sigma^{\ell_{1}}$, for an arbitrary $\ell_{1}$, and the conditional is
 compiled to a public conditional, whereas when the conditional
-expression has type $\kw{bool}^{\mathcal{B}}$, $\ell'$ is also
+expression has type $\kw{bool}^{\mathcal{B}}$, $\ell_{1}$ is also
 $\mathcal{B}$, and the conditional is compiled to a secret conditional
@@
 judgment $ \models e < n$ to prove that the array index is in
-bounds. Section~\ref{sec:impl} discusses our implementation of this check. Rule ({\sc{T-Inp}}) picks a label $m$ for the input. Finally,
+bounds. Section~\ref{sec:impl} discusses our implementation of this
+check. Rule ({\sc{T-Inp}}) picks a label $m$ for the input. Finally,
 the rule ({\sc{T-Sub}}) is the subsumption rule that coerces an expression of
 type $\sigma^{\ell}$ to an expression of type $\sigma^{m}$ using the
-coerce expression. It is important for security that the secrets cannot be coerced to public values and indeed ({\sc{T-Sub}}) does not permit it.
+coerce expression. It is important for security that the secrets
+cannot be coerced to public values and indeed ({\sc{T-Sub}}) does not
+permit it.
 
@@
 and type environment $\Gamma_{1}$. Rule ({\sc{T-Decl}}) picks a label
-$\ell$, and adds the binding for $x$ to the environment. If $\psi = \sigma$, $\psi^\ell = \sigma^\ell$, else if $\psi = \sigma[n]$, $\psi^\ell = \sigma^\ell[n]$. 
+$\ell$, and adds the binding for $x$ to the environment (if $\psi =
+\sigma$, $\psi^\ell = \sigma^\ell$, else if $\psi = \sigma[n]$,
+$\psi^\ell = \sigma^\ell[n]$).
 Rule ({\sc{T-Assign}}) looks up the type of $x$ in $\Gamma$ and compiles $e$ to $\tilde{e}$ of same type.
@@
 rule for the (internal) \kw{while} form ensures that $x$ is mapped in
-$\Gamma$ at type $\kw{uint}^{\mathcal{P}}$, and that $x \notin
-\mathsf{modifies}(s)$--this is necessary for ensuring the termination
-of our semantics.
-
-As mentioned earlier, the intermediate language models the code such as in \figureref{ex-aby} output by our compiler. 
-Next, a program in the intermediate language is evaluated to a circuit that can be executed in the distributed runtime later.
-The evaluation to a circuit computes away the public parts of the program and also \emph{flattens} the arrays so that the circuits are unaware of the array structure.
-Crucially, this phase of the semantics does not have access to the secret inputs.
-Below, we first provide the language for the circuits followed by the evaluation rules.
+$\Gamma$ at type $\kw{uint}^{\mathcal{P}}$, and that the statement $s$
+does not modify $x$ ($x \notin
+\mathsf{modifies}(s)$)--this is necessary for ensuring termination.
+
+As mentioned earlier, the intermediate language models the code such
+as in \figureref{ex-aby} output by our compiler.
+Next, a program in the intermediate language is evaluated to a circuit
+that can be executed in the distributed runtime later. The evaluation
+to a circuit computes away the public parts of the program and also
+\emph{flattens} the arrays so that the circuits are unaware of the
+array structure.
+Crucially, this phase of the semantics does not have access to the
+secret inputs.
+Below, we first provide the language for the circuits followed by the
+evaluation rules.
 
@@
     \ftext{Circuit gate} & \gate & ::= & \wire \mid \kw{in}^{m}_{j} \mid \kw{mult}\:\gate_1\:\gate_{2} \mid \kw{gt}\:\gate_{1}\:\gate_{2}\\
-    & & \mid & \kw{mux}\:\gate\:\gate_{1}\:\gate_{2} \mid \gate \rhd m \mid c\\
+    & & \mid & \kw{mux}\:\gate\:\gate_{1}\:\gate_{2} \mid \langle \ell
+    \rhd m \rangle\:\gate \mid c\\
     \ftext{Sub-circuit} & \val & ::= & \gate \mid [\overline{\gate_{i}}]_{n}\\
@@
                {
-                 \forall i \in \{1, 2\}.\:\widetilde{\rho} \vdash \widetilde{e_{i}} \Downarrow n_{i}
-               }
-               {
-                 \widetilde{\rho} \vdash \widetilde{e_{1}} \mulop_{\mathcal{P}} \widetilde{e_{2}} \Downarrow n_{1} \mulop n_{2}
+                 \forall i \in \{1, 2\}.\:\widetilde{\rho} \vdash \widetilde{e}_{i} \Downarrow n_{i}
+               }
+               {
+                 \widetilde{\rho} \vdash \widetilde{e}_{1} \mulop_{\mathcal{P}} \widetilde{e}_{2} \Downarrow n_{1} \mulop n_{2}
                }
@@
                {
-                 \forall i \in \{1, 2\}.\:\widetilde{\rho} \vdash \widetilde{e_{i}} \Downarrow \gate_{i}
-               }
-               {
-                 \widetilde{\rho} \vdash \widetilde{e_{1}} \mulop_{\mathcal{A}} \widetilde{e_{2}} \Downarrow \kw{mult}\:\gate_{1}\:\gate_{2}
+                 \forall i \in \{1, 2\}.\:\widetilde{\rho} \vdash \widetilde{e}_{i} \Downarrow \gate_{i}
+               }
+               {
+                 \widetilde{\rho} \vdash \widetilde{e}_{1} \mulop_{\mathcal{A}} \widetilde{e}_{2} \Downarrow \kw{mult}\:\gate_{1}\:\gate_{2}
                }
@@
                {
-                 \forall i \in \{1, 2, 3\}.\:\widetilde{\rho} \vdash \widetilde{e_{i}} \Downarrow \gate_{i}
-               }
-               {
-                 \widetilde{\rho} \vdash \lcond{\mathcal{B}}{\widetilde{e_{1}}}{\widetilde{e_{2}}}{\widetilde{e_{3}}} \Downarrow \kw{mux}\:\gate_{1}\:\gate_{2}\:\gate_{3}
+                 \forall i \in \{1, 2, 3\}.\:\widetilde{\rho} \vdash \widetilde{e}_{i} \Downarrow \gate_{i}
+               }
+               {
+                 \widetilde{\rho} \vdash \lcond{\mathcal{B}}{\widetilde{e}_{1}}{\widetilde{e}_{2}}{\widetilde{e}_{3}} \Downarrow \kw{mux}\:\gate_{1}\:\gate_{2}\:\gate_{3}
                }
@@
                {
-                 \widetilde{\rho} \vdash \widetilde{e} \Downarrow \val
-               }
-               {
-                 \widetilde{\rho} \vdash \widetilde{e} \rhd m \Downarrow \val \rhd m
+                 \widetilde{\rho} \vdash \widetilde{e} \Downarrow \gate
+               }
+               {
+                 \widetilde{\rho} \vdash \langle \ell \rhd m \rangle\:
+                 \widetilde{e} \Downarrow \langle \ell \rhd m \rangle\:\gate
                }
@@
                  \widetilde{\rho} \vdash \widetilde{e} \Downarrow c\\\\
-                 c = \top \Rightarrow \widetilde{e'} = \widetilde{e_{1}}\\
-                 c = \bot \Rightarrow \widetilde{e'} = \widetilde{e_{2}}\\\\
-                 \widetilde{\rho} \vdash \widetilde{e'} \Downarrow \val
-               }
-               {
-                 \widetilde{\rho} \vdash \lcond{\mathcal{P}}{\widetilde{e}}{\widetilde{e_{1}}}{\widetilde{e_{2}}} \Downarrow \val
+                 c = \top \Rightarrow \widetilde{e}_{3} = \widetilde{e}_{1}\\
+                 c = \bot \Rightarrow \widetilde{e}_{3} = \widetilde{e}_{2}\\\\
+                 \widetilde{\rho} \vdash \widetilde{e}_{3} \Downarrow \val
+               }
+               {
+                 \widetilde{\rho} \vdash \lcond{\mathcal{P}}{\widetilde{e}}{\widetilde{e}_{1}}{\widetilde{e}_{2}} \Downarrow \val
                }
@@
   \\\\
-    \fbox{$\widetilde{\rho} \vdash \widetilde{s} \Downarrow \widetilde{\rho_1}; \crct$}
+    \fbox{$\widetilde{\rho} \vdash \widetilde{s} \Downarrow \widetilde{\rho}_{1}; \crct$}
   \[
@@
                  \widetilde{\rho} \vdash \widetilde{e} \Downarrow \val\\\\
-                 \left(\val = c\right) \vee \left(\val = [\overline{c_{i}}]_{n}\right)\\\\
-                 \widetilde{\rho}_{1} = \widetilde{\rho}, x \mapsto \val
-               }
-               {
-                 \widetilde{\rho} \vdash \tau\:x = \widetilde{e} \Downarrow \widetilde{\rho}_{1}; \cdot
-               }
-
-               \hspace{0.5cm}
+                 \left(\val = c\right) \vee \left(\val = [\overline{c_{i}}]_{n}\right)
+               }
+               {
+                 \widetilde{\rho} \vdash \tau\:x = \widetilde{e} \Downarrow \widetilde{\rho}, x \mapsto \val; \cdot
+               }
+
+               \hspace{0.2cm}
 
@@
                {
-                 \widetilde{\rho} \vdash \widetilde{e} \Downarrow \gate\quad
-                 \mathsf{fresh}\:\wire\\\\
-                 \widetilde{\rho}_{1} = \widetilde{\rho}, x \mapsto \wire\quad
-                 \crct = \kw{bind}\:\gate\:\wire
-               }
-               {
-                 \widetilde{\rho} \vdash \tau\:x = \widetilde{e} \Downarrow \widetilde{\rho}_{1}; \crct
+                 \widetilde{\rho} \vdash \widetilde{e} \Downarrow \gate\quad\quad
+                 \mathsf{fresh}\:\wire
+               }
+               {
+                 \widetilde{\rho} \vdash \tau\:x = \widetilde{e} \Downarrow \widetilde{\rho}, x \mapsto \wire; \kw{bind}\:\gate\:\wire
                }
@@
                  \widetilde{\rho} \vdash \widetilde{e} \Downarrow [\overline{\gate_{i}}]_{n}\\\\
-                 \forall i \in [n].\:\mathsf{fresh}\:\wire_{i}\\\\
-                 \widetilde{\rho}_{1} = \widetilde{\rho}, x \mapsto [\overline{\wire_{i}}]_{n}\quad
-                 \crct = \overline{\kw{bind}\:\gate_{i}\:\wire_{i}}
-               }
-               {
-                 \widetilde{\rho} \vdash \tau\:x = \widetilde{e} \Downarrow \widetilde{\rho}_{1}; \chi
+                 \forall i \in [n].\:\mathsf{fresh}\:\wire_{i}
+               }
+               {
+                 \widetilde{\rho} \vdash \tau\:x = \widetilde{e} \Downarrow \widetilde{\rho}, x \mapsto [\overline{\wire_{i}}]_{n}; \overline{\kw{bind}\:\gate_{i}\:\wire_{i}}
                }
@@
                  \widetilde{\rho} \vdash x \Downarrow [\overline{\wire_{i}}]_{n}\quad
-                 \widetilde{\rho} \vdash \widetilde{e_{1}} \Downarrow n_{1}\\\\
+                 \widetilde{\rho} \vdash \widetilde{e}_{1} \Downarrow n_{1}\\\\
                  n_{1} < n\quad
                  \mathsf{fresh}\:\wire\quad
-                 \widetilde{\rho} \vdash \widetilde{e_{2}} \Downarrow \gate\\\\
+                 \widetilde{\rho} \vdash \widetilde{e}_{2} \Downarrow \gate\\\\
                  \widetilde{\rho}_{1} = \widetilde{\rho}[x \mapsto ([\overline{\wire_{i}}]_{n}[n_{1} \mapsto \wire])]
@@
                {
-                 \widetilde{\rho} \vdash x[\widetilde{e_{1}}] := \widetilde{e_{2}} \Downarrow \widetilde{\rho}_{1}; \kw{bind}\:\gate\:\wire
+                 \widetilde{\rho} \vdash x[\widetilde{e}_{1}] := \widetilde{e}_{2} \Downarrow \widetilde{\rho}_{1}; \kw{bind}\:\gate\:\wire
                }
@@
 are $\wire$, input gates, \kw{mult}, \kw{gt}, and \kw{mux} gates, 
-coerce gates $\gate \rhd m$, and constant gates.  Sub-circuits $\val$ 
-(generated from $\widetilde e$)
-then consist of gates and arrays of gates.
-
- \figureref{eval-inter-circ} shows the judgments for the evaluation of the intermediate language to a circuit. 
-The runtime environment maps variables to sub-circuits and is defined as follows:
+coerce gates $\langle \ell \rhd m \rangle\:\gate$, and constants.
+Sub-circuits $\val$ (generated from $\widetilde e$) then consist of
+gates and arrays of gates.
+
+\figureref{eval-inter-circ} shows the judgments for the evaluation of
+the intermediate language to a circuit. The runtime environment maps
+variables to sub-circuits and is defined as follows:
 
@@
 the significance of the operator labels. In particular, the rule
-({\sc{S-PMult}}) evaluates a public multiplication $\widetilde{e_{1}}
-\mulop_{\mathcal{P}} \widetilde{e_{2}}$ to $n_{1} \mulop n_{2}$, similar to the
+({\sc{S-PMult}}) evaluates a public multiplication $\widetilde{e}_{1}
+\mulop_{\mathcal{P}} \widetilde{e}_{2}$ to $n_{1} \mulop n_{2}$, similar to the
 source semantics of \figureref{srcsem}. In contrast, the rule
-({\sc{S-SMult}}) evaluates a secret multiplication $\widetilde{e_{1}}
-\mulop_{\mathcal{A}} \widetilde{e_{2}}$ to an arithmetic multiplication
+({\sc{S-SMult}}) evaluates a secret multiplication $\widetilde{e}_{1}
+\mulop_{\mathcal{A}} \widetilde{e}_{2}$ to an arithmetic multiplication
 gate $\kw{mult}\:\gate_{1}\:\gate_{2}$. As mentioned above, the intermediate language
-expressions generated by our compiler never have $\widetilde{e_{1}}
-\mulop_{\mathcal{B}} \widetilde{e_{2}}$, as our compiler is aware that $\mulop$
+expressions generated by our compiler never have $\widetilde{e}_{1}
+\mulop_{\mathcal{B}} \widetilde{e}_{2}$, as our compiler is aware that $\mulop$
 is more performant using an arithmetic circuit compared to a boolean
@@
 is empty. When 
-$\widetilde{e}$ evaluates to a circuit $\gate$, rule
+$\widetilde{e}$ evaluates to a sub-circuit $\gate$, rule
 ({\sc{S-DeclCkt}}) picks a fresh wire $\wire$, adds the mapping $x \mapsto \wire$
@@
 $\kw{bind}\:\gate\:\wire$. Rule ({\sc{S-DeclCktA}}) is analogous for
-$\widetilde{e}$ evaluating to an array of circuits. The variable
+$\widetilde{e}$ evaluating to array of sub-circuits. The variable
 assignment rules (not shown in the figure) are similar. Rule
@@
 
-\subsubsection*{Circuit semantics} Evaluating a program in the intermediate language produces a
-circuit over the secret data to be computed using a \mpc protocol in the distributed setting. With
-our runtime circuit semantics, we model the \emph{functional} aspect of a \mpc protocol, parametrized by cryptographic encoding and decoding functions.
-
-During the circuit evaluation, the wire ranges $r$ are
+\subsubsection*{Circuit semantics} Evaluating a program in the
+intermediate language produces a
+circuit to be computed using a distributed \mpc protocol. With
+our circuit semantics, we model the \emph{functional} aspect
+of a \mpc protocol, parametrized by cryptographic encoding and
+decoding functions.
+
+During the circuit evaluation, the wire ids $w$ are
 mapped to (random) strings $b$. The semantics of these strings is
@@
 More concretely,
-$\mathcal{E}_{m}(c)$ returns a pair of two strings $b_{1}$ and $b_{2}$
+$\mathcal{E}_{m}(c)$ returns a pair of strings $(b_{1}, b_{2})$
  with the property that
@@
 protocol instantiates $\mathcal{E}_{m}$ and $\mathcal{D}_{m}$
-appropriately.
-For ABY protocol \cite{aby}, algorithms $(\mathcal{E}_{\mathcal{A}},\mathcal{D}_{\mathcal{A}})$ (resp.
-$(\mathcal{E}_{\mathcal{B}},\mathcal{D}_{\mathcal{B}})$) correspond to the arithmetic (resp.
-boolean) sharing and reconstruction algorithms.
-
-\figureref{cktsem} gives the judgements for evaluation of circuits by the two parties using the \mpc protocol. The circuit environment  is a map from wires to shares and is defined as follows:
+appropriately. For ABY protocol \cite{aby}, algorithms
+$(\mathcal{E}_{\mathcal{A}},\mathcal{D}_{\mathcal{A}})$
+(resp. $(\mathcal{E}_{\mathcal{B}},\mathcal{D}_{\mathcal{B}})$)
+correspond to the arithmetic (resp. boolean) sharing and
+reconstruction algorithms.
+
+\figureref{cktsem} gives the judgments for evaluation of circuits by
+the two parties using a \mpc protocol. The circuit environment  is a
+map from wires to shares:
 
@@
 where $O$ are the observations (similar to source semantics). 
-Rule ({\sc{C-In}}) is used to evaluate the input gate $\kw{in}^{m}_{j}$ that creates the $m$ type shares of the value $c$ input by party $j$.
+Rule ({\sc{C-In}}) evaluates the input gate $\kw{in}^{m}_{j}$, and
+creates the $m$-type shares of the value $c$ input by party $j$.
 Rule ({\sc{C-Mult}}) evinces the pattern for evaluating circuit
@@
 evaluates, of course, concretely $n_{1}$ and $n_{2}$ are not observed
-by the parties. Rule ({\sc{C-Coerce}}) creates the shares using  $\mathcal{E}_m$. The evaluation
+by the parties. Rule ({\sc{C-Coerce}}) creates the new shares using
+$\mathcal{E}_m$ (the corresponding rule for coercion from
+$\mathcal{P}$ is similar). The evaluation
 of \kw{bind} updates the mapping of $\wire$ in the input environments, and
@@
   \small
-  \fbox{$\widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \gate \Downarrow b_{1}, b_{2}$}
+  \fbox{$\widehat{\rho}_{1}, \widehat{\rho}_{2} \vdash \gate \Downarrow b_{1}, b_{2}$}
   \[
@@
                {
-                 \widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \kw{in}^{m}_{j} \Downarrow b_{1}, b_{2}
+                 \widehat{\rho}_{1}, \widehat{\rho}_{2} \vdash \kw{in}^{m}_{j} \Downarrow b_{1}, b_{2}
                }
@@
                {
-                 \widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \gate \Downarrow b_{1}, b_{2}\\\\
-                 c = \mathcal{D}_{m_{1}}(b_{1}, b_{2})\\\\
-                 (b'_{1}, b'_{2}) = \mathcal{E}_{m}(c)
-               }
-               {
-                 \widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \gate \rhd m \Downarrow b'_{1}, b'_{2}
-               }
+                 \widehat{\rho}_{1}, \widehat{\rho}_{2} \vdash \gate \Downarrow b_{1}, b_{2}\\\\
+                 (b'_{1}, b'_{2}) = \mathcal{E}_{m}(\mathcal{D}_{m_{1}}(b_{1}, b_{2}))
+               }
+               {
+                 \widehat{\rho}_{1}, \widehat{\rho}_{2} \vdash \langle
+                 m_{1} \rhd m \rangle\:g \Downarrow b'_{1}, b'_{2}
+               }
+
                \\\\
@@
                {
-                 \forall i \in \{1, 2\}.\:\widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \gate_{i} \Downarrow b_{1i}, b_{2i}\quad
+                 \forall i \in \{1, 2\}.\:\widehat{\rho}_{1}, \widehat{\rho}_{2} \vdash \gate_{i} \Downarrow b_{1i}, b_{2i}\quad
                  n_{i} = \mathcal{D}_{\mathcal{A}}(b_{1i}, b_{2i})\\\\
@@
                {
-                 \widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \kw{mult}\:\gate_{1}\:\gate_{2} \Downarrow b_{1}, b_{2}
+                 \widehat{\rho}_{1}, \widehat{\rho}_{2} \vdash \kw{mult}\:\gate_{1}\:\gate_{2} \Downarrow b_{1}, b_{2}
                }
@@
                {
-                 \forall i \in \{1, 2\}.\:\widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \gate_{i} \Downarrow b_{1i}, b_{2i}\quad
+                 \forall i \in \{1, 2\}.\:\widehat{\rho}_{1}, \widehat{\rho}_{2} \vdash \gate_{i} \Downarrow b_{1i}, b_{2i}\quad
                  n_{i} = \mathcal{D}_{\mathcal{B}}(b_{1i}, b_{2i})\\\\
@@
                {
-                 \widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \kw{gt}\:\gate_{1}\:\gate_{2} \Downarrow b_{1}, b_{2}
+                 \widehat{\rho}_{1}, \widehat{\rho}_{2} \vdash \kw{gt}\:\gate_{1}\:\gate_{2} \Downarrow b_{1}, b_{2}
                }
@@
                {
-                 \forall i \in \{1, 2, 3\}.\:\widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \gate_{i} \Downarrow b_{1i}, b_{2i}\quad
+                 \forall i \in \{1, 2, 3\}.\:\widehat{\rho}_{1}, \widehat{\rho}_{2} \vdash \gate_{i} \Downarrow b_{1i}, b_{2i}\quad
                  c_{i} = \mathcal{D}_{\mathcal{B}}(b_{1i}, b_{2i})\\\\
@@
                {
-                 \widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \kw{mux}\:\gate_{1}\:\gate_{2}\:\gate_{3} \Downarrow b_{1}, b_{2}
+                 \widehat{\rho}_{1}, \widehat{\rho}_{3} \vdash \kw{mux}\:\gate_{1}\:\gate_{2}\:\gate_{3} \Downarrow b_{1}, b_{2}
                }
@@
   \\\\
-    \fbox{$\widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \crct \Downarrow \widehat{\rho'_{1}}, \widehat{\rho'_{2}}; O$}
+    \fbox{$\widehat{\rho}_{1}, \widehat{\rho}_{2} \vdash \crct \Downarrow \widehat{\rho}'_{1}, \widehat{\rho}'_{2}; O$}
   \[
@@
                {
-                 \widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \gate \Downarrow b_{1}, b_{2}\\\\
-                 \widehat{\rho'_{1}} = \widehat{\rho_{1}}[\wire \mapsto b_{1}] \quad
-                 \widehat{\rho'_{2}} = \widehat{\rho_{2}}[\wire \mapsto b_{2}]
-               }
-               {
-                 \widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \kw{bind}\:\gate\:\wire \Downarrow \widehat{\rho'_{1}}, \widehat{\rho'_{2}}; \cdot
+                 \widehat{\rho}_{1}, \widehat{\rho}_{2} \vdash \gate \Downarrow b_{1}, b_{2}\\\\
+                 \widehat{\rho}'_{1} = \widehat{\rho}_{1}[\wire \mapsto b_{1}] \quad
+                 \widehat{\rho}'_{2} = \widehat{\rho}_{2}[\wire \mapsto b_{2}]
+               }
+               {
+                 \widehat{\rho}_{1}, \widehat{\rho}_{2} \vdash \kw{bind}\:\gate\:\wire \Downarrow \widehat{\rho}'_{1}, \widehat{\rho}'_{2}; \cdot
                }
@@
                {
-                 \widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \gate \Downarrow b_{1}, b_{2}\\\\
+                 \widehat{\rho}_{1}, \widehat{\rho}_{2} \vdash \gate \Downarrow b_{1}, b_{2}\\\\
                  c = \mathcal{D}_{m}(b_{1}, b_{2})
@@
                {
-                 \widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash \kw{out}\:\gate \Downarrow \widehat{\rho_{1}}, \widehat{\rho_{2}}; c
+                 \widehat{\rho}_{1}, \widehat{\rho}_{2} \vdash \kw{out}\:\gate \Downarrow \widehat{\rho}_{1}, \widehat{\rho}_{2}; c
                }
@@
 %\subsubsection*{Correctness theorem}
-Finally, the correctness theorem is as follows (the environments on the left of $\vdash$ are empty and we elide the environments on the right of $\vdash$ for brevity):
+Finally, the correctness theorem is as follows (the environments on
+the left of $\vdash$ are empty and we elide the environments on the
+right of $\vdash$ for brevity):
 
 \begin{theorem}[Correctness]\label{theorem:correctness}
- $\forall\: s, \widetilde{s}$, if $\vdash s \leadsto \widetilde{s} \mid \_$  then $\exists  O , \chi$, s.t.
-$\vdash s \Downarrow \_; O$, $ \vdash \widetilde{s} \Downarrow \_; \chi$, and $ \vdash \chi \Downarrow \_, \_; O$
+ $\forall\: s, \widetilde{s}$, if $\vdash s \leadsto \widetilde{s}
+  \mid \_$  then $\exists  O , \chi$, s.t.  $\vdash s \Downarrow \_;
+  O$, $ \vdash \widetilde{s} \Downarrow \_; \chi$, and $ \vdash \chi
+  \Downarrow \_, \_; O$.
 \end{theorem}
@@
 The theorem states that if a source statement $s$ is well-typed, and
-compiles to $\widetilde{s}$ in intermediate language, then $s$ terminates in the source
+compiles to $\widetilde{s}$ in the intermediate language, then $s$ terminates in the source
 semantics with observations $O$, $\widetilde{s}$ evaluates to circuit $\crct$, and $\crct$ terminates in the

### 2017-10-31 OaklandSubmission/eval.tex

@@
 (4) Can we specify somewhere the dimensions of these matrices?}
+\vspace{-10pt}
 \[
@@
 \]
-
 Here, $I_j(x)$ is 1 if the $j^{th}$ node is present on the path traversed by $x$
@@
 
- \caption{Partitioning results for CIFAR-10. \minion takes 544 seconds and communicates 9272 Mb to evaluate this DNN.}
+ \caption{Partitioning results for CIFAR-10. \minion takes 544 seconds and communicates 9272 Mb.}
  \label{tab:cifar} 
@@
 \noindent\textbf{Matrix factorization}
-\tool, as a language, is not tied to secure prediction and can express more general computations.
-To demonstrate the expressiveness of \tool, we use it to implement secure matrix factorization~\cite{valeriaMatrix}. Abstractly, given a sparse matrix $\mathcal{M}$ of dimensions
+\tool is not tied to secure prediction and can express more general computations.
+To demonstrate this expressiveness, we  implement secure matrix factorization~\cite{valeriaMatrix}. Abstractly, given a sparse matrix $\mathcal{M}$ of dimensions
 $n\times m$ and $M$ non-zero entries, the goal is to generate a matrix $U$ of dimension $n\times d$ and a matrix

### 2017-10-31 OaklandSubmission/security.tex

@@
 
-We shall assume a cryptographic \mpc backend that securely implements any circuit $\crct$ that is output by our compiler (see \figureref{circuits}).
-In more detail, this means that for any well-typed source program $s$, let $\crct$ be the circuit output by our compiler (as in \theoremref{correctness}).
-We assume that there exists a \mpc protocol $\prot$ that securely realizes the functionality $\crct$ and let $\simu_{\tiny{2pc}}$ be the corresponding simulator  (that runs on $\crct$, input of the corrupt party and the output obtained from trusted functionality for $\crct$).
-We note that the work of \cite{aby} provides such a protocol $\prot$ and simulator $\simu_{\tiny{2pc}}$ for all circuits $\crct$ output by our compiler.
-We are now ready to state and prove our security theorem.
+We shall assume a cryptographic \mpc backend that securely implements
+any circuit $\crct$ that is output by our compiler
+(see \figureref{circuits}). In more detail, this means that for any
+well-typed source program $s$, let $\crct$ be the output circuit output
+(as in \theoremref{correctness}). We assume that there
+exists a \mpc protocol $\prot$ that securely realizes the
+functionality $\crct$ and let $\simu_{\tiny{2pc}}$ be the
+corresponding simulator  (that runs on $\crct$, input of the corrupt
+party and the output obtained from trusted functionality for
+$\crct$). We note that ABY~\cite{aby} provides such a protocol
+$\prot$ and simulator $\simu_{\tiny{2pc}}$ for all circuits $\crct$
+output in our framework. We now state and prove our
+security theorem.
 
 \begin{theorem}[Security]\label{theorem:security}
-Let $s$ be any functionality or program in our source language with outputs (or observations) $O_1$, that is compiled into a circuit $\chi$ (as defined in Theorem \ref{theorem:correctness}).
-Let protocol $\prot$ be the two-party secure computation protocol that securely realizes $\chi$ (as defined above).
-Then, $\prot$ securely realizes $s$.
+Let $s$ be any functionality or program in our source language with
+outputs (or observations) $O_1$, that outputs a circuit
+$\chi$ (as defined in Theorem \ref{theorem:correctness}). Let protocol
+$\prot$ be the two-party secure computation protocol that securely
+realizes $\chi$ (as defined above). Then, $\prot$ securely realizes
+$s$.
 \end{theorem}

### 2017-10-31 OaklandSubmission/eval.tex
minor

@@
 dataset~\cite{cifar} from \minion~\cite{minionn} and matrix factorization~\cite{valeriaMatrix}
-to illustrate the use
-of partitioning in \tool.
+to evaluate partitioning.
 %Finally, we evaluate \tool on the task of matrix factorization~\cite{valeriaMatrix}.
@@
 (4) Can we specify somewhere the dimensions of these matrices?}
-\vspace{-10pt}
+%\vspace{-10pt}
 \[
@@
 %result in many boolean-and gates. %(13k/750k denote 13000/75000). 
-As before columns ``Prev. time'' and ``Prev. comm.'' (columns 4 and 5)
-show the numbers reported
-by Bost et al.~\cite{shafindss}. We observe that \tool generated code
+As before, we compare with Bost et al.~\cite{shafindss} and observe that \tool generated code
 has better performance, despite using a generic 2PC,
@@
 three columns.
-For this DNN, the specialized \minion protocol takes 544 seconds and
-communicates 9272 Mb.
+%For this DNN, the specialized \minion protocol takes 544 seconds and
+%communicates 9272 Mb.
 As with Table~\ref{tab:nn}, \tool generated generic 2PC protocol is
@@
 However, the current programmer effort seems miniscule compared to the mammoth implementation effort
-put in by Nikolaenko et al. (Section 5 of ~\cite{valeriaMatrix}) to scale a boolean circuit based
+put in by Nikolaenko et al. (Section 5 of ~\cite{valeriaMatrix}) to scale a boolean circuits based
 backend to this benchmark.

### 2017-10-31 OaklandSubmission/eval.tex
para changes.

--- 2017-10-31 OaklandSubmission/eval.tex
@@
 \subsection{Secure prediction}
-\noindent\textbf{Standard classifiers}
+\noindent\textbf{Standard classifiers.}
 %\label{sec:shallow}
@@
 \\\\
-\noindent\textbf{Deep neural nets}
+\noindent\textbf{Deep neural nets.}
 We evaluate \tool on the DNNs described in SecureML~\cite{secureml},
@@
 \\\\
-\noindent\textbf{State-of-the-art classifiers}
+\noindent\textbf{State-of-the-art classifiers.}
 Tensorflow~\cite{tensorflow} is a standard machine learning toolkit.

### 2017-10-31 OaklandSubmission/impl.tex
minor

@@
 We use the default ABY configuration for our
-experiment \footnote{We set the security parameter to 128 and use OT
+experiment\footnote{We set the security parameter to 128 and use OT
 extension-based arithmetic multiplication triples generation.}. ABY
@@
 operators are not supported natively by \tool. 
-However, we have implemented integral division in 30 lines of \tool,
+However, we have implemented integral division in 20 lines of \tool,
 while the floating-point support in ABY is under active
@@
 coerced to boolean).
-We leave a more sophisticates type inference procedure for future work.
+We leave a more sophisticated type inference procedure for future work.
 

### 2017-10-31 OaklandSubmission/intro.tex

@@
  expected to use them correctly and efficiently on their own.
- Ideally, a developer would declare the functionality in a general
+ Ideally, a developer would declare the functionality in a, general
  purpose, high-level programming language and a
- tool, e.g. a compiler, would generate an efficient protocol that
+ tool, e.g., a compiler would generate an efficient protocol that
  implements the functionality securely, while hiding the cryptography
@@
 where one party (the server) holds a proprietary classifier to predict
-a label (e.g. a disease, genomics, or spam detection), and the other
+a label (e.g., a disease, genomics, or spam detection), and the other
 party (the client) holds a private input that it wants to run the
@@
 $\cdot^{T}$ is the matrix transpose operator, and $w^Tx$ denotes the
-inner product of  $w^{T}$ and $x$. Alice has following options.
+inner product of  $w$ and $x$. Alice has following options.
 
@@
 boolean circuits. However, here, the programming interface is quite
-low-level -- the programmer is required to first manually split the
+low-level: the programmer is required to first manually split the
 computation into arithmetic and boolean components, and then write the

### 2017-10-31 OaklandSubmission/ex.tex
abstract
example fonts
comment

@@
 
-\input{fig-ex-aby}
-
 \subsubsection*{\tool compiler}
 \tool compiler takes as input a source program and produces a C++
-program as output. Figure~\ref{fig:ex-aby} shows the output code for
+program as output. Figure~\ref{fig:ex-aby}\footnote{This is also how a program written directly in
+  ABY would look like.}. shows the output code for
 the example in Figure~\ref{fig:ex-sml}. The output program
@@
 (\ls{role == SERVER} and \ls{role == CLIENT}), and common code for the
-computation.
+computation
 
@@
 \ls{ycirc} to build the corresponding circuits
-(lines~\ref{line:condyaobegin} to~\ref{line:condyaoend})\footnote{In our usage of ABY, boolean is synonymous
-  with Yao (Section~\ref{sec:impl}).}.
+(lines~\ref{line:condyaobegin} to~\ref{line:condyaoend}) (in our usage
+of ABY, boolean is synonymous with Yao (Section~\ref{sec:impl})).
 
@@
 lines~\ref{line:convacc} and~\ref{line:convb}, the compiler converts
-\ls{a_acc} and \ls{a_b} to boolean representation, before they are
+\ls{a\_acc} and \ls{a\_b} to boolean representation, before they are
 input to the comparison and multiplexer circuits.
+
+\begin{figure}
+\begin{lstlisting}[language=C, mathescape=true]
+//circuit builders for arithmetic and boolean
+$\mathtt{
+Circuit\;*ycirc = s[S\_YAO]\rightarrow GetCircuitBuildRoutine();
+}$
+$\mathtt{
+Circuit\;*acirc = s[S\_ARITH]\rightarrow GetCircuitBuildRoutine();
+}$
+$\dots$
+$\mathtt{
+if(role == SERVER)\;\{ 
+}$
+  //Put gates to read w and b
+$\mathtt{
+\}\;else\;\{\;\;\text{{\color{red}//role == CLIENT}}
+}$  
+  //Put gates to read x
+$\mathtt{
+\}
+}$
+
+$\mathtt{
+for(uint32\_t\;i\;=\;0;\;i\;<\;30;\;i = i + 1)\;\{\;\;\text{\color{red}//acc = w}^{\text{{\color{red}T}}}\text{\color{red}x}~\label{line:dotproductloop}
+}$
+$\mathtt{
+\;\;\;\;share\;*a\_t\_0 = acirc\rightarrow PutMULGate(a\_w[i], a\_x[i]);~\label{line:dotmulgate}
+}$
+$\mathtt{
+\;\;\;\;a\_acc = acirc\rightarrow PutADDGate(a\_acc, a\_t\_0 );~\label{line:dotaddgate}
+}$
+$\mathtt{
+\}
+}$
+
+//convert acc and b from arithmetic to boolean
+$\mathtt{
+share\;*y\_acc = ycirc\rightarrow PutA2YGate(a\_acc);~\label{line:convacc}
+}$
+$\mathtt{
+share\;*y\_b = ycirc\rightarrow PutA2YGate(a\_b);~\label{line:convb}
+}$
+
+$\mathtt{
+share\;*y\_pred = ycirc\rightarrow PutGTGate(y\_acc, y\_b);~\label{line:condyaobegin}
+}$
+$\mathtt{
+uint32\_t\;one = 1;
+}$
+$\mathtt{
+share\;*y\_1 = ycirc\rightarrow PutCONSGate(one, bitlen);
+}$
+$\mathtt{
+uint32\_t\;zero = 0 ;
+}$
+$\mathtt{
+share\;*y\_0 = ycirc\rightarrow PutCONSGate(zero, bitlen);
+}$
+$\mathtt{
+share\;*y\_t = ycirc\rightarrow PutMUXGate(y\_pred, y\_1, y\_0);~\label{line:condyaoend}
+}$
+
+$\mathtt{
+share\;*y\_out = ycirc\rightarrow PutOUTGate(y\_t, CLIENT);
+}$
+$\mathtt{
+party\rightarrow ExecCircuit();
+}$
+
+$\mathtt{
+if(role == CLIENT) \{\;\;\text{\color{red}//only to the client}
+}$
+$\mathtt{
+\;\;\;\;uint32\_t \_o = y\_out\rightarrow get\_clear\_value\langle uint32\_t \rangle();
+}$
+$\mathtt{
+\}
+}$
+\end{lstlisting}
+\caption{\tool compiler (partial) output for Figure~\ref{fig:ex-sml}}
+\label{fig:ex-aby}
+\end{figure}
 
@@
 guarantees.
+
 %% Consider a  cloud service provider Bob that wants to provide a service to diagnose

### 2017-10-31 OaklandSubmission/intro.tex
changes in example
example fonts
spelling in intro

@@
 about the client's input or output, and that the client learns nothing
-about the classifier, beyond what is revelaed by the output
+about the classifier, beyond what is revealed by the output
 label.
@@
 
-Alternatively, Alice can use a tool such as ABY (Demmlet et
+Alternatively, Alice can use a tool such as ABY (Demmler et
 al.~\cite{aby}) that allows
@@
 \begin{figure}
-\begin{lstlisting}[language=C]
-uint w[30] = input1(); uint b = input1();
-uint x[30] = input2();
-uint acc = 0;
-for i in [0:30] { acc =  acc + (w[i] * x[i]); }
-output2((acc > b) ? 1 : 0) //only to party 2
+\begin{lstlisting}[language=C,mathescape=true]
+$\mathtt{
+uint\;w[30] = input1();\;\;uint\;b = input1();
+}$
+$\mathtt{
+uint\;x[30] = input2();
+}$
+$\mathtt{
+uint\;acc = 0;
+}$
+$\mathtt{
+for\;i\;in\;[0:30]\;\{\;acc\;=\;acc\;+\;(w[i]\;\times\;x[i]);\;\}
+}$
+$\mathtt{
+output2((acc > b)\;?\;1\;:\;0)\;\text{{\color{red}//only to party 2}}
+}$
 \end{lstlisting}
@@
 
-\subsubsection*{Cryptogaphic-cost aware compiler} The \tool compiler
+\subsubsection*{Cryptographic-cost aware compiler} The \tool compiler
 compiles a source program to a hybrid computation consisting
@@
 corresponding circuits along with the required inter-conversion
-gates. The outcome is an effcient \mpc protocol combining arithmetic
+gates. The outcome is an efficient \mpc protocol combining arithmetic
 and boolean circuits, while the programmer remains

### 2017-10-31 OaklandSubmission/ex.tex
spelling in ex

@@
 cost-aware'', and appropriately picks either arithmetic or boolean
-circuit representations for different sub-componenets. For example,
+circuit representations for different sub-components. For example,
 the compiler realizes that the inner product computation is more

### 2017-10-31 OaklandSubmission/related.tex
Spelling in related

@@
  Mitchell {\em et al.}~\cite{lambdaps} allow the user to select between Shamir's secret sharing~\cite{sss} and fully homomorphic encryption~\cite{gentry}.
-Unlike \tool, all these tools use either a arithmetic backend or a boolean backend but not a combination of both.
+Unlike \tool, all these tools use either an arithmetic backend or a boolean backend but not a combination of both.
 
@@
 %The reason for doing this was to allow for a mix between arithmetic and boolean computations as a secure computation protocol for a pure boolean circuit would have been very inefficient. 
-This process require deep knowledge of cryptography
+This process requires deep knowledge of cryptography
 to ensure that the protocols are secure.

### 2017-10-31 OaklandSubmission/intro.tex
nit

@@
 programmers to write \mpc protocols. E.g. we provide the first \mpc
-implementation of \bonsai~\cite{bonsai}, and it was programmed in the
+implementation of the \bonsai~\cite{bonsai} classifier, and it was programmed in the
 high-level \tool source language by a non-cryptographer. Second, the
@@
 of the protocols
-generated by \tool are comparable to or better than (up to 19x) their state-of-the-art specialized implementations. Finally, we
-show that partitioning enables us to implement applications
-that require more than 300 million gates (Section~\ref{sec:impl} and
+generated by \tool are comparable to or better than (up to 19x) their state-of-the-art, hand-crafted implementations. Finally, we
+demonstrate the usefulness of partitioning by implementing an application
+that requires more than 300 million gates (Section~\ref{sec:impl} and
 Section~\ref{sec:eval}).
@@
 combinations (e.g. \cite{bogdanov,lone,tasty,autoS}). However, as Demmler et
-al.~\cite{aby} note, due
- to high conversion cost between
+al.~\cite{aby} observe, due
+ to the high conversion cost between
 homomorphic encryption and Yao's garbled circuits, these combined

### 2017-10-31 OaklandSubmission/eval.tex
para heads

@@
 layers is larger than those of convolution layers.
-In general, DNNs are computationaly heavy but provide
+In general, DNNs are computationally heavy but provide
 much better accuracies on computer vision tasks than the classifiers
@@
 usually approximated by floating-point operations. 
-As is standard, we port the classifers to integer manipulating
+As is standard, we port the classifiers to integer manipulating
 programs by scaling the models and rounding~\cite{minionn}. These
@@
 vector.
-We implement this classifer in \tool and present the results in the
+We implement this classifier in \tool and present the results in the
 first row of Table~\ref{tab:tf}. We note that a pure boolean circuit for this simple functionality does not fit in the memory of our machines.
 
-The next classifer in the Tensorflow tutorial is a convolution neural net with two convolutions
+The next classifier in the Tensorflow tutorial is a convolution neural net with two convolutions
 (with 32 output channels) and two fully connected layers with ReLU as the activation function.
@@
 from~\cite{bonsai},
-and show the represenative results in Table~\ref{tab:bonsai}.
+and show the representative results in Table~\ref{tab:bonsai}.
 Out of all the benchmarks from~\cite{bonsai}, the dataset WARD
@@
 \label{sec:pipeeval}
-\noindent\textbf{CIFAR-10}
+\noindent\textbf{CIFAR-10.}
 The largest benchmark of \minion~\cite{minionn} is a DNN for CIFAR-10
@@
 \\\\
-\noindent\textbf{Matrix factorization}
+\noindent\textbf{Matrix factorization.}
 \tool is not tied to secure prediction and can express more general computations.
@@
 The time reported in~\cite{valeriaMatrix}
-for one iteration is 2.9 hours\footnote{~\cite{valeriaMatrix} does not report the network round-trip time.}. This compuation is large enough that we partition each iteration
+for one iteration is 2.9 hours\footnote{~\cite{valeriaMatrix} does not report the network round-trip time.}. This computation is large enough that we partition each iteration
 into three stages. The first stage involves a Batcher~\cite{Batcher} sorting network followed by a linear pass.

### 2017-10-31 OaklandSubmission/ex.tex
color
minor

@@
 expressions (the ternary $?\::$ operator), \ls{for}~loops,
-\ls{if}~statements, and special syntax for input/output.
+\ls{if}~statements, and special syntax for input/output from each party.
 
@@
 \tool compiler takes as input a source program and produces a C++
-program as output. Figure~\ref{fig:ex-aby}\footnote{This is also how a program written directly in
-  ABY would look like.}. shows the output code for
+program as output. Figure~\ref{fig:ex-aby} shows the output code\footnote{This is also how a program written directly in
+  ABY would look like.} for
 the example in Figure~\ref{fig:ex-sml}. The output program
@@
 $\mathtt{
-\}\;else\;\{\;\;\text{{\color{red}//role == CLIENT}}
+\}\;else\;\{\;\;\textbf{{\color{blue}//role == CLIENT}}
 }$  
@@
 $\mathtt{
-for(uint32\_t\;i\;=\;0;\;i\;<\;30;\;i = i + 1)\;\{\;\;\text{\color{red}//acc = w}^{\text{{\color{red}T}}}\text{\color{red}x}~\label{line:dotproductloop}
+for(uint32\_t\;i\;=\;0;\;i\;<\;30;\;i = i + 1)\;\{\;\;\textbf{\color{blue}//acc = w}^{\textbf{{\color{blue}T}}}\textbf{\color{blue}x}~\label{line:dotproductloop}
 }$
@@
 $\mathtt{
-if(role == CLIENT) \{\;\;\text{\color{red}//only to the client}
+if(role == CLIENT) \{\;\;\textbf{\color{blue}//only to the client}
 }$
@@
 public parts of the program, including the array accesses, and
-generates a \mpc circuit comprising of arithmetic and boolean gates,
+generates a  circuit comprising of arithmetic and boolean gates,
 with appropriate conversion gates. The circuit is then evaluated using

### 2017-10-31 OaklandSubmission/intro.tex
color
minor

@@
  using cryptographic techniques. Typical developers lack a deep
- understanding of cryptographic protocols, and therefore cannot be
+ understanding of cryptographic protocols, and cannot be
  expected to use them correctly and efficiently on their own.
@@
 state-of-the-art, let us consider an example underlying many
-secure prediction algorithms. Say Alice wants to write
+secure prediction algorithms. Suppose Alice wants to write
 a \mpc protocol to securely
@@
 $\mathtt{
-output2((acc > b)\;?\;1\;:\;0)\;\text{{\color{red}//only to party 2}}
+output2((acc > b)\;?\;1\;:\;0)\;\textbf{{\color{blue}//only to party 2}}
 }$
@@
 to the rescue; it automatically inserts the required instrumentation
-to ensure security of these intermediate outputs. As
+to ensure security of these intermediate outputs (\sectionref{pipe}). As
 we show in our
@@
 compiled program. The theorem
-guarantees that for a well-typed source program, the two semantics
+guarantees that for all well-typed source programs, the two semantics
 successfully terminate
-(e.g. there are no array index out-of-bounds errors), with identical
+(e.g., there are no array index out-of-bounds errors) with identical
 observable outputs. For
 the security theorem, we formally reduce the security
-of our scheme, against semi-honest (or, ``honest but curious") adversaries,
- to the security of the  \mpc back-end. 
+of our scheme against semi-honest (or, ``honest but curious") adversaries
+ to the semi honest security of the  \mpc back-end. 
  We also prove a formal security theorem against semi-honest  adversaries for our partitioning
-scheme (Section~\ref{sec:ld} and Section~\ref{sec:pipe}).
+scheme (\sectionref{ld} and~\sectionref{pipe}).
 
@@
 demonstrate the usefulness of partitioning by implementing an application
-that requires more than 300 million gates (Section~\ref{sec:impl} and
-Section~\ref{sec:eval}).
+that requires more than 300 million gates (\sectionref{impl} and~\sectionref{eval}).
 

### 2017-10-31 OaklandSubmission/ld.tex
minor

@@
 language as $[\kw{in}_{j}]_{n}$, which can then evaluate using the
-rule ({\sc{E-Arr}}) \footnote{The notation $\forall i \in [n]$ is read
-  as $\forall i \in \{0\dots n - 1\}.$}. The remaining rules are
+rule ({\sc{E-Arr}}). The notation $\forall i \in [n]$ is read
+  as $\forall i \in \{0\dots n - 1\}$. The remaining rules are
 straightforward, and are elided for space reasons.
@@
 $\Gamma$ maps variables $x$ to types $\tau$, says that
-under $\Gamma$, $e$ (in source language) compiles to $\widetilde{e}$
+under $\Gamma$, $e$ (in the source language) compiles to $\widetilde{e}$
 (in the intermediate language), where $\widetilde{e}$ has type $\tau$.
@@
 our compiler is cryptographic cost aware, it never compiles the
-multiplication to boolean multiplication $\mulop_{\mathcal{B}}$. In a similar manner,
+multiplication to boolean multiplication $\mulop_{\mathcal{B}}$ (\sectionref{impl}). In a similar manner,
 rule ({\sc{T-Gt}}) compiles $e_{1} > e_{2}$ to either public comparison,
@@
 values later as part of the circuit semantics). Circuit gates $\gate$
-are $\wire$, input gates, \kw{mult}, \kw{gt}, and \kw{mux} gates, 
-coerce gates $\langle \ell \rhd m \rangle\:\gate$, and constants.
+are wires $\wire$, input gates $\kw{in}^m_j$, multiplication gates \kw{mult}, comparison gates \kw{gt}, and multiplexer \kw{mux} gates, 
+coerce gates $\langle \ell \rhd m \rangle$, and constants.
 Sub-circuits $\val$ (generated from $\widetilde e$) then consist of
-gates and arrays of gates.
+gates and arrays of gates.  A circuit $\crct$ is either empty, \kw{bind}-ing of a
+circuit gate $\gate$ to wire $\wire$, \kw{out} gate, or a sequence of
+circuits.
 
 \figureref{eval-inter-circ} shows the judgments for the evaluation of
-the intermediate language to a circuit. The runtime environment maps
-variables to sub-circuits and is defined as follows:
+the intermediate language to a circuit. The circuit generation environment maps
+variables to sub-circuits:
 
@@
 \begin{array}{rrcl}
-    \ftext{Runtime environment} & \widetilde\rho &::=& \cdot \mid \widetilde\rho,x \mapsto \val\\
+    \ftext{Circuit generation environment} & \widetilde\rho &::=& \cdot \mid \widetilde\rho,x \mapsto \val\\
 \end{array}
@@
 are along similar lines. Rule ({\sc{S-PCond}}) evaluates a public
-conditional to the value from one of the branches, while the rule
-({\sc{S-SCond}}) evaluates to a multiplexer \kw{mux} gate that takes input
-circuits from the guard ($\gate_{1}$) and both the branches
-($\gate_{2}$ and $\gate_{3}$). Once again for performance reasons,
-the target expressions from our compiler do not have
+conditional to the sub-circuit  from one of the branches, while the rule
+({\sc{S-SCond}}) evaluates to a multiplexer \kw{mux} gate that takes as  input
+the sub-circuits from the guard ($\gate_{1}$) and both the branches
+($\gate_{2}$ and $\gate_{3}$). Recall, for performance reasons,
+the  expressions in the intermediate language generated by our compiler do not have
 $\lcond{\mathcal{A}}{e_{1}}{e_{2}}{e_{3}}$. Rules
@@
 the environment $\widetilde{\rho}$  to produce a new
-environment $\widetilde{\rho}_{1}$, and a circuit $\crct$. A circuit $\crct$ (shown
-in \figureref{circuits}) is either empty, \kw{bind}-ing of a
-circuit gate $\gate$ to wire $\wire$, \kw{out} gate, or a sequence of
-circuits. Rules ({\sc{S-DeclC}}), ({\sc{S-DeclCkt}}), and
+environment $\widetilde{\rho}_{1}$, and a circuit $\crct$. Rules ({\sc{S-DeclC}}), ({\sc{S-DeclCkt}}), and
 ({\sc{S-DeclCktA}}) show the variable declaration cases. Rule
@@
 to the environment $\widetilde{\rho}$, and outputs the circuit
-$\kw{bind}\:\gate\:\wire$. Rule ({\sc{S-DeclCktA}}) is analogous for
-$\widetilde{e}$ evaluating to array of sub-circuits. The variable
-assignment rules (not shown in the figure) are similar. Rule
+$\kw{bind}\:\gate\:\wire$. The rule ({\sc{S-DeclCktA}}) is analogous for
+$\widetilde{e}$ evaluating to an  array of sub-circuits. The variable
+assignment rules (not shown in the figure) are similar. The rule
 ({\sc{S-WriteCkt}}) shows the case for writing to an array, where the
@@
 $\mathcal{D}_{m}(b_{1}, b_{2}) = c$. 
-The string $b_i$ denotes the $i^{th}$ party's {\it share} of $c$.
+The string $b_j$ denotes the $j^{th}$ party's {\it share} of $c$.
 We assume that the underlying \mpc
@@
 (resp. $(\mathcal{E}_{\mathcal{B}},\mathcal{D}_{\mathcal{B}})$)
-correspond to the arithmetic (resp. boolean) sharing and
+correspond to the arithmetic (resp. boolean) secret-sharing and
 reconstruction algorithms.
@@
 where $O$ are the observations (similar to source semantics). 
+The former judgement evaluates a gate under the environments
+$\widehat{\rho}_j$ and generates shares $b_j$ of the gate's output.
 Rule ({\sc{C-In}}) evaluates the input gate $\kw{in}^{m}_{j}$, and
 creates the $m$-type shares of the value $c$ input by party $j$.
-Rule ({\sc{C-Mult}}) evinces the pattern for evaluating circuit
+Rule ({\sc{C-Mult}}) illustrates the pattern for evaluating circuit
 gates $\gate$. To evaluate $\kw{mult}\:\gate_{1}\:\gate_{2}$, the
@@
   \]
-\caption{Circuit semantics in a distributed runtime(selected rules)}
+\caption{Circuit semantics in a distributed runtime (selected rules)}
 \label{fig:cktsem}
@@
 circuit semantics with the same observations $O$. 
-Therefore, well typed programs always terminate (array indices are always in bounds, there are no unbounded loops, etc.) and the protocol produces the same outputs as the source semantics.
+Therefore, well typed programs always terminate (array indices are always in bounds, there are no unbounded loops, etc.) and the \mpc protocol produces the same outputs as the source program.
 % We first define a judgment for

### 2017-10-31 OaklandSubmission/conclude.tex
backend to implementation
nit
final edits

@@
 The generated protocols comprise combinations
-of arithmetic and boolean circuits and have performance comparable to
-or better than the custom, specialized
+of arithmetic and boolean circuits and have performance comparable to, or better than the custom, specialized
 protocols from previous works.
@@
 partitioning can be modified to provide malicious security by {\it signing} the shares of the intermediate states.
-However, we are not aware of a maliciously secure \mpc backend that works with combinations of
+However, we are not aware of a maliciously secure \mpc implementation for combinations of
 arithmetic and boolean circuits.

### 2017-10-31 OaklandSubmission/ex.tex
final edits

@@
 (\ls{role == SERVER} and \ls{role == CLIENT}), and common code for the
-computation
+computation.
 
@@
 components translate into API calls into our \mpc back-end
-(ABY). E.g., in Figure~\ref{fig:ex-sml}, the \tool
+(ABY). For example, in Figure~\ref{fig:ex-sml}, the \tool
 compiler realizes that the array index \ls{i} in the inner product loop

### 2017-10-31 OaklandSubmission/intro.tex
final edits

@@
 
- Today it is hard for the developers to program secure applications
+ Today it is hard for developers to program secure applications
  using cryptographic techniques. Typical developers lack a deep
@@
 frameworks use cryptographic backends that take as input
-the computation expressed as either a boolean
-circuit~(\cite{yao,gmw}) or an arithmetic circuit~(\cite{sss,viff,gentry}). 
+the computation expressed either as a boolean
+circuit~(\cite{yao,gmw}) or as an arithmetic circuit~(\cite{sss,viff,gentry}). 
 %
@@
 writing large computations in such a framework can be tedious and
-error-prone (a sentiment echoed by Demmler et al.~\cite{aby} themselves).
+subject to errors (a sentiment echoed by Demmler et al.~\cite{aby} themselves).
 

### 2017-10-31 OaklandSubmission/main.tex
minor
smcl
appendix commented
authors
final edits

@@
 % affiliations
-% \author{\IEEEauthorblockN{Michael Shell}
-% \IEEEauthorblockA{School of Electrical and\\Computer Engineering\\
-% Georgia Institute of Technology\\
-% Atlanta, Georgia 30332--0250\\
-% Email: http://www.michaelshell.org/contact.html}
-% \and
-% \IEEEauthorblockN{Homer Simpson}
-% \IEEEauthorblockA{Twentieth Century Fox\\
-% Springfield, USA\\
-% Email: homer@thesimpsons.com}
-% \and
-% \IEEEauthorblockN{James Kirk\\ and Montgomery Scott}
-% \IEEEauthorblockA{Starfleet Academy\\
-% San Francisco, California 96678-2391\\
-% Telephone: (800) 555--1212\\
-% Fax: (888) 555--1212}}
+% \author{\IEEEauthorblockN{Nishanth Chandran \hspace{10pt} Divya Gupta  \hspace{10pt} Aseem Rastogi \hspace{10pt}  Rahul Sharma \hspace{10pt} Shardul Tripathi }
+% \IEEEauthorblockA{Microsoft Research India}
+%}
 
@@
 secure two-party computation (\mpc) framework that generates efficient
-\mpc protocols from high-level, easy-to-write programs.
+\mpc protocols from high-level, easy-to-write, programs.
 \tool provides formal correctness and security guarantees while maintaining performance and scalability.
@@
 %Furthermore,
-Previous language frameworks such as Wysteria, ObliVM, CBMC-GC, etc.
-generate protocols that use either arithmetic or boolean circuits exclusively.
+Previous language frameworks such as CBMC-GC, ObliVM, SMCL, and Wysteria, generate protocols that use either arithmetic or boolean circuits exclusively.
 Our compiler is the first to generate protocols that combine both arithmetic sharing and garbled circuits  for better performance.
@@
 
-\appendix
-\input{app-stash}
+%\appendix
+%\input{app-stash}
 

### 2017-10-31 OaklandSubmission/related.tex
color
final edits

@@
 They conjecture that the problem is NP hard and gave a linear programming based solution as well as a quadratic time greedy heuristic.
-These techniques are not directly applicable to \tool programs because of for-loops and if-conditions.
+These techniques are not directly applicable to \tool programs because of $\mathtt{for}$-loops and $\mathtt{if}$-conditions.
 However, we are exploring whether these ideas can be extended to yield a better type inference.
@@
 %The reason for doing this was to allow for a mix between arithmetic and boolean computations as a secure computation protocol for a pure boolean circuit would have been very inefficient. 
-This process requires deep knowledge of cryptography
-to ensure that the protocols are secure.
+This requires deep knowledge of cryptography
+to ensure the security of the protocols.
 Examples include privacy-preserving classification of medical ElectroCardioGram (ECG)~\cite{barni}, iris and fingerprint recognition~\cite{blanton}, remote diagnostics~\cite{brickell}, private sequence analysis~\cite{franz}, private biometric identification~\cite{huang}, privacy-preserving matrix factorization and ridge-regression~\cite{valeriaMatrix, valeriaRidge}, machine-learning classification~\cite{shafindss}, decision trees and forests~\cite{wu}, neural network training~\cite{secureml} and prediction~\cite{minionn}.

### 2017-10-31 OaklandSubmission/security.tex
final edits

@@
 By \theoremref{correctness}, these outputs $O_1$ are identical to outputs (or observations) $O_2$ of $\crct$ under circuit semantics. 
-Next, $\simu$ runs $\simu_{\tiny{2pc}}$ on $\crct$, input of the corrupt party and $O_2$. 
+Next, $\simu$ runs $\simu_{\tiny{2pc}}$ on $\crct$, input of the corrupt party and $O_2$ (note, while $\simu_{\tiny{2pc}}$ expects output $O_1$ as its input, it is ok for $\simu$ to run $\simu_{\tiny{2pc}}$ on $O_2$ instead as $O_2=O_1$).
 From the security of $\prot$, we have that the simulated view output by $\simu_{\tiny{2pc}}$ is indistinguishable from the real view of the $\prot$ for $\crct$ between the two parties.

### 2017-10-31 OaklandSubmission/eval.tex
minor reorg
final edits
why this no end

@@
 %
-A more interesting classifier is {\em Na\"{i}ve Bayes} that predicts labels
-from the set $[n]$~\cite{shafindss}.
+A more interesting classifier is {\em Na\"{i}ve Bayes}~\cite{shafindss} that predicts labels
+from the set $[n]$.
 Here, the input data point is a {\it feature}
@@
 %The additions in the expression above sum $n$-dimensional vectors.
-%The model size of this classifier is $|P|+|T|$.
-This classifier requires $\Theta(ndF)$ comparisons and additions.
-%Note here  that computing $T(j)(x_j)(i)$ requires a secret look-up in matrix  $T$  based on secret input $x_j$ resulting in $F-1$ comparisons using multiplexers (see
+The model size of this classifier is $\Theta(ndF)$.
+% comparisons and additions. %Note here  that computing $T(j)(x_j)(i)$ requires a secret look-up in matrix  $T$  based on secret input $x_j$ resulting in $F-1$ comparisons using multiplexers (see
   %Section~\ref{sec:impl}).
@@
 %
-A {\em decision tree} of size $N$ takes as input an $x\in\R^d$
+A {\em decision tree} of size $N$ and depth $d$ takes as input an $x\in\R^d$
 %,
@@
 
-\setlength\tabcolsep{2.5pt}
+\setlength\tabcolsep{2.8pt}
 \begin{table}
@@
 \hline
-Dataset & $d$ & $F$ & \thead{Prev. \\ time\\ (s)} & \thead{Prev. \\ Comm\\ (kb)} & \thead{LAN  \\ (s)} & \thead{WAN\\ (s)} & \thead{Comm. \\(kb)} & \#Gates  & LOC\\
+Dataset & $n$ & $F$ & \thead{Prev. \\ time\\ (s)} & \thead{Prev. \\ Comm\\ (kb)} & \thead{LAN  \\ (s)} & \thead{WAN\\ (s)} & \thead{Comm. \\(kb)} & \#Gates  & LOC\\
 \hline
@@
 %\end{table*}
+\setlength\tabcolsep{3.8pt}
 
 \begin{table}[t]
-\begin{tabular}{|c|c|c|c|c|c|c|c|c|}
+\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|}
 \hline
-Dataset  & $N$ & \thead{Prev. \\ time\\ (s)} & \thead{Prev. \\ Comm\\ (kb)} & \thead{LAN  \\ (s)} & \thead{WAN\\ (s)} & \thead{Comm. \\(kb)}& \#Gates & LOC\\
+Dataset  & $d$ & $N$ & \thead{Prev. \\ time\\ (s)} & \thead{Prev. \\ Comm\\ (kb)} & \thead{LAN  \\ (s)} & \thead{WAN\\ (s)} & \thead{Comm. \\(kb)}& \#Gates & LOC\\
 \hline
-Nursery & 4 & 0.3 & 102 & 0.1 & 0.3 & 32 & 3324 & 20\\
+Nursery & 4 & 4 & 0.3 & 102 & 0.1 & 0.3 & 32 & 3324 & 20\\
 \hline
-ECG &  6 & 0.4 & 102 & 0.1 & 0.4 & 49  & 5002 & 20\\
+ECG & 4 &  6 & 0.4 & 102 & 0.1 & 0.4 & 49  & 5002 & 20\\
 \hline
@@
 The results for Na\"{i}ve Bayes are
-in Table~\ref{tab:nb}. As before, $d$ denotes the length of the input
-feature vector $x$ and $F$ is the number of possible values of $x_i$
-for each $i \in [d]$.
+in Table~\ref{tab:nb}. As before, $n$ denotes the number of classes and $F$ is the number of features.
 %Although, there are no multiplications in these benchmarks, these do
@@
 has significantly higher number of boolean-and gates.
-Recall, that square activation can be implemented entirely using
+Note that square activation can be implemented entirely using
 arithmetic gates but ReLU requires boolean-and gates.  For a complete
@@
 Here, $x$ is a 784 length vector obtained from the input image,
-each $W$ is a $10\times 784$ matrix, and each $b$ is a $10$ length
+ $W$ is a $10\times 784$ matrix, and  $b$ is a $10$ length
 vector.
 We implement this classifier in \tool and present the results in the
-first row of Table~\ref{tab:tf}. We note that a pure boolean circuit for this simple functionality does not fit in the memory of our machines.
+first row of Table~\ref{tab:tf}. 
+%We note that a pure boolean circuit for this simple functionality does not fit in the memory of our machines and cannot be evaluated using garbled circuits.
 
@@
 
-\setlength\tabcolsep{6pt}
 \subsection{Secure code partitioning}
 \label{sec:pipeeval}
-\noindent\textbf{CIFAR-10.}
 The largest benchmark of \minion~\cite{minionn} is a DNN for CIFAR-10
@@
 
+\setlength\tabcolsep{6pt}
 \begin{table}
@@
 succeed.
-\\\\
-\noindent\textbf{Matrix factorization.}
+
+\subsection{Matrix factorization.}
 \tool is not tied to secure prediction and can express more general computations.
@@
 The second stage involves sorting and gradient computations and is the heaviest stage.
-The third  stage is similar to the first stage. The results are presented in Table~\ref{tab:factor}.
+The third  stage is similar to the first stage.
+The results are presented in Table~\ref{tab:factor}.
+These circuits have a large depth (column ``depth"); the circuits for secure prediction had depth below 100.
 

### 2017-10-31 OaklandSubmission/impl.tex
why this no end

@@
 operators are not supported natively by \tool. 
-However, we have implemented integral division in 20 lines of \tool,
+However, we have implemented integral division in 30 lines of \tool,
 while the floating-point support in ABY is under active

### 2017-10-31 OaklandSubmission/pipe.tex
nit
why this no end

@@
 can be sequentially executed, using the unmodified underlying \mpc
-backend. The following theorem state the correctness and security of
+backend. The following theorem states the correctness and security of
 the scheme.
@@
 If $\prog_1||\prog_2||\ldots||\prog_k$ is a decomposition of a program $\prog$, then there exists a sequence of programs $\progn_1, \progn_2, \ldots, \progn_k$ and protocols $\prot_1, \prot_2, \ldots, \prot_k$ such that for all $i$, $\prot_i$ securely realizes $\progn_i$ and  $\prot
-= \prot_1\seq\prot_2\dots\seq\prot_k$ securely realizes $s$.
+= \prot_1,\prot_2,\dots,\prot_k$ securely realizes $s$.
 \end{theorem}
@@
 ensures that if $\prog$ is well-typed, then for each $1\leq i \leq k$,
-$\progn_i$ is well-typed. Then, by \theoremref{security}, $\prot_i$
+$\progn_i$ is well-typed. Then, by \theoremref{security}, $\prot_i$,
+the \mpc protocol that evaluates the circuit generated by $\progn_i$,
 securely realizes $\progn_i$. That is, for every $1\leq i \leq k-1$,

### 2017-10-31 OaklandSubmission/security.tex
why this no end

--- 2017-10-31 OaklandSubmission/security.tex
@@
 any circuit $\crct$ that is output by our compiler
-(see \figureref{circuits}). In more detail, this means that for any
-well-typed source program $s$, let $\crct$ be the output circuit output
-(as in \theoremref{correctness}). We assume that there
+(see \figureref{circuits}). This means that for any
+well-typed source program $s$, let $\crct$ be the circuit generated
+as in \theoremref{correctness}. We assume that there
 exists a \mpc protocol $\prot$ that securely realizes the
@@
 \begin{theorem}[Security]\label{theorem:security}
-Let $s$ be any functionality or program in our source language with
-outputs (or observations) $O_1$, that outputs a circuit
+Let $s$ be a well typed program in our source language
+that generates a circuit
 $\chi$ (as defined in Theorem \ref{theorem:correctness}). Let protocol
@@
 Next, $\simu$ runs $\simu_{\tiny{2pc}}$ on $\crct$, input of the corrupt party and $O_2$.
-From the security of $\prot$, we have that the simulated view output by $\simu_{\tiny{2pc}}$ is indistinguishable from the real view of the $\prot$ for $\crct$ between the two parties.
-Hence, the security follows.
+% (note, while $\simu_{\tiny{2pc}}$ expects output $O_1$ as its input, it is ok for $\simu$ to run $\simu_{\tiny{2pc}}$ on $O_2$ instead as $O_2=O_1$). 
+From the security of $\prot$, we have that the simulated view output by $\simu_{\tiny{2pc}}$ is indistinguishable from the real view. Hence, the security follows.
 

### 2017-11-01 OaklandSubmission/ex.tex
nits

@@
 
-Figure~\ref{fig:toolchain} shows an overview of the \tool
+\figureref{toolchain} shows an overview of the \tool
 toolchain. We give a brief overview of each of these phases below.
@@
 \subsubsection*{Source language}
-Consider the example $w^Tx >b$ from Section~\ref{sec:intro}, where
+Consider the example $w^Tx >b$ from \sectionref{intro}, where
 $w$ and $b$ constitute the server's input (a classifier) and $x$ is
-the client's input vector. Figure~\ref{fig:ex-sml} shows the \tool code
+the client's input vector. \figureref{ex-sml} shows the \tool code
 for this example. The code first reads the server's (resp. client's) input using
@@
 \tool compiler takes as input a source program and produces a C++
-program as output. Figure~\ref{fig:ex-aby} shows the output code\footnote{This is also how a program written directly in
+program as output. \figureref{ex-aby} shows the output code\footnote{This is also how a program written directly in
   ABY would look like.} for
-the example in Figure~\ref{fig:ex-sml}. The output program
+the example in \figureref{ex-sml}. The output program
 contains party-specific code for inputs and outputs
@@
 components translate into API calls into our \mpc back-end
-(ABY). For example, in Figure~\ref{fig:ex-sml}, the \tool
+(ABY). For example, in \figureref{ex-sml}, the \tool
 compiler realizes that the array index \ls{i} in the inner product loop
@@
 (lines~\ref{line:condyaobegin} to~\ref{line:condyaoend}) (in our usage
-of ABY, boolean is synonymous with Yao (Section~\ref{sec:impl})).
+of ABY, boolean is synonymous with Yao, we elaborate in \sectionref{impl}).
 
@@
 \end{lstlisting}
-\caption{\tool compiler (partial) output for Figure~\ref{fig:ex-sml}}
+\caption{\tool compiler (partial) output for \figureref{ex-sml}}
 \label{fig:ex-aby}
@@
 easier for a developer to program and get right the code in
-Figure~\ref{fig:ex-sml}, rather than the code in
-Figure~\ref{fig:ex-aby}.
+\figureref{ex-sml}, rather than the code in
+\figureref{ex-aby}.
 % (as we mentioned in Section~\ref{sec:intro},
@@
 programmer to easily modify their code, while the compiler takes care
-of efficiency. For example, consider in Figure~\ref{fig:ex-sml} a
+of efficiency. For example, consider in \figureref{ex-sml} a
 change from multiplication to bitwise-or in the \ls{for} loop. It

### 2017-11-01 OaklandSubmission/impl.tex
nits

@@
 ($\models e < n$
-in ({\sc {T-Read}}) and ({\sc{T-Write}})),
-Figure~\ref{fig:compile}). We take the \tool source program and
+in ({\sc {T-Read}}) and ({\sc{T-Write}}),
+\figureref{compile}). We take the \tool source program and
 translate it as an input to the solver. The solver takes less
@@
 
-The compilation rules of Figure~\ref{fig:compile} can introduce
+The compilation rules of \figureref{compile} can introduce
 repeated coercions from arithmetic to

### 2017-11-01 OaklandSubmission/related.tex
nit

@@
 Wysteria~\cite{wysteria} compiles  programs into boolean circuits and uses a GMW-based backend~\cite{choi,gmw}.
- Mitchell {\em et al.}~\cite{lambdaps} allow the user to select between Shamir's secret sharing~\cite{sss} and fully homomorphic encryption~\cite{gentry}.
+ Mitchell et al.~\cite{lambdaps} allow the user to select between Shamir's secret sharing~\cite{sss} and fully homomorphic encryption~\cite{gentry}.
 Unlike \tool, all these tools use either an arithmetic backend or a boolean backend but not a combination of both.

### 2017-11-01 OaklandSubmission/related.tex
nits

@@
 ObliVM~\cite{oblivm} protects access patterns using an oblivious RAM~\cite{oram1,oram2} and also uses garbled circuits for compute.
-In Secure Multiparty Computation Language (SMCL)~\cite{smcl},  Java like programs are compiled into arithmetic circuits that are then evaluated using the VIFF framework~\cite{viff}.
-Wysteria~\cite{wysteria} compiles  programs into boolean circuits and uses a GMW-based backend~\cite{choi,gmw}.
- Mitchell et al.~\cite{lambdaps} allow the user to select between Shamir's secret sharing~\cite{sss} and fully homomorphic encryption~\cite{gentry}.
-Unlike \tool, all these tools use either an arithmetic backend or a boolean backend but not a combination of both.
+In Secure Multiparty Computation Language (SMCL)~\cite{smcl},  Java
+like programs are compiled into arithmetic circuits that are then
+evaluated using the VIFF
+framework~\cite{viff}. Wysteria~\cite{wysteria} enables programmers to
+write $n$-party mixed-mode programs that combine local, per-party
+computations with secure computations. It compiles secure
+computations to
+boolean circuits and uses a GMW-based backend~\cite{choi,gmw}.
+ Mitchell et al.~\cite{lambdaps} allow the user to select between
+ Shamir's secret sharing~\cite{sss} and fully homomorphic
+ encryption~\cite{gentry}.  Unlike \tool, all these tools use either
+ an arithmetic backend or a boolean backend but not a combination of
+ both.
 

### 2017-11-01 OaklandSubmission/main.tex
nits
nit

@@
 %Furthermore,
-Previous language frameworks such as CBMC-GC, ObliVM, SMCL, and Wysteria, generate protocols that use either arithmetic or boolean circuits exclusively.
-Our compiler is the first to generate protocols that combine both arithmetic sharing and garbled circuits  for better performance.
+Previous language frameworks, such as CBMC-GC, ObliVM, SMCL, and
+Wysteria,
+generate protocols that use either arithmetic or boolean
+circuits exclusively. Our compiler is the first to generate protocols
+that combine both arithmetic sharing and garbled circuits  for better
+performance.
 %We give the first compiler that generates protocols which combine both arithmetic and boolean circuits

### 2018-01-18 usenix18/eval.tex
Usenix template

--- 2017-10-31 OaklandSubmission/eval.tex
@@
 
-We use $[N]$ to denote $\{0,1,\dotsc, N-1\}$.
-Further, given a vector $x\in\R^d$, we say $\mathsf{argmax}\ x = i$ if $x_i = \mathsf{max}\ \{x_0,\ldots,x_{d-1}\}$.
-We focus on the machine learning models for {\it classification}.
-A classifier $C$ uses a trained model to {\it predict} a
+We use $[N]$ to denote $\{0,1,\dotsc, N-1\}$. Further, given a vector
+$x\in\R^d$, we say $\mathsf{argmax}\ x = i$ if $x_i =
+\mathsf{max}\ \{x_0,\ldots,x_{d-1}\}$. Finally, if $A$ is
+a matrix (resp. vector) then we write $f(A)$ for the matrix (resp. vector)
+obtained by applying the scalar function $f$ to each entry of $A$
+pointwise.
+
+We focus on the machine
+learning models for {\it classification}. A classifier $C$ uses a
+trained model to {\it predict} a
 label $\ell$ for an input data point $x$. For example, given a
@@
 unit (ReLU) $f(x)=\mathsf{max}(x,0)$.
-If $A$ is a matrix then the output of $f(A)$ is also a matrix that is
-obtained by applying $f$ to each entry of $A$ pointwise.
 Given an input vector $x$, the predicted label of a DNN is
@@
 matrices $W_j$ and $V_j$, and a vector $\theta_{j}$. The internal node
-$j$  evaluates a predicate $\theta_j^TZx > 0$ to decide whether
+$j$  evaluates a predicate $(\theta_j^T \matmul Z \matmul x) > 0$ to decide whether
 to pass $x$ to the left child $2j+1$ or the right child $2j+2$.
@@
 are matrices of appropriate dimensions. The activation function $f$ is given by $f(y) = y$ if
-$-1 < y < 1$ and $\mathrm{sign}(y)$ otherwise ($f$ on vectors and matrices is defined point-wise on every element of the vector/matrix).
+$-1 < y < 1$ and $\mathrm{sign}(y)$ otherwise.
+%% ($f$ on vectors and
+%% matrices is defined point-wise on every element of the
+%% vector/matrix).
 %\bonsai can be
@@
 convolution layers. In \tool, this benchmark requires partitioning and
-we discuss it in Section~\ref{sec:pipeeval}.
+we discuss it in \sectionref{pipeeval}.
 \\\\

### 2018-01-18 usenix18/fig-ex-aby.tex
Usenix template

--- 2017-10-31 OaklandSubmission/ex.tex
@@
-\section{\tool Overview}
-\label{sec:ex}
-%$w^Tx>0$, mnist logistic regression
-%% \divya{add reference to before and later... as mentioned we would run evaluations on secure prediction of different models. Here we describe scenario in detail..}
-
-\begin{figure}
-  %\vspace{-20pt}
-  %\begin{center}
-  \includegraphics[width=0.45\textwidth]{toolchain}
-  %\end{center}
-  %\vspace{-20pt}
-\caption{\tool toolchain}
-\label{fig:toolchain}
-\end{figure}
-
-Figure~\ref{fig:toolchain} shows an overview of the \tool
-toolchain. We give a brief overview of each of these phases below.
-
-\subsubsection*{Source language}
-Consider the example $w^Tx >b$ from Section~\ref{sec:intro}, where
-$w$ and $b$ constitute the server's input (a classifier) and $x$ is
-the client's input vector. Figure~\ref{fig:ex-sml} shows the \tool code
-for this example. The code first reads the server's (resp. client's) input using
- \ls{input1} (resp. \ls{input2}). It then uses a
-\ls{for} loop to compute \ls{acc}, the inner product of \ls{w} and
-\ls{x}. Finally, the code outputs the result of comparing \ls{acc}
-with \ls{b} only to the client using \ls{output2}.
-
-\tool source language is a simple, imperative language that enables
-the programmers to express \mpc computations in terms of their
-``ideal'' functionalities, without dealing with any cryptographic
-details. The language provides multi-dimensional arrays, conditional
-expressions (the ternary $?\::$ operator), \ls{for}~loops,
-\ls{if}~statements, and special syntax for input/output from each party.
-
-\subsubsection*{\tool compiler}
-\tool compiler takes as input a source program and produces a C++
-program as output. Figure~\ref{fig:ex-aby} shows the output code\footnote{This is also how a program written directly in
-  ABY would look like.} for
-the example in Figure~\ref{fig:ex-sml}. The output program
-contains party-specific code for inputs and outputs
-(\ls{role == SERVER} and \ls{role == CLIENT}), and common code for the
-computation
-
-The compiler splits the input
-program into \emph{public} and \emph{secret} components. The public
-components translate into regular C++ code, while the secret
-components translate into API calls into our \mpc back-end
-(ABY). E.g., in Figure~\ref{fig:ex-sml}, the \tool
-compiler realizes that the array index \ls{i} in the inner product loop
-is public, and hence the access locations need not be hidden.
-Therefore, it compiles the \ls{for}~loop into a C++
-\ls{for}~loop that will be executed in-clear
-(line~\ref{line:dotproductloop}).
-
-Within the secret components, the \tool compiler is ``cryptographic
-cost-aware'', and appropriately picks either arithmetic or boolean
-circuit representations for different sub-components. For example,
-the compiler realizes that the inner product computation is more
-efficient in the arithmetic representation, and therefore it builds
-the corresponding circuit using the arithmetic circuit builder
-\ls{acirc} (lines~\ref{line:dotmulgate} and~\ref{line:dotaddgate}). On
-the other hand, since the comparison with \ls{b}, and the conditional
-expression computation are more efficient in the boolean
-representation, the \tool compiler uses the Yao circuit builder
-\ls{ycirc} to build the corresponding circuits
-(lines~\ref{line:condyaobegin} to~\ref{line:condyaoend}) (in our usage
-of ABY, boolean is synonymous with Yao (Section~\ref{sec:impl})).
-
-Using both arithmetic and boolean representations requires conversions between them.
-The \tool compiler also
-instruments these conversion gates accordingly. For example, in
-lines~\ref{line:convacc} and~\ref{line:convb}, the compiler converts
-\ls{a\_acc} and \ls{a\_b} to boolean representation, before they are
-input to the comparison and multiplexer circuits.
-
 \begin{figure}
@@
 $\mathtt{
-\}\;else\;\{\;\;\textbf{{\color{blue}//role == CLIENT}}
+\}\;else\;\{\;\;\text{{\color{red}//role == CLIENT}}
 }$  
@@
 $\mathtt{
-for(uint32\_t\;i\;=\;0;\;i\;<\;30;\;i = i + 1)\;\{\;\;\textbf{\color{blue}//acc = w}^{\textbf{{\color{blue}T}}}\textbf{\color{blue}x}~\label{line:dotproductloop}
+for(uint32\_t\;i\;=\;0;\;i\;<\;30;\;i = i + 1)\;\{\;\;\text{\color{red}//acc = w}^{\text{{\color{red}T}}}\text{\color{red}x}~\label{line:dotproductloop}
 }$
@@
 $\mathtt{
-if(role == CLIENT) \{\;\;\textbf{\color{blue}//only to the client}
+if(role == CLIENT) \{\;\;\text{\color{red}//only to the client}
 }$
@@
 \end{figure}
-
-\subsubsection*{Circuit generation and evaluation} The next step in
-\tool is to
-compile the output C++ code and execute it. Doing so evaluates away the
-public parts of the program, including the array accesses, and
-generates a  circuit comprising of arithmetic and boolean gates,
-with appropriate conversion gates. The circuit is then evaluated using
-a \mpc protocol.
-
-We can now concretely see the advantages of \tool. Unarguably, it is
-easier for a developer to program and get right the code in
-Figure~\ref{fig:ex-sml}, rather than the code in
-Figure~\ref{fig:ex-aby}.
-% (as we mentioned in Section~\ref{sec:intro},
-%other high-level languages are limited in their use of a single,
-%arithmetic or boolean, crypto back-end). 
-\tool also enables the
-programmer to easily modify their code, while the compiler takes care
-of efficiency. For example, consider in Figure~\ref{fig:ex-sml} a
-change from multiplication to bitwise-or in the \ls{for} loop. It
-turns out that once that's the case, it is more efficient to do both
-the addition and bitwise-or using boolean circuits (if the addition is
-still done using arithmetic, the conversion cost starts to take
-over). In \tool, the programmer simply needs to change one operator in
-the source code, and the compiler generates efficient code that uses
-boolean addition. Whereas, if the programmer was writing ABY
-code, she either has to sacrifice performance, or she would have to
-revisit many parts of the circuit and change them.
-
-To summarize, \tool raises the level of abstraction for the
-programmer, and generates efficient \mpc protocols automatically,
-while its metatheory provides strong correctness and security
-guarantees.
-
-%% Consider a  cloud service provider Bob that wants to provide a service to diagnose
-%% whether a patient has breast cancer or not. Bob trains a machine learning classification model
-%% that results in a vector $w$ (say of length 30) and a scalar $b$.
-%% This model is the intellectual property of Bob and he  wants to keep it secret.
-%% Given a patient's medical report, in the form of a vector $x$ (also of length 30),
-%% the classifer predicts that the patient has breast cancer if $w^Tx>b$.
-%% However, for this task Bob needs access to $x$, which is private data of a customer.
-
-%% A potential customer Alice might not want to reveal $x$ to Bob because of privacy concerns.
-%% And Bob does not want to reveal $w$ and $b$ because then Alice can steal the model.
-\divya{Use HIPAA compliance; as it can leak information of patients that was used in training; more compelling argument.
-See Shafi intro.}
-%% SMC can help Alice and Bob compute $w^Tx>b$ securely, such that Alice receives the classifier's
-%% prediction and Bob learns nothing about $x$. Moreover, Alice does not learn anything more about $w$
-%% and $b$ than what is revealed by the prediction. 
-%% \nc{This is a bit repetitive from the intro, isnt it? Should we have another example perhaps?}
-%% To implement this system, Bob can write the code in \tool and this code is shown in Figure~\ref{fig:ex-sml}.
-%% The expression {\tt input1} reads a value from Bob and {\tt input2} reads from Alice.
-%% The language has arrays and simple loops. A loop {\tt for i = [0:N]} repeats its
-%% body {\tt N} times, the loop counter {\tt i} is assigned {\tt 0} in the first iteration,
-%% and is incremented by one after each iteration. Unbounded {\tt while} loops are problematic
-%% from a cryptographic standpoint and our language only permits these simple {\tt for} loops. 
-%% The first loop in Figure~\ref{fig:ex-sml} reads the model $w$ from Bob, the second loop reads
-%% Alice's medical report $x$, and the third loop computes the dot product $w^Tx$.
-%% The last {\tt output} statement sends the result of the comparison $w^Tx>b$ to Alice.
-%% In particular, the ternary ``{\tt ? :}" operator performs a branch and the result is the second (third)
-%% argument if the first argument is true (false).
-
-%% The \tool compiler is fully automatic and compiles the code described in Figure~\ref{fig:ex-sml} to the C++ code that makes calls to the ABY library  in
-%% Figure~\ref{fig:ex-aby}. An alternative to using \tool is to directly write this C++ code.
-%% However, this code is much more complex than the implementation written in \tool.
-%% Therefore, it is difficult for Alice and bob to verify its correctness and security.
-%% Unlike \tool implementations, the ``secret" variables containing private data ({\tt w,x,b,acc}) need to be handled differently compared to ``public" variables such as loop counters ({\tt i}).
-In particular, the former are manipulated by the ABY library and the latter are manipulated like standard  C++ variables.
-%% Additionally, the C++ code branches on  the special variable
-%% {\tt role}  to decide whether the enclosing code is executed by the {\tt SERVER} Bob or by
-%% the {\tt CLIENT} Alice. In Figure~\ref{fig:ex-sml}, the developer does not need to
-%% manipulate {\tt role} explicitly. In addition to simplifying the implementation, \tool prevents developers from writing buggy code such as\\
-%% \verb+if(role == SERVER) {...manipulate x...}+
-
-%% Furthermore, the code in Figure~\ref{fig:ex-aby} is too low level.
-%% In particular, it is first builds a circuit that represents the computation
-%% $w^Tx>b$ and then executes it via a call to {\tt ExecCircuit}.
-%% To build a circuit, the code adds multiplication gates ({\tt PutMULGate})
-%% and addition gates ({\tt PutADDGate}) for the dot product.
-%% Subsequently, it uses a ``greater than" gate ({\tt PutGTGate})
-%% and a multiplexer gate ({\tt PutMUXGate}) for the comparison with $b$.
-%% For efficiency, the operations for dot product need to be performed using 
-%% arithmetic circuits. However, ABY's arithmetic circuits cannot express
-%% comparisons and these need  boolean circuits.
-%% A program that uses both arithmetic circuit and boolean circuits
-%% requires conversion gates that help  interconvert between arithmetic
-%% representations and boolean representations ({\tt PutA2YGate}).
-%% Bigger computations often require multiple conversions and the
-%% developer effort can quickly become significant.
-
-%% Moreover, it is difficult to maintain the code in Figure~\ref{fig:ex-aby}.
-%% For example, if the multiplication is changed to a bitwise-or then,
-%% in the absence of \tool, an efficiency-conscious developer would need to  use boolean circuits everywhere and would be tasked with removing all arithmetic circuits and the interconversion gates (the cost of  conversion overweighs the gains of performing an arithmetic addition instead of a boolean addition).
-%%  With \tool, the developer needs to change only a single character (replace {\tt *} in Figure~\ref{fig:ex-sml} with {\tt |}) to achieve the same effect. 
-%% On this example and other small examples, we have found that the compiler generated C++ code from a \tool implementation is as efficient as manually written C++ code.
-In particular, for $w^Tx>b$, the compiler automatically generates code that uses arithmetic circuits for dot product and inserts appropriate conversion gates.
-Finally, although this paper uses ABY for evaluation, the  compiler can be easily retargeted to generate code for other cryptography backends.
-
-%% The \tool compiler provides strong static guarantees. For example, a well typed
-%% \tool program is guaranteed to execute to termination without errors.
-A \tool program cannot go into non-termination or dereference illegal memory, i.e., no buffer overflows or underflows can happen at run time.
-Production compilers such as {\tt gcc} provide no such guarantees
-%% for the code in Figure~\ref{fig:ex-aby}.
-%% The primary reason that we are able to achieve these guarantees is because
-%% \tool has been designed to be verifiable and is backed by formal semantics. 
-%% Furthermore, the compiler output is guaranteed to be a cryptographically secure
-%% implementation of the functionality declared by the \tool code. 
-%% Moreover, the \tool compiler also generates a C implementation that can be run natively on a single machine for functional testing.

### 2018-01-18 usenix18/intro.tex
Usenix template

--- 2017-10-31 OaklandSubmission/intro.tex
@@
 high-level computation without thinking about the underlying
-cryptographic details. For example, Figure~\ref{fig:ex-sml}
+cryptographic details. For example, \figureref{ex-sml}
 shows an \tool source
@@
 programmers to write \mpc protocols. E.g. we provide the first \mpc
-implementation of the \bonsai~\cite{bonsai} classifier, and it was programmed in the
+implementation of \bonsai~\cite{bonsai}, and it was programmed in the
 high-level \tool source language by a non-cryptographer. Second, the

### 2018-01-18 usenix18/ld.tex
Usenix template

--- 2017-10-31 OaklandSubmission/ld.tex
@@
 e_{2}$ (we support several other
-operators in the implementation, detailed in Section~\ref{sec:impl}),
+operators in the implementation, detailed in \sectionref{impl}),
 conditionals $\cond{e}{e_{1}}{e_{2}}$,
@@
 language as $[\kw{in}_{j}]_{n}$, which can then evaluate using the
-rule ({\sc{E-Arr}}). The notation $\forall i \in [n]$ is read
-  as $\forall i \in \{0\dots n - 1\}$. The remaining rules are
+rule ({\sc{E-Arr}}) (the notation $\forall i \in [n]$ is read
+  as $\forall i \in \{0\dots n - 1\}$). The remaining rules are
 straightforward, and are elided for space reasons.
@@
 judgment $ \models e < n$ to prove that the array index is in
-bounds. Section~\ref{sec:impl} discusses our implementation of this
+bounds. \sectionref{impl} discusses our implementation of this
 check. Rule ({\sc{T-Inp}}) picks a label $m$ for the input. Finally,
@@
 where $O$ are the observations (similar to source semantics). 
-The former judgement evaluates a gate under the environments
+The former judgment evaluates a gate under the environments
 $\widehat{\rho}_j$ and generates shares $b_j$ of the gate's output.
@@
 %n$, for some $n$. On the other hand, if $x:\kw{uint}^{m} \in
-%\Gamma$ then the judgement checks that $\mathcal{D}_m( \widehat{\rho}_{1}(\widetilde{\rho}(x)),\widehat{\rho}_{2}(\widetilde{\rho}(x)))=\rho(x)$.
+%\Gamma$ then the judgment checks that $\mathcal{D}_m( \widehat{\rho}_{1}(\widetilde{\rho}(x)),\widehat{\rho}_{2}(\widetilde{\rho}(x)))=\rho(x)$.
 %%We elide the judgment, providing the full definition
@@
 %%Furthermore, the final environments are also in the compilation
-%%relation. \divya{Should we call the judgement between environs as compilation relation?}
+%%relation. \divya{Should we call the judgment between environs as compilation relation?}
 %Due to space reasons, we provide the proof in \cite{anonymized}.

### 2018-01-18 usenix18/eval.tex
some formatting changed in the eval sec.

@@
 
-\subsubsection*{Standard classifiers}
+\paragraph{Standard classifiers.}
 A {\em binary linear classifier} is one of the simplest classifiers. Here,
@@
 
-\subsubsection*{Deep neural nets}
+\paragraph{Deep neural nets.}
 The next class of classifiers that we benchmark are deep neural nets
@@
 
-\subsubsection*{State-of-the-art classifiers}
+\paragraph{State-of-the-art classifiers.}
 Finally, there are a class of machine learning classifiers that are
@@
 \subsection{Secure prediction}
-\noindent\textbf{Standard classifiers.}
+\paragraph{Standard classifiers.}
 %\label{sec:shallow}
@@
 \begin{table}
+\footnotesize
 \begin{tabular}{|c|c|c|c|c|c|c|c|c|}
 \hline
-Dataset & $d$  & \thead{Prev. \\ time\\ (s)} & \thead{Prev. \\ Comm\\ (kb)} & \thead{LAN  \\ (s)} & \thead{WAN\\ (s)} & \thead{Comm. \\(kb)}  &  \#Gates & LOC\\
+Dataset & $d$  & \thead{Prev \\ time\\ (s)} & \thead{Prev \\ comm\\ (kb)} & \thead{LAN  \\ (s)} & \thead{WAN\\ (s)} & \thead{Comm \\(kb)}  &  \thead{Num \\ gates} & LOC\\
 \hline
@@
  \caption{Linear classification results. We compare \tool (LAN, WAN, Comm)
- with~\cite{shafindss} (Prev. Time, Prev. Comm).}
+ with~\cite{shafindss} (Prev time, Prev comm).}
  \label{tab:lc} 
@@
 \begin{table}
+\footnotesize
 \begin{tabular}{|c|c|c|c|c|c |c|c|c|c|}
 \hline
-Dataset & $n$ & $F$ & \thead{Prev. \\ time\\ (s)} & \thead{Prev. \\ Comm\\ (kb)} & \thead{LAN  \\ (s)} & \thead{WAN\\ (s)} & \thead{Comm. \\(kb)} & \#Gates  & LOC\\
+Dataset & $n$ & $F$ & \thead{Prev \\ time\\ (s)} & \thead{Prev \\ comm\\ (kb)} & \thead{LAN  \\ (s)} & \thead{WAN\\ (s)} & \thead{Comm \\(kb)} & \thead{Num \\ gates}  & LOC\\
 \hline
@@
 \caption{Na\"{i}ve Bayes results. We compare \tool (LAN, WAN, Comm)
- with~\cite{shafindss} (Prev. Time, Prev. Comm).}
+ with~\cite{shafindss} (Prev time, Prev comm).}
  \label{tab:nb} 
@@
 \begin{table}[t]
+\footnotesize
 \begin{tabular}{|c|c|c|c|c|c|c|c|c|c|}
 \hline
-Dataset  & $d$ & $N$ & \thead{Prev. \\ time\\ (s)} & \thead{Prev. \\ Comm\\ (kb)} & \thead{LAN  \\ (s)} & \thead{WAN\\ (s)} & \thead{Comm. \\(kb)}& \#Gates & LOC\\
+Dataset  & $d$ & $N$ & \thead{Prev \\ time\\ (s)} & \thead{Prev \\ comm\\ (kb)} & \thead{LAN  \\ (s)} & \thead{WAN\\ (s)} & \thead{Comm \\(kb)}& \thead{Num \\ gates} & LOC\\
 \hline
@@
 \caption{Decision tree benchmarks.We compare \tool (LAN, WAN, Comm)
- with~\cite{wu} (Prev. Time, Prev. Comm).}
+ with~\cite{wu} (Prev time, Prev comm).}
  \label{tab:dt} 
@@
 \\\\
-\noindent\textbf{Deep neural nets.}
+\paragraph{Deep neural nets.}
 We evaluate \tool on the DNNs described in SecureML~\cite{secureml},
@@
 \begin{table}
+\footnotesize
 \begin{tabular}{|c|c|c|c |c|c|c|c|c|}
 \hline
-DNN  & \thead{Prev. \\ time\\ (s)} & \thead{Prev. \\ Comm\\ (kb)} & \thead{LAN  \\ (s)} & \thead{WAN\\ (s)} & \thead{Comm. \\(kb)} & \#Gates & \thead{Model\\ Size} & LOC\\
+DNN  & \thead{Prev \\ time\\ (s)} & \thead{Prev \\ comm\\ (kb)} & \thead{LAN  \\ (s)} & \thead{WAN\\ (s)} & \thead{Comm \\(kb)} & \thead{Num \\ gates} & \thead{Model\\ size} & LOC\\
 \hline
@@
  \caption{DNN benchmarks. We compare \tool (LAN, WAN, Comm)
- with~\cite{minionn} (Prev. Time, Prev. Comm).}
+ with~\cite{minionn} (Prev time, Prev comm).}
  \label{tab:nn} 
@@
 \\\\
-\noindent\textbf{State-of-the-art classifiers.}
+\paragraph{State-of-the-art classifiers.}
 Tensorflow~\cite{tensorflow} is a standard machine learning toolkit.
@@
 \begin{table}
+\footnotesize
 \begin{tabular}{|c|c|c|c |c|c|c|c|c|c | c}
 \hline
-Classifier       & \thead{LAN  \\ (s)} & \thead{WAN \\(s)} & \thead{Comm. \\(Mb)}  & \#And & \#Mul & \#Gates & \thead{Model \\size} & LOC\\
+Classifier       & \thead{LAN  \\ (s)} & \thead{WAN \\(s)} & \thead{Comm \\(Mb)}  & \thead{Num \\ And} & \thead{Num \\ Mul} & \thead{Num \\ gates} & \thead{Model \\size} & LOC\\
 \hline
@@
 \begin{table}
+\footnotesize
 \begin{tabular}{|c|c|c|c |c|c|c|c|c|c | c|}
 \hline
-Dataset       & \thead{LAN \\(s)} & \thead{WAN \\(s)} & \thead{Comm. \\ (Mb)}  & \#And & \#Mul & \#Gates & depth & LOC\\
+Dataset       & \thead{LAN \\(s)} & \thead{WAN \\(s)} & \thead{Comm \\ (Mb)}  & \thead{Num \\ And} & \thead{Num \\ Mul} & \thead{Num \\ gates} & depth & LOC\\
 \hline
@@
 \begin{table}
+\footnotesize
 \begin{tabular}{|c|c|c|c |c|c|c|}
 \hline
-           &  LAN (s) & WAN (s) & \thead{Comm. \\ (Mb)}  & \#And & \#Mul & \#Gates \\
+           &  LAN (s) & WAN (s) & \thead{Comm \\ (Mb)}  & \thead{Num \\ And} & \thead{Num \\ Mul} & \thead{Num \\ gates} \\
 \hline
@@
 \begin{table}[t]
+\footnotesize
 \begin{tabular}{|c|c|c|c |c|c| c|}
 \hline
-  Stage         &  LAN (s) & WAN (s) & \thead{Comm. \\ (Mb)}  & depth & \#Gates & LOC\\
+  Stage         &  LAN (s) & WAN (s) & \thead{Comm \\ (Mb)}  & depth & \thead{Num \\ gates} & LOC\\
 \hline

### 2018-01-18 usenix18/ex.tex
some formatting changed in the eval sec.

--- 2017-11-01 OaklandSubmission/ex.tex
@@
 
-\subsubsection*{Source language}
+\paragraph{Source language.}
 Consider the example $w^Tx >b$ from \sectionref{intro}, where
@@
 
-\subsubsection*{\tool compiler}
+\paragraph{\tool compiler.}
 \tool compiler takes as input a source program and produces a C++
@@
 
-\subsubsection*{Circuit generation and evaluation} The next step in
+\paragraph{Circuit generation and evaluation.} The next step in
 \tool is to

### 2018-01-18 usenix18/intro.tex
some formatting changed in the eval sec.

@@
 
-\subsubsection*{Ease of programming} \tool source programs are ideal
+\paragraph{Ease of programming.} \tool source programs are ideal
 functionalities that describe ``what'' computation needs to be done,
@@
 
-\subsubsection*{Cryptographic-cost aware compiler} The \tool compiler
+\paragraph{Cryptographic-cost aware compiler.} The \tool compiler
 compiles a source program to a hybrid computation consisting
@@
 
-\subsubsection*{Scalability (secure code partitioning)} \mpc tools
+\paragraph{Scalability (secure code partitioning).} \mpc tools
 often do not scale to large functionalities. The reason is that
@@
 
-\subsubsection*{Formal guarantees} We prove formal correctness and
+\paragraph{Formal guarantees.} We prove formal correctness and
 security theorems for our compiler. The correctness theorem relates the
@@
 
-\subsubsection*{Evaluation} We have implemented \tool
+\paragraph{Evaluation.} We have implemented \tool
 using ABY~\cite{aby} as the
@@
 
-\subsubsection*{Related Work}
+\paragraph{Related Work.}
 Before ABY~\cite{aby}, several works have proposed

### 2018-01-18 usenix18/ld.tex
some formatting changed in the eval sec.

@@
 
-\subsubsection*{Source language} Our source language
+\paragraph{Source language.} Our source language
  is a simple imperative language shown in \figureref{srclang}. Types
@@
 
-\noindent\textbf{Source semantics.} The runtime semantics for the source language is shown in
+\paragraph{Source semantics.} The runtime semantics for the source language is shown in
 \figureref{srcsem}. These semantics show how a ``trusted third party" computes the outputs when given the inputs of both the parties. 
@@
 
-\subsubsection*{Intermediate language} \figureref{interlang} shows the
+\paragraph{Intermediate language.} \figureref{interlang} shows the
 intermediate language of our compiler. The syntax follows that of the source
@@
 
-\subsubsection*{Source to intermediate compilation} We provide the compilation
+\paragraph{Source to intermediate compilation.} We provide the compilation
 rules in \figureref{compile}. We present the rules in a
@@
 
-\subsubsection*{Evaluation to Circuits}
+\paragraph{Evaluation to Circuits.}
 \figureref{circuits} shows the syntax of circuits. A wire id range
@@
 
-\subsubsection*{Circuit semantics} Evaluating a program in the
+\paragraph{Circuit semantics.} Evaluating a program in the
 intermediate language produces a

### 2018-01-18 usenix18/security.tex
some formatting changed in the eval sec.

--- 2017-10-31 OaklandSubmission/security.tex
@@
-\subsubsection*{Security theorem} The protocols we generate provide
+\paragraph{Security theorem.} The protocols we generate provide
 simulation-based security against a semi-honest adversary, in the

### 2018-02-06 usenix18/pipe.tex
some formatting changed in the eval sec.
code partitioning example

--- 2017-10-31 OaklandSubmission/pipe.tex
@@
 can be sequentially executed, using the unmodified underlying \mpc
-backend. The following theorem states the correctness and security of
-the scheme.
+backend.
+
+\paragraph{Example of code partitioning.} We now illustrate code partitioning through an example.
+Consider the functionality in Figure \ref{fig:largecode}.
+This is a functionality that takes as input two vectors $w$ and $v$ from Alice and two vectors $x$ and $y$ from Bob.
+It computes two inner products $w^Tx$ and $v^Ty$, compares the first value with the second and returns a boolean value (which is $1$ if $w^Tx>v^Ty$ and $0$ otherwise) to Bob.
+Now, if we wish to partition this functionality using secure code partitioning, one possible split is as follows into the following three programs\footnote{All arithmetic is over an appropriate ring in the following discussion.}.
+Partition $1$ (Figure \ref{fig:codepartitioning1}) computes $w^Tx$ and ``secret shares'' the output of this computation between Alice and Bob (Alice's share is $r_1$, a random value, and Bob's share is $o_1 = w^Tx+r_1$).
+Next, partition $2$ (Figure \ref{fig:codepartitioning2}) computes $v^Ty$ and once again provides Alice with $r_2$ and Bob with $o_2 = v^Ty+r_2$.
+Finally, partition $3$ (Figure \ref{fig:codepartitioning3}) compares $o_1-r_1$ with $o_2-r_2$ and provides the output to Bob.
+It is easy to see that the size of the programs $1, 2$ and $3$ (and their corresponding circuits output by the \tool compiler) are smaller than the program in Figure \ref{fig:largecode} and its corresponding circuit, and in particular, smaller than the state that must be maintained between the programs.
+
+\begin{figure}
+\begin{lstlisting}[language=C,mathescape=true]
+$\mathtt{
+uint\;w[30] = input1();\;\;uint\;v[30] = input1();
+}$
+$\mathtt{
+uint\;x[30] = input2();\;\;uint\; y[30] = input2();
+}$
+$\mathtt{
+uint\;acc1 = 0;\;\;uint\;acc2 =0;
+}$
+$\mathtt{
+for\;i\;in\;[0:30]\;
+}$
+$\mathtt{\{acc1\;=\;acc1\;+\;(w[i]\;\times\;x[i]);\;
+}$
+$\mathtt{\;\;acc2\;=\;acc2\;+\;(v[i]\;\times\;y[i]);\}
+}$
+$\mathtt{
+output2((acc1 > acc2)\;?\;1\;:\;0)\;\textbf{{\color{blue}//only to party 2}}
+}$
+\end{lstlisting}
+\caption{\tool code for $w^Tx > v^Ty$}
+\label{fig:largecode}
+\end{figure}
+
+\begin{figure}
+\begin{lstlisting}[language=C,mathescape=true]
+$\mathtt{
+uint\;w[30] = input1();\;\;uint\;r1 = input1();
+}$
+$\mathtt{
+uint\;x[30] = input2();
+}$
+$\mathtt{
+uint\;acc1 = 0;
+}$
+$\mathtt{
+for\;i\;in\;[0:30]\;\{acc1\;=\;acc1\;+\;(w[i]\;\times\;x[i]);\}
+}$
+$\mathtt{
+uint\;o1 = acc1+r1;
+}$
+$\mathtt{
+output2(o1)\;\textbf{{\color{blue}//acc1 is ``secret shared''}}
+}$
+\end{lstlisting}
+\caption{Partition 1: Code for $o_1 = w^Tx+r_1$}
+\label{fig:codepartitioning1}
+\end{figure}
+
+\begin{figure}
+\begin{lstlisting}[language=C,mathescape=true]
+$\mathtt{
+uint\;v[30] = input1();\;\;uint\;r2 = input1();
+}$
+$\mathtt{
+uint\;y[30] = input2();
+}$
+$\mathtt{
+uint\;acc2 =0;
+}$
+$\mathtt{
+for\;i\;in\;[0:30]\;\{acc2\;=\;acc2\;+\;(v[i]\;\times\;y[i]);\;\}
+}$
+$\mathtt{
+uint\;o2 =acc2+r2;
+}$
+$\mathtt{
+output2(o2)\;\textbf{{\color{blue}//acc2 is ``secret shared''}}
+}$
+\end{lstlisting}
+\caption{Partition 2: Code for $o_2 = v^Ty+r_2$}
+\label{fig:codepartitioning2}
+\end{figure}
+
+\begin{figure}
+\begin{lstlisting}[language=C,mathescape=true]
+$\mathtt{
+uint\;r1 = input1();\;\;uint\;r2 = input1();
+}$
+$\mathtt{
+uint\;o1 = input2();\;\;uint\; o2 = input2();
+}$
+$\mathtt{
+uint\;acc3 = o1-r1;\;\;uint\;acc4 =o2-r2;
+}$
+$\mathtt{
+output2((acc3 > acc4)\;?\;1\;:\;0)\;\textbf{{\color{blue}//only to party 2}}
+}$
+\end{lstlisting}
+\caption{Partition 3: Code for $(o_1-r_1)>(o_2-r_2)$}
+\label{fig:codepartitioning3}
+\end{figure}
+
+\paragraph{Security.} The following theorem states the correctness and security of
+code partitioning.
 
@@
 ensures that if $\prog$ is well-typed, then for each $1\leq i \leq k$,
-$\progn_i$ is well-typed. Then, by \theoremref{security}, $\prot_i$,
+$\progn_i$ is well-typed. By \theoremref{security}, $\prot_i$,
 the \mpc protocol that evaluates the circuit generated by $\progn_i$,
 securely realizes $\progn_i$. That is, for every $1\leq i \leq k-1$,
-the protocol $\prot_i$ provides observations $r_i$ to Alice and $o_i$
-to Bob. And, protocol $\prot_k$ provides observation $z$ to both Alice
+the $\prot_i$ provides observations $r_i$ to Alice and $o_i$
+to Bob. Protocol $\prot_k$ provides observation $z$ to both Alice
 and Bob. Finally, since $r_i$ and $o_i$ ($1\leq i\leq k-1$) are
@@
 
-\subsubsection*{Managing circuit sizes using partitioning}
+\paragraph{Managing circuit sizes using partitioning.}
 We use partitioning for programs that require large
@@
 statically estimate the resource usage of a \tool program and we leave
-the construction of such an analysis as future work.
+the construction of such an analysis as future work. \nc{Are we trying to say here which part is manual and which part is automatic? This paragraph is a bit unclear to me.}

### 2018-02-06 usenix18/conclude.tex
Usenix template
minor edits

--- 2017-10-31 OaklandSubmission/conclude.tex
@@
 protocols from previous works.
-The \mpc backed we use provides security against semi-honest or passive adversaries.
+The \mpc backend we use provides security against semi-honest or passive adversaries.
 %our compiler is oblivious to the ``type of adversary'' for the underlying \mpc backend.
@@
 partitioning can be modified to provide malicious security by {\it signing} the shares of the intermediate states.
-However, we are not aware of a maliciously secure \mpc implementation for combinations of
-arithmetic and boolean circuits.
-
+However, we are not aware of efficient maliciously secure \mpc implementation for combinations of
+arithmetic and boolean circuits and hence leave a complete treatment of malicious security to future work.
 % Several areas of improvement remain: a) While the backend that we use~\cite{aby} only provides security against a semi-honest adversary, our compiler is oblivious to the ``type of adversary'' for the underlying backend cryptographic protocol.

### 2018-02-06 usenix18/eval.tex
minor edits

@@
 2.40GHz with 28 GBs of RAM.
+Where we compare our execution times with prior protocols, we match our system and network parameters with those of the prior work (as the code for works such as Bost et al.
+\cite{shafindss} and \minion~\cite{minionn} are not publicly available).
 %
@@
 Since most of our benchmarks
-are related to machine learning, we set up some notation (largely
+are related to machine learning, we set up notation (largely
 standard in machine learning) and describe our benchmarks next.

### 2018-02-07 usenix18/appendix_b.tex
appendix a

\newpage

\section{Example of secure code partitioning}

### 2018-02-07 usenix18/appendix_b.tex
moved partitioning and benchmarks to appendix b and c

@@
-\newpage
+\section{Example of secure code partitioning}\label{app:codepartitioning}
+We now illustrate code partitioning through an example.
+Consider the functionality in Figure \ref{fig:largecode}.
+This is a functionality that takes as input two vectors $w$ and $v$ from Alice and two vectors $x$ and $y$ from Bob.
+It computes two inner products $w^Tx$ and $v^Ty$, compares the first value with the second and returns a boolean value (which is $1$ if $w^Tx>v^Ty$ and $0$ otherwise) to Bob.
+Now, if we wish to partition this functionality using secure code partitioning, one possible split is as follows into the following three programs\footnote{All arithmetic is over an appropriate ring in the following discussion.}.
+Partition $1$ (Figure \ref{fig:codepartitioning1}) computes $w^Tx$ and ``secret shares'' the output of this computation between Alice and Bob (Alice's share is $r_1$, a random value, and Bob's share is $o_1 = w^Tx+r_1$).
+Next, partition $2$ (Figure \ref{fig:codepartitioning2}) computes $v^Ty$ and once again provides Alice with $r_2$ and Bob with $o_2 = v^Ty+r_2$.
+Finally, partition $3$ (Figure \ref{fig:codepartitioning3}) compares $o_1-r_1$ with $o_2-r_2$ and provides the output to Bob.
+It is easy to see that the size of the programs $1, 2$ and $3$ (and their corresponding circuits output by the \tool compiler) are smaller than the program in Figure \ref{fig:largecode} and its corresponding circuit, and in particular, smaller than the state that must be maintained between the programs.
 
-\section{Example of secure code partitioning}
+\begin{figure}
+\begin{lstlisting}[language=C,mathescape=true]
+$\mathtt{
+uint\;w[30] = input1();\;\;uint\;v[30] = input1();
+}$
+$\mathtt{
+uint\;x[30] = input2();\;\;uint\; y[30] = input2();
+}$
+$\mathtt{
+uint\;acc1 = 0;\;\;uint\;acc2 =0;
+}$
+$\mathtt{
+for\;i\;in\;[0:30]\;
+}$
+$\mathtt{\{acc1\;=\;acc1\;+\;(w[i]\;\times\;x[i]);\;
+}$
+$\mathtt{\;\;acc2\;=\;acc2\;+\;(v[i]\;\times\;y[i]);\}
+}$
+$\mathtt{
+output2((acc1 > acc2)\;?\;1\;:\;0)\;\textbf{{\color{blue}//only to party 2}}
+}$
+\end{lstlisting}
+\caption{\tool code for $w^Tx > v^Ty$}
+\label{fig:largecode}
+\end{figure}
+
+\begin{figure}
+\begin{lstlisting}[language=C,mathescape=true]
+$\mathtt{
+uint\;w[30] = input1();\;\;uint\;r1 = input1();
+}$
+$\mathtt{
+uint\;x[30] = input2();
+}$
+$\mathtt{
+uint\;acc1 = 0;
+}$
+$\mathtt{
+for\;i\;in\;[0:30]\;\{acc1\;=\;acc1\;+\;(w[i]\;\times\;x[i]);\}
+}$
+$\mathtt{
+uint\;o1 = acc1+r1;
+}$
+$\mathtt{
+output2(o1)\;\textbf{{\color{blue}//acc1 is ``secret shared''}}
+}$
+\end{lstlisting}
+\caption{Partition 1: Code for $o_1 = w^Tx+r_1$}
+\label{fig:codepartitioning1}
+\end{figure}
+
+\begin{figure}
+\begin{lstlisting}[language=C,mathescape=true]
+$\mathtt{
+uint\;v[30] = input1();\;\;uint\;r2 = input1();
+}$
+$\mathtt{
+uint\;y[30] = input2();
+}$
+$\mathtt{
+uint\;acc2 =0;
+}$
+$\mathtt{
+for\;i\;in\;[0:30]\;\{acc2\;=\;acc2\;+\;(v[i]\;\times\;y[i]);\;\}
+}$
+$\mathtt{
+uint\;o2 =acc2+r2;
+}$
+$\mathtt{
+output2(o2)\;\textbf{{\color{blue}//acc2 is ``secret shared''}}
+}$
+\end{lstlisting}
+\caption{Partition 2: Code for $o_2 = v^Ty+r_2$}
+\label{fig:codepartitioning2}
+\end{figure}
+
+\begin{figure}
+\begin{lstlisting}[language=C,mathescape=true]
+$\mathtt{
+uint\;r1 = input1();\;\;uint\;r2 = input1();
+}$
+$\mathtt{
+uint\;o1 = input2();\;\;uint\; o2 = input2();
+}$
+$\mathtt{
+uint\;acc3 = o1-r1;\;\;uint\;acc4 =o2-r2;
+}$
+$\mathtt{
+output2((acc3 > acc4)\;?\;1\;:\;0)\;\textbf{{\color{blue}//only to party 2}}
+}$
+\end{lstlisting}
+\caption{Partition 3: Code for $(o_1-r_1)>(o_2-r_2)$}
+\label{fig:codepartitioning3}
+\end{figure}

### 2018-02-07 usenix18/appendix_c.tex
moved partitioning and benchmarks to appendix b and c

--- 2018-01-18 usenix18/eval.tex
@@
-\section{Evaluation}
-\label{sec:eval}
-We evaluate \tool on a variety of problems that can fall under the umbrella of {\it secure prediction}, where one
-party (the server) has a machine learning model, and the other party
-(the client) has an input. The goal is to compute the output of the
-model on client's input, with the guarantee that the server
-learns nothing about the input, and the client learns nothing
-about the model beyond what is revealed from the output.
-
-To begin, we first implement the benchmarks from Bost et al.~\cite{shafindss}
-and \minion~\cite{minionn} (both of which study the same setting),
-and show that the performance of the high-level code written in \tool
-is comparable to their hand-crafted protocols.
-%
-Next, we demonstrate the generality and programmability aspects of
-\tool by implementing state-of-the-art machine learning models from
-Tensorflow~\cite{tensorflow} and \bonsai~\cite{bonsai}. Indeed, we
-provide the first 2PC implementation of \bonsai.
-%
-We
-implement a Deep Neural Network (DNN) for CIFAR-10
-dataset~\cite{cifar} from \minion~\cite{minionn} and matrix factorization~\cite{valeriaMatrix}
-to evaluate partitioning.
-%Finally, we evaluate \tool on the task of matrix factorization~\cite{valeriaMatrix}.
-
-We present the numbers for two network settings, a LAN setting and
-a cross-continent WAN setting. The round trip time between the server
-and the client machines in the two settings is 1ms and 40ms respectively. Each
-machine has an Intel(R) Xeon(R) CPU E5-2673 v3 processor running at
-2.40GHz with 28 GBs of RAM.
-%
-%% The  \tool compiler is written in Python and
-%% compiles each of our benchmarks in under a second to C++ code that makes calls to the ABY
-%% library~\cite{aby}. We use an off-the-shelf solver
-%% (SeaHorn~\cite{seahorn}) to check the verification conditions required
-%% to prove that the array indices remain in bound. The solver takes less
-%% than a minute on our largest benchmark.
-Since most of our benchmarks
-are related to machine learning, we set up some notation (largely
-standard in machine learning) and describe our benchmarks next.
-
-\subsection{Benchmarks description}
-%\divya{notation can be moved somewhere... } \\
-
+\section{Description of benchmarks}\label{app:benchmarks}
 We use $[N]$ to denote $\{0,1,\dotsc, N-1\}$. Further, given a vector
@@
 ported classifiers are then implemented in \tool.
-
-\subsection{Secure prediction}
-\paragraph{Standard classifiers.}
-%\label{sec:shallow}
-We evaluate the three standard classifiers, linear, Na\"{i}ve Bayes, and
-decision trees, from~\cite{shafindss} on the following data sets from the UCI
-machine learning repository~\cite{uci}:
- the Wisconsin Breast Cancer data set, 
-Credit Approval data set, Audiology (Standardized) data set, Nursery
-data set, and ECG (electrocardiogram) classification data
-from~\cite{barni}.
-
-The results for linear
-classification are in Table~\ref{tab:lc}.
-The input and the model are both vectors of length $d$. 
-The columns ``Prev. time'' and ``Prev. comm'' show the time and the total
-network communication reported by Bost et al.~\cite{shafindss} for a
-network setting with 40ms
-round trip time, which is same as our WAN setting. The total execution
-time
-of \tool generated code in the LAN and the WAN setting is reported
-next, followed by the total communication.
-We observe that the \tool code performance matches the hand-crafted protocol of Bost
-et al., and the programmer effort in \tool is just 20
-lines (last column in the table) of high-level code in the \tool
-source language. %\aseem{They write circuits,
- % right?} \nc{Actually, they have a specific protocols for dot product and comparison and also say that their developer effort is minimal as they have to write two function calls to these protocols :).}
-% Our execution time is dominated by the multiplications and
-%boolean-ands.
-%The number of boolean-and gates (\#And) remains constant for both the
-%benchmarks as they reflect the number of gates required for a
-%single  comparison, while the number of
-%multiplication gates (\#Mul) is the same as the model size $d$.
-%% The
-%% last column shows the lines of code of \tool program, which is also
-%% independent of the model size.
-
-%\setlength\tabcolsep{7pt}
-%\begin{table*}
-%\begin{tabular}{c|c|c|c|c |c|c|c|c|c|c}
-%Dataset & $d$  & Prev. time (s) & Prev. Comm (kb) & LAN (s) & WAN (s) & Comm. (kb)  & \#And & \#Mul & \#Gates & LOC\\
-%\hline
-%Breast cancer & 30 & 0.3 & 36 & 0.1 & 0.3 & 25 & 95 & 30 & 727 & 20\\
-%\hline
-%Credit & 47 & 0.3 & 41 & 0.1 & 0.3 & 36 & 95 & 47 & 795 & 20\\
-%\hline
-%\end{tabular}
-
-\setlength\tabcolsep{2.8pt}
-\begin{table}
-\footnotesize
-\begin{tabular}{|c|c|c|c|c|c|c|c|c|}
-\hline
-Dataset & $d$  & \thead{Prev \\ time\\ (s)} & \thead{Prev \\ comm\\ (kb)} & \thead{LAN  \\ (s)} & \thead{WAN\\ (s)} & \thead{Comm \\(kb)}  &  \thead{Num \\ gates} & LOC\\
-\hline
-Breast cancer & 30 & 0.3 & 36 & 0.1 & 0.3 & 25  & 727 & 20\\
-\hline
-Credit & 47 & 0.3 & 41 & 0.1 & 0.3 & 36  & 795 & 20\\
-\hline
-\end{tabular}
- \caption{Linear classification results. We compare \tool (LAN, WAN, Comm)
- with~\cite{shafindss} (Prev time, Prev comm).}
- \label{tab:lc} 
-\end{table}
-
-%\subsubsection*{Na\"{i}ve Bayes}
-
-\setlength\tabcolsep{2.5pt}
-\begin{table}
-\footnotesize
-\begin{tabular}{|c|c|c|c|c|c |c|c|c|c|}
-\hline
-Dataset & $n$ & $F$ & \thead{Prev \\ time\\ (s)} & \thead{Prev \\ comm\\ (kb)} & \thead{LAN  \\ (s)} & \thead{WAN\\ (s)} & \thead{Comm \\(kb)} & \thead{Num \\ gates}  & LOC\\
-\hline
-Nursery & 5 & 9 & 1.5 & 0.2 & 0.1 & 0.4 & 0.6  & 73k & 50\\
-\hline
-Audiology & 24 & 70 & 3.9 & 2.0 & 1.5 & 2.9 & 37  & 4219k & 50\\
-\hline
-\end{tabular}
-\caption{Na\"{i}ve Bayes results. We compare \tool (LAN, WAN, Comm)
- with~\cite{shafindss} (Prev time, Prev comm).}
- \label{tab:nb} 
-\end{table}
-
-%\subsubsection*{Decision trees}
-
-%\begin{table*}
-%\begin{tabular}{|c|c|c|c |c|c|c|c|c|c|c|}
-%\hline
-%Dataset  & $N$ & \thead{Prev. \\ time\\ (s)} & \thead{Prev. \\ Comm\\ (kb)} & \thead{LAN  \\ (s)} & \thead{WAN\\ (s)} & \thead{Comm. \\(kb)}& \#And & \#Mul &\#Gates & LOC\\
-%\hline
-%Nursery & 4 & 0.3 & 102 & 0.1 & 0.3 & 32 & 504 & 3 & 3324 & 20\\
-%\hline
-%ECG &  6 & 0.4 & 102 & 0.1 & 0.4 & 49 & 756 & 5 & 5002 & 20\\
-%\hline
-%\end{tabular}
-% \caption{Decision tree benchmarks. We compare our results (columns 5 , 6, 7) with~\cite{wu} (columns 3 and 4)}
-% \label{tab:dt} 
-%\end{table*}
-\setlength\tabcolsep{3.8pt}
-
-\begin{table}[t]
-\footnotesize
-\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|}
-\hline
-Dataset  & $d$ & $N$ & \thead{Prev \\ time\\ (s)} & \thead{Prev \\ comm\\ (kb)} & \thead{LAN  \\ (s)} & \thead{WAN\\ (s)} & \thead{Comm \\(kb)}& \thead{Num \\ gates} & LOC\\
-\hline
-Nursery & 4 & 4 & 0.3 & 102 & 0.1 & 0.3 & 32 & 3324 & 20\\
-\hline
-ECG & 4 &  6 & 0.4 & 102 & 0.1 & 0.4 & 49  & 5002 & 20\\
-\hline
-\end{tabular}
-\caption{Decision tree benchmarks.We compare \tool (LAN, WAN, Comm)
- with~\cite{wu} (Prev time, Prev comm).}
- \label{tab:dt} 
-\end{table}
-
-%dataset  | size | gates | and | mul | depth | length
-
-%dataset | time | comm | timeL | timeW | comm
-%Our decision tree numbers in Table~\ref{tab:dt} further validate
-%this claim. 
-%We compare with the recent work of Wu et al.~\cite{wu} (columns 3 and 4) who
-%describe specialized protocols for trees and forests (that perform
-%better than~\cite{shafindss}).
-%Similar to Bost et al.~\cite{shafindss}, they also remarks that these
-%computations are infeasible with Yao-based 2PC.
-%Again, we observe that
-%the performance of \tool generated generic 2PC
-%code is competitive with specialized protocols. 
-The results for Na\"{i}ve Bayes are
-in Table~\ref{tab:nb}. As before, $n$ denotes the number of classes and $F$ is the number of features.
-%Although, there are no multiplications in these benchmarks, these do
-%have a significant number of comparisons (due to array look-ups using
-%secret indices) that
-%result in many boolean-and gates. %(13k/750k denote 13000/75000). 
-As before, we compare with Bost et al.~\cite{shafindss} and observe that \tool generated code
-has better performance, despite using a generic 2PC,
-as opposed to custom designed protocols developed by Bost et
-al. Moreover, they remark that in their setup, generic Yao-based 2PC
-did not scale to the smallest of their Na\"{i}ve Bayes classifiers, so
-they had to scale down the prediction task, and even then Yao-based \mpc
-was 500x slower. Whereas, we show that by using a
-cryptographic-cost aware compiler, we can scale generic 2PC to real
-prediction tasks, and get performance competitive to or better than the
-specialized protocols. Table~\ref{tab:dt} 
-compares against the more recent work of~\cite{wu}
-on decision trees and further validates this claim.
-\\\\
-\paragraph{Deep neural nets.}
-We evaluate \tool on the DNNs described in SecureML~\cite{secureml},
-Cryptonets~\cite{cryptonets}, and the CNN from \minion~\cite{minionn}. For
-comparison, we consider their implementations from
-\minion~\cite{minionn}, which outperforms their previous
-implementations. Table~\ref{tab:nn} shows the
-results\footnote{\minion does not report the network round-trip time nor the bit-length of their inputs (we use $32$-bit inputs).}.
-We note that for each of these DNNs, \minion provides a
-specialized protocol, while \tool uses a generic 2PC protocol
-(auto) generated from high-level code.
-
-The first benchmark  is the DNN described in
-SecureML~\cite{secureml} (Figure 10 in~\cite{minionn}).
-It has three fully connected layers with square as the activation function.
-%This DNN is light-weight as it requires only arithmetic computations
-%(additions and multiplications) 
-%except for the $\mathsf{argmax}$ at the end of the network.
-Next, we implement the DNN described in Cryptonets~\cite{cryptonets} (Figure 11 in~\cite{minionn})
-in \tool.
-This DNN also uses square as the activation function and has one
-convolution (with 5 output channels) and one fully connected layer. 
-Finally, we implement CNN from \minion (Figure 12 in~\cite{minionn}), that has
-two convolutions (with 16 output channels each) and two fully
-connected layers.  In contrast to the previous two DNNs, it uses
-ReLU for activation and
-has significantly higher number of boolean-and gates.
-Note that square activation can be implemented entirely using
-arithmetic gates but ReLU requires boolean-and gates.  For a complete
-description of these benchmarks and their accuracies, we refer the
-reader to the original references.
-
-%net | minionT | minionC | timeL | timeW | comm | gates | and | mul | length
-
-%\setlength\tabcolsep{6.5pt}
-%\begin{table*}
-%\begin{tabular}{c|c|c|c |c|c|c|c|c|c|c}
-%DNN  & Prev. time (s) & Prev. Comm (Mb) & LAN (s) & WAN (s) & Comm. (Mb)  & \#And & \#Mul & \#Gates & Model Size & LOC\\
-%\hline
-%SecureML   &  1.1 & 15.8 & 0.7 & 1.7  & 76   &  2k   & 119k & 366k   & 119k & 78\\
-%\hline
-%Cryptonets &  1.3 & 47.6 & 0.6 & 1.6  & 70    & 2k    & 108k & 316k & 86k & 88\\
-%\hline
-%CNN        &  9.4 & 657.5& 5.1 & 11.6 & 501  & 1640k & 667k & 9480k & 35k & 154\\
-%\hline
-%\end{tabular}
-%
-% \caption{DNN benchmarks. We compare our results (columns 4 , 5, 6) with~\cite{minionn} (columns 2 and 3)}
-% \label{tab:nn} 
-%\end{table*}
-
-\setlength\tabcolsep{2.5pt}
-\begin{table}
-\footnotesize
-\begin{tabular}{|c|c|c|c |c|c|c|c|c|}
-\hline
-DNN  & \thead{Prev \\ time\\ (s)} & \thead{Prev \\ comm\\ (kb)} & \thead{LAN  \\ (s)} & \thead{WAN\\ (s)} & \thead{Comm \\(kb)} & \thead{Num \\ gates} & \thead{Model\\ size} & LOC\\
-\hline
-SecureML   &  1.1 & 15.8 & 0.7 & 1.7  & 76   &   366k   & 119k & 78\\
-\hline
-Cryptonets &  1.3 & 47.6 & 0.6 & 1.6  & 70    &  316k & 86k & 88\\
-\hline
-CNN        &  9.4 & 657.5& 5.1 & 11.6 & 501  & 9480k & 35k & 154\\
-\hline
-\end{tabular}
-
- \caption{DNN benchmarks. We compare \tool (LAN, WAN, Comm)
- with~\cite{minionn} (Prev time, Prev comm).}
- \label{tab:nn} 
-\end{table}
-
-In Table~\ref{tab:nn}, 
-the column ``Model size'' is the number of parameters in the
-trained model.
-We observe that our performance is competitive with specialized
-\minion protocols, for both the LAN and the WAN settings. Further, lines
-of \tool source code required is still small. 
-We note that \minion and \tool implementations are both based on ABY.
-These results are surprising as \minion has a much more efficient
-preprocessing phase and has SIMD 
-(single instruction multiple data) capabilities that allow it to perform
-matrix operations efficiently. In contrast, \tool focuses on generic
-2PC, and such optimizations are part of our future work.
-\minion also reports performance results on a bigger DNN with 7
-convolution layers. In \tool, this benchmark requires partitioning and
-we discuss it in \sectionref{pipeeval}.
-\\\\
-\paragraph{State-of-the-art classifiers.}
-Tensorflow~\cite{tensorflow} is a standard machine learning toolkit.
-Its introductory tutorial describes two prediction models for handwritten digit recognition
-using the MNIST dataset~\cite{mnist}.
-Each image in this dataset is a greyscale $28\times 28$ image of
-digits 0 to 9.
-The first model that the tutorial describes is a softmax regression
-that provides an accuracy of 92\%. The classifier evaluates
-$ \mathsf{argmax}\ W\matmul x+b$.
-Here, $x$ is a 784 length vector obtained from the input image,
- $W$ is a $10\times 784$ matrix, and  $b$ is a $10$ length
-vector.
-We implement this classifier in \tool and present the results in the
-first row of Table~\ref{tab:tf}. 
-%We note that a pure boolean circuit for this simple functionality does not fit in the memory of our machines and cannot be evaluated using garbled circuits.
-
-The next classifier in the Tensorflow tutorial is a convolution neural net with two convolutions
-(with 32 output channels) and two fully connected layers with ReLU as the activation function.
-This DNN is both bigger and more accurate than the DNNs presented in the previous section.
-In particular, it has an accuracy of 99.2\%.
-Since, we are not aware of any other tools that have used this
-model as a benchmark, we only report numbers for \tool.
-We observe that this DNN can take a minute per prediction in the WAN
-setting and is the largest benchmark that we have evaluated without
-partitioning.
-
-% benchmark | size | gates | and | mul | depth  | length | comm | timeL | timeW
-\setlength\tabcolsep{2.5pt}
-\begin{table}
-\footnotesize
-\begin{tabular}{|c|c|c|c |c|c|c|c|c|c | c}
-\hline
-Classifier       & \thead{LAN  \\ (s)} & \thead{WAN \\(s)} & \thead{Comm \\(Mb)}  & \thead{Num \\ And} & \thead{Num \\ Mul} & \thead{Num \\ gates} & \thead{Model \\size} & LOC\\
-\hline
-Regression &  0.1         & 0.7         & 5            & 2k    & 8k    &  35k    & 8k   & 38\\
-\hline
-CNN        &  30.5        & 60.3        & 2955         & 6082k & 4163k &  42104k & 3226k& 172\\
-\hline
-\end{tabular}
-
- \caption{Tensorflow tutorial benchmarks}
- \label{tab:tf} 
-\end{table}
-\setlength\tabcolsep{3.5pt}
-\begin{table}
-\footnotesize
-\begin{tabular}{|c|c|c|c |c|c|c|c|c|c | c|}
-\hline
-Dataset       & \thead{LAN \\(s)} & \thead{WAN \\(s)} & \thead{Comm \\ (Mb)}  & \thead{Num \\ And} & \thead{Num \\ Mul} & \thead{Num \\ gates} & depth & LOC\\
-\hline
-Chars4k    &  0.1         & 0.7         & 2            & 18k    & 3k    &  85k     & 1   & 89\\
-\hline
-USPS       &  0.2         & 0.9         & 4            & 62k    & 2k    &  285k    & 2   & 156\\
-\hline
-WARD       &  0.3         & 1.1         & 9            & 106k    & 8k    &  506k    & 3   & 283\\
-\hline
-\end{tabular}
-
- \caption{Bonsai benchmarks}
- \label{tab:bonsai} 
-\end{table}
-
-We next present \bonsai~\cite{bonsai} results on three
-standard datasets: character recognition (Chars4k~\cite{campos},
-accuracy 74.71\%), text recognition (USPS~\cite{hull}, accuracy
-94.4\%), and object categorization (WARD~\cite{yang}, accuracy
-95.7\%). 
-The \bonsai models are already over integers and no porting from
-floating-point is required.
-We implement the trained classifiers in \tool for all the benchmarks
-from~\cite{bonsai},
-and show the representative results in Table~\ref{tab:bonsai}.
-Out of all the benchmarks from~\cite{bonsai}, the dataset WARD
-requires the largest model. 
-The column ``depth''  shows the depth of the
-tree used by \bonsai. The size of \tool 
-program grows with the depth of the tree, as the straightforward \tool
-implementation requires a loop for each layer of the
-tree\footnote{We remark here that our current language does not support functions (which we leave for future work) and with this support, LOC would be lower and independent of depth.}.
-
-To summarize, by providing first 2PC implementations of
-state-of-the-art classifiers, we have demonstrated the expressiveness
-of \tool. We discuss scalability next.
-%dataset | size | gates | and | mul | depth | length | comm | timeL | timeW
-
-\subsection{Secure code partitioning}
-\label{sec:pipeeval}
-The largest benchmark of \minion~\cite{minionn} is a DNN for CIFAR-10
-dataset~\cite{cifar}. 
-The classifier's task is to categorize colored ($32\times 32$) images
-into 10 classes. A secure evaluation of this DNN needs more memory
-than what is available
-on our machines. Therefore, we use partitioning and divide the
-computation into seven stages. 
-The first step does a convolution with 64 output channels and a ReLU
-activation. The next four stages together perform a convolution that
-involves multiplying a $64\times 576$ matrix with a 
-$576\times 1024$ matrix. The sixth stage performs a ReLU and a
-convolution. The final stage has four convolutions, five ReLUs, and a
-fully connected layer. The total number of lines of \tool code for
-this benchmark is 336 lines.
-
-\setlength\tabcolsep{6pt}
-\begin{table}
-\footnotesize
-\begin{tabular}{|c|c|c|c |c|c|c|}
-\hline
-           &  LAN (s) & WAN (s) & \thead{Comm \\ (Mb)}  & \thead{Num \\ And} & \thead{Num \\ Mul} & \thead{Num \\ gates} \\
-\hline
-Total      &  265.6       & 647.5        & 40683       & 21m    & 61m    &  337m  \\
-\hline
-Stage 6    &  55.2        & 122.6        & 6744        & 12m    & 10m   &  98m  \\
-\hline
-\end{tabular}
-
- \caption{Partitioning results for CIFAR-10. \minion takes 544 seconds and communicates 9272 Mb.}
- \label{tab:cifar} 
-\end{table}
-
-\begin{table}[t]
-\footnotesize
-\begin{tabular}{|c|c|c|c |c|c| c|}
-\hline
-  Stage         &  LAN (s) & WAN (s) & \thead{Comm \\ (Mb)}  & depth & \thead{Num \\ gates} & LOC\\
-\hline
-1    &  175       & 662        & 29816       & 16370    & 33m    & 500  \\
-\hline
-2    &  193        & 1095        & 31945        & 30916    & 37m & 516 \\
-\hline
-3    &  178        & 627        & 29810        & 16369    & 32m  & 478  \\
-\hline
-Total    &  546      & 2384        & 91571        & --    & 102m & 1494 \\
-\hline
-\end{tabular}
-
- \caption{Partitioning results for matrix factorization. The time reported by~\cite{valeriaMatrix} for this computation is 10440 seconds.}
- \label{tab:factor} 
-\end{table}
-Table~\ref{tab:cifar} shows the end-to-end numbers as
-well as the numbers for the sixth stage, which is the heaviest. 
-The number of gates are in millions, hence the suffix `m' in the last
-three columns.
-%For this DNN, the specialized \minion protocol takes 544 seconds and
-%communicates 9272 Mb.
-As with Table~\ref{tab:nn}, \tool generated generic 2PC protocol is
-competitive with \minion here as well. Therefore,
-we believe that with partitioning, \tool can scale to arbitrary sized
-computations while maintaining performance
-competitive with existing specialized protocols. In particular, for a
-large enough DNN, \minion could run
-out of memory but an appropriately partitioned \tool implementation would still
-succeed.
-
-\subsection{Matrix factorization.}
-\tool is not tied to secure prediction and can express more general computations.
-To demonstrate this expressiveness, we  implement secure matrix factorization~\cite{valeriaMatrix}. Abstractly, given a sparse matrix $\mathcal{M}$ of dimensions
-$n\times m$ and $M$ non-zero entries, the goal is to generate a matrix $U$ of dimension $n\times d$ and a matrix
-$V$ of dimension $d\times m$ such that $\mathcal{M}\approx UV$. This operator is useful in recommender systems.
-In particular, Nikolaenko et al.~\cite{valeriaMatrix} shows how to implement a movie recommender system which does not require users to reveal their data in the clear, i.e., the ratings the users have assigned to movies are kept secret.
-The implementation is a two party computation of an iterative algorithm for matrix factorization (Algorithm~1 in~\cite{valeriaMatrix}).
-This algorithm is based on gradient descent and iteratively converges to a local minima.
-We implement this algorithm in \tool.
-
-To ensure that the algorithm converges to the right local minima, Nikolaenko et al. require
-36 bits of precision. Since ABY supports either 32-bit or 64-bit integers, our \tool implementation
-manipulates 64-bit variables.
-For the matrix $\mathcal{M}$ of  user data, Nikolaenko et al.
-consider $n=940$ users, $m=40$ most popular movies, and $M=14683$ ratings from the MovieLens dataset.
-The time reported in~\cite{valeriaMatrix}
-for one iteration is 2.9 hours\footnote{~\cite{valeriaMatrix} does not report the network round-trip time.}. This computation is large enough that we partition each iteration
-into three stages. The first stage involves a Batcher~\cite{Batcher} sorting network followed by a linear pass.
-The second stage involves sorting and gradient computations and is the heaviest stage.
-The third  stage is similar to the first stage.
-The results are presented in Table~\ref{tab:factor}.
-These circuits have a large depth (column ``depth"); the circuits for secure prediction had depth below 100.
-
-We observe that in the LAN setting, we are about 19 times faster than~\cite{valeriaMatrix} and in the WAN
-setting we are about 4 times faster. The main source of these significant speedups is that, unlike~\cite{valeriaMatrix}, \tool does not need to convert the functionality into boolean circuits. 
-However, this benchmark requires more lines of code than the previous benchmarks
-because of  Batcher's sort (450 lines of \tool code in each stage).
-However, the current programmer effort seems miniscule compared to the mammoth implementation effort
-put in by Nikolaenko et al. (Section 5 of ~\cite{valeriaMatrix}) to scale a boolean circuits based
-backend to this benchmark.
-
-%stage | time | comm | gates | mul | add | depth

### 2018-02-07 usenix18/eval.tex
moved partitioning and benchmarks to appendix b and c

@@
 %% than a minute on our largest benchmark.
-Since most of our benchmarks
-are related to machine learning, we set up notation (largely
-standard in machine learning) and describe our benchmarks next.
+Most of our benchmarks
+are related to machine learning and we set up (largely standard) notation and describe our benchmarks in \appendixref{benchmarks}.
 
-\subsection{Benchmarks description}
+%\subsection{Benchmarks description}
 %\divya{notation can be moved somewhere... } \\
-
-We use $[N]$ to denote $\{0,1,\dotsc, N-1\}$. Further, given a vector
-$x\in\R^d$, we say $\mathsf{argmax}\ x = i$ if $x_i =
-\mathsf{max}\ \{x_0,\ldots,x_{d-1}\}$. Finally, if $A$ is
-a matrix (resp. vector) then we write $f(A)$ for the matrix (resp. vector)
-obtained by applying the scalar function $f$ to each entry of $A$
-pointwise.
-
-We focus on the machine
-learning models for {\it classification}. A classifier $C$ uses a
-trained model to {\it predict} a
-label $\ell$ for an input data point $x$. For example, given a
-data point which is a tuple of humidity and temperature 
-a classifier can predict a label ``will rain'' or ``will not rain''.
-The {\it model size} of a classifier is the number of parameters in the model.
-For example, the model size of the classifier in~\figureref{ex-sml} is $|w|+1=31$.
-The {\it
-  accuracy} of a classifier refers to the fraction of data points that
-the classifier labels correctly from a given set of test data
-points.
-
-\paragraph{Standard classifiers.}
-A {\em binary linear classifier} is one of the simplest classifiers. Here,
-the input is a data point $x\in\R^d$,
-and the model is a vector $w\in\R^d$.
- The possible labels are
-$\ell\in\{\mathit{true},\mathit{false}\}$ and the classifier is
-$C_w\equiv w^Tx>0$.
-%, where $w^T$ denotes the transpose
-%of matrix $w$, $\matmul$ denotes matrix multiplication (with a vector in this case) and $a>b$ returns %$\mathit{true}$ if $a>b$ and $\mathit{false}$ otherwise. 
-%The {\it model size} of this classifier, i.e., the number of parameters in the model is $|w|$, i.e., $d$.
-%This classifier requires $d$ multiplications, $d-1$
-%additions, and a single comparison.
-%
-A more interesting classifier is {\em Na\"{i}ve Bayes}~\cite{shafindss} that predicts labels
-from the set $[n]$.
-Here, the input data point is a {\it feature}
-vector $x=(x_0,x_2,\ldots,x_{d-1})^T$ where each $x_j\in [F]$.
-%The model has two matrices: a vector $P$ of length $n$ s.t. for each
-%$i
-%\in [n]$, $P(i)$ is the log-probability that the output label is $i$.
-%$P(i) \equiv \log p(\ell=i)$, where $p(\ell=i)$ is the likelihood
-%that the output label is $i$ for $i\in[n]$.
-%The other matrix $T$ has size $d\times F\times n$ and for each $j\in
-%[d], k\in [F], i\in [n]$, the entry $T(j)(k)(i)$ is 
-%\equiv\log p(x_j|\ell=i)$, i.e.,
-%the log-probability that  the $j^{th}$ input feature is $k$  conditioned on  the output label being $i$.
-%The classifier $C_{P,T}(x)$ outputs
-%\[
-%{\sf argmax}\ P+\sum_{j=0}^{d-1} T(j)(x_j)
-%\]
-%The additions in the expression above sum $n$-dimensional vectors.
-The model size of this classifier is $\Theta(ndF)$.
-% comparisons and additions. %Note here  that computing $T(j)(x_j)(i)$ requires a secret look-up in matrix  $T$  based on secret input $x_j$ resulting in $F-1$ comparisons using multiplexers (see
-  %Section~\ref{sec:impl}).
-%\nc{Made DG's comment as part of regular text - worth point out}.
-\cmmt{Is there a +1 etc missing or is this exact?} \nc{By our calculation in Sec 5, it seems that this should be $F/2$ multiplexers}
-%
-A {\em decision tree} of size $N$ and depth $d$ takes as input an $x\in\R^d$
-%,
- and 
-%the
-%model consists of a binary tree of depth $d$ with a boolean predicate
-%assigned to each internal node.
-% The root note is the $0^{th}$ node and
-%for an internal node $j \in [N]$, the children nodes are $2j+1$ and
-%$2j+2$. Each internal node $j\in[N]$ at depth $i$ has a predicate
-%$b_{j}^i\equiv x_i\leq w_{j}$. We start evaluating the tree from the
-%root and if the predicate at the current node $b_j^i$ is false (resp.,
-%true) then $x$ is passed to the left (resp., right) child with
-%predicate $b_{2j+1}^{i+1}$ (resp., $b_{2j+2}^{i+1}$). This process is
-%repeated till we reach a leaf. The leaves of the tree are labels and
-%the output label is the leaf visited by this traversal.
-%Such a (binary) decision tree 
-%and can be encoded as a polynomial (of
-%degree linear in depth of the tree) and then
- the prediction task
-reduces to evaluation of a $d$-degree polynomial~\cite{shafindss}.
-% E.g., a binary tree with depth one
-%and size $N=3$ can be encoded as the polynomial
-%$(1-b(x))\ell_0 + b(x)\ell_1$ where $b(x)$ is the predicate at the
-%root, $\ell_0$ is the label of the left leaf and $\ell_1$ is the label
-%of the right leaf.
-
-\paragraph{Deep neural nets.}
-The next class of classifiers that we benchmark are deep neural nets
-or DNNs. A DNN has multiple layers such that each layer computes a
-matrix
-multiplication followed by an {\it activation} function $f$. The most
-common activation functions are square $f(x)=x^2$ and rectifier linear
-unit (ReLU) $f(x)=\mathsf{max}(x,0)$.
-Given an input vector $x$, the predicted label of a DNN is
-\[
- \mathsf{argmax}\ W_N\matmul f_{N-1}(\ldots f_1(W_1\matmul x)\ldots)
-\]
-Here, $f_i$'s are the (public) activation functions, the model
-consists of matrices $W_i$,  $x \in \R^d$ is the input vector, and the
-operator $\matmul$ denotes a matrix multiplication.
-Neural nets usually have one or more fully connected layers, each of
-which multiplies a matrix with a vector.
-Some neural nets have convolution layers and such DNNs are also called
-Convolutional Neural Nets or CNNs.
-For the purpose of this paper, a convolution can be considered as a
-(heavy) matrix-matrix multiplication. The size of matrices manipulated
-by a convolution layer grows linearly with {\it window size}
-(typically 9 or 25), the number of {\it output channels} (typically
-16, 32, or 64), and the size of the matrix input to this layer.
-Therefore, fully connected layers are lighter computation-wise compared
-to convolution layers. However, the model size  of fully connected
-layers is larger than those of convolution layers.
-In general, DNNs are computationally heavy but provide
-much better accuracies on computer vision tasks than the classifiers
-discussed above.
-
-\paragraph{State-of-the-art classifiers.}
-Finally, there are a class of machine learning classifiers that are
-much more efficient than
-DNNs and provide reasonably good accuracies on standard learning
-tasks. \bonsai~\cite{bonsai} is a state-of-the art classifier in this
-class and \tool provides the first 2PC protocol for it.
-\bonsai takes as input $x \in \R^d$, and its model consists of a
-binary tree with $N$ nodes, and a matrix $Z$. Each node $j$ contains
-matrices $W_j$ and $V_j$, and a vector $\theta_{j}$. The internal node
-$j$  evaluates a predicate $(\theta_j^T \matmul Z \matmul x) > 0$ to decide whether
-to pass $x$ to the left child $2j+1$ or the right child $2j+2$.
-The predicted value is
-
-%\aseem{I had a hard time parsing this. (1) Should we use $\cdot$ for
-%  matrix multiplication as we did for DNNs? (2) Precedence of $\circ$
-%  over scalar multiplication (or is it associative so doesn't matter?)
-%(3) The argument to $f$ is a scalar (right?), and so is its output,
-%so then is $\circ$ defined OR is it applying $f$ pointwise to a vector?}
-%\nc{(1) I agree we should do that to be consistent with before.
-(2) I assumed $f()$ is computed to give a vector somehow and $I_j(x)W_j^TZx$ also gives a vector somehow and then we do $\circ$; I might be wrong.
-(3) Hmm..
-I thought it was a vector, but I dont know how.
-(4) Can we specify somewhere the dimensions of these matrices?}
-%\vspace{-10pt}
-\[
-{\sf argmax} \sum_{j = 0}^{N - 1} I_j(x)[ (W_j^T\matmul Z\matmul x) \circ (f(V_j^T\matmul Z\matmul x))] 
-\]
-Here, $I_j(x)$ is 1 if the $j^{th}$ node is present on the path traversed by $x$
-and is zero otherwise. 
-The operation $\circ$ is a pointwise multiplication of two vectors, $W_j$'s and $V_j$'s
-are matrices of appropriate dimensions. The activation function $f$ is given by $f(y) = y$ if
-$-1 < y < 1$ and $\mathrm{sign}(y)$ otherwise.
-%% ($f$ on vectors and
-%% matrices is defined point-wise on every element of the
-%% vector/matrix).
-%\bonsai can be
-%seen as a variant of decision trees where the prediction is a function
-%of the path traversed from root to the leaf and not just the leaf
-%itself.
-
-In the following, we implement these classifiers in \tool and report
-the time taken for making secure predictions. Ideally, the machine
-learning classifiers are mathematical expressions over $\R$ that are
-usually approximated by floating-point operations. 
-As is standard, we port the classifiers to integer manipulating
-programs by scaling the models and rounding~\cite{minionn}. These
-ported classifiers are then implemented in \tool.
 

### 2018-02-07 usenix18/pipe.tex

--- 2017-10-31 OaklandSubmission/pipe.tex
@@
 can be sequentially executed, using the unmodified underlying \mpc
-backend. The following theorem states the correctness and security of
-the scheme.
+backend. An example illustrating the technique of secure code partitioning is provided in  \appendixref{codepartitioning}.
+
+\paragraph{Security.} The following theorem states the correctness and security of
+code partitioning.
 
@@
 
-\subsubsection*{Managing circuit sizes using partitioning}
+\paragraph{Implementing code partitioning.}
 We use partitioning for programs that require large
@@
 statically estimate the resource usage of a \tool program and we leave
-the construction of such an analysis as future work.
+the construction of such an analysis as future work. \nc{I changed the heading.}

### 2018-02-07 usenix18/eval.tex
edits on comparison with minionn and "surprises"

@@
 of \tool source code required is still small. 
-We note that \minion and \tool implementations are both based on ABY.
-These results are surprising as \minion has a much more efficient
-preprocessing phase and has SIMD 
-(single instruction multiple data) capabilities that allow it to perform
-matrix operations efficiently. In contrast, \tool focuses on generic
-2PC, and such optimizations are part of our future work.
+We note that while the \minion implementation is based on the ABY framework, it does not use the framework ``off-the-shelf'' and provide tailor-made optimizations for their applications.
+By contrast, \tool focuses on generic
+2PC, exploiting ABY's performant implementation. Optimizations such as those found in \minion are part of our future work.
 \minion also reports performance results on a bigger DNN with 7

### 2018-02-07 usenix18/ld.tex
some formatting in s3
spacing
formatting
formatting in s3
exactly one column above

@@
 
-\begin{figure}
-  \small
+\begin{figure}[htp]
+  \footnotesize
   \[
@@
 \label{fig:srclang}
-\end{figure}
-
-\begin{figure}
-  \small
-  \fbox{$\rho \vdash e \Downarrow v$}
-  \[
-  \\
-  \begin{array}{c}
-    \inferrule*[lab={\footnotesize{E-Var}}]
-               {
-               }
-               {
-                 \rho \vdash x \Downarrow \rho(x)
-               }
-
-               \hspace{0.1cm}
-
-    \inferrule*[lab={\footnotesize{E-Mult}}]
-               {
-                 \forall i \in \{1, 2\}.\:\rho \vdash e_{i} \Downarrow n_{i}
-               }
-               {
-                 \rho \vdash e_{1} \mulop e_{2} \Downarrow n_{1} \mulop n_{2}
-               }
-
-               \hspace{0.1cm}
-               \inferrule*[lab={\footnotesize{E-Cond}}]
-               {
-                 \rho \vdash e \Downarrow c\\\\
-                 c = \top \Rightarrow e_{3} = e_{1}\\\\
-                 c = \bot \Rightarrow e_{3} = e_{2}\\\\
-                 \rho \vdash e_{3} \Downarrow c_{3}
-               }
-               {
-                 \rho \vdash \cond{e}{e_{1}}{e_{2}} \Downarrow c_{3}
-               }
-
-\\\\
-	 \inferrule*[lab={\footnotesize{E-Read}}]
-               {
-                 \rho \vdash x \Downarrow [\overline{c_{i}}]_{n_{1}} \\\\
-                 \rho \vdash e \Downarrow n \quad n < n_{1}
-               }
-               {
-                 \rho \vdash x[e] \Downarrow c_{n}
-               }
-
-               \hspace{0.1cm}
-    \inferrule*[lab={\footnotesize{E-Arr}}]
-               {
-                 \forall i \in [n].\:\rho \vdash e_{i} \Downarrow c_{i}
-               }
-               {
-                 \rho \vdash [\overline{e_{i}}]_{n} \Downarrow [\overline{c_{i}}]_{n}
-               }
-               \hspace{0.1cm}
-    \inferrule*[lab={\footnotesize{E-Inp}}]
-               {
-               }
-               {
-                 \rho \vdash \kw{in}_{j} \Downarrow c
-               }
-  \end{array}
-  \]
-  \\\\
-    \fbox{$\rho \vdash s \Downarrow \rho_{1}; O$}
-  \[
-  \\
-  \begin{array}{c}
-    \inferrule*[lab={\footnotesize{E-Decl}}]
-               {
-                 \rho \vdash e \Downarrow v
-               }
-               {
-                 \rho \vdash \psi\:x = e \Downarrow \rho, x \mapsto v; \cdot
-               }
-
-               \hspace{0.1cm}
-
-    \inferrule*[lab={\footnotesize{E-LoopT}}]
-               {
-                 \rho(x) > n
-               }
-               {
-                 \rho \vdash \loops{x}{n}{s} \Downarrow \rho; \cdot
-               }
-
-               \\\\
-
-    \inferrule*[lab={\footnotesize{E-LoopI}}]
-               {
-                 \rho(x) \leq n\\\\
-                 \rho \vdash s \Downarrow \rho_{1}; O_{1}\\\\
-                 \rho_{2} = [\rho_{1}]_{\mathsf{dom}(\rho)}[x \mapsto \rho_{1}(x) + 1]\\\\
-                 \rho_{2} \vdash \loops{x}{n}{s} \Downarrow \rho_{3}; O_{2}
-               }
-               {
-                 \rho \vdash \loops{x}{n}{s} \Downarrow \rho_{3}; O_{1}, O_{2}
-               }
-
-               \hspace{0.2cm}
-
-    \inferrule*[lab={\footnotesize{E-If}}]
-               {
-                 \rho \vdash e \Downarrow c\\\\
-                 c = \top \Rightarrow s = s_{1}\\\\
-                 c = \bot \Rightarrow s = s_{2}\\\\
-                 \rho \vdash s \Downarrow \rho_{1}; O
-               }
-               {
-                 \rho \vdash \ite{e}{s_{1}}{s_{2}} \Downarrow \rho_{1}; O
-               }
-
-               \\\\
-
-    \inferrule*[lab={\footnotesize{E-For}}]
-               {
-                 \rho, x \mapsto n_{1} \vdash \loops{x}{n_{2}}{s} \Downarrow \rho_{1}; O
-               }
-               {
-                 \rho \vdash \for{x}{n_{1}}{n_{2}}{s} \Downarrow \rho_{1} - \{x\}; O
-               }
-
-               \hspace{0.2cm}
-
-    \inferrule*[lab={\footnotesize{E-Out}}]
-               {
-                 \rho \vdash e \Downarrow c
-               }
-               {
-                 \rho \vdash \kw{out}\:e \Downarrow \rho; c
-               }
-
-\end{array}
-  \]
-\caption{Source semantics (selected rules)}
-\label{fig:srcsem}
 \end{figure}
@@
 party $j$. The statements $s$ in the language comprise of variable
-declarations and assignments ($\psi\:x = e$ and $x := e$ resp.),
-\kw{for} loops, array writes ($x[e_{1}] := e_{2}$), \kw{if}
-statements, and sequence of statements ($s_{1}; s_{2}$). The statement
+declarations, assignments,
+\kw{for} loops, array writes, \kw{if}
+statements, and sequence of statements. The statement
 $\kw{out}\:e$ denotes revealing the value of $e$ to the
@@
 exposed to the programmer. \\
+
+\begin{figure}[htp]
+  \footnotesize
+  \fbox{$\rho \vdash e \Downarrow v$}\quad\quad
+  \fbox{$\rho \vdash s \Downarrow \rho_{1}; O$}
+  \[
+  \\
+  \begin{array}{c}
+    \inferrule*[lab={\footnotesize{E-Var}}]
+               {
+               }
+               {
+                 \rho \vdash x \Downarrow \rho(x)
+               }
+
+               \hspace{0.1cm}
+
+    \inferrule*[lab={\footnotesize{E-Mult}}]
+               {
+                 \forall i \in \{1, 2\}.\:\rho \vdash e_{i} \Downarrow n_{i}
+               }
+               {
+                 \rho \vdash e_{1} \mulop e_{2} \Downarrow n_{1} \mulop n_{2}
+               }
+
+               \hspace{0.1cm}
+               \inferrule*[lab={\footnotesize{E-Cond}}]
+               {
+                 \rho \vdash e \Downarrow c\\\\
+                 c = \top \Rightarrow e_{3} = e_{1}\\\\
+                 c = \bot \Rightarrow e_{3} = e_{2}\\\\
+                 \rho \vdash e_{3} \Downarrow c_{3}
+               }
+               {
+                 \rho \vdash \cond{e}{e_{1}}{e_{2}} \Downarrow c_{3}
+               }
+
+\\\\
+	 \inferrule*[lab={\footnotesize{E-Read}}]
+               {
+                 \rho \vdash x \Downarrow [\overline{c_{i}}]_{n_{1}} \\\\
+                 \rho \vdash e \Downarrow n \quad n < n_{1}
+               }
+               {
+                 \rho \vdash x[e] \Downarrow c_{n}
+               }
+
+               \hspace{0.1cm}
+    \inferrule*[lab={\footnotesize{E-Arr}}]
+               {
+                 \forall i \in [n].\:\rho \vdash e_{i} \Downarrow c_{i}
+               }
+               {
+                 \rho \vdash [\overline{e_{i}}]_{n} \Downarrow [\overline{c_{i}}]_{n}
+               }
+               \hspace{0.1cm}
+    \inferrule*[lab={\footnotesize{E-Inp}}]
+               {
+               }
+               {
+                 \rho \vdash \kw{in}_{j} \Downarrow c
+               }
+               \\\\
+    \inferrule*[lab={\footnotesize{E-Decl}}]
+               {
+                 \rho \vdash e \Downarrow v
+               }
+               {
+                 \rho \vdash \psi\:x = e \Downarrow \rho, x \mapsto v; \cdot
+               }
+
+               \hspace{0.1cm}
+
+    \inferrule*[lab={\footnotesize{E-LoopT}}]
+               {
+                 \rho(x) > n
+               }
+               {
+                 \rho \vdash \loops{x}{n}{s} \Downarrow \rho; \cdot
+               }
+
+               \\\\
+
+    \inferrule*[lab={\footnotesize{E-LoopI}}]
+               {
+                 \rho(x) \leq n\\\\
+                 \rho \vdash s \Downarrow \rho_{1}; O_{1}\\\\
+                 \rho_{2} = [\rho_{1}]_{\mathsf{dom}(\rho)}[x \mapsto \rho_{1}(x) + 1]\\\\
+                 \rho_{2} \vdash \loops{x}{n}{s} \Downarrow \rho_{3}; O_{2}
+               }
+               {
+                 \rho \vdash \loops{x}{n}{s} \Downarrow \rho_{3}; O_{1}, O_{2}
+               }
+
+               \hspace{0.2cm}
+
+    \inferrule*[lab={\footnotesize{E-If}}]
+               {
+                 \rho \vdash e \Downarrow c\\\\
+                 c = \top \Rightarrow s = s_{1}\\\\
+                 c = \bot \Rightarrow s = s_{2}\\\\
+                 \rho \vdash s \Downarrow \rho_{1}; O
+               }
+               {
+                 \rho \vdash \ite{e}{s_{1}}{s_{2}} \Downarrow \rho_{1}; O
+               }
+
+               \\\\
+
+    \inferrule*[lab={\footnotesize{E-For}}]
+               {
+                 \rho, x \mapsto n_{1} \vdash \loops{x}{n_{2}}{s} \Downarrow \rho_{1}; O
+               }
+               {
+                 \rho \vdash \for{x}{n_{1}}{n_{2}}{s} \Downarrow \rho_{1} - \{x\}; O
+               }
+
+               \hspace{0.2cm}
+
+    \inferrule*[lab={\footnotesize{E-Out}}]
+               {
+                 \rho \vdash e \Downarrow c
+               }
+               {
+                 \rho \vdash \kw{out}\:e \Downarrow \rho; c
+               }
+
+\end{array}
+  \]
+\caption{Source semantics (selected rules)}
+\label{fig:srcsem}
+\end{figure}
 
@@
 $
-\small
+\footnotesize
 \begin{array}{rrcl}
@@
 ({\sc{E-Decl}}) evaluates the expression $e$ to $v$, and returns the
-updated environment $\rho,x \mapsto v$, with empty observations. Rule
-({\sc{E-If}}) evaluates the guard expression, and then evaluates
-either $s_{1}$ or $s_{2}$ accordingly. The $\kw{for}$ statements evaluate
+updated environment $\rho,x \mapsto v$, with empty observations. The
+$\kw{for}$ statements evaluate
 through the internal $\kw{while}$ syntax. Specifically, the rule
@@
 
-\begin{figure}[t]
-  \small
-  \fbox{$\Gamma \vdash e : \tau \leadsto \widetilde{e}$}
-  \[
+\begin{figure}[H]
+  \footnotesize
+  \fbox{$\Gamma \vdash e : \tau \leadsto \widetilde{e}$}\quad\quad
+  \fbox{$\Gamma \vdash s \leadsto \widetilde{s} \mid \Gamma_{1}$}  \[
   \begin{array}{c}
@@
                }
-
-    %% \inferrule*[lab={\footnotesize{T-Cons}}]
-    %%            {
-    %%              \tau = \mathsf{typeof}(c)^{\mathcal{P}}
-    %%            }
-    %%            {
-    %%              \Gamma \vdash c : \tau \leadsto c
-    %%            }
-
-    %%  \inferrule*[lab={\footnotesize{T-Var}}]
-    %%            {
-    %%            }
-    %%            {
-    %%              \Gamma \vdash x : \Gamma(x) \leadsto x
-    %%            }
-\\\\
-
+               \\\\
 	  \inferrule*[lab={\footnotesize{T-Mult}}]
@@
                }
-
-  \end{array}
-  \]
-  \\\\
-    \fbox{$\Gamma \vdash s \leadsto \widetilde{s} \mid \Gamma_{1}$}
-  \[
-  \\
-  \begin{array}{c}
+\\\\
      \inferrule*[lab={\footnotesize{T-Decl}}]
@@
   \]
-\caption{Compilation judgments (selected rules)}
+\caption{Source compilation (selected rules)}
 \label{fig:compile}
@@
 judgment $ \models e < n$ to prove that the array index is in
-bounds. \sectionref{impl} discusses our implementation of this
-check. Rule ({\sc{T-Inp}}) picks a label $m$ for the input. Finally,
+bounds \footnote{\sectionref{impl} discusses our implementation of this
+check.}. Rule ({\sc{T-Inp}}) picks a label $m$ for the input. Finally,
 the rule ({\sc{T-Sub}}) is the subsumption rule that coerces an expression of
@@
 $\kw{uint}^{\mathcal{P}}$, and delegates type checking to the
-\kw{while} form. The rule ({\sc{T-Write}}), similar to ({\sc{T-Read}}),
-checks that the index has type $\kw{uint}^{\mathcal{P}}$, and is in
-bounds. Rule ({\sc{T-Out}})
+\kw{while} form. Rule ({\sc{T-Out}})
 types the expression $e$ at some secret label $m$. Rule ({\sc{T-If}})
@@
 
-As mentioned earlier, the intermediate language models the code such
-as in \figureref{ex-aby} output by our compiler.
-Next, a program in the intermediate language is evaluated to a circuit
-that can be executed in the distributed runtime later. The evaluation
-to a circuit computes away the public parts of the program and also
-\emph{flattens} the arrays so that the circuits are unaware of the
-array structure.
-Crucially, this phase of the semantics does not have access to the
-secret inputs.
-Below, we first provide the language for the circuits followed by the
-evaluation rules.
-
 \begin{figure}
-  \small
+  \footnotesize
   \[
@@
 
-\begin{figure}[t]
-  \small
-  \fbox{$\widetilde{\rho} \vdash \widetilde{e} \Downarrow \val$}
-  \[
-  \\
-  \begin{array}{c}
-    \inferrule*[lab={\footnotesize{S-Var}}]
-               {
-               }
-               {
-                 \widetilde{\rho} \vdash x \Downarrow \widetilde{\rho}(x)
-               }
-
-    \inferrule*[lab={\footnotesize{S-PMult}}]
-               {
-                 \forall i \in \{1, 2\}.\:\widetilde{\rho} \vdash \widetilde{e}_{i} \Downarrow n_{i}
-               }
-               {
-                 \widetilde{\rho} \vdash \widetilde{e}_{1} \mulop_{\mathcal{P}} \widetilde{e}_{2} \Downarrow n_{1} \mulop n_{2}
-               }
-
-    \inferrule*[lab={\footnotesize{S-Read}}]
-               {
-                 \widetilde{\rho} \vdash x \Downarrow [\overline{\wire_i}]_{n_{1}} \\\\
-                 \widetilde{\rho} \vdash \widetilde{e} \Downarrow n \quad n < n_{1}
-               }
-               {
-                 \widetilde{\rho} \vdash x[\widetilde{e}] \Downarrow \wire_n
-               }\\\\
-    \inferrule*[lab={\footnotesize{S-SMult}}]
-               {
-                 \forall i \in \{1, 2\}.\:\widetilde{\rho} \vdash \widetilde{e}_{i} \Downarrow \gate_{i}
-               }
-               {
-                 \widetilde{\rho} \vdash \widetilde{e}_{1} \mulop_{\mathcal{A}} \widetilde{e}_{2} \Downarrow \kw{mult}\:\gate_{1}\:\gate_{2}
-               }
-
-               \hspace{0.3cm}
-
-    \inferrule*[lab={\footnotesize{S-SGt}}]
-               {
-                 \forall i \in \{1, 2\}.\:\widetilde{\rho} \vdash \widetilde{e_{i}} \Downarrow \gate_{i}
-               }
-               {
-                 \widetilde{\rho} \vdash \widetilde{e_{1}} >_{\mathcal{B}} \widetilde{e_{2}} \Downarrow \kw{gt}\:\gate_{1}\:\gate_{2}
-               }\\\\
-               \inferrule*[lab={\footnotesize{S-SCond}}]
-               {
-                 \forall i \in \{1, 2, 3\}.\:\widetilde{\rho} \vdash \widetilde{e}_{i} \Downarrow \gate_{i}
-               }
-               {
-                 \widetilde{\rho} \vdash \lcond{\mathcal{B}}{\widetilde{e}_{1}}{\widetilde{e}_{2}}{\widetilde{e}_{3}} \Downarrow \kw{mux}\:\gate_{1}\:\gate_{2}\:\gate_{3}
-               }
-               \hspace{0.2cm}
-    \inferrule*[lab={\footnotesize{S-Coerce}}]
-               {
-                 \widetilde{\rho} \vdash \widetilde{e} \Downarrow \gate
-               }
-               {
-                 \widetilde{\rho} \vdash \langle \ell \rhd m \rangle\:
-                 \widetilde{e} \Downarrow \langle \ell \rhd m \rangle\:\gate
-               }
-               \\\\
-               %% \inferrule*[lab={\footnotesize{S-Arr}}]
-               %% {
-               %%   \forall i \in \{0 \dots n - 1\}.\:\widetilde{\rho} \vdash \widetilde{e_{i}} \Downarrow \widetilde{w_{i}}
-               %% }
-               %% {
-               %%   \widetilde{\rho} \vdash [\overline{\widetilde{e_{i}}}]_{n} \Downarrow [\overline{\widetilde{w_{i}}}]_{n}
-               %% }
-
-               \inferrule*[lab={\footnotesize{S-PCond}}]
-               {
-                 \widetilde{\rho} \vdash \widetilde{e} \Downarrow c\\\\
-                 c = \top \Rightarrow \widetilde{e}_{3} = \widetilde{e}_{1}\\
-                 c = \bot \Rightarrow \widetilde{e}_{3} = \widetilde{e}_{2}\\\\
-                 \widetilde{\rho} \vdash \widetilde{e}_{3} \Downarrow \val
-               }
-               {
-                 \widetilde{\rho} \vdash \lcond{\mathcal{P}}{\widetilde{e}}{\widetilde{e}_{1}}{\widetilde{e}_{2}} \Downarrow \val
-               }
-               %\hspace{0.5cm}
-
-    \inferrule*[lab={\footnotesize{S-Inp}}]
-               {
-               }
-               {
-                 \widetilde{\rho} \vdash \kw{in}^{m}_{j} \Downarrow \kw{in}^{m}_{j}
-               }
-  \end{array}
-  \]
-  \\\\
-    \fbox{$\widetilde{\rho} \vdash \widetilde{s} \Downarrow \widetilde{\rho}_{1}; \crct$}
-  \[
-  \\
-  \begin{array}{c}
-    \inferrule*[lab={\footnotesize{S-DeclC}}]
-               {
-                 \widetilde{\rho} \vdash \widetilde{e} \Downarrow \val\\\\
-                 \left(\val = c\right) \vee \left(\val = [\overline{c_{i}}]_{n}\right)
-               }
-               {
-                 \widetilde{\rho} \vdash \tau\:x = \widetilde{e} \Downarrow \widetilde{\rho}, x \mapsto \val; \cdot
-               }
-
-               \hspace{0.2cm}
-
-    \inferrule*[lab={\footnotesize{S-DeclCkt}}]
-               {
-                 \widetilde{\rho} \vdash \widetilde{e} \Downarrow \gate\quad\quad
-                 \mathsf{fresh}\:\wire
-               }
-               {
-                 \widetilde{\rho} \vdash \tau\:x = \widetilde{e} \Downarrow \widetilde{\rho}, x \mapsto \wire; \kw{bind}\:\gate\:\wire
-               }
-
-\\\\
-    \inferrule*[lab={\footnotesize{S-DeclCktA}}]
-               {
-                 \widetilde{\rho} \vdash \widetilde{e} \Downarrow [\overline{\gate_{i}}]_{n}\\\\
-                 \forall i \in [n].\:\mathsf{fresh}\:\wire_{i}
-               }
-               {
-                 \widetilde{\rho} \vdash \tau\:x = \widetilde{e} \Downarrow \widetilde{\rho}, x \mapsto [\overline{\wire_{i}}]_{n}; \overline{\kw{bind}\:\gate_{i}\:\wire_{i}}
-               }
-
-               \hspace{0.2cm}
-
-    \inferrule*[lab={\footnotesize{S-Out}}]
-               {
-                 \widetilde{\rho} \vdash \widetilde{e} \Downarrow \gate
-               }
-               {
-                 \widetilde{\rho} \vdash \kw{out}\:\widetilde{e} \Downarrow \widetilde{\rho}; \kw{out}\:\gate
-               }
-
-               \\\\
-    \inferrule*[lab={\footnotesize{S-If}}]
-               {
-                 \widetilde{\rho} \vdash \widetilde{e} \Downarrow c\\\\
-                 c = \top \Rightarrow \widetilde{s} = \widetilde{s_{1}}\\\\
-                 c = \bot \Rightarrow \widetilde{s} = \widetilde{s_{2}}\\\\
-                 \widetilde{\rho} \vdash \widetilde{s} \Downarrow \widetilde{\rho}_{1}; \crct
-               }
-               {
-                 \widetilde{\rho} \vdash \ite{\widetilde{e}}{\widetilde{s_{1}}}{\widetilde{s_{2}}} \Downarrow \widetilde{\rho}_{1}; \crct
-               }
-
-               \hspace{0.5cm}
-
-    \inferrule*[lab={\footnotesize{S-WriteCkt}}]
-               {
-                 \widetilde{\rho} \vdash x \Downarrow [\overline{\wire_{i}}]_{n}\quad
-                 \widetilde{\rho} \vdash \widetilde{e}_{1} \Downarrow n_{1}\\\\
-                 n_{1} < n\quad
-                 \mathsf{fresh}\:\wire\quad
-                 \widetilde{\rho} \vdash \widetilde{e}_{2} \Downarrow \gate\\\\
-                 \widetilde{\rho}_{1} = \widetilde{\rho}[x \mapsto ([\overline{\wire_{i}}]_{n}[n_{1} \mapsto \wire])]
-               }
-               {
-                 \widetilde{\rho} \vdash x[\widetilde{e}_{1}] := \widetilde{e}_{2} \Downarrow \widetilde{\rho}_{1}; \kw{bind}\:\gate\:\wire
-               }
-
-\end{array}
-  \]
-\caption{Evaluation of Intermediate Language to Circuit (selected rules)}
-\label{fig:eval-inter-circ}
-\end{figure}
+As mentioned earlier, the intermediate language models the code such
+as in \figureref{ex-aby} output by our compiler.
+Next, a program in the intermediate language is evaluated to a circuit
+that can be executed in the distributed runtime later. The evaluation
+to a circuit computes away the public parts of the program and also
+\emph{flattens} the arrays so that the circuits are unaware of the
+array structure.
+Crucially, this phase of the semantics does not have access to the
+secret inputs.
+Below, we first provide the language for the circuits followed by the
+evaluation rules.
 
@@
 
+\begin{figure}[h]
+  \footnotesize
+  \fbox{$\widetilde{\rho} \vdash \widetilde{e} \Downarrow \val$}\quad\quad
+  \fbox{$\widetilde{\rho} \vdash \widetilde{s} \Downarrow \widetilde{\rho}_{1}; \crct$}
+  \[
+  \\
+  \begin{array}{c}
+    \inferrule*[lab={\footnotesize{S-Var}}]
+               {
+               }
+               {
+                 \widetilde{\rho} \vdash x \Downarrow \widetilde{\rho}(x)
+               }
+
+    \inferrule*[lab={\footnotesize{S-PMult}}]
+               {
+                 \forall i \in \{1, 2\}.\:\widetilde{\rho} \vdash \widetilde{e}_{i} \Downarrow n_{i}
+               }
+               {
+                 \widetilde{\rho} \vdash \widetilde{e}_{1} \mulop_{\mathcal{P}} \widetilde{e}_{2} \Downarrow n_{1} \mulop n_{2}
+               }
+
+    \inferrule*[lab={\footnotesize{S-Read}}]
+               {
+                 \widetilde{\rho} \vdash x \Downarrow [\overline{\wire_i}]_{n_{1}} \\\\
+                 \widetilde{\rho} \vdash \widetilde{e} \Downarrow n \quad n < n_{1}
+               }
+               {
+                 \widetilde{\rho} \vdash x[\widetilde{e}] \Downarrow \wire_n
+               }\\\\
+    \inferrule*[lab={\footnotesize{S-SMult}}]
+               {
+                 \forall i \in \{1, 2\}.\:\widetilde{\rho} \vdash \widetilde{e}_{i} \Downarrow \gate_{i}
+               }
+               {
+                 \widetilde{\rho} \vdash \widetilde{e}_{1} \mulop_{\mathcal{A}} \widetilde{e}_{2} \Downarrow \kw{mult}\:\gate_{1}\:\gate_{2}
+               }
+
+               \hspace{0.3cm}
+
+    \inferrule*[lab={\footnotesize{S-SGt}}]
+               {
+                 \forall i \in \{1, 2\}.\:\widetilde{\rho} \vdash \widetilde{e_{i}} \Downarrow \gate_{i}
+               }
+               {
+                 \widetilde{\rho} \vdash \widetilde{e_{1}} >_{\mathcal{B}} \widetilde{e_{2}} \Downarrow \kw{gt}\:\gate_{1}\:\gate_{2}
+               }\\\\
+               \inferrule*[lab={\footnotesize{S-SCond}}]
+               {
+                 \forall i \in \{1, 2, 3\}.\:\widetilde{\rho} \vdash \widetilde{e}_{i} \Downarrow \gate_{i}
+               }
+               {
+                 \widetilde{\rho} \vdash \lcond{\mathcal{B}}{\widetilde{e}_{1}}{\widetilde{e}_{2}}{\widetilde{e}_{3}} \Downarrow \kw{mux}\:\gate_{1}\:\gate_{2}\:\gate_{3}
+               }
+               \hspace{0.2cm}
+    \inferrule*[lab={\footnotesize{S-Coerce}}]
+               {
+                 \widetilde{\rho} \vdash \widetilde{e} \Downarrow \gate
+               }
+               {
+                 \widetilde{\rho} \vdash \langle \ell \rhd m \rangle\:
+                 \widetilde{e} \Downarrow \langle \ell \rhd m \rangle\:\gate
+               }
+               \\\\
+               %% \inferrule*[lab={\footnotesize{S-Arr}}]
+               %% {
+               %%   \forall i \in \{0 \dots n - 1\}.\:\widetilde{\rho} \vdash \widetilde{e_{i}} \Downarrow \widetilde{w_{i}}
+               %% }
+               %% {
+               %%   \widetilde{\rho} \vdash [\overline{\widetilde{e_{i}}}]_{n} \Downarrow [\overline{\widetilde{w_{i}}}]_{n}
+               %% }
+
+               \inferrule*[lab={\footnotesize{S-PCond}}]
+               {
+                 \widetilde{\rho} \vdash \widetilde{e} \Downarrow c\\\\
+                 c = \top \Rightarrow \widetilde{e}_{3} = \widetilde{e}_{1}\\
+                 c = \bot \Rightarrow \widetilde{e}_{3} = \widetilde{e}_{2}\\\\
+                 \widetilde{\rho} \vdash \widetilde{e}_{3} \Downarrow \val
+               }
+               {
+                 \widetilde{\rho} \vdash \lcond{\mathcal{P}}{\widetilde{e}}{\widetilde{e}_{1}}{\widetilde{e}_{2}} \Downarrow \val
+               }
+               %\hspace{0.5cm}
+
+    \inferrule*[lab={\footnotesize{S-Inp}}]
+               {
+               }
+               {
+                 \widetilde{\rho} \vdash \kw{in}^{m}_{j} \Downarrow \kw{in}^{m}_{j}
+               }
+               \\\\
+    \inferrule*[lab={\footnotesize{S-DeclC}}]
+               {
+                 \widetilde{\rho} \vdash \widetilde{e} \Downarrow \val\\\\
+                 \left(\val = c\right) \vee \left(\val = [\overline{c_{i}}]_{n}\right)
+               }
+               {
+                 \widetilde{\rho} \vdash \tau\:x = \widetilde{e} \Downarrow \widetilde{\rho}, x \mapsto \val; \cdot
+               }
+
+               \hspace{0.2cm}
+
+    \inferrule*[lab={\footnotesize{S-DeclCkt}}]
+               {
+                 \widetilde{\rho} \vdash \widetilde{e} \Downarrow \gate\quad\quad
+                 \mathsf{fresh}\:\wire
+               }
+               {
+                 \widetilde{\rho} \vdash \tau\:x = \widetilde{e} \Downarrow \widetilde{\rho}, x \mapsto \wire; \kw{bind}\:\gate\:\wire
+               }
+
+\\\\
+    \inferrule*[lab={\footnotesize{S-DeclCktA}}]
+               {
+                 \widetilde{\rho} \vdash \widetilde{e} \Downarrow [\overline{\gate_{i}}]_{n}\\\\
+                 \forall i \in [n].\:\mathsf{fresh}\:\wire_{i}
+               }
+               {
+                 \widetilde{\rho} \vdash \tau\:x = \widetilde{e} \Downarrow \widetilde{\rho}, x \mapsto [\overline{\wire_{i}}]_{n}; \overline{\kw{bind}\:\gate_{i}\:\wire_{i}}
+               }
+
+               \hspace{0.2cm}
+
+    \inferrule*[lab={\footnotesize{S-Out}}]
+               {
+                 \widetilde{\rho} \vdash \widetilde{e} \Downarrow \gate
+               }
+               {
+                 \widetilde{\rho} \vdash \kw{out}\:\widetilde{e} \Downarrow \widetilde{\rho}; \kw{out}\:\gate
+               }
+
+               \\\\
+    \inferrule*[lab={\footnotesize{S-If}}]
+               {
+                 \widetilde{\rho} \vdash \widetilde{e} \Downarrow c\\\\
+                 c = \top \Rightarrow \widetilde{s} = \widetilde{s_{1}}\\\\
+                 c = \bot \Rightarrow \widetilde{s} = \widetilde{s_{2}}\\\\
+                 \widetilde{\rho} \vdash \widetilde{s} \Downarrow \widetilde{\rho}_{1}; \crct
+               }
+               {
+                 \widetilde{\rho} \vdash \ite{\widetilde{e}}{\widetilde{s_{1}}}{\widetilde{s_{2}}} \Downarrow \widetilde{\rho}_{1}; \crct
+               }
+
+               \hspace{0.5cm}
+
+    \inferrule*[lab={\footnotesize{S-WriteCkt}}]
+               {
+                 \widetilde{\rho} \vdash x \Downarrow [\overline{\wire_{i}}]_{n}\quad
+                 \widetilde{\rho} \vdash \widetilde{e}_{1} \Downarrow n_{1}\\\\
+                 n_{1} < n\quad
+                 \mathsf{fresh}\:\wire\quad
+                 \widetilde{\rho} \vdash \widetilde{e}_{2} \Downarrow \gate\\\\
+                 \widetilde{\rho}_{1} = \widetilde{\rho}[x \mapsto ([\overline{\wire_{i}}]_{n}[n_{1} \mapsto \wire])]
+               }
+               {
+                 \widetilde{\rho} \vdash x[\widetilde{e}_{1}] := \widetilde{e}_{2} \Downarrow \widetilde{\rho}_{1}; \kw{bind}\:\gate\:\wire
+               }
+
+\end{array}
+  \]
+\caption{Evaluation of Intermediate Language to Circuit (selected rules)}
+\label{fig:eval-inter-circ}
+\end{figure}
+
 \figureref{eval-inter-circ} shows the judgments for the evaluation of
@@
 \begin{figure}
-  \small
-  \fbox{$\widehat{\rho}_{1}, \widehat{\rho}_{2} \vdash \gate \Downarrow b_{1}, b_{2}$}
+  \footnotesize
+  \fbox{$\widehat{\rho}_{1}, \widehat{\rho}_{2} \vdash \gate \Downarrow b_{1}, b_{2}$}\quad\quad
+  \fbox{$\widehat{\rho}_{1}, \widehat{\rho}_{2} \vdash \crct \Downarrow \widehat{\rho}'_{1}, \widehat{\rho}'_{2}; O$}
   \[
@@
 
-  \end{array}
-  \]
-  \\\\
-    \fbox{$\widehat{\rho}_{1}, \widehat{\rho}_{2} \vdash \crct \Downarrow \widehat{\rho}'_{1}, \widehat{\rho}'_{2}; O$}
-  \[
-  \\
-  \begin{array}{c}
+  \\\\
     \inferrule*[lab={\footnotesize{C-Bind}}]

### 2018-02-07 usenix18/ld.tex
more dirty tricks

@@
 We present only selected parts of our formalization. Full definitions
-and proofs can be found in~\cite{anonymized}.
+and proofs can be found in the Appendix.
 
@@
   \]
-\caption{Source semantics (selected rules)}
+\caption{Source semantics}
 \label{fig:srcsem}
@@
   \]
-\caption{Source compilation (selected rules)}
+\caption{Source compilation}
 \label{fig:compile}
@@
   \]
-\caption{Evaluation of Intermediate Language to Circuit (selected rules)}
+\caption{Evaluation of Intermediate Language to Circuit}
 \label{fig:eval-inter-circ}
@@
   \]
-\caption{Circuit semantics in a distributed runtime (selected rules)}
+\caption{Circuit semantics in a distributed runtime}
 \label{fig:cktsem}

### 2018-02-07 usenix18/pipe.tex
more dirty tricks

@@
 statically estimate the resource usage of a \tool program and we leave
-the construction of such an analysis as future work. \nc{I changed the heading.}
+the construction of such an analysis as future work.% \nc{I changed the heading.}

### 2018-02-07 usenix18/intro.tex
edits on comparison with minionn and "surprises"
space crunches.

@@
   %% \TODO{While taking a pass on intro, check citation correctness or if we are missing any citation.}
-
+\vspace{-0.1in}
 \section{Introduction}
@@
 arithmetic and boolean circuits, and even for an expert cryptographer,
-writing large computations in such a framework can be tedious and
-subject to errors (a sentiment echoed by Demmler et al.~\cite{aby} themselves).
+writing large computations in such a framework can be tedious (a sentiment echoed by Demmler et al.~\cite{aby} themselves).
 
@@
 behind the design of specialized protocols. Using \tool, we empirically  demonstrate
-the surprising fact that generic \mpc implementations are much more efficient than what they were believed to be. 
+the that generic \mpc implementations are much more efficient than what they were believed to be. 
  %(up to a factor of 19x in our
@@
 \end{figure}
-
+\vspace{-0.1in}
 \paragraph{Ease of programming.} \tool source programs are ideal
@@
 shows an \tool source
-program for $w^Tx >b$.
-%
-%\vspace{0.2cm}
-%
-%\vspace{0.2cm}
-The program is quite similar to what a programmer might
+program for $w^Tx >b$. The program is quite similar to what a programmer might
 write in C++ or Java. The simplicity of the
-language comes with its usual benefits: it is easily accessible to the
+language comes with the usual benefits: it is easily accessible to the
 developers, there are fewer avenues for making mistakes, developers
@@
 enjoy these benefits.
-
+\vspace{-0.1in}
 \paragraph{Cryptographic-cost aware compiler.} The \tool compiler
@@
 first such cryptographic-cost aware compiler.
-
+\vspace{-0.1in}
 \paragraph{Scalability (secure code partitioning).} \mpc tools
@@
 %% in \sectionref{pipe}
-
+\vspace{-0.1in}
 \paragraph{Formal guarantees.} We prove formal correctness and
@@
 scheme (\sectionref{ld} and~\sectionref{pipe}).
-
+%\vspace{-0.2in}
 \paragraph{Evaluation.} We have implemented \tool
@@
 that requires more than 300 million gates (\sectionref{impl} and~\sectionref{eval}).
-
+\vspace{-0.1in}
 \paragraph{Related Work.}

### 2018-02-07 usenix18/related.tex
space crunches.

--- 2017-11-01 OaklandSubmission/related.tex
@@
+\vspace{-0.1in}
 \section{Related work}
@@
 The work of Kerschbaum et al.~\cite{autoS} provides a scheme to automatically assign homomorphic encryption or garbled circuits to each operator in a computation that is expressed as a sequence of dyadic operations.
-They conjecture that the problem is NP hard and gave a linear programming based solution as well as a quadratic time greedy heuristic.
+They conjecture that the problem is NP hard and gave a linear programming based solution and a quadratic time greedy heuristic.
 These techniques are not directly applicable to \tool programs because of $\mathtt{for}$-loops and $\mathtt{if}$-conditions.
-However, we are exploring whether these ideas can be extended to yield a better type inference.
+However, we are exploring if these ideas can be extended to yield better type inference.
 %Their performance is much poorer than ABY due to the high 
@@
 Similarly, the TASTY tool~\cite{tasty} provides a way to mix both boolean and arithmetic circuits (using homomorphic encryption for the arithmetic compute).
-Other examples include the VIFF framework~\cite{viff} for arithmetic computations and Sharemind~\cite{sharemind} (that provides secure 3-party boolean computation). 
+Other examples include the VIFF framework~\cite{viff} for arithmetic computations and Sharemind~\cite{sharemind} (secure 3-party boolean computation). 
 %\\\\
@@
 
-The implementations of \mpc backends  have made  tremendous progress in the last decade.
+\mpc backends  have made  tremendous progress in the last decade.
 For example, the circuits can be optimized for depth~\cite{ddkssz15,cbmcgcdepth}, large garbled circuits can be pipelined~\cite{yao-pipe}, and oblivious RAM~\cite{oram1,oram2} can be used to hide access patterns of MIPS code~\cite{mips}.

### 2018-02-08 usenix18/pipe.tex
space crunches.
only a couple of lines more
minor edits.

@@
+\vspace{-0.1in}
 \section{Secure code partitioning}
@@
 $\progn_1,\progn_2,\ldots,\progn_k$ does
-not reveal any more information than $\prog$). More details follow.
+not reveal any more information than $\prog$). More formal details follow, while an example illustrating the technique of secure code partitioning is provided in  \appendixref{codepartitioning}.
 
@@
   protocol in the case where both parties must receive the same
-  output. To do so, we modify $\progn_1$ to output $o_1 = s_1+r_1+r'_1$ to
-  both parties (where $r'_1$ is random and chosen by Bob). We can then
-  appropriately modify the remaining steps as well.}. Alice's output
+  output. %To do so, we modify $\progn_1$ to output $o_1 = s_1+r_1+r'_1$ to
+  %both parties (where $r'_1$ is random and chosen by Bob). We can then
+  %appropriately modify the remaining steps as well.
+  }. 
+  Alice's output
 from $\progn_1$ is $r_1$. Next, every $\progn_i$ ($2\leq i\leq k-1$)
@@
 can be sequentially executed, using the unmodified underlying \mpc
-backend. An example illustrating the technique of secure code partitioning is provided in  \appendixref{codepartitioning}.
+backend. We prove the following theorem for code partitioning:
 
-\paragraph{Security.} The following theorem states the correctness and security of
-code partitioning.
+%% \paragraph{Security.} The following theorem states the correctness and security of
+%% code partitioning.
 
@@
 %output $z$.
-
+\vspace{-0.1in}
 \paragraph{Implementing code partitioning.}

### 2018-02-08 usenix18/ex.tex
formatting
space crunches.
some more minor edits.

@@
+\vspace{-0.1in}
 \section{\tool Overview}
@@
 toolchain. We give a brief overview of each of these phases below.
-
+\vspace{-0.15in}
 \paragraph{Source language.}
 Consider the example $w^Tx >b$ from \sectionref{intro}, where
-$w$ and $b$ constitute the server's input (a classifier) and $x$ is
+$w$ and $b$ are the server's input (a classifier) and $x$ is
 the client's input vector. \figureref{ex-sml} shows the \tool code
@@
 \ls{for} loop to compute \ls{acc}, the inner product of \ls{w} and
-\ls{x}. Finally, the code outputs the result of comparing \ls{acc}
-with \ls{b} only to the client using \ls{output2}.
+\ls{x}. Finally, the code compares \ls{acc}
+with \ls{b} and outputs this only to the client using \ls{output2}.
 
@@
 \ls{if}~statements, and special syntax for input/output from each party.
-
+\vspace{-0.1in}
 \paragraph{\tool compiler.}
@@
 input to the comparison and multiplexer circuits.
-
-\begin{figure}
+\vspace{-0.2in}
+\begin{figure}[h]
 \begin{lstlisting}[language=C, mathescape=true]
@@
 a \mpc protocol.
-
-We can now concretely see the advantages of \tool. Unarguably, it is
-easier for a developer to program and get right the code in
+\vspace{-0.1in}
+\paragraph{Advantages of \tool.} We can now concretely see the advantages of \tool. Unarguably, it is
+easier for a developer to program and get the code right in
 \figureref{ex-sml}, rather than the code in
@@
 change from multiplication to bitwise-or in the \ls{for} loop. It
-turns out that once that's the case, it is more efficient to do both
+turns out that in this case, it is more efficient to do both
 the addition and bitwise-or using boolean circuits (if the addition is
-still done using arithmetic, the conversion cost starts to take
+done using arithmetic, the conversion cost starts to take
 over). In \tool, the programmer simply needs to change one operator in
@@
 boolean addition. Whereas, if the programmer was writing ABY
-code, she either has to sacrifice performance, or she would have to
-revisit many parts of the circuit and change them.
-
-To summarize, \tool raises the level of abstraction for the
+code, she either has to sacrifice performance, or would have to
+revisit many parts of the circuit and change them. In summary, \tool raises the level of abstraction for the
 programmer, and generates efficient \mpc protocols automatically,

### 2018-02-08 usenix18/ex.tex
nits here and there

@@
 \tool compiler takes as input a source program and produces a C++
-program as output. \figureref{ex-aby} shows the output code\footnote{This is also how a program written directly in
-  ABY would look like.} for
-the example in \figureref{ex-sml}. The output program
+program as output. \figureref{ex-aby} shows the output code for
+the example in \figureref{ex-sml} -- this is also how a program written directly in
+  ABY would look like. The output program
 contains party-specific code for inputs and outputs
@@
 input to the comparison and multiplexer circuits.
-\vspace{-0.2in}
+%\vspace{-0.2in}
+
 \begin{figure}[h]

### 2018-03-21 usenix18/rebuttal.txt
reviews and rebuttal

We thank the reviewers for their useful comments. We address a few of them below.
First, our language does indeed allow providing input that is public, by specifying the value to be a constant public value (is this correct and can we write the syntax for it?). An example of public input is loop counter value. This ensures that such values need not be secret-shared leading to high overheads. However, this input is hard-coded during compile time and we currently do not allow users to input such values at runtime (can we say why?)
As noted in prior works ([19], [27], [32]), purely Boolean circuits have reasonable performance when dealing with small input sizes (e.g. Groce et al.); even a simple task such as matrix multiplication can be infeasible to run using garbled circuits when the matrix dimensions are large (e.g. in our machine learning applications). In fact, for these applications, the Yao’s circuit generated would not fit in memory. Additionally, the work of Groce et al. only compares online times with prior work and do not include pre-processing time. We remark that these benchmarks are too small for any appreciable differences in performance between various protocols and we only included them to compare our work with other secure ML classification works. However, we thank the reviewer for the pointer to Groce et al. and shall include it in the paper. Another point to be noted is that our framework can be modified to incorporate further optimized backends (for Boolean, e.g. Groce et al.) and for Arithmetic.
We do not claim that our compiler choices are optimal. Multiplication is always done using arithmetic secret sharing as the cost of matrix multiplication in Boolean shares is much higher; addition is done either over Boolean or Arithmetic shares depending on prior shares that are available and not stale. Obtaining a compiler with better choices is an interesting optimization problem that could potentially consider other parameters such as network latency, bandwidth, compute resources and so on. We leave this exploration to future work.
Secure code partitioning is not a new technique by itself. We only claim that it provides a generic way to execute large programs without going into the details of underlying individual secure protocols. We implement it only in the arithmetic setting currently as share reconstruction is simple 32-bit addition, which has roughly the same cost in both the Boolean and arithmetic settings. 
Our new language is superior to other prior domain-specific language in that it is the only language to allow compilation into a backend that allows mixed arithmetic and Boolean computation. The task of figuring out which part of compute to execute using which type of sharing is unique to our compiler.
We agree with the reviewer that EzPC does not support secret array access and combining the techniques of EzPC and ObliVM to support Boolean, arithmetic and secret array access would be an interesting future direction. However, for our benchmarks (e.g. ML algorithms such as DNNs/CNNs), the access patterns are uniform and so providing for secret array accesses will not improve performance.
We agree with the other comments made by the reviewers and will incorporate them accordingly.
One or two comments have not been addressed and we are slightly over the word limit (500 words).

### 2018-03-21 usenix18/rebuttal.txt
comments

@@
 We thank the reviewers for their useful comments. We address a few of them below.
-First, our language does indeed allow providing input that is public, by specifying the value to be a constant public value (is this correct and can we write the syntax for it?).
-An example of public input is loop counter value.
-This ensures that such values need not be secret-shared leading to high overheads.
-However, this input is hard-coded during compile time and we currently do not allow users to input such values at runtime (can we say why?)
-As noted in prior works ([19], [27], [32]), purely Boolean circuits have reasonable performance when dealing with small input sizes (e.g.
-Groce et al.); even a simple task such as matrix multiplication can be infeasible to run using garbled circuits when the matrix dimensions are large (e.g.
-in our machine learning applications).
-In fact, for these applications, the Yao’s circuit generated would not fit in memory.
-Additionally, the work of Groce et al.
-only compares online times with prior work and do not include pre-processing time.
-We remark that these benchmarks are too small for any appreciable differences in performance between various protocols and we only included them to compare our work with other secure ML classification works.
-However, we thank the reviewer for the pointer to Groce et al.
-and shall include it in the paper.
-Another point to be noted is that our framework can be modified to incorporate further optimized backends (for Boolean, e.g.
-Groce et al.) and for Arithmetic.
-We do not claim that our compiler choices are optimal.
-Multiplication is always done using arithmetic secret sharing as the cost of matrix multiplication in Boolean shares is much higher; addition is done either over Boolean or Arithmetic shares depending on prior shares that are available and not stale.
-Obtaining a compiler with better choices is an interesting optimization problem that could potentially consider other parameters such as network latency, bandwidth, compute resources and so on.
-We leave this exploration to future work.
-Secure code partitioning is not a new technique by itself.
-We only claim that it provides a generic way to execute large programs without going into the details of underlying individual secure protocols.
-We implement it only in the arithmetic setting currently as share reconstruction is simple 32-bit addition, which has roughly the same cost in both the Boolean and arithmetic settings.
-Our new language is superior to other prior domain-specific language in that it is the only language to allow compilation into a backend that allows mixed arithmetic and Boolean computation.
-The task of figuring out which part of compute to execute using which type of sharing is unique to our compiler.
-We agree with the reviewer that EzPC does not support secret array access and combining the techniques of EzPC and ObliVM to support Boolean, arithmetic and secret array access would be an interesting future direction.
-However, for our benchmarks (e.g.
-ML algorithms such as DNNs/CNNs), the access patterns are uniform and so providing for secret array accesses will not improve performance.
-We agree with the other comments made by the reviewers and will incorporate them accordingly.
-One or two comments have not been addressed and we are slightly over the word limit (500 words).
+
+First, our language does indeed allow providing input that is public,
+by specifying the value to be a constant public value (is this correct
+and can we write the syntax for it?). An example of public input is
+loop counter value. This ensures that such values need not be
+secret-shared leading to high overheads. However, this input is
+hard-coded during compile time and we currently do not allow users to
+input such values at runtime (can we say why?)
+
+[AR]: The reviewer is making a distinction between compile time
+constants and public inputs. For example, a public input might be
+implemented as follows:
+
+let x = input1 () in
+let y = declassify x in
+//now on y is public
+
+We can't support this currently (we had this support in Wysteria
+because we were interpreting the programs).
+
+We can rebut by saying: "We don't support it currently, the parties
+would have to do out-of-band communication, and hardcode them in the
+program. For our ML benchmarks, we haven't needed this support."
+
+As noted in prior works ([19], [27], [32]), purely Boolean circuits
+have reasonable performance when dealing with small input sizes
+(e.g. Groce et al.); even a simple task such as matrix multiplication
+can be infeasible to run using garbled circuits when the matrix
+dimensions are large (e.g. in our machine learning applications). In
+fact, for these applications, the Yao’s circuit generated would not
+fit in memory. Additionally, the work of Groce et al. only compares
+online times with prior work and do not include pre-processing
+time. We remark that these benchmarks are too small for any
+appreciable differences in performance between various protocols and
+we only included them to compare our work with other secure ML
+classification works. However, we thank the reviewer for the pointer
+to Groce et al. and shall include it in the paper. Another point to be
+noted is that our framework can be modified to incorporate further
+optimized backends (for Boolean, e.g. Groce et al.) and for
+Arithmetic.
+
+[AR]: I don't understand this well. Is this trying to say that Groce
+et al.'s benchmarks are small, and so their pure Yao approach works
+well, and for large benchmarks it would be terrible? The reviewer also
+says that our benchmarks are simple. Can we be more forceful if indeed
+Groce benchmarks are smaller than ours?
+
+We do not claim that our compiler choices are optimal. Multiplication
+is always done using arithmetic secret sharing as the cost of matrix
+multiplication in Boolean shares is much higher; addition is done
+either over Boolean or Arithmetic shares depending on prior shares
+that are available and not stale. Obtaining a compiler with better
+choices is an interesting optimization problem that could potentially
+consider other parameters such as network latency, bandwidth, compute
+resources and so on. We leave this exploration to future work.
+
+Secure code partitioning is not a new technique by itself. We only
+claim that it provides a generic way to execute large programs without
+going into the details of underlying individual secure protocols. We
+implement it only in the arithmetic setting currently as share
+reconstruction is simple 32-bit addition, which has roughly the same
+cost in both the Boolean and arithmetic settings.
+
+Our new language is superior to other prior domain-specific language
+in that it is the only language to allow compilation into a backend
+that allows mixed arithmetic and Boolean computation. The task of
+figuring out which part of compute to execute using which type of
+sharing is unique to our compiler.
+
+We agree with the reviewer that EzPC does not support secret array
+access and combining the techniques of EzPC and ObliVM to support
+Boolean, arithmetic and secret array access would be an interesting
+future direction. However, for our benchmarks (e.g. ML algorithms such
+as DNNs/CNNs), the access patterns are uniform and so providing for
+secret array accesses will not improve performance.
+We agree with the other comments made by the reviewers and will
+incorporate them accordingly. One or two comments have not been
+addressed and we are slightly over the word limit (500 words).

### 2018-03-21 usenix18/rebuttal.txt
tweaks
DG comments

@@
+[AR]: What is the high-level point of the rebuttal? Worth thinking about that.
+
+[DG]: I agree with Aseem's comment above.
+
+[DG] Regarding the Rev #1, I feel that he is just not convinced that we need a mix of arithmetic and boolean representations for efficiency reasons.
+He seems to come from Yao garbled circuits club and believes that with enough optimizations Yao can work best in all contexts.
+And we need to contest this belief.
+For this, we should give citations of all papers that argued a mix of arithmetic and boolean starting with line of work of HE+Yao, ABY, SecureML, MiniONN, Gazelle and more.
+We should point out that even specialized protocols for NNs and other big tasks did indeed use mix of A and B.
+And this is critical.
+It has been observed by various works that doing large number of multiplications using Yao is terrible (huge size of boolean circuit resulting in big communication complexity etc).
+
+[DG] Then regarding Groce et al.
+we can write that it does better than standard Yao by pushing more stuff from online to offline phase.
+It does no improvement in overall time of a GC based computation.
+And just like all other works, we are intersted in overall end-to-end time and not just online time unlike Groce et al.
+
 We thank the reviewers for their useful comments. We address a few of them below.
@@
 
+[DG] the reason for above is the following.
+Given the ABY code, the circuit generation phase happens on both client and server independently and then ABY compares these circuits for equality.
+If we allowed dynamic public inputs at circuit generation step, this public value would have to be communicated to the other party.
+Currently, it is not supported directly in ABY.
+There is no fundamental reason for this, just that they didnt support this.
+
 [AR]: The reviewer is making a distinction between compile time
-constants and public inputs. For example, a public input might be
+constants and public inputs.
+
+[DG] I agree.
+
+[AR] For example, a public input might be
 implemented as follows:
@@
 
+[DG] This can be one way to implement public values by declassifying it.
+
 We can rebut by saying: "We don't support it currently, the parties
@@
 program. For our ML benchmarks, we haven't needed this support."
+
+[DG] I agree.
 
@@
 fact, for these applications, the Yao’s circuit generated would not
-fit in memory. Additionally, the work of Groce et al. only compares
+fit in memory.
+
+[DG] Above might not be valid argument with Yao pipelining technique. It was just that ABY does not have Yao pipelining supported. I have some suggestions on how to address this above.
+
+Additionally, the work of Groce et al. only compares
 online times with prior work and do not include pre-processing
@@
 
+[DG] We can certainly quantify this.
+For example, Groce et al.
+runs experiments for linear classifier or hyperplane based thing where the length of inner product is 30 or 47.
+Compare this with simplest model for MNIST which is logistic regression, where we need to do an inner product of length 784.
+Also, we can look at numofgates in Tables of experiments, and it is clear that for DNNs, CNNs and matrix factorization, number of gates is just huge.
+Also, here we can note that for DNNs this number of gates is for mix of A and B circuit.
+If we write size of just boolean circuit, it would be order of magnitude bigger.
+Some of these arguments can be used to justify that our benchmarks are much larger and no one has done them earlier.
+
 We do not claim that our compiler choices are optimal. Multiplication
@@
 resources and so on. We leave this exploration to future work.
+
+[DG] We should quantify by saying that for 64-bit integer multiplication, using A shares, we need to do 2 multiplications of uint64.
+Over Yao, the circuit size would 64*64 and would require at least 64*64 AES operations in online phase.
 
@@
 
+[DG] May be say something like "it is the first framework where the compiler chooses between different protocols based on cryptographic costs.
+All earlier DSL's in this space only had a single back-end and there was no choice to be made."
+I think this is a major contribution on the language and compiler front.
+
 We agree with the reviewer that EzPC does not support secret array
@@
 future direction. However, for our benchmarks (e.g. ML algorithms such
-as DNNs/CNNs), the access patterns are uniform and so providing for
+as logistic regresion, linear classifier, DNNs/CNNs), the access patterns are uniform and so providing for
 secret array accesses will not improve performance.

### 2018-03-22 usenix18/rebuttal.txt
new draft incorporating comments

@@
 [DG]: I agree with Aseem's comment above.
+
+[NC]: I think it should be to address Rev 1's main concern that garbled circuit implementations alone might suffice as Divya says below.
 
@@
 addressed and we are slightly over the word limit (500 words).
+
+NEW DRAFT BELOW
+----------------
+
+We thank the reviewers for their useful comments. We address a few of them.
+
+First, our language allows providing public input, by specifying the value to be a compile-time constant value.
+An example is loop counter value.
+However, we do not currently support run-time public values to be input by parties (primarily because the underlying ABY backend does not provide support for this).
+Such values would have to be communicated out-of-band to the other party, and hardcoded into the program.
+We remark that for our ML benchmarks, we haven't needed this support but acknowledge that this would be a useful feature to integrate in the future.
+
+A long sequence of works ([18], [26], [31], [40], [37]) have provided evidence that pure Boolean circuit computation (such as Yao's garbled circuit) does not provide good efficiency, especially when running large-scale arithmetic computation.
+Indeed, even specialized protocols for executing Neural Networks and other more sophisticated ML tasks ([40], [37]) did use a (manual) mix of Arithmetic and Boolean computation for efficiency and remark that doing large number of multiplications using Yao is terrible (due to the large size of the Boolean circuit resulting in high communication).
+
+We thank the reviewer for pointing us to Groce et al.
+that improves on native Yao's garbled circuits by moving more computation to the offline (input independent) phase, and thereby achieving lower timings for the online (input dependent) phase.
+The overall times remain roughly the same.
+As in many other works, we are interested in the overall execution times of our protocols.
+Additionally, the benchmarks provided in Groce et al.
+(and also Bost et al.) are small "warm-up" benchmarks that we provide in our paper for completeness (vector dimensions are 30 and 47).
+The full power of our techniques can be best seen in the ease of running large benchmarks such as Tensorflow neural networks (this is one of our simplest large benchmarks, for which no prior secure computation protocols even exist); the base inner product for this compute is over vectors of length 784.
+The Boolean circuit for this inner product itself would roughly consist of 800k gates (whereas the number of gates in the hybrid approach for the entire computation is only 35k).
+The work of Groce et al.
+can potentially be used as a backend for the Boolean part of the compute in our protocols - we leave this exploration to future work.
+
+We do not claim that our compiler choices are optimal.
+Multiplication is always done using arithmetic secret sharing as the cost of matrix multiplication in Boolean shares is much higher (for 64-bit integer multiplication, using arithmetic shares, we need to do 2 multiplications of uint64.
+Over Yao, the circuit size would be 64*64 and would require at least 4096 AES operations in the online phase.).
+Addition is done either over Boolean or Arithmetic shares depending on prior shares that are available and not stale.
+Obtaining a compiler with better choices is an interesting optimization problem that could potentially consider other parameters such as network latency, bandwidth, and compute resources.
+We leave this exploration to future work.
+
+Secure code partitioning is not a new technique by itself.
+We only claim that it provides a generic way to execute large programs without going into the details of underlying individual secure protocols.
+We implement it only in the arithmetic setting currently as share reconstruction is simple addition, which has roughly the same cost in both the Boolean and arithmetic settings.
+
+Our new language is superior to other prior domain-specific languages in that it is the only language to allow compilation into a backend that allows mixed arithmetic and Boolean computation.
+Ours is the first framework where the compiler chooses between different protocols based on cryptographic costs.
+All earlier DSL's in this space only had a single back-end and there was no choice to be made.
+
+We agree with the reviewer that EzPC does not support secret array access and combining the techniques of EzPC and ObliVM to support Boolean, arithmetic and secret array access would be an interesting future direction.
+However, for our benchmarks (e.g.
+ML algorithms such as logistic regression, linear classifier, DNNs/CNNs), the access patterns are uniform and so providing for secret array accesses will not improve performance.
+
+We agree with the other comments made by the reviewers and will incorporate them accordingly.
+
+-----------------------------------------

### 2018-03-23 usenix18/rebuttal.txt
work

@@
+We thank the reviewers for their comments. We structure the rebuttal
+to address high-level concerns in the first ~500 words, addressing
+remaining concerns in a detailed section.
+
+Comparison with Yao-only approach, Groce et al., and CBMC-GC etc.:
+
+We are not the first to observe that the Yao-only approach does not
+scale well beyond small examples ([18], [26], [31], [40],
+[37]), especially for large-scale arithmetic computations commonly
+found in Machine Learning (ML) algorithms -- the main use case of
+EzPC. Indeed, this observation is one of the main driver behind
+development of specialized MPC protocols in previous work ([40],
+[37]). Therefore, it is critical to combine the  arithmetic and
+boolean circuits for the kind of ML algorithms we apply EzPC to.
+
+We thank Reviewer A for the Groce et al. reference. We present a
+thorough comparison in the detailed section below. At a high-level,
+the two overlapping benchmarks in Groce et al. are quite small, and
+so, it is not surprising that the Yao-only approach works well. It is
+not clear that Groce et al.'s approach would scale to large EzPC
+examples (CIFAR 10 or matrix factorization). Furthermore, Groce et
+al. improve the "online" protocol time at the cost of more "offline"
+computation, while we are interested in minimizing the overall time
+(more below).
+
+As for comparing with other compilers, we compare EzPC with the best
+known previous work, except for the benchmarks such as Bonsai, where
+EzPC is the first MPC implementation. Compilers such as
+CBMC-GC have not been used at the scale of benchmarks that we have
+presented in the paper.
+
+Performance gains:
+
+For large benchmarks, such as CIFAR 10 (Table 7) and Matrix
+Factorization (Table 8), EzPC performs much better than the previous
+best known algorithms (~2x and ~19x resp.). Unsurprisingly, for small
+benchmarks in Tables 1-4, where the baseline time itself is only
+~5seconds, the performance gains of EzPC do not seem as impressive.
+
+Language features:
+
+We agree with the reviewers that the EzPC language itself is quite
+simple, and currently does not support features such as secret
+indices, and public inputs. However, the EzPC design has been driven
+by machine learning, specifically secure prediction case studies. In
+our experience, we have found that even with such a simple language,
+we can implement a lot of such examples.
+
+We also remark that while the EzPC language is simple, its compiler is
+the first compiler that generates a mix of arithmetic and boolean
+circuits, while hiding all the low-level crypto details from the
+programmer.
+
+Cost aware compiler:
+
+As mentioned in Sec. 5, our compiler currently applies certain
+heuristics to choose arithmetic or boolean circuit for different parts
+of the program. Even with such simple heuristics, we are able to
+achieve the demonstrated efficiency and scalability. A more precise
+cost model that factors in the cost of network, input sizes, etc. is
+the immediate future work for us.
+
+==============================================
+
+Detailed responses:
+
+Comparison with Groce et al.:
+
+Groce et al. improve on native Yao's garbled circuits by moving
+more computation to the offline (input independent) phase, and thereby
+achieving lower timings for the online (input dependent) phase. The
+overall times remain roughly the same. As in many other works, we are
+interested in the overall execution times of our
+protocols -- not just the online phase. Additionally, the benchmarks
+provided in Groce et al. (and also Bost et al.) are small "warm-up"
+benchmarks that we provide in our paper for completeness (vector
+dimensions are 30 and 47). The full power of our techniques can be
+best seen in the ease of running large benchmarks such as Tensorflow
+neural networks (this is one of our simplest large benchmarks, for
+which no prior secure computation protocols even exist); the base
+inner product for this compute is over vectors of length 784. The
+Boolean circuit for this inner product itself would roughly consist of
+800k gates (whereas the number of gates in the hybrid approach for the
+entire computation is only 35k). The work of Groce et al. can
+potentially be used as a backend  for the Boolean part of the compute
+in our protocols - we leave this exploration to future work.
+
+Secure code partitioning:
+
+Secure code partitioning is not a new technique by itself. We only
+claim that it provides a generic way to execute large programs without
+going into the details of underlying individual secure protocols. We
+implement it only in the arithmetic setting currently as share
+reconstruction is simple addition, which has roughly the same cost in
+both the Boolean and arithmetic settings.
+
+Secure array accesses:
+
+Combining the techniques of EzPC and ObliVM to support
+Boolean, arithmetic and secret array access would be an interesting
+future direction. However, for ML algorithms such as logistic
+regression, linear classifier, DNNs/CNNs, the access patterns are
+uniform and so providing for secret array accesses will not improve
+performance.
+
+Integrating in an existing compiler:
+
+We found that we can write many secure prediction algorithms
+in an EzPC-like simple language, without pulling in complexities of
+other languages and systems. When we add more features to EzPC, such
+as ORAM support, we could definitely consider the option of
+integrating with some other compiler.
+
+========== END ==========
+
 [AR]: What is the high-level point of the rebuttal? Worth thinking about that.
@@
 
-A long sequence of works ([18], [26], [31], [40], [37]) have provided evidence that pure Boolean circuit computation (such as Yao's garbled circuit) does not provide good efficiency, especially when running large-scale arithmetic computation.
-Indeed, even specialized protocols for executing Neural Networks and other more sophisticated ML tasks ([40], [37]) did use a (manual) mix of Arithmetic and Boolean computation for efficiency and remark that doing large number of multiplications using Yao is terrible (due to the large size of the Boolean circuit resulting in high communication).
+A long sequence of works  have provided evidence that pure Boolean circuit computation (such as Yao's garbled circuit) does not provide good efficiency, especially when running .
+Indeed, even specialized protocols for executing Neural Networks and other more sophisticated ML tasks  did use a (manual) mix of Arithmetic and Boolean computation for efficiency and remark that doing large number of multiplications using Yao is terrible (due to the large size of the Boolean circuit resulting in high communication).
 
-We thank the reviewer for pointing us to Groce et al.
-that improves on native Yao's garbled circuits by moving more computation to the offline (input independent) phase, and thereby achieving lower timings for the online (input dependent) phase.
-The overall times remain roughly the same.
-As in many other works, we are interested in the overall execution times of our protocols.
-Additionally, the benchmarks provided in Groce et al.
-(and also Bost et al.) are small "warm-up" benchmarks that we provide in our paper for completeness (vector dimensions are 30 and 47).
-The full power of our techniques can be best seen in the ease of running large benchmarks such as Tensorflow neural networks (this is one of our simplest large benchmarks, for which no prior secure computation protocols even exist); the base inner product for this compute is over vectors of length 784.
-The Boolean circuit for this inner product itself would roughly consist of 800k gates (whereas the number of gates in the hybrid approach for the entire computation is only 35k).
-The work of Groce et al.
-can potentially be used as a backend for the Boolean part of the compute in our protocols - we leave this exploration to future work.
+We thank the reviewer for pointing us to
 
@@
 
-Secure code partitioning is not a new technique by itself.
-We only claim that it provides a generic way to execute large programs without going into the details of underlying individual secure protocols.
-We implement it only in the arithmetic setting currently as share reconstruction is simple addition, which has roughly the same cost in both the Boolean and arithmetic settings.
-
 Our new language is superior to other prior domain-specific languages in that it is the only language to allow compilation into a backend that allows mixed arithmetic and Boolean computation.
@@
 
-We agree with the reviewer that EzPC does not support secret array access and combining the techniques of EzPC and ObliVM to support Boolean, arithmetic and secret array access would be an interesting future direction.
-However, for our benchmarks (e.g.
-ML algorithms such as logistic regression, linear classifier, DNNs/CNNs), the access patterns are uniform and so providing for secret array accesses will not improve performance.
-
 We agree with the other comments made by the reviewers and will incorporate them accordingly.

### 2018-03-23 usenix18/rebuttal.txt
tweaks

@@
 We are not the first to observe that the Yao-only approach does not
-scale well beyond small examples ([18], [26], [31], [40],
-[37]), especially for large-scale arithmetic computations commonly
-found in Machine Learning (ML) algorithms -- the main use case of
-EzPC. Indeed, this observation is one of the main driver behind
-development of specialized MPC protocols in previous work ([40],
-[37]). Therefore, it is critical to combine the  arithmetic and
-boolean circuits for the kind of ML algorithms we apply EzPC to.
+scale well beyond examples with small number of multiplications ([18],
+[26], [31], [40], [37]), especially for large-scale arithmetic
+computations commonly found in Machine Learning (ML) algorithms -- the
+main use case of EzPC. Indeed, this observation is one of the main
+drivers behind development of specialized MPC protocols in previous
+work ([40], [37]). Therefore, it is critical to combine the
+arithmetic and boolean circuits for the kind of ML algorithms we apply
+EzPC to.
 
@@
 not clear that Groce et al.'s approach would scale to large EzPC
-examples (CIFAR 10 or matrix factorization). Furthermore, Groce et
-al. improve the "online" protocol time at the cost of more "offline"
-computation, while we are interested in minimizing the overall time
-(more below).
+examples (DNNs or matrix factorization). Furthermore, our numbers
+cannot be compared directly with Groce et al. since they report only
+the "online" time of the protocols, while we report the sum of online
+and "offline" times (more below).
 
-As for comparing with other compilers, we compare EzPC with the best
-known previous work, except for the benchmarks such as Bonsai, where
-EzPC is the first MPC implementation. Compilers such as
-CBMC-GC have not been used at the scale of benchmarks that we have
-presented in the paper.
+As for comparing with other compilers, we go a step further and
+compare EzPC with the best known previous specialized protocols,
+except for the benchmarks such as Bonsai, where EzPC is the first MPC
+implementation. Compilers such as CBMC-GC have not been used at the
+scale of benchmarks that we have presented in the paper.
 
@@
 benchmarks in Tables 1-4, where the baseline time itself is only
-~5seconds, the performance gains of EzPC do not seem as impressive.
+~5seconds, the performance gains of EzPC do not seem as
+impressive. TODO: add a line about specialized vs generic.
 
@@
 achieve the demonstrated efficiency and scalability. A more precise
-cost model that factors in the cost of network, input sizes, etc. is
-the immediate future work for us.
+cost model that factors in the cost of network, input sizes,
+etc. would only improve performance (more details below).
 
@@
 in our protocols - we leave this exploration to future work.
+
+Cost aware compiler:
+
+We do not claim that our compiler choices are optimal. Multiplication
+is always done using arithmetic secret sharing as the cost of matrix
+multiplication in Boolean shares is much higher (for 64-bit integer
+multiplication, using arithmetic shares, we need to do 2
+multiplications of uint64. Over Yao, the circuit size would be 64*64
+and would require at least 4096 AES operations in the online
+phase.). Addition is done either over Boolean or Arithmetic shares
+depending on prior shares that are available and not stale.
 
@@
 
-We do not claim that our compiler choices are optimal.
-Multiplication is always done using arithmetic secret sharing as the cost of matrix multiplication in Boolean shares is much higher (for 64-bit integer multiplication, using arithmetic shares, we need to do 2 multiplications of uint64.
-Over Yao, the circuit size would be 64*64 and would require at least 4096 AES operations in the online phase.).
-Addition is done either over Boolean or Arithmetic shares depending on prior shares that are available and not stale.
-Obtaining a compiler with better choices is an interesting optimization problem that could potentially consider other parameters such as network latency, bandwidth, and compute resources.
-We leave this exploration to future work.
-
 Our new language is superior to other prior domain-specific languages in that it is the only language to allow compilation into a backend that allows mixed arithmetic and Boolean computation.

### 2018-03-23 usenix18/rebuttal.txt
finalizing

@@
 We thank the reviewers for their comments. We structure the rebuttal
-to address high-level concerns in the first ~500 words, addressing
-remaining concerns in a detailed section.
+to address main concerns in the first ~500 words, leaving the rest to
+a detailed section.
 
-Comparison with Yao-only approach, Groce et al., and CBMC-GC etc.:
+Comparison with Yao-only, Groce et al., CBMC-GC etc.:
 
@@
 main use case of EzPC. Indeed, this observation is one of the main
-drivers behind development of specialized MPC protocols in previous
+drivers behind development of the specialized MPC protocols in previous
 work ([40], [37]). Therefore, it is critical to combine the
@@
 
-We thank Reviewer A for the Groce et al. reference. We present a
-thorough comparison in the detailed section below. At a high-level,
-the two overlapping benchmarks in Groce et al. are quite small, and
-so, it is not surprising that the Yao-only approach works well. It is
-not clear that Groce et al.'s approach would scale to large EzPC
-examples (DNNs or matrix factorization). Furthermore, our numbers
-cannot be compared directly with Groce et al. since they report only
-the "online" time of the protocols, while we report the sum of online
-and "offline" times (more below).
+We thank Reviewer A for the Groce et al. reference. Firstly, our
+numbers  cannot be compared directly with Groce et al. since they
+report only the "online" time of the protocols, while we report the
+sum of online and "offline" times. Further, the two overlapping
+benchmarks in Groce et al. are quite small, and so, it is not
+surprising that the Yao-only approach works well. It is not clear that
+Groce et al.'s approach would scale to larger EzPC examples, e.g. DNNs
+or matrix factorization (more thorough comparison below).
 
-As for comparing with other compilers, we go a step further and
-compare EzPC with the best known previous specialized protocols,
-except for the benchmarks such as Bonsai, where EzPC is the first MPC
+As for comparing with the other compilers, we go a step further and
+compare with the best known previous specialized protocols,
+except for benchmarks such as Bonsai, where ours is the first MPC
 implementation. Compilers such as CBMC-GC have not been used at the
@@
 
-Performance gains:
+Performance gains not being impressive:
 
@@
 ~5seconds, the performance gains of EzPC do not seem as
-impressive. TODO: add a line about specialized vs generic.
+impressive. But even then we would like to point out that while most
+of the previous work on those benchmarks developed specialized
+protocols, EzPC framework is general and programmer friendly.
 
@@
 
-We agree with the reviewers that the EzPC language itself is quite
-simple, and currently does not support features such as secret
-indices, and public inputs. However, the EzPC design has been driven
-by machine learning, specifically secure prediction case studies. In
-our experience, we have found that even with such a simple language,
-we can implement a lot of such examples.
+EzPC language itself is quite simple and currently does not support
+features such as secret indices and public inputs. However, the
+language design has been driven by machine learning, specifically
+secure prediction, case studies. As we show, even with such a simple
+language, we can implement a lot of such examples.
 
-We also remark that while the EzPC language is simple, its compiler is
-the first compiler that generates a mix of arithmetic and boolean
+We also remark that while the source language is simple, EzPC compiler
+is the first compiler that generates a mix of arithmetic and boolean
 circuits, while hiding all the low-level crypto details from the
@@
 
-Cost aware compiler:
+Cost-awareness
 
-As mentioned in Sec. 5, our compiler currently applies certain
-heuristics to choose arithmetic or boolean circuit for different parts
-of the program. Even with such simple heuristics, we are able to
-achieve the demonstrated efficiency and scalability. A more precise
-cost model that factors in the cost of network, input sizes,
-etc. would only improve performance (more details below).
+Our compiler currently applies certain heuristics to choose arithmetic
+or boolean circuit for different parts of the program (Sec. 5). Even
+with such simple heuristics, we are able to achieve the demonstrated
+efficiency and scalability. A more precise cost model that factors in
+the cost of network, input sizes, etc. would only improve performance
+(more on it below).
 
@@
 integrating with some other compiler.
-
-========== END ==========
-
-[AR]: What is the high-level point of the rebuttal? Worth thinking about that.
-
-[DG]: I agree with Aseem's comment above.
-
-[NC]: I think it should be to address Rev 1's main concern that garbled circuit implementations alone might suffice as Divya says below.
-
-[DG] Regarding the Rev #1, I feel that he is just not convinced that we need a mix of arithmetic and boolean representations for efficiency reasons.
-He seems to come from Yao garbled circuits club and believes that with enough optimizations Yao can work best in all contexts.
-And we need to contest this belief.
-For this, we should give citations of all papers that argued a mix of arithmetic and boolean starting with line of work of HE+Yao, ABY, SecureML, MiniONN, Gazelle and more.
-We should point out that even specialized protocols for NNs and other big tasks did indeed use mix of A and B.
-And this is critical.
-It has been observed by various works that doing large number of multiplications using Yao is terrible (huge size of boolean circuit resulting in big communication complexity etc).
-
-[DG] Then regarding Groce et al.
-we can write that it does better than standard Yao by pushing more stuff from online to offline phase.
-It does no improvement in overall time of a GC based computation.
-And just like all other works, we are intersted in overall end-to-end time and not just online time unlike Groce et al.
-
-We thank the reviewers for their useful comments. We address a few of them below.
-
-First, our language does indeed allow providing input that is public,
-by specifying the value to be a constant public value (is this correct
-and can we write the syntax for it?). An example of public input is
-loop counter value. This ensures that such values need not be
-secret-shared leading to high overheads. However, this input is
-hard-coded during compile time and we currently do not allow users to
-input such values at runtime (can we say why?)
-
-[DG] the reason for above is the following.
-Given the ABY code, the circuit generation phase happens on both client and server independently and then ABY compares these circuits for equality.
-If we allowed dynamic public inputs at circuit generation step, this public value would have to be communicated to the other party.
-Currently, it is not supported directly in ABY.
-There is no fundamental reason for this, just that they didnt support this.
-
-[AR]: The reviewer is making a distinction between compile time
-constants and public inputs.
-
-[DG] I agree.
-
-[AR] For example, a public input might be
-implemented as follows:
-
-let x = input1 () in
-let y = declassify x in
-//now on y is public
-
-We can't support this currently (we had this support in Wysteria
-because we were interpreting the programs).
-
-[DG] This can be one way to implement public values by declassifying it.
-
-We can rebut by saying: "We don't support it currently, the parties
-would have to do out-of-band communication, and hardcode them in the
-program. For our ML benchmarks, we haven't needed this support."
-
-[DG] I agree.
-
-As noted in prior works ([19], [27], [32]), purely Boolean circuits
-have reasonable performance when dealing with small input sizes
-(e.g. Groce et al.); even a simple task such as matrix multiplication
-can be infeasible to run using garbled circuits when the matrix
-dimensions are large (e.g. in our machine learning applications). In
-fact, for these applications, the Yao’s circuit generated would not
-fit in memory.
-
-[DG] Above might not be valid argument with Yao pipelining technique. It was just that ABY does not have Yao pipelining supported. I have some suggestions on how to address this above.
-
-Additionally, the work of Groce et al. only compares
-online times with prior work and do not include pre-processing
-time. We remark that these benchmarks are too small for any
-appreciable differences in performance between various protocols and
-we only included them to compare our work with other secure ML
-classification works. However, we thank the reviewer for the pointer
-to Groce et al. and shall include it in the paper. Another point to be
-noted is that our framework can be modified to incorporate further
-optimized backends (for Boolean, e.g. Groce et al.) and for
-Arithmetic.
-
-[AR]: I don't understand this well. Is this trying to say that Groce
-et al.'s benchmarks are small, and so their pure Yao approach works
-well, and for large benchmarks it would be terrible? The reviewer also
-says that our benchmarks are simple. Can we be more forceful if indeed
-Groce benchmarks are smaller than ours?
-
-[DG] We can certainly quantify this.
-For example, Groce et al.
-runs experiments for linear classifier or hyperplane based thing where the length of inner product is 30 or 47.
-Compare this with simplest model for MNIST which is logistic regression, where we need to do an inner product of length 784.
-Also, we can look at numofgates in Tables of experiments, and it is clear that for DNNs, CNNs and matrix factorization, number of gates is just huge.
-Also, here we can note that for DNNs this number of gates is for mix of A and B circuit.
-If we write size of just boolean circuit, it would be order of magnitude bigger.
-Some of these arguments can be used to justify that our benchmarks are much larger and no one has done them earlier.
-
-We do not claim that our compiler choices are optimal. Multiplication
-is always done using arithmetic secret sharing as the cost of matrix
-multiplication in Boolean shares is much higher; addition is done
-either over Boolean or Arithmetic shares depending on prior shares
-that are available and not stale. Obtaining a compiler with better
-choices is an interesting optimization problem that could potentially
-consider other parameters such as network latency, bandwidth, compute
-resources and so on. We leave this exploration to future work.
-
-[DG] We should quantify by saying that for 64-bit integer multiplication, using A shares, we need to do 2 multiplications of uint64.
-Over Yao, the circuit size would 64*64 and would require at least 64*64 AES operations in online phase.
-
-Secure code partitioning is not a new technique by itself. We only
-claim that it provides a generic way to execute large programs without
-going into the details of underlying individual secure protocols. We
-implement it only in the arithmetic setting currently as share
-reconstruction is simple 32-bit addition, which has roughly the same
-cost in both the Boolean and arithmetic settings.
-
-Our new language is superior to other prior domain-specific language
-in that it is the only language to allow compilation into a backend
-that allows mixed arithmetic and Boolean computation. The task of
-figuring out which part of compute to execute using which type of
-sharing is unique to our compiler.
-
-[DG] May be say something like "it is the first framework where the compiler chooses between different protocols based on cryptographic costs.
-All earlier DSL's in this space only had a single back-end and there was no choice to be made."
-I think this is a major contribution on the language and compiler front.
-
-We agree with the reviewer that EzPC does not support secret array
-access and combining the techniques of EzPC and ObliVM to support
-Boolean, arithmetic and secret array access would be an interesting
-future direction. However, for our benchmarks (e.g. ML algorithms such
-as logistic regresion, linear classifier, DNNs/CNNs), the access patterns are uniform and so providing for
-secret array accesses will not improve performance.
-We agree with the other comments made by the reviewers and will
-incorporate them accordingly. One or two comments have not been
-addressed and we are slightly over the word limit (500 words).
-
-NEW DRAFT BELOW
-----------------
-
-We thank the reviewers for their useful comments. We address a few of them.
-
-First, our language allows providing public input, by specifying the value to be a compile-time constant value.
-An example is loop counter value.
-However, we do not currently support run-time public values to be input by parties (primarily because the underlying ABY backend does not provide support for this).
-Such values would have to be communicated out-of-band to the other party, and hardcoded into the program.
-We remark that for our ML benchmarks, we haven't needed this support but acknowledge that this would be a useful feature to integrate in the future.
-
-A long sequence of works  have provided evidence that pure Boolean circuit computation (such as Yao's garbled circuit) does not provide good efficiency, especially when running .
-Indeed, even specialized protocols for executing Neural Networks and other more sophisticated ML tasks  did use a (manual) mix of Arithmetic and Boolean computation for efficiency and remark that doing large number of multiplications using Yao is terrible (due to the large size of the Boolean circuit resulting in high communication).
-
-We thank the reviewer for pointing us to
-
-Our new language is superior to other prior domain-specific languages in that it is the only language to allow compilation into a backend that allows mixed arithmetic and Boolean computation.
-Ours is the first framework where the compiler chooses between different protocols based on cryptographic costs.
-All earlier DSL's in this space only had a single back-end and there was no choice to be made.
-
-We agree with the other comments made by the reviewers and will incorporate them accordingly.
-
------------------------------------------

### 2018-03-24 usenix18/rebuttal.txt
minor changes

@@
 compare with the best known previous specialized protocols,
-except for benchmarks such as Bonsai, where ours is the first MPC
+except for benchmarks such as Bonsai, where ours is the first 2PC
 implementation. Compilers such as CBMC-GC have not been used at the
@@
 
-Groce et al. improve on native Yao's garbled circuits by moving
-more computation to the offline (input independent) phase, and thereby
-achieving lower timings for the online (input dependent) phase. The
-overall times remain roughly the same. As in many other works, we are
+Groce et al. improves on online time and online communication of native Yao's 
+garbled circuits by moving more computation to the offline (input independent) phase. 
+However, the overall times remain roughly the same. As is standard, we are
 interested in the overall execution times of our
@@
 benchmarks that we provide in our paper for completeness (vector
-dimensions are 30 and 47). The full power of our techniques can be
-best seen in the ease of running large benchmarks such as Tensorflow
-neural networks (this is one of our simplest large benchmarks, for
-which no prior secure computation protocols even exist); the base
-inner product for this compute is over vectors of length 784. The
-Boolean circuit for this inner product itself would roughly consist of
-800k gates (whereas the number of gates in the hybrid approach for the
-entire computation is only 35k). The work of Groce et al. can
-potentially be used as a backend  for the Boolean part of the compute
+dimensions for inner product are 30 and 47).
+The full power of our techniques can be best seen in the ease of 
+running large benchmarks such as logistic regression, and neural networks. 
+These require much larger number of multiplications and no one has 
+attempted to implement these using garbled circuits. For instance, 
+simplest benchmark of logistic regression requires 10 inner products 
+of length 784. The Boolean circuit for one inner product itself would 
+roughly consist of 800k gates (whereas the number of gates in the hybrid
+ approach for the entire computation is only 35k). The work of Groce et al. 
+can potentially be used as a backend for the Boolean part of the compute 
 in our protocols - we leave this exploration to future work.
 
-Cost aware compiler:
+Cryptographic Cost-aware compiler:
 
 We do not claim that our compiler choices are optimal. Multiplication
-is always done using arithmetic secret sharing as the cost of matrix
+is always done using arithmetic secret sharing as the cost of
 multiplication in Boolean shares is much higher (for 64-bit integer
-multiplication, using arithmetic shares, we need to do 2
+multiplication, using arithmetic shares, we need to do only 2
 multiplications of uint64. Over Yao, the circuit size would be 64*64
 and would require at least 4096 AES operations in the online
-phase.). Addition is done either over Boolean or Arithmetic shares
-depending on prior shares that are available and not stale.
+phase.). Addition *does not* have a default way of evaluation and is done 
+either over Boolean or Arithmetic shares depending on available shares of operands.
 
@@
 
-Secure code partitioning is not a new technique by itself. We only
-claim that it provides a generic way to execute large programs without
-going into the details of underlying individual secure protocols. We
-implement it only in the arithmetic setting currently as share
-reconstruction is simple addition, which has roughly the same cost in
-both the Boolean and arithmetic settings.
+Our technique of secure code partitioning, although not entirely new, provides a 
+generic way to execute large programs without going into the specific details of 
+cryptographic protocol. We describe it only in the arithmetic setting but as the
+reviewer correctly points out, same can be done in Boolean setting as well.
 

### 2018-03-24 usenix18/rebuttal.txt
tweaks

@@
 
-Comparison with Yao-only, Groce et al., CBMC-GC etc.:
+Comparison with Groce et al., CBMC-GC, Yao-only etc.:
 
-We are not the first to observe that the Yao-only approach does not
-scale well beyond examples with small number of multiplications ([18],
-[26], [31], [40], [37]), especially for large-scale arithmetic
-computations commonly found in Machine Learning (ML) algorithms -- the
-main use case of EzPC. Indeed, this observation is one of the main
-drivers behind development of the specialized MPC protocols in previous
-work ([40], [37]). Therefore, it is critical to combine the
-arithmetic and boolean circuits for the kind of ML algorithms we apply
-EzPC to.
+We thank Reviewer A for the Groce et al. reference. Our numbers
+should not be directly compared with theirs since they report only the
+"online" time of the protocols, while we report the sum of online and
+"offline" times. Further, the benchmarks in Groce et al. are quite
+small, and so, it is not surprising that the Yao-only approach works
+well. It is unclear that their approach would scale to our larger
+benchmarks, e.g. DNNs or matrix factorization (more thorough
+comparison below).
 
-We thank Reviewer A for the Groce et al. reference. Firstly, our
-numbers  cannot be compared directly with Groce et al. since they
-report only the "online" time of the protocols, while we report the
-sum of online and "offline" times. Further, the two overlapping
-benchmarks in Groce et al. are quite small, and so, it is not
-surprising that the Yao-only approach works well. It is not clear that
-Groce et al.'s approach would scale to larger EzPC examples, e.g. DNNs
-or matrix factorization (more thorough comparison below).
+CBMC-GC takes 30seconds and generates 1.8 million gates to multiply 2
+5x5 matrices. On the other hand, MNIST logistic regression (Table 5 in
+our paper) consists of multiplying 2 192x192 matrices -- and EzPC
+generated protocol takes 0.7 seconds (in WAN) with 35k gates.
 
-As for comparing with the other compilers, we go a step further and
-compare with the best known previous specialized protocols,
-except for benchmarks such as Bonsai, where ours is the first 2PC
-implementation. Compilers such as CBMC-GC have not been used at the
-scale of benchmarks that we have presented in the paper.
+As for the Yao-only approach in general, we are not the first to
+observe that it does not scale well beyond examples with small number
+of multiplications ([18], [26], [31], [40], [37]), especially for
+large-scale arithmetic computations commonly found in Machine Learning
+(ML) algorithms -- the main use case of EzPC. Indeed, this observation
+is one of the main drivers behind development of the specialized MPC
+protocols in previous work ([40], [37]). Therefore, it is critical to
+combine the arithmetic and boolean circuits for the kind of ML
+algorithms we apply EzPC to.
 
@@
 of the previous work on those benchmarks developed specialized
-protocols, EzPC framework is general and programmer friendly.
+protocols, EzPC framework is general, programmer friendly, and does
+not require any crypto expertise.
 

### 2018-03-24 usenix18/rebuttal.txt
minor edits

@@
 
-Performance gains not being impressive:
+Performance gains:
 
@@
 language design has been driven by machine learning, specifically
-secure prediction, case studies. As we show, even with such a simple
+secure prediction case studies. As we show, even with such a simple
 language, we can implement a lot of such examples.
@@
 
-Cost-awareness
+Cost-awareness:
 
 Our compiler currently applies certain heuristics to choose arithmetic
-or boolean circuit for different parts of the program (Sec. 5). Even
+or boolean circuits for different parts of the program (Sec. 5). Even
 with such simple heuristics, we are able to achieve the demonstrated

### 2018-04-27 ccs/conclude.tex
starting ccs repo

%% \section{Conclusion}
%% \label{sec:conclude}
%% We presented \tool, the first
%% cryptographic-cost aware framework that generates efficient and
%% scalable \mpc protocols from high-level programs.
%% The generated protocols comprise combinations
%% of arithmetic and boolean circuits and have performance comparable to, or better than the custom, specialized
%% protocols from previous works.
%% The \mpc backend we use provides security against semi-honest or passive adversaries.
%% %our compiler is oblivious to the ``type of adversary'' for the underlying \mpc backend.
%% Given a \mpc backend secure against malicious or active adversaries, Theorem \ref{theorem:security} 
%% can be extended to obtain a framework with malicious security. Furthermore,  
%% partitioning can be modified to provide malicious security by {\it signing} the shares of the intermediate states. However, we are not aware of efficient maliciously secure \mpc implementation for combinations of
%% arithmetic and boolean circuits and hence leave a complete treatment of malicious security to future work.
%% % Several areas of improvement remain: a) While the backend that we use~\cite{aby} only provides security against a semi-honest adversary, our compiler is oblivious to the ``type of adversary'' for the underlying backend cryptographic protocol. In particular, if our cryptographic backend were maliciously secure, then Theorem \ref{theorem:security} would provide us a framework with malicious security. Furthermore, modifying our code partitioning to provide malicious security is fairly easy (one would instead provide {\em signed} shares of intermediate states in the decomposed programs to obtain malicious security). b) Our compiler converts programs into circuits before executing a \mpc and is efficient for the applications we present. However, one could consider executing programs that exploit secret-dependent memory accesses (such as binary search) to obtain better performance (as in \cite{oblivm}). It would be interesting to see if our techniques can be combined with that of \cite{oblivm}. \nc{Too much? Should we simply say only about the malicious security and stop?}

### 2018-04-27 ccs/intro.tex
starting ccs repo

--- 2018-02-07 usenix18/intro.tex
@@
 behind the design of specialized protocols. Using \tool, we empirically  demonstrate
-the that generic \mpc implementations are much more efficient than what they were believed to be. 
+that generic \mpc implementations are much more efficient than what they were believed to be. 
  %(up to a factor of 19x in our
@@
 the security theorem, we formally reduce the security
-of our scheme against semi-honest (or, ``honest but curious") adversaries
- to the semi honest security of the  \mpc back-end. 
- We also prove a formal security theorem against semi-honest  adversaries for our partitioning
+of our scheme against semi-honest (or, ``honest but curious")
+adversaries to the semi honest security of the  \mpc back-end. The
+theorem provides protection against side-channels arising from
+conditionals and memory access patterns.
+ We also prove a formal security theorem against semi-honest
+ adversaries for our partitioning
 scheme (\sectionref{ld} and~\sectionref{pipe}).

### 2018-04-27 ccs/main.tex
starting ccs repo

--- 2017-11-01 OaklandSubmission/main.tex
@@
-\documentclass[conference,compsoc]{IEEEtran}
+\documentclass[sigconf, anonymous]{acmart}
+
+\fancyhf{} % Remove fancy page headers 
+\fancyhead[C]{Anonymous submission \#9999 to ACM CCS 2017} % TODO: replace 9999 with your paper number
+\fancyfoot[C]{\thepage}
+
+\setcopyright{none} % No copyright notice required for submissions
+\acmConference[Anonymous Submission to ACM CCS 2017]{ACM Conference on Computer and Communications Security}{Due 19 May 2017}{Dallas, Texas}
+\acmYear{2017}
+
+\settopmatter{printacmref=false, printccs=true, printfolios=true} % We want page numbers on submissions
+
+%%\ccsPaper{9999} % TODO: replace with your paper number once obtained
 
@@
 
+%\usepackage{epsfig,endnotes,pdfpages}
+\usepackage{wrapfig}
+\usepackage{float}
 \begin{document}
-\title{\tool: Programmable, Efficient, and Scalable Secure Two-Party Computation}
 
-% author names and affiliations
-% use a multiple column layout for up to three different
-% affiliations
-% \author{\IEEEauthorblockN{Nishanth Chandran \hspace{10pt} Divya Gupta  \hspace{10pt} Aseem Rastogi \hspace{10pt}  Rahul Sharma \hspace{10pt} Shardul Tripathi }
-% \IEEEauthorblockA{Microsoft Research India}
-%}
+\title{\tool: Programmable, Efficient, and Scalable \\ Secure Two-Party Computation}
 
-\maketitle
 \begin{abstract}
-  %% Secure two-party computation (\mpc) allows two mutually distrusting
-  %% parties to compute any function on their joint inputs, without
-  %% revealing anything other than the output of the function to one
-  %% another.
-
-  %While the application and potential impact of secure computation is virtually limitless, writing an efficient and scalable secure protocol is by no means an easy feat.
-% Prior work in designing such protocols, suffered from either a) a lack of programmability (e.g.
-requiring the programmer to write cryptography-aware circuits for the computation); b) a lack of efficiency for several natural computations (such as machine learning classification tasks); or c) required cryptographers to design hand-crafted protocols for various applications.
 We present \tool: a 
@@
 functionalities such as secure  prediction  and matrix factorization.
-%\tool provides formal correctness and security guarantees.
-%% (up to
-%% a 19x factor!).
+
 \end{abstract}
-\IEEEpeerreviewmaketitle
+
+\setlength\intextsep{1pt}
+\setlength{\textfloatsep}{1pt}
+\setlength{\abovecaptionskip}{4pt}
+\setlength{\belowcaptionskip}{4pt}
+
+\maketitle
 
@@
 %\newpage
-\bibliographystyle{abbrv}
-\bibliography{bib}
+
+{\footnotesize \bibliographystyle{acm}
+\bibliography{bib}}
+
+%\bibliographystyle{abbrv}
+%\bibliography{bib}
 %\balance
 
-%\appendix
-%\input{app-stash}
+\appendix
+
+\input{appendix_b}
+
+\input{appendix_c}
+
+\input{appendix_a}
 

### 2018-04-27 ccs/pipe.tex
starting ccs repo

--- 2018-02-08 usenix18/pipe.tex
@@
 \newcommand{\seq}{||}
-\newcommand{\state}{q}
+\newcommand{\stateq}{q}
 
@@
 $\prog_1\seq\prog_2\seq\dots\seq\prog_k$ be a decomposition of $\prog$
-such that the following holds. Define $\state_0 = \bot$ (the public empty
+such that the following holds. Define $\stateq_0 = \bot$ (the public empty
 state). For all
-$1\leq i\leq k-1$, $\prog_i$ takes inputs $x, y$ and $\state_{i-1}$ and
-outputs state $\state_i$. Finally, $\prog_k$ takes inputs $x, y$ and
-$\state_{k-1}$ to  output $z$. It is
+$1\leq i\leq k-1$, $\prog_i$ takes inputs $x, y$ and $\stateq_{i-1}$ and
+outputs state $\stateq_i$. Finally, $\prog_k$ takes inputs $x, y$ and
+$\stateq_{k-1}$ to  output $z$. It is
 possible to decompose any program $\prog$  into such
@@
 sequentially (in a distributed setting)  to obtain
-$\state_1,\dots,\state_{k-1},$ and finally output $z$. At the
+$\stateq_1,\dots,\stateq_{k-1},$ and finally output $z$. At the
 $i^{\textrm{\tiny{th}}}$ step, the parties only need to store
-information proportional to $x,y,\state_{i-1}$ and $\crct_i$ (which is
+information proportional to $x,y,\stateq_{i-1}$ and $\crct_i$ (which is
 much smaller than $\crct$). However, this execution enables the
-parties to learn $\state_i$ (for all $1\leq i\leq k-1$), which
+parties to learn $\stateq_i$ (for all $1\leq i\leq k-1$), which
 completely breaks the security.
@@
 To overcome this problem, we define a sequence of new programs
-$\progn_i$ ($1\leq i\leq k$) as follows. Once again, define $\state_0
+$\progn_i$ ($1\leq i\leq k$) as follows. Once again, define $\stateq_0
 = \bot$. Without
-loss of generality, let all $\state_i$ be values in some additive ring
+loss of generality, let all $\stateq_i$ be values in some additive ring
 $(\mathbb{Z},+)$ (e.g., the additive ring $(\mathbb{Z}_{2^{64}},+)$,
@@
 be the program that takes as input $x,r_1$ from Alice and $y$ from Bob
-(and empty state $\state_0$), and runs $\prog_1$ (as defined above) to
-compute $\state_1$ and then outputs $o_1 = \state_1 + r_1$ {\em only
+(and empty state $\stateq_0$), and runs $\prog_1$ (as defined above) to
+compute $\stateq_1$ and then outputs $o_1 = \stateq_1 + r_1$ {\em only
   to} Bob\footnote{While the description of our protocol here assumes
@@
 runs $\prog_i$
-on inputs $x,y$ and state $\state_{i-1} = (o_{i-1}-r_{i-1})$ (where $-$ denotes
-subtraction in the ring $(\mathbb{Z},+)$) and then outputs $\state_i+r_i$
+on inputs $x,y$ and state $\stateq_{i-1} = (o_{i-1}-r_{i-1})$ (where $-$ denotes
+subtraction in the ring $(\mathbb{Z},+)$) and then outputs $\stateq_i+r_i$
 to Bob and $r_i$ to Alice. The last
 program $\progn_k$ takes inputs $x,y,r_{k-1},o_{k-1}$, runs $\prog_k$
-on inputs $x,y$ and state $\state_{k-1} = (o_{k-1}-r_{k-1})$ and outputs $z$ to both
+on inputs $x,y$ and state $\stateq_{k-1} = (o_{k-1}-r_{k-1})$ and outputs $z$ to both
 parties.

### 2018-04-27 ccs/related.tex
starting ccs repo

--- 2018-02-07 usenix18/related.tex
@@
-\vspace{-0.1in}
+\vspace{-0.2in}
 \section{Related work}
@@
 %\noindent\textbf{Custom \mpc protocols.} 
-Many works have designed specialized, hand-crafted protocols for various \mpc tasks. 
+Many works have designed specialized protocols for various \mpc tasks. 
 %The reason for doing this was to allow for a mix between arithmetic and boolean computations as a secure computation protocol for a pure boolean circuit would have been very inefficient. 
 This requires deep knowledge of cryptography
-to ensure the security of the protocols.
-Examples include privacy-preserving classification of medical ElectroCardioGram (ECG)~\cite{barni}, iris and fingerprint recognition~\cite{blanton}, remote diagnostics~\cite{brickell}, private sequence analysis~\cite{franz}, private biometric identification~\cite{huang}, privacy-preserving matrix factorization and ridge-regression~\cite{valeriaMatrix, valeriaRidge}, machine-learning classification~\cite{shafindss}, decision trees and forests~\cite{wu}, neural network training~\cite{secureml} and prediction~\cite{minionn}.
+to ensure security.
+Examples include classification of medical ElectroCardioGram (ECG)~\cite{barni}, iris and fingerprint recognition~\cite{blanton}, remote diagnostics~\cite{brickell}, sequence analysis~\cite{franz}, biometric identification~\cite{huang}, matrix factorization and ridge-regression~\cite{valeriaMatrix, valeriaRidge}, machine-learning classification~\cite{shafindss}, decision trees/forests~\cite{wu}, neural network training~\cite{secureml} and prediction~\cite{minionn}.
 % As seen from our experiments, the protocols generated by \tool give us better or comparable execution times as many of the above works with minimal programmer overhead.

### 2018-05-04 ccs/ex.tex
yao comparison

--- 2018-02-08 usenix18/ex.tex
@@
 toolchain. We give a brief overview of each of these phases below.
-\vspace{-0.15in}
+
 \paragraph{Source language.}
@@
 \ls{if}~statements, and special syntax for input/output from each party.
-\vspace{-0.1in}
+
 \paragraph{\tool compiler.}
@@
 a \mpc protocol.
-\vspace{-0.1in}
-\paragraph{Advantages of \tool.} We can now concretely see the advantages of \tool. Unarguably, it is
+
+\paragraph{Advantages of \tool} We can now concretely see the advantages of \tool. Unarguably, it is
 easier for a developer to program and get the code right in
@@
 guarantees.
+
+\subsection{Comparison with Yao-only approach}
+Before we provide the formalization of our compiler, we show that for performance it it critical to use a mix of arithmetic and boolean circuits.
+We demonstrate the performance benefits of mixed computation over boolean-only compute by comparing with previous tools CBMC-GC~\cite{cbmcgc} and ObliVM~\cite{oblivm} that are based on Yao's garbled circuits.
+As we already mentioned that we are not the first to observe that Yao-only approach does not scale well beyond examples with small numbers of multiplications \cite{aby,tasty,autos,secureml,minionn,chameleon}, especially for large-scale multiplications commonly found in Machine Learning (ML) algorithms -- the main use case of \tool.
+Indeed, this is one of the main drivers behind development of the specialized MPC protocols in previous works \cite{secureml,minionn,shafindss} with which we compare the performance of our automatically generated protocols in \sectionref{eval}.
+
+The largest example benchmark in CBMC-GC is multiplication of two $8\times 8$ matrices for which it generates <FILL> gates and takes <FILL> seconds to run.
+In contrast, in \tool we use arithmetic sharing for this compute that generates <FILL> gates and completes in <FILL> seconds.
+We tried running CBMC-GC for multiplying two $20\times 20$ matrices and it timed out after 5 hours.
+In contrast, most of our benchmarks are much bigger.
+For example, we run a moderate size DNN for <FILL> in <FILL> seconds which contains a multiplication of <FILL> matrix and <FILL> matrix as a sub-computation.
+On the other hand, ObliVM scales to larger benchmarks with its use of Yao-pipelining \cite{yao-pipe}.
+We ran inner-product computations of varying lengths followed by a comparison in both ObliVM and \tool (see Table~\cite{oblivm-compare}).
+Our experiments show that we are atleast 30x faster than ObliVM due to our use of arithmetic sharing for inner-product computation.
 

### 2018-05-04 ccs/intro.tex
yao comparison

@@
  expected to use them correctly and efficiently on their own.
- Ideally, a developer would declare the functionality in a, general
+ Ideally, a developer would declare the functionality in a general
  purpose, high-level programming language and a
@@
 gate.
+\dg{It is well-known that boolean circuits are not suitable for doing arithmetic operations such as integer multiplications but are unavoidable for boolean operations such as comparison \cite{aby,tasty,autos,secureml,minionn,chameleon}.}
 %
@@
 %
+
 Unfortunately, none of the above frameworks support combinations of
@@
 \end{figure}
-\vspace{-0.1in}
+
 \paragraph{Ease of programming.} \tool source programs are ideal
@@
 enjoy these benefits.
-\vspace{-0.1in}
+
 \paragraph{Cryptographic-cost aware compiler.} The \tool compiler
@@
 first such cryptographic-cost aware compiler.
-\vspace{-0.1in}
+
 \paragraph{Scalability (secure code partitioning).} \mpc tools
@@
 %% in \sectionref{pipe}
-\vspace{-0.1in}
+
 \paragraph{Formal guarantees.} We prove formal correctness and
@@
 scheme (\sectionref{ld} and~\sectionref{pipe}).
-%\vspace{-0.2in}
+
 \paragraph{Evaluation.} We have implemented \tool
@@
 that requires more than 300 million gates (\sectionref{impl} and~\sectionref{eval}).
-\vspace{-0.1in}
+
 \paragraph{Related Work.}

### 2018-05-04 ccs/ex.tex
ccc

@@
 of ABY, boolean is synonymous with Yao, we elaborate in \sectionref{impl}).
+\dg{We discuss the choices made by our compiler below in \sectionref{ccc}.}
 
@@
 
+\subsection{Cryptographic cost-aware compiler}
+\label{sec:ccc}
+\dg{Points to address. 1. hard problem nto solve optimally 2. some decisions are fixed and some are context sensitive 3. our heuristics and their basis}
+One of the main challenges for our compiler is to split the compute between arithmetic and boolean parts well. 
+As mentioned above, \tool is a cryptographic cost-aware compiler and uses the arithmetic and boolean costs of primitive operators and the inter-conversion costs to make the split.
+We analyse these costs by running microbenchmarks for basic operations and interconversions and obtained numers consistent with ABY~\cite{aby} (see Figures 2 and 3 in \cite{aby}).
+Based on these costs, our compiler has some hard constraints.
+For example, multiplication should always be performed in arithmetic circuit as the cost of boolean multiplication is much higher than the cost of coercing from boolean to arithmetic, performing an arithmetic multiplication, and then coercing from arithmetic to boolean.
+However, this is not the case for addition.
+In \tool, if both the operands of addition are boolean shared, we do the addition using boolean circuit, else we use arithmetic circuit.
+Hence, addition may be done in arithmetic or boolean depending on the operands.
+We note that doing an optimal split between arithmetic and boolean is an NP-hard problem and in this work, we use some heuristics based on costs observed using microbenchmarking.
+Moreover, for our machine learning benchmarks (such as linear classifier, DNNs, etc) these choices seem optimal.
+
 %% Consider a  cloud service provider Bob that wants to provide a service to diagnose

### 2018-05-04 ccs/impl.tex
starting ccs repo
ccc

--- 2017-11-01 OaklandSubmission/impl.tex
@@
+\vspace{-0.1in}
 \section{Implementation}
@@
 
+\dg{Remove some parts from here..}
 The compilation rules of \figureref{compile} can introduce
@@
 instead of inserting code to recompute them afresh.
+
+Since all these optimizations are standard compiler
+optimizations~\cite{dragonbook}, we rely on their correctness to
+maintain the correctness and security for the optimized programs.

### 2018-05-06 ccs/conclude.tex
changes, comments

@@
+\section{Conclusion and Future Work}
+\label{sec:conclude}
+We presented \tool, the first
+cryptographic-cost aware framework that generates efficient and
+scalable \mpc protocols from high-level programs.
+The generated protocols comprise combinations
+of arithmetic and boolean circuits and have performance comparable to, or better than the custom, specialized
+protocols from previous works.
+The \mpc backed we use provides security against semi-honest or passive adversaries.
+%our compiler is oblivious to the ``type of adversary'' for the underlying \mpc backend.
+Given a \mpc backend secure against malicious or active adversaries, Theorem \ref{theorem:security} 
+can be extended to obtain a framework with malicious security. Furthermore,  
+partitioning can be modified to provide malicious security by {\it signing} the shares of the intermediate states.
+However, we are not aware of a maliciously secure \mpc implementation for combinations of
+arithmetic and boolean circuits.
+
+\dg{Add other future work such as language enhancement to support public inputs, declassification of intermediate values, secret array indices and support of ORAM.
+But since current focus was ML applications, didn't require these as such.
+Also, crypto backends would keep improving and although our ideas extend to works such as Gazelle, we leave integrating it as a backend in \tool as future work.}
+
 %% \section{Conclusion}

### 2018-05-06 ccs/eval.tex
starting ccs repo
changes, comments

--- 2018-02-07 usenix18/eval.tex
@@
+%\vspace{-0.2in}
 \section{Evaluation}
@@
 2.40GHz with 28 GBs of RAM.
-Where we compare our execution times with prior protocols, we match our system and network parameters with those of the prior work (as the code for works such as Bost et al.
+When we compare our execution times with prior protocols, we match our system and network parameters with those of the prior work (as the code for works such as Bost et al.
 \cite{shafindss} and \minion~\cite{minionn} are not publicly available).
@@
 on decision trees and further validates this claim.
-\\\\
+
 \paragraph{Deep neural nets.}
@@
 we discuss it in \sectionref{pipeeval}.
-\\\\
+
 \paragraph{State-of-the-art classifiers.}
@@
 
-\subsection{Matrix factorization.}
+\subsection{Matrix factorization}
 \tool is not tied to secure prediction and can express more general computations.
@@
 
+\subsection{Subsequent Work}
+\TODO{....}
+
 %stage | time | comm | gates | mul | add | depth

### 2018-05-06 ccs/ex.tex
changes, comments

@@
-\vspace{-0.1in}
 \section{\tool Overview}
@@
 ``ideal'' functionalities, without dealing with any cryptographic
-details. The language provides multi-dimensional arrays, conditional
+details. The language provides multi-dimensional arrays \dg{with public indices}, conditional
 expressions (the ternary $?\::$ operator), \ls{for}~loops,
 \ls{if}~statements, and special syntax for input/output from each party.
+\TODO{Add a justification for not having secret indices, public inputs etc. Unless there is a better place to have those.}
 
@@
 We demonstrate the performance benefits of mixed computation over boolean-only compute by comparing with previous tools CBMC-GC~\cite{cbmcgc} and ObliVM~\cite{oblivm} that are based on Yao's garbled circuits.
-As we already mentioned that we are not the first to observe that Yao-only approach does not scale well beyond examples with small numbers of multiplications \cite{aby,tasty,autos,secureml,minionn,chameleon}, especially for large-scale multiplications commonly found in Machine Learning (ML) algorithms -- the main use case of \tool.
-Indeed, this is one of the main drivers behind development of the specialized MPC protocols in previous works \cite{secureml,minionn,shafindss} with which we compare the performance of our automatically generated protocols in \sectionref{eval}.
+As we already mentioned,  we are not the first to observe that Yao-only approach does not scale well beyond examples with small numbers of multiplications \cite{aby,tasty,autoS,secureml,minionn,chameleon}, especially for large-scale multiplications commonly found in Machine Learning (ML) algorithms -- the main use case of \tool.
+\dg{Speaking loosely,} multiplying two $64$-bit integers using arithmetic sharing requires only $2$ multiplications, whereas Yao requires at least $4096$ AES operations in the online phase.
+Indeed, this is one of the main drivers behind development of various specialized MPC protocols in previous works \cite{secureml,minionn,shafindss} that we compare against in \sectionref{eval}.
 
@@
 In contrast, most of our benchmarks are much bigger.
-For example, we run a moderate size DNN for <FILL> in <FILL> seconds which contains a multiplication of <FILL> matrix and <FILL> matrix as a sub-computation.
+For example, we run the secure computation of a DNN for CIFAR-10 in 265 seconds which requires multiplying a $64\times576$ matrix with a $576\times1024$ matrix as a sub-computation.
 On the other hand, ObliVM scales to larger benchmarks with its use of Yao-pipelining \cite{yao-pipe}.
@@
 \label{sec:ccc}
-\dg{Points to address. 1. hard problem nto solve optimally 2. some decisions are fixed and some are context sensitive 3. our heuristics and their basis}
+\dg{Points to address. 1. hard problem to solve optimally 2. some decisions are fixed and some are context sensitive 3. our heuristics and their basis}
 One of the main challenges for our compiler is to split the compute between arithmetic and boolean parts well. 
 As mentioned above, \tool is a cryptographic cost-aware compiler and uses the arithmetic and boolean costs of primitive operators and the inter-conversion costs to make the split.
-We analyse these costs by running microbenchmarks for basic operations and interconversions and obtained numers consistent with ABY~\cite{aby} (see Figures 2 and 3 in \cite{aby}).
+We analyse these costs by running microbenchmarks for basic operations and interconversions and our observations were consistent with ABY~\cite{aby} (see Figures 2 and 3 in \cite{aby}).
 Based on these costs, our compiler has some hard constraints.
-For example, multiplication should always be performed in arithmetic circuit as the cost of boolean multiplication is much higher than the cost of coercing from boolean to arithmetic, performing an arithmetic multiplication, and then coercing from arithmetic to boolean.
+For example, multiplication should always be performed in arithmetic circuit as the cost of boolean multiplication is much higher than  the cost of coercing from boolean to arithmetic, performing an arithmetic multiplication, and then coercing from arithmetic to boolean \dg{($\sim9x$ time $\&$ $\sim20x$ comm.
+in LAN setting)}.
 However, this is not the case for addition.
@@
 Hence, addition may be done in arithmetic or boolean depending on the operands.
-We note that doing an optimal split between arithmetic and boolean is an NP-hard problem and in this work, we use some heuristics based on costs observed using microbenchmarking.
+Finally, all the boolean operations such as comparison, shifts, multiplexors, etc are done using Yao.
+We note that doing an optimal split between arithmetic and boolean is an \dg{NP-hard} problem and in this work, we use some heuristics based on costs observed using microbenchmarking.
 Moreover, for our machine learning benchmarks (such as linear classifier, DNNs, etc) these choices seem optimal.
+
+\dg{\textsc{People wanted estimate of interconversion costs.
+I don't know where to put it.
+May be implementation?} Interconversion costs: Converting from A to Y requires computing computing a Yao circuit for addition that is linear size and hence, requires linear number of symmetric key operations.
+Similarly, Y2A requires computing a subtraction circuit that is again linear size.}
 

### 2018-05-06 ccs/impl.tex
changes, comments

@@
 
-\dg{Remove some parts from here..}
+\dg{remove addition?}
 The compilation rules of \figureref{compile} can introduce
@@
 an addition then \tool performs the addition in boolean rather than first coercing boolean to arithmetic
-and then performing an arithmetic addition. However, multiplication is always performed in arithmetic
-as the cost of a boolean multiplication is much higher than the cost of coercing from boolean to arithmetic, performing an arithmetic multiplication, and then coercing from arithmetic to boolean~\cite{aby}.
+and then performing an arithmetic addition. 
+%However, multiplication is always performed in arithmetic as the cost of a boolean multiplication is much higher than the cost of coercing from boolean to arithmetic, performing an arithmetic multiplication, and then coercing from arithmetic to boolean~\cite{aby}.
 Other optimizations include the standard ``common subexpression elimination"

### 2018-05-06 ccs/intro.tex
changes, comments

@@
 gate.
-\dg{It is well-known that boolean circuits are not suitable for doing arithmetic operations such as integer multiplications but are unavoidable for boolean operations such as comparison \cite{aby,tasty,autos,secureml,minionn,chameleon}.}
+\dg{It is well-known that boolean circuits are not suitable for doing arithmetic operations such as integer multiplications but are unavoidable for boolean operations such as comparison \cite{aby,tasty,autoS,secureml,minionn,chameleon}.}
 %

### 2018-05-06 ccs/ld.tex
starting ccs repo
changes, comments

--- 2018-02-07 usenix18/ld.tex
@@
 \newcommand{\cond}[3]{\ensuremath{{{#1}\:?\:{#2}\::{#3}}}}
-\newcommand{\for}[4]{\ensuremath{\kw{for}\:{#1}\:\kw{in}\:[{#2}, {#3}]\:\kw{do}\:{#4}}}
+\newcommand{\forl}[4]{\ensuremath{\kw{for}\:{#1}\:\kw{in}\:[{#2}, {#3}]\:\kw{do}\:{#4}}}
 \newcommand{\ite}[3]{\ensuremath{\kw{if}({#1}, {#2}, {#3})}}
@@
 In this section we prove correctness and security of our \tool compiler.
+\dg{Put a warning that this section is technical and not prerequisite for understanding later section.}
 %We formalize our source, intermediate and circuit languages. 
@@
 We present only selected parts of our formalization. Full definitions
-and proofs can be found in the Appendix.
+and auxiliary lemmas can be found in~\appendixref{formal}. We will also
+release complete typeset proofs in a tech report.
 
@@
     & &\mid& [\overline{e_{i}}]_{n} \mid x[e] \mid \kw{in}_{j}\\
-    \ftext{Statement} & s &::=& \psi\:x = e \mid x := e \mid \for{x}{n_{1}}{n_{2}}{s}\\
+    \ftext{Statement} & s &::=& \psi\:x = e \mid x := e \mid \forl{x}{n_{1}}{n_{2}}{s}\\
     & & \mid& x[e_{1}] := e_{2} \mid \ite{e}{s_{1}}{s_{2}} \mid \kw{out}\:e \mid s_{1}; s_{2}\\
@@
 $\kw{out}\:e$ denotes revealing the value of $e$ to the
-parties. The \kw{while} statement is an internal syntax that is not
+parties\footnote{Our language also has statements $\kw{out}_1\:e$ (resp., $\kw{out}_2\:e$) to reveal value of $e$ to only the first (resp., second) party.
+We omit these for brevity.}.
+The \kw{while} statement is an internal syntax that is not
 exposed to the programmer. \\
@@
   \[
-  \\
   \begin{array}{c}
@@
                {
-                 \rho \vdash \for{x}{n_{1}}{n_{2}}{s} \Downarrow \rho_{1} - \{x\}; O
+                 \rho \vdash \forl{x}{n_{1}}{n_{2}}{s} \Downarrow \rho_{1} - \{x\}; O
                }
@@
 \newcommand{\lcond}[4]{\ensuremath{{{#2}\:?_{{#1}}\:{#3}\::{#4}}}}
-%\newcommand{\for}[4]{\ensuremath{\kw{for}\:{#1}\:\kw{in}\:[{#2}, {#3}]\:\kw{do}\:{#4}}}
+%\newcommand{\forl}[4]{\ensuremath{\kw{for}\:{#1}\:\kw{in}\:[{#2}, {#3}]\:\kw{do}\:{#4}}}
 %\newcommand{\ite}[3]{\ensuremath{\kw{if}({#1}, {#2}, {#3})}}
@@
 $\sigma^{\ell}[n]$.
-
 Most of the expression forms $\widetilde{e}$ are same as $e$, except
@@
 
+\paragraph{Source to intermediate compilation.} We provide the compilation
+rules in \figureref{compile}. We present the rules in a
+declarative style, where the rules are non-syntax
+directed, and the labels $\ell$ are chosen
+non-deterministically. \sectionref{impl} describes the label  inference
+scheme in our implementation.
+
 \begin{figure}[H]
@@
                {
-                 \Gamma \vdash \for{x}{n_{1}}{n_{2}}{s} \leadsto \for{x}{n_{1}}{n_{2}}{\widetilde{s}} \mid \Gamma
+                 \Gamma \vdash \forl{x}{n_{1}}{n_{2}}{s} \leadsto \forl{x}{n_{1}}{n_{2}}{\widetilde{s}} \mid \Gamma
                }
@@
 \end{figure}
-
-\paragraph{Source to intermediate compilation.} We provide the compilation
-rules in \figureref{compile}. We present the rules in a
-declarative style, where the rules are non-syntax
-directed, and the labels $\ell$ are chosen
-non-deterministically. \sectionref{impl} describes the label  inference
-scheme in our implementation.
+%\vspace{-0.2in}
 
@@
 
+\paragraph{Correctness Theorem.} We prove that all well typed programs always terminate (array indices are always in bounds, there are no unbounded loops, etc.) and the \mpc protocol produces the same outputs as the source program.
+That is, if a source statement $s$ is well-typed, and compiles to $\widetilde{s}$ in the intermediate language, then $s$ terminates in the source semantics with observations $O$, $\widetilde{s}$ evaluates to circuit $\crct$, and $\crct$ terminates in the circuit semantics with the same observations $O$.
+Formally, the the correctness theorem is as follows (the environments on the left of $\vdash$ are empty and we elide the environments on the right of $\vdash$ for brevity):
+
 %\subsubsection*{Correctness theorem}
+\old{
 Finally, the correctness theorem is as follows (the environments on
 the left of $\vdash$ are empty and we elide the environments on the
-right of $\vdash$ for brevity):
+right of $\vdash$ for brevity):}
 
@@
 
-The theorem states that if a source statement $s$ is well-typed, and
+\old{The theorem states that if a source statement $s$ is well-typed, and
 compiles to $\widetilde{s}$ in the intermediate language, then $s$ terminates in the source
@@
 circuit semantics with the same observations $O$. 
-Therefore, well typed programs always terminate (array indices are always in bounds, there are no unbounded loops, etc.) and the \mpc protocol produces the same outputs as the source program.
+Therefore, well typed programs always terminate (array indices are always in bounds, there are no unbounded loops, etc.) and the \mpc protocol produces the same outputs as the source program.}
+
+\dg{We prove the theorem in Appendix~\ref{}.}
 % We first define a judgment for

### 2018-05-06 ccs/pipe.tex
changes, comments

@@
-\vspace{-0.1in}
+%\vspace{-0.1in}
 \section{Secure code partitioning}
@@
 $\progn_1,\progn_2,\ldots,\progn_k$ does
-not reveal any more information than $\prog$). More formal details follow, while an example illustrating the technique of secure code partitioning is provided in  \appendixref{codepartitioning}.
+not reveal any more information than $\prog$). 
+\dg{Prior works such as FastGC \cite{yao-pipe} and ObliVM \cite{oblivm} have considered the idea of pipelining of circuits for secure computation based on Yao's garbled circuits that allows them to scale to large circuits.
+Our technique of secure code partitioning works with high level programs and does source code transformation.
+This gives a generic way to execute programs with large underlying circuits (even a mix of arithmetic and Boolean) and is not tied to the specific details of the cryptographic protocol used.
+}
+More formal details follow, while an example illustrating the technique of secure code partitioning is provided in  \appendixref{codepartitioning}.
+\dg{We describe it only in the arithmetic setting but the same can be done in the Boolean setting as well.}
 
@@
 compute $\stateq_1$ and then outputs $o_1 = \stateq_1 + r_1$ {\em only
-  to} Bob\footnote{While the description of our protocol here assumes
+  to} Bob\footnote{While the description of the scheme here assumes
   that
@@
 %output $z$.
-\vspace{-0.1in}
+
 \paragraph{Implementing code partitioning.}
 We use partitioning for programs that require large
-circuits. Specifically, the programmer decomposes the program
+circuits. Specifically, we first decompose the program
 $\prog$ into a sequence of small programs
-$\prog_1\seq\dots\seq\prog_k$ manually. And then, \tool generates
+$\prog_1\seq\dots\seq\prog_k$. And then, \tool generates
 sequence of programs $\progn_1,\dots\progn_k$ automatically. We then
@@
 usage after execution of each $\progn_i$.
-Automating the manual decomposition step requires an analysis that can 
-statically estimate the resource usage of a \tool program and we leave
-the construction of such an analysis as future work.% \nc{I changed the heading.}
+\dg{Automating the  decomposition step requires an analysis that can 
+statically estimate the resource usage of a \tool program.
+Resource analysis of programs is a well-known hard problem~\cite{raml}
+ and we describe a heuristic analysis in Appendix~\ref{app:pipe-idea}.}
+% \nc{I changed the heading.}

### 2018-05-06 ccs/related.tex
changes, comments

@@
-\vspace{-0.2in}
 \section{Related work}
@@
 \mpc backends  have made  tremendous progress in the last decade.
-For example, the circuits can be optimized for depth~\cite{ddkssz15,cbmcgcdepth}, large garbled circuits can be pipelined~\cite{yao-pipe}, and oblivious RAM~\cite{oram1,oram2} can be used to hide access patterns of MIPS code~\cite{mips}.
+For example, the circuits can be optimized for depth~\cite{ddkssz15,cbmcgcdepth}, large garbled circuits can be pipelined~\cite{yao-pipe,oblivm}, and oblivious RAM~\cite{oram1,oram2} can be used to hide access patterns of MIPS code~\cite{mips}.
 Incorporating these backends would only improve the performance and scalability of \tool implementations.
@@
 Examples include classification of medical ElectroCardioGram (ECG)~\cite{barni}, iris and fingerprint recognition~\cite{blanton}, remote diagnostics~\cite{brickell}, sequence analysis~\cite{franz}, biometric identification~\cite{huang}, matrix factorization and ridge-regression~\cite{valeriaMatrix, valeriaRidge}, machine-learning classification~\cite{shafindss}, decision trees/forests~\cite{wu}, neural network training~\cite{secureml} and prediction~\cite{minionn}.
+
+\TODO{add Groce et al., Mood et al. Lindell's Libscapi, Chameleon, Gazelle, SPDZ. Reviewers want citations for  3PC papers also. May be we can say they are not 2PC?}
 % As seen from our experiments, the protocols generated by \tool give us better or comparable execution times as many of the above works with minimal programmer overhead.

### 2018-05-06 ccs/eval.tex
comments

@@
 competitive with \minion here as well. Therefore,
-we believe that with partitioning, \tool can scale to arbitrary sized
+we believe that with partitioning, \tool can scale to large
 computations while maintaining performance

### 2018-05-06 ccs/ex.tex
comments

@@
 \label{sec:ccc}
-\dg{Points to address. 1. hard problem to solve optimally 2. some decisions are fixed and some are context sensitive 3. our heuristics and their basis}
 One of the main challenges for our compiler is to split the compute between arithmetic and boolean parts well. 
 As mentioned above, \tool is a cryptographic cost-aware compiler and uses the arithmetic and boolean costs of primitive operators and the inter-conversion costs to make the split.
-We analyse these costs by running microbenchmarks for basic operations and interconversions and our observations were consistent with ABY~\cite{aby} (see Figures 2 and 3 in \cite{aby}).
+Cryptographic cost of converting an arithmetic share to a boolean share requires computing a garbled circuit for addition that is linear in size.
+Similarly, the cost of converting from boolean to arithmetic is computing a subtraction circuit inside garbled circuit that is again linear in size.
+ABY analyses all of these costs by running microbenchmarks for basic operations and interconversions (see Figures 2 and 3 in \cite{aby}).
 Based on these costs, our compiler has some hard constraints.
@@
 Moreover, for our machine learning benchmarks (such as linear classifier, DNNs, etc) these choices seem optimal.
-
-\dg{\textsc{People wanted estimate of interconversion costs.
-I don't know where to put it.
-May be implementation?} Interconversion costs: Converting from A to Y requires computing computing a Yao circuit for addition that is linear size and hence, requires linear number of symmetric key operations.
-Similarly, Y2A requires computing a subtraction circuit that is again linear size.}
 

### 2018-05-06 ccs/impl.tex
comments

@@
 library~\cite{aby}. 
+\TODO{Fix this!}
 ABY provides a GMW-based~\cite{gmw} and a  Yao-based~\cite{yao} protocol for boolean computations.
@@
 and use it in our evaluation.
-We use the default ABY configuration for our
-experiment\footnote{We set the security parameter to 128 and use OT
-extension-based arithmetic multiplication triples generation.}. ABY
+We use $128$ bits of security and OT
+extension-based arithmetic multiplication triplets generation. ABY
 provides multi-threading support (for the offline phase of the \mpc

### 2018-05-06 ccs/intro.tex
comments

@@
 or libraries that lack formal semantics and static guarantees.
+\dg{Cite works like Wysteria~\cite{wysteria} for public inputs and declassification of intermediate values, ObliVM~\cite{oblivm} for hiding secret array indices using Oblivious RAM.
+But since the focus of this work is machine learning applications, we do not concern ourselves with these language complexities.}
 %Can be added to related work

### 2018-05-06 ccs/main.tex
includepdf
changes, comments
comments

@@
 
-%\usepackage{epsfig,endnotes,pdfpages}
+\usepackage{epsfig,endnotes,pdfpages}
 \usepackage{wrapfig}
 \usepackage{float}
+
 \begin{document}
 
-\title{\tool: Programmable, Efficient, and Scalable \\ Secure Two-Party Computation}
+\title{\tool: Programmable, Efficient, and Scalable \\ Secure Two-Party Computation \dg{for Machine Learning}}
 
@@
 secure two-party computation (\mpc) framework that generates efficient
-\mpc protocols from high-level, easy-to-write, programs.
+\mpc protocols from high-level, easy-to-write programs.
 \tool provides formal correctness and security guarantees while maintaining performance and scalability.

### 2018-05-06 ccs/conclude.tex
comments
reviews incorporate

@@
 \label{sec:conclude}
+\dg{Make it consistent with contributions in intro.}
 We presented \tool, the first
@@
 But since current focus was ML applications, didn't require these as such.
-Also, crypto backends would keep improving and although our ideas extend to works such as Gazelle, we leave integrating it as a backend in \tool as future work.}
+Also add functions.
+Also, crypto backends would keep improving and although our ideas extend to works such as Gazelle, we leave integrating it as a backend in \tool as future work.
+
+We will explore the possibility of mechanically verifying the implementation of our compiler.
+
+Tensorflow to \tool.}
 

### 2018-05-06 ccs/eval.tex
reviews incorporate

@@
 \hline
-Dataset & $d$  & \thead{Prev \\ time\\ (s)} & \thead{Prev \\ comm\\ (kb)} & \thead{LAN  \\ (s)} & \thead{WAN\\ (s)} & \thead{Comm \\(kb)}  &  \thead{Num \\ gates} & LOC\\
+Dataset & $d$  & \thead{Prev \\ time\\ (s)} & \thead{Prev \\ comm\\ (KB)} & \thead{LAN  \\ (s)} & \thead{WAN\\ (s)} & \thead{Comm \\(KB)}  &  \thead{Num \\ gates} & LOC\\
 \hline
@@
 \hline
-Dataset & $n$ & $F$ & \thead{Prev \\ time\\ (s)} & \thead{Prev \\ comm\\ (kb)} & \thead{LAN  \\ (s)} & \thead{WAN\\ (s)} & \thead{Comm \\(kb)} & \thead{Num \\ gates}  & LOC\\
+Dataset & $n$ & $F$ & \thead{Prev \\ time\\ (s)} & \thead{Prev \\ comm\\ (KB)} & \thead{LAN  \\ (s)} & \thead{WAN\\ (s)} & \thead{Comm \\(KB)} & \thead{Num \\ gates}  & LOC\\
 \hline
@@
 \hline
-Dataset  & $d$ & $N$ & \thead{Prev \\ time\\ (s)} & \thead{Prev \\ comm\\ (kb)} & \thead{LAN  \\ (s)} & \thead{WAN\\ (s)} & \thead{Comm \\(kb)}& \thead{Num \\ gates} & LOC\\
+Dataset  & $d$ & $N$ & \thead{Prev \\ time\\ (s)} & \thead{Prev \\ comm\\ (KB)} & \thead{LAN  \\ (s)} & \thead{WAN\\ (s)} & \thead{Comm \\(KB)}& \thead{Num \\ gates} & LOC\\
 \hline
@@
 \hline
-DNN  & \thead{Prev \\ time\\ (s)} & \thead{Prev \\ comm\\ (kb)} & \thead{LAN  \\ (s)} & \thead{WAN\\ (s)} & \thead{Comm \\(kb)} & \thead{Num \\ gates} & \thead{Model\\ size} & LOC\\
+DNN  & \thead{Prev \\ time\\ (s)} & \thead{Prev \\ comm\\ (MB)} & \thead{LAN  \\ (s)} & \thead{WAN\\ (s)} & \thead{Comm \\(MB)} & \thead{Num \\ gates} & \thead{Model\\ size} & LOC\\
 \hline
@@
 \hline
-Classifier       & \thead{LAN  \\ (s)} & \thead{WAN \\(s)} & \thead{Comm \\(Mb)}  & \thead{Num \\ And} & \thead{Num \\ Mul} & \thead{Num \\ gates} & \thead{Model \\size} & LOC\\
+Classifier       & \thead{LAN  \\ (s)} & \thead{WAN \\(s)} & \thead{Comm \\(MB)}  & \thead{Num \\ And} & \thead{Num \\ Mul} & \thead{Num \\ gates} & \thead{Model \\size} & LOC\\
 \hline
@@
 \hline
-Dataset       & \thead{LAN \\(s)} & \thead{WAN \\(s)} & \thead{Comm \\ (Mb)}  & \thead{Num \\ And} & \thead{Num \\ Mul} & \thead{Num \\ gates} & depth & LOC\\
+Dataset       & \thead{LAN \\(s)} & \thead{WAN \\(s)} & \thead{Comm \\ (MB)}  & \thead{Num \\ And} & \thead{Num \\ Mul} & \thead{Num \\ gates} & depth & LOC\\
 \hline
@@
 \hline
-           &  LAN (s) & WAN (s) & \thead{Comm \\ (Mb)}  & \thead{Num \\ And} & \thead{Num \\ Mul} & \thead{Num \\ gates} \\
+           &  LAN (s) & WAN (s) & \thead{Comm \\ (MB)}  & \thead{Num \\ And} & \thead{Num \\ Mul} & \thead{Num \\ gates} \\
 \hline
@@
 \hline
-  Stage         &  LAN (s) & WAN (s) & \thead{Comm \\ (Mb)}  & depth & \thead{Num \\ gates} & LOC\\
+  Stage         &  LAN (s) & WAN (s) & \thead{Comm \\ (MB)}  & depth & \thead{Num \\ gates} & LOC\\
 \hline
@@
 out of memory but an appropriately partitioned \tool implementation would still
-succeed.
+succeed. For example, consider a large DNN with $n$ layers such that the circuit for only a single layer can  fit into memory. Then, \tool using partitioning can evaluate this large DNN in $n$ stages.
 
@@
 \subsection{Subsequent Work}
-\TODO{....}
+Subsequent to our work, there have been improvements in the cryptographic backends.
+For example, Gazelle describes specialized protocols for DNNs---which use a 
+lattice-based packed additively homomorphic encryption scheme (PAHE) for arithmetic computations and garbled circuits for boolean computations---that can evaluate the CNN benchmark of Table~\ref{tab:nn} in 0.8 seconds (as opposed to
+5.1 seconds taken by \tool when using ABY). Such advances in cryptography are
+orthogonal to our contributions. In particular, we are interested in using Gazelle
+as our backend when it becomes available. Furthermore, the authors of Gazelle remark:
+``A final, very interesting and ambitious line
+of work would be to build a compiler that allows us to easily
+express arbitrary computations and automatically factor the
+computation into PAHE and two-party primitives.''
+\tool addresses this exact problem of factoring computations into
+arithmetic and boolean parts.
 

### 2018-05-06 ccs/ex.tex
reviews incorporate

@@
 ``ideal'' functionalities, without dealing with any cryptographic
-details. The language provides multi-dimensional arrays \dg{with public indices}, conditional
+details. The language provides multi-dimensional arrays with public indices, conditional
 expressions (the ternary $?\::$ operator), \ls{for}~loops,
 \ls{if}~statements, and special syntax for input/output from each party.
-\TODO{Add a justification for not having secret indices, public inputs etc. Unless there is a better place to have those.}
 
@@
 of ABY, boolean is synonymous with Yao, we elaborate in \sectionref{impl}).
-\dg{We discuss the choices made by our compiler below in \sectionref{ccc}.}
+We elaborate on cryptographic cost awareness in \sectionref{ccc}.
 
@@
 in LAN setting)}.
+Since the performance gap between arithmetic and boolean is large, we believe that this choice is optimal for realistic network settings.
 However, this is not the case for addition.
@@
 We note that doing an optimal split between arithmetic and boolean is an \dg{NP-hard} problem and in this work, we use some heuristics based on costs observed using microbenchmarking. 
-Moreover, for our machine learning benchmarks (such as linear classifier, DNNs, etc) these choices seem optimal.
+For the smaller benchmarks of our evaluation, we wrote ABY code directly and the circuits
+generated by our compiler had about the same size as the  hand written code. 
+%Moreover, for our machine learning benchmarks (such as linear classifier, DNNs, etc) these choices seem optimal.
 

### 2018-05-06 ccs/impl.tex
reviews incorporate

@@
 library~\cite{aby}. 
-\TODO{Fix this!}
-ABY provides a GMW-based~\cite{gmw} and a  Yao-based~\cite{yao} protocol for boolean computations.
-Although \tool can generate code for both, we have observed better performance when using the latter
-and use it in our evaluation.
+ABY provides support for Arithmetic computation based on \cite{autoS}, and boolean computations based on GMW~\cite{gmw} as well as Yao's garbled circuits~\cite{yao}.
+Although \tool can generate code for both kinds of boolean computations, we have observed better performance when using the garbled circuits
+and use it in our evaluation. Hence, \tool generated code uses Arithmetic computation and garbled circuits based boolean computations.
 We use $128$ bits of security and OT
@@
 \figureref{compile}). We take the \tool source program and
-translate it as an input to the solver. The solver takes less
+translate it as an input C program to the solver. The solver takes less
 than a minute on our largest benchmark to verify that all the array
-accesses are in-bounds.
+accesses are in-bounds. This C program can also enable
+validation of \tool generated protocols via differential testing~\cite{mckeeman,frigate}.
 
@@
 
-\dg{remove addition?}
 The compilation rules of \figureref{compile} can introduce

### 2018-05-06 ccs/intro.tex
reviews incorporate

--- 2018-05-06 ccs/intro.tex
@@
 gate.
-\dg{It is well-known that boolean circuits are not suitable for doing arithmetic operations such as integer multiplications but are unavoidable for boolean operations such as comparison \cite{aby,tasty,autoS,secureml,minionn,chameleon}.}
+It is well-known that boolean circuits are not suitable for doing arithmetic operations such as integer multiplications but are unavoidable for boolean operations such as comparison \cite{aby,tasty,autoS,secureml,minionn,chameleon}.
 %
@@
 scalable \mpc protocols from high-level programs devoid of any
-cryptographic details. The generated protocols use combinations of
-arithmetic and boolean circuits and have performance comparable to
+cryptographic details. 
+The generated protocols use combinations of
+arithmetic and boolean circuits that introduce unique compilation challenges.
+Standard compilers do not attempt to reason about or insert non-trivial type casts
+and require the programmers to perform them explicitly. 
+We back \tool with formal semantics that enable the compiler to
+insert type conversions between boolean and arithmetic components
+while maintaining provable correctness, security, and efficiency.
+The automatically generated protocols have performance comparable to
 or better than the custom, specialized
@@
 or libraries that lack formal semantics and static guarantees.
+Finally, we focus on the language
+features necessary to implement machine learning applications.
+In particular, we do not discuss declassification of
+intermediate values or indexing into arrays at secret indices.
+ Handling them requires additional complexity. For example,
+Wysteria~\cite{wysteria} handles the former using dependent types
+and ObliVM~\cite{oblivm} uses Oblivious RAM for the latter.
+These situations arise seldomly in the machine learning applications that we consider:
+intermediate values don't need to be declassified and arrays are traversed in an oblivious manner.
+
+%\dg{Cite works like Wysteria~\cite{wysteria} for public inputs and declassification of intermediate values, ObliVM~\cite{oblivm} for hiding secret array indices using Oblivious RAM.
+But since the focus of this work is machine learning applications, we do not concern ourselves with these language complexities.}
 %Can be added to related work

### 2018-05-06 ccs/ld.tex
reviews incorporate

@@
 In this section we prove correctness and security of our \tool compiler.
-\dg{Put a warning that this section is technical and not prerequisite for understanding later section.}
+The readers who are interested in using the compiler as a black box can
+move directly to  Section~\ref{sec:pipe}.
+%\dg{Put a warning that this section is technical and not prerequisite for understanding later section.}
 %We formalize our source, intermediate and circuit languages. 
@@
 
-\begin{figure}[H]
+\begin{figure}[t]
   \footnotesize
@@
 
-\paragraph{Correctness Theorem.} We prove that all well typed programs always terminate (array indices are always in bounds, there are no unbounded loops, etc.) and the \mpc protocol produces the same outputs as the source program.
+\paragraph{Correctness Theorem.} We prove that all well typed programs always terminate successfully (array indices are always in bounds, there are no unbounded loops, etc.) and the \mpc protocol produces the same outputs as the source program.
 That is, if a source statement $s$ is well-typed, and compiles to $\widetilde{s}$ in the intermediate language, then $s$ terminates in the source semantics with observations $O$, $\widetilde{s}$ evaluates to circuit $\crct$, and $\crct$ terminates in the circuit semantics with the same observations $O$.
-Formally, the the correctness theorem is as follows (the environments on the left of $\vdash$ are empty and we elide the environments on the right of $\vdash$ for brevity):
+Formally, the correctness theorem is as follows (the environments on the left of $\vdash$ are empty and we elide the environments on the right of $\vdash$ for brevity):
 
 %\subsubsection*{Correctness theorem}
-\old{
-Finally, the correctness theorem is as follows (the environments on
-the left of $\vdash$ are empty and we elide the environments on the
-right of $\vdash$ for brevity):}
+
+%Finally, the correctness theorem is as follows (the environments onthe left of $\vdash$ are empty and we elide the environments on the right of $\vdash$ for brevity):
 
@@
 
-\old{The theorem states that if a source statement $s$ is well-typed, and
-compiles to $\widetilde{s}$ in the intermediate language, then $s$ terminates in the source
-semantics with observations $O$, $\widetilde{s}$ evaluates to circuit $\crct$, and $\crct$ terminates in the
-circuit semantics with the same observations $O$. 
-Therefore, well typed programs always terminate (array indices are always in bounds, there are no unbounded loops, etc.) and the \mpc protocol produces the same outputs as the source program.}
-
-\dg{We prove the theorem in Appendix~\ref{}.}
+%%\old{The theorem states that if a source statement $s$ is well-typed, and
+%compiles to $\widetilde{s}$ in the intermediate language, then $s$ terminates in the source
+%semantics with observations $O$, $\widetilde{s}$ evaluates to circuit $\crct$, and $\crct$ terminates in the
+%circuit semantics with the same observations $O$. 
+%Therefore, well typed programs always terminate (array indices are always in bounds, there are no unbounded loops, etc.) and the \mpc protocol produces the same outputs as the source program.}
+
+We prove the theorem in Appendix~\ref{app:formal}.
 % We first define a judgment for

### 2018-05-06 ccs/pipe.tex
minor
reviews incorporate

@@
 \newcommand{\progn}{t}
+\newcommand{\mprog}{u}
 \newcommand{\seq}{||}
@@
 not reveal any more information than $\prog$). 
-\dg{Prior works such as FastGC \cite{yao-pipe} and ObliVM \cite{oblivm} have considered the idea of pipelining of circuits for secure computation based on Yao's garbled circuits that allows them to scale to large circuits.
+Our techniques take inspiration from the idea of pipelining Yao's garbled circuits described in FastGC \cite{yao-pipe}.
+However, unlike FastGC, we do not operate at a circuit level and partitioning is independent of the specific
+\mpc protocol. 
+%
+%Prior works such as FastGC \cite{yao-pipe} and ObliVM \cite{oblivm} have developed similar techniques specific to garbled circuits that work at the level of circuits.
+%
+%considered the idea of pipelining of circuits for secure computation based on Yao's garbled circuits that allows them to scale to large circuits.
 Our technique of secure code partitioning works with high level programs and does source code transformation.
 This gives a generic way to execute programs with large underlying circuits (even a mix of arithmetic and Boolean) and is not tied to the specific details of the cryptographic protocol used.
-}
+%}
 More formal details follow, while an example illustrating the technique of secure code partitioning is provided in  \appendixref{codepartitioning}.
-\dg{We describe it only in the arithmetic setting but the same can be done in the Boolean setting as well.}
 
@@
 on inputs $x,y$ and state $\stateq_{k-1} = (o_{k-1}-r_{k-1})$ and outputs $z$ to both
-parties.
+parties. Although we have used arithmetic sharing here, Boolean sharing can be used to achieve the same effect.
 
@@
 usage after execution of each $\progn_i$.
-\dg{Automating the  decomposition step requires an analysis that can 
+Automating the  decomposition step requires an analysis that can 
 statically estimate the resource usage of a \tool program.
-Resource analysis of programs is a well-known hard problem~\cite{raml}
- and we describe a heuristic analysis in Appendix~\ref{app:pipe-idea}.}
+Resource analysis of high-level programs is a well-known hard problem~\cite{raml}
+and we describe a heuristic analysis.
+
+To build $s_1$, we consider the longest
+prefix of $s$ whose computation {\it size} is below the threshold enforced by the available memory of the machine.
+If $\prog = \prog_1; \prog_r$ then we recurse on $\prog_r$ to obtain $\prog_2,\ldots,\prog_k$.
+For a program $\mprog$, to estimate ${\it size}(\mprog)$, we need to discuss three important cases: if $\mprog\equiv\mprog_1;\mprog_2$
+then ${\it size}(\mprog)={\it size}(\mprog_1)+{\it size}(\mprog_2)$; if $\mprog\equiv\ite{e_1}{\mprog_1}{\mprog_2}$ then ${\it size}(\mprog)=\max({\it size}(\mprog_1),{\it size}(\mprog_2))$; if $s\equiv\forl{i}{n_1}{n_2}{\mprog_1}$ then
+${\it size}(\mprog)=(n_2-n_1){\it size}(\mprog_1)$. If $(n_2-n_1){\it size}(\mprog_1)$ is above the threshold then
+we replace $\mprog$ by 
+\[\forl{i}{n_1}{\frac{n_1+n_2}{2}}{\mprog_1}; \forl{i}{\frac{n_1+n_2}{2}}{n_2}{\mprog_1}\]
+ and recurse to find the prefix again.
+This heuristic analysis is sufficient for the benchmarks discussed in our evaluation.
+
 % \nc{I changed the heading.}

### 2018-05-06 ccs/related.tex
reviews incorporate

--- 2018-04-27 ccs/related.tex
@@
 \mpc backends  have made  tremendous progress in the last decade.
-For example, the circuits can be optimized for depth~\cite{ddkssz15,cbmcgcdepth}, large garbled circuits can be pipelined~\cite{yao-pipe}, and oblivious RAM~\cite{oram1,oram2} can be used to hide access patterns of MIPS code~\cite{mips}.
+For example, the circuits can be optimized for depth~\cite{ddkssz15,cbmcgcdepth}, large garbled circuits can be pipelined~\cite{yao-pipe,oblivm}, online complexity can be reduced at the cost of offline complexity~\cite{groce}, encrypted values output from a garbled circuit can be reused~\cite{ReuseOrLose} and oblivious RAM~\cite{oram1,oram2} can be used to hide access patterns of MIPS code~\cite{mips}.
 Incorporating these backends would only improve the performance and scalability of \tool implementations.
@@
 Examples include classification of medical ElectroCardioGram (ECG)~\cite{barni}, iris and fingerprint recognition~\cite{blanton}, remote diagnostics~\cite{brickell}, sequence analysis~\cite{franz}, biometric identification~\cite{huang}, matrix factorization and ridge-regression~\cite{valeriaMatrix, valeriaRidge}, machine-learning classification~\cite{shafindss}, decision trees/forests~\cite{wu}, neural network training~\cite{secureml} and prediction~\cite{minionn}.
+
 % As seen from our experiments, the protocols generated by \tool give us better or comparable execution times as many of the above works with minimal programmer overhead.

### 2018-05-06 ccs/security.tex
reviews incorporate

--- 2018-01-18 usenix18/security.tex
@@
+%\vspace{-0.2in}
 \paragraph{Security theorem.} The protocols we generate provide
 simulation-based security against a semi-honest adversary, in the
-framework of ~\cite{gmw,can00,can01}. At a very high level, in this
+framework of ~\cite{gmw,can00,can01}\dg{ and provides provable security against all side-channel attacks}. At a very high level, in this
 framework, parties are modeled as non-uniform interactive turing

### 2018-05-06 ccs/conclude.tex
minor

--- 2018-04-27 ccs/conclude.tex
@@
+\section{Conclusion and Future Work}
+\label{sec:conclude}
+
+We presented \tool, the first
+cryptographic-cost aware framework that generates efficient and
+scalable \mpc protocols from high-level programs.
+The generated protocols comprise combinations
+of arithmetic and boolean circuits and have performance comparable to, or better than the previously known custom specialized
+protocols from previous works.
+The compiler is backed by formal semantics that help 
+%it
+%insert type coercions between boolean and arithmetic parts automatically while 
+it maintain correctness, security,
+and efficiency.
+
+Currently, we are working on a front-end to translate Tensorflow code to \tool.
+The aim here is to provide a push button implementation to data scientists that generates secure implementations for their existing models.
+In the future, we would like to extend our security guarantees to malicious adversaries. However, we are currently unaware of a maliciously secure \mpc implementation for combinations of
+arithmetic and boolean circuits.
+The cryptographic backends continue to improve and the modular design of \tool makes it easy to integrate with new backends.
+We will continue integrating with the best available backends.
+Finally, we will explore the possibility of mechanically verifying the implementation of our compiler.
+
 %% \section{Conclusion}

### 2018-05-06 ccs/ex.tex
minor

@@
 As we already mentioned,  we are not the first to observe that Yao-only approach does not scale well beyond examples with small numbers of multiplications \cite{aby,tasty,autoS,secureml,minionn,chameleon}, especially for large-scale multiplications commonly found in Machine Learning (ML) algorithms -- the main use case of \tool.
-\dg{Speaking loosely,} multiplying two $64$-bit integers using arithmetic sharing requires only $2$ multiplications, whereas Yao requires at least $4096$ AES operations in the online phase.
+Roughly, multiplying two $64$-bit integers using arithmetic sharing requires only $2$ multiplications, whereas Yao requires at least $4096$ AES operations in the online phase.
 Indeed, this is one of the main drivers behind development of various specialized MPC protocols in previous works \cite{secureml,minionn,shafindss} that we compare against in \sectionref{eval}.
 
-The largest example benchmark in CBMC-GC is multiplication of two $8\times 8$ matrices for which it generates <FILL> gates and takes <FILL> seconds to run.
+The largest example benchmark in CBMC-GC is multiplication of two $8\times 8$ matrices for which it generates about a million gates and takes about ten seconds to run.
 In contrast, in \tool we use arithmetic sharing for this compute that generates <FILL> gates and completes in <FILL> seconds.
@@
 Based on these costs, our compiler has some hard constraints.
-For example, multiplication should always be performed in arithmetic circuit as the cost of boolean multiplication is much higher than  the cost of coercing from boolean to arithmetic, performing an arithmetic multiplication, and then coercing from arithmetic to boolean \dg{($\sim9x$ time $\&$ $\sim20x$ comm.
-in LAN setting)}.
-Since the performance gap between arithmetic and boolean is large, we believe that this choice is optimal for realistic network settings.
+For example, multiplication should always be performed in arithmetic circuit as the cost of boolean multiplication is much higher than  the cost of coercing from boolean to arithmetic, performing an arithmetic multiplication, and then coercing from arithmetic to boolean ($\sim9x$ more time $\&$ $\sim20x$ more communication in a LAN setting).
+Since the performance gap between arithmetic and boolean multiplication is quite large, we believe that this choice is optimal for realistic network settings.
 However, this is not the case for addition.
@@
 Finally, all the boolean operations such as comparison, shifts, multiplexors, etc are done using Yao.
-We note that doing an optimal split between arithmetic and boolean is an \dg{NP-hard} problem and in this work, we use some heuristics based on costs observed using microbenchmarking. 
+We note that doing an optimal split between arithmetic and boolean is an NP-hard problem and in this work, we use some heuristics based on costs observed using microbenchmarking. 
 For the smaller benchmarks of our evaluation, we wrote ABY code directly and the circuits

### 2018-05-06 ccs/security.tex
minor

--- 2018-01-18 usenix18/security.tex
@@
+%\vspace{-0.2in}
 \paragraph{Security theorem.} The protocols we generate provide
 simulation-based security against a semi-honest adversary, in the
-framework of ~\cite{gmw,can00,can01}. At a very high level, in this
+framework of ~\cite{gmw,can00,can01} and provide provable security against all side-channel attacks.
+At a very high level, in this
 framework, parties are modeled as non-uniform interactive turing

### 2018-05-07 ccs/eval.tex
pass over paper. added partitioning figure.

@@
 accuracy 74.71\%), text recognition (USPS~\cite{hull}, accuracy
-94.4\%), and object categorization (WARD~\cite{yang}, accuracy
+94.4\%), and object categorization (WARD\\~\cite{yang}, accuracy
 95.7\%). 
@@
 out of memory but an appropriately partitioned \tool implementation would still
-succeed. For example, consider a large DNN with $n$ layers such that the circuit for only a single layer can  fit into memory. Then, \tool using partitioning can evaluate this large DNN in $n$ stages.
+succeed.
+For example, consider a large DNN with $n$ layers such that the circuit for only a single layer can  fit into memory.
+Then, \tool using partitioning can evaluate this large DNN in $n$ stages.
+To further illustrate the need for and utility of partitioning, we ran a series of DNNs with different layers with and without partitioning (where partitioning was done for every layer of the DNN).
+For up to 5-layer DNNs, the performance of both settings was comparable, thereby illustrating that partitioning did not cause any overhead to performance.
+For networks that had 6 or more layers, performance of the protocol without partitioning dramatically dropped, while the protocol with partitioning continued to scale linearly.
+These results are illustrated in Figure~\ref{fig:partitioning}.
+
+\begin{figure}
+\includegraphics[width=8cm]{partitioning.png}
+\caption{Comparison of \tool code with and without partitioning. x-axis denotes the number of layers of the DNN, while y-axis denotes time in seconds for the secure protocol.}
+\label{fig:partitioning}
+\end{figure}
 

### 2018-05-07 ccs/ex.tex
pass over paper. added partitioning figure.

@@
 The largest example benchmark in CBMC-GC is multiplication of two $8\times 8$ matrices for which it generates about a million gates and takes about ten seconds to run.
-In contrast, in \tool we use arithmetic sharing for this compute that generates <FILL> gates and completes in <FILL> seconds.
+In contrast, in \tool we use arithmetic sharing for this compute that generates 1218 gates and runs in about 0.1 seconds.
 We tried running CBMC-GC for multiplying two $20\times 20$ matrices and it timed out after 5 hours.
@@
 On the other hand, ObliVM scales to larger benchmarks with its use of Yao-pipelining \cite{yao-pipe}.
-We ran inner-product computations of varying lengths followed by a comparison in both ObliVM and \tool (see Table~\cite{oblivm-compare}).
+We ran inner-product computations of varying lengths followed by a comparison in both ObliVM and \tool (see Table~\cite{oblivm-compare} in the Appendix).
 Our experiments show that we are atleast 30x faster than ObliVM due to our use of arithmetic sharing for inner-product computation.
@@
 Finally, all the boolean operations such as comparison, shifts, multiplexors, etc are done using Yao.
-We note that doing an optimal split between arithmetic and boolean is an NP-hard problem and in this work, we use some heuristics based on costs observed using microbenchmarking. 
+We note that doing an optimal split between arithmetic and boolean is an NP-hard problem and in this work, we use heuristics based on costs observed using microbenchmarking. 
 For the smaller benchmarks of our evaluation, we wrote ABY code directly and the circuits

### 2018-05-07 ccs/impl.tex
pass over paper. added partitioning figure.

@@
 
-Since all these optimizations are standard compiler
-optimizations~\cite{dragonbook}, we rely on their correctness to
+All these optimizations are standard compiler
+optimizations~\cite{dragonbook}, and we rely on their correctness to
 maintain the correctness and security for the optimized programs.

### 2018-05-07 ccs/intro.tex
pass over paper. added partitioning figure.

@@
 $\cdot^{T}$ is the matrix transpose operator, and $w^Tx$ denotes the
-inner product of  $w$ and $x$. Alice has following options.
+inner product of  $w$ and $x$. Alice has the following options.
 
@@
 along. While this
-addresses the scalability concern (the circuit
+addresses the scalability concern (i.e., the circuit
 sizes of the sub-programs are now small enough to fit in the memory),
@@
 and ObliVM~\cite{oblivm} uses Oblivious RAM for the latter.
-These situations arise seldomly in the machine learning applications that we consider:
-intermediate values don't need to be declassified and arrays are traversed in an oblivious manner.
+These situations seldomly arise in the machine learning applications that we consider:
+intermediate values don't need to be declassified and arrays are traversed in an oblivious manner. We provide a detailed survey of related
+work in \sectionref{related}.
 
@@
 %in a richer language with control flow statements such as loops and branches.
- We provide a detailed survey of related
-work in \sectionref{related}.
 

### 2018-05-07 ccs/eval.tex
tweaks

@@
 \label{sec:pipeeval}
+\aseem{Is it worth splitting it into microbenchamrking first, which is
+  Fig. 11, and then comparison with others? Right now, Fig. 11 text
+  just comes at the end, should be highlight it in a paragraph or
+  something?}
 The largest benchmark of \minion~\cite{minionn} is a DNN for CIFAR-10
@@
 out of memory but an appropriately partitioned \tool implementation would still
-succeed.
-For example, consider a large DNN with $n$ layers such that the circuit for only a single layer can  fit into memory.
-Then, \tool using partitioning can evaluate this large DNN in $n$ stages.
-To further illustrate the need for and utility of partitioning, we ran a series of DNNs with different layers with and without partitioning (where partitioning was done for every layer of the DNN).
-For up to 5-layer DNNs, the performance of both settings was comparable, thereby illustrating that partitioning did not cause any overhead to performance.
-For networks that had 6 or more layers, performance of the protocol without partitioning dramatically dropped, while the protocol with partitioning continued to scale linearly.
-These results are illustrated in Figure~\ref{fig:partitioning}.
+succeed. For example, consider a large DNN with $n$ layers such that
+the circuit for only a single layer can  fit into memory. Then, \tool
+using partitioning can evaluate this large DNN in $n$ stages. To
+further illustrate the need for and utility of partitioning, we ran a
+series of DNNs with different layers with and without partitioning
+(where partitioning was done for every layer of the DNN). For up to
+5-layer DNNs, the performance of both settings was comparable, thereby
+illustrating that partitioning did not cause any overhead to
+performance. For networks that had 6 or more layers, performance of
+the protocol without partitioning dramatically dropped, while the
+protocol with partitioning continued to scale linearly. These results
+are illustrated in Figure~\ref{fig:partitioning}.
 
@@
 \subsection{Subsequent Work}
-Subsequent to our work, there have been improvements in the cryptographic backends.
-For example, Gazelle describes specialized protocols for DNNs---which use a 
-lattice-based packed additively homomorphic encryption scheme (PAHE) for arithmetic computations and garbled circuits for boolean computations---that can evaluate the CNN benchmark of Table~\ref{tab:nn} in 0.8 seconds (as opposed to
-5.1 seconds taken by \tool when using ABY). Such advances in cryptography are
-orthogonal to our contributions. In particular, we are interested in using Gazelle
-as our backend when it becomes available. Furthermore, the authors of Gazelle remark:
+Subsequent to our work, \aseem{TODO} et al. \aseem{cite} have
+presented Gazelle, a framework for specialized DNN
+protocols. Gazelle use a lattice-based packed additively homomorphic
+encryption scheme (PAHE) for arithmetic computations and garbled
+circuits for boolean computations. Gazelle can evaluate the CNN benchmark of
+Table~\ref{tab:nn} in 0.8 seconds, as opposed to
+5.1 seconds taken by \tool with the ABY backend. Such advances in
+cryptographic backends are orthogonal to our contributions. In
+particular, once Gazelle is available, we could add it as another
+cryptographic backend to \tool. Furthermore, the authors of Gazelle remark:
 ``A final, very interesting and ambitious line
@@
 express arbitrary computations and automatically factor the
-computation into PAHE and two-party primitives.''
-\tool addresses this exact problem of factoring computations into
-arithmetic and boolean parts.
+computation into PAHE and two-party primitives'' -- the exact problem
+that \tool solves.
 

### 2018-05-07 ccs/ex.tex
tweaks

@@
 (lines~\ref{line:condyaobegin} to~\ref{line:condyaoend}) (in our usage
-of ABY, boolean is synonymous with Yao, we elaborate in \sectionref{impl}).
-We elaborate on cryptographic cost awareness in \sectionref{ccc}.
+of ABY, boolean is synonymous with Yao, we elaborate in
+\sectionref{impl}). We elaborate on cryptographic cost awareness in
+\sectionref{ccc}.
 
@@
 \ls{a\_acc} and \ls{a\_b} to boolean representation, before they are
-input to the comparison and multiplexer circuits.
+input to the comparison and multiplexer circuits. \aseem{Question: why
+  is b arithmetic to begin with? Can this input be boolean and then we
+  don't have to insert a conversion gate for it?}
 %\vspace{-0.2in}
@@
 \subsection{Comparison with Yao-only approach}
+\aseem{We decided to put it here, but reading now, this subsection
+  seems to be a bit misplaced to me. How about moving it to 6.1, the
+  first subsection in the
+  Evaluation? And we can add a couple of lines about it in the
+  introduction evaluation paragraph (We first show that mixed
+  protocols are much more efficient than Yao only by comparing with
+  blah blah). Here it seems quite abrupt, we
+  are talking about EzPC at a high-level, with compiler features
+  etc. and suddenly details about numbers, CIFAR and what not start
+  coming. Seems too soon.}
 Before we provide the formalization of our compiler, we show that for performance it it critical to use a mix of arithmetic and boolean circuits.
@@
 
-\subsection{Cryptographic cost-aware compiler}
-\label{sec:ccc}
-One of the main challenges for our compiler is to split the compute between arithmetic and boolean parts well. 
-As mentioned above, \tool is a cryptographic cost-aware compiler and uses the arithmetic and boolean costs of primitive operators and the inter-conversion costs to make the split.
+%% \subsection{Cryptographic cost-aware compiler}
+%% \label{sec:ccc}
+%% One of the main challenges for our compiler is to split the compute between arithmetic and boolean parts well. 
+%% As mentioned above, \tool is a cryptographic cost-aware compiler and uses the arithmetic and boolean costs of primitive operators and the inter-conversion costs to make the split.
 Cryptographic cost of converting an arithmetic share to a boolean share requires computing a garbled circuit for addition that is linear in size.
@@
 Since the performance gap between arithmetic and boolean multiplication is quite large, we believe that this choice is optimal for realistic network settings.
-However, this is not the case for addition.
+%% However, this is not the case for addition.
 In \tool, if both the operands of addition are boolean shared, we do the addition using boolean circuit, else we use arithmetic circuit.
 Hence, addition may be done in arithmetic or boolean depending on the operands.
-Finally, all the boolean operations such as comparison, shifts, multiplexors, etc are done using Yao.
-We note that doing an optimal split between arithmetic and boolean is an NP-hard problem and in this work, we use heuristics based on costs observed using microbenchmarking. 
-For the smaller benchmarks of our evaluation, we wrote ABY code directly and the circuits
-generated by our compiler had about the same size as the  hand written code. 
+%% Finally, all the boolean operations such as comparison, shifts, multiplexors, etc are done using Yao.
+%% We note that doing an optimal split between arithmetic and boolean is an NP-hard problem and in this work, we use heuristics based on costs observed using microbenchmarking. 
+%% For the smaller benchmarks of our evaluation, we wrote ABY code directly and the circuits
+%% generated by our compiler had about the same size as the  hand written code. 
 %Moreover, for our machine learning benchmarks (such as linear classifier, DNNs, etc) these choices seem optimal.
+
+\section{Cryptographic cost-awareness}
+
+In this section we explain various heuristics that \tool uses to split
+the secret computation in a program between arithmetic and
+boolean. While an optimal split in general is an NP-hard problem
+\aseem{can we cite something?}, \tool heuristics are empirically based
+on micro-benchmarking, and perform well on the various secure
+prediction benchmarks (Section~\ref{sec:eval}).
+
+\paragraph{Micro-benchmarking.} The cost of splitting the compuation
+between arithmetic and boolean involves the cost of operations
+(addition, multiplication, etc.) and the cost of inter-conversion
+gates between arithmetic and boolean. Cryptographic cost of converting
+an arithmetic share to a boolean share requires computing a garbled
+circuit for addition that is linear in size. Similarly, the cost of
+converting from boolean to arithmetic is computing a subtraction
+circuit inside garbled circuit that is again linear in size. Demmler
+et al.~\cite{aby} analyze all these costs by running
+microbenchmarks for basic operations and interconversions (Figures
+2 and 3 in \cite{aby}). \tool heuristics are based on these results.
+
+\paragraph{Multiplication.} For multiplication, \tool always chooses
+arithmetic circuit, as the cost of boolean multiplication is much
+higher than the cost of coercing the operands from boolean to
+arithmetic, performing an arithmetic multiplication, and then coercing
+the result back to boolean ($\sim9x$ more time $\&$ $\sim20x$ more
+communication in a LAN setting). Since the performance gap is quite
+large, we believe that this choice is optimal for realistic network
+settings.
+
+\paragraph{Addition.} However, this is not the case for addition. If
+both the operands of addition are boolean shared, \tool performs
+addition using a boolean circuit, else it uses an arithmetic
+circuit. \aseem{What if one operand is in boolean and the other is in
+  arithmetic?} Hence, addition may be done in arithmetic or boolean
+depending on the operands.
+
+\paragraph{Boolean operations.} Finally, \tool chooses Yao for for all
+the boolean operations such as comparison, shifts, multiplexors,
+etc. \aseem{Can we write a sentence here that Demmler et al. show this
+  is the optimal choice?}
 

### 2018-05-07 ccs/intro.tex
tweaks

@@
  purpose, high-level programming language and a
- tool, e.g., a compiler would generate an efficient protocol that
+ tool, e.g. a compiler, would generate an efficient protocol that
  implements the functionality securely, while hiding the cryptography
@@
 $\cdot^{T}$ is the matrix transpose operator, and $w^Tx$ denotes the
-inner product of  $w$ and $x$. Alice has the following options.
+inner product of  $w^T$ and $x$. Alice has the following options.
 
@@
 Standard compilers do not attempt to reason about or insert non-trivial type casts
-and require the programmers to perform them explicitly. 
+and require the programmers to perform them explicitly. \aseem{This
+last sentence and the bit about unique compilation challenges seems
+unnecessary to me. Compilers do insert type casts automatically. Also
+at this point in the paper the words ``type casts'' don't mean much.}
 We back \tool with formal semantics that enable the compiler to
 insert type conversions between boolean and arithmetic components
-while maintaining provable correctness, security, and efficiency.
+while maintaining provable correctness, security, and
+efficiency. \aseem{Provable efficiency? Perhaps provable should
+be rescoped. And again, type conversions may not be clear at this
+point. I would simply say that EzPC inserts inter-conversion gates
+automatically. It is backed by a formal model and provides
+provable correctness and security, and efficiency.}
 The automatically generated protocols have performance comparable to
@@
 evaluation, partitioning allows us to program large applications
-in \tool.
+in \tool. \aseem{What about the differences from Yao pipelining? Worth
+adding here that it is a backend-agnostic technique?}
 
@@
 the security theorem, we formally reduce the security
-of our scheme against semi-honest (or, ``honest but curious")
+of our scheme against semi-honest (or ``honest but curious")
 adversaries to the semi honest security of the  \mpc back-end. The

### 2018-05-08 ccs/intro.tex
tweaks to intro

@@
 stands for Easy 2 Party Computation.}, the first
-``cryptographic-cost aware'' framework that generates efficient and
-scalable \mpc protocols from high-level programs devoid of any
-cryptographic details. 
-The generated protocols use combinations of
-arithmetic and boolean circuits that introduce unique compilation challenges.
-Standard compilers do not attempt to reason about or insert non-trivial type casts
-and require the programmers to perform them explicitly. \aseem{This
-last sentence and the bit about unique compilation challenges seems
-unnecessary to me. Compilers do insert type casts automatically. Also
-at this point in the paper the words ``type casts'' don't mean much.}
-We back \tool with formal semantics that enable the compiler to
-insert type conversions between boolean and arithmetic components
-while maintaining provable correctness, security, and
-efficiency. \aseem{Provable efficiency? Perhaps provable should
-be rescoped. And again, type conversions may not be clear at this
-point. I would simply say that EzPC inserts inter-conversion gates
-automatically. It is backed by a formal model and provides
-provable correctness and security, and efficiency.}
-The automatically generated protocols have performance comparable to
+``cryptographic-cost aware'' compiler that generates efficient and
+scalable \mpc protocols using combinations of
+arithmetic and boolean circuits. \tool is backed by a formal model
+that enables it to choose arithmetic or boolean representations for
+different parts of
+the program, while automatically inserting inter-conversion gates as
+necessary. In addition to guiding the compiler, the formal model also
+provides strong correctness and
+security theorems. Our comprehensive evaluation shows that the
+automatically generated protocols have performance comparable to
 or better than the custom, specialized
@@
 that generic \mpc implementations are much more efficient than what they were believed to be. 
- %(up to a factor of 19x in our
-%experiments, depending on the functionality) \nc{I added this but I
-%  guess we are saying it many times, but might be nice to highlight
-%  perf early no?} \aseem{My rationale was to save the numbers for
-%  Evaluation summary and remain high-level here.} 
 Below we describe
 the salient features of \tool.
+
+%% that introduce unique compilation challenges.
+%% Standard compilers do not attempt to reason about or insert non-trivial type casts
+%% and require the programmers to perform them explicitly. \aseem{This
+%% last sentence and the bit about unique compilation challenges seems
+%% unnecessary to me. Compilers do insert type casts automatically. Also
+%% at this point in the paper the words ``type casts'' don't mean much.}
+%% We back \tool with formal semantics that enable the compiler to
+%% insert type conversions between boolean and arithmetic components
+%% while maintaining provable correctness, security, and
+%% efficiency. \aseem{Provable efficiency? Perhaps provable should
+%% be rescoped. And again, type conversions may not be clear at this
+%% point. I would simply say that EzPC inserts inter-conversion gates
+%% automatically. It is backed by a formal model and provides
+%% provable correctness and security, and efficiency.}
+
+%% This paper presents \tool\footnote{Read as ``easy peasy'',
+%% stands for Easy 2 Party Computation.}, the first
+%% ``cryptographic-cost aware'' framework that generates efficient and
+%% scalable \mpc protocols from high-level programs devoid of any
+%% cryptographic details. 
+%% The generated protocols use combinations of
+%% arithmetic and boolean circuits that introduce unique compilation challenges.
+%% Standard compilers do not attempt to reason about or insert non-trivial type casts
+%% and require the programmers to perform them explicitly. \aseem{This
+%% last sentence and the bit about unique compilation challenges seems
+%% unnecessary to me. Compilers do insert type casts automatically. Also
+%% at this point in the paper the words ``type casts'' don't mean much.}
+%% We back \tool with formal semantics that enable the compiler to
+%% insert type conversions between boolean and arithmetic components
+%% while maintaining provable correctness, security, and
+%% efficiency. \aseem{Provable efficiency? Perhaps provable should
+%% be rescoped. And again, type conversions may not be clear at this
+%% point. I would simply say that EzPC inserts inter-conversion gates
+%% automatically. It is backed by a formal model and provides
+%% provable correctness and security, and efficiency.}
+%% The automatically generated protocols have performance comparable to
+%% or better than the custom, specialized
+%% protocols from previous works~\cite{shafindss,wu,minionn,secureml,cryptonets,valeriaMatrix}.
+%% In fact, these papers (and others) cite the inefficiency of generic \mpc as the major motivation
+%% behind the design of specialized protocols. Using \tool, we empirically  demonstrate
+%% that generic \mpc implementations are much more efficient than what they were believed to be. 
+%%  %(up to a factor of 19x in our
+%% %experiments, depending on the functionality) \nc{I added this but I
+%% %  guess we are saying it many times, but might be nice to highlight
+%% %  perf early no?} \aseem{My rationale was to save the numbers for
+%% %  Evaluation summary and remain high-level here.} 
+%% Below we describe
+%% the salient features of \tool.
 

### 2018-05-08 ccs/eval.tex
half pass

@@
 of \tool source code required is still small. 
-We note that while the \minion implementation is based on the ABY framework, it does not use the framework ``off-the-shelf'' and provide tailor-made optimizations for their applications.
-By contrast, \tool focuses on generic
-2PC, exploiting ABY's performant implementation. Optimizations such as those found in \minion are part of our future work.
+%We note that while the \minion implementation is based on the ABY framework, it does not use the framework ``off-the-shelf'' and provide tailor-made optimizations for their applications.
+%By contrast, \tool focuses on generic
+%2PC, exploiting ABY's performant implementation. Optimizations such as those found in \minion are part of our future work.
 \minion also reports performance results on a bigger DNN with 7
@@
 95.7\%). 
-The \bonsai models are already over integers and no porting from
-floating-point is required.
+%The \bonsai models are already over integers and no porting from
+%floating-point is required.
 We implement the trained classifiers in \tool for all the benchmarks
@@
 \label{sec:pipeeval}
-\aseem{Is it worth splitting it into microbenchamrking first, which is
-  Fig. 11, and then comparison with others? Right now, Fig. 11 text
-  just comes at the end, should be highlight it in a paragraph or
-  something?}
+%\aseem{Is it worth splitting it into microbenchamrking first, which is
+ % Fig. 11, and then comparison with others? Right now, Fig. 11 text
+ % just comes at the end, should be highlight it in a paragraph or
+ % something?}
 The largest benchmark of \minion~\cite{minionn} is a DNN for CIFAR-10
@@
 out of memory but an appropriately partitioned \tool implementation would still
-succeed. For example, consider a large DNN with $n$ layers such that
-the circuit for only a single layer can  fit into memory. Then, \tool
-using partitioning can evaluate this large DNN in $n$ stages. To
-further illustrate the need for and utility of partitioning, we ran a
-series of DNNs with different layers with and without partitioning
-(where partitioning was done for every layer of the DNN). For up to
-5-layer DNNs, the performance of both settings was comparable, thereby
-illustrating that partitioning did not cause any overhead to
-performance. For networks that had 6 or more layers, performance of
-the protocol without partitioning dramatically dropped, while the
-protocol with partitioning continued to scale linearly. These results
-are illustrated in Figure~\ref{fig:partitioning}.
+succeed.
+
+\paragraph{Scalability.}  To
+illustrate the scalability of partitioning, we evaluate a
+sequence of DNNs with and without partitioning
+(where every layer of each DNN is a separate stage) in~\figureref{partitioning}.
+For DNNs with up to 4 (hidden) layers, the performance is almost idential and the lines overlap, thereby
+illustrating that partitioning did not cause any noticeable performance overheads.
+Memory issues start showing up with the 5 (hidden) layer DNN and the non-partitioned implementation is slower.
+DNNs with 6 or more layers fail to execute (terminate with a ``bus error").
+However, the partitioned implementation scales well to even these large DNNs.
 
@@
 \subsection{Subsequent Work}
-Subsequent to our work, \aseem{TODO} et al. \aseem{cite} have
-presented Gazelle, a framework for specialized DNN
-protocols. Gazelle use a lattice-based packed additively homomorphic
+Subsequent to our work, Juvekar et al.~\cite{gazelle}  have
+presented {\sc Gazelle}, a specialized protocol for DNNs.
+{\sc Gazelle} use a lattice-based packed additively homomorphic
 encryption scheme (PAHE) for arithmetic computations and garbled
-circuits for boolean computations. Gazelle can evaluate the CNN benchmark of
+circuits for boolean computations. {\sc Gazelle} can evaluate the CNN benchmark of
 Table~\ref{tab:nn} in 0.8 seconds, as opposed to
@@
 cryptographic backends are orthogonal to our contributions. In
-particular, once Gazelle is available, we could add it as another
+particular, once {\sc Gazelle} is available, we could add it as another
 cryptographic backend to \tool. Furthermore, the authors of Gazelle remark:

### 2018-05-08 ccs/ex.tex
nit
half pass

@@
 instruments these conversion gates accordingly. For example, in
-lines~\ref{line:convacc} and~\ref{line:convb}, the compiler converts
-\ls{a\_acc} and \ls{a\_b} to boolean representation, before they are
-input to the comparison and multiplexer circuits. \aseem{Question: why
-  is b arithmetic to begin with? Can this input be boolean and then we
-  don't have to insert a conversion gate for it?}
+line~\ref{line:convacc}, the compiler converts
+\ls{a\_acc} to a boolean representation, before it is
+input to the comparison and multiplexer circuits.
 %\vspace{-0.2in}
@@
 share\;*y\_acc = ycirc\rightarrow PutA2YGate(a\_acc);~\label{line:convacc}
-}$
-$\mathtt{
-share\;*y\_b = ycirc\rightarrow PutA2YGate(a\_b);~\label{line:convb}
 }$
@@
 \subsection{Comparison with Yao-only approach}
-\aseem{We decided to put it here, but reading now, this subsection
-  seems to be a bit misplaced to me. How about moving it to 6.1, the
-  first subsection in the
-  Evaluation? And we can add a couple of lines about it in the
-  introduction evaluation paragraph (We first show that mixed
-  protocols are much more efficient than Yao only by comparing with
-  blah blah). Here it seems quite abrupt, we
-  are talking about EzPC at a high-level, with compiler features
-  etc. and suddenly details about numbers, CIFAR and what not start
-  coming. Seems too soon.}
-Before we provide the formalization of our compiler, we show that for performance it it critical to use a mix of arithmetic and boolean circuits.
-We demonstrate the performance benefits of mixed computation over boolean-only compute by comparing with previous tools CBMC-GC~\cite{cbmcgc} and ObliVM~\cite{oblivm} that are based on Yao's garbled circuits.
-As we already mentioned,  we are not the first to observe that Yao-only approach does not scale well beyond examples with small numbers of multiplications \cite{aby,tasty,autoS,secureml,minionn,chameleon}, especially for large-scale multiplications commonly found in Machine Learning (ML) algorithms -- the main use case of \tool.
-Roughly, multiplying two $64$-bit integers using arithmetic sharing requires only $2$ multiplications, whereas Yao requires at least $4096$ AES operations in the online phase.
-Indeed, this is one of the main drivers behind development of various specialized MPC protocols in previous works \cite{secureml,minionn,shafindss} that we compare against in \sectionref{eval}.
+\label{sec:oblivm}
+We show that it is critical to use a mix of arithmetic and boolean circuits  for performance.
+ Previous work has observed that Yao's garbled circuits do not scale to machine learning examples that require a large number of multiplications~\cite{aby,tasty,autoS,secureml,minionn,chameleon}.
+Indeed, this is one of the main drivers behind the development of various specialized MPC protocols in previous works~\cite{secureml,minionn,shafindss} (we compare against them in~\sectionref{eval}) .
+Here, we empirically demonstrate the performance benefits of mixed computations over boolean-only compute by comparing with state of the art Yao-based compilers CBMC-GC~\cite{cbmcgc} and ObliVM~\cite{oblivm}.
 
-The largest example benchmark in CBMC-GC is multiplication of two $8\times 8$ matrices for which it generates about a million gates and takes about ten seconds to run.
-In contrast, in \tool we use arithmetic sharing for this compute that generates 1218 gates and runs in about 0.1 seconds.
-We tried running CBMC-GC for multiplying two $20\times 20$ matrices and it timed out after 5 hours.
-In contrast, most of our benchmarks are much bigger.
-For example, we run the secure computation of a DNN for CIFAR-10 in 265 seconds which requires multiplying a $64\times576$ matrix with a $576\times1024$ matrix as a sub-computation.
-On the other hand, ObliVM scales to larger benchmarks with its use of Yao-pipelining \cite{yao-pipe}.
-We ran inner-product computations of varying lengths followed by a comparison in both ObliVM and \tool (see Table~\cite{oblivm-compare} in the Appendix).
-Our experiments show that we are atleast 30x faster than ObliVM due to our use of arithmetic sharing for inner-product computation.
+The largest benchmark in CBMC-GC is a multiplication of two $8\times 8$ matrices for which it generates about a million gates and takes about ten seconds to run.
+In contrast, \tool uses arithmetic sharing, generates 1218 gates, and runs in less than 0.1 seconds.
+When we tried multiplying two $20\times 20$ matrices with CBMC-GC, it timed out after 5 hours.
+Our benchmarks require much bigger computations (e.g., multiplying a $64\times576$ matrix with a $576\times1024$ matrix).
+Unlike CBMC-GC, ObliVM can scale to larger benchmarks (because it uses Yao-pipelining \cite{yao-pipe}).
+We evaluated the program in \figureref{ex-sml} with vectors of lengths varying between $10^3$ and $10^6$ using both \tool and ObliVM.
+\tool evaluates the inner-product in arithmetic and, empirically, is at least 25x faster than ObliVM.
 
@@
 \section{Cryptographic cost-awareness}
+ \label{sec:ccc}
+In this section, we explain various heuristics that \tool uses to split
+the computation in a program into arithmetic and
+boolean parts. While finding an optimum split  is an NP-hard problem (the predicates in {\tt if} statements can have arbitrary non-linear arithmetic),
+\tool uses heuristics based
+on empirical micro-benchmarking that  perform well in practice (Section~\ref{sec:eval}).
 
-In this section we explain various heuristics that \tool uses to split
-the secret computation in a program between arithmetic and
-boolean. While an optimal split in general is an NP-hard problem
-\aseem{can we cite something?}, \tool heuristics are empirically based
-on micro-benchmarking, and perform well on the various secure
-prediction benchmarks (Section~\ref{sec:eval}).
+The split
+between arithmetic and boolean requires knowing the cost of
+individual operations
+(addition, multiplication, inter-conversion between arithmetic and boolean, etc.).
+Demmler
+et al.~\cite{aby} document these costs by running
+microbenchmarks for basic operations and interconversions (Figures
+2 and 3 in \cite{aby}). \tool heuristics are based on their results.
 
-\paragraph{Micro-benchmarking.} The cost of splitting the compuation
-between arithmetic and boolean involves the cost of operations
-(addition, multiplication, etc.) and the cost of inter-conversion
-gates between arithmetic and boolean. Cryptographic cost of converting
+ Converting
 an arithmetic share to a boolean share requires computing a garbled
-circuit for addition that is linear in size. Similarly, the cost of
-converting from boolean to arithmetic is computing a subtraction
-circuit inside garbled circuit that is again linear in size. Demmler
-et al.~\cite{aby} analyze all these costs by running
-microbenchmarks for basic operations and interconversions (Figures
-2 and 3 in \cite{aby}). \tool heuristics are based on these results.
+circuit for addition. The size of this circuit grows  linearly in the bitwidth of the inputs. Similarly, converting from boolean to arithmetic requires computing a 
+garbled circuit for subtraction, which is also linear. 
+Since each conversion has roughly the same cost as a boolean addition,
+ \tool performs
+addition using a boolean circuit
+if the operands  are boolean shared, else it uses an arithmetic
+circuit.
 
-\paragraph{Multiplication.} For multiplication, \tool always chooses
-arithmetic circuit, as the cost of boolean multiplication is much
-higher than the cost of coercing the operands from boolean to
-arithmetic, performing an arithmetic multiplication, and then coercing
+For multiplication, \tool always chooses
+an arithmetic circuit, as the cost of a boolean multiplication is much
+higher than the cost of converting the operands from boolean to
+arithmetic, performing an arithmetic multiplication, and then converting
 the result back to boolean ($\sim9x$ more time $\&$ $\sim20x$ more
-communication in a LAN setting). Since the performance gap is quite
-large, we believe that this choice is optimal for realistic network
-settings.
-
-\paragraph{Addition.} However, this is not the case for addition. If
-both the operands of addition are boolean shared, \tool performs
-addition using a boolean circuit, else it uses an arithmetic
-circuit. \aseem{What if one operand is in boolean and the other is in
-  arithmetic?} Hence, addition may be done in arithmetic or boolean
-depending on the operands.
-
-\paragraph{Boolean operations.} Finally, \tool chooses Yao for for all
-the boolean operations such as comparison, shifts, multiplexors,
-etc. \aseem{Can we write a sentence here that Demmler et al. show this
-  is the optimal choice?}
+communication in a LAN setting). Theoretically, the size of a boolean multiplication circuit is quadratic in the bitwidth, which causes this performance gap.
+Roughly, multiplying two $64$-bit integers using arithmetic sharing requires only $2$ multiplications, whereas Yao requires at least $4096$ AES operations in the online phase.
+Since this gap is quite large, we believe that this choice is optimal for realistic network
+settings. Finally, \tool chooses boolean circuits for all
+the operations lacking arithmetic support in ABY, e.g., comparisons, bit-shifts, etc.
 

### 2018-05-08 ccs/impl.tex
half pass

@@
 We use an off-the-shelf solver
-(SeaHorn~\cite{seahorn}) to bound check the array indices
-($\models e < n$
+(SeaHorn~\cite{seahorn}) to check that the array indices
+are within bounds ($\models e < n$
 in ({\sc {T-Read}}) and ({\sc{T-Write}}),
@@
 than a minute on our largest benchmark to verify that all the array
-accesses are in-bounds. This C program can also enable
+accesses are in-bounds. This C program also enables
 validation of \tool generated protocols via differential testing~\cite{mckeeman,frigate}.
@@
 Since \tool is aware of the cryptographic costs associated with these coercions,
-it tries to minimize them using several optimizations.
-For instance, if boolean shares of a variable are available in scope and the variable is involved in
-an addition then \tool performs the addition in boolean rather than first coercing boolean to arithmetic
-and then performing an arithmetic addition. 
-%However, multiplication is always performed in arithmetic as the cost of a boolean multiplication is much higher than the cost of coercing from boolean to arithmetic, performing an arithmetic multiplication, and then coercing from arithmetic to boolean~\cite{aby}.
-Other optimizations include the standard ``common subexpression elimination"
+it tries to minimize them using several optimizations, e.g., by
+%For instance, if boolean shares of a variable are available in scope and the %variable is involved in
+%an addition then \tool performs the addition in boolean rather than first %coercing boolean to arithmetic
+%and then performing an arithmetic addition. 
+%However, multiplication is always performed in arithmetic as the cost of a %boolean multiplication is much higher than the cost of coercing from boolean %to arithmetic, performing an arithmetic multiplication, and then coercing from %arithmetic to boolean~\cite{aby}.
+%Other optimizations include
+the standard ``common subexpression elimination"
 optimization~\cite{dragonbook}.
@@
 instead of inserting code to recompute them afresh.
-
-All these optimizations are standard compiler
+These optimizations are standard compiler
 optimizations~\cite{dragonbook}, and we rely on their correctness to
-maintain the correctness and security for the optimized programs.
+maintain the correctness and security of the optimized programs.

### 2018-05-08 ccs/ld.tex
pass over paper. added partitioning figure.
half pass

@@
 The readers who are interested in using the compiler as a black box can
-move directly to  Section~\ref{sec:pipe}.
+move directly to  Section~\ref{sec:pipe} without loss of continuity.
 %\dg{Put a warning that this section is technical and not prerequisite for understanding later section.}
@@
 
-\paragraph{Source language.} Our source language
+\paragraph{Source language.} Our language
  is a simple imperative language shown in \figureref{srclang}. Types

### 2018-05-08 ccs/eval.tex
forgot

@@
 illustrating that partitioning did not cause any noticeable performance overheads.
-Memory issues start showing up with the 5 (hidden) layer DNN and the non-partitioned implementation is slower.
+Memory issues start showing up in the non-partitioned implementation with the 5 layer DNN and it is slower.
+Performance degrades rapidly thereafter and
 DNNs with 6 or more layers fail to execute (terminate with a ``bus error").

### 2018-05-08 ccs/conclude.tex
half pass
forgot

@@
 Currently, we are working on a front-end to translate Tensorflow code to \tool.
-The aim here is to provide a push button implementation to data scientists that generates secure implementations for their existing models.
-In the future, we would like to extend our security guarantees to malicious adversaries. However, we are currently unaware of a maliciously secure \mpc implementation for combinations of
+The aim here is to provide a push button implementation  that generates secure implementations for existing Tensorflow models.
+In the future, we would like to extend our security guarantees to malicious adversaries. 
+ The cryptographic backends continue to improve and the modular design of \tool makes it easy to integrate with the best available backends.
+However, we are currently unaware of a maliciously secure \mpc implementation for combinations of
 arithmetic and boolean circuits.
-The cryptographic backends continue to improve and the modular design of \tool makes it easy to integrate with new backends.
-We will continue integrating with the best available backends.
 Finally, we will explore the possibility of mechanically verifying the implementation of our compiler.

### 2018-05-08 ccs/eval.tex
minor

@@
 of \tool source code required is still small. 
-%We note that while the \minion implementation is based on the ABY framework, it does not use the framework ``off-the-shelf'' and provide tailor-made optimizations for their applications.
-%By contrast, \tool focuses on generic
-%2PC, exploiting ABY's performant implementation. Optimizations such as those found in \minion are part of our future work.
+We note that while the \minion implementation is based on the ABY framework, it does not use ABY ``off-the-shelf'' and performs application-specific optimizations.
+In contrast, \tool focuses on generic
+\mpc and directly exploits the exisiting performant implementations in ABY.
+% Optimizations such as those found in \minion are part of our future work.
 \minion also reports performance results on a bigger DNN with 7
@@
 particular, once {\sc Gazelle} is available, we could add it as another
-cryptographic backend to \tool. Furthermore, the authors of Gazelle remark:
+cryptographic backend to \tool. Furthermore, the authors of {\sc Gazelle} remark:
 ``A final, very interesting and ambitious line

### 2018-05-08 ccs/related.tex
minor

@@
 %The reason for doing this was to allow for a mix between arithmetic and boolean computations as a secure computation protocol for a pure boolean circuit would have been very inefficient. 
-This requires deep knowledge of cryptography
-to ensure security.
-Examples include classification of medical ElectroCardioGram (ECG)~\cite{barni}, iris and fingerprint recognition~\cite{blanton}, remote diagnostics~\cite{brickell}, sequence analysis~\cite{franz}, biometric identification~\cite{huang}, matrix factorization and ridge-regression~\cite{valeriaMatrix, valeriaRidge}, machine-learning classification~\cite{shafindss}, decision trees/forests~\cite{wu}, neural network training~\cite{secureml} and prediction~\cite{minionn}.
+%This requires deep knowledge of cryptography
+%to ensure security.
+ Examples include ~\cite{barni,blanton,brickell,franz,huang,valeriaMatrix, valeriaRidge,shafindss,wu,secureml,minionn}.
+
+%Examples include classification of medical ElectroCardioGram (ECG)~\cite{barni}, iris and fingerprint recognition~\cite{blanton}, remote diagnostics~\cite{brickell}, sequence analysis~\cite{franz}, biometric identification~\cite{huang}, matrix factorization and ridge-regression~\cite{valeriaMatrix, valeriaRidge}, machine-learning classification~\cite{shafindss}, decision trees/forests~\cite{wu}, neural network training~\cite{secureml} and prediction~\cite{minionn}.
 

### 2018-05-09 ccs/eval.tex
spell check

@@
 In contrast, \tool focuses on generic
-\mpc and directly exploits the exisiting performant implementations in ABY.
+\mpc and directly exploits the existing performant implementations in ABY.
 % Optimizations such as those found in \minion are part of our future work.
@@
 (where every layer of each DNN is a separate stage) in~\figureref{partitioning}.
-For DNNs with up to 4 (hidden) layers, the performance is almost idential and the lines overlap, thereby
+For DNNs with up to 4 (hidden) layers, the performance is almost identical and the lines overlap, thereby
 illustrating that partitioning did not cause any noticeable performance overheads.
@@
 because of  Batcher's sort (450 lines of \tool code in each stage).
-However, the current programmer effort seems miniscule compared to the mammoth implementation effort
+However, the current programmer effort seems minuscule compared to the mammoth implementation effort
 put in by Nikolaenko et al. (Section 5 of ~\cite{valeriaMatrix}) to scale a boolean circuits based

### 2018-05-09 ccs/ex.tex
spell check

@@
 
-\subsection{Comparison with Yao-only approach}
+\subsection{Comparison with garbled circuits}
 \label{sec:oblivm}
@@
  Previous work has observed that Yao's garbled circuits do not scale to machine learning examples that require a large number of multiplications~\cite{aby,tasty,autoS,secureml,minionn,chameleon}.
-Indeed, this is one of the main drivers behind the development of various specialized MPC protocols in previous works~\cite{secureml,minionn,shafindss} (we compare against them in~\sectionref{eval}) .
+Indeed, this is one of the main drivers behind the development of various specialized \mpc protocols in previous works~\cite{secureml,minionn,shafindss} (we compare against them in~\sectionref{eval}) .
 Here, we empirically demonstrate the performance benefits of mixed computations over boolean-only compute by comparing with state of the art Yao-based compilers CBMC-GC~\cite{cbmcgc} and ObliVM~\cite{oblivm}.
@@
 an arithmetic share to a boolean share requires computing a garbled
-circuit for addition. The size of this circuit grows  linearly in the bitwidth of the inputs. Similarly, converting from boolean to arithmetic requires computing a 
+circuit for addition. The size of this circuit grows  linearly in the bit-width of the inputs. Similarly, converting from boolean to arithmetic requires computing a 
 garbled circuit for subtraction, which is also linear. 
@@
 the result back to boolean ($\sim9x$ more time $\&$ $\sim20x$ more
-communication in a LAN setting). Theoretically, the size of a boolean multiplication circuit is quadratic in the bitwidth, which causes this performance gap.
+communication in a LAN setting). Theoretically, the size of a boolean multiplication circuit is quadratic in the bit-width, which causes this performance gap.
 Roughly, multiplying two $64$-bit integers using arithmetic sharing requires only $2$ multiplications, whereas Yao requires at least $4096$ AES operations in the online phase.

### 2018-05-09 ccs/main.tex
minor
spell check

@@
 \fancyhf{} % Remove fancy page headers 
-\fancyhead[C]{Anonymous submission \#9999 to ACM CCS 2017} % TODO: replace 9999 with your paper number
+\fancyhead[C]{Anonymous submission \#410 to ACM CCS 2017} % TODO: replace 9999 with your paper number
 \fancyfoot[C]{\thepage}
@@
 
-\title{\tool: Programmable, Efficient, and Scalable \\ Secure Two-Party Computation \dg{for Machine Learning}}
+\title{\tool: Programmable, Efficient, and Scalable \\ Secure Two-Party Computation for Machine Learning}
 

### 2018-05-09 ccs/security.tex
spell check

@@
 At a very high level, in this
-framework, parties are modeled as non-uniform interactive turing
+framework, parties are modeled as non-uniform interactive Turing
 machines (ITMs), with inputs provided by an environment $\env$. An

### 2018-05-09 ccs/ex.tex
nits

@@
 the result back to boolean ($\sim9x$ more time $\&$ $\sim20x$ more
-communication in a LAN setting). Theoretically, the size of a boolean multiplication circuit is quadratic in the bit-width, which causes this performance gap.
+communication in a LAN setting). The size of a boolean multiplication circuit is quadratic in the bit-width, which causes this performance gap.
 Roughly, multiplying two $64$-bit integers using arithmetic sharing requires only $2$ multiplications, whereas Yao requires at least $4096$ AES operations in the online phase.
@@
 the operations lacking arithmetic support in ABY, e.g., comparisons, bit-shifts, etc.
+Further implementation details can be found in~\sectionref{impl}.
 

### 2018-05-09 ccs/impl.tex
nits

@@
 library~\cite{aby}. 
-ABY provides support for Arithmetic computation based on \cite{autoS}, and boolean computations based on GMW~\cite{gmw} as well as Yao's garbled circuits~\cite{yao}.
-Although \tool can generate code for both kinds of boolean computations, we have observed better performance when using the garbled circuits
-and use it in our evaluation. Hence, \tool generated code uses Arithmetic computation and garbled circuits based boolean computations.
+ABY provides support for Arithmetic computations based on \cite{autoS}, and boolean computations based on GMW~\cite{gmw} as well as Yao's garbled circuits~\cite{yao}.
+Although \tool can generate code for both kinds of boolean computations, we have observed better performance when using  garbled circuits
+and use it in our evaluation. Hence, \tool generated code uses Arithmetic computations and garbled circuits based boolean computations.
 We use $128$ bits of security and OT
@@
 These optimizations are standard compiler
-optimizations~\cite{dragonbook}, and we rely on their correctness to
-maintain the correctness and security of the optimized programs.
+optimizations~\cite{dragonbook}, and we rely on their correctness
+(optimizations preserve outputs and well-typedness) to
+maintain the security of the optimized programs.

### 2018-05-09 ccs/ex.tex
nit

@@
 We show that it is critical to use a mix of arithmetic and boolean circuits  for performance.
- Previous work has observed that Yao's garbled circuits do not scale to machine learning examples that require a large number of multiplications~\cite{aby,tasty,autoS,secureml,minionn,chameleon}.
+ Previous works have observed that Yao's garbled circuits do not scale to machine learning examples that require a large number of multiplications~\cite{aby,tasty,autoS,secureml,minionn,chameleon}.
 Indeed, this is one of the main drivers behind the development of various specialized \mpc protocols in previous works~\cite{secureml,minionn,shafindss} (we compare against them in~\sectionref{eval}) .
@@
 the computation in a program into arithmetic and
-boolean parts. While finding an optimum split  is an NP-hard problem (the predicates in {\tt if} statements can have arbitrary non-linear arithmetic),
-\tool uses heuristics based
-on empirical micro-benchmarking that  perform well in practice (Section~\ref{sec:eval}).
+boolean parts. Since finding an optimum split  is an NP-hard problem (the predicates in {\tt if} statements can have arbitrary non-linear arithmetic),
+\tool uses heuristics  that  perform well in practice (Section~\ref{sec:eval}).
 

### 2018-05-09 ccs/eval.tex
para

@@
 \subsection{Secure prediction}
-\paragraph{Standard classifiers.}
+\noindent\textbf{Standard classifiers.}
 %\label{sec:shallow}
@@
 
-\paragraph{Deep neural nets.}
+\noindent\textbf{Deep neural nets.}
 We evaluate \tool on the DNNs described in SecureML~\cite{secureml},
@@
 
-\paragraph{State-of-the-art classifiers.}
+\noindent\textbf{State-of-the-art classifiers.}
 Tensorflow~\cite{tensorflow} is a standard machine learning toolkit.
@@
 
-\paragraph{Scalability.}  To
+\noindent\textbf{Scalability.}  To
 illustrate the scalability of partitioning, we evaluate a
-sequence of DNNs with and without partitioning
-(where every layer of each DNN is a separate stage) in~\figureref{partitioning}.
-For DNNs with up to 4 (hidden) layers, the performance is almost identical and the lines overlap, thereby
-illustrating that partitioning did not cause any noticeable performance overheads.
-Memory issues start showing up in the non-partitioned implementation with the 5 layer DNN and it is slower.
+sequence of DNNs with and without partitioning in~\figureref{partitioning}.
+All layers are identical and partitioning places each layer in a separate stage. 
+For DNNs with up to 4  layers, the performance, with and without partitioning, is almost identical and the lines overlap, thereby
+illustrating that partitioning does not cause any noticeable performance overheads.
+Memory issues start showing up in the non-partitioned implementation of the 5 layer DNN and it is slower.
 Performance degrades rapidly thereafter and
@@
 \includegraphics[width=8cm]{partitioning.png}
-\caption{Comparison of \tool code with and without partitioning. x-axis denotes the number of layers of the DNN, while y-axis denotes time in seconds for the secure protocol.}
+\caption{Comparison of \tool code with and without partitioning. x-axis denotes the number of layers in the DNN, while y-axis denotes time in seconds for the secure protocol.}
 \label{fig:partitioning}

### 2018-05-09 ccs/ex.tex
para

@@
 
-\paragraph{Source language.}
+\noindent\textbf{Source language.}
 Consider the example $w^Tx >b$ from \sectionref{intro}, where
@@
 
-\paragraph{\tool compiler.}
+\noindent\textbf{\tool compiler.}
 \tool compiler takes as input a source program and produces a C++
@@
 
-\paragraph{Circuit generation and evaluation.} The next step in
+\noindent\textbf{Circuit generation and evaluation.} The next step in
 \tool is to
@@
 
-\paragraph{Advantages of \tool} We can now concretely see the advantages of \tool. Unarguably, it is
+\noindent\textbf{Advantages of \tool} We can now concretely see the advantages of \tool. Unarguably, it is
 easier for a developer to program and get the code right in

### 2018-05-09 ccs/intro.tex
half pass
para

@@
 
-\paragraph{Ease of programming.} \tool source programs are ideal
+\noindent\textbf{Ease of programming.} \tool source programs are ideal
 functionalities that describe ``what'' computation needs to be done,
@@
 
-\paragraph{Cryptographic-cost aware compiler.} The \tool compiler
+\noindent\textbf{Cryptographic-cost aware compiler.} The \tool compiler
 compiles a source program to a hybrid computation consisting
@@
 
-\paragraph{Scalability (secure code partitioning).} \mpc tools
+\noindent\textbf{Scalability (secure code partitioning).} \mpc tools
 often do not scale to large functionalities. The reason is that
@@
 evaluation, partitioning allows us to program large applications
-in \tool. \aseem{What about the differences from Yao pipelining? Worth
-adding here that it is a backend-agnostic technique?}
+in \tool. 
+%\aseem{What about the differences from Yao pipelining? Worth
+%adding here that it is a backend-agnostic technique?}
 
@@
 
-\paragraph{Formal guarantees.} We prove formal correctness and
+\noindent\textbf{Formal guarantees.} We prove formal correctness and
 security theorems for our compiler. The correctness theorem relates the
@@
 
-\paragraph{Evaluation.} We have implemented \tool
+\noindent\textbf{Evaluation.} We have implemented \tool
 using ABY~\cite{aby} as the
-cryptographic back-end. We evaluate \tool by implementing a wide range
+cryptographic back-end.
+We compare \tool with Yao-based compilers in~\sectionref{oblivm}
+and with specialized protocols in~\sectionref{eval}. 
+ We evaluate \tool by implementing a wide range
 of secure prediction benchmarks including linear and na\"{i}ve bayes
@@
 
-\paragraph{Related Work.}
+\noindent\textbf{Related Work.}
 Before ABY~\cite{aby}, several works have proposed

### 2018-05-09 ccs/ld.tex
para

@@
 
-\paragraph{Source language.} Our language
+\noindent\textbf{Source language.} Our language
  is a simple imperative language shown in \figureref{srclang}. Types
@@
 
-\paragraph{Source semantics.} The runtime semantics for the source language is shown in
+\noindent\textbf{Source semantics.} The runtime semantics for the source language is shown in
 \figureref{srcsem}. These semantics show how a ``trusted third party" computes the outputs when given the inputs of both the parties. 
@@
 
-\paragraph{Intermediate language.} \figureref{interlang} shows the
+\noindent\textbf{Intermediate language.} \figureref{interlang} shows the
 intermediate language of our compiler. The syntax follows that of the source
@@
 
-\paragraph{Source to intermediate compilation.} We provide the compilation
+\noindent\textbf{Source to intermediate compilation.} We provide the compilation
 rules in \figureref{compile}. We present the rules in a
@@
 
-\paragraph{Evaluation to Circuits.}
+\noindent\textbf{Evaluation to Circuits.}
 \figureref{circuits} shows the syntax of circuits. A wire id range
@@
 
-\paragraph{Circuit semantics.} Evaluating a program in the
+\noindent\textbf{Circuit semantics.} Evaluating a program in the
 intermediate language produces a
@@
 
-\paragraph{Correctness Theorem.} We prove that all well typed programs always terminate successfully (array indices are always in bounds, there are no unbounded loops, etc.) and the \mpc protocol produces the same outputs as the source program.
+\noindent\textbf{Correctness Theorem.} We prove that all well typed programs always terminate successfully (array indices are always in bounds, there are no unbounded loops, etc.) and the \mpc protocol produces the same outputs as the source program.
 That is, if a source statement $s$ is well-typed, and compiles to $\widetilde{s}$ in the intermediate language, then $s$ terminates in the source semantics with observations $O$, $\widetilde{s}$ evaluates to circuit $\crct$, and $\crct$ terminates in the circuit semantics with the same observations $O$.

### 2018-05-09 ccs/pipe.tex
pass over paper. added partitioning figure.
forgot
para

@@
 
-We  now describe our ``secure code partitioning'' technique that
+In this section, we describe our ``secure code partitioning'' technique that
 allows \tool to execute programs that require large circuits.
+Our techniques take inspiration from the idea of pipelining Yao's garbled circuits described in FastGC \cite{yao-pipe}.
+However, unlike FastGC, we do not operate at a circuit level and partitioning is independent of the specific
+\mpc protocol. 
 Let $\prog$ be a program in our source language that
@@
 not reveal any more information than $\prog$). 
-Our techniques take inspiration from the idea of pipelining Yao's garbled circuits described in FastGC \cite{yao-pipe}.
-However, unlike FastGC, we do not operate at a circuit level and partitioning is independent of the specific
-\mpc protocol. 
 %
@@
 
-\paragraph{Implementing code partitioning.}
+\noindent\textbf{Implementing code partitioning.}
 We use partitioning for programs that require large

### 2018-05-09 ccs/security.tex
para

@@
 %\vspace{-0.2in}
-\paragraph{Security theorem.} The protocols we generate provide
+\noindent\textbf{Security theorem.} The protocols we generate provide
 simulation-based security against a semi-honest adversary, in the

### 2018-07-29 ndss/main.tex
added ndss version.

--- 2018-05-09 ccs/main.tex
@@
-\documentclass[sigconf, anonymous]{acmart}
+\documentclass[conference]{IEEEtran}
 
-\fancyhf{} % Remove fancy page headers 
-\fancyhead[C]{Anonymous submission \#410 to ACM CCS 2017} % TODO: replace 9999 with your paper number
-\fancyfoot[C]{\thepage}
-
-\setcopyright{none} % No copyright notice required for submissions
-\acmConference[Anonymous Submission to ACM CCS 2017]{ACM Conference on Computer and Communications Security}{Due 19 May 2017}{Dallas, Texas}
-\acmYear{2017}
-
-\settopmatter{printacmref=false, printccs=true, printfolios=true} % We want page numbers on submissions
-
-%%\ccsPaper{9999} % TODO: replace with your paper number once obtained
-
+\pagestyle{plain}
 \input{headers}
@@
 \title{\tool: Programmable, Efficient, and Scalable \\ Secure Two-Party Computation for Machine Learning}
+
+%\setlength\intextsep{1pt}
+%\setlength{\textfloatsep}{1pt}
+%\setlength{\abovecaptionskip}{4pt}
+%\setlength{\belowcaptionskip}{4pt}
+
+\maketitle
 
@@
 
-\setlength\intextsep{1pt}
-\setlength{\textfloatsep}{1pt}
-\setlength{\abovecaptionskip}{4pt}
-\setlength{\belowcaptionskip}{4pt}
-
-\maketitle
-
 \graphicspath{{./Images/}}
@@
 
-{\footnotesize \bibliographystyle{acm}
-\bibliography{bib}}
+\bibliographystyle{IEEEtranS}
+\bibliography{bib}
 

### 2018-07-29 ndss/eval.tex
added ndss version.
edits to fit in 15 pages. removed appendix.

--- 2018-05-09 ccs/eval.tex
@@
 Most of our benchmarks
-are related to machine learning and we set up (largely standard) notation and describe our benchmarks in \appendixref{benchmarks}.
+are related to machine learning and we use largely standard notation.
+We provide further details on these benchmarks in the full version of this paper.
+%set up (largely standard) notation and describe our benchmarks in \appendixref{benchmarks}.
 

### 2018-07-29 ndss/ld.tex
added ndss version.
edits to fit in 15 pages. removed appendix.

--- 2018-05-09 ccs/ld.tex
@@
 %For the security of our compiler, we rely on the correctness guarantee and  reduce the security to the security of the underlying cryptographic protocol used to compute the generated circuit (see \theoremref{security}).
-We present only selected parts of our formalization. Full definitions
-and auxiliary lemmas can be found in~\appendixref{formal}. We will also
-release complete typeset proofs in a tech report.
+We present only selected parts of our formalization. Full definitions, auxiliary lemmas and complete proofs can be found in the full version of this paper. %We will also
+%release complete typeset proofs in a tech report.
 
@@
 
-We prove the theorem in Appendix~\ref{app:formal}.
+We prove the theorem in the full version. %Appendix~\ref{app:formal}.
 % We first define a judgment for

### 2018-07-29 ndss/pipe.tex
added ndss version.
edits to fit in 15 pages. removed appendix.

--- 2018-05-09 ccs/pipe.tex
@@
 %}
-More formal details follow, while an example illustrating the technique of secure code partitioning is provided in  \appendixref{codepartitioning}.
+More formal details follow, while an example illustrating the technique of secure code partitioning is provided in the full version.% \appendixref{codepartitioning}.
 

### 2018-07-29 ndss/eval.tex
ccs comments.

--- 2018-05-09 ccs/eval.tex
@@
 Most of our benchmarks
-are related to machine learning and we set up (largely standard) notation and describe our benchmarks in \appendixref{benchmarks}.
+are related to machine learning. We set up (largely standard) notation and describe our benchmarks in \appendixref{benchmarks}.
 

### 2018-07-29 ndss/ex.tex
added ndss version.
ccs comments.

--- 2018-05-09 ccs/ex.tex
@@
 (line~\ref{line:dotproductloop}).
+We remark that, in our experiments, assigning public labels to control flow variables is critical for performance and/or security.
+However, we have observed that conservatively creating circuits for other public variables incurs minimal cost in practice.
 
@@
 of ABY, boolean is synonymous with Yao, we elaborate in
-\sectionref{impl}). We elaborate on cryptographic cost awareness in
+\sectionref{impl}).
+At the expression level, \tool provides secret conditionals (using the '?
+:' operator) which are implemented using multiplexers, while secret dependent array accesses can also be implemented using the '?
+:' operator.
+We elaborate on cryptographic cost awareness in
 \sectionref{ccc}.
@@
 
-\noindent\textbf{Advantages of \tool} We can now concretely see the advantages of \tool. Unarguably, it is
+\noindent\textbf{Advantages of \tool.} We can now concretely see the advantages of \tool. Unarguably, it is
 easier for a developer to program and get the code right in
@@
 We evaluated the program in \figureref{ex-sml} with vectors of lengths varying between $10^3$ and $10^6$ using both \tool and ObliVM.
-\tool evaluates the inner-product in arithmetic and, empirically, is at least 25x faster than ObliVM.
+\tool evaluates the inner-product in arithmetic and, empirically, is at least 25x faster than ObliVM on the same hardware.
+
+\subsection{Comparison with handwritten ABY code}
+
+We have evaluated against handwritten ABY code for the examples in the ABY repository and \tool has comparable performance.
+However, these examples are not representative of practical ML tasks.
+Writing sophisticated ML applications such as neural networks in ABY is a very tedious and nearly impossible engineering task even for crypto experts -- automating the same is our main contribution.
 

### 2018-07-29 ndss/intro.tex
added ndss version.
ccs comments.

--- 2018-05-09 ccs/intro.tex
@@
 
+\subsection{Our Contributions}
 This paper presents \tool\footnote{Read as ``easy peasy'',
@@
 
-\noindent\textbf{Ease of programming.} \tool source programs are ideal
+\subsubsection{Ease of programming} \tool source programs are ideal
 functionalities that describe ``what'' computation needs to be done,
@@
 
-\noindent\textbf{Cryptographic-cost aware compiler.} The \tool compiler
+\subsubsection{Cryptographic-cost aware compiler} The \tool compiler
 compiles a source program to a hybrid computation consisting
@@
 
-\noindent\textbf{Scalability (secure code partitioning).} \mpc tools
+\subsubsection{Scalability (secure code partitioning)} \mpc tools
 often do not scale to large functionalities. The reason is that
@@
 
-\noindent\textbf{Formal guarantees.} We prove formal correctness and
+\subsubsection{Formal guarantees} We prove formal correctness and
 security theorems for our compiler. The correctness theorem relates the
@@
 
-\noindent\textbf{Evaluation.} We have implemented \tool
+\subsubsection{Evaluation}We implemented \tool
 using ABY~\cite{aby} as the
@@
 classifiers from Tensorflow~\cite{tensorflow}
-and \bonsai~\cite{bonsai}, and also the matrix factorization example
-from Nikolaenko et al.~\cite{valeriaMatrix}. Our results demonstrate
+and \bonsai~\cite{bonsai}, and also matrix factorization from Nikolaenko et al.~\cite{valeriaMatrix}. Our results demonstrate
 three key points. First, \tool makes it convenient for general
@@
 
-\noindent\textbf{Related Work.}
+\noindent\textbf{Summary.} To summarize the above five points, the novelty of \tool is in implementing cryptographic cost awareness inside a compiler, allowing the programmers to program in a high-level language absent of all the cryptographic details, while the compiler chooses the best circuit representation (boolean or arithmetic) for different parts of the program and also handles interconversions between them -- all this in a provably secure manner.
+Prior compilers such as Wysteria, ObliVM, and CBMC-GC use backends with single representations (either boolean or arithmetic), and so the  challenges associated with multiple representations and interconversions are absent.
+Moreover, as shown in Section~\sectionref{oblivm}, they perform poorly on ML tasks that involve a mix of arithmetic and boolean computations.
+The work of Kerschbaum {\em et al.}~\cite{autoS} integrates a backend with both boolean and arithmetic represenations.
+However it operates over circuits represented as a sequence of dyadic operations, while \tool provides a high-level programming language with loops and branches.
+\tool compilation is automatic and sub-quadratic in the size of generated ABY programs (e.g.
+\figureref{ex-aby}) after switching off memory safety checking.
+In sharp contrast, \cite{autoS} describes approaches that would be doubly exponential and exponential in the length of ABY programs.
+Finally, our partitioning technique, unlike prior work, which is specific to and works on Yao’s garbled circuits, works at the source level and is agnostic to the cryptographic backend.
+
+These features (ease-of-programming, use of multiple crypto backends, and scalability through partitioning) enable us to write and evaluate a wide array of ML algorithms such as matrix factorization with efficient generic 2PC techniques.
+Without \tool, this would have been a near-impossible task and has never been done.
+
+\subsection{Related Work}
 Before ABY~\cite{aby}, several works have proposed

### 2018-07-29 ndss/pipe.tex
ccs comments.

--- 2018-05-09 ccs/pipe.tex
@@
 allows \tool to execute programs that require large circuits.
+The reason to implement partitioning is as follows: when compiling from a high level functionality to a circuit being evaluated securely, the circuits so generated can quite often be larger than the total memory available.
+The secure protocol for evaluating this circuit cannot then be generated without swapping it in and out of main memory, thereby causing large slowdowns (see Section \ref{sec:pipeeval}).
+In order to avoid this problem, we ``split'' the high level functionality into smaller programs that compile into smaller circuits whose secure evaluation can completely fit into main memory.
+These smaller circuits are then evaluated sequentially to evaluate the entire program.
+Naturally, care must be taken to ensure that the partial evaluation of these circuits do not reveal any more information than the final output of the functionality.
+More details follow.
+
 Our techniques take inspiration from the idea of pipelining Yao's garbled circuits described in FastGC \cite{yao-pipe}.
@@
 %}
-More formal details follow, while an example illustrating the technique of secure code partitioning is provided in  \appendixref{codepartitioning}.
+More formal details follow, while an example illustrating the technique of secure code partitioning is provided in the full version. %\appendixref{codepartitioning}.
 
@@
 
+We stress that partitioning is fully automated and is compatible with any 2PC protocol.
+Furthermore, the \tool compiler, with suitable changes to operator costs, can target other backends including 3PC or MPC, or protocols secure against malicious adversaries.
+
 % \nc{I changed the heading.}

### 2018-07-29 ndss/pipe.tex
minor edit on partitioning

--- 2018-05-09 ccs/pipe.tex
@@
 allows \tool to execute programs that require large circuits.
+The reason to implement partitioning is as follows: when compiling from a high level functionality to a circuit being evaluated securely, the circuits so generated can quite often be larger than the total memory available on the machines executing 2PC.
+The secure protocol for evaluating this circuit cannot then be generated without swapping the circuit in and out of main memory, thereby causing large slowdowns (see Section \ref{sec:pipeeval}).
+In order to avoid this problem, we ``split'' the high level functionality into smaller programs that compile into small circuits whose secure evaluation can completely fit into main memory.
+These smaller circuits are then evaluated sequentially to evaluate the entire program.
+Naturally, care must be taken to ensure that the partial evaluation of these circuits do not reveal any more information than the final output of the overall functionality.
+More details follow.
+
 Our techniques take inspiration from the idea of pipelining Yao's garbled circuits described in FastGC \cite{yao-pipe}.

### 2018-08-02 ndss/intro.tex
minor

@@
 circuit~(\cite{yao,gmw}) or as an arithmetic circuit~(\cite{sss,viff,gentry}). 
+%
+\dg{Do we want to say sharemind? It is only 3PC.}
 %
@@
 Prior compilers such as Wysteria, ObliVM, and CBMC-GC use backends with single representations (either boolean or arithmetic), and so the  challenges associated with multiple representations and interconversions are absent.
-Moreover, as shown in Section~\sectionref{oblivm}, they perform poorly on ML tasks that involve a mix of arithmetic and boolean computations.
+Moreover, as shown in \sectionref{oblivm}, they perform poorly on ML tasks that involve a mix of arithmetic and boolean computations.
 The work of Kerschbaum {\em et al.}~\cite{autoS} integrates a backend with both boolean and arithmetic represenations.

### 2018-08-06 ndss/eval.tex
minor

@@
 2.40GHz with 28 GBs of RAM.
-When we compare our execution times with prior protocols, we match our system and network parameters with those of the prior work (as the code for works such as Bost et al.
-\cite{shafindss} and \minion~\cite{minionn} are not publicly available).
+When we compare our execution times with prior protocols, we match our system and network parameters with those of the prior work (these implementations are not publicly available).
 %

### 2018-08-06 ndss/intro.tex
minor

@@
 %
-\dg{Do we want to say sharemind? It is only 3PC.}
+%\dg{Do we want to say sharemind? It is only 3PC.}
 %
@@
 These situations seldomly arise in the machine learning applications that we consider:
-intermediate values don't need to be declassified and arrays are traversed in an oblivious manner. We provide a detailed survey of related
+intermediate values don't need to be declassified and arrays are mostly traversed in an oblivious manner. We provide a detailed survey of related
 work in \sectionref{related}.

### 2018-08-06 ndss/ld.tex
ccs comments.
minor

@@
 %For the security of our compiler, we rely on the correctness guarantee and  reduce the security to the security of the underlying cryptographic protocol used to compute the generated circuit (see \theoremref{security}).
-We present only selected parts of our formalization. Full definitions, auxiliary lemmas and complete proofs can be found in the full version of this paper. %We will also
+We present only selected parts of our formalization.
+Full definitions, auxiliary lemmas and complete proofs (that follow from a straight-forward induction argument once the key lemmas are identified) can be found in the full version of this paper.
+%We will also
 %release complete typeset proofs in a tech report.
@@
 
-We prove the theorem in the full version. %Appendix~\ref{app:formal}.
+We prove the theorem in the full version by induction on the compilation derivation. %Appendix~\ref{app:formal}.
 % We first define a judgment for

### 2018-08-07 ndss/conclude.tex
added ndss version.
comments from the email

--- 2018-05-08 ccs/conclude.tex
@@
 
-Currently, we are working on a front-end to translate Tensorflow code to \tool.
-The aim here is to provide a push button implementation  that generates secure implementations for existing Tensorflow models.
-In the future, we would like to extend our security guarantees to malicious adversaries. 
+Currently, we are working on a front-end to translate Tensorflow code
+to \tool. The aim here is to provide a push button implementation
+that generates secure implementations for existing Tensorflow models. 
+In the future, we would like to extend our security guarantees to
+malicious adversaries. We think that, with suitable
+changes to operator costs, \tool can target other backends including 3PC or
+MPC, or protocols secure against malicious adversaries.
  The cryptographic backends continue to improve and the modular design of \tool makes it easy to integrate with the best available backends.
@@
 arithmetic and boolean circuits.
-Finally, we will explore the possibility of mechanically verifying the implementation of our compiler.
+Finally, we will explore the possibility of mechanically verifying the  compiler implementation.
 

### 2018-08-07 ndss/ex.tex
comments from the email

@@
 components translate into regular C++ code, while the secret
-components translate into API calls into our \mpc back-end
+components translate into API calls into our \mpc backend
 (ABY). For example, in \figureref{ex-sml}, the \tool
@@
 (line~\ref{line:dotproductloop}).
-We remark that, in our experiments, assigning public labels to control flow variables is critical for performance and/or security.
-However, we have observed that conservatively creating circuits for other public variables incurs minimal cost in practice.
 
@@
 the other hand, since the comparison with \ls{b}, and the conditional
-expression computation are more efficient in the boolean
-representation, the \tool compiler uses the Yao circuit builder
-\ls{ycirc} to build the corresponding circuits
+expression (the '? :' operator) computation are more efficient in the
+boolean representation, the \tool compiler uses the Yao circuit
+builder \ls{ycirc} to build the corresponding comparison and
+multiplexer circuits
 (lines~\ref{line:condyaobegin} to~\ref{line:condyaoend}) (in our usage
@@
 \sectionref{impl}).
-At the expression level, \tool provides secret conditionals (using the '?
-:' operator) which are implemented using multiplexers, while secret dependent array accesses can also be implemented using the '?
+
+%% At the expression level, \tool provides secret conditionals  which are implemented using multiplexers, while secret dependent array accesses can also be implemented using the '?
 :' operator.
 We elaborate on cryptographic cost awareness in
-\sectionref{ccc}.
+%% \sectionref{ccc}.
 
@@
 %other high-level languages are limited in their use of a single,
-%arithmetic or boolean, crypto back-end). 
+%arithmetic or boolean, crypto backend). 
 \tool also enables the

### 2018-08-07 ndss/impl.tex
added ndss version.
comments from the email

--- 2018-05-09 ccs/impl.tex
@@
 
-Some of our benchmarks require accessing arrays at secret
-indices. While \tool enforces the array indices to be
-public, secret indices can be encoded in \tool using multiplexers.
+While most ML applications make data oblivious passes over the data,
+some of our benchmarks do require accessing arrays at secret
+indices. \tool enforces the array indices to be
+public, but secret indices can be encoded using multiplexers.
 For example, consider the expression $A[x]$ where $A$ is an array of
@@
 this functionality in \tool as $\cond{x > 0}{A[1]}{A[0]}$. In general,
-a secret access to an array of
-size $n$ requires ${n}-1$ multiplexers in \tool.
+a secret access to an array of size $n$ requires ${n}-1$ multiplexers
+in \tool.
 
@@
 All other variables are assigned arithmetic labels (that can later be
-coerced to boolean).
+coerced to boolean). While assigning public labels to control
+flow variables is critical for
+performance and/or security, assigning arithmetic labels to other
+variables incurs minimal cost in practice.
 We leave a more sophisticated type inference procedure for future work.
@@
 maintain the security of the optimized programs.
+
+\noindent\textbf{Implementing automatic code partitioning.}
+\tool decomposes the source program
+$\prog$ into a sequence of small programs
+$\prog_1\seq\dots\seq\prog_k$, and then appropriately instruments them
+to produce $\progn_1,\dots\progn_k$, as detailed in
+Section~\ref{sec:pipe}. It then
+compiles and executes the $k$ programs
+$\progn_1,\progn_2,\ldots,\progn_k$ sequentially, freeing up memory
+usage after execution of each $\progn_i$.
+Automating the  decomposition step requires an analysis that can 
+statically estimate the resource usage of a \tool program.
+Resource analysis of high-level programs is a well-known hard problem~\cite{raml}
+and we describe a heuristic analysis.
+
+To build $s_1$, we consider the longest
+prefix of $s$ whose computation {\it size} is below the threshold enforced by the available memory of the machine.
+If $\prog = \prog_1; \prog_r$ then we recurse on $\prog_r$ to obtain $\prog_2,\ldots,\prog_k$.
+For a program $\mprog$, to estimate ${\it size}(\mprog)$, we need to discuss three important cases: if $\mprog\equiv\mprog_1;\mprog_2$
+then ${\it size}(\mprog)={\it size}(\mprog_1)+{\it size}(\mprog_2)$; if $\mprog\equiv\ite{e_1}{\mprog_1}{\mprog_2}$ then ${\it size}(\mprog)=\max({\it size}(\mprog_1),{\it size}(\mprog_2))$; if $s\equiv\forl{i}{n_1}{n_2}{\mprog_1}$ then
+${\it size}(\mprog)=(n_2-n_1){\it size}(\mprog_1)$. If $(n_2-n_1){\it size}(\mprog_1)$ is above the threshold then
+we replace $\mprog$ by 
+\[\forl{i}{n_1}{\frac{n_1+n_2}{2}}{\mprog_1}; \forl{i}{\frac{n_1+n_2}{2}}{n_2}{\mprog_1}\]
+ and recurse to find the prefix again.
+This heuristic analysis is sufficient for the benchmarks discussed in
+our evaluation.

### 2018-08-07 ndss/intro.tex
comments from the email

@@
 Most notable of these domain specific languages and systems include \cite{...}.
-All of these works build on a cryptographic back-end that is either entirely boolean \cite{yao,gmw} or arithmetic \cite{homo}.
+All of these works build on a cryptographic backend that is either entirely boolean \cite{yao,gmw} or arithmetic \cite{homo}.
 In fact, most of these works use boolean circuits based scheme for completeness\footnote{Note that comparisons cannot be expressed in arithmetic circuits}.
@@
 
-%%%%%%%Recent works \cite{aby,secureml,minion} have shown that great performance benefit can be obtained by mixing arithmetic and boolean computations in the back-end.
-Certain computations such as integer multiplication are efficient when implemented using arithmetic SMC \cite{gmw} whereas max(x,0) needs to be computed using a boolean back-end \cite{yao}.
+%%%%%%%Recent works \cite{aby,secureml,minion} have shown that great performance benefit can be obtained by mixing arithmetic and boolean computations in the backend.
+Certain computations such as integer multiplication are efficient when implemented using arithmetic SMC \cite{gmw} whereas max(x,0) needs to be computed using a boolean backend \cite{yao}.
 More details later..
@@
 %computation. We build on ABY \cite{aby} that provides the suitable
-%cryptographic back-end in this setting. We provide a formal type
+%cryptographic backend in this setting. We provide a formal type
 %system and prove correctness and security of our compiler. We evaluate
@@
 
-%The programmer is oblivious of the cryptographic back-end being used.
+%The programmer is oblivious of the cryptographic backend being used.
 Our compiler automatically compiles it to a circuit framework consisting to both arithmetic and boolean gates as well share-conversion gates wherever required.
@@
 scalable \mpc protocols using combinations of
-arithmetic and boolean circuits. \tool is backed by a formal model
+arithmetic and boolean circuits. \tool source language is a simple
+imperative language absent of any cryptographic details. The compiler
+is
+backed by a formal model
 that enables it to choose arithmetic or boolean representations for
@@
 show-stopper for applications like secure machine learning, secure
-prediction, etc. that operate on large amounts of data.
+prediction, etc. that operate on large data.
 %\aseem{Can we
@@
 of our scheme against semi-honest (or ``honest but curious")
-adversaries to the semi honest security of the  \mpc back-end. The
+adversaries to the semi honest security of the  \mpc backend. The
 theorem provides protection against side-channels arising from
@@
 using ABY~\cite{aby} as the
-cryptographic back-end.
+cryptographic backend.
 We compare \tool with Yao-based compilers in~\sectionref{oblivm}
@@
 
-\noindent\textbf{Summary.} To summarize the above five points, the novelty of \tool is in implementing cryptographic cost awareness inside a compiler, allowing the programmers to program in a high-level language absent of all the cryptographic details, while the compiler chooses the best circuit representation (boolean or arithmetic) for different parts of the program and also handles interconversions between them -- all this in a provably secure manner.
-Prior compilers such as Wysteria, ObliVM, and CBMC-GC use backends with single representations (either boolean or arithmetic), and so the  challenges associated with multiple representations and interconversions are absent.
-Moreover, as shown in \sectionref{oblivm}, they perform poorly on ML tasks that involve a mix of arithmetic and boolean computations.
-The work of Kerschbaum {\em et al.}~\cite{autoS} integrates a backend with both boolean and arithmetic represenations.
-However it operates over circuits represented as a sequence of dyadic operations, while \tool provides a high-level programming language with loops and branches.
-\tool compilation is automatic and sub-quadratic in the size of generated ABY programs (e.g.
-\figureref{ex-aby}) after switching off memory safety checking.
-In sharp contrast, \cite{autoS} describes approaches that would be doubly exponential and exponential in the length of ABY programs.
-Finally, our partitioning technique, unlike prior work, which is specific to and works on Yao’s garbled circuits, works at the source level and is agnostic to the cryptographic backend.
-
-These features (ease-of-programming, use of multiple crypto backends, and scalability through partitioning) enable us to write and evaluate a wide array of ML algorithms such as matrix factorization with efficient generic 2PC techniques.
-Without \tool, this would have been a near-impossible task and has never been done.
+\noindent\textbf{Summary.} \tool is carefully designed to enable
+non-cryptographers (and cryptographer alike) to write efficient \mpc
+protocols for ML algorithms. With the novel combination of language
+design, cryptographic cost aware compiler, and scalability through
+partitioning, programmers can program in a high-level
+language absent of all the cryptographic details, while the compiler
+chooses the best circuit representation (boolean or arithmetic) for
+different parts of the program and also handles interconversions
+between them -- all this in a provably secure manner. Without \tool,
+implementing \mpc protocols for the wide array of ML algorithms from
+Section~\ref{sec:eval} would be a near-impossible task.
 
 \subsection{Related Work}
-Before ABY~\cite{aby}, several works have proposed
+Compilers
+such as Wysteria~\cite{wysteria}, ObliVM~\cite{oblivm},
+and CBMC-GC~\cite{cbmcgc} use backends with single
+representations (either boolean or arithmetic), and so the challenges
+associated with multiple representations and interconversions are
+absent. Moreover, as shown in \sectionref{oblivm}, they perform poorly
+on ML tasks that involve a mix of arithmetic and boolean
+computations. Before ABY~\cite{aby}, several works have proposed
 combining secure computation protocols based on homomorphic
@@
 protocol. 
-Additionally, these prior works provide informal languages
+Additionally, these prior works are either low-level, or provide
+informal languages
 or libraries that lack formal semantics and static guarantees.
@@
 features necessary to implement machine learning applications.
-In particular, we do not discuss declassification of
-intermediate values or indexing into arrays at secret indices.
- Handling them requires additional complexity. For example,
-Wysteria~\cite{wysteria} handles the former using dependent types
-and ObliVM~\cite{oblivm} uses Oblivious RAM for the latter.
-These situations seldomly arise in the machine learning applications that we consider:
-intermediate values don't need to be declassified and arrays are mostly traversed in an oblivious manner. We provide a detailed survey of related
+For example, we currently do not support declassification of
+intermediate values. Handling them requires additional complexity, e.g.
+using Wysteria~\cite{wysteria} like dependent types.
+These situations seldomly arise in the machine learning applications
+that we consider. We provide a detailed survey of related
 work in \sectionref{related}.
+
+%% To summarize the above five points, the
+%% novelty of \tool is in implementing cryptographic cost awareness
+%% inside a compiler, allowing the .  Finally, our partitioning technique, unlike prior work,
+%% which is specific to and works on Yao’s garbled circuits, works at the
+%% source level and is agnostic to the cryptographic backend.
+
+%% These features (ease-of-programming, use of multiple crypto backends,
+%% and scalability through partitioning) enable us to write and evaluate
+%% a wide array of ML algorithms such as matrix factorization with
+%% efficient generic 2PC techniques. Without \tool, this would have been
+%% a near-impossible task and has never been done.
 
@@
 %% express the function to be computed and uses a mix of arithmetic and
-%% boolean \mpc back-ends. We illustrate the above example of $w^Tx
+%% boolean \mpc backends. We illustrate the above example of $w^Tx
 %% >b$ in our language in \figureref{ex-sml}. 
 %% As is clear from the code, the programmer can remain completely
-%% oblivious of the cryptographic back-end and underlying circuit
+%% oblivious of the cryptographic backend and underlying circuit
 %% representation.
@@
 %% We focus on semi-honest secure two-party computation and use
-%% ABY \cite{aby} as the suitable cryptographic back-end.
+%% ABY \cite{aby} as the suitable cryptographic backend.
 %% We provide a formal type system and operational semantics for our

### 2018-08-07 ndss/ld.tex
comments from the email

@@
 inputs; those are processed by our distributed circuit semantics that
-model the \mpc back-end.
+model the \mpc backend.
 %
@@
 observations corresponding to the values revealed to the parties.
-%Finally, we formalize the distributed circuit semantics (of the cryptographic back-end) that computes the generated circuit in the distributed setting, and like  the source semantics, emits observationscorresponding to the values revealed to the parties.
+%Finally, we formalize the distributed circuit semantics (of the cryptographic backend) that computes the generated circuit in the distributed setting, and like  the source semantics, emits observationscorresponding to the values revealed to the parties.
 
@@
 We combine this correctness theorem and the security of the \mpc
-back-end to prove security of the protocols generated by \tool
+backend to prove security of the protocols generated by \tool
 (\theoremref{security}).
@@
 bounds \footnote{\sectionref{impl} discusses our implementation of this
-check.}. Rule ({\sc{T-Inp}}) picks a label $m$ for the input. Finally,
+check.}. Secret dependent array accesses can be encoded using the
+conditional expressions (see Section~\ref{sec:impl}).  Rule
+({\sc{T-Inp}}) picks a label
+$m$ for the input. Finally,
 the rule ({\sc{T-Sub}}) is the subsumption rule that coerces an expression of

### 2018-08-07 ndss/pipe.tex
comments from the email

@@
 
-\noindent\textbf{Implementing code partitioning.}
-We use partitioning for programs that require large
-circuits. Specifically, we first decompose the program
-$\prog$ into a sequence of small programs
-$\prog_1\seq\dots\seq\prog_k$. And then, \tool generates
-sequence of programs $\progn_1,\dots\progn_k$ automatically. We then
-compile and execute the $k$ programs
-$\progn_1,\progn_2,\ldots,\progn_k$ sequentially, freeing up memory
-usage after execution of each $\progn_i$.
-Automating the  decomposition step requires an analysis that can 
-statically estimate the resource usage of a \tool program.
-Resource analysis of high-level programs is a well-known hard problem~\cite{raml}
-and we describe a heuristic analysis.
-
-To build $s_1$, we consider the longest
-prefix of $s$ whose computation {\it size} is below the threshold enforced by the available memory of the machine.
-If $\prog = \prog_1; \prog_r$ then we recurse on $\prog_r$ to obtain $\prog_2,\ldots,\prog_k$.
-For a program $\mprog$, to estimate ${\it size}(\mprog)$, we need to discuss three important cases: if $\mprog\equiv\mprog_1;\mprog_2$
-then ${\it size}(\mprog)={\it size}(\mprog_1)+{\it size}(\mprog_2)$; if $\mprog\equiv\ite{e_1}{\mprog_1}{\mprog_2}$ then ${\it size}(\mprog)=\max({\it size}(\mprog_1),{\it size}(\mprog_2))$; if $s\equiv\forl{i}{n_1}{n_2}{\mprog_1}$ then
-${\it size}(\mprog)=(n_2-n_1){\it size}(\mprog_1)$. If $(n_2-n_1){\it size}(\mprog_1)$ is above the threshold then
-we replace $\mprog$ by 
-\[\forl{i}{n_1}{\frac{n_1+n_2}{2}}{\mprog_1}; \forl{i}{\frac{n_1+n_2}{2}}{n_2}{\mprog_1}\]
- and recurse to find the prefix again.
-This heuristic analysis is sufficient for the benchmarks discussed in our evaluation.
-
-We stress that partitioning is fully automated and is compatible with any 2PC protocol.
+%% We stress that partitioning is fully automated and is compatible with any 2PC protocol.
 Furthermore, the \tool compiler, with suitable changes to operator costs, can target other backends including 3PC or MPC, or protocols secure against malicious adversaries.

### 2018-08-07 ndss/related.tex
added ndss version.
comments from the email

--- 2018-05-08 ccs/related.tex
@@
 % While, the programmer need not know how ABY implements boolean or arithmetic computation, he or she must know the type of compute (boolean or arithmetic) that is more efficient for specific sub-computations and also explicitly write calls to convert shares from one scheme to another (boolean secret sharing to arithmetic secret sharing and vice-versa).
-The work of Kerschbaum et al.~\cite{autoS} provides a scheme to automatically assign homomorphic encryption or garbled circuits to each operator in a computation that is expressed as a sequence of dyadic operations.
-They conjecture that the problem is NP hard and gave a linear programming based solution and a quadratic time greedy heuristic.
-These techniques are not directly applicable to \tool programs because of $\mathtt{for}$-loops and $\mathtt{if}$-conditions.
-However, we are exploring if these ideas can be extended to yield better type inference.
+Kerschbaum et al.~\cite{autoS} provide a scheme to
+automatically assign homomorphic encryption or garbled circuits to
+each operator in a computation. They conjecture that the problem is NP
+hard and gave a linear programming based solution and a quadratic time
+greedy heuristic. However they operate over circuits represented as a
+sequence of dyadic operations, while \tool provides a high-level
+programming language with loops and branches. \tool compilation is
+automatic and sub-quadratic in the size of generated ABY programs
+(e.g. \figureref{ex-aby}) after switching off memory safety
+checking. In sharp contrast, \cite{autoS} describes approaches that
+would be doubly exponential and exponential in the length of ABY
+programs. As a result, these techniques are not directly applicable to
+\tool. However, we are exploring if these ideas can be extended to
+yield better type inference.
+
 %Their performance is much poorer than ABY due to the high 
@@
 %The reason for doing this was to allow for a mix between arithmetic and boolean computations as a secure computation protocol for a pure boolean circuit would have been very inefficient. 
-%This requires deep knowledge of cryptography
-%to ensure security.
+This requires deep knowledge of cryptography
+to ensure security.
  Examples include ~\cite{barni,blanton,brickell,franz,huang,valeriaMatrix, valeriaRidge,shafindss,wu,secureml,minionn}.

### 2018-08-07 ndss/eval.tex
minor

@@
 Memory issues start showing up in the non-partitioned implementation of the 5 layer DNN and it is slower.
-Performance degrades rapidly thereafter and
+In particular, the circuit for this DNN does not fit in RAM and uses swap space.
+Performance degrades rapidly thereafter: executing the 5 layer DNN and half operations of the sixth layer incurs a 2x slowdown.
 DNNs with 6 or more layers fail to execute (terminate with a ``bus error").

### 2018-08-07 ndss/intro.tex
minor

@@
 machine memory\footnote{Using swap and
-disk space is feasible but it causes huge slowdown.}. This is a
+disk space is feasible but it causes huge slowdown ~\figureref{partitioning}).}. This is a
 show-stopper for applications like secure machine learning, secure

### 2018-08-07 ndss/ex.tex
pass

@@
 multiplexer circuits
-(lines~\ref{line:condyaobegin} to~\ref{line:condyaoend}) (in our usage
+(lines~\ref{line:condyaobegin} to~\ref{line:condyaoend}). In our usage
 of ABY, boolean is synonymous with Yao, we elaborate in
-\sectionref{impl}).
+\sectionref{impl}.
 
@@
 
-We have evaluated against handwritten ABY code for the examples in the ABY repository and \tool has comparable performance.
-However, these examples are not representative of practical ML tasks.
-Writing sophisticated ML applications such as neural networks in ABY is a very tedious and nearly impossible engineering task even for crypto experts -- automating the same is our main contribution.
+We have evaluated against handwritten ABY code for the examples in the
+ABY repository and \tool has comparable performance. However, these
+examples are not representative of practical ML tasks. Writing
+sophisticated ML applications such as neural networks in ABY is a very
+tedious and nearly-impossible engineering task even for crypto experts
+-- automating the same is our main contribution.
 
@@
 
-\section{Cryptographic cost-awareness}
+\section{Cryptographic cost awareness}
  \label{sec:ccc}
@@
 individual operations
-(addition, multiplication, inter-conversion between arithmetic and boolean, etc.).
+(addition, multiplication, interconversion between arithmetic and boolean, etc.).
 Demmler
@@
 settings. Finally, \tool chooses boolean circuits for all
-the operations lacking arithmetic support in ABY, e.g., comparisons, bit-shifts, etc.
-Further implementation details can be found in~\sectionref{impl}.
+the operations lacking arithmetic support in ABY, e.g. comparisons, bit-shifts, etc.
+We provide more implementation details in~\sectionref{impl}.
 

### 2018-08-07 ndss/intro.tex
pass

--- 2018-08-07 ndss/intro.tex
@@
 circuits for all the components manually, including the appropriate
-inter-conversion gates between them. Clearly, writing correct
+interconversion gates between them. Clearly, writing correct
 and efficient protocols in such a framework is beyond an average
@@
 stands for Easy 2 Party Computation.}, the first
-``cryptographic-cost aware'' compiler that generates efficient and
+``cryptographic cost aware'' compiler that generates efficient and
 scalable \mpc protocols using combinations of
@@
 different parts of
-the program, while automatically inserting inter-conversion gates as
+the program, while automatically inserting interconversion gates as
 necessary. In addition to guiding the compiler, the formal model also
@@
 %% be rescoped. And again, type conversions may not be clear at this
-%% point. I would simply say that EzPC inserts inter-conversion gates
+%% point. I would simply say that EzPC inserts interconversion gates
 %% automatically. It is backed by a formal model and provides
@@
 
-\subsubsection{Cryptographic-cost aware compiler} The \tool compiler
+\subsubsection{Cryptographic cost aware compiler} The \tool compiler
 compiles a source program to a hybrid computation consisting
@@
 representations for different sub-parts, and generates the
-corresponding circuits along with the required inter-conversion
+corresponding circuits along with the required interconversion
 gates. The outcome is an efficient \mpc protocol combining arithmetic
@@
 oblivious of all these cryptographic details. Indeed, \tool is the
-first such cryptographic-cost aware compiler.
+first such cryptographic cost aware compiler.
 
@@
 machine memory\footnote{Using swap and
-disk space is feasible but it causes huge slowdown.}. This is a
+disk space is feasible but it causes huge slowdown (\figureref{partitioning}).}. This is a
 show-stopper for applications like secure machine learning, secure
@@
 of our scheme against semi-honest (or ``honest but curious")
-adversaries to the semi honest security of the  \mpc backend. The
+adversaries to the semi-honest security of the  \mpc backend. The
 theorem provides protection against side-channels arising from
@@
 of the protocols
-generated by \tool are comparable to or better than (up to 19x) their state-of-the-art, hand-crafted implementations. Finally, we
+generated by \tool is comparable to or better than (up to 19x) their state-of-the-art, hand-crafted implementations. Finally, we
 demonstrate the usefulness of partitioning by implementing an application

### 2018-08-07 ndss/ld.tex
pass

@@
 
-In this section we prove correctness and security of our \tool compiler.
+In this section we prove correctness and security of the \tool compiler.
 The readers who are interested in using the compiler as a black box can
@@
 %For the security of our compiler, we rely on the correctness guarantee and  reduce the security to the security of the underlying cryptographic protocol used to compute the generated circuit (see \theoremref{security}).
-We present only selected parts of our formalization.
-Full definitions, auxiliary lemmas and complete proofs (that follow from a straight-forward induction argument once the key lemmas are identified) can be found in the full version of this paper.
-%We will also
+We present only selected parts of our formalization. Full definitions,
+auxiliary lemmas, and complete proofs (that follow from a
+straightforward induction argument once the key lemmas are
+identified) can be found in the full version of this paper. %We will
+                                %also
 %release complete typeset proofs in a tech report.
@@
                  \Gamma \vdash e : \kw{uint}^{\mathcal{P}} \leadsto \widetilde{e}\\\\
-                  \models e < n
+                 \Gamma \models e < n
                }
@@
 that the array index $e$ is public, and uses a static bounds checking
-judgment $ \models e < n$ to prove that the array index is in
+judgment $\Gamma \models e < n$ to prove that the array index is in
 bounds \footnote{\sectionref{impl} discusses our implementation of this

### 2018-08-07 ndss/pipe.tex
pass

@@
 In this section, we describe our ``secure code partitioning'' technique that
-allows \tool to execute programs that require large circuits.
-The reason to implement partitioning is as follows: when compiling from a high level functionality to a circuit being evaluated securely, the circuits so generated can quite often be larger than the total memory available on the machines executing 2PC.
-The secure protocol for evaluating this circuit cannot then be generated without swapping the circuit in and out of main memory, thereby causing large slowdowns (see Section \ref{sec:pipeeval}).
-In order to avoid this problem, we ``split'' the high level functionality into smaller programs that compile into small circuits whose secure evaluation can completely fit into main memory.
-These smaller circuits are then evaluated sequentially to evaluate the entire program.
-Naturally, care must be taken to ensure that the partial evaluation of these circuits do not reveal any more information than the final output of the overall functionality.
-More details follow.
+allows \tool to execute programs that require large circuits. The
+reason to implement partitioning is as follows: when compiling from a
+high level functionality to a circuit, the generated circuits
+can quite often be larger than the total memory
+available on the machines executing 2PC. The secure protocol for
+evaluating this circuit cannot then be generated without swapping the
+circuit in and out of main memory, thereby causing large slowdowns
+(see Section \ref{sec:pipeeval}). In order to avoid this problem, we
+``split'' the high level functionality into smaller programs that
+compile into small circuits whose secure evaluation can completely fit
+into main memory. These smaller circuits are then evaluated
+sequentially to evaluate the entire program. Naturally, care must be
+taken to ensure that the partial evaluation of these circuits do not
+reveal any more information than the final output of the overall
+functionality. More details follow.
 

### 2018-08-07 ndss/impl.tex
nits

@@
 (SeaHorn~\cite{seahorn}) to check that the array indices
-are within bounds ($\models e < n$
+are within bounds ($\Gamma \models e < n$
 in ({\sc {T-Read}}) and ({\sc{T-Write}}),

### 2018-08-07 ndss/impl.tex
nit

@@
 If $\prog = \prog_1; \prog_r$ then we recurse on $\prog_r$ to obtain $\prog_2,\ldots,\prog_k$.
-For a program $\mprog$, to estimate ${\it size}(\mprog)$, we need to discuss three important cases: if $\mprog\equiv\mprog_1;\mprog_2$
-then ${\it size}(\mprog)={\it size}(\mprog_1)+{\it size}(\mprog_2)$; if $\mprog\equiv\ite{e_1}{\mprog_1}{\mprog_2}$ then ${\it size}(\mprog)=\max({\it size}(\mprog_1),{\it size}(\mprog_2))$; if $s\equiv\forl{i}{n_1}{n_2}{\mprog_1}$ then
+For a program $\mprog$, to estimate ${\it size}(\mprog)$, we need to
+consider three important cases: if $\mprog\equiv\mprog_1;\mprog_2$
+then ${\it size}(\mprog)={\it size}(\mprog_1)+{\it size}(\mprog_2)$; if $\mprog\equiv\ite{e_1}{\mprog_1}{\mprog_2}$ then ${\it size}(\mprog)=\max({\it size}(\mprog_1),{\it size}(\mprog_2))$;if $\mprog\equiv\forl{i}{n_1}{n_2}{\mprog_1}$ then
 ${\it size}(\mprog)=(n_2-n_1){\it size}(\mprog_1)$. If $(n_2-n_1){\it size}(\mprog_1)$ is above the threshold then

### 2018-08-07 ndss/eval.tex
nits in cryptographic cost aware

@@
 was 500x slower. Whereas, we show that by using a
-cryptographic-cost aware compiler, we can scale generic 2PC to real
+cryptographic cost aware compiler, we can scale generic 2PC to real
 prediction tasks, and get performance competitive to or better than the

### 2018-08-07 ndss/ex.tex
nits in cryptographic cost aware

@@
 Within the secret components, the \tool compiler is ``cryptographic
-cost-aware'', and appropriately picks either arithmetic or boolean
+cost aware'', and appropriately picks either arithmetic or boolean
 circuit representations for different sub-components. For example,
@@
 
-%% \subsection{Cryptographic cost-aware compiler}
+%% \subsection{Cryptographic cost aware compiler}
 %% \label{sec:ccc}
 %% One of the main challenges for our compiler is to split the compute between arithmetic and boolean parts well. 
-%% As mentioned above, \tool is a cryptographic cost-aware compiler and uses the arithmetic and boolean costs of primitive operators and the inter-conversion costs to make the split.
+%% As mentioned above, \tool is a cryptographic cost aware compiler and uses the arithmetic and boolean costs of primitive operators and the inter-conversion costs to make the split.
 Cryptographic cost of converting an arithmetic share to a boolean share requires computing a garbled circuit for addition that is linear in size.

### 2018-08-07 ndss/pipe.tex
some changes in s5

@@
 reason to implement partitioning is as follows: when compiling from a
-high level functionality to a circuit, the generated circuits
+high level functionality to a circuit, the generated circuits,
+especially for ML applications,
 can quite often be larger than the total memory
-available on the machines executing 2PC. The secure protocol for
-evaluating this circuit cannot then be generated without swapping the
-circuit in and out of main memory, thereby causing large slowdowns
-(see Section \ref{sec:pipeeval}). In order to avoid this problem, we
+available on the machines executing the 2PC protocol. The secure
+protocol for evaluating this circuit cannot then be generated without
+swapping the circuit in and out of the main memory, thereby causing large slowdowns
+(see Section \ref{sec:pipeeval}). To avoid this problem, we use
+partitioning to
 ``split'' the high level functionality into smaller programs that
-compile into small circuits whose secure evaluation can completely fit
-into main memory. These smaller circuits are then evaluated
-sequentially to evaluate the entire program. Naturally, care must be
+compile into small circuits that can completely fit
+into the main memory. These small circuits are then evaluated
+sequentially, with proper threading of the intermediate outputs, to
+evaluate the entire program. Naturally, care must be
 taken to ensure that the partial evaluation of these circuits do not
 reveal any more information than the final output of the overall
-functionality. More details follow.
+functionality.
 
-Our techniques take inspiration from the idea of pipelining Yao's garbled circuits described in FastGC \cite{yao-pipe}.
-However, unlike FastGC, we do not operate at a circuit level and partitioning is independent of the specific
-\mpc protocol. 
-Let $\prog$ be a program in our source language that
+Concretely, let $\prog$ be a program in our source language that
 generates a circuit $\crct$. For some programs, the circuit $\crct$
@@
   for most machines the memory limit is hit first.} and fail to
-execute. Partitioning enables us to 
-execute such programs via a source to source transformation that is
-oblivious to the underlying \mpc backend. Partitioning decomposes the
+execute.
+%% Partitioning enables us to 
+%% execute such programs via a source to source transformation that is
+%% oblivious to the underlying \mpc backend.
+Partitioning decomposes the
 program $\prog$ into a sequence of smaller \tool programs
-$\progn_1,\progn_2,\ldots,\progn_k$ (as defined below) such
+$\progn_1,\progn_2,\ldots,\progn_k$, as described below, such
 that the circuit size
@@
 not reveal any more information than $\prog$).
+
+Our techniques take inspiration from the idea of pipelining Yao's garbled circuits described in FastGC \cite{yao-pipe}.
+However, unlike FastGC, we do not operate at a circuit level and the
+scheme itself is independent of the specific
+\mpc protocol. 
 %
@@
 %}
-More formal details follow, while an example illustrating the technique of secure code partitioning is provided in the full version. %\appendixref{codepartitioning}.
+We now describe the instrumentation to ensure the security of the
+partitioning scheme.
+%More formal details follow, while an example illustrating the technique of secure code partitioning is provided in the full version. %\appendixref{codepartitioning}.
 
-Let $\prog$ be a program that takes (secret) inputs $x$ from Alice and
+Let $\prog$ takes (secret) inputs $x$ from Alice and
 $y$ from Bob and produces an output $z$ to both parties. Let

### 2018-08-07 ndss/appendix_c.tex
added ndss version.
ccs comments.
herculean and appendix change.

--- 2018-02-07 usenix18/appendix_c.tex
@@
+%\newpage
+
 \section{Description of benchmarks}\label{app:benchmarks}
+
+\subsection{Benchmarks}
+
 We use $[N]$ to denote $\{0,1,\dotsc, N-1\}$. Further, given a vector
@@
  The possible labels are
-$\ell\in\{\mathit{true},\mathit{false}\}$ and the classifier is
+$\ell\in \{ true, false \}$ and the classifier is
 $C_w\equiv w^Tx>0$.
@@
 
-In the following, we implement these classifiers in \tool and report
+We implement these classifiers in \tool and report
 the time taken for making secure predictions. Ideally, the machine
